### 2022-12-22 doc/al_benchmark_paper.tex
more agents

\documentclass[]{article}
\usepackage[a4paper, total={6.5in, 8.5in}]{geometry}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algorithmic}

\title{Towards Comparable Active Learning}

\begin{document}

\maketitle

\section{Introduction}

\subsection{Contribution}
\begin{itemize}
	\item Benchmark suit for Active Learning
	\item Properly optimized classification models
	\item Multiple use-cases: Single Dataset, Dataset Transfer
	\item Multiple Domains: Tabular, Image, Text
\end{itemize}

\section{Related Work}

\section{Overview}

\section{Methodology}

\subsection{Classification Model}
The classifier is constructed according to two kinds of information.
The general class of model (Dense, Convolutional, Attention, ...), and the configuration of the model (number of layers, size of each layer, ...). \\
The model class and exact configuration is determined by the dataset, i.e. tabular datasets will prescribe a dense model.
If special capabilities of the model are needed (i.e. Monte-Carlo Dropout), an extension of the given model class can be provided to the framework. \\ [1mm]
To ensure comparability between models, the model's configuration should not be changed or an additional evaluation of the new configuration should be conducted to compare the baseline expressivity.

\subsection{State Space}
Since every AL agent needs a different state space our environment exposes a callback-function that gives the agent full control over how the state is constructed. \\ [1mm]
The callback includes the following information:
\begin{itemize}
	\item The current sample of IDs that point to the presented unlabeled datapoints
	\item The entire labeled dataset $\mathcal{L}$
	\item The entire unlabeled dataset $\mathcal{U}$
	\item A histogram of labeled points per class (simple count)
	\item The current classification model including all model weights
	\item The current optimizer including it's full state
\end{itemize}
Every agent needs to implement this callback that transforms the given information into a state tensor that will be directly consumed by the agent to make it's prediction.

\bibliographystyle{plain}
\bibliography{main.bib}

\end{document}

### 2023-01-05 core/agent.py
meta data

@@ -26,3 +26,7 @@ class BaseAgent(ABC):
     @abstractmethod
     def predict(self, state:Union[Tensor, dict], greed:float=0.0)->Tensor:
         pass
+
+
+    def get_meta_data(self)->str:
+        return f"{self.name}"

### 2023-01-05 core/classifier.py

@@ -3,23 +3,23 @@ import torch.nn.functional as F
 from torch import Tensor
 import torch
 
-def get_dense_sequential_model(input_size:int, num_classes:int, hidden_sizes:tuple)->nn.Sequential:
-    if len(hidden_sizes) > 0:
-        inpt = nn.Linear(input_size, hidden_sizes[0])
-        hidden = list()
-        for i in range(len(hidden_sizes)):
-            hidden.append(nn.Linear(hidden_sizes[max(0, i - 1)], hidden_sizes[i]))
-            hidden.append(nn.ReLU(inplace=True))
-        out = nn.Linear(hidden_sizes[-1], num_classes)
-        model = nn.Sequential(inpt, nn.ReLU(inplace=True),
-                              *hidden,
-                              out)
-    else:
-        model = nn.Sequential(nn.Linear(input_size, num_classes),
-                              nn.ReLU(inplace=True))
-    example_forward_input = torch.rand(1, input_size)
-    model = torch.jit.trace_module(model, {'forward': example_forward_input})
-    return model
+# def get_dense_sequential_model(input_size:int, num_classes:int, hidden_sizes:tuple)->nn.Sequential:
+#     if len(hidden_sizes) > 0:
+#         inpt = nn.Linear(input_size, hidden_sizes[0])
+#         hidden = list()
+#         for i in range(len(hidden_sizes)):
+#             hidden.append(nn.Linear(hidden_sizes[max(0, i - 1)], hidden_sizes[i]))
+#             hidden.append(nn.ReLU(inplace=True))
+#         out = nn.Linear(hidden_sizes[-1], num_classes)
+#         model = nn.Sequential(inpt, nn.ReLU(inplace=True),
+#                               *hidden,
+#                               out)
+#     else:
+#         model = nn.Sequential(nn.Linear(input_size, num_classes),
+#                               nn.ReLU(inplace=True))
+#     example_forward_input = torch.rand(1, input_size)
+#     model = torch.jit.trace_module(model, {'forward': example_forward_input})
+#     return model
 
 class DenseModel(nn.Module):
     def __init__(self, input_size:int, num_classes:int, hidden_sizes:tuple):

### 2023-01-05 core/data.py

@@ -15,7 +15,7 @@ class BaseDataset(ABC):
                  classifier_batch_size:int,
                  cache_folder:str="~/.al_benchmark/datasets",
                  device=None,
-                 class_fitting_mode:Literal["from_scratch", "finetuning", "single_sample"]="finetuning"):
+                 class_fitting_mode:Literal["from_scratch", "finetuning"]="finetuning"):
         assert isinstance(budget, int) and budget > 0, f"The budget {budget} is invalid"
         if device is None:
             self.device = torch.device('cuda' if torch.cuda.is_available() else 'cpu')
@@ -111,6 +111,9 @@ class BaseDataset(ABC):
                 if type(value) == torch.Tensor:
                     setattr(self, attr, value.to(device))
         return self
+
+    def get_meta_data(self)->str:
+        return f"{self.name}"
 
 
 

### 2023-01-05 core/environment.py

@@ -20,11 +20,7 @@ class ALGame(gym.Env):
         self.budget = dataset.budget
         self.sample_size = labeled_sample_size
         self.fitting_mode = dataset.class_fitting_mode
-        self.next_case_weight = 1.0
-        if self.fitting_mode == "single_sample":
-            self.loss = nn.CrossEntropyLoss(reduction="none")
-        else:
-            self.loss = nn.CrossEntropyLoss()
+        self.loss = nn.CrossEntropyLoss()
         self.create_state_callback = create_state_callback
 
         # set gym observation space and action space
@@ -49,10 +45,6 @@ class ALGame(gym.Env):
             self.initial_weights = self.classifier.state_dict()
             self.optimizer = self.dataset.get_optimizer(self.classifier)
             self.reset_al_pool()
-
-            # variables for single sample fitting mode
-            self.next_case_weight = 1.0
-            self.total_weight = len(self.x_labeled)
         # first training of the model should be done from scratch
         self._fit_classifier(from_scratch=True)
         self.initial_test_accuracy = self.current_test_accuracy
@@ -95,48 +87,6 @@ class ALGame(gym.Env):
         return next_state, reward, done, truncated, {}
 
 
-    def _fit_classifier_single_sample(self):
-        last_x = self.x_labeled[-1]
-        last_y = self.y_labeled[-1]
-        random_id = np.random.choice(list(range(len(self.x_labeled)-1)))
-        batch_x = torch.stack([self.x_labeled[random_id], last_x], dim=0)
-        batch_y = torch.stack([self.y_labeled[random_id], last_y], dim=0)
-        self.next_case_weight += 1.0 / self.total_weight
-        self.total_weight += self.next_case_weight
-
-        # update
-        yHat = self.classifier(batch_x)
-        weights = torch.Tensor([1.0, self.next_case_weight]).to(self.device)
-        loss_value = weights * self.loss(yHat, batch_y)
-        loss_value = loss_value.mean()
-        self.optimizer.zero_grad()
-        loss_value.backward()
-        self.optimizer.step()
-
-        # eval
-        with torch.no_grad():
-            test_dataloader = DataLoader(TensorDataset(self.x_test, self.y_test), batch_size=100)
-            loss_sum = 0.0
-            total = 0.0
-            correct = 0.0
-            for batch_x, batch_y in test_dataloader:
-                yHat = self.classifier(batch_x)
-                predicted = torch.argmax(yHat, dim=1)
[…]

### 2023-01-05 datasets/splice.py

@@ -5,7 +5,7 @@ import torch.nn as nn
 import numpy as np
 from sklearn.datasets import load_svmlight_file
 from core.data import BaseDataset, normalize, postprocess_svm_data
-from core.classifier import DenseModel, get_dense_sequential_model
+from core.classifier import DenseModel#, get_dense_sequential_model
 
 class Splice(BaseDataset):
 
@@ -78,12 +78,12 @@ class Splice(BaseDataset):
 
     def get_classifier(self, hidden_dims :Tuple[int] =(24, 12)) -> nn.Module:
         input_size = self.x_test.size(1)
-        return get_dense_sequential_model(input_size=input_size,
-                           num_classes=self.y_test.size(1),
-                           hidden_sizes=hidden_dims)
-        # model = DenseModel(input_size=input_size,
+        # return get_dense_sequential_model(input_size=input_size,
         #                    num_classes=self.y_test.size(1),
         #                    hidden_sizes=hidden_dims)
+        model = DenseModel(input_size=input_size,
+                           num_classes=self.y_test.size(1),
+                           hidden_sizes=hidden_dims)
         # with torch.no_grad():
         #     example_forward_input = torch.rand(1, input_size)
         #     example_encode_input = torch.rand(1, input_size)
@@ -97,4 +97,10 @@ class Splice(BaseDataset):
 
 
     def get_optimizer(self, model) -> torch.optim.Optimizer:
-        return torch.optim.NAdam(model.parameters(), lr=0.001)
+        return torch.optim.NAdam(model.parameters(), lr=0.001,
+                                 weight_decay=0.0)
+
+    def get_meta_data(self)->str:
+        s = super().get_meta_data()
+        return f"{s}\n" \
+               f"Budget: {self.budget}"

### 2023-01-25 doc/research_presentation/research_presentation.tex
Splice SGD test

\documentclass[aspectratio=169, 11pt, invertlogo]{ismll-slides}
% - logo=[all, frametitle, headline, none]
% - logotitlepage=[true, false]
% - invertlogo: where to use white instead of black text (on transparent background)
% - nopdfpagenumbers: disable adding a number for each frame to pdf toc

%%%%%%%%%%%%%%%%
% BEAMER THEME %
%%%%%%%%%%%%%%%%
\usetheme[block=fill, background=light, progressbar=foot, numbering=counter]{metropolis}
%\usetheme{Madrid}
\setbeamertemplate{blocks}[default]
\setbeamercovered{transparent}

%%%%%%%%%%%%%%%%
% BIBLIOGRAPHY %
%%%%%%%%%%%%%%%%
% NOTE: Bibliography should be compiled with biber!
\addbibresource{slides-bibliography.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE PAGE INFORMATION %
%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Active Learning Benchmark}
\subtitle{Towards Comparable Active Learning}
\date{February 31, 2021}
\institute{Information Systems and Machine Learning Lab (ISMLL)\\Institute for Computer Science \\ University of Hildesheim}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Competitor Paper}
	Some Benchmark papers for DAL exist already :
	\begin{itemize}
		\item Focus on Image Classification
		\item Focus on best possible classifier performance
		\begin{itemize}
			\item Data Augmentation
			\item Type of Optimizer
			\item Semi-Supervised Learning
		\end{itemize}
	\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Classifier}
	Adam vs. SGD?
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Cross Validation}
	pass
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Evaluation Metrics}
	pass
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Single Domain vs. Transfer}
	pass
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{True Hyperparameters}
	Sample Size
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[allowframebreaks]%{References}
%
\printbibliography%[heading=none]
%
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2023-01-25 datasets/splice.py

@@ -48,8 +48,9 @@ class Splice(BaseDataset):
 
 
     def get_optimizer(self, model, lr=0.001, weight_decay=0.0) -> torch.optim.Optimizer:
-        return torch.optim.NAdam(model.parameters(), lr=lr,
-                                 weight_decay=weight_decay)
+        # return torch.optim.NAdam(model.parameters(), lr=lr,
+        #                          weight_decay=weight_decay)
+        return torch.optim.SGD(model.parameters(), lr=lr, weight_decay=weight_decay)
 
     def get_meta_data(self) ->str:
         s = super().get_meta_data() + '\n'

### 2023-01-26 doc/research_presentation/research_presentation.tex
cifar10 resnet

@@
 % NOTE: Bibliography should be compiled with biber!
-\addbibresource{slides-bibliography.bib}
+%\bibliography{slides-bibliography}
+%\addbibresource{slides-bibliography}
 
@@
 
+\begin{frame}[fragile]{Motivation}
+	The DAL result landscape is terrible \\ [2mm]
+	Every paper uses different datasets and use-cases \\ [1mm]
+	Every paper has a different classification model and training regime \\ [1mm]
+	No one-fits-all evaluation metric 
+\end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{frame}[fragile]{Competitor Paper}
-	Some Benchmark papers for DAL exist already :
+\begin{frame}[fragile]{Competitor Papers}
+	Some Benchmark papers for DAL exist already:
 	\begin{itemize}
@@
 	\end{itemize}
-
 \end{frame}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}[fragile]{Classifier}
-	Adam vs. SGD?
+\begin{frame}[fragile]{Main Idea}
+	We focus on the AL algorithms themselves:
+	\begin{itemize}
+		\item When we find the "best" algorithm, classification performance will come naturally
+		\item AL is lacking reproducible research, not strong classification models (random/uncertainty sampling does a good job already)
+		\item We incorporate different domains instead of focusing on images
+		\item We reduce the amount of true hyperparameters by design
+	\end{itemize}
 \end{frame}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}[fragile]{Cross Validation}
-	pass
+\begin{frame}[fragile]{Theoretical Problems}
+	Some conflicting results from other papers:
+	\begin{itemize}
+		\item Data augmentations (seemingly) can replace diversity components from AL algorithms (under investigation currently)
+	\end{itemize}
 \end{frame}
@@
 
+\begin{frame}[fragile]{Scope}
+	The framework is planned to include
+	\begin{itemize}
+		\item Three different domains (Tabular, Image, Text)
+		\item SOTA baseline methods 
+		\item Preprocessed datasets (train/test split and seed set)
+		\item Tuned classifiers for each dataset
+		\item Set logging and evaluation procedures
+		\item (Single Domain and Domain Transfer use cases)
+	\end{itemize}
+	The framework will \textbf{not} include
+	\begin{itemize}
+		\item Batch Active Learning
+	\end{itemize}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Design Decisions}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{The AL Cycle}
+	\begin{figure}
+		\centering
+		\includegraphics[width=\linewidth]{pics/al_cycle.png}
+	\end{figure}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{Datasets}
+	Each dataset should be preprocessed - The features are normalized, the target are one-hot encoded and \textbf{the seed set should be fixed}
+
+	Datasets should be selected by their "potential"
+
+	Each dataset needs an oracle curve that sets an upper bound performance
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{Classifier}
+	The classification model should be governed by the dataset
+
+	There is no need to have the same model for every dataset, as long as the model in question is suited well for the data
+
+	Simpler models with less dynamic behaviour are better - SOTA performance is not so important
+
+	The model needs have tuned hyperparameters for each dataset (full dataset or subset?)
+
+	The expectation is that good AL algorithms will also work well on SOTA models
+	%Adam vs. SGD?
+\end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{frame}[fragile]{Evaluation Metrics}
-	pass
+	Multiple options are available
+	\begin{itemize}
+		\item Accuracy* / F1-Score
+		\item AUC
+		\item Advantage over random sampling*
+		\item Regret from the oracle curve
+		\item Relative performance (0\% = random / 100\% = oracle)
+	\end{itemize}
 \end{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{Cross Validation}
+	Splice: 
+	random 0.874 +- 0.01
+	margin 0.885 +- 0.015
+	difference 0.011
+	Changing the data vs. changing the model initialization
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{Hyperparameters}
+	Data HPs are fixed by the dataset (budget, seed set, ...)
+
+	Classifier HPs are set with BO per dataset (architecture, learning rate, regularization, ...)
+
+	Environment HPs are choosen with respect to the classifiers (training loop, ...)
+
+	Each AL agent defines its own state space with the information from the environment
+
+	This leaves very few "true" hyperparameters: 
+	\begin{itemize}
+		\item Unlabeled Sample Size
+		\item \# of cross validation trials
+	\end{itemize}
+\end{frame}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}[fragile]{True Hyperparameters}
-	Sample Size
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}[allowframebreaks]%{References}
-%
-\printbibliography%[heading=none]
-%
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \end{document}

### 2023-01-26 bo_optimize_classifier.py

@@ -25,7 +25,7 @@ bounds = torch.Tensor([
     [-100, 100],
     # Hidden size multiplier: This sets the size of each hidden layer by
     # size = layer number x multiplier      or
-    # size = (1/layer number) x multiplier  if multiplier < 0
+    # size = (1/layer number) x -multiplier  if multiplier < 0
 ]).T
 
 norm_bounds = torch.stack([
@@ -158,3 +158,9 @@ if __name__ == '__main__':
     os.makedirs(folder, exist_ok=True)
     filename = os.path.join(folder, filename)
     df.to_csv(filename)
+    # write some meta information
+    filename = os.path.join(folder, "meta.txt")
+    with open(filename, "w") as f:
+        if hasattr(dataset, "get_meta_data"):
+            f.write("# Dataset: \n")
+            f.write(f"{dataset.get_meta_data()} \n\n")

### 2023-01-26 datasets/cifar10.py

@@ -27,9 +27,11 @@ class Cifar10(BaseDataset):
 
 
     def get_classifier(self, hidden_dims :Tuple[int] =(24, 12)) -> nn.Module:
-        model = ConvolutionalModel(input_size=self.x_shape,
-                                   num_classes=self.n_classes,
-                                   hidden_sizes=hidden_dims)
+        from core.resnet import ResNet18
+        model = ResNet18()
+        # model = ConvolutionalModel(input_size=self.x_shape,
+        #                            num_classes=self.n_classes,
+        #                            hidden_sizes=hidden_dims)
         return model
 
 

### 2023-01-26 testing.py

@@ -1,3 +1,5 @@
+import sys
+
 import datasets.dna
 import experiment_util as util
 import os, argparse
@@ -38,9 +40,15 @@ torch.random.manual_seed(args.seed)
 SAMPLE_SIZE = 20
 
 # dataset = datasets.DNA(cache_folder="../datasets")
-dataset = datasets.Splice(cache_folder="../datasets")
-# dataset = datasets.Cifar10(cache_folder="../datasets")
+# dataset = datasets.Splice(cache_folder="../datasets")
+dataset = datasets.Cifar10(cache_folder="../datasets")
 dataset = dataset.to(util.device)
+
+from core.classifier import fit_and_evaluate
+acc = fit_and_evaluate(dataset, lr=0.1, weight_decay=0.0, batch_size=128, hidden_sizes=None)
+print("Final Acc", acc)
+sys.exit(0)
+
 env = ALGame(dataset,
              SAMPLE_SIZE,
              AgentClass.create_state_callback,

### 2023-01-27 doc/research_presentation/research_presentation.tex

@@
 \begin{frame}[fragile]{Cross Validation}
-	Splice: 
-	random 0.874 +- 0.01
-	margin 0.885 +- 0.015
-	difference 0.011
-	Changing the data vs. changing the model initialization
+	\begin{columns}
+		\begin{column}{0.3\linewidth}
+			Splice (Tabular): \\
+			Random: $0.874 \pm 0.01$ \\
+			Margin: $0.885 \pm 0.015$ \\
+			Difference: $0.011$
+			%Changing the data vs. changing the model initialization
+		\end{column}
+		\hspace{3mm}
+		\begin{column}{0.7\linewidth}
+			\includegraphics[width=0.9\linewidth]{pics/splice_performance.png}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{Cross Validation}
+	AL can cross validate in two ways
+	\begin{itemize}
+		\item Changing the seed set
+		\item Changing the presented sub-samples of unlabeled points
+		\item Changing model initialization
+	\end{itemize}
+	Currently, I keep the seed set fixed \\ [2mm]
+	Is there value in fixing everything? (Model checkpoints, pre-sampling data*, etc.)
 \end{frame}

### 2023-01-30 doc/research_presentation/research_presentation.tex
dna runs
docs

@@
 		\item Batch Active Learning
+		\item Elaborate Training (Data Augmentation, etc.)
 	\end{itemize}
@@
 	Each dataset needs an oracle curve that sets an upper bound performance
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{Dataset Potential}
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.7\linewidth]{pics/splice_oracle.png}
+	\end{figure}
 \end{frame}
@@
 \begin{frame}[fragile]{Cross Validation}
+AL can cross validate in two ways
+\begin{itemize}
+	\item Changing the seed set
+	\item Changing the presented sub-samples of unlabeled points
+	\item Changing model initialization
+\end{itemize}
+Currently, I keep the seed set fixed \\ [2mm]
+Is there value in fixing everything? (Model checkpoints, pre-sampling data*, etc.)
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{Cross Validation}
 	\begin{columns}
@@
 \begin{frame}[fragile]{Cross Validation}
-	AL can cross validate in two ways
-	\begin{itemize}
-		\item Changing the seed set
-		\item Changing the presented sub-samples of unlabeled points
-		\item Changing model initialization
-	\end{itemize}
-	Currently, I keep the seed set fixed \\ [2mm]
-	Is there value in fixing everything? (Model checkpoints, pre-sampling data*, etc.)
+	\begin{columns}
+		\begin{column}{0.5\linewidth}
+			Random Sampling
+			\includegraphics[width=\linewidth]{pics/splice_random_convergence.jpg}
+		\end{column}
+		\hspace{3mm}
+		\begin{column}{0.5\linewidth}
+			Uncertainty Sampling
+			\includegraphics[width=\linewidth]{pics/splice_margin_convergence.jpg}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{Cross Validation}
+	\begin{figure}
+		\centering
+		\includegraphics[width=.7\linewidth]{pics/splice_margin100_convergence.jpg}
+	\end{figure}
 \end{frame}
@@
 \begin{frame}[fragile]{Hyperparameters}
-	Data HPs are fixed by the dataset (budget, seed set, ...)
+	Data HPs are fixed for each dataset (budget, seed set, ...)
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{The End}
+	Cheers
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{Classifier Training}
+	Labeled points are unweighted
+
+	Always at least one epoch of training
+
+	Aggressive early stopping with patience $0$
+
+	Adam (NAdam) optimizer, since SGD proves to be too slow (needs more epochs)
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}

### 2023-01-31 core/environment.py
cifar10 preliminary tests

@@ -1,4 +1,5 @@
 from typing import Callable, Tuple
+from time import time
 import numpy as np
 import torch
 import torch.nn as nn
@@ -27,7 +28,11 @@ class ALGame(gym.Env):
 
         # set gym observation space and action space
         self.current_test_accuracy = 0.0
+        start_time = time()
         state = self.reset()
+        duration = time() - start_time
+        print("Estimated time per run (h): %1.1f"%(duration * self.budget / 60 / 60))
+
         if isinstance(state, dict):
             self.observation_space = dict()
             for key, value in state.items():

### 2023-01-31 generate_dataset_leaderboard.py

@@ -1,0 +1,37 @@
+import os
+import pandas as pd
+import numpy as np
+
+for dataset in os.listdir("runs"):
+    dataset_folder = os.path.join("runs", dataset)
+    leaderboard = pd.DataFrame(columns=["Agent", "Acc", "Improvement", "AUC"])
+
+    agent_idx = 0
+    for agent in os.listdir(dataset_folder):
+        agent_folder = os.path.join(dataset_folder, agent)
+        acc_file = os.path.join(agent_folder, "accuracies.csv")
+        if os.path.exists(acc_file):
+            accuracies = pd.read_csv(acc_file, header=0, index_col=0)
+            values = accuracies.values
+            final_acc = values[-1, :]
+            final_acc = (np.mean(final_acc), np.std(final_acc))
+            improvement = values[-1, :] - values[0, :]
+            improvement = (np.mean(improvement), np.std(improvement))
+            auc = np.sum(values, axis=0) / values.shape[0]
+            auc = (np.mean(auc), np.std(auc))
+            leaderboard = pd.concat([
+                leaderboard,
+                pd.DataFrame({
+                    "Agent": agent,
+                    "Acc": "%1.4f +- %1.4f"%final_acc,
+                    "Improvement": "%1.4f +- %1.4f"%improvement,
+                    "AUC": "%1.4f +- %1.4f"%auc,
+                }, index=[agent_idx])
+            ])
+            agent_idx += 1
+
+    if len(leaderboard) > 0:
+        leaderboard_file = os.path.join(dataset_folder, "leaderboard.csv")
+        if os.path.exists(leaderboard_file):
+            os.remove(leaderboard_file)
+        leaderboard.to_csv(leaderboard_file)

### 2023-01-31 datasets/cifar10.py
cifar10 preliminary tests

@@ -7,7 +7,7 @@ from core.classifier import ConvolutionalModel
 
 class Cifar10(BaseDataset):
 
-    def __init__(self, budget=2000, initial_points_per_class=1, classifier_batch_size=128,
+    def __init__(self, budget=2000, initial_points_per_class=1, classifier_batch_size=256,
                  data_file="cifar10_al.pt",
                  cache_folder:str="~/.al_benchmark/datasets"):
         super().__init__(budget, initial_points_per_class, classifier_batch_size, data_file, cache_folder)

### 2023-01-31 doc/research_presentation/research_presentation.tex
cifar10 preliminary tests
presentation

@@
 		\item Set logging and evaluation procedures
-		\item (Single Domain and Domain Transfer use cases)
+		\item (Ensemble Approaches)
+		\item (Single Domain and Domain-Transfer use cases)
 	\end{itemize}
@@
 		\includegraphics[width=.7\linewidth]{pics/splice_margin100_convergence.jpg}
+		\caption*{Uncertainty Sampling}
 	\end{figure}
@@
 
+\begin{frame}[fragile]{Ensemble Approaches}
+	Uncertainty Sampling profits a lot from ensembles
+
+	The logical thing is to build the final classifier as an ensemble \\ [7mm]
+	%
+	How do you compare single model approaches to ensemble approaches?
+
+	Prefix a single evaluation model out of the ensemble?
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}[fragile]{Single Domain vs. Transfer}
-	pass
+	Single Domain: Fit AL agent on the same dataset that you evaluate on (not useful in practice)
+
+	Domain Transfer: Fit AL agent on domain with large resources and evaluate agent on domain with few resources
+
+	Often used in NLP (fit agent on English data and evaluate for Spanish / German)
 \end{frame}
@@
 \begin{frame}[fragile]{Classifier Training}
-	Labeled points are unweighted
+	\begin{table}
+		\centering
+		\begin{tabular}{l || l}
+			\textbf{General Setup} & \textbf{Speed Optimizations} \\
+			\hline
+			Cross-Entropy Loss & Fine tuning \\ [1mm]
+			No case weights & Agressive Early Stopping \\[1mm]
+			At least one epoch & \\[1mm]
+			NAdam optimizer &
+		\end{tabular}
+	\end{table}
+\end{frame}
 
-	Always at least one epoch of training
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-	Aggressive early stopping with patience $0$
-
-	Adam (NAdam) optimizer, since SGD proves to be too slow (needs more epochs)
+\begin{frame}[fragile]{Some Results}
+	\begin{table}
+		\centering
+		\begin{tabular}{l || c | c }
+			\textbf{Agent} & \textbf{Acc} & \textbf{AUC} \\
+			\hline
+			Oracle & 0.8806 $\pm$ 0.0141 & 0.8530 $\pm$ 0.0132 \\
+			SAR & 0.8839 $\pm$ 0.0142 & 0.8339 $\pm$ 0.0118 \\
+			Margin & 0.8849 $\pm$ 0.0112 & 0.8309 $\pm$ 0.0141 \\
+			Entropy & 0.8822 $\pm$ 0.0156 & 0.8285 $\pm$ 0.0154 \\
+			Coreset & 0.8788 $\pm$ 0.0105 & 0.8246 $\pm$ 0.0108 \\
+			Random & 0.8744 $\pm$ 0.0144 & 0.8152 $\pm$ 0.0129
+		\end{tabular}
+		\caption*{Results for Splice}
+	\end{table}
 \end{frame}

### 2023-01-31 doc/research_presentation/research_presentation.tex
presentation

@@
 
-\begin{frame}[fragile]{Hyperparameters}
-	Data HPs are fixed for each dataset (budget, seed set, ...)
+\begin{frame}[fragile]{Hyperparameters: Dataset}
+	Budget is set to allow convergence
 
-	Classifier HPs are set with BO per dataset (architecture, learning rate, regularization, ...)
+	Seed set is fixed to 1 point per class unless the cold start problem is to pronounced
 
-	Environment HPs are choosen with respect to the classifiers (training loop, ...)
+	Classifier batch size / LR / regularization is optimized for performance \\
+	Image datasets might require a trade-off for speed (Cifar10 currently needs 8 days)
+\end{frame}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{Hyperparameters: Agent}
 	Each AL agent defines its own state space with the information from the environment
 
+	Other hyperparameters for baselines need to be investigated
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]{True Hyperparameters}
 	This leaves very few "true" hyperparameters: 
@@
 
+\appendix
+
 \begin{frame}[fragile]{Classifier Training}
@@
 			Cross-Entropy Loss & Fine tuning \\ [1mm]
-			No case weights & Agressive Early Stopping \\[1mm]
+			No case weights & Aggressive Early Stopping \\[1mm]
 			At least one epoch & \\[1mm]
@@
 
-\begin{frame}[fragile]{Some Results}
+\begin{frame}[fragile]{Results for Splice}
 	\begin{table}

### 2023-02-03 doc/al_benchmark_paper.tex
meta data
upper bound

@@
 
+\subsection{Problem Description}
+We assume a dataset $\mathcal{D}$ consisting of a large pool of unlabeled data $\mathcal{U}$ and a small set of labeled data called the seed set $\mathcal{S}$: $\mathcal{D} := (\mathcal{U}, \mathcal{S}); \hspace{2mm} |\mathcal{S}| \ll |\mathcal{U|}$.
+For the purpose of this work, only fully labeled datasets are used and the labels for $\mathcal{U}$ are suppressed until they are selected for labeling.
+We consider only classification problems, hence the instances of a dataset have the form $x \in \mathbb{R}^M$ for $\mathcal{U}$ and $(x, y) \in (\mathbb{R}^M, \mathbb{R}^C)$ for $\mathcal{S}$.
+To perform classification, a model $\phi_\theta := \mathbb{R}^M \rightarrow \mathbb{R}^C$ is used.
+To fit the model, it is parameterized by $\theta$ and subjected to loss $l(y, \phi_\theta(x)) := \mathbb{R}^C \times \mathbb{R}^C \rightarrow \mathbb{R}$.
+For this work, categorical cross-entropy (CE) is used.
+
 \section{Methodology}
@@
 	\item The entire unlabeled dataset $\mathcal{U}$
-	\item A histogram of labeled points per class (simple count)
+	\item A histogram of labeled points per class (count)
+	\item The available budget
+	\item Number of added datapoints $|\mathcal{L}| - |\mathcal{S}|$
+	\item The initial validation accuracy and current validation accuracy
 	\item The current classification model including all model weights

### 2023-02-03 core/classifier.py

@@ -76,23 +76,29 @@ class ConvolutionalModel(nn.Module):
 
 
 def fit_and_evaluate(dataset:BaseDataset,
-                     lr:float, weight_decay:float, batch_size:int,
-                     hidden_sizes:tuple,
+                     lr:float=None, weight_decay:float=None, batch_size:int=None,
+                     hidden_sizes:tuple=None,
                      disable_progess_bar:bool=False,
                      max_epochs:int=1000):
     loss = nn.CrossEntropyLoss()
-    model = dataset.get_classifier(hidden_dims=hidden_sizes)
+    if hidden_sizes is not None:
+        model = dataset.get_classifier(hidden_dims=hidden_sizes)
+    else:
+        model = dataset.get_classifier()
     model = model.to(dataset.device)
-    optimizer = dataset.get_optimizer(model, lr=lr, weight_decay=weight_decay)
+    if lr is not None and weight_decay is not None:
+        optimizer = dataset.get_optimizer(model, lr=lr, weight_decay=weight_decay)
+    else:
+        optimizer = dataset.get_optimizer(model)
 
     train_dataloader = DataLoader(TensorDataset(dataset.x_unlabeled, dataset.y_unlabeled),
                                   batch_size=batch_size,
                                   shuffle=True, num_workers=4)
-    test_dataloader = DataLoader(TensorDataset(dataset.x_test, dataset.y_test), batch_size=256,
+    test_dataloader = DataLoader(TensorDataset(dataset.x_test, dataset.y_test), batch_size=512,
                                  num_workers=4)
 
     class EarlyStopping:
-        def __init__(self, patience=3):
+        def __init__(self, patience=5):
             self.patience = patience
             self.best_loss = torch.inf
             self.steps_without_improvement = 0

### 2023-02-16 core/classifier.py
upper bound fix

@@ -90,6 +90,8 @@ def fit_and_evaluate(dataset:BaseDataset,
         optimizer = dataset.get_optimizer(model, lr=lr, weight_decay=weight_decay)
     else:
         optimizer = dataset.get_optimizer(model)
+    if batch_size is None:
+        batch_size = dataset.classifier_batch_size
 
     train_dataloader = DataLoader(TensorDataset(dataset.x_unlabeled, dataset.y_unlabeled),
                                   batch_size=batch_size,

### 2023-02-16 evaluate.py

@@ -18,7 +18,7 @@ AgentClass = get_agent_by_name(args.agent)
 DatasetClass = get_dataset_by_name(args.dataset)
 
 if args.sub_run_id is not None:
-    print("Sub-run ID given. This will override the seed")
+    print(f"Sub-run ID {args.sub_run_id} given. This will override the seed")
     numpy.random.seed(args.sub_run_id)
     torch.random.manual_seed(args.sub_run_id)
 else:

### 2023-02-16 doc/al_benchmark_paper.tex
upper bound fix
data rework

@@
 \subsection{Problem Description}
-We assume a dataset $\mathcal{D}$ consisting of a large pool of unlabeled data $\mathcal{U}$ and a small set of labeled data called the seed set $\mathcal{S}$: $\mathcal{D} := (\mathcal{U}, \mathcal{S}); \hspace{2mm} |\mathcal{S}| \ll |\mathcal{U|}$.
+We assume a dataset $\mathcal{D} := (\mathcal{U}, \mathcal{L})$ consisting of a large pool of unlabeled data $\mathcal{U}$ and a small set of labeled data called the seed set with $|\mathcal{L}| \ll |\mathcal{U|}$.
 For the purpose of this work, only fully labeled datasets are used and the labels for $\mathcal{U}$ are suppressed until they are selected for labeling.
-We consider only classification problems, hence the instances of a dataset have the form $x \in \mathbb{R}^M$ for $\mathcal{U}$ and $(x, y) \in (\mathbb{R}^M, \mathbb{R}^C)$ for $\mathcal{S}$.
+For evaluation purposes we also assume a held-out test set $(x_{test}, y_{test}) \in (\mathbb{R}^M, \mathbb{R}^C)$
+We consider only classification problems, hence the instances of a dataset have the form $x \in \mathbb{R}^M$ for $\mathcal{U}$ and $(x, y) \in (\mathbb{R}^M, \mathbb{R}^C)$ for $\mathcal{L}$.
 To perform classification, a model $\phi_\theta := \mathbb{R}^M \rightarrow \mathbb{R}^C$ is used.
@@
 For this work, categorical cross-entropy (CE) is used.
+For evaluating classification performance, we use Accuracy \\ [1mm]
+Active learning is defined as sequentially selecting single instances of $x_i \in \mathcal{U}$, requesting their label $y_i$ and adding them to the labeled pool $\mathcal{L} := \mathcal{L} \cup (x_i, y_i)$ until a fixed budget $B$ is exhausted $i := 0 \ldots B$.
+After each added instance the classification model is retrained according to section \ref{sec:training_the_classifier} and its performance is measured on a held-out test set.
+The quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
+We employ normalized AUC:
+\begin{equation}
+	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \phi, B) := \frac{1}{B} \sum_{i=1}^{B} acc(y_{test}, \phi_i(x_{test}))
+\end{equation}
+Where $\phi_i$ is the retrained classification model after the i-\textit{th} instance was selected
 
@@
 
+\subsection{Training the Classifier}\label{sec:training_the_classifier}
+\subsection{Evaluation}\label{sec:evaluation}
+
+\section{Ablation Studies}
+\begin{itemize}
+	\item Weird drop of performance for multiples of batch size (drop\_last in DataLoader)
+	\item Reduction of the test set for speed
+\end{itemize}
+
 \bibliographystyle{plain}

### 2023-02-16 core/classifier.py

@@ -86,19 +86,20 @@ def fit_and_evaluate(dataset:BaseDataset,
     else:
         model = dataset.get_classifier()
     model = model.to(dataset.device)
+
     if lr is not None and weight_decay is not None:
         optimizer = dataset.get_optimizer(model, lr=lr, weight_decay=weight_decay)
     else:
         optimizer = dataset.get_optimizer(model)
+
     if batch_size is None:
         batch_size = dataset.classifier_batch_size
 
-    train_dataloader = DataLoader(TensorDataset(dataset.x_unlabeled, dataset.y_unlabeled),
+    train_dataloader = DataLoader(TensorDataset(dataset.x_train, dataset.y_train),
                                   batch_size=batch_size,
                                   shuffle=True, num_workers=4)
     test_dataloader = DataLoader(TensorDataset(dataset.x_test, dataset.y_test), batch_size=512,
                                  num_workers=4)
-
     class EarlyStopping:
         def __init__(self, patience=5):
             self.patience = patience

### 2023-02-17 doc/al_benchmark_paper.tex
agent state space rework

@@
 For evaluating classification performance, we use Accuracy \\ [1mm]
-Active learning is defined as sequentially selecting single instances of $x_i \in \mathcal{U}$, requesting their label $y_i$ and adding them to the labeled pool $\mathcal{L} := \mathcal{L} \cup (x_i, y_i)$ until a fixed budget $B$ is exhausted $i := 0 \ldots B$.
+Active learning is defined as sequentially removing single instances $x_i \in \mathcal{U}_i$, requesting their label $y_i$ and adding them to the labeled pool $\mathcal{L}_{i+1} := \mathcal{L}_i \cup (x_i, y_i)$ until a fixed budget $B$ is exhausted $i := 0 \ldots B$.
 After each added instance the classification model is retrained according to section \ref{sec:training_the_classifier} and its performance is measured on a held-out test set.
 The quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
-We employ normalized AUC:
+We employ the normalized area under the accuracy curve (AUC):
 \begin{equation}
@@
 \end{equation}
-Where $\phi_i$ is the retrained classification model after the i-\textit{th} instance was selected
+Where $\phi_i$ is the retrained classification model after the i-\textit{th} instance was selected. \\ [1mm]
+We define the active learning process as an adapted reinforcement learning loop $(S, A, \tau, \Omega)$ where an environment will expose a state $s \in S$ to an agent $\Omega$, which will choose actions $a \in A$.
+One iteration consists of the environment sampling a subset of size $\tau$ of unlabeled instances $x_i \underset{\tau}{\sim} \mathcal{U}_i$, constructing the state $s_i := \omega(u_i)$ and presenting it to the agent to select an action $a_i := \Omega(s_i)$.
+The action $a_i$ is the index of the selected instance in $u_i$ out of all possible indices $A := [0 \ldots \tau]$.
+This process is repeated $B$ times $i := [0 \ldots B]$.
 

### 2023-02-17 core/agent.py

@@ -12,19 +12,15 @@ class BaseAgent(ABC):
         self.name = str(self.__class__).split('.')[-1][:-2]
         print(f"Loaded Agent: {self.name}")
 
-    @classmethod
-    @abstractmethod
-    def create_state_callback(cls, state_ids:list[int],
-                              x_unlabeled:Tensor,
-                              x_labeled:Tensor, y_labeled:Tensor,
-                              per_class_instances:dict,
-                              budget:int, added_images:int,
-                              initial_test_acc:float, current_test_acc:float,
-                              classifier:Module, optimizer:Optimizer)->Union[Tensor, dict]:
-        pass
 
     @abstractmethod
-    def predict(self, state:Union[Tensor, dict], greed:float=0.0)->Tensor:
+    def predict(self, state_ids:list[int],
+                      x_unlabeled:Tensor,
+                      x_labeled:Tensor, y_labeled:Tensor,
+                      per_class_instances:dict,
+                      budget:int, added_images:int,
+                      initial_test_acc:float, current_test_acc:float,
+                      classifier:Module, optimizer:Optimizer)->Union[Tensor, dict]:
         pass
 
 

### 2023-02-17 core/environment.py

@@ -12,7 +12,6 @@ class ALGame(gym.Env):
 
     def __init__(self, dataset:BaseDataset,
                  labeled_sample_size,
-                 create_state_callback:Callable,
                  device=None):
         if device is None:
             self.device = torch.device('cuda' if torch.cuda.is_available() else 'cpu')
@@ -23,7 +22,6 @@ class ALGame(gym.Env):
         self.sample_size = labeled_sample_size
         self.fitting_mode = dataset.class_fitting_mode
         self.loss = nn.CrossEntropyLoss()
-        self.create_state_callback = create_state_callback
 
         # set gym observation space and action space
         self.current_test_accuracy = 0.0
@@ -36,14 +34,14 @@ class ALGame(gym.Env):
         if isinstance(state, dict):
             self.observation_space = dict()
             for key, value in state.items():
-                self.observation_space[key] = gym.spaces.Box(-np.inf, np.inf, shape=[value.size(1),])
+                self.observation_space[key] = gym.spaces.Box(-np.inf, np.inf, shape=[len(value),])
         else:
-            self.observation_space = gym.spaces.Box(-np.inf, np.inf, shape=[state.size(1),])
+            self.observation_space = gym.spaces.Box(-np.inf, np.inf, shape=[len(state),])
         self.action_space = gym.spaces.Discrete(self.sample_size)
         self.spec = gym.envs.registration.EnvSpec("RlAl-v0", reward_threshold=np.inf, entry_point="ALGame")
 
 
-    def reset(self, *args, **kwargs)->Tuple[torch.Tensor, dict]:
+    def reset(self, *args, **kwargs)->list:
         with torch.no_grad():
             self.n_interactions = 0
             self.added_images = 0
@@ -59,25 +57,23 @@ class ALGame(gym.Env):
         return self.create_state()
 
 
-    def _send_state_to_device(self, state):
-        if isinstance(state, dict):
-            result = {}
-            for k,v in state.items():
-                result[k] = v.to(self.device)
-        else:
-            result = state.to(self.device)
-        return result
+    def _send_state_to_device(self, state:list):
+        for i, component in enumerate(state):
+            if isinstance(component, torch.Tensor):
+                state[i] = component.to(self.device)
+        return state
 
 
     def create_state(self):
-        state = self.create_state_callback(self.state_ids,
-                                           self.x_unlabeled,
-                                           self.x_labeled, self.y_labeled,
-                                           self.per_class_instances,
-                                           self.budget, self.added_images,
[…]

### 2023-02-17 evaluate.py

@@ -7,10 +7,10 @@ from core.helper_functions import *
 
 parser = argparse.ArgumentParser()
 parser.add_argument("--run_id", type=int, default=1)
-parser.add_argument("--agent", type=str, default="margin")
-parser.add_argument("--dataset", type=str, default="cifar10")
+parser.add_argument("--agent", type=str, default="entropy")
+parser.add_argument("--dataset", type=str, default="splice")
 parser.add_argument("--sample_size", type=int, default=20)
-parser.add_argument("--restarts", type=int, default=10)
+parser.add_argument("--restarts", type=int, default=50)
 args = parser.parse_args()
 
 numpy.random.seed(args.run_id)
@@ -23,7 +23,6 @@ dataset = DatasetClass(cache_folder="../datasets")
 dataset = dataset.to(util.device)
 env = core.ALGame(dataset,
                   args.sample_size,
-                  AgentClass.create_state_callback,
                   device=util.device)
 agent = AgentClass()
 base_path = os.path.join("runs", dataset.name, agent.name)
@@ -37,7 +36,7 @@ with core.EnvironmentLogger(env, log_path, util.is_cluster, args.restarts) as en
         dataset.reset()
         state = env.reset()
         while not done:
-            action = agent.predict(state)
+            action = agent.predict(*state)
             state, reward, done, truncated, info = env.step(action.item())
 
 # collect results from all runs

### 2023-02-21 doc/al_benchmark_paper.tex
upper bound

--- 2023-02-16 doc/al_benchmark_paper.tex
@@
 \subsection{Contribution}
-\begin{itemize}
-	\item Benchmark suit for Active Learning
-	\item Properly optimized classification models
-	\item Multiple use-cases: Single Dataset, Dataset Transfer
-	\item Multiple Domains: Tabular, Image, Text
-\end{itemize}
 
@@
 
-\subsection{Problem Description}
+\subsection{Problem Description / Delineation}
+\textbf{Basic Classification}\\
 We assume a dataset $\mathcal{D} := (\mathcal{U}, \mathcal{L})$ consisting of a large pool of unlabeled data $\mathcal{U}$ and a small set of labeled data called the seed set with $|\mathcal{L}| \ll |\mathcal{U|}$.
@@
 We consider only classification problems, hence the instances of a dataset have the form $x \in \mathbb{R}^M$ for $\mathcal{U}$ and $(x, y) \in (\mathbb{R}^M, \mathbb{R}^C)$ for $\mathcal{L}$.
-To perform classification, a model $\phi_\theta := \mathbb{R}^M \rightarrow \mathbb{R}^C$ is used.
-To fit the model, it is parameterized by $\theta$ and subjected to loss $l(y, \phi_\theta(x)) := \mathbb{R}^C \times \mathbb{R}^C \rightarrow \mathbb{R}$.
+To perform classification, a model $\hat y_\theta := \mathbb{R}^M \rightarrow \mathbb{R}^C$ is used.
+To fit the model, it is parameterized by $\theta$ and subjected to loss $l(y, \hat y_\theta(x)) := \mathbb{R}^C \times \mathbb{R}^C \rightarrow \mathbb{R}$.
 For this work, categorical cross-entropy (CE) is used.
 For evaluating classification performance, we use Accuracy \\ [1mm]
-Active learning is defined as sequentially selecting single instances of $x_i \in \mathcal{U}$, requesting their label $y_i$ and adding them to the labeled pool $\mathcal{L} := \mathcal{L} \cup (x_i, y_i)$ until a fixed budget $B$ is exhausted $i := 0 \ldots B$.
+%
+\textbf{Pool-based AL with single instances (non-batch setting)}\\
+Active learning is defined as sequentially removing single instances $x_i \in \mathcal{U}_i\hspace{1mm}; \hspace{2mm} \mathcal{U}_{i+1} := \mathcal{U}_i / \{x_i\}$, requesting their label $y_i$ and adding them to the labeled pool $\mathcal{L}_{i+1} := \mathcal{L}_i \cup (x_i, y_i)$ until a fixed budget $B$ is exhausted $i := 0 \ldots B$.
 After each added instance the classification model is retrained according to section \ref{sec:training_the_classifier} and its performance is measured on a held-out test set.
 The quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
-We employ normalized AUC:
+We employ the normalized area under the accuracy curve (AUC):
 \begin{equation}
-	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \phi, B) := \frac{1}{B} \sum_{i=1}^{B} acc(y_{test}, \phi_i(x_{test}))
+	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} acc(y_{test}, \hat y_i(x_{test}))
 \end{equation}
-Where $\phi_i$ is the retrained classification model after the i-\textit{th} instance was selected
+Where $\hat y_i$ is the retrained classification model after the i-\textit{th} instance was selected. \\ [1mm]
+%
+\textbf{Framing AL as RL}\\
+We define the active learning process as an adapted reinforcement learning loop $(S, A, \tau, \Omega, \omega)$ where an environment iteratively will expose a state $s \in S$ to an agent $\Omega$, which will choose actions $a \in A$.
+For each iteration $i$ the environment samples a subset of size $\tau$ of unlabeled instances $x_i \underset{\tau}{\sim} \mathcal{U}_i$, constructs the state $s_i := \omega(x_i)$ and presents it to the agent to select an action $a_i := \Omega(s_i)$.
+The action $a_i$ is the index of the selected instance in $u_i$ out of all possible indices $A := [0 \ldots \tau]$.
+This process is repeated $B$ times $i := [0 \ldots B]$.
 

### 2023-02-21 compute_upper_bound.py

@@ -6,8 +6,8 @@ from core.helper_functions import *
 
 parser = argparse.ArgumentParser()
 parser.add_argument("--run_id", type=int, default=1)
-parser.add_argument("--dataset", type=str, default="cifar10")
-parser.add_argument("--restarts", type=int, default=50)
+parser.add_argument("--dataset", type=str, default="splice")
+parser.add_argument("--restarts", type=int, default=5) # TODO
 args = parser.parse_args()
 
 numpy.random.seed(args.run_id)
@@ -24,13 +24,17 @@ save_meta_data(log_path, None, None, dataset)
 
 accuracies = []
 for _ in range(args.restarts):
-    acc = fit_and_evaluate(dataset)
-    accuracies.append(acc)
+    accs = fit_and_evaluate(dataset)
+    accuracies.append(accs)
+# save final accuracies for performance comparisons
 data_dict = {}
 for k, v in enumerate(accuracies):
-    data_dict[k] = [v]
+    data_dict[k] = [v[-1]]
 df = pd.DataFrame(data_dict)
 df.to_csv(os.path.join(log_path, "accuracies.csv"))
+
+# save learning curves for training evaluation
+plot_learning_curves(accuracies, os.path.join(log_path, "learning_curves.jpg"))
 
 # collect results from all runs
 collect_results(base_path, "run_")

### 2023-02-21 core/classifier.py

@@ -79,7 +79,7 @@ def fit_and_evaluate(dataset:BaseDataset,
                      lr:float=None, weight_decay:float=None, batch_size:int=None,
                      hidden_sizes:tuple=None,
                      disable_progess_bar:bool=False,
-                     max_epochs:int=1000):
+                     max_epochs:int=4000):
     loss = nn.CrossEntropyLoss()
     if hidden_sizes is not None:
         model = dataset.get_classifier(hidden_dims=hidden_sizes)
@@ -115,7 +115,8 @@ def fit_and_evaluate(dataset:BaseDataset,
                 self.best_loss = loss_val
             return False
 
-    early_stop = EarlyStopping()
+    all_accs = []
+    early_stop = EarlyStopping(patience=7)
     iterator = tqdm(range(max_epochs), disable=disable_progess_bar)
     for e in iterator:
         for batch_x, batch_y in train_dataloader:
@@ -139,6 +140,7 @@ def fit_and_evaluate(dataset:BaseDataset,
             if early_stop.check_stop(loss_sum):
                 print(f"Early stop after {e} epochs")
                 break
-            iterator.set_postfix({"val loss": loss_sum, "val acc": correct / total})
-    accuracy = correct / total
-    return accuracy
+            current_acc = correct / total
+            all_accs.append(current_acc)
+            iterator.set_postfix({"val loss": "%1.4f"%loss_sum, "val acc": "%1.4f"%current_acc})
+    return all_accs

### 2023-02-21 core/helper_functions.py

@@ -66,6 +66,20 @@ def plot_mean_std_development(inpt:list, title:str, out_file:str=None):
     plt.close(fig)
 
 
+def plot_learning_curves(list_of_accs:list, out_file:str=None):
+    for accs in list_of_accs:
+        x = range(len(accs))
+        plt.plot(x, accs, alpha=0.7)
+
+    plt.xlabel("epochs")
+    plt.ylabel("validation accuracy")
+    if out_file is None:
+        plt.show()
+    else:
+        plt.savefig(out_file, dpi=100, bbox_inches='tight')
+    plt.clf()
+
+
 def collect_results(base_path, folder_prefix):
     def check_for_nan_cols(df: pd.DataFrame):
         cleaned_pd = df.copy(deep=True)

### 2023-02-23 doc/al_benchmark_paper.tex
docs

@@
 \usepackage{algorithm}
-\usepackage{algorithmic}
+%\usepackage{algorithmic}
+\usepackage{algpseudocode}
 
@@
 \textbf{Basic Classification}\\
-We assume a dataset $\mathcal{D} := (\mathcal{U}, \mathcal{L})$ consisting of a large pool of unlabeled data $\mathcal{U}$ and a small set of labeled data called the seed set with $|\mathcal{L}| \ll |\mathcal{U|}$.
-For the purpose of this work, only fully labeled datasets are used and the labels for $\mathcal{U}$ are suppressed until they are selected for labeling.
-For evaluation purposes we also assume a held-out test set $(x_{test}, y_{test}) \in (\mathbb{R}^M, \mathbb{R}^C)$
-We consider only classification problems, hence the instances of a dataset have the form $x \in \mathbb{R}^M$ for $\mathcal{U}$ and $(x, y) \in (\mathbb{R}^M, \mathbb{R}^C)$ for $\mathcal{L}$.
-To perform classification, a model $\hat y_\theta := \mathbb{R}^M \rightarrow \mathbb{R}^C$ is used.
-To fit the model, it is parameterized by $\theta$ and subjected to loss $l(y, \hat y_\theta(x)) := \mathbb{R}^C \times \mathbb{R}^C \rightarrow \mathbb{R}$.
+We assume a dataset $\mathcal{D} := (x_i, y_i) \hspace{0.5mm}; \hspace{1mm} i := 0 \ldots N$ consisting of instances $x_i \in \mathbb{R}^M$ and corresponding $y_i \in \mathbb{R}^C$.
+For evaluation purposes we assume a held-out test set $\mathcal{D}^{\text{test}}$ with the same characteristics.
+We consider classification problems with one-hot encoded classes, hence $C$ models the number of classes.
+To perform classification, a model $\hat y_\theta : \mathbb{R}^M \rightarrow \mathbb{R}^C$ is used.
+To fit the model, it is parameterized by $\theta$ and subjected to loss $\ell: \mathbb{R}^C \times \mathbb{R}^C \rightarrow \mathbb{R}$.
 For this work, categorical cross-entropy (CE) is used.
-For evaluating classification performance, we use Accuracy \\ [1mm]
+For evaluating classification performance, we use accuracy on the test set $\operatorname{Acc}(\mathcal{D}^{\text{test}}, \hat{y}_\theta)$. \\ [1mm]
 %
 \textbf{Pool-based AL with single instances (non-batch setting)}\\
-Active learning is defined as sequentially removing single instances $x_i \in \mathcal{U}_i\hspace{1mm}; \hspace{2mm} \mathcal{U}_{i+1} := \mathcal{U}_i / \{x_i\}$, requesting their label $y_i$ and adding them to the labeled pool $\mathcal{L}_{i+1} := \mathcal{L}_i \cup (x_i, y_i)$ until a fixed budget $B$ is exhausted $i := 0 \ldots B$.
+To construct the active learning setting, we suppress the labels $y_i$ of $\mathcal{D}$ to form the unlabeled pool $\mathcal{U} := x_i \hspace{0.5mm}; \hspace{1mm} i := 0 \ldots N$ and form and initial labeled pool $\mathcal{L}$ by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label.
+The result of this so-called "seeding" process is $\mathcal{L} := (x_i, y_i) \hspace{0.5mm}; \hspace{1mm} i := 0 \ldots k*C$. \\
+Active learning is defined as sequentially removing single instances $x_i \in \mathcal{U}_i\hspace{1mm}; \hspace{2mm} \mathcal{U}_{i+1} := \mathcal{U}_i \setminus´ \{x_i\}$, recovering their label $y_i$ and adding them to the labeled pool $\mathcal{L}_{i+1} := \mathcal{L}_i \cup (x_i, y_i)$ until a fixed budget $B$ is exhausted $i := 0 \ldots B$.
 After each added instance the classification model is retrained according to section \ref{sec:training_the_classifier} and its performance is measured on a held-out test set.
@@
 \begin{equation}
-	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} acc(y_{test}, \hat y_i(x_{test}))
+	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(y_{test}, \hat y_i(x_{test}))
 \end{equation}
@@
 We define the active learning process as an adapted reinforcement learning loop $(S, A, \tau, \Omega, \omega)$ where an environment iteratively will expose a state $s \in S$ to an agent $\Omega$, which will choose actions $a \in A$.
-For each iteration $i$ the environment samples a subset of size $\tau$ of unlabeled instances $x_i \underset{\tau}{\sim} \mathcal{U}_i$, constructs the state $s_i := \omega(x_i)$ and presents it to the agent to select an action $a_i := \Omega(s_i)$.
-The action $a_i$ is the index of the selected instance in $u_i$ out of all possible indices $A := [0 \ldots \tau]$.
-This process is repeated $B$ times $i := [0 \ldots B]$.
+For each iteration $i$ the environment samples a subset of size $\tau$ of unlabeled instances $x^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$, constructs the state $s^{(i)} := \omega(x^{(i)})$ and presents it to the agent to select an action $a^{(i)} := \Omega(s^{(i)})$.
+The action $a^{(i)}$ is the index of the selected instance in $u^{(i)}$ out of all possible indices $A := [1 \ldots \tau]$.
+This process is repeated $B$ times $i := [1 \ldots B]$.
+
+\begin{minipage}{0.59\linewidth}
+	\begin{algorithm}[H]
+		\caption{Active Learning}\label{alg:active_learning}
+		\begin{algorithmic}[1]
+			\Require $\mathcal{U}$ \Comment{Unlabeled Pool}
+			\Require $\tau$ \Comment{Unlabeled Sample Size}
+			\Require $\Omega$ \Comment{AL Agent}
+			\Require $\omega$ \Comment{Environment State function}
+			\State $\mathcal{L}^{(1)} \gets \operatorname{seed}(\mathcal{U})$  \Comment{Create the initial labeled set}
+			\State $\mathcal{U}^{(1)} \gets \mathcal{U}$
+			\For{$i := 1 \ldots B$}
+			\State $u^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$
+			\State $s^{(i)} \gets \omega(u^{(i)})$
+			\State $a^{(i)} \gets \Omega(s^{(i)})$ \Comment{$a^{(i)}$ is an index inside of $u^{(i)}$}
+			\State $y^{(i)} \gets \operatorname{oracle}(u^{(i)}_{a})$ \Comment{$u^{(i)}_{a}$ is shorthand for $u^{(i)}_{a^{(i)}}$}
+			\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_a, y^{(i)})\}$
+			\State $\mathcal{U}^{(i+1)} \gets \mathcal{U}^{(i)} \setminus \{u^{(i)}_a\}$
+			\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i+1)})$  \Comment{$\operatorname{Retrain}(\mathcal{L}^{(i+1)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i+1)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
+			\EndFor
+			\State
+			\Return $\frac{1}{B} \sum_{i=1}^{B} \text{acc}^{(i)}$
+		\end{algorithmic}
+	\end{algorithm}
+\end{minipage}
+\hspace{3mm}
+\begin{minipage}{0.35\linewidth}
+	\begin{algorithm}[H]
+		\caption{Retrain}\label{alg:retrain}
+		\begin{algorithmic}[1]
+			\Require $\mathcal{L}$ \Comment{Labeled Pool}
+			\Require $\mathcal{L}^\text{test}$ \Comment{Labeled Test Data}
+			\Require $\hat y_\theta$ \Comment{Classification Model}
+			\Require $e^\text{max}$ \Comment{Maximum Epochs}
+			\State $\text{loss}^* \gets \infty$
+			\For{$i := 1 \ldots e^{\text{max}}$}
+			\State $\theta_{i+1} \gets \theta_i - \eta \nabla_\theta \ell(\mathcal{L}, \hat y_{\theta})$
+			\State $\text{loss}_i \gets \ell(\mathcal{L}^\text{test}, \hat y_{\theta})$
+			\If{$\text{loss}_i < \text{loss}^*$}
+			\State $\text{loss}^* \gets \text{loss}_i$
+			\Else
+			\State Break
+			\EndIf
+			\EndFor
+			\State
+			\Return Acc($\mathcal{L}^\text{test}, \hat y_{\theta}$)
+		\end{algorithmic}
+	\end{algorithm}
+\end{minipage}
 
@@
 
+\appendix
+
+\section{Comparison of different sample sizes $\tau$}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\linewidth]{img/tau_ablation.png}
+\end{figure}
+
 \end{document}

### 2023-02-24 doc/al_benchmark_paper.tex
data folder rework

@@
 \textbf{Basic Classification}\\
-We assume a dataset $\mathcal{D} := (x_i, y_i) \hspace{0.5mm}; \hspace{1mm} i := 0 \ldots N$ consisting of instances $x_i \in \mathbb{R}^M$ and corresponding $y_i \in \mathbb{R}^C$.
+We assume a dataset $\mathcal{D} := (x_i, y_i) \hspace{0.5mm}; \hspace{1mm} i := 1 \ldots N$ consisting of instances $x_i \in \mathbb{R}^M$ and corresponding $y_i \in \mathbb{R}^C$.
 For evaluation purposes we assume a held-out test set $\mathcal{D}^{\text{test}}$ with the same characteristics.
@@
 \textbf{Pool-based AL with single instances (non-batch setting)}\\
-To construct the active learning setting, we suppress the labels $y_i$ of $\mathcal{D}$ to form the unlabeled pool $\mathcal{U} := x_i \hspace{0.5mm}; \hspace{1mm} i := 0 \ldots N$ and form and initial labeled pool $\mathcal{L}$ by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label.
-The result of this so-called "seeding" process is $\mathcal{L} := (x_i, y_i) \hspace{0.5mm}; \hspace{1mm} i := 0 \ldots k*C$. \\
-Active learning is defined as sequentially removing single instances $x_i \in \mathcal{U}_i\hspace{1mm}; \hspace{2mm} \mathcal{U}_{i+1} := \mathcal{U}_i \setminus´ \{x_i\}$, recovering their label $y_i$ and adding them to the labeled pool $\mathcal{L}_{i+1} := \mathcal{L}_i \cup (x_i, y_i)$ until a fixed budget $B$ is exhausted $i := 0 \ldots B$.
-After each added instance the classification model is retrained according to section \ref{sec:training_the_classifier} and its performance is measured on a held-out test set.
+To construct the active learning setting, we suppress the labels $y_i$ of $\mathcal{D}$ to form the unlabeled pool $\mathcal{U} := u_i \hspace{0.5mm}; \hspace{1mm} i := 1 \ldots N$ and form and initial labeled pool $\mathcal{L}$ by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label.
+The result of this so-called "seeding" process is $\mathcal{L} := (u_i, y_i) \hspace{0.5mm}; \hspace{1mm} i := 1 \ldots k*C$. \\
+Active learning is defined as sequentially removing single instances $u^{(i)} \in \mathcal{U}^{(i)}\hspace{1mm}; \hspace{2mm} \mathcal{U}^{(i+1)} := \mathcal{U}^{(i)} \setminus´ \{u^{(i)}\}$, recovering their label $y^{(i)}$ and adding them to the labeled pool $\mathcal{L}^{(i+1)} := \mathcal{L}^{(i)} \cup (u^{(i)}, y^{(i)})$ until a fixed budget $B$ is exhausted $i := 1 \ldots B$.
+After each added instance the classification model is retrained according to section \ref{sec:training_the_classifier} and its performance is measured on the held-out test set $\mathcal{D}^{\text{test}}$.
 The quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
@@
 We define the active learning process as an adapted reinforcement learning loop $(S, A, \tau, \Omega, \omega)$ where an environment iteratively will expose a state $s \in S$ to an agent $\Omega$, which will choose actions $a \in A$.
-For each iteration $i$ the environment samples a subset of size $\tau$ of unlabeled instances $x^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$, constructs the state $s^{(i)} := \omega(x^{(i)})$ and presents it to the agent to select an action $a^{(i)} := \Omega(s^{(i)})$.
+For each iteration $i$ the environment samples a subset of size $\tau$ of unlabeled instances $u^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$, constructs the state $s^{(i)} := \omega(u^{(i)})$ and presents it to the agent to select an action $a^{(i)} := \Omega(s^{(i)})$.
 The action $a^{(i)}$ is the index of the selected instance in $u^{(i)}$ out of all possible indices $A := [1 \ldots \tau]$.

### 2023-03-15 doc/al_benchmark_paper.tex
Docs

@@
 			\For{$i := 1 \ldots B$}
+				\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$  \Comment{$\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
 				\State $u^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$
@@
 				\State $a^{(i)} \gets \Omega(s^{(i)})$ \Comment{$a^{(i)}$ is an index inside of $u^{(i)}$}
-			\State $y^{(i)} \gets \operatorname{oracle}(u^{(i)}_{a})$ \Comment{$u^{(i)}_{a}$ is shorthand for $u^{(i)}_{a^{(i)}}$}
+				\State $y^{(i)} \gets \operatorname{label}(u^{(i)}_{a})$ \Comment{$u^{(i)}_{a}$ is shorthand for $u^{(i)}_{a^{(i)}}$}
 				\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_a, y^{(i)})\}$
 				\State $\mathcal{U}^{(i+1)} \gets \mathcal{U}^{(i)} \setminus \{u^{(i)}_a\}$
-			\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i+1)})$  \Comment{$\operatorname{Retrain}(\mathcal{L}^{(i+1)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i+1)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
 			\EndFor
@@
 
+\begin{algorithm}
+	\caption{Oracle}\label{alg:oracle}
+	\begin{algorithmic}[1]
+		\Require $\mathcal{U}$ \Comment{Unlabeled Pool}
+		\Require $\tau$ \Comment{Unlabeled Sample Size}
+		\Require $\Omega$ \Comment{AL Agent}
+		\Require $\omega$ \Comment{Environment State function}
+		\State $\mathcal{L}^{(1)} \gets \operatorname{seed}(\mathcal{U})$  \Comment{Create the initial labeled set}
+		\State $\mathcal{U}^{(1)} \gets \mathcal{U}$
+		\For{$i := 1 \ldots B$} 
+			\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$  \Comment{$\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
+			\State $u^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$
+			\State $r* \gets -\infty$
+			\State $j* \gets -1$
+			\For{$j := 1 \ldots \tau$} \Comment{Testing every unlabeled point}
+				\State $y^{(j)} \gets \operatorname{label}(u^{(i)}_{j})$
+				\State $\mathcal{L}^{(j)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_j, y^{(j)})\}$
+				\State $\text{acc}^{(j)} \gets \operatorname{Retrain}(\mathcal{L}^{(j)})$  
+				\State $r^{(j)} \gets \text{acc}^{(j)} - \text{acc}^{(i)}$
+				\If{$r^{(j)} > r^*$} \Comment{Select point with largest increase in performance}
+					\State $r* \gets r^{(j)}$
+					\State $j* \gets j$
+				\EndIf  
+			\EndFor
+			\State $y^{(i)} \gets \operatorname{label}(u^{(i)}_{j*})$
+			\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_{j*}, y^{(i)})\}$
+			\State $\mathcal{U}^{(i+1)} \gets \mathcal{U}^{(i)} \setminus \{u^{(i)}_{j*}\}$
+		\EndFor
+		\State
+		\Return $\frac{1}{B} \sum_{i=1}^{B} \text{acc}^{(i)}$
+	\end{algorithmic}
+\end{algorithm}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\linewidth]{img/oracle_performance.png}
+	\caption{Performance drop for Oracle in late stages of Active Learning.}
+\end{figure}
+
 \section{Methodology}
@@
 \subsection{State Space}
-Since every AL agent needs a different state space our environment exposes a callback-function that gives the agent full control over how the state is constructed. \\ [1mm]
-The callback includes the following information:
+Since every AL agent needs a different state space our environment exposes a full state to the agent, so that the agent has full control of what information will be used. \\ [1mm]
+The state can include the following information:
 \begin{itemize}
-	\item The current sample of IDs that point to the presented unlabeled datapoints
 	\item The entire labeled dataset $\mathcal{L}$
@@
 	\item The current optimizer including it's full state
+	\item The current sample of unlabeled points 
 \end{itemize}
-Every agent needs to implement this callback that transforms the given information into a state tensor that will be directly consumed by the agent to make it's prediction.
 

### 2023-03-15 experiment_util.py

@@ -16,5 +16,6 @@ os.environ["HDF5_USE_FILE_LOCKING"]="FALSE"
 
 import torch
 device = torch.device('cuda' if torch.cuda.is_available() else 'cpu')
+print(f"Used Torch Device: {device}")
 is_cluster = 'miniconda3' in sys.prefix
 

### 2023-04-06 doc/al_benchmark_paper.tex
cifar10 experiments

@@
 \usepackage{algorithm}
-%\usepackage{algorithmic}
 \usepackage{algpseudocode}
@@
 \section{Related Work}
+Many different algorithms have been proposed for active learning. 
+In this work we focus on those approaches that have shown consistent results over the years as well as newer approaches that have demonstrated significant lifts in their initial experiments.
+AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
+Geometric approaches include CoreSet \cite{sener2017active} and TypiClust \cite{hacohen2022active}, which use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
+Uncertainty-based approaches include classic uncertainty sampling (based on Shannon-Entropy and the margin-score), BALD \cite{kirsch2019batchbald} and BADGE \cite{ashdeep}, which use metrics to measure the classifiers state.
+\\ [1mm]
+%
+Some previous work also aimed to provide a benchmark suite
 
@@
 \begin{itemize}
-	\item Weird drop of performance for multiples of batch size (drop\_last in DataLoader)
 	\item Reduction of the test set for speed

### 2023-04-06 evaluate.py

@@ -12,8 +12,8 @@ parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--agent_seed", type=int, default=1)
 parser.add_argument("--pool_seed", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
-parser.add_argument("--agent", type=str, default="coreset")
-parser.add_argument("--dataset", type=str, default="splice")
+parser.add_argument("--agent", type=str, default="margin")
+parser.add_argument("--dataset", type=str, default="cifar10")
 parser.add_argument("--encoded", type=bool, default=True)
 parser.add_argument("--sample_size", type=int, default=20)
 parser.add_argument("--restarts", type=int, default=50)

### 2023-04-11 ray_tune.py
cifar10 MLP tune

@@ -15,8 +15,8 @@ from ray import tune
 
 parser = argparse.ArgumentParser()
 parser.add_argument("--data_folder", type=str, required=True)
-parser.add_argument('--dataset', type=str, default="dna")
-parser.add_argument('--num_trials', type=int, default=100)
+parser.add_argument('--dataset', type=str, default="cifar10")
+parser.add_argument('--num_trials', type=int, default=200)
 parser.add_argument('--max_conc_trials', type=int, default=15)
 
 def evaluate_encoded_classification_config(raytune_config, DatasetClass, config_file, cache_folder, benchmark_folder):

### 2023-04-12 doc/al_benchmark_paper.tex
cifar10 MLP tune
cifar10 MLP head

@@
 \usepackage{algpseudocode}
+\usepackage{color}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Related Work}
+{\color{red} Version: Braindump}\\
 Many different algorithms have been proposed for active learning. 
@@
 %
-Some previous work also aimed to provide a benchmark suite
+Some previous work also aimed to provide a benchmark suite for active learning:
+The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
+While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call "Label Efficiency" and provide experiments on many common configurations for data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Overview}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Problem Description / Delineation}
@@
 We employ the normalized area under the accuracy curve (AUC):
-\begin{equation}
+\begin{equation}\label{eq:auc}
 	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(y_{test}, \hat y_i(x_{test}))
@@
 	\begin{algorithm}[H]
-		\caption{Active Learning}\label{alg:active_learning}
+		\caption{Active Learning {\color{red} OLD}}\label{alg:active_learning}
 		\begin{algorithmic}[1]
@@
 \begin{algorithm}
-	\caption{Oracle}\label{alg:oracle}
+	\caption{Oracle {\color{red} OLD}}\label{alg:oracle}
 	\begin{algorithmic}[1]
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\linewidth]{img/oracle_performance.png}
-	\caption{Performance drop for Oracle in late stages of Active Learning.}
-\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Datasets}
+{\color{red} Version: 1.0}\\
+In this work we use 4 different datasets from two domains.
+For vector data, we use DNA and Splice (Source: \cite{libsvmtools}) and for image data we use FashionMNIST \cite{xiao2017fashion} and Cifar10 \cite{krizhevsky2009learning}. \\
+We would like to point out that these datasets can be considered "toy-datasets" and therefore not relevant for practical purposes.
+This might be true if we aimed to develop novel classification models on these dataset, however we are solely focused on comparing different AL algorithms in this paper.
+Our core assumption is that a well-performing algorithm in our benchmark will transfer well into more practical use-cases. \\ [1mm]
+Adapting the experimental setting from \cite{hacohen2022active} we offer all our datasets in the raw setting as well as pre-encoded by a fixed embedding model that was trained in an unsupervised manner.
+The pre-encoded datasets offer multiple advantages in that they require smaller classification models (i.e.
+small MLPs) and in general much less sampled datapoints to archive close to upper bound performance (measured with the full dataset).
+They unify experimental setups for different datasets and therefore make them much more comparable.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methodology}
 
-\subsection{Classification Model}
-The classifier is constructed according to two kinds of information.
-The general class of model (Dense, Convolutional, Attention, ...), and the configuration of the model (number of layers, size of each layer, ...). \\
-The model class and exact configuration is determined by the dataset, i.e. tabular datasets will prescribe a dense model.
-If special capabilities of the model are needed (i.e. Monte-Carlo Dropout), an extension of the given model class can be provided to the framework. \\ [1mm]
-To ensure comparability between models, the model's configuration should not be changed or an additional evaluation of the new configuration should be conducted to compare the baseline expressivity.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{State Space}
+{\color{red} Version: 0.1}\\
 Since every AL agent needs a different state space our environment exposes a full state to the agent, so that the agent has full control of what information will be used. \\ [1mm]
@@
 \begin{itemize}
-	\item The entire labeled dataset $\mathcal{L}$
-	\item The entire unlabeled dataset $\mathcal{U}$
+	\item The entire labeled dataset $\mathcal{L}^{(i)}$
+	\item The entire unlabeled dataset $\mathcal{U}^{(i)}$
 	\item A histogram of labeled points per class (count)
-	\item The available budget
+	\item The available budget $B$
 	\item Number of added datapoints $|\mathcal{L}| - |\mathcal{S}|$
-	\item The initial validation accuracy and current validation accuracy
-	\item The current classification model including all model weights
+	\item The initial validation accuracy $\operatorname{acc}^{(0)}$ and current validation accuracy $\operatorname{acc}^{(i)}$
+	\item The current classification model including all model weights $\theta^{(i)}$
 	\item The current optimizer including it's full state
-	\item The current sample of unlabeled points 
+	\item The current sample of unlabeled points $u^{(i)}$
 \end{itemize}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Training the Classifier}\label{sec:training_the_classifier}
+Generally, the classification model can be trained in two ways.
+Either you reset the parameters after each AL iteration and train the classifier from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or you retain the previous state and only fine-tune the classifier on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
+In this work we use the fine-tuning method for raw datasets to save computation, while we use the from-scratch training for embedded dataset, since they have very small classifiers and this method generally produces better results.
+Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping after that.
+The early stopping has patience 0, so it will stop as soon as the validation loss does no longer decrease.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Evaluation}\label{sec:evaluation}
+{\color{red} Version: 0.1}\\
+We compare different AL algorithms based on their archived AUC score (Eq. \ref{eq:auc}).
+This score does incorporate performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
+A good AL algorithm should be able to perform well in both scenarios. \\
+{\color{red} TODO}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Splice Embedded}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\linewidth]{img/eval_splice_embedded}
+	\caption{Results for all algorithms on the pre-encoded Splice dataset}
+\end{figure}
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\linewidth]{img/eval_splice_embedded_cd}
+\caption{Critical Difference Diagram for Splice where every restart of the algorithm is one sample for the Wilcoxon-Holm method}
+\end{figure}
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\linewidth]{img/eval_macro}
+\caption{Critical Difference Diagram for Splice, DNA, USPS and Cifar10}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Ablation Studies}
 \begin{itemize}
+	\item Setting $\tau$ to $|\mathcal{U}|$
 	\item Reduction of the test set for speed

### 2023-04-12 compute_upper_bound.py
pretext task now optimizes for linear head classification instead of kNN

@@ -8,7 +8,7 @@ parser = argparse.ArgumentParser()
 parser.add_argument("--data_folder", type=str, required=True)
 parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
-parser.add_argument("--dataset", type=str, default="usps")
+parser.add_argument("--dataset", type=str, default="cifar10")
 parser.add_argument("--encoded", type=bool, default=True)
 parser.add_argument("--restarts", type=int, default=3)
 args = parser.parse_args()

### 2023-04-12 core/helper_functions.py

@@ -14,12 +14,14 @@ from core.data import BaseDataset
 import datasets
 
 class EarlyStopping:
-    def __init__(self, patience=7):
+    def __init__(self, patience=7, lower_is_better=True):
         self.patience = patience
-        self.best_loss = torch.inf
+        self.lower_is_better = lower_is_better
+        self.best_loss = torch.inf if lower_is_better else -torch.inf
         self.steps_without_improvement = 0
     def check_stop(self, loss_val):
-        if loss_val >= self.best_loss:
+        if (self.lower_is_better     and loss_val >= self.best_loss) or \
+           (not self.lower_is_better and loss_val <= self.best_loss):
             self.steps_without_improvement += 1
             if self.steps_without_improvement > self.patience:
                 return True

### 2023-04-12 sim_clr/evaluate.py

@@ -1,5 +1,7 @@
 from training import AverageMeter
 import torch
+import torch.nn as nn
+from core.helper_functions import EarlyStopping
 
 @torch.no_grad()
 def contrastive_evaluate(val_loader, model, memory_bank, device):
@@ -17,3 +19,48 @@ def contrastive_evaluate(val_loader, model, memory_bank, device):
         top1.update(acc1.item(), images.size(0))
 
     return top1.avg
+
+
+def linear_evaluate(train_loader, val_loader, model, embedding_dim, n_classes, device):
+    model.eval()
+    class_head = nn.Linear(embedding_dim, n_classes).to(device)
+    optimizer = torch.optim.Adam(class_head.parameters(), lr=0.05)
+    loss = nn.CrossEntropyLoss().to(device)
+    max_acc = 0.0
+    all_accs = []
+    early_stop = EarlyStopping(patience=5, lower_is_better=False)
+    for epoch in range(30):
+        for batch in train_loader:
+            x = batch['image']
+            y = batch['target']
+            with torch.no_grad():
+                x_enc = model(x).detach()
+            y_hat = class_head(x_enc)
+            loss_value = loss(y_hat, y)
+            optimizer.zero_grad()
+            loss_value.backward()
+            optimizer.step()
+        # validation
+        with torch.no_grad():
+            total = 0.0
+            correct = 0.0
+            for batch in val_loader:
+                batch_x, batch_y = batch["image"], batch["target"]
+                x_enc = model(batch_x)
+                y_hat = class_head(x_enc)
+                predicted = torch.argmax(y_hat, dim=1)
+                total += batch_y.size(0)
+                correct += (predicted == batch_y).sum().item()
+                # class_loss = self.loss(y_hat, torch.argmax(batch_y.long(), dim=1))
+                # loss_sum += class_loss.detach().cpu().numpy()
+            accuracy = correct / total
+            all_accs.append(accuracy)
+            # early stop on test with patience of 0
+            if early_stop.check_stop(accuracy):
+                # print("early stop", len(all_accs))
+                # print(["%1.4f"%(a) for a in all_accs])
+                return accuracy * 100.0
+            max_acc = max(accuracy, max_acc)
+    # print(["%1.4f"%(a) for a in all_accs[:10]])
+    # print(["%1.4f"%(a) for a in all_accs[-10:]])
+    return accuracy * 100.0

### 2023-04-14 download_all_datasets.py
fashion mnist encoder training GPU fix

@@ -1,5 +1,7 @@
 from core.helper_functions import get_dataset_by_name
 import argparse
+import yaml
+import numpy as np
 
 
 parser = argparse.ArgumentParser()
@@ -17,8 +19,11 @@ all_names = [
 for name in all_names:
     print("##########################################")
     print(f"downloading {name}...")
+    with open(f"configs/{name}.yaml", 'r') as f:
+        config = yaml.load(f, yaml.Loader)
+    pool_rng = np.random.default_rng(1)
     DatasetClass = get_dataset_by_name(name)
-    dataset = DatasetClass(cache_folder=args.data_folder)
+    dataset = DatasetClass(args.data_folder, config, pool_rng, encoded=False)
 
 print("\n")
 print("> all datasets downloaded")

### 2023-04-17 doc/al_benchmark_paper.tex
pretext task now optimizes for linear head classification instead of kNN
fashion mnist encoder training GPU fix
fashion mnist pretext 800 epochs

@@
 While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call "Label Efficiency" and provide experiments on many common configurations for data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
+The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
+They test for several different properties of the models including robustness, response to compression techniques and final performance.
 
@@
 \section{Overview}
+Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with an extensive experiments section.
+\\
+{\color{red} TODO} We include in this table any methodological paper that experiments on at least two domains.
+\begin{table}[h]
+	\centering
+	\begin{tabular}{l | c c c c c}
+		Paper & \# Datasets & Domains & Scenarios & Oracle & RL Alg. \\
+		\hline
+		Beck et al. \cite{beck2021effective} & 4 & 1 & 3 & - & - \\
+		Hu et al. \cite{hu2021towards} & 5 & 2 & 1 & - & - \\
+		Li et al. \cite{li2022empirical} & 5 & 1 & 1 & - & - \\
+		\textbf{Ours} & 5 & 2 & 2 & \checkmark & -
+	\end{tabular}
+	\caption{Comparison of our benchmark with the existing literature}
+	\label{tab:benchmark_comparison}
+\end{table}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Reproducibility}\label{sec:reproducibility}
+A big focus in this work is to provide an experimental setup that is fully reproducible independent from the dataset, classification model or AL algorithm used.
+In our opinion, an evaluation on one dataset with a given seed should always be done on the same sequence of batches $x_0, \ldots, x_B$.
+Even though different AL algorithms will pick different samples from these batches, making them unavailable for sampling in later batches, the theoretical decision tree for every possible choice in every iteration $i$ should stay the same.
+This behavior is not possible with the default choice of setting a global seed at the start of the experiment, since a single additional random draw from the random number generator completely changes the decision tree for the batches.
+This additional random number might be drawn during the initialization of the classification model, or the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
+This problem also applies to the initialization of the classification models $\theta$, the initialization and querying of $Omega$ and the drawn mini batches during the training of $\theta$.
+The desired control over all these processes can be archived by assigning a separate seeded random number generator to all these processes.
+To the best of our knowledge, we are the first work that discusses this issue and proposes a solution for it.
+We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-04-17 doc/al_benchmark_paper.tex
dropout classifier

@@
 \\
-{\color{red} TODO} We include in this table any methodological paper that experiments on at least two domains.
+{\color{red} TODO} We include in this table any methodological paper that experiments on at least two domains. \\
+{\color{red} TODO} Define AL scenarios (really hard)
 \begin{table}[h]
@@
 \subsection{Problem Description / Delineation}
+\textbf{Lars' Version} \\
+Given
+\begin{itemize}
+	\item a number $B\in\mathcal{N}$ (called budget),
+	\item two spaces $\mathcal{X}$ and $\mathcal{Y}$,  {\tiny e.g., $\mathcal{X}:=\mathcal{R}^M, \mathcal{Y}:=\mathcal{R}^T$},
+	\item a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of
+	sequences of pairs $(x,y)$  from an unknown distribution $p$
+	(called datasets),
+	%  {\tiny with $|\mathcal{D}_n|\geq B$ for all $n\in 1{:}N$,}
+	{\tiny with $p(\mathcal{D})=0$ for $|\mathcal{D}|<B$,}
+	\item a function $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ (called loss), and
+	\item a function $A:  (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$
+	(called learning algorithm),
+\end{itemize}
+find a function
+\begin{align*}
+	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
+	\quad\quad \text{\tiny (equivariant in the second argument)}
+\end{align*}
+% which is equivariant in the second argument, 
+called acquisition function,
+s.t. the expected loss of a model learned on
+all predictors 
+plus $B$ sequentially acquired
+targets is minimal:
+\begin{align*}
+	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
+	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
+	% \operatorname{avg} \{
+	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
+	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
+	\ell(y, \hat y(x)) 
+	\\
+	\text{with }
+	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
+	\\ 
+	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
+	\quad b\in 1{:}B
+\end{align*}
+\textbf{Notes:}
+\begin{itemize}
+	\item We would need to switch from lowest expected loss to highest AUC
+	\item 
+\end{itemize}
+\vspace{5mm}
+\textbf{Thorben's Version}\\
 \textbf{Basic Classification}\\
@@
 \subsection{Reproducibility}\label{sec:reproducibility}
+{\color{red} Version: 1.0}\\
 A big focus in this work is to provide an experimental setup that is fully reproducible independent from the dataset, classification model or AL algorithm used.
@@
 We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Comparibility}\label{sec:comparibility}
+{\color{red} Version: braindump}\\
+We split the evaluation model that generates accuracy scores and the model that generates features for the AL agent. \\
+This allows for dropout / ensemble approaches to invest more compute to generate features \\
+The normal evaluation model is available as a fallback of course \\
+This way, the comparison is always fair
 

### 2023-04-17 agents/bald.py

@@ -5,12 +5,70 @@ from torch import Tensor
 from torch.nn import Module
 from torch.optim import Optimizer
 from core.agent import BaseAgent
+from core.classifier import construct_model
 
 class BALD(BaseAgent):
 
-    def __init__(self, agent_rng, dropout_trials=5):
-        super().__init__(agent_rng)
+    def __init__(self, agent_rng, config, dropout_trials=5):
+        super().__init__(agent_rng, config)
+        assert "current_run_info" in config and "embedded" in config["current_run_info"]
         self.dropout_trials = dropout_trials
+        self.dropout_model = None
+        self.dropout_model_rng = torch.Generator()
+        self.dropout_model_rng.manual_seed(1)
+
+
+    def _create_dropout_model(self, x_shape, y_shape):
+        if self.config["current_run_info"]["embedded"]:
+            self.dropout_model, _ = construct_model(self.dropout_model_rng, x_shape, y_shape[-1], self.config["classifier_embedded"])
+        else:
+            self.dropout_model, _ = construct_model(self.dropout_model_rng, x_shape, y_shape[-1], self.config["classifier"])
+        self.initial_weights = self.dropout_model.state_dict()
+
+    def _fit_dropout_model(self, x, y):
+        if self.config["current_run_info"]["embedded"]:
+            self.dropout_model.load_state_dict(self.initial_weights)
+
+        drop_last = 64 < len(self.x_labeled)
+        train_dataloader = DataLoader(TensorDataset(self.x_labeled, self.y_labeled),
+                                      batch_size=self.dataset.classifier_batch_size,
+                                      drop_last=drop_last,
+                                      generator=self.data_loader_rng,
+                                      # num_workers=4, # dropped for CUDA compat
+                                      shuffle=True)
+        test_dataloader = DataLoader(TensorDataset(self.dataset.x_test, self.dataset.y_test), batch_size=512,
+                                     # num_workers=4 # dropped for CUDA compat
+                                     )
+
+        lastLoss = torch.inf
+        for e in range(epochs):
+            for batch_x, batch_y in train_dataloader:
+                yHat = self.classifier(batch_x)
+                loss_value = self.loss(yHat, batch_y)
+                self.optimizer.zero_grad()
+                loss_value.backward()
+                self.optimizer.step()
+            # early stopping on test
+            with torch.no_grad():
+                loss_sum = 0.0
[…]

### 2023-04-17 core/data.py

@@ -145,44 +145,20 @@ class BaseDataset(ABC):
                    dataset["x_test"], dataset["y_test"]
         return None
 
-    def _construct_model(self, model_rng, model_config, add_head=True) -> Tuple[Module, int]:
-        '''
-        Constructs the model by name and additional parameters
-        Returns model and its output dim
-        '''
-        model_type = model_config["type"].lower()
-        if model_type == "linear":
-            from core.classifier import SeededLinear
-            return nn.Sequential(SeededLinear(model_rng, self.x_shape[-1], self.n_classes)), \
-                   self.n_classes
-        elif model_type == "resnet18":
-            from core.resnet import ResNet18
-            return ResNet18(num_classes=self.n_classes, in_channels=self.x_shape[0],
-                            add_head=add_head), \
-                   self.n_classes if add_head else 512
-        elif model_type == "mlp":
-            from core.classifier import DenseModel
-            return DenseModel(model_rng,
-                              input_size=self.x_shape[-1],
-                              num_classes=self.n_classes,
-                              hidden_sizes=model_config["hidden"],
-                              add_head=add_head), \
-                   self.n_classes if add_head else model_config["hidden"][-1]
-        else:
-            raise NotImplementedError
-
     def get_classifier(self, model_rng) -> Module:
+        from core.classifier import construct_model
         if self.encoded:
-            model, _ = self._construct_model(model_rng, self.config["classifier_embedded"])
+            model, _ = construct_model(model_rng, self.x_shape, self.n_classes, self.config["classifier_embedded"])
         else:
-            model, _ = self._construct_model(model_rng, self.config["classifier"])
+            model, _ = construct_model(model_rng, self.x_shape, self.n_classes, self.config["classifier"])
         return model
 
     def get_pretext_encoder(self, config: dict, seed=1) -> nn.Module:
         from sim_clr.encoder import ContrastiveModel
+        from core.classifier import construct_model
         model_rng = torch.Generator()
         model_rng.manual_seed(seed)
-        backbone, out_dim = self._construct_model(model_rng, config["pretext_encoder"], add_head=False)
+        backbone, out_dim = construct_model(model_rng, self.x_shape, self.n_classes, config["pretext_encoder"], add_head=False)
         config["pretext_encoder"]["encoder_dim"] = out_dim
         model = ContrastiveModel({'backbone': backbone, 'dim': config["pretext_encoder"]["encoder_dim"]},
                                  head="mlp", features_dim=config["pretext_encoder"]["feature_dim"])

### 2023-04-17 evaluate.py

@@ -26,6 +26,8 @@ max_run_id = run_id + args.restarts
 while run_id < max_run_id:
     with open(f"configs/{args.dataset}.yaml", 'r') as f:
         config = yaml.load(f, yaml.Loader)
+    config["current_run_info"] = dict()
+    config["current_run_info"]["embedded"] = args.encoded
 
     pool_rng = np.random.default_rng(args.pool_seed + run_id)
     agent_rng = np.random.default_rng(args.agent_seed)
@@ -43,7 +45,7 @@ while run_id < max_run_id:
                       model_seed=model_seed,
                       data_loader_seed=data_loader_seed,
                       device=util.device)
-    agent = AgentClass(agent_rng)
+    agent = AgentClass(agent_rng, config)
 
     if args.experiment_postfix is not None:
         base_path = os.path.join("runs", dataset.name, f"{agent.name}_{args.experiment_postfix}")

### 2023-04-19 compute_upper_bound.py
encoded argument fix

@@ -9,9 +9,10 @@ parser.add_argument("--data_folder", type=str, required=True)
 parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
 parser.add_argument("--dataset", type=str, default="splice")
-parser.add_argument("--encoded", type=bool, default=False)
-parser.add_argument("--restarts", type=int, default=3)
+parser.add_argument("--encoded", type=int, default=0)
+parser.add_argument("--restarts", type=int, default=15)
 args = parser.parse_args()
+args.encoded = bool(args.encoded)
 
 run_id = args.run_id
 max_run_id = run_id + args.restarts

### 2023-04-19 evaluate_oracle.py

@@ -11,11 +11,12 @@ parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--pool_seed", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
 parser.add_argument("--dataset", type=str, default="splice")
-parser.add_argument("--encoded", type=bool, default=True)
+parser.add_argument("--encoded", type=int, default=1)
 parser.add_argument("--sample_size", type=int, default=20)
 parser.add_argument("--restarts", type=int, default=3)
 parser.add_argument("--store_dataset", type=bool, default=False)
 args = parser.parse_args()
+args.encoded = bool(args.encoded)
 
 run_id = args.run_id
 max_run_id = run_id + args.restarts

### 2023-04-20 doc/al_benchmark_paper.tex
encoded argument fix
tuned usps

@@
 \end{table}
+
+Manual NIPS Benchmark paper eval (out of 6):
+\begin{itemize}
+	\item Has new method: 1
+	\item Multiple domains: 2
+	\item $> 5$ datasets: 3
+	\item $> 5$ baselines: 4
+\end{itemize}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{State Space}
-{\color{red} Version: 0.1}\\
-Since every AL agent needs a different state space our environment exposes a full state to the agent, so that the agent has full control of what information will be used. \\ [1mm]
-The state can include the following information:
-\begin{itemize}
-	\item The entire labeled dataset $\mathcal{L}^{(i)}$
-	\item The entire unlabeled dataset $\mathcal{U}^{(i)}$
-	\item A histogram of labeled points per class (count)
-	\item The available budget $B$
-	\item Number of added datapoints $|\mathcal{L}| - |\mathcal{S}|$
-	\item The initial validation accuracy $\operatorname{acc}^{(0)}$ and current validation accuracy $\operatorname{acc}^{(i)}$
-	\item The current classification model including all model weights $\theta^{(i)}$
-	\item The current optimizer including it's full state
-	\item The current sample of unlabeled points $u^{(i)}$
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Training the Classifier}\label{sec:training_the_classifier}
-Generally, the classification model can be trained in two ways.
-Either you reset the parameters after each AL iteration and train the classifier from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or you retain the previous state and only fine-tune the classifier on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
-In this work we use the fine-tuning method for raw datasets to save computation, while we use the from-scratch training for embedded dataset, since they have very small classifiers and this method generally produces better results.
-Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping after that.
-The early stopping has patience 0, so it will stop as soon as the validation loss does no longer decrease.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation}\label{sec:evaluation}
-{\color{red} Version: 0.1}\\
-We compare different AL algorithms based on their archived AUC score (Eq. \ref{eq:auc}).
-This score does incorporate performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
-A good AL algorithm should be able to perform well in both scenarios. \\
-{\color{red} TODO}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Reproducibility}\label{sec:reproducibility}
@@
 This way, the comparison is always fair
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evaluation}\label{sec:evaluation}
+{\color{red} Version: 0.1}\\
+We compare different AL algorithms based on their archived AUC score (Eq. \ref{eq:auc}).
+This score does incorporate performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
+A good AL algorithm should be able to perform well in both scenarios. \\
+{\color{red} TODO}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Implementation Details}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{State Space}
+{\color{red} Version: 0.1 Remove analogies to RL}\\
+Since every AL agent needs a different state space our environment exposes a full state to the agent, so that the agent has full control of what information will be used. \\ [1mm]
+The state can include the following information:
+\begin{itemize}
+	\item The entire labeled dataset $\mathcal{L}^{(i)}$
+	\item The entire unlabeled dataset $\mathcal{U}^{(i)}$
+	\item A histogram of labeled points per class (count)
+	\item The available budget $B$
+	\item Number of added datapoints $|\mathcal{L}| - |\mathcal{S}|$
+	\item The initial validation accuracy $\operatorname{acc}^{(0)}$ and current validation accuracy $\operatorname{acc}^{(i)}$
+	\item The current classification model including all model weights $\theta^{(i)}$
+	\item The current optimizer including it's full state
+	\item The current sample of unlabeled points $u^{(i)}$
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Choosing the Classifier}\label{sec:choosing_the_classifier}
+Traditionally, the classifier is chosen per dataset so that it is capable of solving the dataset close to the SOTA performance reported in the literature.
+Similar to our hypothesis in section \ref{sec:evaluation} we hypothesize that AL algorithms will perform similarly on small classifiers and more complex ones, so that the overall ranking of algorithms stays the same.
+{\color{red} TODO: Transform this into full hypothesis incl. formatting?}\\
+On the basis of this hypothesis we opt to use smaller classifiers that still solve the dataset to a reasonable degree.
+Smaller classifiers also require fewer labeled datapoints to acquire performance close to the upper bound.
+This behavior can be observed in Figure {\color{red} TODO}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Training the Classifier}\label{sec:training_the_classifier}
+{\color{red} Version: 0.1 Full rework needed}\\
+Generally, the classification model can be trained in two ways.
+Either you reset the parameters after each AL iteration and train the classifier from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or you retain the previous state and only fine-tune the classifier on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
+In this work we use the fine-tuning method for raw datasets to save computation, while we use the from-scratch training for embedded dataset, since they have very small classifiers and this method generally produces better results.
+Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping after that.
+The early stopping has patience 0, so it will stop as soon as the validation loss does no longer decrease.
 

### 2023-04-20 doc/al_benchmark_paper.tex
resnet dropout cifar10 linear raytune

@@
 \subsection{Comparibility}\label{sec:comparibility}
-{\color{red} Version: braindump}\\
+{\color{red} Version: brain dump}\\
 We split the evaluation model that generates accuracy scores and the model that generates features for the AL agent. \\
@@
 Smaller classifiers also require fewer labeled datapoints to acquire performance close to the upper bound.
-This behavior can be observed in Figure {\color{red} TODO}
+This behavior can be observed in Figure {\color{red} TODO} \\ [1mm]
+For every dataset the chosen architecture's hyperparameters are optimized by to archive maximum upper bound performance.
+For an overview of architectures and hyperparameters please refer to Table \ref{tab:architectures}.
+\begin{table}
+	\centering
+	\begin{tabular}{l || l | l l l l}
+		Dataset & Architecture & Optimizer & LR & Weight Decay & Dropout \\
+		\hline
+		Splice & [24, 12] & NAdam & 0.0012 & 5.9e-5 & 0 \\
+		DNA & [24, 12] & NAdam & 0.0398 & 3.6e-5 & 0 \\
+		USPS & [24, 12] & Adam & 0.0081 & 1.5e-6 & 0 \\
+		FashionMnist & linear & - & - & - & - \\
+		Cifar10 & linear & - & - & - & - \\
+	\end{tabular}
+	\caption{Classifier architectures and optimized hyperparameters per dataset. Numbers in brackets signify a MLP with corresponding hidden layers.}
+	\label{tab:architectures}
+\end{table}
 

### 2023-04-20 agents/bald.py

@@ -11,7 +11,7 @@ class BALD(BaseAgent):
 
     def __init__(self, agent_rng, config, dropout_trials=5):
         super().__init__(agent_rng, config)
-        assert "current_run_info" in config and "embedded" in config["current_run_info"]
+        assert "current_run_info" in config and "encoded" in config["current_run_info"]
         self.dropout_trials = dropout_trials
         self.dropout_model = None
 
@@ -20,7 +20,7 @@ class BALD(BaseAgent):
         """
         Add dropout to classification model
         """
-        class_key = "classifier_embedded" if config["current_run_info"]["embedded"] else "classifier"
+        class_key = "classifier_embedded" if config["current_run_info"]["encoded"] else "classifier"
         if "dropout" not in config[class_key]:
             config[class_key]["dropout"] = 0.2
 

### 2023-04-20 core/resnet.py

@@ -37,41 +37,11 @@ class BasicBlock(nn.Module):
         return out
 
 
-class Bottleneck(nn.Module):
-    expansion = 4
-
-    def __init__(self, in_planes, planes, stride=1):
-        super(Bottleneck, self).__init__()
-        self.conv1 = nn.Conv2d(in_planes, planes, kernel_size=1, bias=False)
-        self.bn1 = nn.BatchNorm2d(planes)
-        self.conv2 = nn.Conv2d(planes, planes, kernel_size=3,
-                               stride=stride, padding=1, bias=False)
-        self.bn2 = nn.BatchNorm2d(planes)
-        self.conv3 = nn.Conv2d(planes, self.expansion *
-                               planes, kernel_size=1, bias=False)
-        self.bn3 = nn.BatchNorm2d(self.expansion*planes)
-
-        self.shortcut = nn.Sequential()
-        if stride != 1 or in_planes != self.expansion*planes:
-            self.shortcut = nn.Sequential(
-                nn.Conv2d(in_planes, self.expansion*planes,
-                          kernel_size=1, stride=stride, bias=False),
-                nn.BatchNorm2d(self.expansion*planes)
-            )
-
-    def forward(self, x):
-        out = F.relu(self.bn1(self.conv1(x)))
-        out = F.relu(self.bn2(self.conv2(out)))
-        out = self.bn3(self.conv3(out))
-        out += self.shortcut(x)
-        out = F.relu(out)
-        return out
-
-
 class ResNet(nn.Module):
-    def __init__(self, block, num_blocks, num_classes=10, in_channels=3, add_head=True):
+    def __init__(self, block, num_blocks, num_classes=10, in_channels=3, dropout=None, add_head=True):
         super(ResNet, self).__init__()
         self.in_planes = 64
+        self.dropout = dropout
 
         self.conv1 = nn.Conv2d(in_channels, 64, kernel_size=3,
                                stride=1, padding=1, bias=False)
@@ -95,10 +65,18 @@ class ResNet(nn.Module):
     def _encode(self, x):
         out = F.relu(self.bn1(self.conv1(x)))
         out = self.layer1(out)
+        if self.dropout is not None:
+            out = F.dropout2d(out, p=self.dropout)
         out = self.layer2(out)
+        if self.dropout is not None:
+            out = F.dropout2d(out, p=self.dropout)
         out = self.layer3(out)
+        if self.dropout is not None:
[…]

### 2023-04-20 raytune/embedded_classification.py

@@ -13,14 +13,15 @@ def evaluate_encoded_classification_config(raytune_config, DatasetClass, config_
     with open(config_file, 'r') as f:
         config = yaml.load(f, yaml.Loader)
 
-    hidden_dims = [raytune_config["h1"]]
-    if raytune_config["h2"] > 0:
-        hidden_dims.append(raytune_config["h2"])
-    config["classifier_embedded"]["hidden"] = hidden_dims
+    # hidden_dims = [raytune_config["h1"]]
+    # if raytune_config["h2"] > 0:
+    #     hidden_dims.append(raytune_config["h2"])
+    # config["classifier_embedded"]["hidden"] = hidden_dims
     config["dataset_embedded"]["encoder_checkpoint"] = join(benchmark_folder, config["dataset_embedded"]["encoder_checkpoint"])
     config["optimizer_embedded"]["type"] = raytune_config["type"]
     config["optimizer_embedded"]["lr"] = raytune_config["lr"]
     config["optimizer_embedded"]["weight_decay"] = raytune_config["weight_decay"]
+    config["classifier"]["dropout"] = raytune_config["dropout"]
 
     loss_sum = 0.0
     acc_sum = 0.0
@@ -81,8 +82,9 @@ def tune_encoded_classification(num_samples, max_conc_trials, log_folder, config
         "type": tune.choice(["NAdam", "Adam", "SGD"]),
         "lr": tune.loguniform(1e-6, 1e-1),
         "weight_decay": tune.loguniform(1e-8, 1e-3),
-        "h1": tune.choice([12, 24, 48]),
-        "h2": tune.choice([0, 12, 24, 48]),
+        "dropout": tune.choice([0.0, 0.05, 0.1, 0.2, 0.3]),
+        # "h1": tune.choice([12, 24, 48]),
+        # "h2": tune.choice([0, 12, 24, 48]),
     }
 
     # fixes some parameters of the function

### 2023-04-21 doc/al_benchmark_paper.tex
updated fashion_mnist

@@
 \subsection{Problem Description / Delineation}
-\textbf{Lars' Version} \\
-Given
-\begin{itemize}
-	\item a number $B\in\mathcal{N}$ (called budget),
-	\item two spaces $\mathcal{X}$ and $\mathcal{Y}$,  {\tiny e.g., $\mathcal{X}:=\mathcal{R}^M, \mathcal{Y}:=\mathcal{R}^T$},
-	\item a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of
-	sequences of pairs $(x,y)$  from an unknown distribution $p$
-	(called datasets),
-	%  {\tiny with $|\mathcal{D}_n|\geq B$ for all $n\in 1{:}N$,}
-	{\tiny with $p(\mathcal{D})=0$ for $|\mathcal{D}|<B$,}
-	\item a function $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ (called loss), and
-	\item a function $A:  (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$
-	(called learning algorithm),
-\end{itemize}
-find a function
+{\color{red} Version: 1.0}\\
+Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$.
+\\
+Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
 \begin{align*}
@@
 % which is equivariant in the second argument, 
-called acquisition function,
-s.t. the expected loss of a model learned on
-all predictors 
-plus $B$ sequentially acquired
-targets is minimal:
+called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal:
 \begin{align*}
@@
 \end{align*}
-\textbf{Notes:}
-\begin{itemize}
-	\item We would need to switch from lowest expected loss to highest AUC
-	\item 
-\end{itemize}
-\vspace{5mm}
-\textbf{Thorben's Version}\\
-\textbf{Basic Classification}\\
-We assume a dataset $\mathcal{D} := (x_i, y_i) \hspace{0.5mm}; \hspace{1mm} i := 1 \ldots N$ consisting of instances $x_i \in \mathbb{R}^M$ and corresponding $y_i \in \mathbb{R}^C$.
-For evaluation purposes we assume a held-out test set $\mathcal{D}^{\text{test}}$ with the same characteristics.
-We consider classification problems with one-hot encoded classes, hence $C$ models the number of classes.
-To perform classification, a model $\hat y_\theta : \mathbb{R}^M \rightarrow \mathbb{R}^C$ is used.
-To fit the model, it is parameterized by $\theta$ and subjected to loss $\ell: \mathbb{R}^C \times \mathbb{R}^C \rightarrow \mathbb{R}$.
-For this work, categorical cross-entropy (CE) is used.
-For evaluating classification performance, we use accuracy on the test set $\operatorname{Acc}(\mathcal{D}^{\text{test}}, \hat{y}_\theta)$. \\ [1mm]
-%
+(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
+Since combinatorial problem of finding the optimal subset $\mathcal{D}_\text{train}$ is computational not feasible, we apply two relaxations to the problem. 
+Firstly, we allow sequential construction of $\mathcal{D}_\text{train}$ and secondly, we limit the choice of data points in every iteration to a random sample of size $\tau$. \\ [2mm]
+
 \textbf{Pool-based AL with single instances (non-batch setting)}\\

### 2023-04-21 collect_oracle_results.py

@@ -1,44 +1,0 @@
-import argparse
-import sys, shutil
-import os
-from os.path import join, exists
-import pandas as pd
-
-def check_for_nan_cols(df:pd.DataFrame):
-    cleaned_pd = df.copy(deep=True)
-    for col_name in df:
-        if df[col_name].isnull().values.any():
-            cleaned_pd.drop(columns=[col_name], inplace=True)
-    return cleaned_pd
-
-
-parser = argparse.ArgumentParser()
-parser.add_argument("--dataset", type=str, default="DNA")
-args = parser.parse_args()
-
-dataset_folder = join("runs", args.dataset)
-assert exists(dataset_folder)
-
-output_folder = join(dataset_folder, "Oracle")
-if exists(output_folder):
-    resp = input(f"{output_folder} already exists. Do you want to remove it? (y/n)")
-    if resp != "y":
-        sys.exit(0)
-    shutil.rmtree(output_folder)
-os.makedirs(output_folder)
-
-result_acc = pd.DataFrame()
-result_loss = pd.DataFrame()
-
-for run_folder in os.listdir(dataset_folder):
-    if run_folder.startswith("Oracle_"):
-        accuracies = pd.read_csv(join(dataset_folder, run_folder, "accuracies.csv"), header=0, index_col=0)
-        accuracies = check_for_nan_cols(accuracies)
-        result_acc = pd.concat([result_acc, accuracies], axis=1, ignore_index=True)
-        losses = pd.read_csv(join(dataset_folder, run_folder, "losses.csv"), header=0, index_col=0)
-        losses = check_for_nan_cols(losses)
-        result_loss = pd.concat([result_loss, losses], axis=1, ignore_index=True)
-
-result_acc.to_csv(join(output_folder, "accuracies.csv"))
-result_loss.to_csv(join(output_folder, "losses.csv"))
-

### 2023-04-28 doc/al_benchmark_paper.tex
TopV2 datasets

@@
 		Li et al. \cite{li2022empirical} & 5 & 1 & 1 & - & - \\
+		Zhou et al. \cite{zhou2021towards} & 3 & 2 & 1 & \checkmark & - \\
 		\textbf{Ours} & 5 & 2 & 2 & \checkmark & -
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Problem Description / Delineation}
+\subsection{Problem Description}
 {\color{red} Version: 1.0}\\
@@
 Firstly, we allow sequential construction of $\mathcal{D}_\text{train}$ and secondly, we limit the choice of data points in every iteration to a random sample of size $\tau$. \\ [2mm]
+%
+To construct the active learning setting, we suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$.
+\\ [2mm]
 
-\textbf{Pool-based AL with single instances (non-batch setting)}\\
-To construct the active learning setting, we suppress the labels $y_i$ of $\mathcal{D}$ to form the unlabeled pool $\mathcal{U} := u_i \hspace{0.5mm}; \hspace{1mm} i := 1 \ldots N$ and form and initial labeled pool $\mathcal{L}$ by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label.
-The result of this so-called "seeding" process is $\mathcal{L} := (u_i, y_i) \hspace{0.5mm}; \hspace{1mm} i := 1 \ldots k*C$. \\
-Active learning is defined as sequentially removing single instances $u^{(i)} \in \mathcal{U}^{(i)}\hspace{1mm}; \hspace{2mm} \mathcal{U}^{(i+1)} := \mathcal{U}^{(i)} \setminus´ \{u^{(i)}\}$, recovering their label $y^{(i)}$ and adding them to the labeled pool $\mathcal{L}^{(i+1)} := \mathcal{L}^{(i)} \cup (u^{(i)}, y^{(i)})$ until a fixed budget $B$ is exhausted $i := 1 \ldots B$.
-After each added instance the classification model is retrained according to section \ref{sec:training_the_classifier} and its performance is measured on the held-out test set $\mathcal{D}^{\text{test}}$.
-The quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
+Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
 We employ the normalized area under the accuracy curve (AUC):
@@
 \subsection{Splice Embedded}
+
+\begin{table}[]
+	\begin{tabular}{l|lll}
+		& Splice        & DNA           & USPS          \\
+		\hline
+		Oracle          & 0.835 +- 0.01 & 0.879 +- 0.01 & 0.870 +- 0.01 \\
+		SAL\_23\_01\_16 & 0.808 +- 0.01 & 0.857 +- 0.02 & 0.865 +- 0.01 \\
+		Coreset\_Greedy & 0.808 +- 0.01 & 0.853 +- 0.02 & 0.849 +- 0.01 \\
+		MarginScore     & 0.806 +- 0.01 & 0.855 +- 0.02 & 0.864 +- 0.01 \\
+		ShannonEntropy  & 0.808 +- 0.01 & 0.855 +- 0.02 & 0.864 +- 0.01 \\
+		RandomAgent     & 0.797 +- 0.01 & 0.823 +- 0.02 & 0.831 +- 0.01
+	\end{tabular}
+\end{table}
+%
+\begin{table}[]
+	\begin{tabular}{l|ll}
+		& Cifar10 & FashionMnist \\
+		\hline
+		Oracle          & 0.718 +- 0.01  & 0.721 +- 0.01       \\
+		SAL\_23\_01\_16 & 0.660 +- 0.01  & 0.667 +- 0.01       \\
+		Coreset\_Greedy & 0.668 +- 0.01  & 0.674 +- 0.01       \\
+		MarginScore     & 0.674 +- 0.01  & 0.677 +- 0.01       \\
+		ShannonEntropy  & 0.671 +- 0.01  & 0.680 +- 0.01       \\
+		RandomAgent     & 0.632 +- 0.02  & 0.633 +- 0.01      
+	\end{tabular}
+\end{table}
+%
 \begin{figure}[H]

### 2023-04-28 classifiers/classifier.py

@@ -1,58 +1,13 @@
 from typing import Tuple
-import math
 import torch
 import torch.nn as nn
-from torch.nn.init import _calculate_fan_in_and_fan_out, calculate_gain, _calculate_correct_fan
 import torch.nn.functional as F
 from torch import Tensor
 from core.data import BaseDataset
 from torch.utils.data import TensorDataset, DataLoader
 from tqdm import tqdm
-
-def kaiming_uniform_seeded(
-    rng, tensor: torch.Tensor, a: float = 0, mode: str = 'fan_in', nonlinearity: str = 'leaky_relu'
-):
-
-    if torch.overrides.has_torch_function_variadic(tensor):
-        return torch.overrides.handle_torch_function(
-            kaiming_uniform_seeded,
-            (tensor,),
-            tensor=tensor,
-            a=a,
-            mode=mode,
-            nonlinearity=nonlinearity)
-
-    if 0 in tensor.shape:
-        print("Initializing zero-element tensors is a no-op")
-        return tensor
-    fan = _calculate_correct_fan(tensor, mode)
-    gain = calculate_gain(nonlinearity, a)
-    std = gain / math.sqrt(fan)
-    bound = math.sqrt(3.0) * std  # Calculate uniform bounds from standard deviation
-    with torch.no_grad():
-        return tensor.uniform_(-bound, bound, generator=rng)
-
-
-
-class SeededLinear(nn.Linear):
-    """
-    Replaces the default kaiming initialization with a seeded version of itself
-    """
-
-    def __init__(self, model_rng, *args, **kwargs):
-        self.model_rng = model_rng
-        super().__init__(*args, **kwargs)
-
-    def reset_parameters(self) -> None:
-        # Setting a=sqrt(5) in kaiming_uniform is the same as initializing with
-        # uniform(-1/sqrt(in_features), 1/sqrt(in_features)). For details, see
-        # https://github.com/pytorch/pytorch/issues/57109
-        kaiming_uniform_seeded(self.model_rng, self.weight, a=math.sqrt(5))
-        if self.bias is not None:
-            fan_in, _ = _calculate_fan_in_and_fan_out(self.weight)
-            bound = 1 / math.sqrt(fan_in) if fan_in > 0 else 0
-            with torch.no_grad():
-                self.bias.uniform_(-bound, bound, generator=self.model_rng)
+from classifiers.seeded_layers import SeededLinear
[…]

### 2023-04-28 classifiers/seeded_layers.py

@@ -1,0 +1,72 @@
+import math
+import torch
+import torch.nn as nn
+from torch.nn.init import _calculate_fan_in_and_fan_out, calculate_gain, _calculate_correct_fan
+
+def kaiming_uniform_seeded(
+    rng, tensor: torch.Tensor, a: float = 0, mode: str = 'fan_in', nonlinearity: str = 'leaky_relu'
+):
+
+    if torch.overrides.has_torch_function_variadic(tensor):
+        return torch.overrides.handle_torch_function(
+            kaiming_uniform_seeded,
+            (tensor,),
+            tensor=tensor,
+            a=a,
+            mode=mode,
+            nonlinearity=nonlinearity)
+
+    if 0 in tensor.shape:
+        print("Initializing zero-element tensors is a no-op")
+        return tensor
+    fan = _calculate_correct_fan(tensor, mode)
+    gain = calculate_gain(nonlinearity, a)
+    std = gain / math.sqrt(fan)
+    bound = math.sqrt(3.0) * std  # Calculate uniform bounds from standard deviation
+    with torch.no_grad():
+        return tensor.uniform_(-bound, bound, generator=rng)
+
+
+
+class SeededLinear(nn.Linear):
+    """
+    Replaces the default kaiming initialization with a seeded version of itself
+    """
+
+    def __init__(self, model_rng, *args, **kwargs):
+        self.model_rng = model_rng
+        super().__init__(*args, **kwargs)
+
+    def reset_parameters(self) -> None:
+        # Setting a=sqrt(5) in kaiming_uniform is the same as initializing with
+        # uniform(-1/sqrt(in_features), 1/sqrt(in_features)). For details, see
+        # https://github.com/pytorch/pytorch/issues/57109
+        kaiming_uniform_seeded(self.model_rng, self.weight, a=math.sqrt(5))
+        if self.bias is not None:
+            fan_in, _ = _calculate_fan_in_and_fan_out(self.weight)
+            bound = 1 / math.sqrt(fan_in) if fan_in > 0 else 0
+            with torch.no_grad():
+                self.bias.uniform_(-bound, bound, generator=self.model_rng)
+
+
+class SeededConv2d(nn.Conv2d):
+    """
+    Replaces the default kaiming initialization with a seeded version of itself
+    """
+
+    def __init__(self, model_rng, *args, **kwargs):
+        self.model_rng = model_rng
+        super().__init__(*args, **kwargs)
+
[…]

### 2023-04-28 compute_upper_bound.py

@@ -1,7 +1,7 @@
 import yaml
 import experiment_util as util
 import argparse
-from core.classifier import fit_and_evaluate
+from classifiers.classifier import fit_and_evaluate
 from core.helper_functions import *
 
 parser = argparse.ArgumentParser()

### 2023-04-28 datasets/__init__.py

@@ -7,5 +7,8 @@ from datasets.usps import USPS
 from datasets.cifar10 import Cifar10
 from datasets.fashion_mnist import FashionMnist
 
+# Text Datasets
+from datasets.topv2 import TopV2
+
 # Toy Datasets
 from datasets.toydata1 import ToyData1

### 2023-04-28 datasets/cifar10.py

@@ -1,12 +1,7 @@
-from typing import Tuple
-import yaml
 import torch
-import torch.nn as nn
-from torch.utils.data import Dataset, TensorDataset, DataLoader
+from torch.utils.data import Dataset
 import torchvision
 from torchvision import transforms
-from core.resnet import ResNet18
-from sim_clr.encoder import ContrastiveModel
 from core.data import BaseDataset, postprocess_torch_dataset, convert_to_channel_first, subsample_data
 
 class Cifar10(BaseDataset):
@@ -27,6 +22,7 @@ class Cifar10(BaseDataset):
         high = 255.0
         self.x_train = self.x_train / (high / 2.0) - 1.0
         self.x_test = self.x_test / (high / 2.0) - 1.0
+        self._convert_data_to_tensors()
         print("Download successful")
 
     def load_pretext_data(self)->tuple[Dataset, Dataset]:

### 2023-04-28 datasets/fashion_mnist.py

@@ -1,11 +1,7 @@
-from typing import Tuple
 import torch
-import torch.nn as nn
 from torch.utils.data import Dataset
 import torchvision
 from torchvision import transforms
-from core.resnet import ResNet18
-from sim_clr.encoder import ContrastiveModel
 from core.data import BaseDataset, postprocess_torch_dataset, convert_to_channel_first, subsample_data
 
 class FashionMnist(BaseDataset):
@@ -30,6 +26,7 @@ class FashionMnist(BaseDataset):
         high = 255.0
         self.x_train = self.x_train / (high / 2.0) - 1.0
         self.x_test = self.x_test / (high / 2.0) - 1.0
+        self._convert_data_to_tensors()
         print("Download successful")
 
 

### 2023-04-28 datasets/toydata1.py

@@ -1,15 +1,7 @@
-from typing import Tuple, Union, Callable
-import os
-from os.path import exists
-import torch
-import torch.nn as nn
+from typing import Tuple, Union
 import numpy as np
-import pandas as pd
-from sklearn.datasets import load_svmlight_file
-from sklearn.model_selection import train_test_split
-from core.data import BaseDataset, normalize, postprocess_svm_data
-from core.classifier import DenseModel
-import requests
+from core.data import BaseDataset
+
 
 class ToyData1(BaseDataset):
 

### 2023-04-28 evaluate.py

@@ -13,8 +13,8 @@ parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--agent_seed", type=int, default=1)
 parser.add_argument("--pool_seed", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
-parser.add_argument("--agent", type=str, default="entropy")
-parser.add_argument("--dataset", type=str, default="dna")
+parser.add_argument("--agent", type=str, default="margin")
+parser.add_argument("--dataset", type=str, default="topv2")
 parser.add_argument("--encoded", type=int, default=0)
 parser.add_argument("--sample_size", type=int, default=20)
 parser.add_argument("--restarts", type=int, default=10)
@@ -45,6 +45,7 @@ while run_id < max_run_id:
 
     # Inject additional configuration into the dataset config (See BALD agent)
     AgentClass.inject_config(config)
+    DatasetClass.inject_config(config)
 
     dataset = DatasetClass(args.data_folder, config, pool_rng, args.encoded)
     dataset = dataset.to(util.device)

### 2023-04-28 generate_result_table.py

@@ -1,0 +1,37 @@
+import os
+from os.path import exists, join
+import pandas as pd
+import numpy as np
+
+agents = ["Oracle", "SAL_23_01_16", "Coreset_Greedy", "MarginScore", "ShannonEntropy", "RandomAgent"]
+vector_data = ["Splice", "DNA", "USPS"]
+img_data = ["Cifar10Encoded", "FashionMnistEncoded"]
+
+def generate_table(agents, datasets, out_file):
+    result_table = pd.DataFrame(index=agents, columns=datasets)
+    for dataset in datasets:
+        dataset_folder = os.path.join("runs", dataset)
+        if not os.path.isdir(dataset_folder):
+            continue
+        agent_aucs = dict()
+        for a in agents:
+            agent_aucs[a] = "-"
+
+        for agent in agents:
+            agent_folder = join(dataset_folder, agent)
+            if exists(agent_folder):
+                acc_file = join(agent_folder, "accuracies.csv")
+                if exists(acc_file):
+                    accuracies = pd.read_csv(acc_file, header=0, index_col=0)
+                    values = accuracies.values
+                    auc = np.sum(values, axis=0) / values.shape[0]
+                    result_table[dataset][agent] = "%1.3f +- %1.2f"%(np.mean(auc).item(), np.std(auc).item())
+
+    if len(result_table) > 0:
+        result_table_file = os.path.join("runs", out_file)
+        if os.path.exists(result_table_file):
+            os.remove(result_table_file)
+        result_table.to_csv(result_table_file)
+
+generate_table(agents, vector_data, "result_vector.csv")
+generate_table(agents, img_data, "result_img.csv")

### 2023-05-02 doc/al_benchmark_paper.tex
TopV2 raytune

@@
 \section{Introduction}
+Among others, the authors of \cite{zhou2021towards} have pointed out severe inconsistencies in results of AL papers in recent years. 
+In their supplementary materials they conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while all following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm.
+The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art results by significantly outperforming everyone else, while the vast majority of results proves to be non-reproducible.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Oracle}\label{sec:oracle}
+{\color{red} Version: 1.0}\\
+Posing active learning as a sequence ordering problem, the oracle sequence for a given combination of dataset, model and training procedure would be the sequence that induces the highest AUC score for a given budget.
+However, since this combinatorial problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence.
+\cite{zhou2021towards} has used simulated annealing to search for the best sequence and used the best found solution after a fixed time budget. 
+Even though the reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
+\\
+In this paper we propose a greedy oracle algorithm the constructs an approximation of the optimal sequence in an iterative fashion.
+Our oracle simply tests every data point in the provided sample of unlabeled points by fitting the classifier and directly measuring the resulting validation performance.
+The point with the best validation performance is selected and added to the labeled pool for that iteration.
+We noticed that this oracle is overfitting on the validation set, resulting in stagnating or even decreasing performance curves in later AL iterations.
+To circumvent this problem, we introduced margin sampling as a fallback option for the oracle.
+Whenever the oracle does not find an unlabeled point that results in an increase in performance (indicating an overfitting position), it defaults to margin sampling in that iteration.
+The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 		\hline
-		Oracle          & 0.835 +- 0.01 & 0.879 +- 0.01 & 0.870 +- 0.01 \\
-		SAL\_23\_01\_16 & 0.808 +- 0.01 & 0.857 +- 0.02 & 0.865 +- 0.01 \\
-		Coreset\_Greedy & 0.808 +- 0.01 & 0.853 +- 0.02 & 0.849 +- 0.01 \\
-		MarginScore     & 0.806 +- 0.01 & 0.855 +- 0.02 & 0.864 +- 0.01 \\
-		ShannonEntropy  & 0.808 +- 0.01 & 0.855 +- 0.02 & 0.864 +- 0.01 \\
-		RandomAgent     & 0.797 +- 0.01 & 0.823 +- 0.02 & 0.831 +- 0.01
+		Oracle          & 0.835 $\pm$ 0.01 & 0.879 $\pm$ 0.01 & 0.870 $\pm$ 0.01 \\
+		SAL\_23\_01\_16 & 0.808 $\pm$ 0.01 & 0.857 $\pm$ 0.02 & 0.865 $\pm$ 0.01 \\
+		Coreset\_Greedy & 0.808 $\pm$ 0.01 & 0.853 $\pm$ 0.02 & 0.849 $\pm$ 0.01 \\
+		MarginScore     & 0.806 $\pm$ 0.01 & 0.855 $\pm$ 0.02 & 0.864 $\pm$ 0.01 \\
+		ShannonEntropy  & 0.808 $\pm$ 0.01 & 0.855 $\pm$ 0.02 & 0.864 $\pm$ 0.01 \\
+		RandomAgent     & 0.797 $\pm$ 0.01 & 0.823 $\pm$ 0.02 & 0.831 $\pm$ 0.01
 	\end{tabular}
@@
 		\hline
-		Oracle          & 0.718 +- 0.01  & 0.721 +- 0.01       \\
-		SAL\_23\_01\_16 & 0.660 +- 0.01  & 0.667 +- 0.01       \\
-		Coreset\_Greedy & 0.668 +- 0.01  & 0.674 +- 0.01       \\
-		MarginScore     & 0.674 +- 0.01  & 0.677 +- 0.01       \\
-		ShannonEntropy  & 0.671 +- 0.01  & 0.680 +- 0.01       \\
-		RandomAgent     & 0.632 +- 0.02  & 0.633 +- 0.01      
+		Oracle          & 0.718 $\pm$ 0.01  & 0.721 $\pm$ 0.01       \\
+		SAL\_23\_01\_16 & 0.660 $\pm$ 0.01  & 0.667 $\pm$ 0.01       \\
+		Coreset\_Greedy & 0.668 $\pm$ 0.01  & 0.674 $\pm$ 0.01       \\
+		MarginScore     & 0.674 $\pm$ 0.01  & 0.677 $\pm$ 0.01       \\
+		ShannonEntropy  & 0.671 $\pm$ 0.01  & 0.680 $\pm$ 0.01       \\
+		RandomAgent     & 0.632 $\pm$ 0.02  & 0.633 $\pm$ 0.01      
 	\end{tabular}

### 2023-05-02 datasets/topv2.py

@@ -119,6 +119,6 @@ if __name__ == '__main__':
     with open(f"../configs/topv2.yaml", 'r') as f:
         config = yaml.load(f, yaml.Loader)
     dataset = TopV2("../../datasets", config, np.random.default_rng(0), encoded=False)
-    model, retain_graph = dataset.get_classifier(model_rng=torch.Generator())
+    model = dataset.get_classifier(model_rng=torch.Generator())
     test_point = dataset.x_train[0].unsqueeze(0)
     model(test_point)

### 2023-05-02 results/generate_result_table.py

@@ -3,14 +3,10 @@ from os.path import exists, join
 import pandas as pd
 import numpy as np
 
-agents = ["Oracle", "SAL_23_01_16", "Coreset_Greedy", "MarginScore", "ShannonEntropy", "RandomAgent"]
-vector_data = ["Splice", "DNA", "USPS"]
-img_data = ["Cifar10Encoded", "FashionMnistEncoded"]
-
 def generate_table(agents, datasets, out_file):
     result_table = pd.DataFrame(index=agents, columns=datasets)
     for dataset in datasets:
-        dataset_folder = os.path.join("runs", dataset)
+        dataset_folder = os.path.join("../runs", dataset)
         if not os.path.isdir(dataset_folder):
             continue
         agent_aucs = dict()
@@ -28,10 +24,15 @@ def generate_table(agents, datasets, out_file):
                     result_table[dataset][agent] = "%1.3f +- %1.2f"%(np.mean(auc).item(), np.std(auc).item())
 
     if len(result_table) > 0:
-        result_table_file = os.path.join("runs", out_file)
+        result_table_file = os.path.join("../runs", out_file)
         if os.path.exists(result_table_file):
             os.remove(result_table_file)
         result_table.to_csv(result_table_file)
+
+
+agents = ["Oracle", "SAL_23_01_16", "Coreset_Greedy", "MarginScore", "ShannonEntropy", "RandomAgent"]
+vector_data = ["Splice", "DNA", "USPS"]
+img_data = ["Cifar10Encoded", "FashionMnistEncoded"]
 
 generate_table(agents, vector_data, "result_vector.csv")
 generate_table(agents, img_data, "result_img.csv")

### 2023-05-05 classifiers/classifier.py
LSTM seeded layers News dataset

@@ -104,11 +104,13 @@ class ConvolutionalModel(nn.Module):
         return x
 
 
-def construct_model(model_rng, x_shape, n_classes, model_config, add_head=True) -> Tuple[nn.Module, int]:
+def construct_model(model_rng, dataset:BaseDataset, model_config:dict, add_head=True) -> Tuple[nn.Module, int]:
         '''
         Constructs the model by name and additional parameters
         Returns model and its output dim
         '''
+        x_shape = dataset.x_shape
+        n_classes = dataset.n_classes
         model_type = model_config["type"].lower()
         dropout = model_config["dropout"] if "dropout" in model_config else None
         if model_type == "linear":
@@ -130,8 +132,10 @@ def construct_model(model_rng, x_shape, n_classes, model_config, add_head=True) 
                               add_head=add_head), \
                    n_classes if add_head else model_config["hidden"][-1]
         elif model_type == "bilstm":
+            assert hasattr(dataset, "embedding_data_file"), "Dataset is missing the embedding file. This is specific to text datasets."
+            embedding_data = torch.load(dataset.embedding_data_file)
             return BiLSTMModel(model_rng,
-                               emb_dim=x_shape[-1],
+                               embedding_data=embedding_data,
                                num_classes=n_classes,
                                dropout=dropout), \
                    n_classes if add_head else model_config["hidden"][-1]

### 2023-05-05 classifiers/seeded_layers.py

@@ -70,3 +70,72 @@ class SeededConv2d(nn.Conv2d):
                 self.bias.uniform_(-bound, bound, generator=self.model_rng)
                 # init.uniform_(self.bias, -bound, bound)
 
+
+class SeededEmbedding(nn.Embedding):
+    def __init__(self, model_rng, *args, **kwargs):
+        self.model_rng = model_rng
+        super().__init__(*args, **kwargs)
+
+    def reset_parameters(self) -> None:
+        with torch.no_grad():
+            self.weight.normal_(mean=0., std=1., generator=self.model_rng)
+            self._fill_padding_idx_with_zero()
+
+    @classmethod
+    def from_pretrained(cls, model_rng, embeddings, freeze=True, padding_idx=None,
+                        max_norm=None, norm_type=2., scale_grad_by_freq=False,
+                        sparse=False):
+        r"""Creates Embedding instance from given 2-dimensional FloatTensor.
+
+        Args:
+            embeddings (Tensor): FloatTensor containing weights for the Embedding.
+                First dimension is being passed to Embedding as ``num_embeddings``, second as ``embedding_dim``.
+            freeze (bool, optional): If ``True``, the tensor does not get updated in the learning process.
+                Equivalent to ``embedding.weight.requires_grad = False``. Default: ``True``
+            padding_idx (int, optional): If specified, the entries at :attr:`padding_idx` do not contribute to the gradient;
+                                         therefore, the embedding vector at :attr:`padding_idx` is not updated during training,
+                                         i.e. it remains as a fixed "pad".
+            max_norm (float, optional): See module initialization documentation.
+            norm_type (float, optional): See module initialization documentation. Default ``2``.
+            scale_grad_by_freq (bool, optional): See module initialization documentation. Default ``False``.
+            sparse (bool, optional): See module initialization documentation.
+
+        Examples::
+
+            >>> # FloatTensor containing pretrained weights
+            >>> weight = torch.FloatTensor([[1, 2.3, 3], [4, 5.1, 6.3]])
+            >>> embedding = nn.Embedding.from_pretrained(weight)
+            >>> # Get embeddings for index 1
+            >>> input = torch.LongTensor([1])
+            >>> # xdoctest: +IGNORE_WANT("non-deterministic")
+            >>> embedding(input)
+            tensor([[ 4.0000,  5.1000,  6.3000]])
+        """
[…]

### 2023-05-05 compute_upper_bound.py

@@ -8,7 +8,7 @@ parser = argparse.ArgumentParser()
 parser.add_argument("--data_folder", type=str, required=True)
 parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
-parser.add_argument("--dataset", type=str, default="topv2")
+parser.add_argument("--dataset", type=str, default="news")
 parser.add_argument("--encoded", type=int, default=0)
 parser.add_argument("--restarts", type=int, default=5)
 args = parser.parse_args()

### 2023-05-05 core/data.py

@@ -158,9 +158,9 @@ class BaseDataset(ABC):
     def get_classifier(self, model_rng) -> Module:
         from classifiers.classifier import construct_model
         if self.encoded:
-            model, _= construct_model(model_rng, self.x_shape, self.n_classes, self.config["classifier_embedded"])
+            model, _= construct_model(model_rng, self, self.config["classifier_embedded"])
         else:
-            model, _ = construct_model(model_rng, self.x_shape, self.n_classes, self.config["classifier"])
+            model, _ = construct_model(model_rng, self, self.config["classifier"])
         return model
 
     def get_pretext_encoder(self, config: dict, seed=1) -> nn.Module:
@@ -357,6 +357,11 @@ def convert_to_channel_first(train: Union[Tensor, np.ndarray], test: Union[Tenso
         test = test.permute(0, 3, 1, 2)
     return train, test
 
+
+def to_one_hot(labels:np.ndarray)->np.ndarray:
+    one_hot = np.zeros((len(labels), labels.max() + 1))
+    one_hot[np.arange(len(labels)), labels] = 1
+    return one_hot
 
 def postprocess_torch_dataset(train, test) -> Tuple:
     x_train, y_train = train.data, np.array(train.targets)

### 2023-05-05 datasets/__init__.py

@@ -9,6 +9,7 @@ from datasets.fashion_mnist import FashionMnist
 
 # Text Datasets
 from datasets.topv2 import TopV2
+from datasets.news import News
 
 # Toy Datasets
 from datasets.toydata1 import ToyData1

### 2023-05-05 datasets/news.py

@@ -1,0 +1,145 @@
+import os
+from os.path import exists
+import numpy as np
+import pandas as pd
+import tqdm
+import torch
+import torch.nn as nn
+from torch.utils.data import Dataset
+from torch.nn.utils.rnn import pack_padded_sequence
+from torchvision import transforms
+from core.data import BaseDataset, to_one_hot
+import requests, zipfile
+import json
+import nltk
+from string import punctuation
+
+class News(BaseDataset):
+
+    def __init__(self, cache_folder:str, config:dict, pool_rng, encoded:bool,
+                 data_file="news_al.pt"):
+        assert not encoded, "News Categories only supports un-encoded version (which uses GLOVE embeddings)"
+        self.raw_zip_file = os.path.join(cache_folder, "news_categories.zip")
+        self.raw_unzipped_file = os.path.join(cache_folder, "news_categories/News_Category_Dataset_v3.json")
+        self.raw_embedding_file = os.path.join(cache_folder, "news_categories/glove_embeddings.json")
+        self.raw_train_file = os.path.join(cache_folder, "news_train.txt")
+        self.raw_test_file = os.path.join(cache_folder, "news_test.txt")
+        self.embedding_data_file = os.path.join(cache_folder, "news_al_embeddings.pt")
+        fitting_mode = "from_scratch" if encoded else "finetuning"
+        super().__init__(cache_folder, config, pool_rng, encoded,
+                         data_file, fitting_mode)
+
+
+    def _download_data(self, target_to_one_hot=True, max_sentence_len=80):
+        zip_url = "https://www.ismll.uni-hildesheim.de/personen/twerner/news_categories.zip"
+
+        print("This download can take up to 10 minutes (108MB)")
+        if not exists(self.raw_zip_file):
+            r = requests.get(zip_url)
+            with open(self.raw_zip_file, 'wb') as f:
+                f.write(r.content)
+
+        if not exists(self.raw_unzipped_file):
+            z = zipfile.ZipFile(self.raw_zip_file)
+            z.extract('News_Category_Dataset_v3.json', os.path.join(self.cache_folder, "news_categories"))
+            z.extract('glove_embeddings.json', os.path.join(self.cache_folder, "news_categories"))
+
+        # We are using the top 15 categories as indicated by Kaggle: https://www.kaggle.com/datasets/rmisra/news-category-dataset
[…]

### 2023-05-05 evaluate.py

@@ -14,7 +14,7 @@ parser.add_argument("--agent_seed", type=int, default=1)
 parser.add_argument("--pool_seed", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
 parser.add_argument("--agent", type=str, default="margin")
-parser.add_argument("--dataset", type=str, default="topv2")
+parser.add_argument("--dataset", type=str, default="news")
 parser.add_argument("--encoded", type=int, default=0)
 parser.add_argument("--sample_size", type=int, default=20)
 parser.add_argument("--restarts", type=int, default=10)

### 2023-05-05 doc/al_benchmark_paper.tex
LSTM seeded layers News dataset
News experiments

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}
-{\color{red} Version: 1.0}\\
-In this work we use 4 different datasets from two domains.
-For vector data, we use DNA and Splice (Source: \cite{libsvmtools}) and for image data we use FashionMNIST \cite{xiao2017fashion} and Cifar10 \cite{krizhevsky2009learning}. \\
-We would like to point out that these datasets can be considered "toy-datasets" and therefore not relevant for practical purposes.
-This might be true if we aimed to develop novel classification models on these dataset, however we are solely focused on comparing different AL algorithms in this paper.
-Our core assumption is that a well-performing algorithm in our benchmark will transfer well into more practical use-cases. \\ [1mm]
-Adapting the experimental setting from \cite{hacohen2022active} we offer all our datasets in the raw setting as well as pre-encoded by a fixed embedding model that was trained in an unsupervised manner.
-The pre-encoded datasets offer multiple advantages in that they require smaller classification models (i.e.
-small MLPs) and in general much less sampled datapoints to archive close to upper bound performance (measured with the full dataset).
-They unify experimental setups for different datasets and therefore make them much more comparable.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Datasets}\label{sec:datasets}
+{\color{red} Version: 1.0}\\
+For all our datasets we use the pre-defined train / test splits, if given. 
+In the remaining cases, we define test sets upfront and store them into separate files to keep them fixed across all experiments.
+The validation set is split during experiment-time and depends on the dataset-seed.
+
+\paragraph{Tabular}
+We use \textbf{Splice}, \textbf{DNA} and \textbf{USPS} from LibSVMTools \cite{libsvmtools}.
+All three datasets are normalized between [0, 1].
+
+\paragraph{Image}
+We use \textbf{FashionMNIST} \cite{xiao2017fashion} and \textbf{Cifar10} \cite{krizhevsky2009learning}.
+Both datasets are normalized between [-1, 1].
+
+\paragraph{Text}
+We use \textbf{News Category} \cite{misra2022news} and \textbf{TopV2} \cite{chen-etal-2020-low-resource}.
+For News Category we use  the 15 most common categories as indicated by its Kaggle site.
+We additionally drop sentences above 80 words to reduce the necessary padding (retaining 99,86\% of the data).
+For TopV2, we are only using the "alarm" domain.
+Both datasets are encoded with pre-trained GloVe embeddings \cite{pennington2014glove}.
+Since neither set provided a fixed test set, we split random 5000 datapoints into a test set. \\ [3mm]
+%
+We would like to point out that these datasets can be considered "toy-datasets" and therefore not relevant for practical purposes.
+This might be true if we aimed to develop novel classification models on these datasets, however we are solely focused on comparing different AL algorithms in this paper.
+Our core assumption is that a well-performing algorithm in our benchmark will transfer well into more practical use-cases. \\ [1mm]
+Adapting the experimental setting from \cite{hacohen2022active} we offer all our datasets in the raw setting as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning.
+The text datasets are an exception, as they are only offered in their encoded form.
+The pre-encoded datasets offer multiple advantages in that they require smaller classification models (i.e.
+small MLPs) and in general much less sampled datapoints to archive close to upper bound performance (measured with the full dataset).
+They unify experimental setups for different datasets and therefore make them much more comparable.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Splice Embedded}

### 2023-05-11 doc/al_benchmark_paper.tex
state space rework

@@
 
-Manual NIPS Benchmark paper eval (out of 6):
-\begin{itemize}
-	\item Has new method: 1
-	\item Multiple domains: 2
-	\item $> 5$ datasets: 3
-	\item $> 5$ baselines: 4
-\end{itemize}
+%Manual NIPS Benchmark paper eval (out of 6):
+%\begin{itemize}
+%	\item Has new method: 1
+%	\item Multiple domains: 2
+%	\item $> 5$ datasets: 3
+%	\item $> 5$ baselines: 4
+%\end{itemize}
 
@@
 (\textit{We would need to switch from lowest expected loss to highest AUC}) \\
-Since combinatorial problem of finding the optimal subset $\mathcal{D}_\text{train}$ is computational not feasible, we apply two relaxations to the problem. 
-Firstly, we allow sequential construction of $\mathcal{D}_\text{train}$ and secondly, we limit the choice of data points in every iteration to a random sample of size $\tau$. \\ [2mm]
+Since combinatorial problem of finding the optimal subset $\mathcal{D}_\text{train}$ is computational not feasible, we allow sequential construction of $\mathcal{D}_\text{train}$. \\ [2mm]
 %
@@
 
+% TODO: remove tau from pseudocode
 \begin{minipage}{0.59\linewidth}
 	\begin{algorithm}[H]
-		\caption{Active Learning {\color{red} OLD}}\label{alg:active_learning}
+		\caption{Active Learning}\label{alg:active_learning}
 		\begin{algorithmic}[1]
@@
 			\For{$i := 1 \ldots B$}
-				\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$  \Comment{$\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
-				\State $u^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$
-				\State $s^{(i)} \gets \omega(u^{(i)})$
-				\State $a^{(i)} \gets \Omega(s^{(i)})$ \Comment{$a^{(i)}$ is an index inside of $u^{(i)}$}
-				\State $y^{(i)} \gets \operatorname{label}(u^{(i)}_{a})$ \Comment{$u^{(i)}_{a}$ is shorthand for $u^{(i)}_{a^{(i)}}$}
+				\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$  %\Comment{$\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
+				\State $u^{(i)} \sim \text{unif}(1:|\mathcal{U}^{(i)}|)$
+				\State $a^{(i)} \gets \Omega(u^{(i)}) \hspace{2mm}|\hspace{2mm} a \in 1:|u^{(i)}|$ 
+				%\Comment{$a^{(i)}$ is an index inside of $u^{(i)}$}
+				\State $y^{(i)} \gets \operatorname{label}(u^{(i)}_{a})$ 
+				%\Comment{$u^{(i)}_{a}$ is shorthand for $u^{(i)}_{a^{(i)}}$}
 				\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_a, y^{(i)})\}$
@@
 \begin{algorithm}
-	\caption{Oracle {\color{red} OLD}}\label{alg:oracle}
+	\caption{Oracle}\label{alg:oracle}
 	\begin{algorithmic}[1]
@@
 			\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$  \Comment{$\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
-			\State $u^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$
+			\State $u^{(i)} \sim \text{unif}(1:|\mathcal{U}^{(i)}|)$
 			\State $r* \gets -\infty$
@@
 This behavior is not possible with the default choice of setting a global seed at the start of the experiment, since a single additional random draw from the random number generator completely changes the decision tree for the batches.
-This additional random number might be drawn during the initialization of the classification model, or the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
-This problem also applies to the initialization of the classification models $\theta$, the initialization and querying of $Omega$ and the drawn mini batches during the training of $\theta$.
+This additional random number might be drawn during the initialization of the classification model, the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
+This problem also applies to the initialization of the classification model $\theta$, the initialization and querying of $\Omega$ and the drawn mini batches during the training of $\theta$.
 The desired control over all these processes can be archived by assigning a separate seeded random number generator to all these processes.
 To the best of our knowledge, we are the first work that discusses this issue and proposes a solution for it.
-We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparibility}\label{sec:comparibility}
-{\color{red} Version: brain dump}\\
-We split the evaluation model that generates accuracy scores and the model that generates features for the AL agent. \\
-This allows for dropout / ensemble approaches to invest more compute to generate features \\
-The normal evaluation model is available as a fallback of course \\
-This way, the comparison is always fair
+We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
+In summary, we introduce three different seeds: $s_\Omega$ for the acquisition function, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
+Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
 
@@
 \subsection{Evaluation}\label{sec:evaluation}
-{\color{red} Version: 0.1}\\
-We compare different AL algorithms based on their archived AUC score (Eq. \ref{eq:auc}).
+{\color{red} Version: 1.0}\\
+We compare different AL algorithms based on their median AUC score (Eq. \ref{eq:auc}) across multiple restarts of the experiment.
 This score does incorporate performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
-A good AL algorithm should be able to perform well in both scenarios. \\
-{\color{red} TODO}
+A good AL algorithm should be able to perform well in both scenarios. \\ [1mm]
+Since AUC is dependent on the chosen budget, we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
+In this work, we choose the budget per dataset to be the first point at which the oracle manages to reach a percentage of the upper bound performance.
+Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets.
+We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: {\color{red}TODO}. \\ [1mm]
+Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
+To create one slice in Fig.
+\ref{fig:restarts}, we used all our runs for the Splice dataset and drew multiple sub-sets of that respective size (X axis), effectively simulating an experiment with that many restarts.
+Calculating the AUC values for every drawn run allows us to draw the box-plot.
+Each box represents the variance of an evaluation if conducted with the respective number of restarts.
+We can clearly observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{img/ablation_restarts.png}
+	\caption{Random draws from Splice with different numbers of repetitions.}
+	\label{fig:restarts}
+\end{figure}
+To combat is uncertain evaluation, we propose to repeat every experiment 50 times, to arrive at the true median AUC for each algorithm.
 
@@
 However, since this combinatorial problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence.
-\cite{zhou2021towards} has used simulated annealing to search for the best sequence and used the best found solution after a fixed time budget. 
-Even though the reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
+\cite{zhou2021towards} has used simulated annealing to search for the optimal sequence and used the best  solution found after a fixed time budget. 
+Even though their reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
 \\
-In this paper we propose a greedy oracle algorithm the constructs an approximation of the optimal sequence in an iterative fashion.
+In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
 Our oracle simply tests every data point in the provided sample of unlabeled points by fitting the classifier and directly measuring the resulting validation performance.
@@
 \section{Implementation Details}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{State Space}
-{\color{red} Version: 0.1 Remove analogies to RL}\\
-Since every AL agent needs a different state space our environment exposes a full state to the agent, so that the agent has full control of what information will be used. \\ [1mm]
-The state can include the following information:
-\begin{itemize}
-	\item The entire labeled dataset $\mathcal{L}^{(i)}$
-	\item The entire unlabeled dataset $\mathcal{U}^{(i)}$
-	\item A histogram of labeled points per class (count)
-	\item The available budget $B$
-	\item Number of added datapoints $|\mathcal{L}| - |\mathcal{S}|$
-	\item The initial validation accuracy $\operatorname{acc}^{(0)}$ and current validation accuracy $\operatorname{acc}^{(i)}$
-	\item The current classification model including all model weights $\theta^{(i)}$
-	\item The current optimizer including it's full state
-	\item The current sample of unlabeled points $u^{(i)}$
-\end{itemize}
 

### 2023-05-11 agents/__init__.py

@@ -4,3 +4,4 @@ from agents.random_agent import RandomAgent
 from agents.sal import SAL
 from agents.shannon_entropy import ShannonEntropy
 from agents.bald import BALD
+from agents.typiclust import TypiClust

### 2023-05-11 agents/bald.py

@@ -8,8 +8,8 @@ from core.agent import BaseAgent
 
 class BALD(BaseAgent):
 
-    def __init__(self, agent_rng, config, dropout_trials=5):
-        super().__init__(agent_rng, config)
+    def __init__(self, agent_seed, config, dropout_trials=5):
+        super().__init__(agent_seed, config)
         assert "current_run_info" in config and "encoded" in config["current_run_info"]
         self.dropout_trials = dropout_trials
         self.dropout_model = None
@@ -24,15 +24,17 @@ class BALD(BaseAgent):
             config[class_key]["dropout"] = 0.2
 
 
-    def predict(self, state_ids: list[int],
-                      x_unlabeled: Tensor,
+    def predict(self, x_unlabeled: Tensor,
                       x_labeled: Tensor, y_labeled: Tensor,
                       per_class_instances: dict,
                       budget:int, added_images:int,
                       initial_test_acc:float, current_test_acc:float,
-                      classifier: Module, optimizer: Optimizer) -> Union[Tensor, dict]:
+                      classifier: Module, optimizer: Optimizer,
+                      sample_size=100) -> Union[Tensor, dict]:
 
         with torch.no_grad():
+            replacement_needed = len(x_unlabeled) < sample_size
+            state_ids = self.agent_rng.choice(len(x_unlabeled), sample_size, replace=replacement_needed)
             classifier.train()
             for m in classifier.modules():
                 if isinstance(m, torch.nn.BatchNorm2d) or isinstance(m, torch.nn.BatchNorm1d):
@@ -61,4 +63,4 @@ class BALD(BaseAgent):
 
             u_x = g_x - f_x
 
-        return torch.argmax(u_x, dim=0)
+        return state_ids[torch.argmax(u_x, dim=0)]

### 2023-05-11 agents/coreset.py

@@ -14,19 +14,21 @@ class Coreset_Greedy(BaseAgent):
     Taken from https://github.com/svdesai/coreset-al
     """
 
-    def predict(self, state_ids: list[int],
-                      x_unlabeled: Tensor,
+    def predict(self, x_unlabeled: Tensor,
                       x_labeled: Tensor, y_labeled: Tensor,
                       per_class_instances: dict,
                       budget:int, added_images:int,
                       initial_test_acc:float, current_test_acc:float,
-                      classifier: Module, optimizer: Optimizer) -> Union[Tensor, dict]:
+                      classifier: Module, optimizer: Optimizer,
+                      sample_size=100) -> Union[Tensor, dict]:
 
         assert hasattr(classifier, "_encode"), "The provided model needs the '_encode' function"
         with torch.no_grad():
+            replacement_needed = len(x_unlabeled) < sample_size
+            state_ids = self.agent_rng.choice(len(x_unlabeled), sample_size, replace=replacement_needed)
             candidates = classifier._encode(x_unlabeled[state_ids])
             centers = classifier._encode(x_labeled)
             dist = pairwise_distances(candidates.detach().cpu(), centers.detach().cpu(), metric='euclidean')
             dist = np.min(dist, axis=1).reshape(-1, 1)
             dist = torch.from_numpy(dist)
-        return torch.argmax(dist, dim=0)
+        return state_ids[torch.argmax(dist, dim=0)]

### 2023-05-11 agents/margin.py

@@ -8,19 +8,21 @@ from core.agent import BaseAgent
 
 class MarginScore(BaseAgent):
 
-    def predict(self, state_ids: list[int],
-                      x_unlabeled: Tensor,
+    def predict(self, x_unlabeled: Tensor,
                       x_labeled: Tensor, y_labeled: Tensor,
                       per_class_instances: dict,
                       budget:int, added_images:int,
                       initial_test_acc:float, current_test_acc:float,
-                      classifier: Module, optimizer: Optimizer) -> Union[Tensor, dict]:
+                      classifier: Module, optimizer: Optimizer,
+                      sample_size=100) -> Union[Tensor, dict]:
 
         with torch.no_grad():
+            replacement_needed = len(x_unlabeled) < sample_size
+            state_ids = self.agent_rng.choice(len(x_unlabeled), sample_size, replace=replacement_needed)
             x_sample = x_unlabeled[state_ids]
             pred = classifier(x_sample).detach()
             pred = torch.softmax(pred, dim=1)
             two_highest, _ = pred.topk(2, dim=1)
             bVsSB = 1 - (two_highest[:, -2] - two_highest[:, -1])
             torch.unsqueeze(bVsSB, dim=-1)
-        return torch.argmax(bVsSB, dim=0)
+        return state_ids[torch.argmax(bVsSB, dim=0)]

### 2023-05-11 agents/random_agent.py

@@ -8,12 +8,11 @@ from core.agent import BaseAgent
 
 class RandomAgent(BaseAgent):
 
-    def predict(self, state_ids: list[int],
-                      x_unlabeled: Tensor,
+    def predict(self, x_unlabeled: Tensor,
                       x_labeled: Tensor, y_labeled: Tensor,
                       per_class_instances: dict,
                       budget:int, added_images:int,
                       initial_test_acc:float, current_test_acc:float,
                       classifier: Module, optimizer: Optimizer) -> Union[Tensor, dict]:
-        idx = int(self.agent_rng.random()*len(state_ids))
+        idx = int(self.agent_rng.random()*len(x_unlabeled))
         return torch.Tensor([[idx]]).int()

### 2023-05-11 agents/sal.py

@@ -113,8 +113,8 @@ class SAL(BaseAgent):
     Top-Level Shell for the framework
     """
 
-    def __init__(self, agent_rng, config, file="sar_23_01_16.pth", device=None):
-        super().__init__(agent_rng, config)
+    def __init__(self, agent_seed, config, file="sar_23_01_16.pth", device=None):
+        super().__init__(agent_seed, config)
         self.name += '_23_01_16'
 
         if device == None:
@@ -167,15 +167,17 @@ class SAL(BaseAgent):
         return state.cpu()
 
 
-    def predict(self, state_ids: list[int],
-                      x_unlabeled: Tensor,
+    def predict(self, x_unlabeled: Tensor,
                       x_labeled: Tensor, y_labeled: Tensor,
                       per_class_instances: dict,
                       budget:int, added_images:int,
                       initial_test_acc:float, current_test_acc:float,
-                      classifier: Module, optimizer: Optimizer) -> Union[Tensor, dict]:
+                      classifier: Module, optimizer: Optimizer,
+                      sample_size=100) -> Union[Tensor, dict]:
 
         with torch.no_grad():
+            replacement_needed = len(x_unlabeled) < sample_size
+            state_ids = self.agent_rng.choice(len(x_unlabeled), sample_size, replace=replacement_needed)
             sample_x = x_unlabeled[state_ids]
             sample_features = self._get_sample_features(sample_x, classifier, y_labeled.shape[1])
             interal_features = self._get_internal_features(initial_test_acc, current_test_acc, added_images, budget)
@@ -187,7 +189,7 @@ class SAL(BaseAgent):
                 "info": Batch(),
             })
             result = self.agent(data)
-        return torch.from_numpy(result.act)
+        return state_ids[torch.from_numpy(result.act)]
 
     def get_meta_data(self)->str:
         s = super().get_meta_data()

### 2023-05-11 agents/shannon_entropy.py

@@ -8,19 +8,21 @@ from core.agent import BaseAgent
 
 class ShannonEntropy(BaseAgent):
 
-    def predict(self, state_ids: list[int],
-                      x_unlabeled: Tensor,
+    def predict(self, x_unlabeled: Tensor,
                       x_labeled: Tensor, y_labeled: Tensor,
                       per_class_instances: dict,
                       budget:int, added_images:int,
                       initial_test_acc:float, current_test_acc:float,
-                      classifier: Module, optimizer: Optimizer) -> Union[Tensor, dict]:
+                      classifier: Module, optimizer: Optimizer,
+                      sample_size=100) -> Union[Tensor, dict]:
 
         with torch.no_grad():
+            replacement_needed = len(x_unlabeled) < sample_size
+            state_ids = self.agent_rng.choice(len(x_unlabeled), sample_size, replace=replacement_needed)
             x_sample = x_unlabeled[state_ids]
             pred = classifier(x_sample).detach()
             pred = torch.softmax(pred, dim=1)
             eps = 1e-7
             entropy = -torch.mean(pred * torch.log(eps + pred) + (1 + eps - pred) * torch.log(1 + eps - pred), dim=1)
             entropy = torch.unsqueeze(entropy, dim=-1)
-        return torch.argmax(entropy, dim=0)
+        return state_ids[torch.argmax(entropy, dim=0)]

### 2023-05-11 agents/typiclust.py

@@ -1,0 +1,101 @@
+from typing import Union
+import numpy as np
+import pandas as pd
+import faiss
+import torch
+from sklearn.cluster import MiniBatchKMeans, KMeans
+from core.agent import BaseAgent
+from torch import Tensor
+from torch.nn import Module
+from torch.optim import Optimizer
+
+
+
+
+class TypiClust(BaseAgent):
+    MIN_CLUSTER_SIZE = 5
+    MAX_NUM_CLUSTERS = 500
+    K_NN = 20
+
+    def predict(self, x_unlabeled:Tensor,
+                      x_labeled:Tensor, y_labeled:Tensor,
+                      per_class_instances:dict,
+                      budget:int, added_images:int,
+                      initial_test_acc:float, current_test_acc:float,
+                      classifier:Module, optimizer:Optimizer)->Union[Tensor, dict]:
+
+        num_clusters = min(len(x_labeled) + 1, self.MAX_NUM_CLUSTERS)
+        all_data = torch.concat([x_unlabeled, x_labeled], dim=0)
+        clusters = self._kmeans(all_data, num_clusters=num_clusters)
+        labels = np.copy(clusters)
+        # counting cluster sizes and number of labeled samples per cluster
+        cluster_ids, cluster_sizes = np.unique(labels, return_counts=True)
+        cluster_ids, cluster_sizes = self._fill_in_zero_size_clusters(cluster_ids, cluster_sizes)
+        existing_indices = np.arange(len(x_unlabeled), len(x_unlabeled)+len(x_labeled))
+        cluster_labeled_counts = np.bincount(labels[existing_indices], minlength=len(cluster_ids))
+        clusters_df = pd.DataFrame({'cluster_id': cluster_ids,
+                                    'cluster_size': cluster_sizes,
+                                    'existing_count': cluster_labeled_counts,
+                                    'neg_cluster_size': -1 * cluster_sizes})
+        # drop too small clusters
+        clusters_df = clusters_df[clusters_df.cluster_size > self.MIN_CLUSTER_SIZE]
+
+        # sort clusters by lowest number of existing samples, and then by cluster sizes (large to small)
+        clusters_df = clusters_df.sort_values(['existing_count', 'neg_cluster_size'])
+        labels[existing_indices] = -1
+
+        # pick the most typical example from the top-ranked cluster
+        cluster_id = clusters_df.iloc[0].cluster_id
+        indices = (labels == cluster_id).nonzero()[0]
+        rel_feats = all_data[indices]
[…]

### 2023-05-11 core/helper_functions.py

@@ -164,5 +164,7 @@ def get_agent_by_name(name:str)->Union[Callable, BaseAgent]:
         return agents.BALD
     elif name == "sal":
         return agents.SAL
+    elif name == "typiclust":
+        return agents.TypiClust
     else:
         raise ValueError(f"Agent name '{name}' not recognized")

### 2023-05-12 doc/neurips_data_2023.tex
TopV2 / News update

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_data_2023

% ready for submission
\usepackage{neurips_data_2023}

% to compile a preprint version, add the [preprint] option, e.g.:
%     \usepackage[preprint]{neurips_data_2023}
% This will indicate that the work is currently under review.

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_data_2023}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_data_2023}

% Submissions to the datasets and benchmarks are typically non anonymous,
% but anonymous submissions are allowed. If you feel that you must submit 
% anonymously, you can compile an anonymous version by adding the [anonymous] 
% option, e.g.:
%     \usepackage[anonymous]{neurips_data_2023}
% This will hide all author names.

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\title{Formatting Instructions For the NeurIPS 2023 \\ Track on Datasets and Benchmarks}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  David S.~Hippocampus\thanks{Use footnote for providing further information
    about author (webpage, alternative address)---\emph{not} for acknowledging
    funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

\section{Submissions to the NeurIPS 2023 Track on Datasets and Benchmarks}

Please read the instructions below carefully and follow them faithfully.

\subsection{Style}

Papers must be prepared according to the
instructions presented here. Papers may only be up to {\bf nine} pages long,
including figures. Additional pages \emph{containing only acknowledgments and
references} are allowed. Papers that exceed the page limit will not be
reviewed, or in any other way considered for presentation at the conference.

Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
NeurIPS website as indicated below. Please make sure you use the current files
and not previous versions. Tweaking the style files may be grounds for
rejection.

\subsection{Retrieval of style files}

The style files for the NeurIPS Track on Datasets and Benchmarks and other information are available on the World Wide Web at
\begin{center}
  \url{http://www.neurips.cc/Conferences/2023/CallForDatasetsBenchmarks}
\end{center}
The file \verb+neurips_data_2023.pdf+ contains these instructions and illustrates the
various formatting requirements your submission must satisfy.

The only supported style file for NeurIPS 2023 is \verb+neurips_data_2023.sty+,
written for \LaTeXe{}. There are no supported style sheets for Microsoft Word, RTF, or other formats. The \LaTeX{} style file contains three optional arguments: \verb+final+, which
creates a camera-ready copy, \verb+preprint+, which creates a preprint for
submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
\verb+natbib+ package for you in case of package clash.

\paragraph{Preprint option}
If you wish to post a preprint of your work online, e.g., on arXiv, using the
NeurIPS style, please use the \verb+preprint+ option. This will create a version of your work with the text ``Preprint. Work in progress.''
in the footer. This version may be distributed as you see fit. Please \textbf{do
  not} use the \verb+final+ option, which should \textbf{only} be used for
papers accepted to the NeurIPS Track on Datasets and Benchmarks.

At submission time, please omit the \verb+final+ and \verb+preprint+
options. This will add line numbers to aid
review. Please do \emph{not} refer to these line numbers in your paper as they
will be removed during generation of camera-ready copies. Note that submissions to the NeurIPS Track on Datasets and Benchmarks are reviewed in a single-blind fashion and therefore not anonymous. This is because datasets can typically not be shared in a non-anonymous way. If you feel strongly that your work should be submitted anonymously, please use the \verb+anonymous+ option. This will create a version of your work with all author names hidden.

The file \verb+neurips_data_021.tex+ may be used as a ``shell'' for writing your
paper. All you have to do is replace the author, title, abstract, and text of
the paper with your own.

The formatting instructions contained in these style files are summarized in
Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

Authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper nouns),
flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the heading in
bold, flush left, and inline with the text, with the heading followed by 1\,em
of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.  Citations may be
author/year or numeric, as long as you maintain internal consistency.  As to the
format of the references themselves, any style is acceptable as long as it is
used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations appropriate for
use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may add the
following before loading the \verb+neurips_2023+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add the optional
argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{neurips_2023}
\end{verbatim}

As submission is double blind, refer to your own published work in the third
person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
previous work [4].'' If you cite your other papers that are not widely available
(e.g., a journal paper under review), use anonymous author names in the
citation, e.g., an author of the form ``A.\ Anonymous.''

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote, indicate
footnotes with a number\footnote{Sample of the first footnote.} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table number and
title always appear before the table.  See Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical rules.} We
strongly suggest the use of the \verb+booktabs+ package, which allows for
typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Final instructions}

Do not change any aspects of the formatting parameters in the style files.  In
particular, do not modify the width or length of the rectangle the text should
fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and not, for
example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file must only
contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
  available out-of-the-box on most Linux machines.

\item The IEEE has recommendations for generating PDF files whose fonts are also
  acceptable for NeurIPS. Please see
  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
  "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
  the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
figure width as a multiple of the line width as in the example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
necessary.

\subsection{Suppressing line numbers in supplementary materials}
If you need to suppress line numbers in the supplementary materials because 
they interfere with the text, for instance because you are including a 
data sheet in 2-column format, you can do so by
placing the following command before it:
\begin{verbatim}
    \let\linenumbers\nolinenumbers\nolinenumbers
\end{verbatim}

\begin{ack}
Use unnumbered first level headings for the acknowledgments. All acknowledgments
go at the end of the paper before the list of references. Moreover, you are required to declare
funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2023/PaperInformation/FundingDisclosure}.
You can use the \texttt{ack} environment provided in the style file. As opposed to the main NeurIPS track, acknowledgements do not need to be hidden.
\end{ack}

\section*{References}

References follow the acknowledgments. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references.
Note that the Reference section does not count towards the page limit.
\medskip

{
\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Checklist}

%%% BEGIN INSTRUCTIONS %%%
The checklist follows the references.  Please
read the checklist guidelines carefully for information on how to answer these
questions.  For each question, change the default \answerTODO{} to \answerYes{},
\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
justification to your answer}, either by referencing the appropriate section of
your paper or providing a brief inline description.  For example:
\begin{itemize}
  \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
  \item Did you include the license to the code and datasets? \answerNA{}
\end{itemize}
Please do not modify the questions and only use the provided macros for your
answers.  Note that the Checklist section does not count towards the page
limit.  In your paper, please delete this instructions block and only keep the
Checklist section heading above along with the questions/answers below.
%%% END INSTRUCTIONS %%%

\begin{enumerate}

\item For all authors...
\begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \answerTODO{}
  \item Did you describe the limitations of your work?
    \answerTODO{}
  \item Did you discuss any potential negative societal impacts of your work?
    \answerTODO{}
  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
    \answerTODO{}
\end{enumerate}

\item If you are including theoretical results...
\begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    \answerTODO{}
	\item Did you include complete proofs of all theoretical results?
    \answerTODO{}
\end{enumerate}

\item If you ran experiments (e.g. for benchmarks)...
\begin{enumerate}
  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
    \answerTODO{}
  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
    \answerTODO{}
	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
    \answerTODO{}
	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
    \answerTODO{}
\end{enumerate}

\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
\begin{enumerate}
  \item If your work uses existing assets, did you cite the creators?
    \answerTODO{}
  \item Did you mention the license of the assets?
    \answerTODO{}
  \item Did you include any new assets either in the supplemental material or as a URL?
    \answerTODO{}
  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
    \answerTODO{}
  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
    \answerTODO{}
\end{enumerate}

\item If you used crowdsourcing or conducted research with human subjects...
\begin{enumerate}
  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
    \answerTODO{}
  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
    \answerTODO{}
  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
    \answerTODO{}
\end{enumerate}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Appendix}

Include extra information in the appendix. This section will often be part of the supplemental material. Please see the call on the NeurIPS website for links to additional guides on dataset publication.

\begin{enumerate}

\item Submission introducing new datasets must include the following in the supplementary materials:
\begin{enumerate}
  \item Dataset documentation and intended uses. Recommended documentation frameworks include datasheets for datasets, dataset nutrition labels, data statements for NLP, and accountability frameworks.
  \item URL to website/platform where the dataset/benchmark can be viewed and downloaded by the reviewers.
  \item Author statement that they bear all responsibility in case of violation of rights, etc., and confirmation of the data license.
  \item Hosting, licensing, and maintenance plan. The choice of hosting platform is yours, as long as you ensure access to the data (possibly through a curated interface) and will provide the necessary maintenance.
\end{enumerate}

\item To ensure accessibility, the supplementary materials for datasets must include the following:
\begin{enumerate}
  \item Links to access the dataset and its metadata. This can be hidden upon submission if the dataset is not yet publicly available but must be added in the camera-ready version. In select cases, e.g when the data can only be released at a later date, this can be added afterward. Simulation environments should link to (open source) code repositories.
  \item The dataset itself should ideally use an open and widely used data format. Provide a detailed explanation on how the dataset can be read. For simulation environments, use existing frameworks or explain how they can be used.
  \item Long-term preservation: It must be clear that the dataset will be available for a long time, either by uploading to a data repository or by explaining how the authors themselves will ensure this.
  \item Explicit license: Authors must choose a license, ideally a CC license for datasets, or an open source license for code (e.g. RL environments).
  \item Add structured metadata to a dataset's meta-data page using Web standards (like schema.org and DCAT): This allows it to be discovered and organized by anyone. If you use an existing data repository, this is often done automatically.
  \item Highly recommended: a persistent dereferenceable identifier (e.g. a DOI minted by a data repository or a prefix on identifiers.org) for datasets, or a code repository (e.g. GitHub, GitLab,...) for code. If this is not possible or useful, please explain why.
\end{enumerate}

\item For benchmarks, the supplementary materials must ensure that all results are easily reproducible. Where possible, use a reproducibility framework such as the ML reproducibility checklist, or otherwise guarantee that all results can be easily reproduced, i.e. all necessary datasets, code, and evaluation procedures must be accessible and documented.

\item For papers introducing best practices in creating or curating datasets and benchmarks, the above supplementary materials are not required.
\end{enumerate}

\end{document}

### 2023-05-12 datasets/topv2.py

@@ -21,6 +21,7 @@ class TopV2(BaseDataset):
         self.raw_unzipped_file = os.path.join(cache_folder, "optimal_al_data/intent_classification/data/TOP.pkl")
         self.raw_train_file = os.path.join(cache_folder, "topv2_train.txt")
         self.raw_test_file = os.path.join(cache_folder, "topv2_test.txt")
+        self.embedding_data_file = os.path.join(cache_folder, "topv2_al_embeddings.pt")
         fitting_mode = "from_scratch" if encoded else "finetuning"
         super().__init__(cache_folder, config, pool_rng, encoded,
                          data_file, fitting_mode)
@@ -39,37 +40,54 @@ class TopV2(BaseDataset):
             z = zipfile.ZipFile(self.raw_zip_file)
             z.extract('intent_classification/data/TOP.pkl', os.path.join(self.cache_folder, "optimal_al_data"))
 
-        print("Extracting and Embedding...")
+        print("Extracting and Processing...")
         if exists(self.raw_unzipped_file):
             data = pickle.load(open(self.raw_unzipped_file, "rb"))
             all_points = data[self.config["dataset"]["domain"].lower()]["raw"]
-            embeddings = data['GLOVE_EMBEDDING']
+            embedding_dict = data['GLOVE_EMBEDDING']
             x = [p[0] for p in all_points]
-            x = self._embed_texts(x, embeddings)
             y = [p[1] for p in all_points]
+            # x = self._embed_texts(x, embeddings)
+
+            # Save embedding dict as numpy array
+            embeddings = np.array(list(embedding_dict.values())).astype(np.float16)
+            num_embedding, emb_dim = embeddings.shape
+            unk_idx = num_embedding
+            pad_idx = num_embedding + 1
+            word_embedding_data = torch.cat((torch.tensor(embeddings).float(),
+                                             torch.randn(1, emb_dim) * 0.01,   # <unk>
+                                             torch.zeros(1, emb_dim)           # <pad>
+                                            ), dim=0)
+            torch.save(word_embedding_data, self.embedding_data_file)
+
+            token_x = self._tokenize_texts(x, embedding_dict)
+            token_x = self._pad_texts(token_x, pad_idx)
+            token_x = torch.stack(token_x)
+
             one_hot_y = np.zeros((len(y), max(y) + 1))
             one_hot_y[np.arange(len(y)), y] = 1
             all_ids = list(range(len(x)))
[…]

### 2023-05-12 doc/al_benchmark_paper.tex
lstm model _encode function

@@
 \documentclass[]{article}
-\usepackage[a4paper, total={6.5in, 8.5in}]{geometry}
+%\usepackage[a4paper, total={6.5in, 8.5in}]{geometry}
+\usepackage{color}
 \usepackage{hyperref}
@@
 \usepackage{algpseudocode}
-\usepackage{color}
+\usepackage{neurips_data_2023}
 
 \title{Towards Comparable Active Learning}
+\author{%
+	David S.~Hippocampus\thanks{Use footnote for providing further information
+		about author (webpage, alternative address)---\emph{not} for acknowledging
+		funding agencies.} \\
+	Department of Computer Science\\
+	Cranberry-Lemon University\\
+	Pittsburgh, PA 15213 \\
+	\texttt{hippo@cs.cranberry-lemon.edu} \\
+	% examples of more authors
+	% \And
+	% Coauthor \\
+	% Affiliation \\
+	% Address \\
+	% \texttt{email} \\
+	% \AND
+	% Coauthor \\
+	% Affiliation \\
+	% Address \\
+	% \texttt{email} \\
+	% \And
+	% Coauthor \\
+	% Affiliation \\
+	% Address \\
+	% \texttt{email} \\
+	% \And
+	% Coauthor \\
+	% Affiliation \\
+	% Address \\
+	% \texttt{email} \\
+}
 
@@
 \maketitle
+
+\begin{abstract}
+	The abstract paragraph should be indented $\frac{1}{2}$~inch (3~picas) on
+	both the left- and right-hand margins. Use 10~point type, with a vertical
+	spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
+	bold, and in point size 12. Two line spaces precede the abstract. The abstract
+	must be limited to one paragraph.
+\end{abstract}
 
@@
 
-\subsection{Contribution}
+\subsection{Contributions}
+\begin{enumerate}
+	\item Evaluation of Active Learning algorithms on datasets from 4 different domains, including synthetic data that highlights principled shortcomings of existing approaches.
+	\item Novel experimental protocol for seeding the experiment with 3 different seeds to allow full control and reproducibility and analysis of how many restarts are required to converge to the true median performance reliably.
+	\item Simple algorithm for an Oracle-Curve that can be constructed greedily and does not rely on search.
+\end{enumerate}
 
@@
 
-\section{Comparison of different sample sizes $\tau$}
+\section{Comparison of different sample sizes}
 \begin{figure}[H]

### 2023-05-17 doc/al_benchmark_paper.tex

@@
 \section{Introduction}
+{\color{red} Version: Braindump}\\
 Among others, the authors of \cite{zhou2021towards} have pointed out severe inconsistencies in results of AL papers in recent years. 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sampling Strategies}\label{sec:sampling_strategies}
+\paragraph{Uncertainty Sampling}
+\paragraph{BALD}
+\paragraph{BADGE}
+\paragraph{Coreset}
+\paragraph{TypiClust}
+%
+\subsubsection{Honorable Mentions}
+\paragraph{Learning Loss for AL}
+Introduces an updated classification model training via an auxiliary loss and therefore cannot be compared fairly against classification models without this boosted training regime.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Choosing the Classifier}\label{sec:choosing_the_classifier}

### 2023-05-18 doc/al_benchmark_paper.tex

@@
 \paragraph{Learning Loss for AL}
-Introduces an updated classification model training via an auxiliary loss and therefore cannot be compared fairly against classification models without this boosted training regime.
+Introduces an updated classification model training with an auxiliary loss and therefore cannot be compared fairly against classification models without this boosted training regime.
 

### 2023-05-18 doc/al_benchmark_paper.tex
finishing the runs

@@
 \section{Introduction}
-{\color{red} Version: Braindump}\\
+{\color{red} Version: 1.0}\\
+Active Learning (AL) plays an important role in our society that applies machine learning to more and more areas and therefore has a high demand for labeled data in more and more areas.
+A problem that concerns academic researchers and practitioners in businesses alike and even could be extended to education in schools and hobbyists around the world.
+On top of providing a principled way to labeled unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
+Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
 Among others, the authors of \cite{zhou2021towards} have pointed out severe inconsistencies in results of AL papers in recent years. 
@@
 (\textit{We would need to switch from lowest expected loss to highest AUC}) \\
-Since combinatorial problem of finding the optimal subset $\mathcal{D}_\text{train}$ is computational not feasible, we allow sequential construction of $\mathcal{D}_\text{train}$. \\ [2mm]
+Since combinatorial problem of finding the optimal subset $\mathcal{D}_\text{train}$ is computational not feasible, we allow sequential construction of $\mathcal{D}_\text{train}$. \\
 %
-To construct the active learning setting, we suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$.
-\\ [2mm]
-
+To construct the active learning setting from existing labeled datasets, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$.
+\\
+%
 Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
@@
 \end{equation}
-Where $\hat y_i$ is the retrained classification model after the i-\textit{th} instance was selected. \\ [1mm]
-%
-\textbf{Framing AL as RL}\\
-We define the active learning process as an adapted reinforcement learning loop $(S, A, \tau, \Omega, \omega)$ where an environment iteratively will expose a state $s \in S$ to an agent $\Omega$, which will choose actions $a \in A$.
-For each iteration $i$ the environment samples a subset of size $\tau$ of unlabeled instances $u^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$, constructs the state $s^{(i)} := \omega(u^{(i)})$ and presents it to the agent to select an action $a^{(i)} := \Omega(s^{(i)})$.
-The action $a^{(i)}$ is the index of the selected instance in $u^{(i)}$ out of all possible indices $A := [1 \ldots \tau]$.
-This process is repeated $B$ times $i := [1 \ldots B]$.
+Where $\hat y_i$ is the retrained classification model after the i-\textit{th} instance was selected.
 
-% TODO: remove tau from pseudocode
-\begin{minipage}{0.59\linewidth}
+\begin{minipage}{0.49\linewidth}
 	\begin{algorithm}[H]
-		\caption{Active Learning}\label{alg:active_learning}
+		\caption{Oracle}\label{alg:oracle}
 		\begin{algorithmic}[1]
@@
 			\Require $\tau$ \Comment{Unlabeled Sample Size}
-			\Require $\Omega$ \Comment{AL Agent}
-			\Require $\omega$ \Comment{Environment State function}
-			\State $\mathcal{L}^{(1)} \gets \operatorname{seed}(\mathcal{U})$  \Comment{Create the initial labeled set}
-			\State $\mathcal{U}^{(1)} \gets \mathcal{U}$
+			\State $\mathcal{L}^{(1)} \gets \operatorname{seed}(\mathcal{U})$  %\Comment{Create the initial labeled set}
+			\State $\mathcal{U}^{(1)} \gets \mathcal{U} \setminus \mathcal{L}^{(1)}$
 			\For{$i := 1 \ldots B$} 
 			\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$  %\Comment{$\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
-				\State $u^{(i)} \sim \text{unif}(1:|\mathcal{U}^{(i)}|)$
-				\State $a^{(i)} \gets \Omega(u^{(i)}) \hspace{2mm}|\hspace{2mm} a \in 1:|u^{(i)}|$ 
-				%\Comment{$a^{(i)}$ is an index inside of $u^{(i)}$}
-				\State $y^{(i)} \gets \operatorname{label}(u^{(i)}_{a})$ 
-				%\Comment{$u^{(i)}_{a}$ is shorthand for $u^{(i)}_{a^{(i)}}$}
-				\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_a, y^{(i)})\}$
-				\State $\mathcal{U}^{(i+1)} \gets \mathcal{U}^{(i)} \setminus \{u^{(i)}_a\}$
+			\State $u^{(i)} \underset{\tau}{\sim} \text{unif}(1:|\mathcal{U}^{(i)}|)$
+			\State $r* \gets 0$
+			\For{$j := 1 \ldots \tau$} %\Comment{Testing every unlabeled point}
+			\State $y^{(j)} \gets \operatorname{label}(u^{(i)}_{j})$
+			\State $\mathcal{L}^{(j)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_j, y^{(j)})\}$
+			\State $\text{acc}^{(j)} \gets \operatorname{Retrain}(\mathcal{L}^{(j)})$  
+			\State $r^{(j)} \gets \text{acc}^{(j)} - \text{acc}^{(i)}$
+			\If{$r^{(j)} > r^*$} %\Comment{Select point with largest increase in performance}
+			\State $r* \gets r^{(j)}$
+			\State $j* \gets j$
+			\EndIf  
+			\EndFor
+			\If{$r^* = 0$}
+			\State $j* \gets \operatorname{margin}(u^{(i)}, \hat y_\theta)$
+			\EndIf
+			\State $y^{(i)} \gets \operatorname{label}(u^{(i)}_{j*})$
+			\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_{j*}, y^{(i)})\}$
+			\State $\mathcal{U}^{(i+1)} \gets \mathcal{U}^{(i)} \setminus \{u^{(i)}_{j*}\}$
 			\EndFor
@@
 \hspace{3mm}
-\begin{minipage}{0.35\linewidth}
+\begin{minipage}{0.5\linewidth}
 	\begin{algorithm}[H]
@@
 			\Require $\mathcal{L}$ \Comment{Labeled Pool}
-			\Require $\mathcal{L}^\text{test}$ \Comment{Labeled Test Data}
-			\Require $\hat y_\theta$ \Comment{Classification Model}
+			\Require $\mathcal{D}_\text{test}$ \Comment{Test Data}
+			\Require $\hat y_\theta$ \Comment{Class. Model}
 			\Require $e^\text{max}$ \Comment{Maximum Epochs}
@@
 	\end{algorithm}
+	In Alg.
+\ref{alg:oracle} $\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{D}_\text{test}, \hat y_\theta, e^\text{max})$, $\operatorname{label}(u^{(i)}_{j})$ recovers the label of instance $u^{(i)}_{j}$ and $\operatorname{margin}(u^{(i)}, \hat y_\theta)$ applies margin sampling to select as a fallback, when the oracle has not found another option that induces a positive change in classification performance.
 \end{minipage}
-
-\begin{algorithm}
-	\caption{Oracle}\label{alg:oracle}
-	\begin{algorithmic}[1]
-		\Require $\mathcal{U}$ \Comment{Unlabeled Pool}
-		\Require $\tau$ \Comment{Unlabeled Sample Size}
-		\Require $\Omega$ \Comment{AL Agent}
-		\Require $\omega$ \Comment{Environment State function}
-		\State $\mathcal{L}^{(1)} \gets \operatorname{seed}(\mathcal{U})$  \Comment{Create the initial labeled set}
-		\State $\mathcal{U}^{(1)} \gets \mathcal{U}$
-		\For{$i := 1 \ldots B$} 
-			\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$  \Comment{$\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
-			\State $u^{(i)} \sim \text{unif}(1:|\mathcal{U}^{(i)}|)$
-			\State $r* \gets -\infty$
-			\State $j* \gets -1$
-			\For{$j := 1 \ldots \tau$} \Comment{Testing every unlabeled point}
-				\State $y^{(j)} \gets \operatorname{label}(u^{(i)}_{j})$
-				\State $\mathcal{L}^{(j)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_j, y^{(j)})\}$
-				\State $\text{acc}^{(j)} \gets \operatorname{Retrain}(\mathcal{L}^{(j)})$  
-				\State $r^{(j)} \gets \text{acc}^{(j)} - \text{acc}^{(i)}$
-				\If{$r^{(j)} > r^*$} \Comment{Select point with largest increase in performance}
-					\State $r* \gets r^{(j)}$
-					\State $j* \gets j$
-				\EndIf  
-			\EndFor
-			\State $y^{(i)} \gets \operatorname{label}(u^{(i)}_{j*})$
-			\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_{j*}, y^{(i)})\}$
-			\State $\mathcal{U}^{(i+1)} \gets \mathcal{U}^{(i)} \setminus \{u^{(i)}_{j*}\}$
-		\EndFor
-		\State
-		\Return $\frac{1}{B} \sum_{i=1}^{B} \text{acc}^{(i)}$
-	\end{algorithmic}
-\end{algorithm}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Reproducibility}\label{sec:reproducibility}
-{\color{red} Version: 1.0}\\
-A big focus in this work is to provide an experimental setup that is fully reproducible independent from the dataset, classification model or AL algorithm used.
-In our opinion, an evaluation on one dataset with a given seed should always be done on the same sequence of batches $x_0, \ldots, x_B$.
-Even though different AL algorithms will pick different samples from these batches, making them unavailable for sampling in later batches, the theoretical decision tree for every possible choice in every iteration $i$ should stay the same.
-This behavior is not possible with the default choice of setting a global seed at the start of the experiment, since a single additional random draw from the random number generator completely changes the decision tree for the batches.
-This additional random number might be drawn during the initialization of the classification model, the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
-This problem also applies to the initialization of the classification model $\theta$, the initialization and querying of $\Omega$ and the drawn mini batches during the training of $\theta$.
-The desired control over all these processes can be archived by assigning a separate seeded random number generator to all these processes.
-To the best of our knowledge, we are the first work that discusses this issue and proposes a solution for it.
-We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
-In summary, we introduce three different seeds: $s_\Omega$ for the acquisition function, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
-Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Evaluation}\label{sec:evaluation}
@@
 We compare different AL algorithms based on their median AUC score (Eq. \ref{eq:auc}) across multiple restarts of the experiment.
-This score does incorporate performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
+Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split to mimic the leave-one-out protocol for cross-validation.
+The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
 A good AL algorithm should be able to perform well in both scenarios. \\ [1mm]
 Since AUC is dependent on the chosen budget, we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
-In this work, we choose the budget per dataset to be the first point at which the oracle manages to reach a percentage of the upper bound performance.
+In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
 Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets.
-We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: {\color{red}TODO}. \\ [1mm]
+We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
 Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
-To create one slice in Fig.
-\ref{fig:restarts}, we used all our runs for the Splice dataset and drew multiple sub-sets of that respective size (X axis), effectively simulating an experiment with that many restarts.
-Calculating the AUC values for every drawn run allows us to draw the box-plot.
-Each box represents the variance of an evaluation if conducted with the respective number of restarts.
+To create Fig. \ref{fig:restarts} we used all our 50 runs from the margin/random sampling algorithm on the splice dataset.
+From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
+One AUC value like this corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
+To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
+Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
 We can clearly observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
@@
 To combat is uncertain evaluation, we propose to repeat every experiment 50 times, to arrive at the true median AUC for each algorithm.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Reproducibility}\label{sec:reproducibility}
+{\color{red} Version: 1.0}\\
+A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
+In our opinion, an evaluation on one dataset with a given seed should always be done on the same validation split as well as using the same random state for all included systems like the mini batch sampler for the classifier model training.
+Even though different AL algorithms will pick different samples, making them unavailable for sampling in earlier or later batches, the theoretical decision tree for every possible choice in every iteration $i$ should stay the same.
+Since every possible trajectory cannot be precomputed and stored to disk, we need to resort to seeding the evaluation.
+The default choice of setting a global seed at the start of the experiment is not sufficient here, since a single additional random draw from the random number generator completely changes the behavior of all other systems.
+This additional random number might be drawn during the initialization of the classification model, the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
+This problem also applies to the initialization of the classification model $\theta$ and the initialization and querying of $\Omega$.
+The desired control over all these processes can only be archived by assigning a separate random number generator to all these processes.
+To the best of our knowledge, we are the first work that discusses this issue and proposes a solution for it.
+We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
+In summary, we introduce three different seeds: $s_\Omega$ for the acquisition function, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
+Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
 
@@
 
+\section{Alternative Plot for Restarts Ablation}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\linewidth]{img/ablation_restarts_2.png}
+\end{figure}
+
 \section{Comparison of different sample sizes}
@@
 
+\section{AL Pseudocode}
+\begin{algorithm}[H]
+	\caption{Active Learning}\label{alg:active_learning}
+	\begin{algorithmic}[1]
+		\Require $\mathcal{U}$ \Comment{Unlabeled Pool}
+		\Require $\tau$ \Comment{Unlabeled Sample Size}
+		\Require $\Omega$ \Comment{AL Agent}
+		\Require $\omega$ \Comment{Environment State function}
+		\State $\mathcal{L}^{(1)} \gets \operatorname{seed}(\mathcal{U})$  \Comment{Create the initial labeled set}
+		\State $\mathcal{U}^{(1)} \gets \mathcal{U}$
+		\For{$i := 1 \ldots B$}
+		\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$  %\Comment{$\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
+		%\State $u^{(i)} \sim \text{unif}(1:|\mathcal{U}^{(i)}|)$
+		\State $a^{(i)} \gets \Omega(\mathcal{U}^{(i)}) \hspace{1mm};\hspace{2mm} a \in 1:|\mathcal{U}^{(i)}|$ 
+		%\Comment{$a^{(i)}$ is an index inside of $u^{(i)}$}
+		\State $y^{(i)} \gets \operatorname{label}(\mathcal{U}^{(i)}_{a})$ 
+		%\Comment{$u^{(i)}_{a}$ is shorthand for $u^{(i)}_{a^{(i)}}$}
+		\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(\mathcal{U}^{(i)}_a, y^{(i)})\}$
+		\State $\mathcal{U}^{(i+1)} \gets \mathcal{U}^{(i)} \setminus \{\mathcal{U}^{(i)}_a\}$
+		\EndFor
+		\State
+		\Return $\frac{1}{B} \sum_{i=1}^{B} \text{acc}^{(i)}$
+	\end{algorithmic}
+\end{algorithm}
+
 \end{document}

### 2023-05-18 results/generate_result_table.py

@@ -6,7 +6,7 @@ import numpy as np
 def generate_table(agents, datasets, out_file):
     result_table = pd.DataFrame(index=agents, columns=datasets)
     for dataset in datasets:
-        dataset_folder = os.path.join("../runs", dataset)
+        dataset_folder = os.path.join("runs", dataset)
         if not os.path.isdir(dataset_folder):
             continue
         agent_aucs = dict()
@@ -22,17 +22,21 @@ def generate_table(agents, datasets, out_file):
                     values = accuracies.values
                     auc = np.sum(values, axis=0) / values.shape[0]
                     result_table[dataset][agent] = "%1.3f +- %1.2f"%(np.mean(auc).item(), np.std(auc).item())
+            else:
+                print(f"{agent} is missing for {dataset}")
 
     if len(result_table) > 0:
-        result_table_file = os.path.join("../runs", out_file)
+        result_table_file = os.path.join("runs", out_file)
         if os.path.exists(result_table_file):
             os.remove(result_table_file)
         result_table.to_csv(result_table_file)
 
 
-agents = ["Oracle", "SAL_23_01_16", "Coreset_Greedy", "MarginScore", "ShannonEntropy", "RandomAgent"]
+agents = ["Oracle", "SAL_23_01_16", "Coreset_Greedy", "TypiClust", "MarginScore", "ShannonEntropy", "RandomAgent", "Badge", "BALD", ]
 vector_data = ["Splice", "DNA", "USPS"]
 img_data = ["Cifar10Encoded", "FashionMnistEncoded"]
+text_data = ["TopV2", "News"]
 
 generate_table(agents, vector_data, "result_vector.csv")
 generate_table(agents, img_data, "result_img.csv")
+generate_table(agents, text_data, "result_text.csv")

### 2023-05-19 doc/al_benchmark_paper.tex
typiclust bugfix normalized embeddings

@@
 		\hline
-		Oracle          & 0.835 $\pm$ 0.01 & 0.879 $\pm$ 0.01 & 0.870 $\pm$ 0.01 \\
-		SAL\_23\_01\_16 & 0.808 $\pm$ 0.01 & 0.857 $\pm$ 0.02 & 0.865 $\pm$ 0.01 \\
-		Coreset\_Greedy & 0.808 $\pm$ 0.01 & 0.853 $\pm$ 0.02 & 0.849 $\pm$ 0.01 \\
-		MarginScore     & 0.806 $\pm$ 0.01 & 0.855 $\pm$ 0.02 & 0.864 $\pm$ 0.01 \\
-		ShannonEntropy  & 0.808 $\pm$ 0.01 & 0.855 $\pm$ 0.02 & 0.864 $\pm$ 0.01 \\
-		RandomAgent     & 0.797 $\pm$ 0.01 & 0.823 $\pm$ 0.02 & 0.831 $\pm$ 0.01
+		Oracle          & 0.830 +- 0.01 & 0.836 +- 0.02 & 0.823 +- 0.01 \\
+		SAL & 0.799 +- 0.01 & 0.797 +- 0.03 & 0.809 +- 0.01 \\
+		Coreset & 0.800 +- 0.01 & 0.795 +- 0.03 & 0.787 +- 0.02 \\
+		TypiClust       & 0.790 +- 0.01 & 0.771 +- 0.04 & 0.761 +- 0.02 \\
+		MarginScore     & 0.797 +- 0.02 & 0.795 +- 0.04 & 0.808 +- 0.01 \\
+		ShannonEntropy  & 0.799 +- 0.02 & 0.794 +- 0.04 & 0.807 +- 0.01 \\
+		RandomAgent     & 0.788 +- 0.01 & 0.765 +- 0.03 & 0.772 +- 0.01 \\
+		Badge           & 0.807 +- 0.01 & 0.769 +- 0.06 & 0.797 +- 0.02 \\
+		BALD            & 0.811 +- 0.01 & 0.743 +- 0.04 & 0.717 +- 0.05
 	\end{tabular}
+	\caption{AUC values for all algorithms on the tabular datasets. Higher is better.}
 \end{table}
 %
-\begin{table}[]
-	\begin{tabular}{l|ll}
-		& Cifar10 & FashionMnist \\
-		\hline
-		Oracle          & 0.718 $\pm$ 0.01  & 0.721 $\pm$ 0.01       \\
-		SAL\_23\_01\_16 & 0.660 $\pm$ 0.01  & 0.667 $\pm$ 0.01       \\
-		Coreset\_Greedy & 0.668 $\pm$ 0.01  & 0.674 $\pm$ 0.01       \\
-		MarginScore     & 0.674 $\pm$ 0.01  & 0.677 $\pm$ 0.01       \\
-		ShannonEntropy  & 0.671 $\pm$ 0.01  & 0.680 $\pm$ 0.01       \\
-		RandomAgent     & 0.632 $\pm$ 0.02  & 0.633 $\pm$ 0.01      
-	\end{tabular}
-\end{table}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\linewidth]{img/eval_dna}
+	\caption{Results for all algorithms on the DNA dataset}
+\end{figure}
 %
-\begin{figure}[H]
+\begin{figure}
 \centering
-	\includegraphics[width=0.8\linewidth]{img/eval_splice_embedded}
-	\caption{Results for all algorithms on the pre-encoded Splice dataset}
-\end{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.8\linewidth]{img/eval_splice_embedded_cd}
-\caption{Critical Difference Diagram for Splice where every restart of the algorithm is one sample for the Wilcoxon-Holm method}
-\end{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.8\linewidth]{img/eval_macro}
-\caption{Critical Difference Diagram for Splice, DNA, USPS and Cifar10}
+	\includegraphics[width=0.8\linewidth]{img/macro_al.png}
+	\caption{Critical Difference Diagram for all datasets and algorithms}
 \end{figure}
@@
 \appendix
+
+\section{Other result tables}
+\begin{table}[H]
+	\begin{tabular}{l|lll}
+		& SpliceEncoded & DNAEncoded    & USPSEncoded   \\
+		\hline
+		Oracle          & 0.754 +- 0.02 & 0.726 +- 0.02 & 0.674 +- 0.01 \\
+		SAL & 0.675 +- 0.03 & 0.640 +- 0.04 & 0.634 +- 0.01 \\
+		Coreset & 0.690 +- 0.02 & 0.644 +- 0.05 & 0.607 +- 0.02 \\
+		TypiClust       & 0.695 +- 0.02 & 0.660 +- 0.03 & 0.643 +- 0.01 \\
+		MarginScore     & 0.675 +- 0.03 & 0.643 +- 0.05 & 0.632 +- 0.02 \\
+		ShannonEntropy  & 0.673 +- 0.03 & 0.638 +- 0.05 & 0.626 +- 0.02 \\
+		RandomAgent     & 0.680 +- 0.03 & 0.633 +- 0.04 & 0.594 +- 0.02 \\
+		Badge           & 0.670 +- 0.04 & 0.600 +- 0.07 & 0.597 +- 0.02 \\
+		BALD            & 0.660 +- 0.04 & 0.597 +- 0.06 & 0.652 +- 0.01
+	\end{tabular}
+	\caption{AUC values for all algorithms on the encoded tabular datasets. Higher is better.}
+\end{table}
+%
+\begin{table}[H]
+	\begin{tabular}{l|ll}
+		& Cifar10Encoded & FashionMnistEncoded \\
+		\hline
+		Oracle          & 0.699 +- 0.01  & 0.721 +- 0.01       \\
+		SAL & 0.623 +- 0.01  & 0.667 +- 0.01       \\
+		Coreset & 0.641 +- 0.01  & 0.674 +- 0.01       \\
+		TypiClust       & 0.627 +- 0.01  & 0.638 +- 0.01       \\
+		MarginScore     & 0.641 +- 0.01  & 0.676 +- 0.01       \\
+		ShannonEntropy  & 0.637 +- 0.02  & 0.680 +- 0.01       \\
+		RandomAgent     & 0.598 +- 0.02  & 0.633 +- 0.01       \\
+		Badge           & 0.645 +- 0.02  & 0.681 +- 0.01       \\
+		BALD            & 0.633 +- 0.01  & 0.666 +- 0.01      
+	\end{tabular}
+	\caption{AUC values for all algorithms on the encoded image datasets. Higher is better.}
+\end{table}
+%
+\begin{table}[H]
+	\begin{tabular}{l|ll}
+		& TopV2         & News          \\
+		\hline
+		Oracle          & 0.860 +- 0.01 & 0.448 +- 0.01 \\
+		SAL & 0.831 +- 0.01 & 0.312 +- 0.01 \\
+		Coreset & 0.816 +- 0.02 & 0.355 +- 0.02 \\
+		TypiClust       & 0.718 +- 0.02 & 0.314 +- 0.01 \\
+		MarginScore     & 0.821 +- 0.02 & 0.357 +- 0.01 \\
+		ShannonEntropy  & 0.794 +- 0.02 & 0.342 +- 0.01 \\
+		RandomAgent     & 0.776 +- 0.02 & 0.349 +- 0.01 \\
+		Badge           & 0.824 +- 0.01 & 0.343 +- 0.01 \\
+		BALD            & 0.825 +- 0.02 & 0.347 +- 0.01
+	\end{tabular}
+\caption{AUC values for all algorithms on the text datasets. Higher is better.}
+\end{table}
 

### 2023-05-19 core/data.py

@@ -138,7 +138,9 @@ class BaseDataset(ABC):
             for x in tqdm(test_loader):
                 x_enc = model(x[0])
                 enc_test = torch.cat([enc_test, x_enc], dim=0)
-            self.x_train, self.y_train, self.x_test, self.y_test = enc_train, y_train, enc_test, y_test
+            enc_train, enc_test = normalize(enc_train, enc_test, mode="min_max")
+            self.x_train, self.y_train, self.x_test, self.y_test = to_torch(enc_train, torch.float32), y_train, \
+                                                                   to_torch(enc_test, torch.float32), y_test
 
     def _load_data(self, encoded) -> Union[None, Tuple]:
         """
@@ -169,7 +171,7 @@ class BaseDataset(ABC):
         from classifiers.classifier import construct_model
         model_rng = torch.Generator()
         model_rng.manual_seed(seed)
-        backbone, out_dim = construct_model(model_rng, self.x_shape, self.n_classes, config["pretext_encoder"], add_head=False)
+        backbone, out_dim = construct_model(model_rng, self, config["pretext_encoder"], add_head=False)
         config["pretext_encoder"]["encoder_dim"] = out_dim
         model = ContrastiveModel({'backbone': backbone, 'dim': config["pretext_encoder"]["encoder_dim"]},
                                  head="mlp", features_dim=config["pretext_encoder"]["feature_dim"])

### 2023-05-19 agents/typiclust.py
typiclust bugfix normalized embeddings

@@ -66,7 +66,7 @@ class TypiClust(BaseAgent):
 
     def _get_mean_nn_dist(self, features, num_neighbors, return_indices=False):
         distances, indices = self._get_nn(features, num_neighbors)
-        if distances.shape[1] > 0:
+        if distances.shape[1] > 1:
             mean_distance = distances.mean(axis=1)
         else:
             mean_distance = np.zeros(len(distances))

### 2023-05-19 download_all_datasets.py

@@ -6,7 +6,9 @@ import numpy as np
 
 parser = argparse.ArgumentParser()
 parser.add_argument("--data_folder", type=str, required=True)
+parser.add_argument("--encode", type=int, default=0)
 args = parser.parse_args()
+args.encode = bool(args.encode)
 
 all_names = [
     "splice",
@@ -26,7 +28,9 @@ for name in all_names:
         config = yaml.load(f, yaml.Loader)
     pool_rng = np.random.default_rng(1)
     DatasetClass = get_dataset_by_name(name)
-    dataset = DatasetClass(args.data_folder, config, pool_rng, encoded=False)
+    DatasetClass(args.data_folder, config, pool_rng, encoded=False)
+    if args.encode:
+        DatasetClass(args.data_folder, config, pool_rng, encoded=True)
 
 print("\n")
-print("> all datasets downloaded")
+print("> all datasets prepared")

### 2023-05-22 agents/typiclust.py
splice validation run

@@ -28,7 +28,7 @@ class TypiClust(BaseAgent):
         state_ids = self.agent_rng.choice(len(x_unlabeled), sample_size, replace=False)
         num_clusters = min(len(x_labeled) + 1, self.MAX_NUM_CLUSTERS)
         if len(x_unlabeled.size()) > 2:
-            assert hasattr(classifier, "_encode"), "When using TypiClust with un-embedded data, the classifier needs the _encode method"
+            assert hasattr(classifier, "_encode"), "When using TypiClust with images or text, the classifier needs the _encode method"
             all_data = torch.concat([self._embed(x_unlabeled[state_ids], classifier),
                                      self._embed(x_labeled, classifier)], dim=0).cpu()
         else:

### 2023-05-23 doc/al_benchmark_paper.tex
typiclust bugfix normalized embeddings
splice validation run
img data normalization paper update

@@
 \begin{abstract}
-	The abstract paragraph should be indented $\frac{1}{2}$~inch (3~picas) on
-	both the left- and right-hand margins. Use 10~point type, with a vertical
-	spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
-	bold, and in point size 12. Two line spaces precede the abstract. The abstract
-	must be limited to one paragraph.
+	Active Learning (AL) is a technique used in machine learning to improve data efficiency by selecting the most informative samples for annotation. 
+	In this paper we address the issue of inconsistent results in active learning literature.
+	Authors of previous papers are constantly reporting significant performance improvements in their original work, while subsequent literature fails to reproduce those results.
+This inconsistency leads to a chaotic landscape of AL algorithms.
+	We propose the first AL benchmark that tests algorithms in all three major domains of tabular, image and text data.
+	Furthermore, we discuss a central problem for reproducing AL experiments with the default seeding setup that depends on a single seed per experiment.
+	...
 \end{abstract}
@@
 The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art results by significantly outperforming everyone else, while the vast majority of results proves to be non-reproducible.
+\\ [1mm]
+%
+To allow comparison between
 
@@
 	\includegraphics[width=\linewidth]{img/ablation_restarts.png}
-	\caption{Random draws from Splice with different numbers of repetitions.}
+	\caption{Random draws from an experimental distribution on the Splice datasets with different numbers of repetitions.
+Each point on the Y-axis represents a cross-validated result that could have been reported in a paper.
+This analysis shows the drastic differences in performance one could observe even when repeating an experiment 2-10 times.}
 	\label{fig:restarts}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Available Information}
+At each iteration $i$ the AL algorithm needs to pick an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \theta^{(i)}, \text{opt}_\theta\}$, where $\theta^{(i)}$ is the current classifier and $\text{opt}_\theta$ is the optimizer used to fit $\theta^{(i)}$.
+We allow algorithms to derive additional information of this set like predictions of the classifier, K-Means clustering or even training new classifiers.
+However, the algorithm may not incorporate external information like other datasets, queries to recover additional labels, or the test/validation set.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Sampling Strategies}\label{sec:sampling_strategies}
+We selected AL algorithms that have good performances reported by multiple different sources.
+To ensure a fair comparison we fixed the training process of our classification model as well as the set of available information for the algorithm and selected only those that can work under these restrictions:
 \paragraph{Uncertainty Sampling} 
-\paragraph{BALD}
-\paragraph{BADGE}
-\paragraph{Coreset}
+Tries to find the sample that the classifier is most uncertain about. For our benchmark we use entropy and margin (a.k.a. best-vs-second-best) sampling.
+\paragraph{BALD \cite{kirsch2019batchbald}}
+Applies the query-by-committee strategy of model ensembles to a single model by interpreting the model's parameters as distributions and then sample multiple outputs from them via Monte-Carlo dropout.
+\paragraph{BADGE \cite{ashdeep}} Uses gradient embeddings of unlabeled points to select samples where the classifier is expected to change a lot.
+The higher the magnitude of the gradient the higher is the expected improvement in model performance.
+\paragraph{Coreset \cite{sener2017active}}
+Employs K-Means clustering to try to cover all {\color{red}modalities (is this the right word?)} that are observed in the data.
+Selects the unlabeled sample that is the furthest away from all cluster centers.
+In this work we use the greedy variant of Coreset and we also encode the data with the current classifier $\theta_i$ before clustering.
 \paragraph{TypiClust}
+Relies on clustering similar to Coreset but proposes a new measure called "Typicality" to select unlabeled centers.
+Tries to select points that are in the densest regions of clusters that do not contain labeled samples yet.
+For this work we also use the current classifier $\theta_i$ to encode data if necessary (raw images or text).
 %
@@
 \paragraph{Learning Loss for AL}
-Introduces an updated classification model training with an auxiliary loss and therefore cannot be compared fairly against classification models without this boosted training regime.
+Introduces an updated training of the classification model with an auxiliary loss and therefore cannot be compared fairly against classification models without this boosted training regime.
 
@@
 Smaller classifiers also require fewer labeled datapoints to acquire performance close to the upper bound.
-This behavior can be observed in Figure {\color{red} TODO} \\ [1mm]
 For every dataset the chosen architecture's hyperparameters are optimized by to archive maximum upper bound performance.
-For an overview of architectures and hyperparameters please refer to Table \ref{tab:architectures}.
-\begin{table}
-	\centering
-	\begin{tabular}{l || l | l l l l}
-		Dataset & Architecture & Optimizer & LR & Weight Decay & Dropout \\
-		\hline
-		Splice & [24, 12] & NAdam & 0.0012 & 5.9e-5 & 0 \\
-		DNA & [24, 12] & NAdam & 0.0398 & 3.6e-5 & 0 \\
-		USPS & [24, 12] & Adam & 0.0081 & 1.5e-6 & 0 \\
-		FashionMnist & linear & - & - & - & - \\
-		Cifar10 & linear & - & - & - & - \\
-	\end{tabular}
-	\caption{Classifier architectures and optimized hyperparameters per dataset. Numbers in brackets signify a MLP with corresponding hidden layers.}
-	\label{tab:architectures}
-\end{table}
+For an overview of architectures and hyperparameters please refer to Appendix \ref{app:hyperparameters}.
 
@@
 \subsection{Training the Classifier}\label{sec:training_the_classifier}
-{\color{red} Version: 0.1 Full rework needed}\\
+{\color{red} Version: 0.1}\\
 Generally, the classification model can be trained in two ways.
@@
 We use \textbf{FashionMNIST} \cite{xiao2017fashion} and \textbf{Cifar10} \cite{krizhevsky2009learning}.
-Both datasets are normalized between [-1, 1].
+Both datasets are normalized according to their standard protocols.
 
@@
 For News Category we use  the 15 most common categories as indicated by its Kaggle site.
-We additionally drop sentences above 80 words to reduce the necessary padding (retaining 99,86\% of the data).
+We additionally drop sentences above 80 words to reduce the needed padding (retaining 99,86\% of the data).
 For TopV2, we are only using the "alarm" domain.
 Both datasets are encoded with pre-trained GloVe embeddings \cite{pennington2014glove}.
-Since neither set provided a fixed test set, we split random 5000 datapoints into a test set. \\ [3mm]
+Since neither dataset provided a fixed test set, we randomly split 7000 datapoints into a test set. \\ [3mm]
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Splice Embedded}
+\subsection{Results}
 
 \begin{table}[]
+	\centering
 	\begin{tabular}{l|lll}
@@
 	\end{tabular}
-	\caption{AUC values for all algorithms on the tabular datasets. Higher is better.}
+	\caption{AUC values for all algorithms on the tabular datasets. Higher is better.
+	{\color{red}Is the STD even helping us here? Since we repeated 50 times, we can use the values as "true medians"} }
 \end{table}
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{img/eval_dna}
-	\caption{Results for all algorithms on the DNA dataset}
+	\includegraphics[width=0.49\linewidth]{img/eval_dna}
+	\includegraphics[width=0.49\linewidth]{img/eval_splice.png}
+	\caption{Results for all algorithms on Splice and DNA, both from the tabular domain. Even within one domain, the performance of the same algorithm can vary drastically.}
+	\label{fig:eval_vector}
 \end{figure}
@@
 \centering
-	\includegraphics[width=0.8\linewidth]{img/macro_al.png}
-	\caption{Critical Difference Diagram for all datasets and algorithms}
+	\includegraphics[width=0.49\linewidth]{img/vector.png}
+	\includegraphics[width=0.49\linewidth]{img/vector_enc.png}
+	\includegraphics[width=0.49\linewidth]{img/text.png}
+	\includegraphics[width=0.49\linewidth]{img/all_data.png}
+	\caption{Critical Difference Diagram for all algorithms grouped by domain and all domains combined.
+Ranks are computed based on median AUC for each algorithm and dataset combination.
+Lower ranks are better.}
+	\label{fig:cd_diagrams}
 \end{figure}
+%
+From Fig. \ref{fig:eval_vector} we notice drastically different qualities for the same AL algorithm for different datasets.
+We would like to highlight that both datasets are tabular from the medical domain with similar number of features and classes, yet we see that i.e.
+BALD is the best algorithm for Splice and the worst algorithm for DNA.
+These inconsistencies are present between the datasets of all our tested domains, further highlighting the difficulties for comparing AL algorithms in terms of average performance.
+In order to provide a meaningful analysis of which algorithm can be expected to perform best on average we ranked the algorithm for each dataset based on their median AUC and displayed these rankings in critical difference diagrams \cite{IsmailFawaz2018deep}.
+In Fig. \ref{fig:cd_diagrams} we report the rankings split by domain as well as across all domains (excluding the toydata).
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Ablation Studies}
-\begin{itemize}
-	\item Setting $\tau$ to $|\mathcal{U}|$
-	\item Reduction of the test set for speed
-\end{itemize}
+\section{Conclusion}
+
+\newpage
 
@@
 
+\section{Hyperparameters per Dataset}\label{app:hyperparameters}
+\begin{table}[H]
+	\centering
+	\begin{tabular}{l || l | l l l l}
+		Dataset & Classifier & Optimizer & LR & Weight Decay & Dropout \\
+		\hline
+		Splice & [24, 12] & NAdam & 1.2e-3 & 5.9e-5 & 0 \\
+		SpliceEnc. & [12, 48] & NAdam & 6.2e-4 & 5.9e-6 & 0 \\
+		DNA & [24, 12] & NAdam & 3.9e-2 & 3.6e-5 & 0 \\
+		DNAEnc & [24] & NAdam & 1.6e-3 & 4e-4 & 0 \\
+		USPS & [24, 12] & Adam & 8.1e-3 & 1.5e-6 & 0 \\
+		USPS & [48, 24] & NAdam & 7.8e-3 & 1.9e-6 & 0 \\
+		FashionMnist & ResNet18 & NAdam & 1e-3 & 0 & 0 \\
+		FashionMnistEnc & linear & Adam & 1.6e-3 & 1e-5 & 5e-2 \\
+		Cifar10 & ResNet18 & NAdam & 1e-3 & 0 & 0 \\
+		Cifar10Enc & linear & NAdam & 1.7e-3 & 2.3e-5 & 0 \\
+		TopV2 & BiLSTM & NAdam & 1.5e-3 & 1.7e-7 & 5e-2 \\
+		News & BiLSTM & NAdam & 1.5e-3 & 1.7e-7 & 5e-2 \\
+	\end{tabular}
+	\caption{Classifier architectures and optimized hyperparameters per dataset. Numbers in brackets signify a MLP with corresponding hidden layers.}
+	\label{tab:architectures}
+\end{table}
+
 \section{Comparison of different sample sizes}

### 2023-05-24 doc/al_benchmark_paper.tex
mnist dataset

@@
 \\
-{\color{red} TODO} We include in this table any methodological paper that experiments on at least two domains. \\
-{\color{red} TODO} Define AL scenarios (really hard)
+{\color{red} TODO} We include in this table any methodological paper that experiments on at least two domains.
 \begin{table}[h]
@@
 	\begin{tabular}{l | c c c c c}
-		Paper & \# Datasets & Domains & Scenarios & Oracle & RL Alg. \\
+		Paper & Sampling & \# Datasets & Domains & Algorithms & Oracle \\
 		\hline
-		Beck et al. \cite{beck2021effective} & 4 & 1 & 3 & - & - \\
-		Hu et al. \cite{hu2021towards} & 5 & 2 & 1 & - & - \\
-		Li et al. \cite{li2022empirical} & 5 & 1 & 1 & - & - \\
-		Zhou et al. \cite{zhou2021towards} & 3 & 2 & 1 & \checkmark & - \\
-		\textbf{Ours} & 5 & 2 & 2 & \checkmark & -
+		Beck et al. \cite{beck2021effective} & batch & 4 & 1 & 7 & - \\
+		Hu et al. \cite{hu2021towards} & batch & 5 & 2 & 13 & - \\
+		Li et al. \cite{li2022empirical} & batch & 5 & 1 & 13 & - \\
+		Zhou et al. \cite{zhou2021towards} & batch & 3 & 2 & 2 & \checkmark \\
+		\textbf{Ours} & single & 9 & 4 & 6 & \checkmark 
 	\end{tabular}
@@
 
+\section{Limitations and Future Work}
+No batch AL \\
+No learned algorithms \\
+No improved classifier training (semi-supervised, etc.)
+
 \newpage
@@
 		Splice & [24, 12] & NAdam & 1.2e-3 & 5.9e-5 & 0 \\
-		SpliceEnc. & [12, 48] & NAdam & 6.2e-4 & 5.9e-6 & 0 \\
+		SpliceEnc. & linear & NAdam & 6.2e-4 & 5.9e-6 & 0 \\
 		DNA & [24, 12] & NAdam & 3.9e-2 & 3.6e-5 & 0 \\
-		DNAEnc & [24] & NAdam & 1.6e-3 & 4e-4 & 0 \\
+		DNAEnc & linear & NAdam & 1.6e-3 & 4e-4 & 0 \\
 		USPS & [24, 12] & Adam & 8.1e-3 & 1.5e-6 & 0 \\
-		USPS & [48, 24] & NAdam & 7.8e-3 & 1.9e-6 & 0 \\
+		USPS & linear & NAdam & 7.8e-3 & 1.9e-6 & 0 \\
 		FashionMnist & ResNet18 & NAdam & 1e-3 & 0 & 0 \\

### 2023-05-24 datasets/__init__.py

@@ -5,6 +5,7 @@ from datasets.usps import USPS
 
 # Image Datasets
 from datasets.cifar10 import Cifar10
+from datasets.mnist import Mnist
 from datasets.fashion_mnist import FashionMnist
 
 # Text Datasets

### 2023-05-24 datasets/mnist.py

@@ -1,0 +1,59 @@
+import torch
+from torch.utils.data import Dataset
+import torchvision
+from torchvision import transforms
+from core.data import BaseDataset, postprocess_torch_dataset, convert_to_channel_first, subsample_data
+
+class Mnist(BaseDataset):
+    def __init__(self, cache_folder:str, config:dict, pool_rng, encoded:bool,
+                 data_file="mnist_al.pt",):
+        # TODO: decide on a budget
+        fitting_mode = "from_scratch" if encoded else "finetuning"
+        super().__init__(cache_folder, config, pool_rng, encoded,
+                         data_file, fitting_mode)
+
+
+    def _download_data(self, test_data_fraction=0.5):
+        train = torchvision.datasets.MNIST(root=self.cache_folder, train=True, download=True)
+        test = torchvision.datasets.MNIST(root=self.cache_folder, train=False, download=True)
+        x_train, self.y_train, x_test, y_test = postprocess_torch_dataset(train, test)
+        # add an explicit color dimension for compatibility
+        x_train = torch.unsqueeze(x_train, -1)
+        x_test = torch.unsqueeze(x_test, -1)
+        x_test, self.y_test = subsample_data(x_test, y_test, test_data_fraction, self.pool_rng)
+        self.x_train, self.x_test = convert_to_channel_first(x_train, x_test)
+        # normalize pixel values from [0..255] to [0..1]
+        high = 255.0
+        self.x_train = self.x_train / high
+        self.x_test = self.x_test / high
+        self._convert_data_to_tensors()
+        print("Download successful")
+
+
+    def load_pretext_data(self)->tuple[Dataset, Dataset]:
+        train_dataset = torchvision.datasets.MNIST(root=self.cache_folder, train=True, download=True)
+        val_dataset = torchvision.datasets.MNIST(root=self.cache_folder, train=False, download=True)
+        train_dataset.targets = torch.Tensor(train_dataset.targets).int()
+        val_dataset.targets = torch.Tensor(val_dataset.targets).int()
+        return (train_dataset, val_dataset)
+
+    def get_pretext_transforms(self, config:dict)->transforms.Compose:
+        return transforms.Compose([
+                transforms.RandomResizedCrop(size=28),
+                transforms.RandomHorizontalFlip(),
+                transforms.ToTensor(),
+            ])
+
+    def get_pretext_validation_transforms(self, config:dict)->transforms.Compose:
+        return transforms.Compose([
[…]

### 2023-05-24 doc/al_benchmark_paper.tex
paper update

@@
 \author{%
-	David S.~Hippocampus\thanks{Use footnote for providing further information
-		about author (webpage, alternative address)---\emph{not} for acknowledging
-		funding agencies.} \\
-	Department of Computer Science\\
-	Cranberry-Lemon University\\
-	Pittsburgh, PA 15213 \\
-	\texttt{hippo@cs.cranberry-lemon.edu} \\
+	Thorben Werner 
+	\thanks{Information Systems and Machine Learning Lab (ISMLL)} 
+	\\
+	Institute of Computer Science\\
+	University of Hildesheim\\
+	Universitätsplatz 1, 31141 Hildesheim \\
+	\texttt{werner@ismll.de} \\
 	% examples of more authors
-	% \And
-	% Coauthor \\
-	% Affiliation \\
-	% Address \\
-	% \texttt{email} \\
-	% \AND
-	% Coauthor \\
-	% Affiliation \\
-	% Address \\
-	% \texttt{email} \\
+	\And
+	Johannes Burchert$^*$ \\
+	Institute of Computer Science\\
+	University of Hildesheim\\
+	Universitätsplatz 1, 31141 Hildesheim \\
+	\texttt{burchert@ismll.de} \\
+	\AND
+	Prof. Lars Schmidt-Thieme$^*$ \\
+	Institute of Computer Science\\
+	University of Hildesheim\\
+	Universitätsplatz 1, 31141 Hildesheim \\
+	\texttt{schmidt-thieme@ismll.uni-hildesheim.de} \\
 	% \And
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-{\color{red} Version: Braindump}\\
-Many different algorithms have been proposed for active learning. 
-In this work we focus on those approaches that have shown consistent results over the years as well as newer approaches that have demonstrated significant lifts in their initial experiments.
-AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
-Geometric approaches include CoreSet \cite{sener2017active} and TypiClust \cite{hacohen2022active}, which use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
-Uncertainty-based approaches include classic uncertainty sampling (based on Shannon-Entropy and the margin-score), BALD \cite{kirsch2019batchbald} and BADGE \cite{ashdeep}, which use metrics to measure the classifiers state.
-\\ [1mm]
-%
-Some previous work also aimed to provide a benchmark suite for active learning:
-The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
-While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call "Label Efficiency" and provide experiments on many common configurations for data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
-The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
-They test for several different properties of the models including robustness, response to compression techniques and final performance.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Overview}
-Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with an extensive experiments section.
-\\
-{\color{red} TODO} We include in this table any methodological paper that experiments on at least two domains.
+We constrain our work on pool-based active learning where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
+Specifically, we are not experimenting on so-called batch active learning, where at each iteration multiple unlabeled samples are chosen at the same time.
+Even though batch AL is the more active research domain, it does not have a principled advantage over single-sample AL except speed of computation.
+Not only is the problem of optimizing a portfolio of unlabeled samples more complicated to solve, the algorithms also have systematically less information per sample to work with.
+For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
+Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with experiments on at least two data domains.
 \begin{table}[h]
@@
 \end{minipage}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Related Work}
+{\color{red} Version: Braindump}\\
+Many different algorithms have been proposed for active learning. 
+In this work we focus on those approaches that have shown consistent results over the years as well as newer approaches that have demonstrated significant lifts in their initial experiments.
+AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
+Geometric approaches include CoreSet \cite{sener2017active} and TypiClust \cite{hacohen2022active}, which use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
+Uncertainty-based approaches include classic uncertainty sampling (based on Shannon-Entropy and the margin-score), BALD \cite{kirsch2019batchbald} and BADGE \cite{ashdeep}, which use metrics to measure the classifiers state.
+\\ [1mm]
+%
+Some previous work also aimed to provide a benchmark suite for active learning:
+The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
+While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call "Label Efficiency" and provide experiments on many common configurations for data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
+The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
+They test for several different properties of the models including robustness, response to compression techniques and final performance.
 

### 2023-05-25 agents/coreset.py
typiclust always encodes now

@@ -6,6 +6,7 @@ from torch import Tensor
 from torch.nn import Module
 from torch.optim import Optimizer
 from core.agent import BaseAgent
+from torch.utils.data import TensorDataset, DataLoader
 
 
 class Coreset_Greedy(BaseAgent):
@@ -26,9 +27,23 @@ class Coreset_Greedy(BaseAgent):
         with torch.no_grad():
             sample_size = min(sample_size, len(x_unlabeled))
             state_ids = self.agent_rng.choice(len(x_unlabeled), sample_size, replace=False)
-            candidates = classifier._encode(x_unlabeled[state_ids])
-            centers = classifier._encode(x_labeled)
+            candidates = self._embed(x_unlabeled[state_ids], classifier)
+            centers = self._embed(x_labeled, classifier)
             dist = pairwise_distances(candidates.detach().cpu(), centers.detach().cpu(), metric='euclidean')
             dist = np.min(dist, axis=1).reshape(-1, 1)
             dist = torch.from_numpy(dist)
         return state_ids[torch.argmax(dist, dim=0)]
+
+    def _embed(self, x:Tensor, model:Module)->Tensor:
+        with torch.no_grad():
+            loader = DataLoader(TensorDataset(x),
+                                batch_size=256)
+            emb_x = None
+            for batch in loader:
+                batch = batch[0]
+                emb_batch = model._encode(batch)
+                if emb_x is None:
+                    emb_dim = emb_batch.size(-1)
+                    emb_x = torch.zeros((0, emb_dim)).to(emb_batch.device)
+                emb_x = torch.cat([emb_x, emb_batch])
+        return emb_x

### 2023-05-26 doc/al_benchmark_paper.tex
typiclust always encodes now
big splice raytune

@@
 			\Require $\mathcal{L}$ \Comment{Labeled Pool}
+			\Require $\mathcal{D}_\text{val}$ \Comment{Validation Data}
 			\Require $\mathcal{D}_\text{test}$ \Comment{Test Data}
@@
 			\State $\theta_{i+1} \gets \theta_i - \eta \nabla_\theta \ell(\mathcal{L}, \hat y_{\theta})$
-			\State $\text{loss}_i \gets \ell(\mathcal{L}^\text{test}, \hat y_{\theta})$
+			\State $\text{loss}_i \gets \ell(\mathcal{D}^\text{val}, \hat y_{\theta})$
 			\If{$\text{loss}_i < \text{loss}^*$}
@@
 			\State
-			\Return Acc($\mathcal{L}^\text{test}, \hat y_{\theta}$)
+			\Return Acc($\mathcal{D}^\text{test}, \hat y_{\theta}$)
 		\end{algorithmic}
@@
 A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
-In our opinion, an evaluation on one dataset with a given seed should always be done on the same validation split as well as using the same random state for all included systems like the mini batch sampler for the classifier model training.
+Given a seed, an evaluation on one dataset should always be done on the same validation split as well as the same random state for all included systems, like the mini batch sampler for model training or the initialization for the classifier itself.
 Even though different AL algorithms will pick different samples, making them unavailable for sampling in earlier or later batches, the theoretical decision tree for every possible choice in every iteration $i$ should stay the same.
-Since every possible trajectory cannot be precomputed and stored to disk, we need to resort to seeding the evaluation.
+Since every possible trajectory cannot be precomputed and stored to disk, we need to resort to seeding.
 The default choice of setting a global seed at the start of the experiment is not sufficient here, since a single additional random draw from the random number generator completely changes the behavior of all other systems.
-This additional random number might be drawn during the initialization of the classification model, the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
-This problem also applies to the initialization of the classification model $\theta$ and the initialization and querying of $\Omega$.
-The desired control over all these processes can only be archived by assigning a separate random number generator to all these processes.
+This additional random number might be drawn during the initialization of the classification model or the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
+%This problem also applies to the initialization of the classification model $\theta$ and the initialization and querying of $\Omega$.
+The desired control only be archived by assigning a separate random number generator to all these processes.
 To the best of our knowledge, we are the first work that discusses this issue and proposes a solution for it.
@@
 We selected AL algorithms that have good performances reported by multiple different sources.
-To ensure a fair comparison we fixed the training process of our classification model as well as the set of available information for the algorithm and selected only those that can work under these restrictions:
-\paragraph{Uncertainty Sampling} 
-Tries to find the sample that the classifier is most uncertain about. For our benchmark we use entropy and margin (a.k.a. best-vs-second-best) sampling.
-\paragraph{BALD \cite{kirsch2019batchbald}}
-Applies the query-by-committee strategy of model ensembles to a single model by interpreting the model's parameters as distributions and then sample multiple outputs from them via Monte-Carlo dropout.
-\paragraph{BADGE \cite{ashdeep}} Uses gradient embeddings of unlabeled points to select samples where the classifier is expected to change a lot.
-The higher the magnitude of the gradient the higher is the expected improvement in model performance.
-\paragraph{Coreset \cite{sener2017active}}
+To ensure a fair comparison we fixed the training process of our classification model as well as the set of available information for the algorithm and selected only those that can work under these restrictions:\\
+\textbf{Uncertainty Sampling} 
+Tries to find the sample that the classifier is most uncertain about. For our benchmark we use entropy and margin (a.k.a. best-vs-second-best) sampling.\\
+\textbf{BALD \cite{kirsch2019batchbald}}
+Applies the query-by-committee strategy of model ensembles to a single model by interpreting the model's parameters as distributions and then sample multiple outputs from them via Monte-Carlo dropout.\\
+\textbf{BADGE \cite{ashdeep}} Uses gradient embeddings of unlabeled points to select samples where the classifier is expected to change a lot.
+The higher the magnitude of the gradient the higher is the expected improvement in model performance.\\
+\textbf{Coreset \cite{sener2017active}}
 Employs K-Means clustering to try to cover all {\color{red}modalities (is this the right word?)} that are observed in the data.
 Selects the unlabeled sample that is the furthest away from all cluster centers.
-In this work we use the greedy variant of Coreset and we also encode the data with the current classifier $\theta_i$ before clustering.
-\paragraph{TypiClust}
+Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\theta_i$.
+In this work we use the greedy variant of Coreset.\\
+\textbf{TypiClust \cite{hacohen2022active}}
 Relies on clustering similar to Coreset but proposes a new measure called "Typicality" to select unlabeled centers.
 Tries to select points that are in the densest regions of clusters that do not contain labeled samples yet.
-For this work we also use the current classifier $\theta_i$ to encode data if necessary (raw images or text).
+Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\theta_i$.
+It has to be pointed out that TypiClust was designed for low-budget scenarios, but we think it is still worthwhile to test and compare this algorithm with practically relevant budgets.
 %
 \subsubsection{Honorable Mentions}
-\paragraph{Learning Loss for AL}
+\textbf{Learning Loss for AL}
 Introduces an updated training of the classification model with an auxiliary loss and therefore cannot be compared fairly against classification models without this boosted training regime.

### 2023-05-26 doc/al_benchmark_paper.tex
batch Mnist BALD

@@
 The early stopping has patience 0, so it will stop as soon as the validation loss does no longer decrease.
+Even though the use of a fully labeled validation set might be regarded as impractical, since such a set will never exist during deployment, we strongly advocate for using it to control the classifier training.
+In this work we use the validation set to optimize the hyperparameters of the classifier and reduce overfitting with early stopping the training process in every iteration.
 
@@
 No learned algorithms \\
-No improved classifier training (semi-supervised, etc.)
+No SOTA classifier training (data augmentation, semi-supervised, etc.)
 
 \newpage
+
+\begin{ack}
+	Funded by the Lower Saxony Ministry of Science and Culture under grant number ZN3492 within the Lower Saxony “Vorab“ of the Volkswagen Foundation and supported by the Center for Digital Innovations (ZDIN).
+\end{ack}
 

### 2023-05-26 core/environment.py

@@ -87,9 +87,10 @@ class ALGame(gym.Env):
             with torch.no_grad():
                 self._add_point_to_labeled_pool(a)
 
-            # fit classification model
-            reward = self.fit_classifier()
-
+            if self.fitting_mode == "finetuning":
+                reward = self.fit_classifier()
+        if self.fitting_mode == "from_scratch":
+                reward = self.fit_classifier()
         next_state = self.create_state()
         done = self.added_images >= self.budget
         truncated = False

### 2023-05-30 doc/al_benchmark_paper.tex
big DNA experiments

@@
 \usepackage{algorithm}
-\usepackage{algpseudocode}
+\usepackage[noend]{algpseudocode}
 \usepackage{neurips_data_2023}
@@
 	Thorben Werner 
-	\thanks{Information Systems and Machine Learning Lab (ISMLL)} 
+	\thanks{Institute of Computer Science - Information Systems and Machine Learning Lab (ISMLL)} 
 	\\
-	Institute of Computer Science\\
 	University of Hildesheim\\
-	Universitätsplatz 1, 31141 Hildesheim \\
+	Universitätsplatz 1\, 31141 Hildesheim \\
 	\texttt{werner@ismll.de} \\
@@
 	Johannes Burchert$^*$ \\
-	Institute of Computer Science\\
 	University of Hildesheim\\
@@
 	Prof. Lars Schmidt-Thieme$^*$ \\
-	Institute of Computer Science\\
 	University of Hildesheim\\
@@
 	We propose the first AL benchmark that tests algorithms in all three major domains of tabular, image and text data.
-	Furthermore, we discuss a central problem for reproducing AL experiments with the default seeding setup that depends on a single seed per experiment.
-	...
+	Furthermore, we discuss overlooked problems for reproducing AL experiments with the default seeding setup that depends on a single seed per experiment and provide evidence for the correct amount of repetitions for AL algorithms that reliably converge to the true median performance.
+	We report empirical results for 6 algorithms on 7 datasets and aggregate them into a ranking of AL algorithms via Critical-Difference Diagrams.
 \end{abstract}
@@
 The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art results by significantly outperforming everyone else, while the vast majority of results proves to be non-reproducible.
-\\ [1mm]
-%
-To allow comparison between
 
@@
 
-\begin{minipage}{0.49\linewidth}
+\begin{minipage}{0.4\linewidth}
 	\begin{algorithm}[H]
@@
 		\begin{algorithmic}[1]
-			\Require $\mathcal{U}$ \Comment{Unlabeled Pool}
-			\Require $\tau$ \Comment{Unlabeled Sample Size}
-			\State $\mathcal{L}^{(1)} \gets \operatorname{seed}(\mathcal{U})$  %\Comment{Create the initial labeled set}
-			\State $\mathcal{U}^{(1)} \gets \mathcal{U} \setminus \mathcal{L}^{(1)}$
-			\For{$i := 1 \ldots B$} 
-			\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$  %\Comment{$\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
+			\Ensure Retrain, margin
+			\Require $\mathcal{U}^{(i)}, \mathcal{L}^{(i)}, \tau, \hat y_\theta$ 
+			\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$ 
 			\State $u^{(i)} \underset{\tau}{\sim} \text{unif}(1:|\mathcal{U}^{(i)}|)$
-			\State $r* \gets 0$
-			\For{$j := 1 \ldots \tau$} %\Comment{Testing every unlabeled point}
-			\State $y^{(j)} \gets \operatorname{label}(u^{(i)}_{j})$
-			\State $\mathcal{L}^{(j)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_j, y^{(j)})\}$
-			\State $\text{acc}^{(j)} \gets \operatorname{Retrain}(\mathcal{L}^{(j)})$  
-			\State $r^{(j)} \gets \text{acc}^{(j)} - \text{acc}^{(i)}$
-			\If{$r^{(j)} > r^*$} %\Comment{Select point with largest increase in performance}
-			\State $r* \gets r^{(j)}$
-			\State $j* \gets j$
+			\State $r^* \gets 0$
+			\For{$t := 1 \ldots \tau$}
+			\State $\mathcal{K} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_t, y_t)\}$
+			\State $\text{acc} \gets \operatorname{Retrain}(\mathcal{K}, \hat y_\theta)$  
+			\State $r \gets \text{acc} - \text{acc}^{(i)}$
+			\If{$r > r^*$} 
+			\State $r^* \gets r$
+			\State $t^* \gets t$
+			\EndIf  
+			\If{$r^* = 0$}
+			\State $t^* \gets \operatorname{margin}(u^{(i)}, \hat y_\theta)$
 			\EndIf
 			\EndFor
-			\If{$r^* = 0$}
-			\State $j* \gets \operatorname{margin}(u^{(i)}, \hat y_\theta)$
-			\EndIf
-			\State $y^{(i)} \gets \operatorname{label}(u^{(i)}_{j*})$
-			\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_{j*}, y^{(i)})\}$
-			\State $\mathcal{U}^{(i+1)} \gets \mathcal{U}^{(i)} \setminus \{u^{(i)}_{j*}\}$
-			\EndFor
-			\State
-			\Return $\frac{1}{B} \sum_{i=1}^{B} \text{acc}^{(i)}$
+			\Return $t^*$
 		\end{algorithmic}
@@
 \hspace{3mm}
-\begin{minipage}{0.5\linewidth}
-	\begin{algorithm}[H]
-		\caption{Retrain}\label{alg:retrain}
-		\begin{algorithmic}[1]
-			\Require $\mathcal{L}$ \Comment{Labeled Pool}
-			\Require $\mathcal{D}_\text{val}$ \Comment{Validation Data}
-			\Require $\mathcal{D}_\text{test}$ \Comment{Test Data}
-			\Require $\hat y_\theta$ \Comment{Class. Model}
-			\Require $e^\text{max}$ \Comment{Maximum Epochs}
-			\State $\text{loss}^* \gets \infty$
-			\For{$i := 1 \ldots e^{\text{max}}$}
-			\State $\theta_{i+1} \gets \theta_i - \eta \nabla_\theta \ell(\mathcal{L}, \hat y_{\theta})$
-			\State $\text{loss}_i \gets \ell(\mathcal{D}^\text{val}, \hat y_{\theta})$
-			\If{$\text{loss}_i < \text{loss}^*$}
-			\State $\text{loss}^* \gets \text{loss}_i$
-			\Else
-			\State Break
-			\EndIf
-			\EndFor
-			\State
-			\Return Acc($\mathcal{D}^\text{test}, \hat y_{\theta}$)
-		\end{algorithmic}
-	\end{algorithm}
+\begin{minipage}{0.59\linewidth}
 	In Alg.
-\ref{alg:oracle} $\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{D}_\text{test}, \hat y_\theta, e^\text{max})$, $\operatorname{label}(u^{(i)}_{j})$ recovers the label of instance $u^{(i)}_{j}$ and $\operatorname{margin}(u^{(i)}, \hat y_\theta)$ applies margin sampling to select as a fallback, when the oracle has not found another option that induces a positive change in classification performance.
+\ref{alg:oracle} $\operatorname{Retrain}(\mathcal{L}^{(i)}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns the accuracy on the test set $\mathcal{D}_\text{test}$.
+\\
+	$y_t$ is shorthand for the corresponding label of $u^{(i)}_{t}$ that can be recovered from the dataset labels. \\
+	When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$).
 \end{minipage}
@@
 In the remaining cases, we define test sets upfront and store them into separate files to keep them fixed across all experiments.
-The validation set is split during experiment-time and depends on the dataset-seed.
-
-\paragraph{Tabular}
+The validation set is split during experiment-time and depends on the dataset-seed.\\
+\textbf{Tabular:}
 We use \textbf{Splice}, \textbf{DNA} and \textbf{USPS} from LibSVMTools \cite{libsvmtools}.
-All three datasets are normalized between [0, 1].
-
-\paragraph{Image}
+All three datasets are normalized between [0, 1]. \\
+\textbf{Image:}
 We use \textbf{FashionMNIST} \cite{xiao2017fashion} and \textbf{Cifar10} \cite{krizhevsky2009learning}.
-Both datasets are normalized according to their standard protocols.
-
-\paragraph{Text}
+Both datasets are normalized according to their standard protocols. \\
+\textbf{Text:}
 We use \textbf{News Category} \cite{misra2022news} and \textbf{TopV2} \cite{chen-etal-2020-low-resource}.
@@
 Both datasets are encoded with pre-trained GloVe embeddings \cite{pennington2014glove}.
-Since neither dataset provided a fixed test set, we randomly split 7000 datapoints into a test set. \\ [3mm]
+Since neither dataset provided a fixed test set, we randomly split 7000 datapoints into a test set. \\ [1mm]
 %
@@
 This might be true if we aimed to develop novel classification models on these datasets, however we are solely focused on comparing different AL algorithms in this paper.
-Our core assumption is that a well-performing algorithm in our benchmark will transfer well into more practical use-cases. \\ [1mm]
+Our core assumption is that a well-performing algorithm in our benchmark will also transfer into more practical use-cases. \\ [1mm]
 Adapting the experimental setting from \cite{hacohen2022active} we offer all our datasets in the raw setting as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning.
 The text datasets are an exception, as they are only offered in their encoded form.
-The pre-encoded datasets offer multiple advantages in that they require smaller classification models (i.e.
-small MLPs) and in general much less sampled datapoints to archive close to upper bound performance (measured with the full dataset).
-They unify experimental setups for different datasets and therefore make them much more comparable.
+The pre-encoded datasets enable us to test our single-sample algorithms on more complex datasets like Cifar10 and FashionMnist without the need of sampling $>2000$ datapoints before we can reach our upper bound performance.
+The embedding model was trained with the SimCLR \cite{chen2020simple} algorithm. 
+For Cifar10 we adapt the reported hyperparameters from \cite{hacohen2022active} and for the tabular datasets we use random search to optimize the hyperparameters.
+The quality of embeddings during pretext training was measured after each epoch by attaching a linear classification head and evaluating this classifier for test accuracy, mirroring our AL setup for embedded datasets.
 
@@
 In order to provide a meaningful analysis of which algorithm can be expected to perform best on average we ranked the algorithm for each dataset based on their median AUC and displayed these rankings in critical difference diagrams \cite{IsmailFawaz2018deep}.
-In Fig. \ref{fig:cd_diagrams} we report the rankings split by domain as well as across all domains (excluding the toydata).
+In Fig. \ref{fig:cd_diagrams} we report the rankings split by domain as well as across all domains (excluding the toydata). \\ [2mm]
+%
+BALD performs bad with linear classifiers since they are trained without dropout and cannot cope well with missing inputs. \\
+TypiClust is better with embedded data not only due to lower budgets. On other datasets it is not able to outperform other algorithms in early stages
 
@@
 No learned algorithms \\
-No SOTA classifier training (data augmentation, semi-supervised, etc.)
+No SOTA classifier training (data augmentation, semi-supervised, etc.) \\
+SimCRL as Pretext task works better for images
 
@@
 
+\begin{algorithm}[H]
+	\caption{Retrain}\label{alg:retrain}
+	\begin{algorithmic}[1]
+		\Require $\mathcal{L}$ \Comment{Labeled Pool}
+		\Require $\mathcal{D}_\text{val}$ \Comment{Validation Data}
+		\Require $\mathcal{D}_\text{test}$ \Comment{Test Data}
+		\Require $\hat y_\theta$ \Comment{Class. Model}
+		\Require $e^\text{max}$ \Comment{Maximum Epochs}
+		\State $\text{loss}^* \gets \infty$
+		\For{$i := 1 \ldots e^{\text{max}}$}
+		\State $\theta_{i+1} \gets \theta_i - \eta \nabla_\theta \ell(\mathcal{L}, \hat y_{\theta})$
+		\State $\text{loss}_i \gets \ell(\mathcal{D}^\text{val}, \hat y_{\theta})$
+		\If{$\text{loss}_i < \text{loss}^*$}
+		\State $\text{loss}^* \gets \text{loss}_i$
+		\Else
+		\State Break
+		\EndIf
+		\EndFor
+		\State
+		\Return Acc($\mathcal{D}^\text{test}, \hat y_{\theta}$)
+	\end{algorithmic}
+\end{algorithm}
+
 \end{document}

### 2023-05-30 evaluate.py

@@ -15,8 +15,8 @@ parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--agent_seed", type=int, default=1)
 parser.add_argument("--pool_seed", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
-parser.add_argument("--agent", type=str, default="batchrandom")
-parser.add_argument("--dataset", type=str, default="mnist")
+parser.add_argument("--agent", type=str, default="bald")
+parser.add_argument("--dataset", type=str, default="fashionmnist")
 parser.add_argument("--encoded", type=int, default=0)
 # parser.add_argument("--sample_size", type=int, default=20)
 parser.add_argument("--restarts", type=int, default=10)

### 2023-05-31 doc/al_benchmark_paper.tex
divergingsin exp

@@
 \usepackage{graphicx}
+\usepackage{wrapfig}
 \usepackage{float}
@@
 \begin{abstract}
-	Active Learning (AL) is a technique used in machine learning to improve data efficiency by selecting the most informative samples for annotation. 
+	%Active Learning (AL) is a technique used in machine learning to improve data efficiency by selecting the most informative samples for annotation. 
 	In this paper we address the issue of inconsistent results in active learning literature.
-	Authors of previous papers are constantly reporting significant performance improvements in their original work, while subsequent literature fails to reproduce those results.
+	Authors of previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results.
 This inconsistency leads to a chaotic landscape of AL algorithms.
-	We propose the first AL benchmark that tests algorithms in all three major domains of tabular, image and text data.
+	We propose the first AL benchmark that tests algorithms in three major domains of tabular, image and text data.
 	Furthermore, we discuss overlooked problems for reproducing AL experiments with the default seeding setup that depends on a single seed per experiment and provide evidence for the correct amount of repetitions for AL algorithms that reliably converge to the true median performance.
@@
 \section{Introduction}
-{\color{red} Version: 1.0}\\
 Active Learning (AL) plays an important role in our society that applies machine learning to more and more areas and therefore has a high demand for labeled data in more and more areas.
@@
 For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
+A performance comparison of batch AL and single-sample AL can be found in Fig.
+\ref{fig:ablation_bald}, which reproduces the message of Figure 1 from the paper \cite{gal2017deep} that proposed BALD, one of the SOTA algorithms for AL.
+We can see that the batched version of BALD \cite{krizhevsky2009learning} can at most perform on-par with the single-sample algorithm.
+Fig. \ref{fig:ablation_bald} also serves as a proof of concept for our provided code base.
 Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with experiments on at least two data domains.
@@
 \subsection{Problem Description}
-{\color{red} Version: 1.0}\\
+Given two spaces $\mathcal{X}:=\mathbb{R}^M$ and $\mathcal{Y}:=\mathbb{R}^C$ where a samples is drawn from an unknown data distribution $(x, y) \sim p$ with $x \in \mathcal{X}, y \in \mathcal{Y}$.
+We call $\mathcal{L} := \{ (x, y) \}^*$ the labeled pool and $\mathcal{U} := \{ x \}^*$ the unlabeled pool.
+Their domains are $\Omega := \mathbb{R}^M \times \mathbb{R}^C$ and $\Lambda := \mathbb{R}^M$ respectively. \\
+A function $\hat y: \mathcal{X} \rightarrow \mathcal{Y}$ is called a classifier and a function $\ell: \mathcal{Y} \rightarrow \mathcal{Y}$ is called a loss.
+Given a scalar $B := \mathbb{R}; \hspace{1mm} B < |\mathcal{U}|$ called budget, we want to find an acquisition function $a: \Omega \times \Lambda \mapsto \mathcal{U}^B$ that creates a subset of $\mathcal{U}$ so that the expected loss on our data is minimal.
+$$
+\underset{(x,y)\sim p}{\mathbb{E}} \ell(y, \hat y(x))
+$$
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Lars' Problem Description}
 Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$.
@@
 	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
-	\quad\quad \text{\tiny (equivariant in the second argument)}
 \end{align*}
-% which is equivariant in the second argument, 
-called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal:
+called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal.
 \begin{align*}
@@
 \end{align*}
-(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
-Since combinatorial problem of finding the optimal subset $\mathcal{D}_\text{train}$ is computational not feasible, we allow sequential construction of $\mathcal{D}_\text{train}$. \\
+%(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
+Even though the acquisition function in principle could output the full subset of $\mathcal{D}_\text{train}$, the combinatorial problem is computationally not feasible and we allow sequential construction of the subset as a relaxation of the problem.
+\\
 %
@@
 \begin{equation}\label{eq:auc}
-	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(y_{test}, \hat y_i(x_{test}))
+	\operatorname{auc}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
 \end{equation}
 Where $\hat y_i$ is the retrained classification model after the i-\textit{th} instance was selected.
-
-\begin{minipage}{0.4\linewidth}
-	\begin{algorithm}[H]
-		\caption{Oracle}\label{alg:oracle}
-		\begin{algorithmic}[1]
-			\Ensure Retrain, margin
-			\Require $\mathcal{U}^{(i)}, \mathcal{L}^{(i)}, \tau, \hat y_\theta$ 
-			\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$ 
-			\State $u^{(i)} \underset{\tau}{\sim} \text{unif}(1:|\mathcal{U}^{(i)}|)$
-			\State $r^* \gets 0$
-			\For{$t := 1 \ldots \tau$}
-			\State $\mathcal{K} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_t, y_t)\}$
-			\State $\text{acc} \gets \operatorname{Retrain}(\mathcal{K}, \hat y_\theta)$  
-			\State $r \gets \text{acc} - \text{acc}^{(i)}$
-			\If{$r > r^*$} 
-			\State $r^* \gets r$
-			\State $t^* \gets t$
-			\EndIf  
-			\If{$r^* = 0$}
-			\State $t^* \gets \operatorname{margin}(u^{(i)}, \hat y_\theta)$
-			\EndIf
-			\EndFor
-			\Return $t^*$
-		\end{algorithmic}
-	\end{algorithm}
-\end{minipage}
-\hspace{3mm}
-\begin{minipage}{0.59\linewidth}
-	In Alg.
-\ref{alg:oracle} $\operatorname{Retrain}(\mathcal{L}^{(i)}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns the accuracy on the test set $\mathcal{D}_\text{test}$.
-\\
-	$y_t$ is shorthand for the corresponding label of $u^{(i)}_{t}$ that can be recovered from the dataset labels. \\
-	When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$).
-\end{minipage}
 
@@
 Many different algorithms have been proposed for active learning. 
-In this work we focus on those approaches that have shown consistent results over the years as well as newer approaches that have demonstrated significant lifts in their initial experiments.
+In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
 AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
-Geometric approaches include CoreSet \cite{sener2017active} and TypiClust \cite{hacohen2022active}, which use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
-Uncertainty-based approaches include classic uncertainty sampling (based on Shannon-Entropy and the margin-score), BALD \cite{kirsch2019batchbald} and BADGE \cite{ashdeep}, which use metrics to measure the classifiers state.
-\\ [1mm]
+Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
+They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
+This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
+Uncertainty-based approaches use metrics to measure the classifiers state.
+Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
+This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
 %
@@
 The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
-While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call "Label Efficiency" and provide experiments on many common configurations for data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
+While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call "Label Efficiency" and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
 The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
@@
 \subsection{Evaluation}\label{sec:evaluation}
-{\color{red} Version: 1.0}\\
 We compare different AL algorithms based on their median AUC score (Eq. \ref{eq:auc}) across multiple restarts of the experiment.
@@
 \subsection{Reproducibility}\label{sec:reproducibility}
-{\color{red} Version: 1.0}\\
 A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
@@
 Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
+% This seeding setup enables a reproduction of the oracle curve even when the classifier or AL algorithm is different.
 
@@
 \subsection{Oracle}\label{sec:oracle}
-{\color{red} Version: 1.0}\\
 Posing active learning as a sequence ordering problem, the oracle sequence for a given combination of dataset, model and training procedure would be the sequence that induces the highest AUC score for a given budget.
@@
 Even though their reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
-\\
+\begin{minipage}{0.47\textwidth}
+	\begin{algorithm}[H]
+		\caption{Oracle}\label{alg:oracle}
+		\begin{algorithmic}[1]
+			\Require $\mathcal{U}, \mathcal{L}, \mathcal{Y}, \mathcal{D}_\text{test}$ Train, Margin, $\tau, \hat y_\theta$ 
+			\State $\text{acc} \gets \operatorname{Train}(\mathcal{L}, \mathcal{D}_\text{test}, \hat y_\theta)$ 
+			%\State $u \underset{\tau}{\sim} \text{unif}(1:|\mathcal{U}|)$
+			\State $r^* \gets 0$
+			\For{$t := 1 \ldots \tau$}
+			\State $\mathcal{L}' \gets \mathcal{L}^{(i)} \cup \{(\mathcal{U}_t, y_t)\}$
+			\State $\text{acc}' \gets \operatorname{Train}(\mathcal{L}', \mathcal{D}_\text{test}, \hat y_\theta)$  
+			\State $r \gets \text{acc} - \text{acc}'$
+			\If{$r > r^*$} 
+			\State $r^* \gets r$
+			\State $u^* \gets \mathcal{U}_t$
+			\EndIf
+			\EndFor
+			\If{$r^* = 0$}
+			\State $u^* \gets \operatorname{margin}(\mathcal{U}, \hat y_\theta)$
+			\EndIf
+			\Return $u^*$
+		\end{algorithmic}
+	\end{algorithm}
+\end{minipage}
+\hspace{2mm}
+\begin{minipage}{0.5\linewidth}
+	\begin{figure}[H]
+		\includegraphics[width=\linewidth]{img/ablation_bald}
+		\caption{Shows a reproduction of the performance analysis of BALD from the original authors \cite{gal2017deep} in comparison to entropy sampling and an adaptation of BALD for batch AL \cite{kirsch2019batchbald} with a batch size of 40.}
+		\label{fig:ablation_bald}
+	\end{figure}
+\end{minipage} \\[2mm]
 In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
@@
 The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.
+In the algorithm $\operatorname{Retrain}(\mathcal{L}^{(i)}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns the accuracy on the test set $\mathcal{D}_\text{test}$. \\ 
+$y_t$ is shorthand for the corresponding label of $u^{(i)}_{t}$ that can be recovered from the dataset labels. \\
+When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$). \\
+Alg. \ref{alg:oracle} replaces the acquisition function in the AL process.
 
@@
 \subsection{Training the Classifier}\label{sec:training_the_classifier}
-{\color{red} Version: 0.1}\\
 Generally, the classification model can be trained in two ways.
@@
 \subsection{Datasets}\label{sec:datasets}
-{\color{red} Version: 1.0}\\
 For all our datasets we use the pre-defined train / test splits, if given. 

### 2023-05-31 compute_upper_bound.py

@@ -8,7 +8,7 @@ parser = argparse.ArgumentParser()
 parser.add_argument("--data_folder", type=str, required=True)
 parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
-parser.add_argument("--dataset", type=str, default="splice")
+parser.add_argument("--dataset", type=str, default="divergingsin")
 parser.add_argument("--encoded", type=int, default=0)
 parser.add_argument("--restarts", type=int, default=15)
 parser.add_argument("--patience", type=int, default=20)

### 2023-05-31 doc/al_benchmark_paper.tex
reverted to small classifiers for splice and dna

@@
 
+\section{Comparison of Different Classifier Sizes}
+We tested two different classifier sizes in Splice and DNA:
+\begin{itemize}
+	\item Small: [24, 12] (2400 parameters)
+	\item Big: [24, 48, 48] (5700 parameters)
+\end{itemize}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.49\linewidth]{img/supp_dna_small.png}
+	\includegraphics[width=0.49\linewidth]{img/supp_dna_big.png}
+	\includegraphics[width=0.49\linewidth]{img/supp_micro_dna_small.png}
+	\includegraphics[width=0.49\linewidth]{img/supp_micro_dna_big.png}
+	\caption{Comparison of small and big classifiers for the DNA dataset}
+\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.49\linewidth]{img/supp_splice_small.png}
+	\includegraphics[width=0.49\linewidth]{img/supp_splice_big.png}
+	\includegraphics[width=0.49\linewidth]{img/supp_micro_splice_small.png}
+	\includegraphics[width=0.49\linewidth]{img/supp_micro_splice_big.png}
+	\caption{Comparison of small and big classifiers for the Splice dataset}
+\end{figure}
+
 \section{Comparison of different sample sizes}

### 2023-05-31 doc/al_benchmark_paper.tex
paper update

@@
 Traditionally, the classifier is chosen per dataset so that it is capable of solving the dataset close to the SOTA performance reported in the literature.
-Similar to our hypothesis in section \ref{sec:evaluation} we hypothesize that AL algorithms will perform similarly on small classifiers and more complex ones, so that the overall ranking of algorithms stays the same.
-{\color{red} TODO: Transform this into full hypothesis incl. formatting?}\\
-On the basis of this hypothesis we opt to use smaller classifiers that still solve the dataset to a reasonable degree.
-Smaller classifiers also require fewer labeled datapoints to acquire performance close to the upper bound.
+Since we are not interested in archiving a new SOTA in any classification problem, we opt to use smaller classifiers for the following reasons:
+Smaller classifiers generally (i) exhibit more stable training behavior and (ii) on average require less sampled datapoints to reach the their upper bound performance on the full dataset.
 For every dataset the chosen architecture's hyperparameters are optimized by to archive maximum upper bound performance.
+One desired characteristic of small classifiers is that the ranking of AL algorithms should stay the same when switching to larger models.
+A small analysis of this behavior can be found in Appendix \ref{app:classifier_size}.
+We found that the ranking of AL algorithms unfortunately does change, but we did not observe systematics that benefit one or few specific algorithms.
+We therefore rely on the different data domains to provide classification models of different sizes and archetypes to cover all of the use-cases.
 For an overview of architectures and hyperparameters please refer to Appendix \ref{app:hyperparameters}.
@@
 \subsection{Training the Classifier}\label{sec:training_the_classifier}
-Generally, the classification model can be trained in two ways.
+The classification model can be trained in two ways.
 Either you reset the parameters after each AL iteration and train the classifier from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or you retain the previous state and only fine-tune the classifier on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
@@
 The early stopping has patience 0, so it will stop as soon as the validation loss does no longer decrease.
-Even though the use of a fully labeled validation set might be regarded as impractical, since such a set will never exist during deployment, we strongly advocate for using it to control the classifier training.
+Even though the use of a fully labeled validation set might be regarded as impractical, since such a set will never exist during deployment, we strongly advocate for using it in benchmarks to control the classifier training.
 In this work we use the validation set to optimize the hyperparameters of the classifier and reduce overfitting with early stopping the training process in every iteration.
@@
 For News Category we use  the 15 most common categories as indicated by its Kaggle site.
-We additionally drop sentences above 80 words to reduce the needed padding (retaining 99,86\% of the data).
+We additionally drop sentences above 80 words to reduce the padding needed (retaining 99,86\% of the data).
 For TopV2, we are only using the "alarm" domain.
-Both datasets are encoded with pre-trained GloVe embeddings \cite{pennington2014glove}.
+Both datasets are encoded with pre-trained GloVe (Common Crawl 840B Tokens) embeddings \cite{pennington2014glove}.
 Since neither dataset provided a fixed test set, we randomly split 7000 datapoints into a test set. \\ [1mm]
@@
 We would like to point out that these datasets can be considered "toy-datasets" and therefore not relevant for practical purposes.
-This might be true if we aimed to develop novel classification models on these datasets, however we are solely focused on comparing different AL algorithms in this paper.
+This might be true if we aimed to develop novel classification models on these datasets, however, similar to our argumentation for picking smaller classifiers, we are solely focused on comparing different AL algorithms in this paper.
 Our core assumption is that a well-performing algorithm in our benchmark will also transfer into more practical use-cases. \\ [1mm]
@@
 	\centering
+	\includegraphics[width=0.49\linewidth]{img/eval_splice}
+	\hspace{1mm}
 	\includegraphics[width=0.49\linewidth]{img/eval_dna}
-	\includegraphics[width=0.49\linewidth]{img/eval_splice.png}
 	\caption{Results for all algorithms on Splice and DNA, both from the tabular domain. Even within one domain, the performance of the same algorithm can vary drastically.}
@@
 
-\section{Comparison of Different Classifier Sizes}
+\section{Comparison of Different Classifier Sizes}\label{app:classifier_size}
 We tested two different classifier sizes in Splice and DNA:

### 2023-05-31 doc/al_benchmark_paper.tex
paper update

@@
 	%Active Learning (AL) is a technique used in machine learning to improve data efficiency by selecting the most informative samples for annotation. 
-	In this paper we address the issue of inconsistent results in active learning literature.
-	Authors of previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results.
+	In this paper we address the issue of inconsistent results in active learning (AL) literature.
+	Previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results.
 This inconsistency leads to a chaotic landscape of AL algorithms.
-	We propose the first AL benchmark that tests algorithms in three major domains of tabular, image and text data.
-	Furthermore, we discuss overlooked problems for reproducing AL experiments with the default seeding setup that depends on a single seed per experiment and provide evidence for the correct amount of repetitions for AL algorithms that reliably converge to the true median performance.
-	We report empirical results for 6 algorithms on 7 datasets and aggregate them into a ranking of AL algorithms via Critical-Difference Diagrams.
+	To the best of our knowledge, we propose the first AL benchmark that tests algorithms in 3 major domains; Tabular, Image and Text.
+	Furthermore, we discuss overlooked problems for reproducing AL experiments that can lead to unfair comparisons and increased variance in the results.
+	We propose a framework for choosing hyperparameters of AL setups and accurately controlling the experiments.
+	Using this framework, we report empirical results for 6 algorithms on 7 datasets and aggregate them into a domain-specific ranking of AL algorithms.
 \end{abstract}
@@
 \begin{enumerate}
-	\item Evaluation of Active Learning algorithms on datasets from 4 different domains, including synthetic data that highlights principled shortcomings of existing approaches.
+	\item Evaluation of Active Learning algorithms on datasets from 3 different domains, including synthetic data that highlights principled shortcomings of existing approaches.
 	\item Novel experimental protocol for seeding the experiment with 3 different seeds to allow full control and reproducibility and analysis of how many restarts are required to converge to the true median performance reliably.
@@
 \end{enumerate}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{(From Vijaya) Problem Description}
+Given $n = l+u$ data points with $l \in \mathbb{N}$ many labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ many unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathbb{R}^M \to \mathbb{R}^C$ that can label $x$.
+We call $x \in \mathbb{R}^M$, $y \in \mathbb{R}^C$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$.
+Find an acquisition function $F: \mathcal{U},\mathcal{L}, b \mapsto \mathcal{U}_b$ that selects $b$ many examples from $\mathcal{U}$ such that $\mathcal{U}_b \subseteq \mathcal{U}$, $|\mathcal{U}_b|=b$ for labeling and the expected loss $\ell: \mathbb{R}^C \times \mathbb{R}^C \to \mathbb{R}$ (cross entropy loss) of the machine learning algorithm $\hat{y}$ (ex.
+classification) trained on $\mathcal{L} \cup \{\left(x, A(x)\right) \mid x \in \mathcal{U}_b\}$ is minimal:
+$$\min \quad \mathbb{E}_{(x,y) \sim \rho} \ell(y, \hat{y}(x))$$
+Finding the optimal subset $\mathcal{U}_b$ is a combinatorial problem that is computationally not feasible for large $b$, so we allow sequential construction of the subset as a relaxation of the problem.
+%To construct the active learning setting from a labeled dataset $\mathcal{D} = \{(x_1, y_1),\ldots, (x_n,y_n)\}$, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U}$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class, removing them from $\mathcal{U}$ and recovering their label $\mathcal{L} := \{(x_1, A(x_1)),\ldots, (x_{k*C},A(x_{k*C}))\} \quad s.t.
+\quad \mathcal{U} \cap \mathcal{L} = \emptyset$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{(From Lars) Problem Description}
+Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$.
+\\
+Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
+\begin{align*}
+	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
+\end{align*}
+called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal.
+%\begin{align*}
+%	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
+%	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
+%	% \operatorname{avg} \{
+%	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
+%	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
+%	\ell(y, \hat y(x)) 
+%	\\
+%	\text{with }
+%	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
+%	\\ 
+%	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
+%	\quad b\in 1{:}B
+%\end{align*}
+%(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
+Even though the acquisition function in principle could output the full subset of $\mathcal{D}_\text{train}$, the combinatorial problem is computationally not feasible for large $B$ and we allow sequential construction of the subset as a relaxation of the problem.
+\\
+%
+%To construct the active learning setting from existing labeled datasets, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$.
 
@@
 		Zhou et al. \cite{zhou2021towards} & batch & 3 & 2 & 2 & \checkmark \\
-		\textbf{Ours} & single & 9 & 4 & 6 & \checkmark 
+		\textbf{Ours} & single & 7 & 3 & 6 & \checkmark 
 	\end{tabular}
@@
 %\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Problem Description}
-Given two spaces $\mathcal{X}:=\mathbb{R}^M$ and $\mathcal{Y}:=\mathbb{R}^C$ where a samples is drawn from an unknown data distribution $(x, y) \sim p$ with $x \in \mathcal{X}, y \in \mathcal{Y}$.
-We call $\mathcal{L} := \{ (x, y) \}^*$ the labeled pool and $\mathcal{U} := \{ x \}^*$ the unlabeled pool.
-Their domains are $\Omega := \mathbb{R}^M \times \mathbb{R}^C$ and $\Lambda := \mathbb{R}^M$ respectively. \\
-A function $\hat y: \mathcal{X} \rightarrow \mathcal{Y}$ is called a classifier and a function $\ell: \mathcal{Y} \rightarrow \mathcal{Y}$ is called a loss.
-Given a scalar $B := \mathbb{R}; \hspace{1mm} B < |\mathcal{U}|$ called budget, we want to find an acquisition function $a: \Omega \times \Lambda \mapsto \mathcal{U}^B$ that creates a subset of $\mathcal{U}$ so that the expected loss on our data is minimal.
-$$
-\underset{(x,y)\sim p}{\mathbb{E}} \ell(y, \hat y(x))
-$$
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Lars' Problem Description}
-Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$.
-\\
-Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
-\begin{align*}
-	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
-\end{align*}
-called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal.
-\begin{align*}
-	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
-	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
-	% \operatorname{avg} \{
-	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
-	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
-	\ell(y, \hat y(x)) 
-	\\
-	\text{with }
-	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
-	\\ 
-	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
-	\quad b\in 1{:}B
-\end{align*}
-%(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
-Even though the acquisition function in principle could output the full subset of $\mathcal{D}_\text{train}$, the combinatorial problem is computationally not feasible and we allow sequential construction of the subset as a relaxation of the problem.
-\\
-%
-To construct the active learning setting from existing labeled datasets, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$.
-\\
-%
-Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
-We employ the normalized area under the accuracy curve (AUC):
-\begin{equation}\label{eq:auc}
-	\operatorname{auc}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
-\end{equation}
-Where $\hat y_i$ is the retrained classification model after the i-\textit{th} instance was selected.
 
@@
 The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
-While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call "Label Efficiency" and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
+While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
 The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
@@
 \subsection{Evaluation}\label{sec:evaluation}
-We compare different AL algorithms based on their median AUC score (Eq. \ref{eq:auc}) across multiple restarts of the experiment.
-Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split to mimic the leave-one-out protocol for cross-validation.
+Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
+We employ the normalized area under the accuracy curve (AUC):
+\begin{equation}\label{eq:auc}
+	\operatorname{AUC}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
+\end{equation}
+where $\hat y_i$ is the (re-)trained classification model after the i-\textit{th} iteration.
+To mimic the leave-one-out protocol for cross-validation we will restart each experiment multiple times.
+Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split.
 The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
@@
 Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
-To create Fig. \ref{fig:restarts} we used all our 50 runs from the margin/random sampling algorithm on the splice dataset.
+To create Fig. \ref{fig:restarts} we used 50 restarts from the margin/random sampling algorithm on the splice dataset.
 From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
-One AUC value like this corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
+One of these median AUC values corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
 To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
 Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
-We can clearly observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
+We can observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
 \begin{figure}
@@
 \end{figure}
-To combat is uncertain evaluation, we propose to repeat every experiment 50 times, to arrive at the true median AUC for each algorithm.
+To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
+One way to reduce the number of necessary repetitions is to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
+We ultimately decided in favor of high variance and high number of repetitions as the high variance accurately reflects real world applications of AL.
 
@@
 A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
+For a fair comparison of two AL algorithms, both algorithms should receive equal starting conditions in terms of train / validation split, initialization of classifier and even the state of minor systems like the optimizer of mini-batch sampler.
+Even though some implementations might have their own solution to some of these problems, to the best of our knowledge
+
+%Decided against a fixed time budget
 Given a seed, an evaluation on one dataset should always be done on the same validation split as well as the same random state for all included systems, like the mini batch sampler for model training or the initialization for the classifier itself.
@@
 \subsection{Oracle}\label{sec:oracle}
-Posing active learning as a sequence ordering problem, the oracle sequence for a given combination of dataset, model and training procedure would be the sequence that induces the highest AUC score for a given budget.
-However, since this combinatorial problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence.
+Posing active learning as a combinatorial problem, the oracle set $\mathcal{U}_b$ for a given dataset, model and training procedure would be the set that induces the highest AUC score for a given budget.
+However, since this problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence.
 \cite{zhou2021towards} has used simulated annealing to search for the optimal sequence and used the best  solution found after a fixed time budget. 
@@
 In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
-Our oracle simply tests every data point in the provided sample of unlabeled points by fitting the classifier and directly measuring the resulting validation performance.
-The point with the best validation performance is selected and added to the labeled pool for that iteration.
-We noticed that this oracle is overfitting on the validation set, resulting in stagnating or even decreasing performance curves in later AL iterations.
+Our oracle simply evaluated every data point in the provided sample of unlabeled points by fitting the classifier and directly measuring the resulting test performance.
+The point with the best test performance is selected and added to the labeled pool for that iteration.
+We noticed that this oracle is overfitting on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
 To circumvent this problem, we introduced margin sampling as a fallback option for the oracle.
@@
 	\end{tabular}
-	\caption{AUC values for all algorithms on the tabular datasets. Higher is better.
-	{\color{red}Is the STD even helping us here? Since we repeated 50 times, we can use the values as "true medians"} }
+	\caption{Median AUC values and standard deviation for all algorithms on the tabular datasets. Higher is better.}
 \end{table}

### 2023-05-31 results/generate_result_tables.py

@@ -3,17 +3,9 @@ from os.path import exists, join
 import pandas as pd
 import numpy as np
 
-def _pad_nans_with_last_value(df:pd.DataFrame):
-    max_len = len(df)
-    for col in df:
-        diff = max_len - sum(pd.notna(df[col]))
-        if diff > 0:
-            last_val = df[col][sum(pd.notna(df[col])) - 1]
-            df[col] = pd.concat([df[col].iloc[:-diff], pd.Series([last_val]*diff)], ignore_index=True)
-    return df
-
 def generate_table(agents, datasets, out_file):
     result_table = pd.DataFrame(index=agents, columns=datasets)
+    result_table["Sum"] = [0.0]*len(agents)
     for dataset in datasets:
         dataset_folder = os.path.join("runs", dataset)
         if not os.path.isdir(dataset_folder):
@@ -28,21 +20,24 @@ def generate_table(agents, datasets, out_file):
                 acc_file = join(agent_folder, "accuracies.csv")
                 if exists(acc_file):
                     accuracies = pd.read_csv(acc_file, header=0, index_col=0)
-                    accuracies = _pad_nans_with_last_value(accuracies)
                     values = accuracies.values
                     auc = np.sum(values, axis=0) / values.shape[0]
-                    result_table[dataset][agent] = "%1.3f +- %1.2f"%(np.mean(auc).item(), np.std(auc).item())
+                    result_table.loc[agent, "Sum"] += np.median(auc).item()
+                    result_table.loc[agent, dataset] = "%1.3f +- %1.2f"%(np.median(auc).item(), np.std(auc).item())
             else:
                 print(f"{agent} is missing for {dataset}")
 
     if len(result_table) > 0:
+        result_table = result_table.sort_values("Sum", ascending=False)
+        result_table = result_table.drop(columns=["Sum"])
+
         result_table_file = os.path.join("results", out_file)
         if os.path.exists(result_table_file):
             os.remove(result_table_file)
         result_table.to_csv(result_table_file)
 
 
-agents = ["Oracle", "SAL_23_01_16", "Coreset_Greedy", "TypiClust", "MarginScore", "ShannonEntropy", "RandomAgent", "Badge", "BALD", ]
+agents = ["Oracle", "Coreset_Greedy", "TypiClust", "MarginScore", "ShannonEntropy", "RandomAgent", "Badge", "BALD", ]
[…]

### 2023-06-01 doc/al_benchmark_paper.tex
paper update

@@
 $$\min \quad \mathbb{E}_{(x,y) \sim \rho} \ell(y, \hat{y}(x))$$
-Finding the optimal subset $\mathcal{U}_b$ is a combinatorial problem that is computationally not feasible for large $b$, so we allow sequential construction of the subset as a relaxation of the problem.
+Finding the optimal subset $\mathcal{U}_b$ is a combinatorial problem that is computationally not feasible for large $b$, so we allow sequential construction of the subset.
 %To construct the active learning setting from a labeled dataset $\mathcal{D} = \{(x_1, y_1),\ldots, (x_n,y_n)\}$, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U}$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class, removing them from $\mathcal{U}$ and recovering their label $\mathcal{L} := \{(x_1, A(x_1)),\ldots, (x_{k*C},A(x_{k*C}))\} \quad s.t.
@@
 %(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
-Even though the acquisition function in principle could output the full subset of $\mathcal{D}_\text{train}$, the combinatorial problem is computationally not feasible for large $B$ and we allow sequential construction of the subset as a relaxation of the problem.
+Even though the acquisition function in principle could output the full subset of $\mathcal{D}_\text{train}$, the combinatorial problem is computationally not feasible for large $B$ and we allow sequential construction of the subset.
 \\
@@
 	\begin{tabular}{l | c c c c c}
-		Paper & Sampling & \# Datasets & Domains & Algorithms & Oracle \\
+		Paper & Sampling & \#Datasets & \#Domains & \#Algorithms & Oracle \\
 		\hline
@@
 	\end{tabular}
-	\caption{Comparison of our benchmark with the existing literature}
+	\caption{Comparison of our benchmark with the existing literature. Oracle curves serve as an approximation of the best possible AL algorithm.}
 	\label{tab:benchmark_comparison}
@@
 {\color{red} Version: Braindump}\\
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[width=\linewidth]{img/ablation_bald}
+	\caption{Shows a reproduction of the performance analysis of BALD from the original authors \cite{gal2017deep} in comparison to entropy sampling and an adaptation of BALD for batch AL \cite{kirsch2019batchbald} with a batch size of 40.}
+	\label{fig:ablation_bald}
+\end{wrapfigure}
 Many different algorithms have been proposed for active learning. 
@@
 The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
-A good AL algorithm should be able to perform well in both scenarios. \\ [1mm]
-Since AUC is dependent on the chosen budget, we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
+A good AL algorithm should be able to perform well in both scenarios. 
+Since AL has high variance and is prone to outliers, we propose to report median AUC instead of mean AUC values. \\ [1mm]
+The AUC is dependent on the chosen budget, so we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
 In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
@@
 A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
-For a fair comparison of two AL algorithms, both algorithms should receive equal starting conditions in terms of train / validation split, initialization of classifier and even the state of minor systems like the optimizer of mini-batch sampler.
-Even though some implementations might have their own solution to some of these problems, to the best of our knowledge
-
-%Decided against a fixed time budget
-Given a seed, an evaluation on one dataset should always be done on the same validation split as well as the same random state for all included systems, like the mini batch sampler for model training or the initialization for the classifier itself.
-Even though different AL algorithms will pick different samples, making them unavailable for sampling in earlier or later batches, the theoretical decision tree for every possible choice in every iteration $i$ should stay the same.
-Since every possible trajectory cannot be precomputed and stored to disk, we need to resort to seeding.
-The default choice of setting a global seed at the start of the experiment is not sufficient here, since a single additional random draw from the random number generator completely changes the behavior of all other systems.
-This additional random number might be drawn during the initialization of the classification model or the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
-%This problem also applies to the initialization of the classification model $\theta$ and the initialization and querying of $\Omega$.
-The desired control only be archived by assigning a separate random number generator to all these processes.
-To the best of our knowledge, we are the first work that discusses this issue and proposes a solution for it.
+For a fair comparison of two AL algorithms, both algorithms should receive equal starting conditions in terms of train / validation split, initialization of classifier and even the state of minor systems like the optimizer or mini-batch sampler.
+Even though some implementations might have their own solution to some of these problems, to the best of our knowledge no previous work has discussed this topic in detail.
+At the core of this problem is the seeding utility in PyTorch, Tensorflow and other framework, whose default choice is a single global seed.
+Since many system draw random numbers from this seed all of them influence each other to a point where a single additional draw can completely change the model initialization or data split, depending on the order in the implementation.
+Even though you could find workarounds like re-setting the seed multiple times, this problem also extends to every AL iteration and the systems within.
+We propose an implementation that creates a separate Random Number Generator (RNG) for each of these systems to ensure equal testing conditions even when the AL algorithm, dataset or classifier changes.
 We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
-In summary, we introduce three different seeds: $s_\Omega$ for the acquisition function, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
+In summary, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
 Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
-% This seeding setup enables a reproduction of the oracle curve even when the classifier or AL algorithm is different.
+Some algorithms require a subsample to be drawn from $\mathcal{U}$ in order to reduce the computational cost in each iteration, while others need access to the full unlabeled pool (i.e.
+for effective clustering).
+If a subsample is required, it will be drawn from $s_\Omega$ and therefore will not influence other systems in the experiments.
+For each algorithm, we decided if subsampling is required based on our available hardware, but decided against setting a fixed time limit per experiment, since this would introduce unnecessary complexity into the benchmark.
+An overview of selected hyperparameters per AL algorithm can be found in Appendix \ref{app:agent_hyperparameters}.
 
@@
 Even though their reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
-\begin{minipage}{0.47\textwidth}
-	\begin{algorithm}[H]
+\begin{algorithm}
 	\caption{Oracle}\label{alg:oracle}
@@
 \end{algorithm}
-\end{minipage}
-\hspace{2mm}
-\begin{minipage}{0.5\linewidth}
-	\begin{figure}[H]
-		\includegraphics[width=\linewidth]{img/ablation_bald}
-		\caption{Shows a reproduction of the performance analysis of BALD from the original authors \cite{gal2017deep} in comparison to entropy sampling and an adaptation of BALD for batch AL \cite{kirsch2019batchbald} with a batch size of 40.}
-		\label{fig:ablation_bald}
-	\end{figure}
-\end{minipage} \\[2mm]
 In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
@@
 When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$). \\
-Alg. \ref{alg:oracle} replaces the acquisition function in the AL process.
+Alg. \ref{alg:oracle} replaces the algorithm in the AL process.
 
@@
 We allow algorithms to derive additional information of this set like predictions of the classifier, K-Means clustering or even training new classifiers.
-However, the algorithm may not incorporate external information like other datasets, queries to recover additional labels, or the test/validation set.
+However, the algorithm may not incorporate external information like other datasets, queries to recover additional labels, additional training steps for the classifier, or the test/validation set.
 
@@
 \textbf{Coreset \cite{sener2017active}}
-Employs K-Means clustering to try to cover all {\color{red}modalities (is this the right word?)} that are observed in the data.
+Employs K-Means clustering to try to cover the whole data distribution.
 Selects the unlabeled sample that is the furthest away from all cluster centers.
@@
 
-\begin{table}[]
-	\centering
-	\begin{tabular}{l|lll}
-		& Splice        & DNA           & USPS          \\
-		\hline
-		Oracle          & 0.830 +- 0.01 & 0.836 +- 0.02 & 0.823 +- 0.01 \\
-		SAL & 0.799 +- 0.01 & 0.797 +- 0.03 & 0.809 +- 0.01 \\
-		Coreset & 0.800 +- 0.01 & 0.795 +- 0.03 & 0.787 +- 0.02 \\
-		TypiClust       & 0.790 +- 0.01 & 0.771 +- 0.04 & 0.761 +- 0.02 \\
-		MarginScore     & 0.797 +- 0.02 & 0.795 +- 0.04 & 0.808 +- 0.01 \\
-		ShannonEntropy  & 0.799 +- 0.02 & 0.794 +- 0.04 & 0.807 +- 0.01 \\
-		RandomAgent     & 0.788 +- 0.01 & 0.765 +- 0.03 & 0.772 +- 0.01 \\
-		Badge           & 0.807 +- 0.01 & 0.769 +- 0.06 & 0.797 +- 0.02 \\
-		BALD            & 0.811 +- 0.01 & 0.743 +- 0.04 & 0.717 +- 0.05
-	\end{tabular}
-	\caption{Median AUC values and standard deviation for all algorithms on the tabular datasets. Higher is better.}
-\end{table}
+%\begin{table}[]
+%	\centering
+%	\begin{tabular}{l|lll}
+%		& Splice        & DNA           & USPS          \\
+%		\hline
+%		Oracle          & 0.830 +- 0.01 & 0.836 +- 0.02 & 0.823 +- 0.01 \\
+%		SAL & 0.799 +- 0.01 & 0.797 +- 0.03 & 0.809 +- 0.01 \\
+%		Coreset & 0.800 +- 0.01 & 0.795 +- 0.03 & 0.787 +- 0.02 \\
+%		TypiClust       & 0.790 +- 0.01 & 0.771 +- 0.04 & 0.761 +- 0.02 \\
+%		MarginScore     & 0.797 +- 0.02 & 0.795 +- 0.04 & 0.808 +- 0.01 \\
+%		ShannonEntropy  & 0.799 +- 0.02 & 0.794 +- 0.04 & 0.807 +- 0.01 \\
+%		RandomAgent     & 0.788 +- 0.01 & 0.765 +- 0.03 & 0.772 +- 0.01 \\
+%		Badge           & 0.807 +- 0.01 & 0.769 +- 0.06 & 0.797 +- 0.02 \\
+%		BALD            & 0.811 +- 0.01 & 0.743 +- 0.04 & 0.717 +- 0.05
+%	\end{tabular}
+%	\caption{Median AUC values and standard deviation for all algorithms on the tabular datasets. Higher is better.}
+%\end{table}
 %
@@
 	\hspace{1mm}
-	\includegraphics[width=0.49\linewidth]{img/eval_dna}
+	\includegraphics[width=0.49\linewidth]{img/eval_dna}\\[2mm]
+	\includegraphics[width=0.49\linewidth]{img/micro_splice}
+	\hspace{1mm}
+	\includegraphics[width=0.49\linewidth]{img/micro_dna}
 	\caption{Results for all algorithms on Splice and DNA, both from the tabular domain. Even within one domain, the performance of the same algorithm can vary drastically.}
@@
 \centering
-	\includegraphics[width=0.49\linewidth]{img/vector.png}
-	\includegraphics[width=0.49\linewidth]{img/vector_enc.png}
-	\includegraphics[width=0.49\linewidth]{img/text.png}
+	\includegraphics[width=0.49\linewidth]{img/macro_vector}
+	\includegraphics[width=0.49\linewidth]{img/macro_vector_enc}
+	\includegraphics[width=0.49\linewidth]{img/macro_img}
+	\includegraphics[width=0.49\linewidth]{img/macro_img_enc}
+	\includegraphics[width=0.49\linewidth]{img/macro_text}
 	\includegraphics[width=0.49\linewidth]{img/all_data.png}
@@
 
-\section{Other result tables}
-\begin{table}[H]
-	\begin{tabular}{l|lll}
-		& SpliceEncoded & DNAEncoded    & USPSEncoded   \\
-		\hline
-		Oracle          & 0.754 +- 0.02 & 0.726 +- 0.02 & 0.674 +- 0.01 \\
-		SAL & 0.675 +- 0.03 & 0.640 +- 0.04 & 0.634 +- 0.01 \\
-		Coreset & 0.690 +- 0.02 & 0.644 +- 0.05 & 0.607 +- 0.02 \\
-		TypiClust       & 0.695 +- 0.02 & 0.660 +- 0.03 & 0.643 +- 0.01 \\
-		MarginScore     & 0.675 +- 0.03 & 0.643 +- 0.05 & 0.632 +- 0.02 \\
-		ShannonEntropy  & 0.673 +- 0.03 & 0.638 +- 0.05 & 0.626 +- 0.02 \\
-		RandomAgent     & 0.680 +- 0.03 & 0.633 +- 0.04 & 0.594 +- 0.02 \\
-		Badge           & 0.670 +- 0.04 & 0.600 +- 0.07 & 0.597 +- 0.02 \\
-		BALD            & 0.660 +- 0.04 & 0.597 +- 0.06 & 0.652 +- 0.01
-	\end{tabular}
-	\caption{AUC values for all algorithms on the encoded tabular datasets. Higher is better.}
-\end{table}
-%
-\begin{table}[H]
-	\begin{tabular}{l|ll}
-		& Cifar10Encoded & FashionMnistEncoded \\
-		\hline
-		Oracle          & 0.699 +- 0.01  & 0.721 +- 0.01       \\
-		SAL & 0.623 +- 0.01  & 0.667 +- 0.01       \\
-		Coreset & 0.641 +- 0.01  & 0.674 +- 0.01       \\
-		TypiClust       & 0.627 +- 0.01  & 0.638 +- 0.01       \\
-		MarginScore     & 0.641 +- 0.01  & 0.676 +- 0.01       \\
-		ShannonEntropy  & 0.637 +- 0.02  & 0.680 +- 0.01       \\
-		RandomAgent     & 0.598 +- 0.02  & 0.633 +- 0.01       \\
-		Badge           & 0.645 +- 0.02  & 0.681 +- 0.01       \\
-		BALD            & 0.633 +- 0.01  & 0.666 +- 0.01      
-	\end{tabular}
-	\caption{AUC values for all algorithms on the encoded image datasets. Higher is better.}
-\end{table}
-%
-\begin{table}[H]
-	\begin{tabular}{l|ll}
-		& TopV2         & News          \\
-		\hline
-		Oracle          & 0.860 +- 0.01 & 0.448 +- 0.01 \\
-		SAL & 0.831 +- 0.01 & 0.312 +- 0.01 \\
-		Coreset & 0.816 +- 0.02 & 0.355 +- 0.02 \\
-		TypiClust       & 0.718 +- 0.02 & 0.314 +- 0.01 \\
-		MarginScore     & 0.821 +- 0.02 & 0.357 +- 0.01 \\
-		ShannonEntropy  & 0.794 +- 0.02 & 0.342 +- 0.01 \\
-		RandomAgent     & 0.776 +- 0.02 & 0.349 +- 0.01 \\
-		Badge           & 0.824 +- 0.01 & 0.343 +- 0.01 \\
-		BALD            & 0.825 +- 0.02 & 0.347 +- 0.01
-	\end{tabular}
-\caption{AUC values for all algorithms on the text datasets. Higher is better.}
-\end{table}
+\section{All Results}\label{app:all_results}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.49\linewidth]{img/eval_splice}
+	\includegraphics[width=0.49\linewidth]{img/eval_splice_enc} \\ [2mm]
+	\includegraphics[width=0.49\linewidth]{img/micro_splice.jpg}
+	\includegraphics[width=0.49\linewidth]{img/micro_splice_enc.jpg} \\ [4mm]
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.49\linewidth]{img/eval_dna}
+	\includegraphics[width=0.49\linewidth]{img/eval_dna_enc} \\ [2mm]
+	\includegraphics[width=0.49\linewidth]{img/micro_dna.jpg}
+	\includegraphics[width=0.49\linewidth]{img/micro_dna_enc.jpg} \\ [4mm]
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.49\linewidth]{img/eval_usps}
+	\includegraphics[width=0.49\linewidth]{img/eval_usps_enc} \\ [2mm]
+	\includegraphics[width=0.49\linewidth]{img/micro_usps.jpg}
+	\includegraphics[width=0.49\linewidth]{img/micro_usps_enc.jpg} \\ [4mm]
+\end{figure}
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.49\linewidth]{img/eval_cifar10}
+\includegraphics[width=0.49\linewidth]{img/eval_cifar10_enc} \\ [2mm]
+\includegraphics[width=0.49\linewidth]{img/micro_cifar10.jpg}
+\includegraphics[width=0.49\linewidth]{img/micro_cifar10_enc.jpg} \\ [4mm]
+\end{figure}
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.49\linewidth]{img/eval_fmnist}
+\includegraphics[width=0.49\linewidth]{img/eval_fmnist_enc} \\ [2mm]
+\includegraphics[width=0.49\linewidth]{img/micro_fmnist.jpg}
+\includegraphics[width=0.49\linewidth]{img/micro_fmnist_enc.jpg} \\ [4mm]
+\end{figure}
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.49\linewidth]{img/eval_topv2}
+\includegraphics[width=0.49\linewidth]{img/eval_news} \\ [2mm]
+\includegraphics[width=0.49\linewidth]{img/micro_topv2.jpg}
+\includegraphics[width=0.49\linewidth]{img/micro_news.jpg} \\ [4mm]
+\end{figure}
 
@@
 \end{figure}
+
+\section{Hyperparameters per AL Algorithm}\label{app:agent_hyperparameters}
+\begin{table}[H]
+	TODO
+\end{table}
 

### 2023-06-02 doc/al_benchmark_paper.tex
batch bald 50

@@
 
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\EE}{\mathbb{E}}
+\newcommand{\N}{\mathbb{N}}
+\newcommand{\D}{\mathcal{D}}
+\newcommand{\X}{\mathcal{X}}
+\newcommand{\Y}{\mathcal{Y}}
+\newcommand{\U}{\mathcal{U}}
+\newcommand{\LL}{\mathcal{L}}
+\newcommand{\test}{\text{test}}
+\newcommand{\train}{\text{train}}
+
 \title{Towards Comparable Active Learning}
 \author{%
-	Thorben Werner 
-	\thanks{Institute of Computer Science - Information Systems and Machine Learning Lab (ISMLL)} 
-	\\
-	University of Hildesheim\\
-	Universitätsplatz 1\, 31141 Hildesheim \\
-	\texttt{werner@ismll.de} \\
+	Anonymous Authors
+	%Thorben Werner 
+	%\thanks{Institute of Computer Science - Information Systems and Machine Learning Lab (ISMLL)} 
+	%\\
+	%University of Hildesheim\\
+	%Universitätsplatz 1\, 31141 Hildesheim \\
+	%\texttt{werner@ismll.de} \\
 	% examples of more authors
-	\And
-	Johannes Burchert$^*$ \\
-	University of Hildesheim\\
-	Universitätsplatz 1, 31141 Hildesheim \\
-	\texttt{burchert@ismll.de} \\
-	\AND
-	Prof. Lars Schmidt-Thieme$^*$ \\
-	University of Hildesheim\\
-	Universitätsplatz 1, 31141 Hildesheim \\
-	\texttt{schmidt-thieme@ismll.uni-hildesheim.de} \\
-	% \And
-	% Coauthor \\
-	% Affiliation \\
-	% Address \\
-	% \texttt{email} \\
-	% \And
-	% Coauthor \\
-	% Affiliation \\
-	% Address \\
-	% \texttt{email} \\
+	%\And
+	%Johannes Burchert$^*$ \\
+	%University of Hildesheim\\
+	%Universitätsplatz 1, 31141 Hildesheim \\
+	%\texttt{burchert@ismll.de} \\
+	%\AND
+	%Prof. Lars Schmidt-Thieme$^*$ \\
+	%University of Hildesheim\\
+	%Universitätsplatz 1, 31141 Hildesheim \\
+	%\texttt{schmidt-thieme@ismll.uni-hildesheim.de}
 }
@@
 \begin{abstract}
-	%Active Learning (AL) is a technique used in machine learning to improve data efficiency by selecting the most informative samples for annotation. 
-	In this paper we address the issue of inconsistent results in active learning (AL) literature.
-	Previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results.
+	Active Learning has received significant attention in the field of machine learning for its potential in selecting the most informative samples for labeling, thereby reducing data annotation costs.
+However, the lack of reproducibility of performance gains reported in recent literature has created a chaotic landscape in Active Learning research.
+	This paper addresses these issues of inconsistent results in active learning (AL) literature.
+	%Previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results.
 This inconsistency leads to a chaotic landscape of AL algorithms.
-	To the best of our knowledge, we propose the first AL benchmark that tests algorithms in 3 major domains; Tabular, Image and Text.
-	Furthermore, we discuss overlooked problems for reproducing AL experiments that can lead to unfair comparisons and increased variance in the results.
-	We propose a framework for choosing hyperparameters of AL setups and accurately controlling the experiments.
-	Using this framework, we report empirical results for 6 algorithms on 7 datasets and aggregate them into a domain-specific ranking of AL algorithms.
+	To the best of our knowledge, we propose the first AL benchmark that tests algorithms in 3 major domains: Tabular, Image and Text.
+	Furthermore, we highlight overlooked problems for reproducing AL experiments that can lead to unfair comparisons and increased variance in the results.
+	To tackle these challenges, we propose an experimental protocol to accurately control the experiments.
+	We report empirical results for 6 widely used algorithms on 7 datasets and aggregate them into a domain-specific ranking of AL algorithms.
 \end{abstract}
@@
 \section{Introduction}
-Active Learning (AL) plays an important role in our society that applies machine learning to more and more areas and therefore has a high demand for labeled data in more and more areas.
-A problem that concerns academic researchers and practitioners in businesses alike and even could be extended to education in schools and hobbyists around the world.
-On top of providing a principled way to labeled unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
-Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
-Among others, the authors of \cite{zhou2021towards} have pointed out severe inconsistencies in results of AL papers in recent years. 
-In their supplementary materials they conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while all following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm.
+Deep neural networks (NN) have produced state-of-the-art results on many important supervised learning tasks.
+Since Deep NNs usually require large amounts of labeled training data, Active Learning (AL) can be used to instead to select the most informative samples out of a large pool of unlabeled data, so that only these samples need to be labeled.
+It has been shown that a small labeled set of this nature can be used to train well-performing models \cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}. \\
+%On top of providing a principled way to labeled unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
+%Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
+In the last decade many different algorithms for AL have been proposed.
+Even though, almost every method has reported lifts of their approach over random selection and all predecessors \footnote{out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not establish a new SOTA performance in their results sections}, AL research faces two central difficulties:
+(i) The experiments are often carried out on different datasets and model architectures, hindering direct comparison,
+(ii) The reported results prove to be very difficult to reproduce.
+While multiple benchmark suites have been proposed to solve (i), to the best of our knowledge, we are the first to compare AL algorithms on all 3 data domains of vector, image and text.
+Regarding (ii), \cite{zhou2021towards} has pointed out severe inconsistencies in results of AL papers in recent years. 
+They conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm (\cite{zhou2021towards} Appendix A).
 The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art results by significantly outperforming everyone else, while the vast majority of results proves to be non-reproducible.
-
-\subsection{Contributions}
-\begin{enumerate}
-	\item Evaluation of Active Learning algorithms on datasets from 3 different domains, including synthetic data that highlights principled shortcomings of existing approaches.
-	\item Novel experimental protocol for seeding the experiment with 3 different seeds to allow full control and reproducibility and analysis of how many restarts are required to converge to the true median performance reliably.
-	\item Simple algorithm for an Oracle-Curve that can be constructed greedily and does not rely on search.
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{(From Vijaya) Problem Description}
-Given $n = l+u$ data points with $l \in \mathbb{N}$ many labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ many unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathbb{R}^M \to \mathbb{R}^C$ that can label $x$.
-We call $x \in \mathbb{R}^M$, $y \in \mathbb{R}^C$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$.
-Find an acquisition function $F: \mathcal{U},\mathcal{L}, b \mapsto \mathcal{U}_b$ that selects $b$ many examples from $\mathcal{U}$ such that $\mathcal{U}_b \subseteq \mathcal{U}$, $|\mathcal{U}_b|=b$ for labeling and the expected loss $\ell: \mathbb{R}^C \times \mathbb{R}^C \to \mathbb{R}$ (cross entropy loss) of the machine learning algorithm $\hat{y}$ (ex.
-classification) trained on $\mathcal{L} \cup \{\left(x, A(x)\right) \mid x \in \mathcal{U}_b\}$ is minimal:
-$$\min \quad \mathbb{E}_{(x,y) \sim \rho} \ell(y, \hat{y}(x))$$
-Finding the optimal subset $\mathcal{U}_b$ is a combinatorial problem that is computationally not feasible for large $b$, so we allow sequential construction of the subset.
-%To construct the active learning setting from a labeled dataset $\mathcal{D} = \{(x_1, y_1),\ldots, (x_n,y_n)\}$, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U}$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class, removing them from $\mathcal{U}$ and recovering their label $\mathcal{L} := \{(x_1, A(x_1)),\ldots, (x_{k*C},A(x_{k*C}))\} \quad s.t.
-\quad \mathcal{U} \cap \mathcal{L} = \emptyset$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{(From Lars) Problem Description}
-Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$.
 \\
-Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
-\begin{align*}
-	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
-\end{align*}
-called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal.
-%\begin{align*}
-%	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
-%	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
-%	% \operatorname{avg} \{
-%	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
-%	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
-%	\ell(y, \hat y(x)) 
-%	\\
-%	\text{with }
-%	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
-%	\\ 
-%	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
-%	\quad b\in 1{:}B
-%\end{align*}
-%(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
-Even though the acquisition function in principle could output the full subset of $\mathcal{D}_\text{train}$, the combinatorial problem is computationally not feasible for large $B$ and we allow sequential construction of the subset.
-\\
-%
-%To construct the active learning setting from existing labeled datasets, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Overview}
-We constrain our work on pool-based active learning where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
-Specifically, we are not experimenting on so-called batch active learning, where at each iteration multiple unlabeled samples are chosen at the same time.
-Even though batch AL is the more active research domain, it does not have a principled advantage over single-sample AL except speed of computation.
-Not only is the problem of optimizing a portfolio of unlabeled samples more complicated to solve, the algorithms also have systematically less information per sample to work with.
-For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
-A performance comparison of batch AL and single-sample AL can be found in Fig.
-\ref{fig:ablation_bald}, which reproduces the message of Figure 1 from the paper \cite{gal2017deep} that proposed BALD, one of the SOTA algorithms for AL.
-We can see that the batched version of BALD \cite{krizhevsky2009learning} can at most perform on-par with the single-sample algorithm.
-Fig. \ref{fig:ablation_bald} also serves as a proof of concept for our provided code base.
-Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with experiments on at least two data domains.
-\begin{table}[h]
+In this work we propose an evaluation protocol that was designed to cope with high variance in the performances of AL algorithms as well as being fully controllable regardless of the combination of dataset, model and AL algorithm.\\
+We focus our work on single-sample pool-based AL where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
+Specifically, we are not experimenting on so-called Batch Active Learning, where at each iteration multiple unlabeled samples are chosen at the same time.
+Even though most proposed algorithms (and benchmarks) only do Batch AL, it does not have a principled advantage over single-sample AL except speed of computation.
+The problem of optimizing a portfolio of unlabeled samples more complicated to solve and the algorithms also have systematically less information per sample to work with.
+%For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
+A performance comparison of batch AL and single-sample AL can be found in Fig. \ref{fig:ablation_bald}.
+We can see that for two established algorithms BALD \cite{krizhevsky2009learning} and BADGE \cite{ashdeep} Batch AL can at most perform on-par with single-sample AL. \\
+Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature.
+The table also includes methodological AL papers with experiments on at least two data domains.
+\begin{table}
 	\centering
@@
 
-%Manual NIPS Benchmark paper eval (out of 6):
-%\begin{itemize}
-%	\item Has new method: 1
-%	\item Multiple domains: 2
-%	\item $> 5$ datasets: 3
-%	\item $> 5$ baselines: 4
-%\end{itemize}
+\subsection*{Contributions}
+\begin{enumerate}
+	\item Evaluation of Active Learning algorithms on datasets from 3 different domains, including synthetic data that highlights principled shortcomings of existing approaches.
+	\item Novel experimental protocol for seeding the experiment with 3 different seeds to allow full control, reproducibility and analysis of how many restarts are required to converge to the true median performance reliably.
+	\item Simple algorithm for an Oracle that can be constructed greedily and does not rely on search.
+\end{enumerate}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Problem Description}
+Given two spaces $\mathcal{X}, \mathcal{Y}$, $n=l+u$ data points with $l \in \mathbb{N}$ labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathbb{R}^\mathcal{X} \to \mathbb{R}^\mathcal{Y}$ that can label $x$.
+We call $x \in \mathcal{X}$, $y \in \mathcal{Y}$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$. 
+Find an acquisition function $a: \U^{(i)},\LL^{(i)} \mapsto x^{(i)}$ that iteratively selects the next unlabeled point $x^{(i)} \in \U^{(i)}$ for labeling
+\begin{align*}
+	\LL^{(i+1)} &\gets \LL^{(i)} \cup \{\left(x^{(i)}, A(x^{(i)})\right)\} \\
+	\U^{(i+1)} &\gets \U^{(i)} \setminus x^{(i)} %, i \in 1{:}B
+\end{align*}
+so that the expected loss $\ell: \mathbb{R}^\mathcal{Y} \times \mathbb{R}^\mathcal{Y} \to \mathbb{R}$ of the machine learning algorithm $\hat y$ after $B$ iterations is minimal: 
+$$\min \quad \mathbb{E}_{(x,y) \sim \rho} \ell(y, \hat{y}(\LL^{(B)}))$$
+This formulation is a special case of the problem formulation in Appendix \ref{app:problem_formulation}.
+
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Related Work}
-{\color{red} Version: Braindump}\\
-\begin{wrapfigure}{r}{0.5\textwidth}
+\begin{figure}
 	\centering
-	\includegraphics[width=\linewidth]{img/ablation_bald}
+	\includegraphics[width=0.49\linewidth]{img/ablation_bald}
+	\includegraphics[width=0.49\linewidth]{img/ablation_bald}
 	\caption{Shows a reproduction of the performance analysis of BALD from the original authors \cite{gal2017deep} in comparison to entropy sampling and an adaptation of BALD for batch AL \cite{kirsch2019batchbald} with a batch size of 40.}
 	\label{fig:ablation_bald}
-\end{wrapfigure}
-Many different algorithms have been proposed for active learning. 
-In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
-AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
-Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
-They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
-This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
-Uncertainty-based approaches use metrics to measure the classifiers state.
-Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
-This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
+\end{figure}
+%Many different algorithms have been proposed for Active Learning. 
+%In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
+%AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
+%Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
+%They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
+%This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
+%Uncertainty-based approaches use metrics to measure the classifiers state.
+%Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
+%This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
 %
-Some previous work also aimed to provide a benchmark suite for active learning:
-The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
-While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
-The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
+Multiple benchmark suites have already been proposed for Active Learning:
+The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus exclusively on Batch AL in the image domain.
+While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
+The authors of \cite{hu2021towards} study models that are trained with AL techniques in the image and text domain.
 They test for several different properties of the models including robustness, response to compression techniques and final performance.
+\cite{zhou2021towards} proposed an oracle algorithm for AL that uses Simulated Annealing search to approximate a solution for the optimal subset of labeled data.
+Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
+The employed AL algorithms for our experiments are introduced in Section \ref{sec:sampling_strategies}.
 
@@
 \subsection{Evaluation}\label{sec:evaluation}
-Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
+Following \cite{zhou2021towards}, the quality of an AL algorithm is evaluated by an ``anytime" protocol that incorporates classification performance at every iteration, as opposed to evaluating final performance after the budget is exhausted.
 We employ the normalized area under the accuracy curve (AUC):
@@
 The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
-A good AL algorithm should be able to perform well in both scenarios. 
-Since AL has high variance and is prone to outliers, we propose to report median AUC instead of mean AUC values. \\ [1mm]
-The AUC is dependent on the chosen budget, so we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
+AL algorithms have to perform well in both scenarios. 
+Since AL performance has high variance and is prone to outliers, we propose to aggregate AUC values with the median instead of mean. \\ [1mm]
+Since the AUC depends on the chosen budget, we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
 In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
-Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets.
+Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets
+\footnote{Since we only consider single-sample AL, increasing the budget comes at a high computational cost. Therefore we could not freely increase the budget above 2000.}.
 We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
@@
 	\centering
-	\includegraphics[width=\linewidth]{img/ablation_restarts.png}
-	\caption{Random draws from an experimental distribution on the Splice datasets with different numbers of repetitions.
+	\includegraphics[width=\linewidth]{img/ablation_restarts}
+	\caption{Random draws from an experimental distribution on the Splice dataset with different numbers of repetitions.
 Each point on the Y-axis represents a cross-validated result that could have been reported in a paper.
@@
 \end{figure}
-To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
+To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repetitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
 One way to reduce the number of necessary repetitions is to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
 We ultimately decided in favor of high variance and high number of repetitions as the high variance accurately reflects real world applications of AL.
+Fixing for example the classifier initialization to random draws from a constant seed does not have a corresponding use-case in the real world.
 
@@
 For a fair comparison of two AL algorithms, both algorithms should receive equal starting conditions in terms of train / validation split, initialization of classifier and even the state of minor systems like the optimizer or mini-batch sampler.
-Even though some implementations might have their own solution to some of these problems, to the best of our knowledge no previous work has discussed this topic in detail.
-At the core of this problem is the seeding utility in PyTorch, Tensorflow and other framework, whose default choice is a single global seed.
-Since many system draw random numbers from this seed all of them influence each other to a point where a single additional draw can completely change the model initialization or data split, depending on the order in the implementation.
+Additionally, previous works have noted a significant adverse effect of training stochasticity on the performance of AL algorithms (\cite{zhou2021towards, lowell2018practical}).
+Even though implementations might have their own solution to some of these problems, to the best of our knowledge no previous work has discussed this topic in detail.
+At the core of this problem is the seeding utility in PyTorch, Tensorflow and other frameworks, whose default choice is a single global seed.
+Since many systems draw random numbers from this seed, all of them influence each other to a point where a single additional draw can completely change the model initialization or data split, depending on the ordering of these two in the implementation.
 Even though you could find workarounds like re-setting the seed multiple times, this problem also extends to every AL iteration and the systems within.
@@
 \subsection{Oracle}\label{sec:oracle}
-Posing active learning as a combinatorial problem, the oracle set $\mathcal{U}_b$ for a given dataset, model and training procedure would be the set that induces the highest AUC score for a given budget.
+Posing Active Learning as a combinatorial problem, the oracle set $\mathcal{U}_b$ for a given dataset, model and training procedure would be the set that induces the highest AUC score for a given budget.
 However, since this problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence. 
@@
 Even though their reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
-\begin{algorithm}
+In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
+Our oracle simply evaluates every data point $u_k = \operatorname{unif(\mathcal{U} | k = 1 \ldots \tau)}$ in a subsample of unlabeled points by fitting the classifier on $\mathcal{L}^{(i)} \cup u_k$ and directly measuring the resulting test performance.
+The point with the best test performance is selected and added to the labeled pool for that iteration.
+We noticed that this oracle is overfitting on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
+This can happen for example, if the oracle picked a labeled set that enables the classifier to correctly classifier a big portion of easy samples in the test set, but now fails to find the next single unlabeled point that would enable the classifier to succeed on one of the hard samples in the test set.
+This leads to a situation, where the selected data point is basically random.\\
+\begin{minipage}{0.5\linewidth}
+
+	To circumvent this problem, we introduced margin sampling \cite{wang2014new} as a fallback option for the oracle.
+	Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
+	The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.
+	In the algorithm $\operatorname{Retrain}(\mathcal{L}^{(i)}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns the accuracy on the test set $\mathcal{D}_\text{test}$. \\ 
+	$y_t$ is shorthand for the corresponding label of $u^{(i)}_{t}$ that can be recovered from the dataset labels. \\
+	When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$). \\
+	Alg. \ref{alg:oracle} replaces the algorithm in the AL process.
+\end{minipage}
+\hspace{2mm}
+\begin{minipage}{0.45\linewidth}
+	\begin{algorithm}[H]
 		\caption{Oracle}\label{alg:oracle}
@@
 	\end{algorithm}
-In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
-Our oracle simply evaluated every data point in the provided sample of unlabeled points by fitting the classifier and directly measuring the resulting test performance.
-The point with the best test performance is selected and added to the labeled pool for that iteration.
-We noticed that this oracle is overfitting on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
-To circumvent this problem, we introduced margin sampling as a fallback option for the oracle.
-Whenever the oracle does not find an unlabeled point that results in an increase in performance (indicating an overfitting position), it defaults to margin sampling in that iteration.
-The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.
-In the algorithm $\operatorname{Retrain}(\mathcal{L}^{(i)}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns the accuracy on the test set $\mathcal{D}_\text{test}$. \\ 
-$y_t$ is shorthand for the corresponding label of $u^{(i)}_{t}$ that can be recovered from the dataset labels. \\
-When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$). \\
-Alg. \ref{alg:oracle} replaces the algorithm in the AL process.
+\end{minipage}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Available Information}
+%\subsection{Available Information}
 At each iteration $i$ the AL algorithm needs to pick an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \theta^{(i)}, \text{opt}_\theta\}$, where $\theta^{(i)}$ is the current classifier and $\text{opt}_\theta$ is the optimizer used to fit $\theta^{(i)}$.
@@
 	\includegraphics[width=0.49\linewidth]{img/macro_text}
-	\includegraphics[width=0.49\linewidth]{img/all_data.png}
+	\includegraphics[width=0.49\linewidth]{img/macro_all.jpg}
 	\caption{Critical Difference Diagram for all algorithms grouped by domain and all domains combined.
@@
 
-\section{Conclusion}
+\section{Discussion}
+Domains are very different \\
+Even within one domain we have stark differences (Fig. \ref{fig:eval_vector}) \\
+Best 3 ranks for all domains are 3, 3.1 and 3.4 ... No clear winner \\
+At least on average everything is better that random \\
+BADGE: "Separately, we notice that diversity sampling only seems to work well when either the model has good architectural priors (inductive biases) built in, or when the data are easy to learn.
+Otherwise, penultimate layer representations are not meaningful, and diverse sampling can be deleterious.
+For this reason, CORESET often performs worse than random on sufficiently complex data when not using a convolutional network (Figure 3b)" \\
 
@@
 \appendix
+
+\section{Problem Formulation}\label{app:problem_formulation}
+Given
+\begin{itemize}
+	\item a number $B\in\N$ (called budget),
+	\item two spaces $\X$ and $\Y$,  {\tiny e.g., $\X:=\R^M, \Y:=\R^T$},
+	\item a sample $\D_1,\ldots,\D_N \in (\X\times \Y)^*$ of
+	sequences of pairs $(x,y)$  from an unknown distribution $p$
+	(called datasets),
+	%  {\tiny with $|\D_n|\geq B$ for all $n\in 1{:}N$,}
+	{\tiny with $p(\D)=0$ for $|\D|<B$,}
+	\item a function $\ell:\Y\times\Y\rightarrow\R$ (called loss), and
+	\item a function $\hat y:  (\X\times \Y)^* \times \X^* \rightarrow \Y^\X$
+	(called learning algorithm), \\
+	{\tiny where $\Y^\X$ is the space of all function from $\X$ to $\Y$}
+\end{itemize}
+find a function
+\vspace*{-0.5cm}
+\begin{align*}
+	a: (\X\times \Y)^* \times \X^* &\rightarrow \N
+	\quad\quad \text{\tiny (with $a(\D,X) \leq |X|$)} \\
+	& \text{\tiny where $a(\D,X)$ selects an unlabeled instance from $X$}
+\end{align*}
+% which is equivariant in the second argument,
+
+called acquisition function,
+s.t. the expected loss of a model learned on all predictors plus $B$ sequentially acquired targets
+is minimal:
+\begin{align*}
+	\min\ \EE\   &  \{
+	% \operatorname{avg}\limits_{(x,y)\in\D\test} \ell(y, \hat y(x))
+	\ell(\hat y, \D\test)
+	\mid \D\sim p, (\D\train,\D\test):= \text{split}(\D) \}
+	\\
+	\text{with }
+	\hat y:= & A( (\D_{\train_{n_1}},\ldots,\D_{\train_{n_B}}), \D\train|_{\X})
+	\\ 
+	n_b := & a( (\D_{\train_{n_1}},\ldots,\D_{\train_{n_{b-1}}}), \D\train|_{\X}) ,
+	\quad b\in 1{:}B
+\end{align*}
 
@@
 	\centering
-	\includegraphics[width=\linewidth]{img/ablation_restarts_2.png}
+	\includegraphics[width=\linewidth]{img/ablation_restarts_2}
 \end{figure}

### 2023-06-02 agents/bald.py

@@ -68,7 +68,7 @@ class BALD(BaseAgent):
 
 class BatchBALD(BaseAgent):
 
-    def __init__(self, agent_seed, config, dropout_trials=5, batch_size=40):
+    def __init__(self, agent_seed, config, dropout_trials=5, batch_size=50):
         super().__init__(agent_seed, config)
         assert "current_run_info" in config and "encoded" in config["current_run_info"]
         self.dropout_trials = dropout_trials

### 2023-06-02 doc/al_benchmark_paper.tex

--- 2023-06-01 doc/al_benchmark_paper.tex
@@
 \section{Related Work}
-{\color{red} Version: Braindump}\\
 \begin{wrapfigure}{r}{0.5\textwidth}
@@
 \section{Conclusion}
+Domains are very different \\
+Even within one domain we have stark differences (Fig. \ref{fig:eval_vector}) \\
+Best 3 ranks for all domains are 3, 3.1 and 3.4 ... No clear winner \\
+At least on average everything is better that random
 

### 2023-06-02 agents/__init__.py
batch badge

@@ -4,6 +4,6 @@ from agents.random_agent import RandomAgent, BatchRandomAgent
 from agents.sal import SAL
 from agents.shannon_entropy import ShannonEntropy
 from agents.bald import BALD, BatchBALD
-from agents.badge import Badge
+from agents.badge import Badge, BatchBadge
 from agents.typiclust import TypiClust
 

### 2023-06-02 agents/badge.py

@@ -29,11 +29,11 @@ class Badge(BaseAgent):
                       budget:int, added_images:int,
                       initial_test_acc:float, current_test_acc:float,
                       classifier: nn.Module, optimizer: Optimizer,
-                      sample_size=100) -> Union[Tensor, dict]:
+                      sample_size=8000) -> Union[Tensor, dict]:
 
         assert hasattr(classifier, "_encode"), "The provided model needs the '_encode' function"
-        replacement_needed = len(x_unlabeled) < sample_size
-        state_ids = self.agent_rng.choice(len(x_unlabeled), sample_size, replace=replacement_needed)
+        sample_size = min(len(x_unlabeled), sample_size)
+        state_ids = self.agent_rng.choice(len(x_unlabeled), sample_size, replace=False)
 
         gradEmbedding = self._get_grad_embedding(x_unlabeled[state_ids], classifier, sample_size)
         chosen = self._init_centers(gradEmbedding, 1)[0]
@@ -79,16 +79,17 @@ class Badge(BaseAgent):
 
     def _init_centers(self, X, K):
         ind = torch.argmax(torch.linalg.norm(X, dim=1)).item()
-        mu = [X[ind]]
+        X_arr = X.cpu().numpy()
+        mu = [X_arr[ind]]
         indsAll = [ind]
-        centInds = [0.] * len(X)
+        centInds = [0.] * len(X_arr)
         cent = 0
         while len(mu) < K:
             if len(mu) == 1:
-                D2 = pairwise_distances(X, mu).ravel().astype(float)
+                D2 = pairwise_distances(X_arr, mu).ravel().astype(float)
             else:
-                newD = pairwise_distances(X, [mu[-1]]).ravel().astype(float)
-                for i in range(len(X)):
+                newD = pairwise_distances(X_arr, [mu[-1]]).ravel().astype(float)
+                for i in range(len(X_arr)):
                     if D2[i] >  newD[i]:
                         centInds[i] = cent
                         D2[i] = newD[i]
@@ -97,7 +98,7 @@ class Badge(BaseAgent):
             Ddist = (D2 ** 2)/ sum(D2 ** 2)
             customDist = stats.rv_discrete(name='custm', values=(np.arange(len(D2)), Ddist))
             ind = customDist.rvs(size=1)[0]
-            mu.append(X[ind])
+            mu.append(X_arr[ind])
             indsAll.append(ind)
             cent += 1
         # gram = np.matmul(X[indsAll], X[indsAll].T)
[…]

### 2023-06-02 core/helper_functions.py

@@ -167,6 +167,8 @@ def get_agent_by_name(name:str)->Union[Callable, BaseAgent]:
         return agents.BatchRandomAgent
     elif name == "batchbald":
         return agents.BatchBALD
+    elif name == "batchbadge":
+        return agents.BatchBadge
     else:
         raise ValueError(f"Agent name '{name}' not recognized")
 

### 2023-06-02 evaluate.py

@@ -15,7 +15,7 @@ parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--agent_seed", type=int, default=1)
 parser.add_argument("--pool_seed", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
-parser.add_argument("--agent", type=str, default="batchbald")
+parser.add_argument("--agent", type=str, default="badge")
 parser.add_argument("--dataset", type=str, default="splice")
 parser.add_argument("--encoded", type=int, default=0)
 # parser.add_argument("--sample_size", type=int, default=20)

### 2023-06-04 doc/al_benchmark_paper.tex
batch badge
paper update

--- 2023-06-02 doc/al_benchmark_paper.tex
@@
 In the last decade many different algorithms for AL have been proposed.
-Even though, almost every method has reported lifts of their approach over random selection and all predecessors \footnote{out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not establish a new SOTA performance in their results sections}, AL research faces two central difficulties:
+Even though, almost every method has reported lifts all it's predecessors 
+\footnote{out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not claim a new SOTA performance in their result section}
+, AL research faces two central difficulties:
 (i) The experiments are often carried out on different datasets and model architectures, hindering direct comparison,
-(ii) The reported results prove to be very difficult to reproduce.
+(ii) the reported results prove to be very difficult to reproduce.
 While multiple benchmark suites have been proposed to solve (i), to the best of our knowledge, we are the first to compare AL algorithms on all 3 data domains of vector, image and text.
@@
 They conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm (\cite{zhou2021towards} Appendix A).
-The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art results by significantly outperforming everyone else, while the vast majority of results proves to be non-reproducible.
+The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art performance
+%by significantly outperforming everyone else
+, while the vast majority of results proves to be non-reproducible. \\
+In this work we propose an evaluation protocol that was designed to cope with the high variance in the performances of AL algorithms as well as being fully controllable regardless of the combination of dataset, model and AL algorithm.\\
+We focus our work on single-sample pool-based AL where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
+Specifically, we are not experimenting on so-called batch Active Learning, where at each iteration multiple unlabeled samples are chosen at the same time.
+Even though most proposed algorithms (and benchmarks) are batched, they do not have a principled advantage over single-sample AL except speed of computation.
+The problem of optimizing a portfolio of unlabeled samples in each iteration is more complicated to solve and the algorithms have systematically less information per sample to work with.
+A performance comparison of batch AL and single-sample AL can be found in Fig. \ref{fig:ablation_bald}.
+We can see that for two established algorithms BALD \cite{krizhevsky2009learning} and BADGE \cite{ashdeep} Batch AL can at most perform on-par with single-sample AL, independent of wether algorithm was originally designed for batch AL or not.
 \\
-In this work we propose an evaluation protocol that was designed to cope with high variance in the performances of AL algorithms as well as being fully controllable regardless of the combination of dataset, model and AL algorithm.\\
-We focus our work on single-sample pool-based AL where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
-Specifically, we are not experimenting on so-called Batch Active Learning, where at each iteration multiple unlabeled samples are chosen at the same time.
-Even though most proposed algorithms (and benchmarks) only do Batch AL, it does not have a principled advantage over single-sample AL except speed of computation.
-The problem of optimizing a portfolio of unlabeled samples more complicated to solve and the algorithms also have systematically less information per sample to work with.
-%For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
-A performance comparison of batch AL and single-sample AL can be found in Fig. \ref{fig:ablation_bald}.
-We can see that for two established algorithms BALD \cite{krizhevsky2009learning} and BADGE \cite{ashdeep} Batch AL can at most perform on-par with single-sample AL. \\
 Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature.
-The table also includes methodological AL papers with experiments on at least two data domains.
 \begin{table}
@@
 	\item Evaluation of Active Learning algorithms on datasets from 3 different domains, including synthetic data that highlights principled shortcomings of existing approaches.
-	\item Novel experimental protocol for seeding the experiment with 3 different seeds to allow full control, reproducibility and analysis of how many restarts are required to converge to the true median performance reliably.
+	\item Novel experimental protocol for controlling AL experiments to allow full reproducibility.
 	\item Simple algorithm for an Oracle that can be constructed greedily and does not rely on search.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Problem Description}
+\section{Problem Description}
 Given two spaces $\mathcal{X}, \mathcal{Y}$, $n=l+u$ data points with $l \in \mathbb{N}$ labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathbb{R}^\mathcal{X} \to \mathbb{R}^\mathcal{Y}$ that can label $x$.
 We call $x \in \mathcal{X}$, $y \in \mathcal{Y}$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$. 
-Find an acquisition function $a: \U^{(i)},\LL^{(i)} \mapsto x^{(i)}$ that iteratively selects the next unlabeled point $x^{(i)} \in \U^{(i)}$ for labeling
+Find an acquisition function $\Omega: \U^{(i)},\LL^{(i)} \mapsto x^{(i)}$ that iteratively selects the next unlabeled point $x^{(i)} \in \U^{(i)}$ for labeling
 \begin{align*}
@@
 	\centering
-	\includegraphics[width=0.49\linewidth]{img/ablation_bald}
-	\includegraphics[width=0.49\linewidth]{img/ablation_bald}
-	\caption{Shows a reproduction of the performance analysis of BALD from the original authors \cite{gal2017deep} in comparison to entropy sampling and an adaptation of BALD for batch AL \cite{kirsch2019batchbald} with a batch size of 40.}
+	\includegraphics[width=0.49\linewidth]{img/ablation_splice_bald}
+	\includegraphics[width=0.49\linewidth]{img/ablation_splice_badge}
+	\caption{Performance comparison of BALD \cite{gal2017deep} with it's batch AL adaptation \cite{kirsch2019batchbald} and BADGE \cite{ashdeep} with it's single-sample AL adaptation ($k=1$ in \cite{ashdeep} Alg.
+1).
+	In either scenario, the batched algorithm can at most perform on-par with it's single-sample analog, independent of wether it was originally designed for batch AL.}
 	\label{fig:ablation_bald}
 \end{figure}
-%Many different algorithms have been proposed for Active Learning. 
-%In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
-%AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
-%Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
-%They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
-%This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
-%Uncertainty-based approaches use metrics to measure the classifiers state.
-%Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
-%This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
-%
 Multiple benchmark suites have already been proposed for Active Learning:
@@
 AL algorithms have to perform well in both scenarios. 
-Since AL performance has high variance and is prone to outliers, we propose to aggregate AUC values with the median instead of mean. \\ [1mm]
+Since AL performance inhibits high variance and is prone to outliers, we propose to aggregate AUC values with their median instead of mean. \\ [1mm]
 Since the AUC depends on the chosen budget, we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
 In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
-Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets
+Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use-cases need different percentages to produce sensible budgets
 \footnote{Since we only consider single-sample AL, increasing the budget comes at a high computational cost. Therefore we could not freely increase the budget above 2000.}.
-We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
+We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\%, \textbf{Text}: 95\%, \textbf{Embedded Data}: 90\%. \\ [1mm]
 Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
-To create Fig. \ref{fig:restarts} we used 50 restarts from the margin/random sampling algorithm on the splice dataset.
+To create Fig. \ref{fig:restarts} we used 50 restarts from the margin/random acquisition function on the Splice dataset.
 From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
@@
 To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repetitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
-One way to reduce the number of necessary repetitions is to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
+One way to reduce the number of necessary repetitions would be to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
 We ultimately decided in favor of high variance and high number of repetitions as the high variance accurately reflects real world applications of AL.
-Fixing for example the classifier initialization to random draws from a constant seed does not have a corresponding use-case in the real world.
+%Fixing for example the classifier initialization to random draws from a constant seed does not have a corresponding use-case in the real world.
 
@@
 \subsection{Reproducibility}\label{sec:reproducibility}
-A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
-For a fair comparison of two AL algorithms, both algorithms should receive equal starting conditions in terms of train / validation split, initialization of classifier and even the state of minor systems like the optimizer or mini-batch sampler.
-Additionally, previous works have noted a significant adverse effect of training stochasticity on the performance of AL algorithms (\cite{zhou2021towards, lowell2018practical}).
-Even though implementations might have their own solution to some of these problems, to the best of our knowledge no previous work has discussed this topic in detail.
-At the core of this problem is the seeding utility in PyTorch, Tensorflow and other frameworks, whose default choice is a single global seed.
+Previous works have noted adverse effects of training stochasticity on the evaluation of AL algorithms (\cite{zhou2021towards, lowell2018practical}).
+Both papers observed that an actively sampled labeled set does not generalize well between model architectures or even different initializations of the same model.
+To cater for this we aim to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
+For a fair comparison of two AL algorithms, both algorithms need to receive equal starting conditions in terms of train / validation split, initialization of classifier and even the state of minor systems like the optimizer or mini-batch sampler.
+Even though different implementations might have their own solution to some of these problems, to the best of our knowledge no previous work has discussed this topic in detail.
+The main obstacle for ensuring reproducibility is the seeding utility in PyTorch, Tensorflow and other frameworks, whose default choice is a single global seed.
 Since many systems draw random numbers from this seed, all of them influence each other to a point where a single additional draw can completely change the model initialization or data split, depending on the ordering of these two in the implementation.
-Even though you could find workarounds like re-setting the seed multiple times, this problem also extends to every AL iteration and the systems within.
+Even though some workarounds exists, like re-setting the seed multiple times, this problem is not limited to the initialization phase, but also extends to the AL iterations and the systems within.
 We propose an implementation that creates a separate Random Number Generator (RNG) for each of these systems to ensure equal testing conditions even when the AL algorithm, dataset or classifier changes.
@@
 \subsection{Oracle}\label{sec:oracle}
-Posing Active Learning as a combinatorial problem, the oracle set $\mathcal{U}_b$ for a given dataset, model and training procedure would be the set that induces the highest AUC score for a given budget.
+\begin{minipage}{0.5\linewidth}
+	Posing Active Learning as a combinatorial problem, the oracle set $\mathcal{U}_b$ for a given dataset, model and training procedure is the set that induces the highest AUC score for a given budget.
 	However, since this problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence. 
-\cite{zhou2021towards} has used simulated annealing to search for the optimal sequence and used the best  solution found after a fixed time budget. 
+	\cite{zhou2021towards} has used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
 	Even though their reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
-In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
-Our oracle simply evaluates every data point $u_k = \operatorname{unif(\mathcal{U} | k = 1 \ldots \tau)}$ in a subsample of unlabeled points by fitting the classifier on $\mathcal{L}^{(i)} \cup u_k$ and directly measuring the resulting test performance.
-The point with the best test performance is selected and added to the labeled pool for that iteration.
-We noticed that this oracle is overfitting on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
-This can happen for example, if the oracle picked a labeled set that enables the classifier to correctly classifier a big portion of easy samples in the test set, but now fails to find the next single unlabeled point that would enable the classifier to succeed on one of the hard samples in the test set.
-This leads to a situation, where the selected data point is basically random.\\
-\begin{minipage}{0.5\linewidth}
-
-	To circumvent this problem, we introduced margin sampling \cite{wang2014new} as a fallback option for the oracle.
-	Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
-	The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.
-	In the algorithm $\operatorname{Retrain}(\mathcal{L}^{(i)}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns the accuracy on the test set $\mathcal{D}_\text{test}$. \\ 
-	$y_t$ is shorthand for the corresponding label of $u^{(i)}_{t}$ that can be recovered from the dataset labels. \\
-	When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$). \\
-	Alg. \ref{alg:oracle} replaces the algorithm in the AL process.
+	In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
 \end{minipage}
@@
 			\State $r^* \gets 0$
-			\For{$t := 1 \ldots \tau$}
-			\State $\mathcal{L}' \gets \mathcal{L}^{(i)} \cup \{(\mathcal{U}_t, y_t)\}$
+			\For{$k := 1 \ldots \tau$}
+			\State $\mathcal{L}' \gets \mathcal{L}^{(i)} \cup \{(u_k, A(u_k))\}$
 			\State $\text{acc}' \gets \operatorname{Train}(\mathcal{L}', \mathcal{D}_\text{test}, \hat y_\theta)$  
@@
 			\State $r^* \gets r$
-			\State $u^* \gets \mathcal{U}_t$
+			\State $u^* \gets u_k$
 			\EndIf
@@
 	\end{algorithm}
-\end{minipage}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implementation Details}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Available Information}
-At each iteration $i$ the AL algorithm needs to pick an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \theta^{(i)}, \text{opt}_\theta\}$, where $\theta^{(i)}$ is the current classifier and $\text{opt}_\theta$ is the optimizer used to fit $\theta^{(i)}$.
-We allow algorithms to derive additional information of this set like predictions of the classifier, K-Means clustering or even training new classifiers.
-However, the algorithm may not incorporate external information like other datasets, queries to recover additional labels, additional training steps for the classifier, or the test/validation set.
+\end{minipage} \\ [1mm]
+Our oracle simply evaluates every data point $u_k = \operatorname{unif(\mathcal{U} | k = 1 \ldots \tau)}$ in a subsample of unlabeled points by fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup u_k$ and directly measuring the resulting test performance.
+The data point with the best test performance is selected and added to the labeled pool for that iteration.
+We noticed that this oracle is overfitting on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
+This can happen for example, if the oracle picked a labeled set that enables the classifier to correctly classifier a big portion of easy samples in the test set, but now fails to find the next single unlabeled point that would enable the classifier to succeed on one of the hard samples in the test set.
+This leads to a situation, where the selected data point is basically random.\\
+To circumvent this problem, we introduced margin sampling \cite{wang2014new} as a fallback option for the oracle.
+Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
+The resulting greedy algorithm constructs an approximation of the optimal labeled set that consistently outperforms all other algorithms by a significant margin, while requiring relatively low computational cost (scaling linearly in $\tau$).
+The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.
+In the algorithm $\operatorname{Train}(\LL, \D_\text{test}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns it's accuracy  on $\D_\text{test}$. \\
+Alg. \ref{alg:oracle} replaces the acquisition function in the AL process.
 
@@
 We selected AL algorithms that have good performances reported by multiple different sources.
-To ensure a fair comparison we fixed the training process of our classification model as well as the set of available information for the algorithm and selected only those that can work under these restrictions:\\
+To ensure a fair comparison, we fixed the training process of our classification model as well as the set of available information for the algorithms and selected only those that can work under these restrictions:\\
 \textbf{Uncertainty Sampling} 
-Tries to find the sample that the classifier is most uncertain about. For our benchmark we use entropy and margin (a.k.a. best-vs-second-best) sampling.\\
+Tries to find the sample that the classifier is most uncertain about by computing heuristics of the class probabilities.
+For our benchmark we use entropy and margin (a.k.a.
+best-vs-second-best) sampling.\\
 \textbf{BALD \cite{kirsch2019batchbald}}
-Applies the query-by-committee strategy of model ensembles to a single model by interpreting the model's parameters as distributions and then sample multiple outputs from them via Monte-Carlo dropout.\\
+Applies the query-by-committee strategy of model ensembles to a single model by interpreting the classifier's parameters as distributions and then sample multiple outputs from them via Monte-Carlo dropout.\\
 \textbf{BADGE \cite{ashdeep}} Uses gradient embeddings of unlabeled points to select samples where the classifier is expected to change a lot.
@@
 Selects the unlabeled sample that is the furthest away from all cluster centers.
-Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\theta_i$.
+Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\hat y$.
 In this work we use the greedy variant of Coreset.\\
 \textbf{TypiClust \cite{hacohen2022active}}
-Relies on clustering similar to Coreset but proposes a new measure called "Typicality" to select unlabeled centers.
-Tries to select points that are in the densest regions of clusters that do not contain labeled samples yet.
-Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\theta_i$.
+Relies on clustering similar to Coreset but proposes a new measure called "Typicality" to select unlabeled samples.
+Selects points that are in the densest regions of clusters that do not contain labeled samples yet.
+Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\hat y$.
 It has to be pointed out that TypiClust was designed for low-budget scenarios, but we think it is still worthwhile to test and compare this algorithm with practically relevant budgets.
 %
-\subsubsection{Honorable Mentions}
+\subsubsection*{Honorable Mentions}
 \textbf{Learning Loss for AL}
@@
 For every dataset the chosen architecture's hyperparameters are optimized by to archive maximum upper bound performance.
-One desired characteristic of small classifiers is that the ranking of AL algorithms should stay the same when switching to larger models.
+One desired characteristic of these small classifiers is that the ranking of AL algorithms should stay the same when switching to larger models.
 A small analysis of this behavior can be found in Appendix \ref{app:classifier_size}.
-We found that the ranking of AL algorithms unfortunately does change, but we did not observe systematics that benefit one or few specific algorithms.
+We found that the ranking of AL algorithms unfortunately does change slightly, but we did not observe systematics that benefit one or few specific algorithms.
 We therefore rely on the different data domains to provide classification models of different sizes and archetypes to cover all of the use-cases.
 For an overview of architectures and hyperparameters please refer to Appendix \ref{app:hyperparameters}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Implementation Details}
+At each iteration $i$ the AL algorithm picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \theta^{(i)}, \text{opt}_\theta\}$, where $\theta^{(i)}$ is the current classifier and $\text{opt}_\theta$ is the optimizer used to fit $\theta^{(i)}$.
+We allow algorithms to derive additional information of this set like predictions of the classifier, K-Means clustering or even training new classifiers.
+However, the algorithm may not incorporate external information like other datasets, queries to recover additional labels, additional training steps for the classifier, or the test/validation set.
 
@@
 
+% Missing:
+% Seed set
+
 \section{Discussion}
@@
 \begin{align*}
-	a: (\X\times \Y)^* \times \X^* &\rightarrow \N
+	\Omega: (\X\times \Y)^* \times \X^* &\rightarrow \N
 	\quad\quad \text{\tiny (with $a(\D,X) \leq |X|$)} \\

### 2023-06-04 doc/al_benchmark_paper.tex
paper update

@@
 \begin{enumerate}
-	\item Evaluation of Active Learning algorithms on datasets from 3 different domains, including synthetic data that highlights principled shortcomings of existing approaches.
-	\item Novel experimental protocol for controlling AL experiments to allow full reproducibility.
-	\item Simple algorithm for an Oracle that can be constructed greedily and does not rely on search.
+	\item Evaluation of Active Learning algorithms on datasets from 3 different domains
+	\item Two novel synthetic dataset that highlight principled shortcomings of existing AL algorithms
+	\item Novel experimental protocol for controlling AL experiments to allow full reproducibility
+	\item Simple algorithm for an Oracle that can be constructed greedily and does not rely on search
 \end{enumerate}

### 2023-06-05 doc/al_benchmark_paper.tex
paper update

@@
 	\cite{zhou2021towards} has used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
-	Even though their reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
+	Even though their reported performance curves display a significant lift over all other acquisition functions, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
 	In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implementation Details}
-At each iteration $i$ the AL algorithm picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \theta^{(i)}, \text{opt}_\theta\}$, where $\theta^{(i)}$ is the current classifier and $\text{opt}_\theta$ is the optimizer used to fit $\theta^{(i)}$.
-We allow algorithms to derive additional information of this set like predictions of the classifier, K-Means clustering or even training new classifiers.
-However, the algorithm may not incorporate external information like other datasets, queries to recover additional labels, additional training steps for the classifier, or the test/validation set.
-
+\section{Experiments}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Training the Classifier}\label{sec:training_the_classifier}
+\subsection{Implementation Details}
+At each iteration $i$ the AL algorithm picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \hat y^{(i)}, \text{opt}_{\hat y}\}$, where $\hat y^{(i)}$ is the current classifier and $\text{opt}_{\hat y}$ is the optimizer used to fit $\hat y^{(i)}$.
+We allow acquisition functions to derive additional information of this set like predictions of the classifier, K-Means clustering or even training new classifiers.
+However, the algorithm may not incorporate external information like other datasets, queries to recover additional labels, additional training steps for $\hat y$, or the test/validation set. \\ [1mm]
 The classification model can be trained in two ways.
@@
 In this work we use the validation set to optimize the hyperparameters of the classifier and reduce overfitting with early stopping the training process in every iteration.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiments}
 
@@
 The text datasets are an exception, as they are only offered in their encoded form.
-The pre-encoded datasets enable us to test our single-sample algorithms on more complex datasets like Cifar10 and FashionMnist without the need of sampling $>2000$ datapoints before we can reach our upper bound performance.
+The pre-encoded datasets enable us to test our single-sample algorithms on more complex datasets like Cifar10 and FashionMnist.
 The embedding model was trained with the SimCLR \cite{chen2020simple} algorithm. 
@@
 The quality of embeddings during pretext training was measured after each epoch by attaching a linear classification head and evaluating this classifier for test accuracy, mirroring our AL setup for embedded datasets.
+\\ [1mm]
+Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being raw FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.
 
@@
 
-%\begin{table}[]
+\begin{table}[]
+	\hspace{-5mm}
+	\begin{tabular}{l|lllll|c}
+		& Vector       & Vector Enc. & Image        & Image Enc. & Text         & All          \\
+		\hline
+		Oracle          & 1.07 $\pm$ 0.02 & 1.13 $\pm$ 0.01    & 1.07 $\pm$ 0.04 & 1.16 $\pm$ 0.02   & 1.16 $\pm$ 0.07 & 1.12 $\pm$ 0.05 \\
+		Margin     & 1.03 $\pm$ 0.02 & 1.03 $\pm$ 0.03    & 1.01 $\pm$ 0.01 & 1.07 $\pm$ 0.00   & 1.04 $\pm$ 0.00 & 1.04 $\pm$ 0.02 \\
+		Badge           & 1.02 $\pm$ 0.01 & 0.98 $\pm$ 0.04    & 1.01 $\pm$ 0.00 & 1.07 $\pm$ 0.00   & 1.04 $\pm$ 0.01 & 1.02 $\pm$ 0.04 \\
+		Entropy  & 1.03 $\pm$ 0.02 & 0.99 $\pm$ 0.01    & 1.00 $\pm$ 0.01 & 1.07 $\pm$ 0.00   & 1.02 $\pm$ 0.00 & 1.02 $\pm$ 0.03 \\
+		Coreset & 1.02 $\pm$ 0.01 & 0.99 $\pm$ 0.04    & 1.00 $\pm$ 0.00 & 1.07 $\pm$ 0.00   & 1.03 $\pm$ 0.02 & 1.02 $\pm$ 0.04 \\
+		TypiClust       & 1.03 $\pm$ 0.03 & 1.05 $\pm$ 0.04    & 0.99 $\pm$ 0.02 & 1.03 $\pm$ 0.01   & 0.98 $\pm$ 0.04 & 1.02 $\pm$ 0.04 \\
+		Random     & 1.00 $\pm$ 0.00 & 1.00 $\pm$ 0.00    & 1.00 $\pm$ 0.00 & 1.00 $\pm$ 0.00   & 1.00 $\pm$ 0.00 & 1.00 $\pm$ 0.00 \\
+		BALD            & 0.98 $\pm$ 0.03 & 0.90 $\pm$ 0.04    & 0.99 $\pm$ 0.03 & 1.06 $\pm$ 0.01   & 1.04 $\pm$ 0.01 & 0.98 $\pm$ 0.07
+	\end{tabular}
+	\caption{Performances for each algorithm per domain normalized by the performance of random sampling. Higher is better, scores below 1 indicate worse-than-random performance.}
+	\label{tab:results}
+\end{table}
+%
+%\begin{figure}
 %	\centering
-%	\begin{tabular}{l|lll}
-%		& Splice        & DNA           & USPS          \\
-%		\hline
-%		Oracle          & 0.830 +- 0.01 & 0.836 +- 0.02 & 0.823 +- 0.01 \\
-%		SAL & 0.799 +- 0.01 & 0.797 +- 0.03 & 0.809 +- 0.01 \\
-%		Coreset & 0.800 +- 0.01 & 0.795 +- 0.03 & 0.787 +- 0.02 \\
-%		TypiClust       & 0.790 +- 0.01 & 0.771 +- 0.04 & 0.761 +- 0.02 \\
-%		MarginScore     & 0.797 +- 0.02 & 0.795 +- 0.04 & 0.808 +- 0.01 \\
-%		ShannonEntropy  & 0.799 +- 0.02 & 0.794 +- 0.04 & 0.807 +- 0.01 \\
-%		RandomAgent     & 0.788 +- 0.01 & 0.765 +- 0.03 & 0.772 +- 0.01 \\
-%		Badge           & 0.807 +- 0.01 & 0.769 +- 0.06 & 0.797 +- 0.02 \\
-%		BALD            & 0.811 +- 0.01 & 0.743 +- 0.04 & 0.717 +- 0.05
-%	\end{tabular}
-%	\caption{Median AUC values and standard deviation for all algorithms on the tabular datasets. Higher is better.}
-%\end{table}
-%
-\begin{figure}
-	\centering
-	\includegraphics[width=0.49\linewidth]{img/eval_splice}
-	\hspace{1mm}
-	\includegraphics[width=0.49\linewidth]{img/eval_dna}\\[2mm]
-	\includegraphics[width=0.49\linewidth]{img/micro_splice}
-	\hspace{1mm}
-	\includegraphics[width=0.49\linewidth]{img/micro_dna}
-	\caption{Results for all algorithms on Splice and DNA, both from the tabular domain. Even within one domain, the performance of the same algorithm can vary drastically.}
-	\label{fig:eval_vector}
-\end{figure}
+%	\includegraphics[width=0.49\linewidth]{img/eval_splice}
+%	\hspace{1mm}
+%	\includegraphics[width=0.49\linewidth]{img/eval_dna}\\[2mm]
+%	\includegraphics[width=0.49\linewidth]{img/micro_splice}
+%	\hspace{1mm}
+%	\includegraphics[width=0.49\linewidth]{img/micro_dna}
+%	\caption{Results for all algorithms on Splice and DNA, both from the tabular domain. Even within one domain, the performance of the same algorithm can vary drastically.}
+%	\label{fig:eval_vector}
+%\end{figure}
 %
@@
 %
-From Fig. \ref{fig:eval_vector} we notice drastically different qualities for the same AL algorithm for different datasets.
-We would like to highlight that both datasets are tabular from the medical domain with similar number of features and classes, yet we see that i.e.
-BALD is the best algorithm for Splice and the worst algorithm for DNA.
-These inconsistencies are present between the datasets of all our tested domains, further highlighting the difficulties for comparing AL algorithms in terms of average performance.
-In order to provide a meaningful analysis of which algorithm can be expected to perform best on average we ranked the algorithm for each dataset based on their median AUC and displayed these rankings in critical difference diagrams \cite{IsmailFawaz2018deep}.
-In Fig. \ref{fig:cd_diagrams} we report the rankings split by domain as well as across all domains (excluding the toydata). \\ [2mm]
-%
+We report our results in two different fashions: Table \ref{tab:results} contains mean AUC performances per data domain.
+To make performances from different datasets comparable we normalized them with the random performance of each dataset.
+Figure \ref{fig:cd_diagrams} aggregates the performances per domain by counting wins and deriving a ranking of algorithms.
+This is closely related to the evaluation protocol of \cite{ashdeep}, who displayed a pairwise penalty matrix for all algorithms. \\
+Based these two evaluations we make the following observations:\\
+(i) The ranking of both reports does not match exactly.
+This is due to Table \ref{tab:results} being more sensitive to outlier performances (i.e. BALD being worse than random due to it's bad performance on embedded vector data). \\
+(ii) Both evaluations make it clear that no one-fits-all algorithm for all domains seems to exist, with only margin sampling being significantly better that random in Tab.
+\ref{tab:results} and the best 3 ranks for all domains in Fig.
+\ref{fig:cd_diagrams} being close together.
+\\
+(iii) Even though in Fig.
+\ref{fig:cd_diagrams} some domains show evidence of one algorithm being systematically better than the rest (i.e.
+Image), the actual differences in Tab.
+\ref{tab:results} seem to be negligible.
+\\
+(iv) Not only do different domains favor different AL algorithms, we noticed that even within the same domain datasets can induce drastically different behaviors (Compare Appendix \ref{app:all_results} Splice/DNA/USPS).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+Based on the findings of the previous section a couple of things deserve to be highlighted:
+(i) None of the algorithms display performance values that could be thought of a reproduction of the reported values from their papers.
 BALD performs bad with linear classifiers since they are trained without dropout and cannot cope well with missing inputs. \\
 TypiClust is better with embedded data not only due to lower budgets. On other datasets it is not able to outperform other algorithms in early stages
-
-% Missing:
-% Seed set
-
-\section{Discussion}
-Domains are very different \\
-Even within one domain we have stark differences (Fig. \ref{fig:eval_vector}) \\
-Best 3 ranks for all domains are 3, 3.1 and 3.4 ... No clear winner \\
-At least on average everything is better that random \\
 BADGE: "Separately, we notice that diversity sampling only seems to work well when either the model has good architectural priors (inductive biases) built in, or when the data are easy to learn.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Limitations and Future Work}

### 2023-06-05 results/generate_result_tables.py

@@ -3,6 +3,20 @@ from os.path import exists, join
 import pandas as pd
 import numpy as np
 
+domains = ["Vector", "Vector Embedded", "Image", "Image Embedded", "Text"]
+agents = ["RandomAgent", "Oracle", "Coreset_Greedy", "TypiClust", "MarginScore", "ShannonEntropy", "Badge", "BALD", ]
+agents_no_random = agents[1:]
+vector_data = ["Splice", "DNA", "USPS"]
+enc_vector_data = ["SpliceEncoded", "DNAEncoded", "USPSEncoded"]
+img_data = ["Cifar10", "FashionMnist"]
+enc_img_data = ["Cifar10Encoded", "FashionMnistEncoded"]
+text_data = ["TopV2", "News"]
+
+def get_auc(acc_file):
+    accuracies = pd.read_csv(acc_file, header=0, index_col=0)
+    values = accuracies.values
+    return np.sum(values, axis=0) / values.shape[0]
+
 def generate_table(agents, datasets, out_file):
     result_table = pd.DataFrame(index=agents, columns=datasets)
     result_table["Sum"] = [0.0]*len(agents)
@@ -19,9 +33,7 @@ def generate_table(agents, datasets, out_file):
             if exists(agent_folder):
                 acc_file = join(agent_folder, "accuracies.csv")
                 if exists(acc_file):
-                    accuracies = pd.read_csv(acc_file, header=0, index_col=0)
-                    values = accuracies.values
-                    auc = np.sum(values, axis=0) / values.shape[0]
+                    auc = get_auc(acc_file)
                     result_table.loc[agent, "Sum"] += np.median(auc).item()
                     result_table.loc[agent, dataset] = "%1.3f +- %1.2f"%(np.median(auc).item(), np.std(auc).item())
             else:
@@ -37,13 +49,60 @@ def generate_table(agents, datasets, out_file):
         result_table.to_csv(result_table_file)
 
 
-agents = ["Oracle", "Coreset_Greedy", "TypiClust", "MarginScore", "ShannonEntropy", "RandomAgent", "Badge", "BALD", ]
-vector_data = ["Splice", "DNA", "USPS"]
-enc_vector_data = ["SpliceEncoded", "DNAEncoded", "USPSEncoded"]
-img_data = ["Cifar10Encoded", "FashionMnistEncoded"]
-text_data = ["TopV2", "News"]
+def new_list():
+    return list()
 
[…]

### 2023-06-05 doc/al_benchmark_paper.tex
paper update

@@
 \section{Discussion}
-Based on the findings of the previous section a couple of things deserve to be highlighted:
-(i) None of the algorithms display performance values that could be thought of a reproduction of the reported values from their papers.
-BALD performs bad with linear classifiers since they are trained without dropout and cannot cope well with missing inputs. \\
-TypiClust is better with embedded data not only due to lower budgets. On other datasets it is not able to outperform other algorithms in early stages
-BADGE: "Separately, we notice that diversity sampling only seems to work well when either the model has good architectural priors (inductive biases) built in, or when the data are easy to learn.
-Otherwise, penultimate layer representations are not meaningful, and diverse sampling can be deleterious.
-For this reason, CORESET often performs worse than random on sufficiently complex data when not using a convolutional network (Figure 3b)" \\
+Our results empirically reinforce what the authors of \cite{zhou2021towards} have noticed in their meta-analysis already:
+When slightly altering the dataset, model or training conditions, most AL algorithms are reduced to uncertainty sampling or worse.
+Additionally, some results remain non-reproducible even on the same dataset and model combination (Compare Appendix \ref{app:all_results} Cifar10 with \cite{hacohen2022active} or \cite{sener2017active}).
+On the other hand we were able to also confirm some previous results i.e. for the NewsCategory dataset (Compare Appendix \ref{app:all_results} News with \cite{hu2021towards} Fig 3g). \\ [1mm]
+Regarding meta analysis of our results we have the following conclusions:\\
+(i) BALD seems to be performing especially bad when the input dimensionality is low and the applied classifier is too simple. (i.e. (Encoded) Vector). \\
+(ii) TypiClust seemingly holds its claim to work well with extremely low budgets (Encoded Vector $B=\{50,60,600\}$), but fails to push the classifier close to upper bound performance even with moderate budgets in other domains.
+\\
+(iii) The authors of \cite{ashdeep} noted that diversity sampling and adjacent methods, like Coreset and TypiClust, only work well when the classifier has correct inductive biases (correct architecture, tuned hyperparameters).
+Since our architectures are extensively tuned, the performance of these algorithms might be slightly inflated.
 
@@
 \section{Limitations and Future Work}
-No batch AL \\
-No learned algorithms \\
-No SOTA classifier training (data augmentation, semi-supervised, etc.) \\
-SimCRL as Pretext task works better for images
+In the current state, a couple of aspects/use-cases are missing from our proposed benchmark and are considered future work: \\
+(i) We elected to not provide any experiments on batch AL, but the provided implementation is already capable of doing that and we plan to extend the benchmark to include batch sizes larger than 1 in our experiments.
+\\
+(ii) Due to difficulties with reproducing results and high computational setup costs, we did not include learned acquisition functions like \cite{shao2019learning}, \cite{vu2019learning} or \cite{konyushkova2017learning} \\
+(iii) Our training process for the classifier does not include techniques that would be present in real-world scenarios, like data augmentation, semi-supervised learning, or auxiliary tasks.
+We opted for simpler training process to be able to extensively tune those fewer hyperparameters and have full control over the training.\\
+(iv) We used SimCLR as pretext task for our vector datasets, which primarily relies on data augmentation.
+While images have a plethora of well-defined augmentation techniques, we had to rely on Gaussian noise to augment the vector data.
+Other pretext tasks, like autoencoders have potential to provide better encodings of the data.
 

### 2023-06-08 doc/al_benchmark_paper.tex
paper update

@@
 \footnote{out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not claim a new SOTA performance in their result section}
+\let\thefootnote\relax\footnotetext{Code available at: \url{https://github.com/wernerth94/comparable-active-learning}}
 , AL research faces two central difficulties:
@@
 		\begin{algorithmic}[1]
-			\Require $\mathcal{U}, \mathcal{L}, \mathcal{Y}, \mathcal{D}_\text{test}$ Train, Margin, $\tau, \hat y_\theta$ 
+			\Require $\mathcal{U}, \mathcal{L}, A, \mathcal{D}_\text{test}$ Train, Margin, $\tau, \hat y_\theta$ 
 			\State $\text{acc} \gets \operatorname{Train}(\mathcal{L}, \mathcal{D}_\text{test}, \hat y_\theta)$ 
@@
 	\centering
-	\includegraphics[width=0.49\linewidth]{img/eval_splice}
-	\includegraphics[width=0.49\linewidth]{img/eval_splice_enc} \\ [2mm]
-	\includegraphics[width=0.49\linewidth]{img/micro_splice.jpg}
-	\includegraphics[width=0.49\linewidth]{img/micro_splice_enc.jpg} \\ [4mm]
+	\includegraphics[width=0.4\linewidth]{img/eval_splice}
+	\includegraphics[width=0.4\linewidth]{img/eval_splice_enc} \\ [2mm]
+	\includegraphics[width=0.3\linewidth]{img/micro_splice.jpg}
+	\includegraphics[width=0.3\linewidth]{img/micro_splice_enc.jpg} \\ [4mm]
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.49\linewidth]{img/eval_dna}
-	\includegraphics[width=0.49\linewidth]{img/eval_dna_enc} \\ [2mm]
-	\includegraphics[width=0.49\linewidth]{img/micro_dna.jpg}
-	\includegraphics[width=0.49\linewidth]{img/micro_dna_enc.jpg} \\ [4mm]
+	\includegraphics[width=0.4\linewidth]{img/eval_dna}
+	\includegraphics[width=0.4\linewidth]{img/eval_dna_enc} \\ [2mm]
+	\includegraphics[width=0.3\linewidth]{img/micro_dna.jpg}
+	\includegraphics[width=0.3\linewidth]{img/micro_dna_enc.jpg} \\ [4mm]
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.49\linewidth]{img/eval_usps}
-	\includegraphics[width=0.49\linewidth]{img/eval_usps_enc} \\ [2mm]
-	\includegraphics[width=0.49\linewidth]{img/micro_usps.jpg}
-	\includegraphics[width=0.49\linewidth]{img/micro_usps_enc.jpg} \\ [4mm]
+	\includegraphics[width=0.4\linewidth]{img/eval_usps}
+	\includegraphics[width=0.4\linewidth]{img/eval_usps_enc} \\ [2mm]
+	\includegraphics[width=0.3\linewidth]{img/micro_usps.jpg}
+	\includegraphics[width=0.3\linewidth]{img/micro_usps_enc.jpg} \\ [4mm]
 \end{figure}
@@
 \centering
-\includegraphics[width=0.49\linewidth]{img/eval_cifar10}
-\includegraphics[width=0.49\linewidth]{img/eval_cifar10_enc} \\ [2mm]
-\includegraphics[width=0.49\linewidth]{img/micro_cifar10.jpg}
-\includegraphics[width=0.49\linewidth]{img/micro_cifar10_enc.jpg} \\ [4mm]
+\includegraphics[width=0.4\linewidth]{img/eval_cifar10}
+\includegraphics[width=0.4\linewidth]{img/eval_cifar10_enc} \\ [2mm]
+\includegraphics[width=0.3\linewidth]{img/micro_cifar10.jpg}
+\includegraphics[width=0.3\linewidth]{img/micro_cifar10_enc.jpg} \\ [4mm]
 \end{figure}
@@
 \centering
-\includegraphics[width=0.49\linewidth]{img/eval_fmnist}
-\includegraphics[width=0.49\linewidth]{img/eval_fmnist_enc} \\ [2mm]
-\includegraphics[width=0.49\linewidth]{img/micro_fmnist.jpg}
-\includegraphics[width=0.49\linewidth]{img/micro_fmnist_enc.jpg} \\ [4mm]
+\includegraphics[width=0.4\linewidth]{img/eval_fmnist}
+\includegraphics[width=0.4\linewidth]{img/eval_fmnist_enc} \\ [2mm]
+\includegraphics[width=0.3\linewidth]{img/micro_fmnist.jpg}
+\includegraphics[width=0.3\linewidth]{img/micro_fmnist_enc.jpg} \\ [4mm]
 \end{figure}
@@
 \centering
-\includegraphics[width=0.49\linewidth]{img/eval_topv2}
-\includegraphics[width=0.49\linewidth]{img/eval_news} \\ [2mm]
-\includegraphics[width=0.49\linewidth]{img/micro_topv2.jpg}
-\includegraphics[width=0.49\linewidth]{img/micro_news.jpg} \\ [4mm]
+\includegraphics[width=0.4\linewidth]{img/eval_topv2}
+\includegraphics[width=0.4\linewidth]{img/eval_news} \\ [2mm]
+\includegraphics[width=0.3\linewidth]{img/micro_topv2.jpg}
+\includegraphics[width=0.3\linewidth]{img/micro_news.jpg} \\ [4mm]
 \end{figure}
@@
 \begin{table}[H]
-	TODO
+	\centering
+	\begin{tabular}{l || l | l}
+		Algorithm & Sample Size & \\
+		\hline
+		BADGE & 100 & \\
+		BALD & 100 & Dropout Trials: 5 \\
+		Coreset & 8000 & \\
+		TypiClust & 10000 & Min Cluster Size: 5, Max \# Clusters: 500 \\
+		Margin & 8000 & \\
+		Entropy & 8000 & \\
+	\end{tabular}
 \end{table}
@@
 	\centering
-	\begin{tabular}{l || l | l l l l}
-		Dataset & Classifier & Optimizer & LR & Weight Decay & Dropout \\
+	\begin{tabular}{l || l l l }
+		Dataset & Seed Set & Budget & Val Split \\
 		\hline
-		Splice & [24, 12] & NAdam & 1.2e-3 & 5.9e-5 & 0 \\
-		SpliceEnc. & linear & NAdam & 6.2e-4 & 5.9e-6 & 0 \\
-		DNA & [24, 12] & NAdam & 3.9e-2 & 3.6e-5 & 0 \\
-		DNAEnc & linear & NAdam & 1.6e-3 & 4e-4 & 0 \\
-		USPS & [24, 12] & Adam & 8.1e-3 & 1.5e-6 & 0 \\
-		USPS & linear & NAdam & 7.8e-3 & 1.9e-6 & 0 \\
-		FashionMnist & ResNet18 & NAdam & 1e-3 & 0 & 0 \\
-		FashionMnistEnc & linear & Adam & 1.6e-3 & 1e-5 & 5e-2 \\
-		Cifar10 & ResNet18 & NAdam & 1e-3 & 0 & 0 \\
-		Cifar10Enc & linear & NAdam & 1.7e-3 & 2.3e-5 & 0 \\
-		TopV2 & BiLSTM & NAdam & 1.5e-3 & 1.7e-7 & 5e-2 \\
-		News & BiLSTM & NAdam & 1.5e-3 & 1.7e-7 & 5e-2 \\
+		Splice & 1 & 400 & 0.2 \\
+		SpliceEnc. & 1 & 60 & 0.2 \\
+		DNA & 1 & 300 & 0.2 \\
+		DNAEnc & 1 & 40 & 0.2 \\
+		USPS & 1 & 400 & 0.2 \\
+		USPSEnc & 1 & 600 & 0.2 \\
+		FashionMnist & 100 & 2000 & 0.04 \\
+		FashionMnistEnc & 1 & 500 & 0.04 \\
+		Cifar10 & 100 & 2000 & 0.04  \\
+		Cifar10Enc & 1 & 350 & 0.04  \\
+		TopV2 & 1 & 125 & 0.25  \\
+		News & 1 & 1500 & 0.03  \\
+	\end{tabular}
+	\caption{Size of the seed set is given by number of labeled sample per class.}
+	\label{tab:architectures}
+\end{table}
+\begin{table}[H]
+\centering
+\begin{tabular}{l || l | l l l l l}
+	Dataset & Classifier & Optimizer & LR & Weight Decay & Dropout & Batch Size \\
+	\hline
+	Splice & [24, 12] & NAdam & 1.2e-3 & 5.9e-5 & 0 & 43 \\
+	SpliceEnc. & linear & NAdam & 6.2e-4 & 5.9e-6 & 0 & 64 \\
+	DNA & [24, 12] & NAdam & 3.9e-2 & 3.6e-5 & 0 & 64 \\
+	DNAEnc & linear & NAdam & 1.6e-3 & 4e-4 & 0 & 64 \\
+	USPS & [24, 12] & Adam & 8.1e-3 & 1.5e-6 & 0 & 43 \\
+	USPSEnc & linear & NAdam & 7.8e-3 & 1.9e-6 & 0 & 64 \\
+	FashionMnist & ResNet18 & NAdam & 1e-3 & 0 & 0 & 64 \\
+	FashionMnistEnc & linear & Adam & 1.6e-3 & 1e-5 & 5e-2 & 64 \\
+	Cifar10 & ResNet18 & NAdam & 1e-3 & 0 & 0 & 64 \\
+	Cifar10Enc & linear & NAdam & 1.7e-3 & 2.3e-5 & 0 & 64 \\
+	TopV2 & BiLSTM & NAdam & 1.5e-3 & 1.7e-7 & 5e-2 & 64 \\
+	News & BiLSTM & NAdam & 1.5e-3 & 1.7e-7 & 5e-2 & 64 \\
 \end{tabular}
@@
 \end{itemize}
-\begin{figure}
+\begin{figure}[H]
 	\centering
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[H]
 	\centering
@@
 \end{figure}
-
-\section{Comparison of different sample sizes}
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.7\linewidth]{img/tau_ablation.png}
-\end{figure}
+Even though the ranking between the two sizes of classifier does change, we observed no systematic that would favor one algorithm. 
+We therefore rely on the different datasets to provide classifiers of varying size and architecture.
 
@@
 	\begin{algorithmic}[1]
-		\Require $\mathcal{U}$ \Comment{Unlabeled Pool}
-		\Require $\tau$ \Comment{Unlabeled Sample Size}
-		\Require $\Omega$ \Comment{AL Agent}
-		\Require $\omega$ \Comment{Environment State function}
-		\State $\mathcal{L}^{(1)} \gets \operatorname{seed}(\mathcal{U})$  \Comment{Create the initial labeled set}
-		\State $\mathcal{U}^{(1)} \gets \mathcal{U}$
+		\Require $\LL, \U, \D_\text{test}, \operatorname{Train}, \operatorname{Seed}, \hat y$
+		\Require $\Omega$ \Comment{Acquisition Function}
+		\State $\LL^{(1)} \gets \operatorname{Seed}(\U)$  \Comment{Create the initial labeled set}
+		\State $\U^{(1)} \gets \U$
 		\For{$i := 1 \ldots B$}
-		\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$  %\Comment{$\operatorname{Retrain}(\mathcal{L}^{(i)})$ is shorthand for $\operatorname{Retrain}(\mathcal{L}^{(i)}, \mathcal{L}^\text{test}, \hat y_\theta, e^\text{max})$}
-		%\State $u^{(i)} \sim \text{unif}(1:|\mathcal{U}^{(i)}|)$
-		\State $a^{(i)} \gets \Omega(\mathcal{U}^{(i)}) \hspace{1mm};\hspace{2mm} a \in 1:|\mathcal{U}^{(i)}|$ 
-		%\Comment{$a^{(i)}$ is an index inside of $u^{(i)}$}
-		\State $y^{(i)} \gets \operatorname{label}(\mathcal{U}^{(i)}_{a})$ 
-		%\Comment{$u^{(i)}_{a}$ is shorthand for $u^{(i)}_{a^{(i)}}$}
-		\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(\mathcal{U}^{(i)}_a, y^{(i)})\}$
-		\State $\mathcal{U}^{(i+1)} \gets \mathcal{U}^{(i)} \setminus \{\mathcal{U}^{(i)}_a\}$
+		\State $\text{acc}^{(i)} \gets \operatorname{Train}(\LL^{(i)})$ 
+		\State $a^{(i)} \gets \Omega(\mathcal{U}^{(i)})$ 
+		\State $\mathcal{L}^{(i+1)} \gets \mathcal{L}^{(i)} \cup \{(\U^{(i)}_a, A(\U^{(i)}_{a}))\}$
+		\State $\U^{(i+1)} \gets \U^{(i)} \setminus \{\U^{(i)}_a\}$
 		\EndFor
@@
 	\begin{algorithmic}[1]
-		\Require $\mathcal{L}$ \Comment{Labeled Pool}
-		\Require $\mathcal{D}_\text{val}$ \Comment{Validation Data}
-		\Require $\mathcal{D}_\text{test}$ \Comment{Test Data}
-		\Require $\hat y_\theta$ \Comment{Class. Model}
-		\Require $e^\text{max}$ \Comment{Maximum Epochs}
+		\Require $\LL, \D_\text{val}, \D_\text{test}$
+		\Require $\hat y, e_\text{max}$
 		\State $\text{loss}^* \gets \infty$
 		\For{$i := 1 \ldots e^{\text{max}}$}
-		\State $\theta_{i+1} \gets \theta_i - \eta \nabla_\theta \ell(\mathcal{L}, \hat y_{\theta})$
-		\State $\text{loss}_i \gets \ell(\mathcal{D}^\text{val}, \hat y_{\theta})$
+		\State $\hat y_{i+1} \gets \hat y_i - \eta \nabla_{\hat y} \ell(\mathcal{L}, \hat y)$
+		\State $\text{loss}_i \gets \ell(\mathcal{D}^\text{val}, \hat y)$
 		\If{$\text{loss}_i < \text{loss}^*$}
@@
 		\State
-		\Return Acc($\mathcal{D}^\text{test}, \hat y_{\theta}$)
+		\Return Acc($\mathcal{D}^\text{test}, \hat y$)
 	\end{algorithmic}

### 2023-06-08 core/helper_functions.py

@@ -174,5 +174,5 @@ def get_agent_by_name(name:str)->Union[Callable, BaseAgent]:
 
 
 if __name__ == '__main__':
-    base_path = "runs/Mnist/BatchRandomAgent"
+    base_path = "runs/Splice/BatchBALD_100"
     collect_results(base_path, "run_")

### 2023-06-08 results/generate_result_tables.py

@@ -35,9 +35,44 @@ def generate_table(agents, datasets, out_file):
                 if exists(acc_file):
                     auc = get_auc(acc_file)
                     result_table.loc[agent, "Sum"] += np.median(auc).item()
-                    result_table.loc[agent, dataset] = "%1.3f +- %1.2f"%(np.median(auc).item(), np.std(auc).item())
+                    result_table.loc[agent, dataset] = "%1.3f $\pm$ %1.2f"%(np.median(auc).item(), np.std(auc).item())
             else:
                 print(f"{agent} is missing for {dataset}")
+
+    if len(result_table) > 0:
+        result_table = result_table.sort_values("Sum", ascending=False)
+        result_table = result_table.drop(columns=["Sum"])
+
+        result_table_file = os.path.join("results", out_file)
+        if os.path.exists(result_table_file):
+            os.remove(result_table_file)
+        result_table.to_csv(result_table_file)
+
+def normalized_table(agents, dataset, out_file):
+    result_table = pd.DataFrame(index=agents, columns=[dataset])
+    result_table["Sum"] = [0.0]*len(agents)
+    dataset_folder = os.path.join("runs", dataset)
+    if not os.path.isdir(dataset_folder):
+        raise ValueError
+
+    acc_file = join(dataset_folder, "RandomAgent", "accuracies.csv")
+    random_auc = np.median(get_auc(acc_file)).item()
+
+    agent_aucs = dict()
+    for a in agents:
+        agent_aucs[a] = "-"
+
+    for agent in agents:
+        agent_folder = join(dataset_folder, agent)
+        if exists(agent_folder):
+            acc_file = join(agent_folder, "accuracies.csv")
+            if exists(acc_file):
+                auc = get_auc(acc_file)
+                auc /= random_auc
+                result_table.loc[agent, "Sum"] += np.median(auc).item()
+                result_table.loc[agent, dataset] = "%1.2f $\pm$ %1.2f"%(np.median(auc).item(), np.std(auc).item())
+        else:
+            print(f"{agent} is missing for {dataset}")
 
     if len(result_table) > 0:
         result_table = result_table.sort_values("Sum", ascending=False)
@@ -101,8 +136,27 @@ def generate_domain_table(agents, out_file):
 
 generate_domain_table(agents, "result_domains.csv")
 
[…]

### 2024-05-30 doc/al_benchmark_paper.tex
paper

@@
 %\usepackage[a4paper, total={6.5in, 8.5in}]{geometry}
-\usepackage{color}
+
+%\usepackage{PRIMEarxiv}
+
+\usepackage{xcolor}
 \usepackage{hyperref}
@@
 \usepackage[noend]{algpseudocode}
-\usepackage{neurips_data_2023}
+\usepackage[normalem]{ulem}
+\useunder{\uline}{\ul}{}
+\usepackage{makecell}
+\usepackage{neurips_data_2024}
+
+\graphicspath{/}
 
@@
 
+%\pagestyle{fancy}
+%\thispagestyle{empty}
+%\rhead{ \textit{ }} 
+%\fancyhead[LO]{Running Title for Header}
+
 \title{Towards Comparable Active Learning}
@@
 	Anonymous Authors
-	%Thorben Werner 
-	%\thanks{Institute of Computer Science - Information Systems and Machine Learning Lab (ISMLL)} 
-	%\\
-	%University of Hildesheim\\
-	%Universitätsplatz 1\, 31141 Hildesheim \\
-	%\texttt{werner@ismll.de} \\
-	% examples of more authors
-	%\And
-	%Johannes Burchert$^*$ \\
-	%University of Hildesheim\\
-	%Universitätsplatz 1, 31141 Hildesheim \\
-	%\texttt{burchert@ismll.de} \\
-	%\AND
-	%Prof. Lars Schmidt-Thieme$^*$ \\
-	%University of Hildesheim\\
-	%Universitätsplatz 1, 31141 Hildesheim \\
-	%\texttt{schmidt-thieme@ismll.uni-hildesheim.de}
+%	Thorben Werner 
+%	\thanks{Institute of Computer Science - Information Systems and Machine Learning Lab (ISMLL)} 
+%	\\
+%	University of Hildesheim\\
+%	Universitätsplatz 1\, 31141 Hildesheim \\
+%	\texttt{werner@ismll.de} \\
+%	\And
+%	Johannes Burchert$^*$ \\
+%	University of Hildesheim\\
+%	Universitätsplatz 1, 31141 Hildesheim \\
+%	\texttt{burchert@ismll.de} \\
+%	\And
+%	Maximilian Stubbemann$^*$ \\
+%	University of Hildesheim\\
+%	Universitätsplatz 1, 31141 Hildesheim \\
+%	\texttt{stubbemann@ismll.de} \\
+%	\AND
+%	Lars Schmidt-Thieme$^*$ \\
+%	University of Hildesheim\\
+%	Universitätsplatz 1, 31141 Hildesheim \\
+%	\texttt{schmidt-thieme@ismll.uni-hildesheim.de}
 }
@@
 
+% Review point
+% - Restarting process seems to be hard to understand (probably in terms of the seeding)
+% - Justification for datasets and algorithms
+
 \maketitle
@@
 \begin{abstract}
-	Active Learning has received significant attention in the field of machine learning for its potential in selecting the most informative samples for labeling, thereby reducing data annotation costs.
-However, the lack of reproducibility of performance gains reported in recent literature has created a chaotic landscape in Active Learning research.
-	This paper addresses these issues of inconsistent results in active learning (AL) literature.
-	%Previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results.
-This inconsistency leads to a chaotic landscape of AL algorithms.
-	To the best of our knowledge, we propose the first AL benchmark that tests algorithms in 3 major domains: Tabular, Image and Text.
-	Furthermore, we highlight overlooked problems for reproducing AL experiments that can lead to unfair comparisons and increased variance in the results.
-	To tackle these challenges, we propose an experimental protocol to accurately control the experiments.
-	We report empirical results for 6 widely used algorithms on 7 datasets and aggregate them into a domain-specific ranking of AL algorithms.
+Active Learning has received significant attention in the field of machine learning. Its goal is to select the most informative samples for labeling, thereby reducing data annotation costs. 
+However, it has been brought to attention multiple times that reported lifts from literature generalize poorly and display high variance, leading to an inconclusive landscape in Active Learning research.
+Based on recent insights for reliable evaluation for Active Learning, this work extends experimentation from the commonly used image domain to a wide spectrum of domains.
+%This work follows up on recent benchmarking suites for Active Learning, applying the proposed solutions and evaluation protocols to a wide spectrum of domains.
+Additionally, we provide an analysis of how many repetitions an Active Learning experiment needs in order to derive conclusive results and propose that previous benchmarks have not met the necessary number of repetitions.
+%evidence that existing benchmarks might not have repeated their experiments often enough.
+To the best of our knowledge, we propose the first AL benchmark that applies state-of-the-art evaluation on algorithms in 3 major domains: Tabular, Image, and Text as well as synthetic data.
+We report empirical results for 10 widely used algorithms on 7 real-world and 2 synthetic datasets and aggregate them into domain-specific and overall rankings of AL algorithms.
 \end{abstract}
@@
 Deep neural networks (NN) have produced state-of-the-art results on many important supervised learning tasks.
-Since Deep NNs usually require large amounts of labeled training data, Active Learning (AL) can be used to instead to select the most informative samples out of a large pool of unlabeled data, so that only these samples need to be labeled.
+Since Deep NNs usually require large amounts of labeled training data, Active Learning (AL) can be used to instead select the most informative samples out of a large pool of unlabeled data, so that only these samples need to be labeled.
 It has been shown that a small labeled set of this nature can be used to train well-performing models \cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}. \\
-%On top of providing a principled way to labeled unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
+On top of providing a principled way to label unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
 %Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
-In the last decade many different algorithms for AL have been proposed.
-Even though, almost every method has reported lifts all it's predecessors 
-\footnote{out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not claim a new SOTA performance in their result section}
-\let\thefootnote\relax\footnotetext{Code available at: \url{https://github.com/wernerth94/comparable-active-learning}}
-, AL research faces two central difficulties:
+In the last decade, many different algorithms for AL have been proposed.
+Even though, almost every method has reported lifts over all its predecessors,
+\footnote{Out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not claim a new SOTA performance in their result section.}
+AL research faces four central difficulties:
+{\let\thefootnote\relax\footnotetext{Code available at: anonymous}}
+%\let\thefootnote\relax\footnotetext{Code available at: \url{https://github.com/wernerth94/comparable-active-learning}}
 (i) The experiments are often carried out on different datasets and model architectures, hindering direct comparison,
-(ii) the reported results prove to be very difficult to reproduce.
-While multiple benchmark suites have been proposed to solve (i), to the best of our knowledge, we are the first to compare AL algorithms on all 3 data domains of vector, image and text.
-Regarding (ii), \cite{zhou2021towards} has pointed out severe inconsistencies in results of AL papers in recent years. 
+(ii) generalize poorly across different domains,
+(iii) the reported results can be subject to very high variance across restarts and
+(iv) are not always compared against important baselines like margin sampling \cite{wang2014new}.
+While multiple benchmark suites have been proposed to solve (i), to the best of our knowledge, we are the first to report results on all 3 data domains of tabular, image and text.
+Additionally, we provide synthetic datasets to highlight principled shortcoming of existing AL algorithms.
+Regarding (ii) and (iii), \cite{zhou2021towards} has pointed out severe inconsistencies in results of AL papers in recent years. 
 They conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm (\cite{zhou2021towards} Appendix A).
-The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art performance
-%by significantly outperforming everyone else
-, while the vast majority of results proves to be non-reproducible. \\
-In this work we propose an evaluation protocol that was designed to cope with the high variance in the performances of AL algorithms as well as being fully controllable regardless of the combination of dataset, model and AL algorithm.\\
-We focus our work on single-sample pool-based AL where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
-Specifically, we are not experimenting on so-called batch Active Learning, where at each iteration multiple unlabeled samples are chosen at the same time.
-Even though most proposed algorithms (and benchmarks) are batched, they do not have a principled advantage over single-sample AL except speed of computation.
-The problem of optimizing a portfolio of unlabeled samples in each iteration is more complicated to solve and the algorithms have systematically less information per sample to work with.
-A performance comparison of batch AL and single-sample AL can be found in Fig. \ref{fig:ablation_bald}.
-We can see that for two established algorithms BALD \cite{krizhevsky2009learning} and BADGE \cite{ashdeep} Batch AL can at most perform on-par with single-sample AL, independent of wether algorithm was originally designed for batch AL or not.
-\\
+These outlined issues lead to an inconclusive landscape of AL algorithms, where the vast majority of reported lifts are neither statistically significant, nor prove to be generalizable.
+This makes it very hard to identify the best AL algorithm, or even identifying state-of-the-art algorithms.
+In this work we propose an evaluation protocol that was designed to handle the high variance in the performances of AL algorithms as well as being fully controllable regardless of the combination of dataset, model and AL algorithm.
+We base our work largely on \cite{ji2023randomness}, following their guidelines for a reliable evaluation of AL algorithms, while extending the number of evaluated data domains from 1 to 5. \\
+We focus on pool-based AL where a pool of unlabeled samples is fixed at the start of each experiment and one or more samples are chosen sequentially.
+In addition to the default scenario of selecting a batch of samples every iteration we incorporate the single sample case into our benchmark.
+Batched algorithms (and benchmarks) do not have a principled advantage over single-sample AL except for speed of computation.
+The problem of optimizing a portfolio of unlabeled samples in each iteration is more complicated to solve and the algorithms have systematically less information per sample to work with leading to a generally worse performance, that impacts some algorithms more than others.
+%In Fig. \ref{fig:batch_vs_single} we present a comparison of Badge \cite{ashdeep} and Coreset\cite{sener2017active}, in the batch and single-sample setting.
+%Even though both algorithms were designed for batch sampling, both perform better in the single-sample setting.
+We propose single-sample AL as an important tool to identify the best acquisition function, rather than the best combination of acquisition function and diversity regularization. \\
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=\linewidth]{img/ablation_batch_single}
+%    \caption{Comparison of single-sample vs.
+batch selection for two acquisition functions that were specifically designed for batch selection.
+Even when designed for batch selection, acquisition functions tend to perform better in the single sample setting.}
+%    \label{fig:batch_vs_single}
+%\end{figure}
 Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature. 
+We offer our datasets in two versions - normal and encoded.
+An encoded dataset was pre-encoded by an self-supervised encoder model, providing a different use case for active learning.
+Our synthetic and text datasets (which use word embeddings in the normal setting) are exempt from this, bringing our total dataset count to 14 across 5 domains (Tabular, Image, Text, Synthetic, Encoded).
 \begin{table}
 	\centering
-	\begin{tabular}{l | c c c c c}
-		Paper & Sampling & \#Datasets & \#Domains & \#Algorithms & Oracle \\
+        \caption{Comparison of our benchmark with the existing literature.
+Oracle curves serve as an approximation of the best possible AL algorithm.
+Including the encoded versions of our datasets we reach 14 datasets.
+"Semi" indicates whether the paper is employing any form of self- or semi-supervised learning.
+A "-" for repetitions means that we could not determine how often each experiment is repeated in the respective framework.}
+	\label{tab:benchmark_comparison}
+\begin{tabular}{lccccllllcl}
 \hline
-		Beck et al. \cite{beck2021effective} & batch & 4 & 1 & 7 & - \\
-		Hu et al. \cite{hu2021towards} & batch & 5 & 2 & 13 & - \\
-		Li et al. \cite{li2022empirical} & batch & 5 & 1 & 13 & - \\
-		Zhou et al. \cite{zhou2021towards} & batch & 3 & 2 & 2 & \checkmark \\
-		\textbf{Ours} & single & 7 & 3 & 6 & \checkmark 
+Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
+\multicolumn{1}{l|}{Beck et al. \cite{beck2021effective}} & batch & 4 & 7 & \checkmark & - & - & - & - & - & - \\
+\multicolumn{1}{l|}{Hu et al. \cite{hu2021towards}} & batch & 5 & 13 & \checkmark & \checkmark & - & - & - & - & 3 \\
+\multicolumn{1}{l|}{Zhou et al. \cite{zhou2021towards}} & batch & 3 & 2 & \checkmark & \checkmark & - & - & - & \checkmark & 5 \\
+\multicolumn{1}{l|}{Zhan et al. \cite{zhan2021comparative}} & sngl+batch & 35 & 18 & - & - & \checkmark & \checkmark & - & \checkmark & 10-100 \\
+\multicolumn{1}{l|}{Munjal et al. \cite{munjal2022towards}} & batch & 2 & 8 & \checkmark & - & - & - & - & - & 3 \\
+\multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\
+\multicolumn{1}{l|}{Ji et al. \cite{ji2023randomness}} & batch & 3 & 8 & \checkmark & - & - & - & - & - & - \\
+\multicolumn{1}{l|}{Luth et al. \cite{luth2024navigating}} & batch & 4 & 5 & \checkmark & - & - & - & \checkmark & - & 3 \\
+\multicolumn{1}{l|}{\textbf{Ours}} & sngl+batch & 9(14) & 10 & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & 50 
 \end{tabular}
-	\caption{Comparison of our benchmark with the existing literature. Oracle curves serve as an approximation of the best possible AL algorithm.}
-	\label{tab:benchmark_comparison}
+
 \end{table}
@@
 \begin{enumerate}
-	\item Evaluation of Active Learning algorithms on datasets from 3 different domains
-	\item Two novel synthetic dataset that highlight principled shortcomings of existing AL algorithms
-	\item Novel experimental protocol for controlling AL experiments to allow full reproducibility
-	\item Simple algorithm for an Oracle that can be constructed greedily and does not rely on search
+	\item Evaluation of Active Learning algorithms on datasets from 5 different domains
+	\item Two novel synthetic datasets that highlight principled shortcomings of existing AL algorithms
+	\item Efficient and performant algorithm for an oracle that can be constructed greedily and does not rely on search
+        \item Evidence that previous works might have not repeated their experiments often enough to provide conclusive results
 \end{enumerate}
@@
 \section{Problem Description}
-Given two spaces $\mathcal{X}, \mathcal{Y}$, $n=l+u$ data points with $l \in \mathbb{N}$ labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathbb{R}^\mathcal{X} \to \mathbb{R}^\mathcal{Y}$ that can label $x$.
+Given two spaces $\X, \Y$, $n=l+u$ data points with $l \in \mathbb{N}$ labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a model $\hat y: \X \to \Y$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathcal{X} \to \mathcal{Y}$ that can label $x$ (we consider only hard labels in the one-hot format).
 We call $x \in \mathcal{X}$, $y \in \mathcal{Y}$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$. 
-Find an acquisition function $\Omega: \U^{(i)},\LL^{(i)} \mapsto x^{(i)}$ that iteratively selects the next unlabeled point $x^{(i)} \in \U^{(i)}$ for labeling
+Find an acquisition function $\Omega: \U^{(i)},\LL^{(i)} \mapsto x^{(i)} \in \U^{(i)}$ that iteratively selects the next unlabeled point $x^{(i)}$ for labeling
 \begin{align*}
@@
 \end{align*}
-so that the expected loss $\ell: \mathbb{R}^\mathcal{Y} \times \mathbb{R}^\mathcal{Y} \to \mathbb{R}$ of the machine learning algorithm $\hat y$ after $B$ iterations is minimal: 
-$$\min \quad \mathbb{E}_{(x,y) \sim \rho} \ell(y, \hat{y}(\LL^{(B)}))$$
-This formulation is a special case of the problem formulation in Appendix \ref{app:problem_formulation}.
+with $\U^{(0)} = \operatorname{seed}(\U, s)$ and $\LL^{(0)} = \{\left(\U^{(0)}, A(\U^{(0)})\right)\}$, where $\operatorname{seed}(\U, s)$ selects $s$ points per class for the initial labeled set. \\
+So that the average expected loss $\ell: \mathcal{Y} \times \mathcal{Y} \to \mathbb{R}$ of a machine learning algorithm fitting $\hat y^{(i)}$ on the respective labeled set $\LL^{(i)}$ is minimal: 
+$$\min \quad \frac{1}{B} \sum\limits_{i=0}^B \mathbb{E}_{(x,y) \sim \rho} \ell(y, \hat y^{(i)})$$
 
-\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Related Work}
-\begin{figure}
-	\centering
-	\includegraphics[width=0.49\linewidth]{img/ablation_splice_bald}
-	\includegraphics[width=0.49\linewidth]{img/ablation_splice_badge}
-	\caption{Performance comparison of BALD \cite{gal2017deep} with it's batch AL adaptation \cite{kirsch2019batchbald} and BADGE \cite{ashdeep} with it's single-sample AL adaptation ($k=1$ in \cite{ashdeep} Alg.
-1).
-	In either scenario, the batched algorithm can at most perform on-par with it's single-sample analog, independent of wether it was originally designed for batch AL.}
-	\label{fig:ablation_bald}
-\end{figure}
 Multiple benchmark suites have already been proposed for Active Learning:
-The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus exclusively on Batch AL in the image domain.
-While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
-The authors of \cite{hu2021towards} study models that are trained with AL techniques in the image and text domain.
+The authors of \cite{beck2021effective}, \cite{munjal2022towards}, \cite{li2022empirical}, \cite{ji2023randomness} and \cite{luth2024navigating} focus exclusively on batch AL in the image domain.
+While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training, and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
+The authors of \cite{hu2021towards} study models that are trained on actively learned datasets in the image and text domain.
 They test for several different properties of the models including robustness, response to compression techniques and final performance.
@@
 Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
-The employed AL algorithms for our experiments are introduced in Section \ref{sec:sampling_strategies}.
+The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also discussed the problems of evaluating AL algorithms under many forms of variance.
+We largely adapt the proposed guidelines of \cite{ji2023randomness} and extend their work to multiple domains, batch sizes and comparisons. 
+This work also pays attention to the so-called "pitfalls" of AL evaluation mentioned in \cite{luth2024navigating} with the exception of P1 - Controlling the data distribution of the tested datasets.
+We leave this aspect for future versions of this benchmark. \\
+Two preceding benchmarks have also proposed an oracle algorithm to find a near-optimal labeled set $\LL$, however, both algorithms rely on search and are for that reason considerably slower than our proposed greedy oracle.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation}\label{sec:evaluation}
-Following \cite{zhou2021towards}, the quality of an AL algorithm is evaluated by an ``anytime" protocol that incorporates classification performance at every iteration, as opposed to evaluating final performance after the budget is exhausted.
+\subsection{Evaluation Protocol}\label{sec:evaluation}
+Following \cite{zhou2021towards}, the quality of an AL algorithm is evaluated by an ``anytime protocol" that incorporates classification performance at every iteration, as opposed to evaluating final performance after the budget is exhausted.
 We employ the normalized area under the accuracy curve (AUC):
 \begin{equation}\label{eq:auc}
-	\operatorname{AUC}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
+	\operatorname{AUC}(\D_\test, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\D_\test, \hat y^{(i)})
 \end{equation}
-where $\hat y_i$ is the (re-)trained classification model after the i-\textit{th} iteration.
-To mimic the leave-one-out protocol for cross-validation we will restart each experiment multiple times.
-Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split.
 The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
 AL algorithms have to perform well in both scenarios. 
-Since AL performance inhibits high variance and is prone to outliers, we propose to aggregate AUC values with their median instead of mean. \\ [1mm]
-Since the AUC depends on the chosen budget, we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
-In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
-Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use-cases need different percentages to produce sensible budgets
-\footnote{Since we only consider single-sample AL, increasing the budget comes at a high computational cost. Therefore we could not freely increase the budget above 2000.}.
-We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\%, \textbf{Text}: 95\%, \textbf{Embedded Data}: 90\%. \\ [1mm]
-Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
-To create Fig. \ref{fig:restarts} we used 50 restarts from the margin/random acquisition function on the Splice dataset.
-From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
-One of these median AUC values corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
-To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
-Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
-We can observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
+\\ [1mm]
+Since AUC is still influenced by the budget, we define a set of rules to set this hyperparameter upfront, so that we are not favoring a subset of algorithms by handcrafting a budget.
+In this work, we choose the budget per dataset to be the first point at which one of 2 stopping conditions apply: (i) an algorithm (except Oracle) manages to reach 99\% of the full-dataset-performance (using the smallest query size) or (ii) the best algorithm (except oracle) did not improve the classifier's accuracy by at least 2\% in the last 20\% of iterations.
+\\ [1mm]
+%\footnote{Since we only consider single-sample AL, increasing the budget comes at a high computational cost. Therefore we could not freely increase the budget above 2000.} 
+To mimic the leave-one-out protocol for cross-validation, we will restart each experiment multiple times.
+Each restart retains the train/test split (often given by the dataset itself), but introduces a new validation split that is sampled from the entire dataset (not just $\LL$).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Why we need 50 restarts}\label{sec:restarts}
+To evaluate how many restarts are necessary to obtain conclusive and reproducible results in an AL experiment, we set the number of restarts of the margin sampling baseline for the Splice dataset to 100.
+This allows us firstly, to obtain a very strong estimation of the "true" average performance of margin sampling on this dataset and secondly, to draw subsets from this pool of 100 runs.
+Setting the size of our draws to $\alpha$ and sampling uniformly, we can approximate a cross-validation process with $\alpha$ restarts.
+Each of these draws can be interpreted as a reported result in AL literature where the authors employed $\alpha$ restarts.
+%We provide evidence in Fig. \ref{fig:restarts} that previous works might not have evaluated their experiments with a sufficient number of restarts.
+Figure \ref{fig:restarts} shows the "true" mean performance of margin sampling (green) in relation to random sampling (black) and the oracle performance (red).
+We display 5 random draws of size $\alpha$ in blue.
+We can observe that even for a relatively high number of restarts the variance between the samples is extremely high, resulting in some performance curves being worse that random and some being significantly better.
+When setting $\alpha = 50$ we observe all samples to converge close to the true mean performance. 
+For that reason, we employ 50 restarts of every experiment in this work.
 \begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{img/ablation_restarts}
-	\caption{Random draws from an experimental distribution on the Splice dataset with different numbers of repetitions.
-Each point on the Y-axis represents a cross-validated result that could have been reported in a paper.
-This analysis shows the drastic differences in performance one could observe even when repeating an experiment 2-10 times.}
+        \hspace{-20mm}
+	\includegraphics[width=1.24\linewidth]{img/ablation_restarts}
+	\caption{Random draws from a pool of 100 runs for margin sampling on the Splice dataset with different numbers of repetitions ($\alpha=\{5,10,50\}$).
+Green curves signify the true mean performance of all 100 runs, while the samples are blue.
+Even with 5 or 10 repetitions, we can observe that single draws for margin sampling display below-random performance (black), while the true mean should be above random.}
 	\label{fig:restarts}
 \end{figure}
-To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repetitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
-One way to reduce the number of necessary repetitions would be to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
-We ultimately decided in favor of high variance and high number of repetitions as the high variance accurately reflects real world applications of AL.
-%Fixing for example the classifier initialization to random draws from a constant seed does not have a corresponding use-case in the real world.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Reproducibility}\label{sec:reproducibility}
-Previous works have noted adverse effects of training stochasticity on the evaluation of AL algorithms (\cite{zhou2021towards, lowell2018practical}).
-Both papers observed that an actively sampled labeled set does not generalize well between model architectures or even different initializations of the same model.
-To cater for this we aim to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
-For a fair comparison of two AL algorithms, both algorithms need to receive equal starting conditions in terms of train / validation split, initialization of classifier and even the state of minor systems like the optimizer or mini-batch sampler.
-Even though different implementations might have their own solution to some of these problems, to the best of our knowledge no previous work has discussed this topic in detail.
-The main obstacle for ensuring reproducibility is the seeding utility in PyTorch, Tensorflow and other frameworks, whose default choice is a single global seed.
-Since many systems draw random numbers from this seed, all of them influence each other to a point where a single additional draw can completely change the model initialization or data split, depending on the ordering of these two in the implementation.
-Even though some workarounds exists, like re-setting the seed multiple times, this problem is not limited to the initialization phase, but also extends to the AL iterations and the systems within.
-We propose an implementation that creates a separate Random Number Generator (RNG) for each of these systems to ensure equal testing conditions even when the AL algorithm, dataset or classifier changes.
-We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
-In summary, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
-Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
-Some algorithms require a subsample to be drawn from $\mathcal{U}$ in order to reduce the computational cost in each iteration, while others need access to the full unlabeled pool (i.e.
-for effective clustering).
-If a subsample is required, it will be drawn from $s_\Omega$ and therefore will not influence other systems in the experiments.
-For each algorithm, we decided if subsampling is required based on our available hardware, but decided against setting a fixed time limit per experiment, since this would introduce unnecessary complexity into the benchmark.
-An overview of selected hyperparameters per AL algorithm can be found in Appendix \ref{app:agent_hyperparameters}.
+\subsection{Seeding vs. Restarts}\label{sec:reproducibility}
+Considering the high computational cost of 50 repetitions, another approach to ensure reproducibility would be to reduce the amount of variance in the experiment by keeping as many subsystems (weight initialization, data splits, etc.) as possible fixed with specialized seeding.
+\\
+We describe a novel seeding strategy in Appendix \ref{app:seeding_strategy} that creates 3 separate Random Number Generators (RNG) based on 3 different seeds.
+In summary, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini-batch sampling, and $s_\theta$ for model initialization and sampling of dropout masks.
+Unless stated otherwise, we will keep $s_\Omega$ fixed, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
+While this seeding strategy is capable of controlling the amount variance in the experiment, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), reducing its value in practice and providing a bad approximation of the quality of the AL algorithm.
+Hence, we opt for letting the subsystems vary (by increasing $s_\mathcal{D}$ and $s_\theta$) and combine that with a high number of restarts to obtain a good average of the generalization performance of each AL algorithm.
+\\
+Where a high number of restarts is computationally not feasible, we advise to additionally keep either $s_\mathcal{D}$ or $s_\theta$ (or both) fixed.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Oracle}\label{sec:oracle}
-\begin{minipage}{0.5\linewidth}
-	Posing Active Learning as a combinatorial problem, the oracle set $\mathcal{U}_b$ for a given dataset, model and training procedure is the set that induces the highest AUC score for a given budget.
-	However, since this problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence. 
-	\cite{zhou2021towards} has used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
+\subsection{Datasets}\label{sec:datasets}
+A detailed description of the preprocessing of each dataset can be found in Appendix \ref{app:hyperparameters}. \\ [1mm]
+\textbf{Tabular:}
+AL research conducted on tabular data is sparse (only \cite{ashdeep} from the considered baseline papers). 
+We, therefore, introduce a set of tabular datasets that we selected according to the following criteria:
+(i) They should be solvable by medium-sized models in under 1000 samples, (ii) the gap between most AL algorithms and random sampling should be significant (potential for AL is present) and (iii) the gap between the AL algorithms and our oracle should also be significant (research on these datasets can produce further lifts).
+We use \textbf{Splice}, \textbf{DNA} and \textbf{USPS} from LibSVMTools \cite{libsvmtools}.\\
+\textbf{Image:}
+We use \textbf{FashionMNIST} \cite{xiao2017fashion} and \textbf{Cifar10} \cite{krizhevsky2009learning}, since both are widely used in AL literature.\\
+\textbf{Text:}
+We use \textbf{News Category} \cite{misra2022news} and \textbf{TopV2} \cite{chen-etal-2020-low-resource}.
+Text datasets have seen less attention in AL research, but most of the papers that evaluate on text (\cite{hu2021towards}, \cite{zhou2021towards}) use at least one of these datasets.\\ [1mm]
+%
+We would like to point out that these datasets are selected for speed of computation (both in terms of number of features and necessary budget to solve the dataset). 
+However, similar to our argumentation for picking smaller classifiers, we are solely focused on comparing different AL algorithms in this paper and do not aim to develop novel classification models on these datasets.
+Our assumption is that a well-performing algorithm in our benchmark will also generalize well to larger real-world datasets, because we included multiple different data domains and classifier sizes in our experiments.
+\\ [1mm]
+Adapting the experimental setting from \cite{hacohen2022active}, we offer all our datasets in the un-encoded (normal) setting as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning.
+The text datasets are an exception to this, as they are only offered in their encoded form.
+The pre-encoded datasets enable us to test single-sample algorithms on more complex datasets like Cifar10 and FashionMnist.
+They also serve the purpose of investigating the interplay between self-supervised learning techniques and AL, as well as alleviating the cold-start problem described in \cite{luth2024navigating} as they require a way smaller seed set.
+The classification model for every encoded dataset is a single linear layer with softmax activation.
+The embedding model was trained with the SimCLR \cite{chen2020simple} algorithm adopting the protocol from \cite{hacohen2022active}. 
+%For Cifar10 and FashionMnist we adapt the reported hyperparameters from \cite{hacohen2022active} and for the tabular datasets we use random search to optimize the hyperparameters.
+To ensure that enough information from the data is encoded by our embedding model, the quality of embeddings during pretext training was measured after each epoch.
+We attached a linear classification head to the encoder, fine-tuned it to the data and evaluated this classifier for test accuracy, mirroring our AL setup for embedded datasets. 
+The checkpoints for each encoder model will be provided together with the framework. \\ [1mm]
+\textbf{Synthetic Data}
+Existing AL algorithms can broadly be categorized into two types, uncertainty \cite{wang2014new, gal2017deep} - and geometric-approaches \cite{sener2017active, hacohen2022active, ashdeep}. 
+Both types have principled shortcomings in terms of the utilized information that makes them unsuitable for certain data distributions.
+To test for these shortcomings, we created two synthetic datasets, namely "Honeypot" and "Diverging Sin", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively.
+To avoid algorithms memorizing these datasets they are generated during the experiment, depending on $s_\D$. \\ [1mm]
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.4\linewidth]{img/diverging_sin.jpg}
+	\includegraphics[width=0.4\linewidth]{img/honeypot.jpg}
+	\caption{Our two synthetic datasets.}
+        \label{fig:synthDataAppendix}
+\end{figure}
+Honeypot creates to two easy to distinguish clusters with 150 samples each and one overlapping "honeypot" that represents a noisy region of the dataset with potentially miss-labeled, miss-measured or generally adverse samples.
+This honeypot contains 150 of each class, creating a balance of 50\% beneficial samples and 50\% adverse samples in the dataset.
+The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its negative impact on purely uncertainty based acquisition functions.
+Diverging Sine samples the datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.
+This creates a challenging region one the left hand side, where a lot of datapoints need to be sampled and an easy region on the right hand side, where very few datapoints are enough. 
+The repeating nature of a sin function encourages diversity based acquisition functions to equally sample the entire length, drastically oversampling the right hand side of the dataset.
+Each class has 500 datapoints. \\ [1mm]
+Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Classification Model}\label{sec:choosing_the_classifier}
+\begin{wraptable}{R}{0.55\textwidth}
+    \vspace{-0.7cm}
+    \caption{Overview of available batch sizes for each dataset}
+    \vspace{0.1cm}
+    \label{tab:batch_sizes}
+    \begin{tabular}{l|c|c|c|c|c|c|c}
+                    & 1 & 5 & 20 & 50 & 100 & 500 & 1000 \\
+         \hline
+        Enc. DNA    & o & o &&&&& \\
+        Enc. Splice & o & o & o & o &&&\\
+        TopV2       & o & o & o & o &&& \\
+        Splice      & o & o & o & o & o && \\
+        DNA         & o & o & o & o & o && \\
+        USPS        & o & o & o & o & o && \\
+        Enc. USPS   & o & o & o & o & o && \\
+        Enc. Cifar10& o & o & o & o & o && \\
+        Enc. FMnist & o & o & o & o & o && \\
+        News        &&& o & o & o & o &\\
+        FMnist      &&&&&& o & o\\
+        Cifar10     &&&&&& o & o \\
+    \end{tabular}
+    \vspace{-0.35cm}
+\end{wraptable}
+The model $\hat y$ can be trained in two ways.
+Either the parameters of the model are reset to a fixed initial setting $\hat y^{(0)}$ after each AL iteration and the classifier is trained from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or the previous state $\hat y^{(i-1)}$ is retained and the classifier is fine-tuned on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
+In this work, we use the fine-tuning method for un-encoded datasets to save computational time, while we use the from-scratch training for embedded datasets since they have very small classifiers and this approach generally produces better results.
+Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping with a patience of 2 afterwards.\\
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Batch Sizes}\label{sec:batch_sizes}
+We selected batch sizes for each dataset to accommodate the widest range possible that results in a reasonable runtime for low batch sizes and allows for at least 4 round of data acquisition for high batch sizes.
+The available batch sizes per dataset can be found in Table \ref{tab:batch_sizes}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Realism vs. Variance}\label{sec:realism}
+We would like to point out that some design choices for this framework prohibit direct transfer of our results to practical applications. 
+This is a conscious choice, as we think that this is a necessary trade-off between realism and experiment variance.
+We would like to highlight the following design decisions: \\
+(i) Creating test and validation splits from the full dataset rather than only the labeled seed set.
+Fully fledged test and validation splits are unobtainable in practice, but they provide not only a better approximation of algorithm performance, but also a better foundation for hyperparameter tuning, which is bound to reduce variance in the experiment.
+\\
+(ii) Choosing smaller classifiers instead of SOTA models.
+Since we are not interested in archiving a new SOTA in any classification problem, we instead opt to use smaller classifiers for the following reasons:
+Smaller classifiers generally exhibit more stable training behavior, on average require fewer sampled datapoints to reach their full-dataset-performance and have faster training times.
+For every dataset, the chosen architecture's hyperparameters are optimized to archive maximum full-dataset performance.
+Generally, we use MLPs for tabular, RestNet18 for image and BiLSTMs for text datasets.
+Every encoded dataset is classified by a single linear layer with softmax activation.
+For an overview of architectures and hyperparameters please refer to Appendix \ref{app:hyperparameters}. \\
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Greedy Oracle Algorithm}\label{sec:oracle}
+%\begin{wrapfigure}{R}{0.36\textwidth}
+%    \vspace{-0.83cm}
+%    \begin{minipage}{0.36\textwidth}
+%       Space for the Oracle Algorithm
+%    \vspace{-0.83cm}
+%\end{minipage}
+%\end{wrapfigure}
+Posing Active Learning as a combinatorial problem, the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure is the set that induces the highest AUC score for a given budget.
+However, since this problem is computationally infeasible for realistic datasets, previous works have proposed approximations to this oracle sequence. 
+\cite{zhou2021towards} used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
 Even though their reported performance curves display a significant lift over all other acquisition functions, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
-	In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
-\end{minipage}
-\hspace{2mm}
-\begin{minipage}{0.45\linewidth}
-	\begin{algorithm}[H]
-		\caption{Oracle}\label{alg:oracle}
-		\begin{algorithmic}[1]
-			\Require $\mathcal{U}, \mathcal{L}, A, \mathcal{D}_\text{test}$ Train, Margin, $\tau, \hat y_\theta$ 
-			\State $\text{acc} \gets \operatorname{Train}(\mathcal{L}, \mathcal{D}_\text{test}, \hat y_\theta)$ 
-			%\State $u \underset{\tau}{\sim} \text{unif}(1:|\mathcal{U}|)$
-			\State $r^* \gets 0$
-			\For{$k := 1 \ldots \tau$}
-			\State $\mathcal{L}' \gets \mathcal{L}^{(i)} \cup \{(u_k, A(u_k))\}$
-			\State $\text{acc}' \gets \operatorname{Train}(\mathcal{L}', \mathcal{D}_\text{test}, \hat y_\theta)$  
-			\State $r \gets \text{acc} - \text{acc}'$
-			\If{$r > r^*$} 
-			\State $r^* \gets r$
-			\State $u^* \gets u_k$
-			\EndIf
-			\EndFor
-			\If{$r^* = 0$}
-			\State $u^* \gets \operatorname{margin}(\mathcal{U}, \hat y_\theta)$
-			\EndIf
-			\Return $u^*$
-		\end{algorithmic}
-	\end{algorithm}
-\end{minipage} \\ [1mm]
-Our oracle simply evaluates every data point $u_k = \operatorname{unif(\mathcal{U} | k = 1 \ldots \tau)}$ in a subsample of unlabeled points by fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup u_k$ and directly measuring the resulting test performance.
+In this paper, we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
+Our oracle algorithm evaluates every data point $u_k = \operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$ in a subsample of unlabeled points by fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test performance.
 The data point with the best test performance is selected and added to the labeled pool for that iteration.
 We noticed that this oracle is overfitting on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
-This can happen for example, if the oracle picked a labeled set that enables the classifier to correctly classifier a big portion of easy samples in the test set, but now fails to find the next single unlabeled point that would enable the classifier to succeed on one of the hard samples in the test set.
-This leads to a situation, where the selected data point is basically random.\\
-To circumvent this problem, we introduced margin sampling \cite{wang2014new} as a fallback option for the oracle.
+This can happen, for example, if the oracle picked a labeled set that enables the classifier to correctly classify a big portion of easy samples in the test set, but now fails to find the next single unlabeled point that would enable the classifier to succeed on one of the hard samples.
+This leads to a situation, where no point can incur an increase in test performance and therefore the selected data point can be considered random.
+To circumvent this problem, we use margin sampling \cite{wang2014new} as a fallback option for the oracle.
 Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
-The resulting greedy algorithm constructs an approximation of the optimal labeled set that consistently outperforms all other algorithms by a significant margin, while requiring relatively low computational cost (scaling linearly in $\tau$).
-The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.
-In the algorithm $\operatorname{Train}(\LL, \D_\text{test}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns it's accuracy  on $\D_\text{test}$. \\
-Alg. \ref{alg:oracle} replaces the acquisition function in the AL process.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sampling Strategies}\label{sec:sampling_strategies}
-We selected AL algorithms that have good performances reported by multiple different sources.
-To ensure a fair comparison, we fixed the training process of our classification model as well as the set of available information for the algorithms and selected only those that can work under these restrictions:\\
-\textbf{Uncertainty Sampling} 
-Tries to find the sample that the classifier is most uncertain about by computing heuristics of the class probabilities.
-For our benchmark we use entropy and margin (a.k.a.
-best-vs-second-best) sampling.\\
-\textbf{BALD \cite{kirsch2019batchbald}}
-Applies the query-by-committee strategy of model ensembles to a single model by interpreting the classifier's parameters as distributions and then sample multiple outputs from them via Monte-Carlo dropout.\\
-\textbf{BADGE \cite{ashdeep}} Uses gradient embeddings of unlabeled points to select samples where the classifier is expected to change a lot.
-The higher the magnitude of the gradient the higher is the expected improvement in model performance.\\
-\textbf{Coreset \cite{sener2017active}}
-Employs K-Means clustering to try to cover the whole data distribution.
-Selects the unlabeled sample that is the furthest away from all cluster centers.
-Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\hat y$.
-In this work we use the greedy variant of Coreset.\\
-\textbf{TypiClust \cite{hacohen2022active}}
-Relies on clustering similar to Coreset but proposes a new measure called "Typicality" to select unlabeled samples.
-Selects points that are in the densest regions of clusters that do not contain labeled samples yet.
-Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\hat y$.
-It has to be pointed out that TypiClust was designed for low-budget scenarios, but we think it is still worthwhile to test and compare this algorithm with practically relevant budgets.
-%
-\subsubsection*{Honorable Mentions}
-\textbf{Learning Loss for AL}
-Introduces an updated training of the classification model with an auxiliary loss and therefore cannot be compared fairly against classification models without this boosted training regime.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Choosing the Classifier}\label{sec:choosing_the_classifier}
-Traditionally, the classifier is chosen per dataset so that it is capable of solving the dataset close to the SOTA performance reported in the literature.
-Since we are not interested in archiving a new SOTA in any classification problem, we opt to use smaller classifiers for the following reasons:
-Smaller classifiers generally (i) exhibit more stable training behavior and (ii) on average require less sampled datapoints to reach the their upper bound performance on the full dataset.
-For every dataset the chosen architecture's hyperparameters are optimized by to archive maximum upper bound performance.
-One desired characteristic of these small classifiers is that the ranking of AL algorithms should stay the same when switching to larger models.
-A small analysis of this behavior can be found in Appendix \ref{app:classifier_size}.
-We found that the ranking of AL algorithms unfortunately does change slightly, but we did not observe systematics that benefit one or few specific algorithms.
-We therefore rely on the different data domains to provide classification models of different sizes and archetypes to cover all of the use-cases.
-For an overview of architectures and hyperparameters please refer to Appendix \ref{app:hyperparameters}.
+The resulting greedy algorithm constructs an approximation of the optimal labeled set that consistently outperforms all other algorithms by a significant margin, while requiring relatively low computational cost ($\mathcal{O}(B\tau)$).
+We fix $\tau = 20$ in this work, as this gave us already a significant lift and we expect diminishing returns for larger $\tau$.
+The pseudocode for our oracle can be found in App. \ref{app:pseudocode}.
+%In the algorithm $\operatorname{Train}(\LL, \hat y_\theta)$ trains the classification model $\hat y_\theta$ on $\LL$. \\
+%Alg. \ref{alg:oracle} replaces the acquisition function $\Omega$ in the AL loop (Appendix \ref{app:pseudocode} Alg. \ref{alg:active_learning}).
+Even though our proposed algorithm is more efficient than other approaches, the computational costs for high budget datasets like Cifa10 and FashionMnist meant that we could not compute the oracle for all 10000 datapoints.
+To still provide an oracle for these two datasets, we select two points per iteration instead of one and stop the oracle computation at a budget of 5000.
+The rest of the curve is forecast with a simple linear regression that asymptotically approaches the upper bound performance of the dataset. 
+A detailed description can be found in App. \ref{app:oracle_forecasting}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Implementation Details}
-At each iteration $i$ the AL algorithm picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \hat y^{(i)}, \text{opt}_{\hat y}\}$, where $\hat y^{(i)}$ is the current classifier and $\text{opt}_{\hat y}$ is the optimizer used to fit $\hat y^{(i)}$.
-We allow acquisition functions to derive additional information of this set like predictions of the classifier, K-Means clustering or even training new classifiers.
-However, the algorithm may not incorporate external information like other datasets, queries to recover additional labels, additional training steps for $\hat y$, or the test/validation set. \\ [1mm]
-The classification model can be trained in two ways.
-Either you reset the parameters after each AL iteration and train the classifier from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or you retain the previous state and only fine-tune the classifier on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
-In this work we use the fine-tuning method for raw datasets to save computation, while we use the from-scratch training for embedded dataset, since they have very small classifiers and this method generally produces better results.
-Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping after that.
-The early stopping has patience 0, so it will stop as soon as the validation loss does no longer decrease.
-Even though the use of a fully labeled validation set might be regarded as impractical, since such a set will never exist during deployment, we strongly advocate for using it in benchmarks to control the classifier training.
-In this work we use the validation set to optimize the hyperparameters of the classifier and reduce overfitting with early stopping the training process in every iteration.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}\label{sec:datasets}
-For all our datasets we use the pre-defined train / test splits, if given. 
-In the remaining cases, we define test sets upfront and store them into separate files to keep them fixed across all experiments.
-The validation set is split during experiment-time and depends on the dataset-seed.\\
-\textbf{Tabular:}
-We use \textbf{Splice}, \textbf{DNA} and \textbf{USPS} from LibSVMTools \cite{libsvmtools}.
-All three datasets are normalized between [0, 1]. \\
-\textbf{Image:}
-We use \textbf{FashionMNIST} \cite{xiao2017fashion} and \textbf{Cifar10} \cite{krizhevsky2009learning}.
-Both datasets are normalized according to their standard protocols. \\
-\textbf{Text:}
-We use \textbf{News Category} \cite{misra2022news} and \textbf{TopV2} \cite{chen-etal-2020-low-resource}.
-For News Category we use  the 15 most common categories as indicated by its Kaggle site.
-We additionally drop sentences above 80 words to reduce the padding needed (retaining 99,86\% of the data).
-For TopV2, we are only using the "alarm" domain.
-Both datasets are encoded with pre-trained GloVe (Common Crawl 840B Tokens) embeddings \cite{pennington2014glove}.
-Since neither dataset provided a fixed test set, we randomly split 7000 datapoints into a test set. \\ [1mm]
+\subsection{Implementation Details}\label{sec:implementation_details}
+At each iteration $i$ the acquisition function $\Omega$ picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \hat y^{(i)}, \text{opt}_{\hat y}\}$, where $\text{opt}_{\hat y}$ is the optimizer used to fit $\hat y^{(i)}$.
+This set grants full access to the labeled and unlabeled set, as well as all parameters of the classifier and the optimizer.
+Additionally, we provide meta-information, like the size of the seed set through $|\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|$, the remaining budget though the addition of $B$ and the classifiers potential though $\text{acc}^{(1)}$ and $\text{acc}^{(i)}$.
+We allow acquisition functions to derive information from this set, e.g.
+predictions of the classifier $\hat y^{(i)}(x); \hspace{2mm} x \in \mathcal{U}^{(i)} \cup \mathcal{L}^{(i)}$, clustering, or even training additional models.
+However, the algorithm may not incorporate external information e.g. other datasets, queries to recover additional labels, additional training steps for $\hat y$, or the test/validation set. \\
+For our study we selected acquisition functions with good performances reported by multiple different sources that can work with the set of information stated above.
+For a list of all acquisition functions, please refer to Table \ref{tab:results}, with detailed descriptions being found in Appendix \ref{app:acquisition_functions}. \\ [1mm]
 %
-We would like to point out that these datasets can be considered "toy-datasets" and therefore not relevant for practical purposes.
-This might be true if we aimed to develop novel classification models on these datasets, however, similar to our argumentation for picking smaller classifiers, we are solely focused on comparing different AL algorithms in this paper.
-Our core assumption is that a well-performing algorithm in our benchmark will also transfer into more practical use-cases. \\ [1mm]
-Adapting the experimental setting from \cite{hacohen2022active} we offer all our datasets in the raw setting as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning.
-The text datasets are an exception, as they are only offered in their encoded form.
-The pre-encoded datasets enable us to test our single-sample algorithms on more complex datasets like Cifar10 and FashionMnist.
-The embedding model was trained with the SimCLR \cite{chen2020simple} algorithm. 
-For Cifar10 we adapt the reported hyperparameters from \cite{hacohen2022active} and for the tabular datasets we use random search to optimize the hyperparameters.
-The quality of embeddings during pretext training was measured after each epoch by attaching a linear classification head and evaluating this classifier for test accuracy, mirroring our AL setup for embedded datasets.
-\\ [1mm]
-Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being raw FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.
+%This work employs a validation split that is produced from the source dataset itself before active learning instead of splitting it from the seed set or omitting it altogether.
+%Even though the use of a fully labeled validation set might be regarded as impractical, since such a set will never exist during deployment, we strongly advocate for using it in \textbf{research}, as it can be used for hyperparameter tuning, resulting in more stable hyperparameters and reducing the overall training stochasticity.
+%We optimize all hyperparameters on our fully labeled validation sets using grid search.
 
@@
 \subsection{Results}
-
+%
 \begin{table}[]
-	\hspace{-5mm}
-	\begin{tabular}{l|lllll|c}
-		& Vector       & Vector Enc. & Image        & Image Enc. & Text         & All          \\
+	\caption{Performances for acquisition functions on each dataset and aggregated for un-encoded and encoded datasets.
+Performance is shown as average ranks over restarts (1.0 is the best rank).
+Algorithms are sorted by aggregated performance on un-encoded datasets.
+For brevity, we only display per-dataset standard deviations for Splice.
+All standard deviations lie within $[0.001, 0.024]$.}
+	\label{tab:results}
+	\centering
+        %\resizebox{\columnwidth}{!}{
+\begin{tabular}{l|lllllll|ll}
+          & Splice & DNA  & USPS & Cfr10 & FMnist & TopV2 & News  & Un-encoded & Encoded \\
           \hline
-		Oracle          & 1.07 $\pm$ 0.02 & 1.13 $\pm$ 0.01    & 1.07 $\pm$ 0.04 & 1.16 $\pm$ 0.02   & 1.16 $\pm$ 0.07 & 1.12 $\pm$ 0.05 \\
-		Margin     & 1.03 $\pm$ 0.02 & 1.03 $\pm$ 0.03    & 1.01 $\pm$ 0.01 & 1.07 $\pm$ 0.00   & 1.04 $\pm$ 0.00 & 1.04 $\pm$ 0.02 \\
-		Badge           & 1.02 $\pm$ 0.01 & 0.98 $\pm$ 0.04    & 1.01 $\pm$ 0.00 & 1.07 $\pm$ 0.00   & 1.04 $\pm$ 0.01 & 1.02 $\pm$ 0.04 \\
-		Entropy  & 1.03 $\pm$ 0.02 & 0.99 $\pm$ 0.01    & 1.00 $\pm$ 0.01 & 1.07 $\pm$ 0.00   & 1.02 $\pm$ 0.00 & 1.02 $\pm$ 0.03 \\
-		Coreset & 1.02 $\pm$ 0.01 & 0.99 $\pm$ 0.04    & 1.00 $\pm$ 0.00 & 1.07 $\pm$ 0.00   & 1.03 $\pm$ 0.02 & 1.02 $\pm$ 0.04 \\
-		TypiClust       & 1.03 $\pm$ 0.03 & 1.05 $\pm$ 0.04    & 0.99 $\pm$ 0.02 & 1.03 $\pm$ 0.01   & 0.98 $\pm$ 0.04 & 1.02 $\pm$ 0.04 \\
-		Random     & 1.00 $\pm$ 0.00 & 1.00 $\pm$ 0.00    & 1.00 $\pm$ 0.00 & 1.00 $\pm$ 0.00   & 1.00 $\pm$ 0.00 & 1.00 $\pm$ 0.00 \\
-		BALD            & 0.98 $\pm$ 0.03 & 0.90 $\pm$ 0.04    & 0.99 $\pm$ 0.03 & 1.06 $\pm$ 0.01   & 1.04 $\pm$ 0.01 & 0.98 $\pm$ 0.07
+Oracle & 1.0$\pm$0.011 & 1.0 & 1.0 & 3.9 & 7.3 & 1.0 & 1.0 & 2.3 & 1.9 \\
+Margin & 3.9$\pm$0.019 & 4.2 & \textbf{2.1} & 4.7 & \textbf{2.9} & \textbf{2.4} & 2.7 & 3.3 & 3.9 \\
+Badge & \textbf{3.0$\pm$0.014} & 6.2 & 2.9 & \textbf{3.5} & 3.2 & 3.3 & \textbf{2.5} & 3.5 & 4.9 \\
+DSA & 7.1$\pm$0.016 & 7.1 & 7.1 & 3.8 & 3.5 & 5.7 & 5.9 & 5.8 & 6.2 \\
+CoreGCN & 6.6$\pm$0.012 & 4.8 & 9.4 & 6.1 & 4.8 & 4.0 & 5.4 & 5.9 & 7.7 \\
+BALD & 3.9$\pm$0.011 & 4.6 & 5.3 & 10.6 & 5.8 & 7.2 & 3.8 & 5.9 & 6.9 \\
+Entropy & 6.4$\pm$0.018 & \textbf{3.8} & 7.2 & 6.0 & 3.3 & 9.1 & 8.2 & 6.3 & 6.0 \\
+LSA & 5.8$\pm$0.009 & 6.6 & 5.2 & 6.1 & 9.2 & 7.0 & 6.0 & 6.6 & 6.9 \\
+Coreset & 6.8$\pm$0.01 & 8.8 & 9.6 & 5.2 & 5.3 & 8.0 & 9.4 & 7.6 & 6.6 \\
+Random & 8.6$\pm$0.01 & 9.0 & 5.2 & 6.9 & 9.7 & 7.4 & 6.6 & 7.6 & 6.4 \\
+TypiClust & 8.3$\pm$0.009 & 9.8 & 11.0 & 6.4 & 9.6 & 11.0 & 10.6 & 9.5 & 8.6
 \end{tabular}
-	\caption{Performances for each algorithm per domain normalized by the performance of random sampling. Higher is better, scores below 1 indicate worse-than-random performance.}
-	\label{tab:results}
+        %}
 \end{table}
-%
 %\begin{figure}
 %    \centering
-%	\includegraphics[width=0.49\linewidth]{img/eval_splice}
-%	\hspace{1mm}
-%	\includegraphics[width=0.49\linewidth]{img/eval_dna}\\[2mm]
-%	\includegraphics[width=0.49\linewidth]{img/micro_splice}
-%	\hspace{1mm}
-%	\includegraphics[width=0.49\linewidth]{img/micro_dna}
-%	\caption{Results for all algorithms on Splice and DNA, both from the tabular domain. Even within one domain, the performance of the same algorithm can vary drastically.}
-%	\label{fig:eval_vector}
+%    \caption{Results for all acquisition functions on un-encoded USPS (top) and the encoded FashionMnist (bottom) for two query sizes.}
+%    \label{fig:main_body_result}
+%    \includegraphics[width=0.9\linewidth]{img/main_body_result.jpg}
 %\end{figure}
@@
     \centering
-	\includegraphics[width=0.49\linewidth]{img/macro_vector}
-	\includegraphics[width=0.49\linewidth]{img/macro_vector_enc}
-	\includegraphics[width=0.49\linewidth]{img/macro_img}
-	\includegraphics[width=0.49\linewidth]{img/macro_img_enc}
-	\includegraphics[width=0.49\linewidth]{img/macro_text}
-	\includegraphics[width=0.49\linewidth]{img/macro_all.jpg}
-	\caption{Critical Difference Diagram for all algorithms grouped by domain and all domains combined.
-Ranks are computed based on median AUC for each algorithm and dataset combination.
-Lower ranks are better.}
-	\label{fig:cd_diagrams}
+    \caption{Ranks of each acquisition function aggregated by domain.}
+    \label{fig:ranks_by_domain}
+    \includegraphics[width=0.49\linewidth]{img/macro_vector.jpg}
+    \includegraphics[width=0.49\linewidth]{img/macro_text.jpg}
+    \includegraphics[width=0.49\linewidth]{img/macro_img.jpg}
+    \includegraphics[width=0.49\linewidth]{img/macro_enc.jpg}
 \end{figure}
 %
-We report our results in two different fashions: Table \ref{tab:results} contains mean AUC performances per data domain.
-To make performances from different datasets comparable we normalized them with the random performance of each dataset.
-Figure \ref{fig:cd_diagrams} aggregates the performances per domain by counting wins and deriving a ranking of algorithms.
-This is closely related to the evaluation protocol of \cite{ashdeep}, who displayed a pairwise penalty matrix for all algorithms. \\
-Based these two evaluations we make the following observations:\\
-(i) The ranking of both reports does not match exactly.
-This is due to Table \ref{tab:results} being more sensitive to outlier performances (i.e. BALD being worse than random due to it's bad performance on embedded vector data). \\
-(ii) Both evaluations make it clear that no one-fits-all algorithm for all domains seems to exist, with only margin sampling being significantly better that random in Tab.
-\ref{tab:results} and the best 3 ranks for all domains in Fig.
-\ref{fig:cd_diagrams} being close together.
-\\
-(iii) Even though in Fig.
-\ref{fig:cd_diagrams} some domains show evidence of one algorithm being systematically better than the rest (i.e.
-Image), the actual differences in Tab.
-\ref{tab:results} seem to be negligible.
-\\
-(iv) Not only do different domains favor different AL algorithms, we noticed that even within the same domain datasets can induce drastically different behaviors (Compare Appendix \ref{app:all_results} Splice/DNA/USPS).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-Our results empirically reinforce what the authors of \cite{zhou2021towards} have noticed in their meta-analysis already:
-When slightly altering the dataset, model or training conditions, most AL algorithms are reduced to uncertainty sampling or worse.
-Additionally, some results remain non-reproducible even on the same dataset and model combination (Compare Appendix \ref{app:all_results} Cifar10 with \cite{hacohen2022active} or \cite{sener2017active}).
-On the other hand we were able to also confirm some previous results i.e. for the NewsCategory dataset (Compare Appendix \ref{app:all_results} News with \cite{hu2021towards} Fig 3g). \\ [1mm]
-Regarding meta analysis of our results we have the following conclusions:\\
-(i) BALD seems to be performing especially bad when the input dimensionality is low and the applied classifier is too simple. (i.e. (Encoded) Vector). \\
-(ii) TypiClust seemingly holds its claim to work well with extremely low budgets (Encoded Vector $B=\{50,60,600\}$), but fails to push the classifier close to upper bound performance even with moderate budgets in other domains.
-\\
-(iii) The authors of \cite{ashdeep} noted that diversity sampling and adjacent methods, like Coreset and TypiClust, only work well when the classifier has correct inductive biases (correct architecture, tuned hyperparameters).
-Since our architectures are extensively tuned, the performance of these algorithms might be slightly inflated.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Limitations and Future Work}
-In the current state, a couple of aspects/use-cases are missing from our proposed benchmark and are considered future work: \\
-(i) We elected to not provide any experiments on batch AL, but the provided implementation is already capable of doing that and we plan to extend the benchmark to include batch sizes larger than 1 in our experiments.
-\\
-(ii) Due to difficulties with reproducing results and high computational setup costs, we did not include learned acquisition functions like \cite{shao2019learning}, \cite{vu2019learning} or \cite{konyushkova2017learning} \\
-(iii) Our training process for the classifier does not include techniques that would be present in real-world scenarios, like data augmentation, semi-supervised learning, or auxiliary tasks.
-We opted for simpler training process to be able to extensively tune those fewer hyperparameters and have full control over the training.\\
-(iv) We used SimCLR as pretext task for our vector datasets, which primarily relies on data augmentation.
-While images have a plethora of well-defined augmentation techniques, we had to rely on Gaussian noise to augment the vector data.
-Other pretext tasks, like autoencoders have potential to provide better encodings of the data.
+\begin{figure}
+    \centering
+    \caption{Results for all acquisition functions on both synthetic datasets.}
+    \label{fig:main_body_result}
+    \includegraphics[width=0.9\linewidth]{img/synth_trajectories.jpg}
+\end{figure}
+Storing the results of every run in a separate file allows us to create various aggregates.
+Apart from plotting standard performance curves and reporting their AUC values per dataset (Fig.
+\ref{fig:main_body_result}), we primarily rely on ranks to aggregate the performance of an acquisition function across datasets.
+For each dataset, the AUC values of all acquisition functions are sorted and assigned a rank based on position, with the best rank being 1.
+These ranks can safely be averages across datasets as they are no longer subjected to scaling differences of each dataset.
+In Table \ref{tab:results} we provide the rank of each acquisition function per dataset and averaged for each (un-)encoded dataset. \\
+Additionally, we provide aggregated ranks per data domain in Fig. \ref{fig:ranks_by_domain}, revealing varying performances of acquisition functions like DSA/LSA, BALD and CoreGCN.
+The leading algorithms Margin-Sampling and BADGE show very stable performance across all tested data domains. \\ [1mm]
+Results for the Honeypot dataset reveal expected shortcomings of uncertainty sampling algorithms like margin and entropy sampling, but also show that BADGE is underperforming for this dataset. 
+This is an important result, as margin sampling and BADGE were evaluated to be the best performing acquisition functions for real-world data.
+Both being highly vulnerable to adverse samples or simply measurement noise highlights the need for further research in this area. \\
+Results for Diverging Sine also confirm expected behaviour, as clustering algorithms (Coreset, TypiClust) fall behind uncertainty algorithms (Entropy-, Margin-Sampling), with the exception of BADGE.
+The fact that BADGE is able to perform well on this dataset highlights the importance of embeddings in the clustering algorithms, as the so-called gradient embedding from BADGE seems to be able to encode uncertainty information, guiding the selection into the left hand regions of the dataset.
+We provide additional evidence for the importance of the embeddings by testing a version of Coreset on this dataset that does not use the embeddings produced by the classification model, but rather clusters the data directly.
+This "Coreset Raw" algorithm performs notably worse than its embedding-based counterpart. \\
+We strongly advocate to test newly proposed AL algorithms not only on our benchmark suite of real data, but also pay close attention to Honeypot and Diverging Sine to reveal principled shortcomings of the algorithm.
 
 \newpage
-
-\begin{ack}
-	Funded by the Lower Saxony Ministry of Science and Culture under grant number ZN3492 within the Lower Saxony “Vorab“ of the Volkswagen Foundation and supported by the Center for Digital Innovations (ZDIN).
-\end{ack}
+\paragraph{Acknowledgement}
+anonymous
+%TODO
+%Funded by the Lower Saxony Ministry of Science and Culture under grant number ZN3492 within the Lower Saxony “Vorab“ of the Volkswagen Foundation and supported by the Center for Digital Innovations (ZDIN).
 
 \bibliographystyle{plain}
-\bibliography{main.bib}
+\bibliography{al_benchmark_paper.bib}
 
 \appendix
+\newpage
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Acquistion Functions}\label{app:acquisition_functions}
+\textbf{Uncertainty Sampling} 
+tries to find the sample that the classifier is most uncertain about by computing heuristics of the class probabilities.
+For our benchmark, we use entropy and margin (a.k.a.
+best-vs-second-best) sampling.\\
+\textbf{BALD \cite{kirsch2019batchbald}}
+applies the query-by-committee strategy of model ensembles to a single model by interpreting the classifier's parameters as distributions and then sample multiple outputs from them via Monte-Carlo dropout.\\
+\textbf{BADGE \cite{ashdeep}} uses gradient embeddings of unlabeled points to select samples where the classifier is expected to change a lot.
+The higher the magnitude of the gradient the higher the expected improvement in model performance.\\
+\textbf{Coreset \cite{sener2017active}}
+employs K-Means clustering trying to cover the whole data distribution.
+Selects the unlabeled sample that is the furthest away from all cluster centers.
+Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\hat y$.
+In this work, we use the greedy variant of Coreset.\\
+\textbf{TypiClust \cite{hacohen2022active}}
+relies on clustering similar to Coreset, but proposes a new measure called ``Typicality" to select unlabeled samples.
+It selects points that are in the densest regions of clusters that do not contain labeled samples yet.
+Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\hat y$.
+It has to be pointed out that TypiClust was designed for low-budget scenarios, but we think it is still worthwhile to test and compare this algorithm with higher budgets. \\
+\textbf{Core-GCN \cite{caramalau2021sequential}} TODO \\
+\textbf{DSA/LSA \cite{kim2019guiding}} TODO \\ [2mm]
+%
+\textbf{Excluded Algorithms}\\
+\textbf{Learning Loss for AL \cite{yoo2019learning}}
+Introduces an updated training of the classification model with an auxiliary loss and therefore cannot be compared fairly against classification models without this boosted training regime.\\ [1mm]
+%
+\textbf{Reinforcement Learning Algorithms} \\
+TODO \\
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{All Results}\label{app:all_results}
+% Splice
+\begin{minipage}{0.65\linewidth}
 \begin{figure}[H]
     \centering
-	\includegraphics[width=0.4\linewidth]{img/eval_splice}
-	\includegraphics[width=0.4\linewidth]{img/eval_splice_enc} \\ [2mm]
-	\includegraphics[width=0.3\linewidth]{img/micro_splice.jpg}
-	\includegraphics[width=0.3\linewidth]{img/micro_splice_enc.jpg} \\ [4mm]
+    \includegraphics[width=\linewidth]{img/eval_splice}\\ [2mm]
+    \includegraphics[width=\linewidth]{img/micro_splice.jpg}
 \end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+    &Splice \\
+    \hline
+    Oracle&0.811 $\pm$ 0.010\\
+    BALD&0.785 $\pm$ 0.013\\
+    Coreset&0.778 $\pm$ 0.014\\
+    Entropy&0.774 $\pm$ 0.016\\
+    Margin&0.773 $\pm$ 0.016\\
+    Badge&0.770 $\pm$ 0.016\\
+    Random&0.768 $\pm$ 0.014\\
+    TypiClust&0.766 $\pm$ 0.014\\
+\end{tabular}
+\end{minipage}
+%DNA
+\begin{minipage}{0.65\linewidth}
 \begin{figure}[H]
     \centering
-	\includegraphics[width=0.4\linewidth]{img/eval_dna}
-	\includegraphics[width=0.4\linewidth]{img/eval_dna_enc} \\ [2mm]
-	\includegraphics[width=0.3\linewidth]{img/micro_dna.jpg}
-	\includegraphics[width=0.3\linewidth]{img/micro_dna_enc.jpg} \\ [4mm]
+	\includegraphics[width=\linewidth]{img/eval_dna} \\ [2mm]
+        \includegraphics[width=\linewidth]{img/micro_dna.jpg}
 \end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&DNA \\
+\hline
+Oracle&0.842 $\pm$ 0.021\\
+Margin&0.807 $\pm$ 0.035\\
+Entropy&0.805 $\pm$ 0.038\\
+Coreset&0.796 $\pm$ 0.028\\
+Badge&0.789 $\pm$ 0.056\\
+TypiClust&0.788 $\pm$ 0.036\\
+Random&0.768 $\pm$ 0.024\\
+BALD&0.749 $\pm$ 0.044\\
+\end{tabular}
+\end{minipage}
+%USPS
+\begin{minipage}{0.65\linewidth}
 \begin{figure}[H]
     \centering
-	\includegraphics[width=0.4\linewidth]{img/eval_usps}
-	\includegraphics[width=0.4\linewidth]{img/eval_usps_enc} \\ [2mm]
-	\includegraphics[width=0.3\linewidth]{img/micro_usps.jpg}
-	\includegraphics[width=0.3\linewidth]{img/micro_usps_enc.jpg} \\ [4mm]
+	\includegraphics[width=\linewidth]{img/eval_usps} \\ [2mm]
+	\includegraphics[width=\linewidth]{img/micro_usps.jpg}
 \end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&USPS\\
+\hline
+TypiClust&0.830 $\pm$ 0.007\\
+Oracle&0.823 $\pm$ 0.011\\
+Margin&0.809 $\pm$ 0.013\\
+Entropy&0.807 $\pm$ 0.013\\
+Badge&0.795 $\pm$ 0.018\\
+Coreset&0.788 $\pm$ 0.017\\
+Random&0.774 $\pm$ 0.012\\
+BALD&0.725 $\pm$ 0.050\\
+\end{tabular}
+\end{minipage}
+%Splice Enc
+\begin{minipage}{0.65\linewidth}
 \begin{figure}[H]
     \centering
-\includegraphics[width=0.4\linewidth]{img/eval_cifar10}
-\includegraphics[width=0.4\linewidth]{img/eval_cifar10_enc} \\ [2mm]
-\includegraphics[width=0.3\linewidth]{img/micro_cifar10.jpg}
-\includegraphics[width=0.3\linewidth]{img/micro_cifar10_enc.jpg} \\ [4mm]
+    \includegraphics[width=\linewidth]{img/eval_splice_enc} \\ [2mm]
+    \includegraphics[width=\linewidth]{img/micro_splice_enc.jpg}
 \end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&SpliceEncoded \\
+\hline
+Oracle&0.728 $\pm$ 0.022\\
+Coreset&0.648 $\pm$ 0.027\\
+TypiClust&0.645 $\pm$ 0.042\\
+Random&0.643 $\pm$ 0.036\\
+Entropy&0.636 $\pm$ 0.033\\
+Margin&0.636 $\pm$ 0.033\\
+Badge&0.627 $\pm$ 0.040\\
+BALD&0.565 $\pm$ 0.049\\
+\end{tabular}
+\end{minipage}
+%DNA Enc
+\begin{minipage}{0.65\linewidth}
 \begin{figure}[H]
     \centering
-\includegraphics[width=0.4\linewidth]{img/eval_fmnist}
-\includegraphics[width=0.4\linewidth]{img/eval_fmnist_enc} \\ [2mm]
-\includegraphics[width=0.3\linewidth]{img/micro_fmnist.jpg}
-\includegraphics[width=0.3\linewidth]{img/micro_fmnist_enc.jpg} \\ [4mm]
+    \includegraphics[width=\linewidth]{img/eval_dna_enc}\\ [2mm]
+    \includegraphics[width=\linewidth]{img/micro_dna_enc.jpg} 
 \end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&DNAEncoded\\
+\hline
+Oracle&0.709 $\pm$ 0.023\\
+TypiClust&0.672 $\pm$ 0.029\\
+Margin&0.648 $\pm$ 0.047\\
+Badge&0.647 $\pm$ 0.037\\
+Coreset&0.640 $\pm$ 0.041\\
+Entropy&0.629 $\pm$ 0.062\\
+Random&0.626 $\pm$ 0.035\\
+BALD&0.594 $\pm$ 0.039\\
+\end{tabular}
+\end{minipage}
+%USPS Enc
+\begin{minipage}{0.65\linewidth}
 \begin{figure}[H]
     \centering
-\includegraphics[width=0.4\linewidth]{img/eval_topv2}
-\includegraphics[width=0.4\linewidth]{img/eval_news} \\ [2mm]
-\includegraphics[width=0.3\linewidth]{img/micro_topv2.jpg}
-\includegraphics[width=0.3\linewidth]{img/micro_news.jpg} \\ [4mm]
+    \includegraphics[width=\linewidth]{img/eval_usps_enc}\\ [2mm]
+    \includegraphics[width=\linewidth]{img/micro_usps_enc.jpg} 
+\end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&USPSEncoded\\
+\hline
+Oracle&0.522 $\pm$ 0.021\\
+TypiClust&0.507 $\pm$ 0.025\\
+Margin&0.496 $\pm$ 0.030\\
+Random&0.468 $\pm$ 0.025\\
+Entropy&0.459 $\pm$ 0.021\\
+Badge&0.440 $\pm$ 0.026\\
+Coreset&0.435 $\pm$ 0.027\\
+BALD&0.402 $\pm$ 0.052\\
+\end{tabular}
+\end{minipage}
+%Cifar10
+\begin{minipage}{0.65\linewidth}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{img/eval_cifar10}\\ [2mm]
+    \includegraphics[width=\linewidth]{img/micro_cifar10.jpg}
+\end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&Cifar10\\
+\hline
+Oracle&0.500 $\pm$ 0.010\\
+Badge&0.453 $\pm$ 0.012\\
+Coreset&0.453 $\pm$ 0.009\\
+Margin&0.451 $\pm$ 0.010\\
+Random&0.450 $\pm$ 0.012\\
+Entropy&0.449 $\pm$ 0.010\\
+TypiClust&0.436 $\pm$ 0.010\\
+BALD&0.436 $\pm$ 0.010\\
+\end{tabular}
+\end{minipage}
+%FMnist
+\begin{minipage}{0.65\linewidth}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{img/eval_fmnist}\\ [2mm]
+    \includegraphics[width=\linewidth]{img/micro_fmnist.jpg}
+\end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&FashionMnist\\
+\hline
+Oracle&0.862 $\pm$ 0.003\\
+BALD&0.854 $\pm$ 0.003\\
+Margin&0.851 $\pm$ 0.003\\
+Badge&0.851 $\pm$ 0.003\\
+Entropy&0.847 $\pm$ 0.004\\
+TypiClust&0.846 $\pm$ 0.004\\
+Coreset&0.840 $\pm$ 0.004\\
+Random&0.837 $\pm$ 0.004\\
+\end{tabular}
+\end{minipage}
+%Cifar10 Encoded
+\begin{minipage}{0.65\linewidth}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{img/eval_cifar10_enc} \\ [2mm]
+    \includegraphics[width=\linewidth]{img/micro_cifar10_enc.jpg} 
+\end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&Cifar10Encoded\\
+\hline
+Oracle&0.714 $\pm$ 0.007\\
+Entropy&0.654 $\pm$ 0.013\\
+Coreset&0.653 $\pm$ 0.012\\
+Badge&0.653 $\pm$ 0.012\\
+BALD&0.650 $\pm$ 0.016\\
+Margin&0.647 $\pm$ 0.012\\
+TypiClust&0.636 $\pm$ 0.009\\
+Random&0.607 $\pm$ 0.013\\
+\end{tabular}
+\end{minipage}
+%FashionMnist Encoded
+\begin{minipage}{0.65\linewidth}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{img/eval_fmnist_enc} \\ [2mm]
+\includegraphics[width=\linewidth]{img/micro_fmnist_enc.jpg} 
+\end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&FashionMnistEncoded\\
+\hline
+Oracle&0.732 $\pm$ 0.006\\
+Coreset&0.686 $\pm$ 0.008\\
+Badge&0.685 $\pm$ 0.008\\
+Entropy&0.684 $\pm$ 0.009\\
+Margin&0.682 $\pm$ 0.011\\
+BALD&0.668 $\pm$ 0.009\\
+TypiClust&0.652 $\pm$ 0.009\\
+Random&0.640 $\pm$ 0.011\\
+\end{tabular}
+\end{minipage}
+%TopV2
+\begin{minipage}{0.65\linewidth}
+\begin{figure}[H]
+    \centering
+\includegraphics[width=\linewidth]{img/eval_topv2} \\[2mm]
+\includegraphics[width=\linewidth]{img/micro_topv2.jpg}
+\end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&TopV2\\
+\hline
+Oracle&0.862 $\pm$ 0.006\\
+BALD&0.831 $\pm$ 0.013\\
+Badge&0.826 $\pm$ 0.015\\
+Coreset&0.823 $\pm$ 0.016\\
+Margin&0.822 $\pm$ 0.015\\
+TypiClust&0.805 $\pm$ 0.015\\
+Entropy&0.801 $\pm$ 0.025\\
+Random&0.787 $\pm$ 0.015\\
+\end{tabular}
+\end{minipage}
+%News
+\begin{minipage}{0.65\linewidth}
+\begin{figure}[H]
+    \centering
+\includegraphics[width=\linewidth]{img/eval_news} \\[2mm]
+\includegraphics[width=\linewidth]{img/micro_news.jpg}
+\end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&News\\
+\hline
+Oracle&0.502 $\pm$ 0.005\\
+Margin&0.427 $\pm$ 0.007\\
+BALD&0.421 $\pm$ 0.008\\
+Badge&0.420 $\pm$ 0.011\\
+Entropy&0.416 $\pm$ 0.010\\
+Coreset&0.415 $\pm$ 0.011\\
+Random&0.409 $\pm$ 0.008\\
+TypiClust&0.385 $\pm$ 0.010\\
+\end{tabular}
+\end{minipage}
+%ThreeClust
+\begin{minipage}{0.65\linewidth}
+\begin{figure}[H]
+    \centering
+\includegraphics[width=\linewidth]{img/eval_threeclust.pdf} \\[2mm]
+\includegraphics[width=\linewidth]{img/micro_threeclust.jpg}
+\end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&ThreeClust \\
+\hline
+Oracle&0.722 $\pm$ 0.097\\
+Coreset&0.698 $\pm$ 0.058\\
+TypiClust&0.697 $\pm$ 0.055\\
+Entropy&0.682 $\pm$ 0.098\\
+Random&0.672 $\pm$ 0.067\\
+Margin&0.669 $\pm$ 0.095\\
+Badge&0.524 $\pm$ 0.086\\
+BALD&0.507 $\pm$ 0.050\\
+\end{tabular}
+\end{minipage}
+%DivergingSin
+\begin{minipage}{0.65\linewidth}
+\begin{figure}[H]
+    \centering
+\includegraphics[width=\linewidth]{img/eval_divergingsin.pdf} \\[2mm]
+\includegraphics[width=\linewidth]{img/micro_divsin.jpg}
+\end{figure}
+\end{minipage}
+\begin{minipage}{0.29\linewidth}
+\begin{tabular}{c|c}
+&DivergingSin\\
+\hline
+Oracle&0.948 $\pm$ 0.198\\
+Entropy&0.936 $\pm$ 0.202\\
+TypiClust&0.930 $\pm$ 0.196\\
+Margin&0.929 $\pm$ 0.201\\
+Random&0.919 $\pm$ 0.191\\
+Badge&0.914 $\pm$ 0.202\\
+Coreset&0.914 $\pm$ 0.197\\
+BALD&0.661 $\pm$ 0.167\\
+\end{tabular}
+\end{minipage}
+
+\section{Seeding Strategy}\label{app:seeding_strategy}
+We aim to provide an experimental setup that is fully reproducible independent of the dataset, classification model, or AL algorithm used.
+For a fair comparison of two AL algorithms, both algorithms need to receive equal starting conditions in terms of train/validation split, initialization of classifier, and even the state of minor systems like the optimizer or mini-batch sampler.
+Even though different implementations might have their own solution to some of these problems, only \cite{ji2023randomness} has described and implemented a fully reproducible pipeline for AL evaluation.
+The term reproducibility in this work is used as a synonym not only for the reproducibility of an experiment (a final result given a seed), but also the reproducibility of all subsystems independent of each other.
+The seed for one subsystem should always reproduce the behavior of this subsystem independent of all other subsystems and their seeds.
+The main obstacle for ensuring reproducibility is the seeding utility in PyTorch, Tensorflow, and other frameworks, whose default choice is a single global seed.
+Since many subsystems draw random numbers from this seed, all of them influence each other to a point where a single additional draw can completely change the model initialization, data split or the order of training batches.
+Even though some workarounds exist, e.g. re-setting the seed multiple times, this problem is not limited to the initialization phase, but also extends to the AL iterations and the systems within.
+We propose an implementation that creates separate Random Number Generators (RNGs) for each of these systems to ensure equal testing conditions even when the AL algorithm, dataset, or classifier changes.
+We hypothesize that the insufficient setup with global seeds contributes to the ongoing problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
+In summary, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini-batch sampling, and $s_\theta$ for model initialization and sampling of dropout masks.
+Unless stated otherwise, we will keep $s_\Omega$ fixed, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
+Some algorithms require a subsample to be drawn from $\mathcal{U}$ in order to reduce the computational cost in each iteration, while others need access to the full unlabeled pool (e.g.
+for effective clustering).
+If a subsample is required, it will be drawn from $s_\Omega$ and therefore will not influence other systems in the experiments.
+For each algorithm, we decided if subsampling is required based on our available hardware, but decided against setting a fixed time limit per experiment, since this would introduce unnecessary complexity into the benchmark.
+An overview of selected hyperparameters per AL algorithm can be found in Appendix \ref{app:agent_hyperparameters}. \\
+\textbf{Note:} Even though we decoupled the subsystems via the described seeds, the subsystems can still influence each other in a practical sense. 
+For example, keeping $s_\mathcal{D}$ fixed does not mean that always the same sequence of samples from $\mathcal{U}$ (if subsamples are drawn) are shown to all acquisition functions. 
+This is practically impossible, as different acquisition functions pick different $x^{(i)}$.
+However, the hypothetical \textbf{tree} of all possible sequences of samples from $\mathcal{U}$ remains the same, granting every acquisition function equal possibilities.
+
+\section{Number of Restarts Ablation for "Training from Scratch"}\label{app:restarts_from_scratch}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{img/ablation_cifar_restarts}
+    \caption{Created by the protocol described in Sec.
+\ref{sec:evaluation} for the Encoded Cifar10 dataset, with the classifier being trained from scratch.
+We observe similar characteristics, where a low number of restarts leads to extreme variance in the reported results.}
 \end{figure}
 
-\section{Alternative Plot for Restarts Ablation}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\linewidth]{img/ablation_restarts_2}
-\end{figure}
+\section{Oracle Curve Forecasting}\label{app:oracle_forecasting}
+TODO
 
@@
 \begin{table}[H]
+    \caption{Selected hyperparameters for all tested acquisition functions. Last column indicates the source of our implementation.}
 	\centering
-	\begin{tabular}{l || l | l}
-		Algorithm & Sample Size & \\
+	\begin{tabular}{l || l | l | l}
+		Algorithm & \makecell{Sample\\Size} & Other & Source\\
 		\hline
-		BADGE & 100 & \\
-		BALD & 100 & Dropout Trials: 5 \\
-		Coreset & 8000 & \\
-		TypiClust & 10000 & Min Cluster Size: 5, Max \# Clusters: 500 \\
-		Margin & 8000 & \\
-		Entropy & 8000 & \\
+		BADGE & 100 & & Based on \cite{ashdeep, curelab}\\
+		BALD & 100 & Dropout Trials: 5 & Based on \cite{pycls} \\
+		Coreset & 8000 & & Own \\
+		TypiClust & 10000 & \makecell[tl]{Min Cluster Size: 5\\Max \# Clusters: 500} & Based on \cite{hacohen2022active} \\
+		Margin & 8000 & & Own\\
+		Entropy & 8000 &  & Own\\
 	\end{tabular}
@@
 
-\section{Hyperparameters per Dataset}\label{app:hyperparameters}
+\section{Hyperparameters and Preprocessing per Dataset}\label{app:hyperparameters}
+For all our datasets we use the pre-defined train/test splits, if given. 
+In the remaining cases, we define test sets upfront and store them into separate files to keep them fixed across all experiments.
+The validation set is split in the experiment run itself and depends on the dataset-seed.\\
+\textbf{Tabular:}
+We use \textbf{Splice}, \textbf{DNA} and \textbf{USPS} from LibSVMTools \cite{libsvmtools}.
+All three datasets are normalized between [0, 1]. \\
+\textbf{Image:}
+We use \textbf{FashionMNIST} \cite{xiao2017fashion} and \textbf{Cifar10} \cite{krizhevsky2009learning}, since both are widely used in AL literature.
+Both datasets are normalized according to their standard protocols. \\
+\textbf{Text:}
+We use \textbf{News Category} \cite{misra2022news} and \textbf{TopV2} \cite{chen-etal-2020-low-resource}.
+For News Category we use  the 15 most common categories as indicated by its Kaggle site.
+We additionally drop sentences above 80 words to reduce the padding needed (retaining 99,86\% of the data).
+For TopV2, we are only using the "alarm" domain.
+Both datasets are encoded with pre-trained GloVe (Common Crawl 840B Tokens) embeddings \cite{pennington2014glove}.
+Since neither dataset provided a fixed test set, we randomly split 7000 datapoints into a test set.
+%
 \begin{table}[H]
@@
 	\caption{Size of the seed set is given by number of labeled sample per class.}
-	\label{tab:architectures}
+	\label{tab:architecture_hps}
 \end{table}
@@
 
-\section{Comparison of Different Classifier Sizes}\label{app:classifier_size}
-We tested two different classifier sizes in Splice and DNA:
-\begin{itemize}
-	\item Small: [24, 12] (2400 parameters)
-	\item Big: [24, 48, 48] (5700 parameters)
-\end{itemize}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.49\linewidth]{img/supp_dna_small.png}
-	\includegraphics[width=0.49\linewidth]{img/supp_dna_big.png}
-	\includegraphics[width=0.49\linewidth]{img/supp_micro_dna_small.png}
-	\includegraphics[width=0.49\linewidth]{img/supp_micro_dna_big.png}
-	\caption{Comparison of small and big classifiers for the DNA dataset}
-\end{figure}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.49\linewidth]{img/supp_splice_small.png}
-	\includegraphics[width=0.49\linewidth]{img/supp_splice_big.png}
-	\includegraphics[width=0.49\linewidth]{img/supp_micro_splice_small.png}
-	\includegraphics[width=0.49\linewidth]{img/supp_micro_splice_big.png}
-	\caption{Comparison of small and big classifiers for the Splice dataset}
-\end{figure}
-Even though the ranking between the two sizes of classifier does change, we observed no systematic that would favor one algorithm. 
-We therefore rely on the different datasets to provide classifiers of varying size and architecture.
-
-\section{AL Pseudocode}
+\section{AL Pseudocode}\label{app:pseudocode}
 \begin{algorithm}[H]
-	\caption{Active Learning}\label{alg:active_learning}
+	\caption{Active Learning Loop}\label{alg:active_learning}
 	\begin{algorithmic}[1]
@@
 
+\begin{algorithm}[H]
+\caption{Acquire Oracle $\Omega$}\label{alg:oracle}
+\begin{algorithmic}[1]
+    \Require $\mathcal{U}, \mathcal{L}, A, \mathcal{D}_\text{test}, \tau, \hat y_\theta$ 
+    \Require Train, Margin, Acc
+    \State $\text{acc}^0 \gets \text{acc}^* \gets \operatorname{Acc}(\mathcal{D}_\text{test}, \hat y_\theta)$ 
+    \For{$k := 1 \ldots \tau$}
+        \State $u_k = \operatorname{unif}(\U)$
+    \State $\mathcal{L}' \gets \mathcal{L}^{(i)} \cup \{(u_k, A(u_k))\}$
+    \State $\hat y'_\theta \gets \operatorname{Train}(\mathcal{L}', \hat y_\theta)$
+        \State $\text{acc}' \gets \text{Acc}(\D_\test, \hat y'_\theta)$
+    %\State $r \gets \text{acc}' - \text{acc}$
+    \If{$\text{acc}' > \text{acc}^*$} 
+    \State $\text{acc}^* \gets \text{acc}'$
+    \State $u^* \gets u_k$
+    \EndIf
+    \EndFor
+    \If{$\text{acc}^0 = \text{acc}^*$}
+    \State $u^* \gets \operatorname{Margin}(\mathcal{U}, \hat y_\theta)$
+    \EndIf
+    \Return $u^*$
+\end{algorithmic}
+\end{algorithm}
+Alg. \ref{alg:oracle} replaces the acquisition function $\Omega$ in the AL loop (Alg. \ref{app:pseudocode} line 5).
+
 \end{document}

### 2024-05-31 doc/al_benchmark_paper.tex
paper

@@
 	\label{tab:benchmark_comparison}
+	\resizebox{\columnwidth}{!}{
 \begin{tabular}{lccccllllcl}
@@
 \end{tabular}
-
+}
 \end{table}
@@
 	\caption{Random draws from a pool of 100 runs for margin sampling on the Splice dataset with different numbers of repetitions ($\alpha=\{5,10,50\}$).
-Green curves signify the true mean performance of all 100 runs, while the samples are blue.
+Green curves are the mean performance of all 100 runs, while the samples are blue.
 Even with 5 or 10 repetitions, we can observe that single draws for margin sampling display below-random performance (black), while the true mean should be above random.}
@@
 To test for these shortcomings, we created two synthetic datasets, namely "Honeypot" and "Diverging Sin", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively.
-To avoid algorithms memorizing these datasets they are generated during the experiment, depending on $s_\D$. \\ [1mm]
+To avoid algorithms memorizing these datasets they are generated during the experiment, depending on $s_\D$. \\
 \begin{figure}[H]
@@
 	\includegraphics[width=0.4\linewidth]{img/honeypot.jpg}
-	\caption{Our two synthetic datasets.}
+	\caption{Synthetic "Honeypot" and "Diverging Sine" datasets. The decision boundary is not part of the dataset and serves only as a visual guide.}
         \label{fig:synthDataAppendix}
@@
 \subsection{Classification Model}\label{sec:choosing_the_classifier}
-\begin{wraptable}{R}{0.55\textwidth}
+\begin{wraptable}{R}{0.65\textwidth}
     \vspace{-0.7cm}
@@
     \label{tab:batch_sizes}
-    \begin{tabular}{l|c|c|c|c|c|c|c}
-                    & 1 & 5 & 20 & 50 & 100 & 500 & 1000 \\
-         \hline
-        Enc. DNA    & o & o &&&&& \\
-        Enc. Splice & o & o & o & o &&&\\
-        TopV2       & o & o & o & o &&& \\
-        Splice      & o & o & o & o & o && \\
-        DNA         & o & o & o & o & o && \\
-        USPS        & o & o & o & o & o && \\
-        Enc. USPS   & o & o & o & o & o && \\
-        Enc. Cifar10& o & o & o & o & o && \\
-        Enc. FMnist & o & o & o & o & o && \\
-        News        &&& o & o & o & o &\\
-        FMnist      &&&&&& o & o\\
-        Cifar10     &&&&&& o & o \\
+    \begin{tabular}{l|c|c|c|c|c|c|c|c}
+                    & B & 1 & 5 & 20 & 50 & 100 & 500 & 1000 \\
+         \hline
+        Enc. DNA    & 40 & o & o &&&&& \\
+        Honeypot    & 60 & o & o &&&&& \\
+        Diverging Sine& 60 & o & o &&&&& \\
+        Enc. Splice & 100 & o & o & o & o &&&\\
+        TopV2       & 200 & o & o & o & o &&& \\
+        Splice      & 400 & o & o & o & o & o && \\
+        DNA         & 300 & o & o & o & o & o && \\
+        USPS        & 400 & o & o & o & o & o && \\
+        Enc. Cifar10& 450 & o & o & o & o & o && \\
+        Enc. FMnist & 500 & o & o & o & o & o && \\
+        Enc. USPS   & 600 & o & o & o & o & o && \\
+        News        & 3K &&& o & o & o & o &\\
+        FMnist      & 10K &&&&&& o & o\\
+        Cifar10     & 10K &&&&&& o & o \\
     \end{tabular}
@@
 In this work, we use the fine-tuning method for un-encoded datasets to save computational time, while we use the from-scratch training for embedded datasets since they have very small classifiers and this approach generally produces better results.
-Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping with a patience of 2 afterwards.\\
+Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping with a patience of 2 afterwards.
 
@@
 Every encoded dataset is classified by a single linear layer with softmax activation.
-For an overview of architectures and hyperparameters please refer to Appendix \ref{app:hyperparameters}. \\
+For an overview of architectures and hyperparameters please refer to Appendix \ref{app:hyperparameters}.
 
@@
 \begin{table}[]
-	\caption{Performances for acquisition functions on each dataset and aggregated for un-encoded and encoded datasets.
+	\caption{Performances for acquisition functions on real-world datasets, aggregated for un-encoded and encoded datasets.
 Performance is shown as average ranks over restarts (1.0 is the best rank).
@@
 For brevity, we only display per-dataset standard deviations for Splice.
-All standard deviations lie within $[0.001, 0.024]$.}
+All standard deviations lie within $[0.0007, 0.024]$.}
 	\label{tab:results}
 	\centering
-        %\resizebox{\columnwidth}{!}{
+        \resizebox{\columnwidth}{!}{
 \begin{tabular}{l|lllllll|ll}
@@
           \hline
-Oracle & 1.0$\pm$0.011 & 1.0 & 1.0 & 3.9 & 7.3 & 1.0 & 1.0 & 2.3 & 1.9 \\
-Margin & 3.9$\pm$0.019 & 4.2 & \textbf{2.1} & 4.7 & \textbf{2.9} & \textbf{2.4} & 2.7 & 3.3 & 3.9 \\
-Badge & \textbf{3.0$\pm$0.014} & 6.2 & 2.9 & \textbf{3.5} & 3.2 & 3.3 & \textbf{2.5} & 3.5 & 4.9 \\
-DSA & 7.1$\pm$0.016 & 7.1 & 7.1 & 3.8 & 3.5 & 5.7 & 5.9 & 5.8 & 6.2 \\
-CoreGCN & 6.6$\pm$0.012 & 4.8 & 9.4 & 6.1 & 4.8 & 4.0 & 5.4 & 5.9 & 7.7 \\
-BALD & 3.9$\pm$0.011 & 4.6 & 5.3 & 10.6 & 5.8 & 7.2 & 3.8 & 5.9 & 6.9 \\
-Entropy & 6.4$\pm$0.018 & \textbf{3.8} & 7.2 & 6.0 & 3.3 & 9.1 & 8.2 & 6.3 & 6.0 \\
-LSA & 5.8$\pm$0.009 & 6.6 & 5.2 & 6.1 & 9.2 & 7.0 & 6.0 & 6.6 & 6.9 \\
-Coreset & 6.8$\pm$0.01 & 8.8 & 9.6 & 5.2 & 5.3 & 8.0 & 9.4 & 7.6 & 6.6 \\
-Random & 8.6$\pm$0.01 & 9.0 & 5.2 & 6.9 & 9.7 & 7.4 & 6.6 & 7.6 & 6.4 \\
-TypiClust & 8.3$\pm$0.009 & 9.8 & 11.0 & 6.4 & 9.6 & 11.0 & 10.6 & 9.5 & 8.6
-\end{tabular}
-        %}
+Oracle 	& 1.0$\pm$0.011 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.9 \\
+Margin 	& 3.9$\pm$0.019 & 4.2 & \textbf{2.1} & 4.4 & \textbf{2.9} & \textbf{2.4} & 2.7 & 3.2 & 3.9 \\
+Badge  	& \textbf{3.0$\pm$0.014} & 6.2 & 2.9 & \textbf{3.4} & 3.2 & 3.3 & \textbf{2.5} & 3.5 & 4.9 \\
+DSA 	& 7.1$\pm$0.016 & 7.1 & 7.1 & 3.6 & 3.5 & 5.7 & 5.9 & 5.7 & 6.2 \\
+BALD 	& 3.9$\pm$0.011 & 4.6 & 5.3 & 10.1 & 5.8 & 7.2 & 3.8 & 5.8 & 6.9 \\
+CoreGCN & 6.6$\pm$0.012 & 4.8 & 9.4 & 5.7 & 4.8 & 4.0 & 5.4 & 5.8 & 7.7 \\
+Entropy & 6.4$\pm$0.018 & \textbf{3.8} & 7.2 & 5.7 & 3.4 & 9.1 & 8.2 & 6.3 & 6.0 \\
+LSA 	& 5.8$\pm$0.009 & 6.6 & 5.2 & 5.8 & 8.7 & 7.0 & 6.0 & 6.4 & 6.9 \\
+Random 	& 8.6$\pm$0.01 & 9.0 & 5.2 & 6.5 & 9.2 & 7.4 & 6.6 & 7.5 & 6.4 \\
+Coreset & 6.8$\pm$0.01 & 8.8 & 9.6 & 4.9 & 5.3 & 8.0 & 9.4 & 7.5 & 6.6 \\
+TypiClust & 8.3$\pm$0.009 & 9.8 & 11.0 & 6.0 & 9.1 & 11.0 & 10.6 & 9.4 & 8.6
+\end{tabular}
+       	}
 \end{table}

### 2024-05-31 core/helper_functions.py

@@ -177,13 +177,14 @@ def _insert_oracle_forecast(df:pd.DataFrame):
         full_range = df[df["dataset"] == dataset]["iteration"].unique().tolist()
         oracle_max = oracle_data["iteration"].max()
         if oracle_max < max(full_range):
+            _, upper_bound, _ = _load_eval_data(dataset, None, "UpperBound")
             oracle_data = oracle_data.sort_values(["iteration"], ascending=True)
             # construct prediction horizon
             oracle_iterations = df_oracles["iteration"].unique().tolist()
             oracle_stepsize = round(float(max(oracle_iterations) - min(oracle_iterations)) / len(oracle_iterations), 0)
             testing_range = range(max(oracle_iterations), max(full_range), int(oracle_stepsize))
             x, y = _get_oracle_regression(x=oracle_data["iteration"].to_list(), y=oracle_data["auc"].to_list(),
-                                          x_test=testing_range, upper_bound=None)
+                                          x_test=testing_range, upper_bound=upper_bound)
             for trial in df_oracles["trial"].unique().tolist():
                 forecast_data = pd.DataFrame({
                     "dataset": [dataset]*len(x),
@@ -198,6 +199,7 @@ def _insert_oracle_forecast(df:pd.DataFrame):
 
 
 def _get_oracle_regression(x, y, x_test, upper_bound):
+    upper_bound = upper_bound[0]
     max_value = float(max(x_test))
     cutoff = int(len(x)*0.5)
     x_train = np.array(x[cutoff:], dtype=float) / max_value # drop the first 50 percent of data
@@ -205,13 +207,18 @@ def _get_oracle_regression(x, y, x_test, upper_bound):
     lr_model = LinearRegression()
     lr_model.fit(x_train.reshape(-1, 1), y_train)
     x_test = [i for i in x_test if i >= max(x)]
-    x_test_array = np.array(x_test, dtype=float) / max_value
-    # x_test = np.arange(max(x), max(x_test))
+    # x_test_array = np.array(x_test, dtype=float) / max_value
+    x_test_array = np.arange(max(x), max(x_test)+1) / max_value
     y_test = lr_model.predict(x_test_array.reshape(-1, 1))
-    for i in range(1, len(y_test)):
-        multiplicator = 1 / i
[…]

### 2024-05-31 results/cd_diagram.py

@@ -429,18 +429,18 @@ if __name__ == '__main__':
     # df = prepare_df(df)
     # draw_cd_diagram(df, title="Tabular", alpha=10000.0, file="doc/img/macro_vector.jpg")
 
-    df = combine_agents_into_df(["TopV2", "News"], include_oracle=True)
+    # df = combine_agents_into_df(["TopV2", "News"], include_oracle=True)
+    # df = prepare_df(df)
+    # draw_cd_diagram(df, title="Text", alpha=10000.0, file="doc/img/macro_text.jpg")
+
+    df = combine_agents_into_df(["Cifar10", "FashionMnist"], include_oracle=True)
     df = prepare_df(df)
-    draw_cd_diagram(df, title="Text", alpha=10000.0, file="doc/img/macro_text.jpg")
+    draw_cd_diagram(df, title="Image", alpha=10000.0, file="doc/img/macro_img.jpg")
 
-    # df = combine_agents_into_df(["Cifar10", "FashionMnist"], include_oracle=True)
-    # df = prepare_df(df)
-    # draw_cd_diagram(df, title="Image", alpha=10000.0, file="doc/img/macro_img.jpg")
-
-    # df = combine_agents_into_df(["Cifar10Encoded", "DNAEncoded", "FashionMnistEncoded",
-    #                                      "SpliceEncoded", "USPSEncoded"], include_oracle=True)
-    # df = prepare_df(df)
-    # draw_cd_diagram(df, title="Encoded", alpha=10000.0, file="doc/img/macro_enc.jpg")
+    df = combine_agents_into_df(["Cifar10Encoded", "DNAEncoded", "FashionMnistEncoded",
+                                         "SpliceEncoded", "USPSEncoded"], include_oracle=True)
+    df = prepare_df(df)
+    draw_cd_diagram(df, title="Encoded", alpha=10000.0, file="doc/img/macro_enc.jpg")
 
     # Single Dataset
     # df = combine_agents_into_df(dataset="Splice", max_loaded_runs=50)

### 2024-05-31 doc/al_benchmark_paper.tex
least confident baseline

@@
 
-% Review point
-% - Restarting process seems to be hard to understand (probably in terms of the seeding)
-% - Justification for datasets and algorithms
+% TODOs nach Priorität
+% !! Least-Confident Baseline
+% !! "Overfitting on test" für Oracle: Sanity-Check und eventuell komplett streichen
+% !! Ergebnis-Tabelle mit nur 3 runs neu berechnen und Unterschiede zeigen
+% !! Introduction neu schreiben: Wir brauchen eine klare Benennung der Ziele und bessere Abgrenzung zu anderen papern
+% Advantage-Matrix von Ji et. al. neu aufarbeiten. Wo sind die Unterschiede? Ist das CD-Diagramm besser/schlechter?
+% Eventuell Batchsize 1 von den anderen Ergebnissen separieren. Der Vergleich macht mehr Sinn, als einfach die Batchsizes zu mitteln
+% Infos und Ergebnisse für die syntetischen Datensätze in ein Kapitel bündeln und direkt auf die Contribution beziehen
+% Oracle Performance mit einer Zahl benennen: z.B. Da wird ~20% lift zu allen Baselines haben, brauchen wir kein aufwendigeres Oracle
+% Contributions neu sortieren und sehr viel ausführlicher beschreiben
+% Unsere Rankings mit Literatur vergleichen. Wie anders sind wir? -> Evtl. besten Alg. pro Benchmark sammeln und vergleichen
+% Causalitäten checken: Viele Dinge werden einfach genannt, ohne das der Grund richtig eingeführt wird
+% "reproducible" als Aufhänger für 50 repeats ist nicht intuitiv
+% Fig. 1: Schlechte Performance am Anfang verwirrt nur. Entweder BADGE benutzen, anders plotten, oder im Text in Kontext setzen
+% paired-t-test im Critical Difference Diagram herausstellen und die Relevanz für paper zeigen
+% Eventuell Honeypot und Diverging Sine direkt in den Contributions nennen
+% Oracle Forecast im Appendix erklären 
+% Tabelle mit AUCs in den Appendix
 
@@
 \subsection{Why we need 50 restarts}\label{sec:restarts}
-To evaluate how many restarts are necessary to obtain conclusive and reproducible results in an AL experiment, we set the number of restarts of the margin sampling baseline for the Splice dataset to 100.
+To evaluate how many restarts are necessary to obtain conclusive and reproducible results in an AL experiment, we set the number of restarts fore margin sampling on the Splice dataset to 100.
 This allows us firstly, to obtain a very strong estimation of the "true" average performance of margin sampling on this dataset and secondly, to draw subsets from this pool of 100 runs.
@@
 We describe a novel seeding strategy in Appendix \ref{app:seeding_strategy} that creates 3 separate Random Number Generators (RNG) based on 3 different seeds.
-In summary, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini-batch sampling, and $s_\theta$ for model initialization and sampling of dropout masks.
+In short, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini-batch sampling, and $s_\theta$ for model initialization and sampling of dropout masks.
 Unless stated otherwise, we will keep $s_\Omega$ fixed, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
-While this seeding strategy is capable of controlling the amount variance in the experiment, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), reducing its value in practice and providing a bad approximation of the quality of the AL algorithm.
+While this seeding strategy is capable of controlling the amount variance in the experiment, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), reducing its value in practice and providing a bad approximation of the quality of an AL algorithm.
 Hence, we opt for letting the subsystems vary (by increasing $s_\mathcal{D}$ and $s_\theta$) and combine that with a high number of restarts to obtain a good average of the generalization performance of each AL algorithm.
@@
 We attached a linear classification head to the encoder, fine-tuned it to the data and evaluated this classifier for test accuracy, mirroring our AL setup for embedded datasets. 
-The checkpoints for each encoder model will be provided together with the framework. \\ [1mm]
+The checkpoint of each encoder model will be provided together with the framework. \\ [1mm]
 \textbf{Synthetic Data}
@@
 Both types have principled shortcomings in terms of the utilized information that makes them unsuitable for certain data distributions.
-To test for these shortcomings, we created two synthetic datasets, namely "Honeypot" and "Diverging Sin", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively.
+To test for these shortcomings, we created two synthetic datasets, namely "Honeypot" and "Diverging Sine", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively.
 To avoid algorithms memorizing these datasets they are generated during the experiment, depending on $s_\D$. \\
@@
 Honeypot creates to two easy to distinguish clusters with 150 samples each and one overlapping "honeypot" that represents a noisy region of the dataset with potentially miss-labeled, miss-measured or generally adverse samples.
-This honeypot contains 150 of each class, creating a balance of 50\% beneficial samples and 50\% adverse samples in the dataset.
+This honeypot contains 150 samples of each class, creating a balance of 50\% beneficial samples and 50\% adverse samples in the dataset.
 The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its negative impact on purely uncertainty based acquisition functions.

### 2024-05-31 agents/__init__.py

@@ -2,6 +2,7 @@ from agents.coreset import Coreset_Greedy
 from agents.margin import MarginScore
 from agents.random_agent import RandomAgent
 from agents.shannon_entropy import ShannonEntropy
+from agents.least_confident import LeastConfident
 from agents.bald import BALD
 from agents.badge import Badge
 from agents.typiclust import TypiClust

### 2024-05-31 core/helper_functions.py

@@ -411,6 +411,8 @@ def get_agent_by_name(name:str)->Callable:
         return agents.ShannonEntropy
     elif name == "margin":
         return agents.MarginScore
+    elif name == "leastconfident":
+        return agents.LeastConfident
     elif name == "coreset":
         return agents.Coreset_Greedy
     elif name == "coresetraw":

### 2024-05-31 evaluate.py

@@ -15,7 +15,7 @@ parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--agent_seed", type=int, default=1)
 parser.add_argument("--pool_seed", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
-parser.add_argument("--agent", type=str, default="margin")
+parser.add_argument("--agent", type=str, default="leastconfident")
 parser.add_argument("--dataset", type=str, default="usps")
 parser.add_argument("--query_size", type=int, default=1)
 parser.add_argument("--encoded", type=int, default=1)

### 2024-06-02 doc/al_benchmark_paper.tex
paper und ergebnisse

@@
 
+% Introduction:
+% What is the problem?
+% - Ergebnisse schwierig zu reproduzieren
+% - Vorherige Benchmarks haben dem Versuchsaufbau zu wenig Beachtung geschenkt - erst Ji und Luth haben verlässliche Benchmarks beschrieben
+% - keiner (!) der Benchmarks hat beide unserer Top-2 Methoden (Margin und Badge)
+% - Es werden maximal zwei Domänen abgedeckt
+% - Bisherige Oracle-Algorithmen sind extrem rechenaufwenig
+% What are we adding to it?
+% - Wir implementieren die Guidelines von Ji et al
+% - Wir erweitern die Anzahl der Domänen von 1 auf 5 und führen mehr Wiederholungen ein
+% - Bessere Abdeckung der relevanten Algorithmen (von 8 (Ji) auf 11)
+% - Unser Oracle ist viel leichter zu berechnen
+% Why is that important?
+% - Wir zeigen, dass die Wiederholungen wichtig sind, um reproduzierbare Ergebnisse zu liefern
+% - Wir zeigen, dass es signifikante Unterschiede zwischen den Domänen gibt
+% - Wir zeigen, dass Margin und Badge zwei starke Baselines sind, die immer dabei sein müssen
+% - Unser Oracle hat x% lift über alle Algorithmen und ist somit ausreichend
+
 % TODOs nach Priorität
-% !! Least-Confident Baseline
-% !! "Overfitting on test" für Oracle: Sanity-Check und eventuell komplett streichen
-% !! Ergebnis-Tabelle mit nur 3 runs neu berechnen und Unterschiede zeigen
+% !! "Overfitting on test" für Oracle: Sanity-Check durchführen und eventuell komplett streichen
 % !! Introduction neu schreiben: Wir brauchen eine klare Benennung der Ziele und bessere Abgrenzung zu anderen papern 
+% !! Sec 4 Methodology ist sehr chaotisch. Wir müssen neu sortieren und unwichtige Dinge streichen
+% Contributions neu sortieren und ausführlicher beschreiben
+% Methode zum CD-Diagramm genau erklären (paired-t-test etc.) und die Relevanz fürs paper zeigen
 % Advantage-Matrix von Ji et. al. neu aufarbeiten. Wo sind die Unterschiede? Ist das CD-Diagramm besser/schlechter?
+% Gibt es eine Conclusion für die Encoded-Datasets? Im Moment werden die nicht besprochen
 % Eventuell Batchsize 1 von den anderen Ergebnissen separieren. Der Vergleich macht mehr Sinn, als einfach die Batchsizes zu mitteln
 % Infos und Ergebnisse für die syntetischen Datensätze in ein Kapitel bündeln und direkt auf die Contribution beziehen
+% Tabelle 1 für Algorithmen statt Domänen. Das zeigt, dass immer entweder BADGE oder Margin (unsere Top 2) fehlt
 % Oracle Performance mit einer Zahl benennen: z.B. Da wird ~20% lift zu allen Baselines haben, brauchen wir kein aufwendigeres Oracle
-% Contributions neu sortieren und sehr viel ausführlicher beschreiben
 % Unsere Rankings mit Literatur vergleichen. Wie anders sind wir? -> Evtl. besten Alg. pro Benchmark sammeln und vergleichen
@@
 % Fig. 1: Schlechte Performance am Anfang verwirrt nur. Entweder BADGE benutzen, anders plotten, oder im Text in Kontext setzen
-% paired-t-test im Critical Difference Diagram herausstellen und die Relevanz für paper zeigen
 % Eventuell Honeypot und Diverging Sine direkt in den Contributions nennen
@@
 % Tabelle mit AUCs in den Appendix
+
+%%%%%%%%%%%%%%%%%%
+% DONE
+%%%%%%%%%%%%%%%%%
+% !! Ergebnis-Tabelle mit nur 3 runs neu berechnen und Unterschiede zeigen
+% !! Least-Confident Baseline
+% Conclusion: Es gibt Unterschiede bei der Top 1 zwischen den Domänen, aber die Top 3 ist konstant
 
@@
 \multicolumn{1}{l|}{Luth et al. \cite{luth2024navigating}} & batch & 4 & 5 & \checkmark & - & - & - & \checkmark & - & 3 \\
-\multicolumn{1}{l|}{\textbf{Ours}} & sngl+batch & 9(14) & 10 & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & 50 
+\multicolumn{1}{l|}{\textbf{Ours}} & sngl+batch & 9(14) & 11 & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & 50 
 \end{tabular}
@@
 \subsection*{Contributions}
+%Maybe absorb this into the introduction text
 \begin{enumerate}
-	\item Evaluation of Active Learning algorithms on datasets from 5 different domains
-	\item Two novel synthetic datasets that highlight principled shortcomings of existing AL algorithms
-	\item Efficient and performant algorithm for an oracle that can be constructed greedily and does not rely on search
-        \item Evidence that previous works might have not repeated their experiments often enough to provide conclusive results
+	\item[C1] Study on the reproducibility of AL experiments, providing evidence that previous works might have not repeated their experiments often enough to provide reliable results
+	\item[C2] Efficient and performant algorithm for an oracle that can be constructed greedily and does not rely on search
+	\item[C3] Two novel synthetic datasets that highlight principled shortcomings of of the top-performing AL algorithms
+    \item[C4] Extending the evaluation of Active Learning algorithms to 5 data domains and revealing significant differences in algorithm performance between them
 \end{enumerate}
@@
 When setting $\alpha = 50$ we observe all samples to converge close to the true mean performance. 
-For that reason, we employ 50 restarts of every experiment in this work.
 \begin{figure}
@@
 \end{figure}
+In addition to this motivating example, we carried out our main evaluation (Tab. \ref{tab:results}) multiple times by uniformly sampling 3 random from our 50 available runs and comparing the results.
+We found significant differences in the performance of acquisition functions on individual datasets, as well as permutations in the final ranking.
+This partly explains the ongoing difficulties in reproducing results for AL experiments and benchmarks.
+This details can be found in App. \ref{app:rank_difference}.
+For this benchmark we always employ 50 restarts of the same experiment.
 
@@
 %\end{wrapfigure}
+
+%an oracle set was so far always described as the optimal set and therefore
 Posing Active Learning as a combinatorial problem, the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure is the set that induces the highest AUC score for a given budget.
@@
 The data point with the best test performance is selected and added to the labeled pool for that iteration.
-We noticed that this oracle is overfitting on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
-This can happen, for example, if the oracle picked a labeled set that enables the classifier to correctly classify a big portion of easy samples in the test set, but now fails to find the next single unlabeled point that would enable the classifier to succeed on one of the hard samples.
-This leads to a situation, where no point can incur an increase in test performance and therefore the selected data point can be considered random.
+We noticed that this oracle is over-specializing on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
+This can happen, for example, if the oracle picked a labeled set that enables the classifier to correctly classify a big portion of easy samples in the test set, but now fails to find the next \textbf{single} unlabeled point that would enable the classifier to succeed on one of the hard samples.
+This leads to a situation, where no point can immediately incur an increase in test performance and therefore the selected data point can be considered random.
 To circumvent this problem, we use margin sampling \cite{wang2014new} as a fallback option for the oracle.
@@
 Algorithms are sorted by aggregated performance on un-encoded datasets.
-For brevity, we only display per-dataset standard deviations for Splice.
-All standard deviations lie within $[0.0007, 0.024]$.}
+}
 	\label{tab:results}
@@
 \begin{tabular}{l|lllllll|ll}
-          & Splice & DNA  & USPS 		 	& Cfr10 & FMnist & TopV2 & News  & Un-encoded & Encoded \\
-          \hline
-Oracle 	& 1.0$\pm$0.011 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.9 \\
-Margin 	& 3.9$\pm$0.019 & 4.2 & \textbf{2.1} & 4.4 & \textbf{2.9} & \textbf{2.4} & 2.7 & 3.2 & 3.9 \\
-Badge  	& \textbf{3.0$\pm$0.014} & 6.2 & 2.9 & \textbf{3.4} & 3.2 & 3.3 & \textbf{2.5} & 3.5 & 4.9 \\
-DSA 	& 7.1$\pm$0.016 & 7.1 & 7.1 & 3.6 & 3.5 & 5.7 & 5.9 & 5.7 & 6.2 \\
-BALD 	& 3.9$\pm$0.011 & 4.6 & 5.3 & 10.1 & 5.8 & 7.2 & 3.8 & 5.8 & 6.9 \\
-CoreGCN & 6.6$\pm$0.012 & 4.8 & 9.4 & 5.7 & 4.8 & 4.0 & 5.4 & 5.8 & 7.7 \\
-Entropy & 6.4$\pm$0.018 & \textbf{3.8} & 7.2 & 5.7 & 3.4 & 9.1 & 8.2 & 6.3 & 6.0 \\
-LSA 	& 5.8$\pm$0.009 & 6.6 & 5.2 & 5.8 & 8.7 & 7.0 & 6.0 & 6.4 & 6.9 \\
-Random 	& 8.6$\pm$0.01 & 9.0 & 5.2 & 6.5 & 9.2 & 7.4 & 6.6 & 7.5 & 6.4 \\
-Coreset & 6.8$\pm$0.01 & 8.8 & 9.6 & 4.9 & 5.3 & 8.0 & 9.4 & 7.5 & 6.6 \\
-TypiClust & 8.3$\pm$0.009 & 9.8 & 11.0 & 6.0 & 9.1 & 11.0 & 10.6 & 9.4 & 8.6
+          & Splice & DNA  & USPS 		 	& Cfr10 & FMnist & TopV2 & News  & Un-enc. & Enc. \\
+          \hline
+Oracle  & 1.0 $\pm$ 0.01 & 1.0 $\pm$ 0.01 & 1.0 $\pm$ 0.0   & 1.0 $\pm$ 0.0   & 1.0 $\pm$ 0.0   & 1.0 $\pm$ 0.01 & 1.0 $\pm$ 0.0  & 1.0 & 2.0 \\
+Margin  & 6.6 $\pm$ 0.02 & 4.3 $\pm$ 0.01 & 2.1 $\pm$ 0.01  & 6.3 $\pm$ 0.01  & 4.4 $\pm$ 0.0   & 2.4 $\pm$ 0.01 & 3.7 $\pm$ 0.0  & 4.3 & 4.2 \\
+Badge   & 5.2 $\pm$ 0.01 & 6.3 $\pm$ 0.01 & 2.9 $\pm$ 0.01  & 5.2 $\pm$ 0.01  & 4.7 $\pm$ 0.0   & 3.3 $\pm$ 0.01 & 3.5 $\pm$ 0.0  & 4.5 & 5.4 \\
+LeastConf & 9.2 $\pm$ 0.02 & 10.3 $\pm$ 0.02 & 8.1 $\pm$ 0.02  & 2.1 $\pm$ 0.01 & 4.0 $\pm$ 0.0   & 7.9 $\pm$ 0.02  & 3.0 $\pm$ 0.01  & 6.4  & 6.5 \\
+DSA     & 7.4 $\pm$ 0.02 & 7.3 $\pm$ 0.01 & 7.5 $\pm$ 0.01  & 5.4 $\pm$ 0.01  & 5.1 $\pm$ 0.0   & 6.0 $\pm$ 0.02 & 7.3 $\pm$ 0.01 & 6.6 & 6.7 \\
+BALD    & 4.0 $\pm$ 0.01 & 4.7 $\pm$ 0.01 & 5.4 $\pm$ 0.01  & 12.0 $\pm$ 0.01 & 7.6 $\pm$ 0.0   & 7.6 $\pm$ 0.02 & 5.0 $\pm$ 0.0  & 6.6 & 7.6 \\
+CoreGCN & 6.9 $\pm$ 0.01 & 4.9 $\pm$ 0.01 & 10.4 $\pm$ 0.01 & 7.6 $\pm$ 0.01  & 6.5 $\pm$ 0.01  & 4.0 $\pm$ 0.01 & 6.8 $\pm$ 0.0  & 6.7 & 8.2 \\
+Entropy & 6.6 $\pm$ 0.02 & 3.9 $\pm$ 0.01 & 7.6 $\pm$ 0.01  & 7.6 $\pm$ 0.01  & 4.9 $\pm$ 0.01  & 9.8 $\pm$ 0.02 & 9.6 $\pm$ 0.0  & 7.1 & 6.5 \\
+LSA     & 6.1 $\pm$ 0.01 & 6.8 $\pm$ 0.01 & 5.3 $\pm$ 0.01  & 7.7 $\pm$ 0.01  & 10.6 $\pm$ 0.01 & 7.5 $\pm$ 0.01 & 7.3 $\pm$ 0.01 & 7.3 & 7.5 \\
+Random  & 9.0 $\pm$ 0.01 & 9.3 $\pm$ 0.01 & 5.3 $\pm$ 0.01  & 8.4 $\pm$ 0.01  & 11.1 $\pm$ 0.0  & 7.9 $\pm$ 0.01 & 8.0 $\pm$ 0.0  & 8.4 & 6.9 \\
+Coreset        & 7.1 $\pm$ 0.01 & 9.0 $\pm$ 0.01  & 10.5 $\pm$ 0.01 & 6.8 $\pm$ 0.01 & 7.1 $\pm$ 0.0   & 8.5 $\pm$ 0.02  & 10.8 $\pm$ 0.01 & 8.5  & 7.2 \\
+TypiClust      & 8.8 $\pm$ 0.01 & 10.2 $\pm$ 0.01 & 12.0 $\pm$ 0.02 & 7.9 $\pm$ 0.01 & 11.0 $\pm$ 0.01 & 12.0 $\pm$ 0.02 & 12.0 $\pm$ 0.01 & 10.5 & 9.2
 \end{tabular}
@@
 \end{table}
-%\begin{figure}
-%    \centering
-%    \caption{Results for all acquisition functions on un-encoded USPS (top) and the encoded FashionMnist (bottom) for two query sizes.}
-%    \label{fig:main_body_result}
-%    \includegraphics[width=0.9\linewidth]{img/main_body_result.jpg}
-%\end{figure}
 %
@@
     \centering
-    \caption{Ranks of each acquisition function aggregated by domain.}
+    \caption{Ranks of each acquisition function aggregated by domain.
+Horizontal bars indicate a \textbf{non}-significant rank difference.
+The significance is tested via a paired-t-test with $\alpha=0.05$.}
     \label{fig:ranks_by_domain}
@@
     \label{fig:main_body_result}
-    \includegraphics[width=0.9\linewidth]{img/synth_trajectories.jpg}
+    \includegraphics[width=0.49\linewidth]{img/micro_honeypot.jpg}
+    \includegraphics[width=0.49\linewidth]{img/micro_diverging_sin.jpg}
 \end{figure}
@@
 These ranks can safely be averages across datasets as they are no longer subjected to scaling differences of each dataset.
-In Table \ref{tab:results} we provide the rank of each acquisition function per dataset and averaged for each (un-)encoded dataset. \\
+In Table \ref{tab:results} we provide the rank of each acquisition function per dataset and averaged for each (un-)encoded dataset. 
+Please note, that for Tab \ref{tab:results} we are averaging not only over runs, but also over query sizes per dataset. 
+For the results per query size, please refer to App. \ref{app:ranks_by_query_size}. \\
 Additionally, we provide aggregated ranks per data domain in Fig. \ref{fig:ranks_by_domain}, revealing varying performances of acquisition functions like DSA/LSA, BALD and CoreGCN.
 The leading algorithms Margin-Sampling and BADGE show very stable performance across all tested data domains. \\ [1mm]
-Results for the Honeypot dataset reveal expected shortcomings of uncertainty sampling algorithms like margin and entropy sampling, but also show that BADGE is underperforming for this dataset. 
-This is an important result, as margin sampling and BADGE were evaluated to be the best performing acquisition functions for real-world data.
-Both being highly vulnerable to adverse samples or simply measurement noise highlights the need for further research in this area. \\
-Results for Diverging Sine also confirm expected behaviour, as clustering algorithms (Coreset, TypiClust) fall behind uncertainty algorithms (Entropy-, Margin-Sampling), with the exception of BADGE.
-The fact that BADGE is able to perform well on this dataset highlights the importance of embeddings in the clustering algorithms, as the so-called gradient embedding from BADGE seems to be able to encode uncertainty information, guiding the selection into the left hand regions of the dataset.
-We provide additional evidence for the importance of the embeddings by testing a version of Coreset on this dataset that does not use the embeddings produced by the classification model, but rather clusters the data directly.
-This "Coreset Raw" algorithm performs notably worse than its embedding-based counterpart. \\
-We strongly advocate to test newly proposed AL algorithms not only on our benchmark suite of real data, but also pay close attention to Honeypot and Diverging Sine to reveal principled shortcomings of the algorithm.
+Results for the Honeypot dataset reveal expected shortcomings of uncertainty sampling algorithms like margin, entropy and least confident sampling as well as BALD.
+In addition, BADGE is underperforming for this dataset compared to real-world data. 
+Results for Diverging Sine also confirm expected behavior, as clustering algorithms (Coreset, TypiClust) fall behind uncertainty algorithms (Entropy-, Margin-Sampling), with the exception of BADGE. \\
+We provide a very small ablation study on the importance of the embeddings by testing a version of Coreset and TypiClust on this dataset that does not use the embeddings produced by the classification model, but rather clusters the data directly.
+"Coreset Raw" and "TypiClust Raw" both perform worse than their embedding-based counterpart.
+
+\section{Conclusion}
+As stated in contribution C4, our results on real-world data shows significant differences in the performance of the tested algorithms between data domains.
+Not only do some algorithms overperform on some domains (like least confidence sampling on Images), but the Top-3 of algorithms (except Oracle) does not contain the same three algorithms for any two domains.
+This highlights the dire need for diverse data domains in AL benchmarking.
+Furthermore, our results on Honeypot reveal principled shortcomings for the two best algorithms in BADGE and margin sampling.
+Both are vulnerable to adverse samples or simply measurement noise, which highlights the need for further research in this area. \\
+Finally, the fact that BADGE is able to perform well on Diverging Sine highlights the importance of embeddings for the clustering algorithms, as the so-called gradient embedding from BADGE seems to be able to encode uncertainty information, guiding the selection into the left hand regions of the dataset.
+We also show that embeddings are generally useful for this dataset, by providing results for "Coreset Raw" and "TypiClust Raw". \\ [1mm]
+In conclusion, we strongly advocate to test newly proposed AL algorithms not only on a wide variety of real data domains, but also to pay close attention to the Honeypot and Diverging Sine datasets to reveal principled shortcomings of the algorithm in question.
+
+%\section{Future Work}
+% Wahrscheinlich werden wir keinen Platz dafür haben
+% Wie sehr hängen die Unterschiede am Classifier und nicht an den Daten? -> Mehrere Classifier pro Domäne
+% Mehr Datensätze pro Domäne
+% Neuere Algorithmen. (Neuster von 2022)
 
@@
 
-\section{Problem Formulation}\label{app:problem_formulation}
-Given
-\begin{itemize}
-	\item a number $B\in\N$ (called budget),
-	\item two spaces $\X$ and $\Y$,  {\tiny e.g., $\X:=\R^M, \Y:=\R^T$},
-	\item a sample $\D_1,\ldots,\D_N \in (\X\times \Y)^*$ of
-	sequences of pairs $(x,y)$  from an unknown distribution $p$
-	(called datasets),
-	%  {\tiny with $|\D_n|\geq B$ for all $n\in 1{:}N$,}
-	{\tiny with $p(\D)=0$ for $|\D|<B$,}
-	\item a function $\ell:\Y\times\Y\rightarrow\R$ (called loss), and
-	\item a function $\hat y:  (\X\times \Y)^* \times \X^* \rightarrow \Y^\X$
-	(called learning algorithm), \\
-	{\tiny where $\Y^\X$ is the space of all function from $\X$ to $\Y$}
-\end{itemize}
-find a function
-\vspace*{-0.5cm}
-\begin{align*}
-	\Omega: (\X\times \Y)^* \times \X^* &\rightarrow \N
-	\quad\quad \text{\tiny (with $a(\D,X) \leq |X|$)} \\
-	& \text{\tiny where $a(\D,X)$ selects an unlabeled instance from $X$}
-\end{align*}
+%\section{Problem Formulation}\label{app:problem_formulation}
+%Given
+%\begin{itemize}
+%	\item a number $B\in\N$ (called budget),
+%	\item two spaces $\X$ and $\Y$,  {\tiny e.g., $\X:=\R^M, \Y:=\R^T$},
+%	\item a sample $\D_1,\ldots,\D_N \in (\X\times \Y)^*$ of
+%	sequences of pairs $(x,y)$  from an unknown distribution $p$
+%	(called datasets),
+%	%  {\tiny with $|\D_n|\geq B$ for all $n\in 1{:}N$,}
+%	{\tiny with $p(\D)=0$ for $|\D|<B$,}
+%	\item a function $\ell:\Y\times\Y\rightarrow\R$ (called loss), and
+%	\item a function $\hat y:  (\X\times \Y)^* \times \X^* \rightarrow \Y^\X$
+%	(called learning algorithm), \\
+%	{\tiny where $\Y^\X$ is the space of all function from $\X$ to $\Y$}
+%\end{itemize}
+%find a function
+%\vspace*{-0.5cm}
+%\begin{align*}
+%	\Omega: (\X\times \Y)^* \times \X^* &\rightarrow \N
+%	\quad\quad \text{\tiny (with $a(\D,X) \leq |X|$)} \\
+%	& \text{\tiny where $a(\D,X)$ selects an unlabeled instance from $X$}
+%\end{align*}
 % which is equivariant in the second argument,
 
-called acquisition function,
-s.t. the expected loss of a model learned on all predictors plus $B$ sequentially acquired targets
-is minimal:
-\begin{align*}
-	\min\ \EE\   &  \{
-	% \operatorname{avg}\limits_{(x,y)\in\D\test} \ell(y, \hat y(x))
-	\ell(\hat y, \D\test)
-	\mid \D\sim p, (\D\train,\D\test):= \text{split}(\D) \}
-	\\
-	\text{with }
-	\hat y:= & A( (\D_{\train_{n_1}},\ldots,\D_{\train_{n_B}}), \D\train|_{\X})
-	\\ 
-	n_b := & a( (\D_{\train_{n_1}},\ldots,\D_{\train_{n_{b-1}}}), \D\train|_{\X}) ,
-	\quad b\in 1{:}B
-\end{align*}
+%called acquisition function,
+%s.t. the expected loss of a model learned on all predictors plus $B$ sequentially acquired targets
+%is minimal:
+%\begin{align*}
+%	\min\ \EE\   &  \{
+%	% \operatorname{avg}\limits_{(x,y)\in\D\test} \ell(y, \hat y(x))
+%	\ell(\hat y, \D\test)
+%	\mid \D\sim p, (\D\train,\D\test):= \text{split}(\D) \}
+%	\\
+%	\text{with }
+%	\hat y:= & A( (\D_{\train_{n_1}},\ldots,\D_{\train_{n_B}}), \D\train|_{\X})
+%	\\ 
+%	n_b := & a( (\D_{\train_{n_1}},\ldots,\D_{\train_{n_{b-1}}}), \D\train|_{\X}) ,
+%	\quad b\in 1{:}B
+%\end{align*}
 
@@
 \textbf{Reinforcement Learning Algorithms} \\
-TODO \\
+%TODO \\
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{All Results}\label{app:all_results}
+\section{Difference of Ranks with 3 Repetitions}\label{app:rank_difference}
+Table \ref{tab:rank_diff_1} and Table \ref{tab:rank_diff_2} follow the exact same computation of ranks that created the main result (Table \ref{tab:results}) with the only difference being a reduced number of runs per acquisition function.
+For each table we uniformly sampled 3 runs from the available 50 per acquisition function. \\
+We can observe significant differences between the two tables: \\
+{\color{purple}Purple}: A multitude of rank differences of acquisition functions for specific datasets, some as high as 4.7 ranks for TypiClust on the Splice dataset \\
+{\color{olive}Olive}: Well separated acquisition functions in Tab. \ref{tab:rank_diff_2} (Margin and BADGE) are almost indistinguishable in Tab \ref{tab:rank_diff_1} \\
+{\color{red}Red}: BALD lost 2 places in the overall ranking and Entropy gained 2 \\ [1mm]
+Even though the overall ordering of acquisition functions stayed relatively unchanged due to the averaging across many datasets, each individual dataset was subject to drastic permutations.
+This highlights the need for many repetitions in AL experiments.
+\begin{table}[H]
+	\centering
+	\caption{Ranks of all acquisition functions per dataset. First random draw of 3 runs from the overall pool of 50.}
+	\label{tab:rank_diff_1}
+	\resizebox{\columnwidth}{!}{
+	\begin{tabular}{l|lllllll|ll}
+		               & Splice & DNA  & USPS & Cifar10 & FMnist & TopV2 & News & Unencoded & Encoded \\
+		               \hline
+		Oracle         & 1.0    & 1.0  & 1.0  & 1.0     & 1.0          & 1.0   & 1.0  & 1.0       & 2.1     \\
+		Margin         & 6.0    & 7.3  & 2.0  & 6.7     & 5.3          & 2.3   & 3.3  & {\color{olive}4.7}& 4.4     \\
+		Badge          & 6.0    & 7.3  & 3.0  & 6.7     & 5.0          & 3.3   & 4.0  & {\color{olive}5.0}& 5.3     \\
+		{\color{red}BALD}& 3.3    & 4.7  & 5.3  & 12.0    & 7.0          & 6.3   & 4.3  & 6.1       & 7.9     \\
+		CoreGCN        & 8.7    & 3.7  & 10.7 & 6.3     & 5.3          & 4.0   & 7.7  & 6.6       & 9.1     \\
+		DSA            & 8.3    & 6.3  & 7.7  & 7.7     & 4.3          & 6.7   & 6.7  & 6.8       & 6.1     \\
+		LeastConf      & 10.0   & 12.0 & 8.0  & 3.0     & 4.3          & 9.3   & 2.3  & 7.0       & 6.7     \\
+		LSA            & 5.7    & 6.7  & 5.3  & 6.7     & 10.7         & 7.7   & 7.0  & 7.1       & 6.3     \\
+		Entropy        & 11.0   & 3.3  & 7.3  & 4.0     & 6.7          & 8.3   & 9.7  & 7.2       & 7.0     \\
+		Random         & 7.7    & 8.7  & 5.3  & 8.0     & 11.0         & 8.0   & 9.0  & 8.2       & 6.3     \\
+		Coreset        & 4.7    & 10.3 & 10.3 & 7.7     & 6.0          & 9.0   & 11.0 & 8.4       & 7.2     \\
+		TypiClust      & {\color{purple}5.7}& 6.7  & 12.0 & 8.3     & 11.3         & 12.0  & 12.0 & 9.7       & 9.7    
+	\end{tabular}
+	}
+\end{table}
+%
+\begin{table}[H]
+	\centering
+	\caption{Ranks of all acquisition functions per dataset. Second random draw of 3 runs from the overall pool of 50.}
+	\label{tab:rank_diff_2}
+	\resizebox{\columnwidth}{!}{
+	\begin{tabular}{l|lllllll|ll}
+		               & Splice & DNA  & USPS & Cifar10 & FMnist & TopV2 & News & Unencoded & Encoded \\
+		\hline
+		Oracle         & 1.0    & 1.0  & 1.0  & 1.0     & 1.0          & 1.0   & 1.0  & 1.0       & 2.4     \\
+		Margin         & 6.0    & 3.3  & 2.0  & 5.7     & 2.0          & 2.0   & 4.3  & {\color{olive}3.6}& 3.8     \\
+		Badge          & 6.0    & 9.0  & 3.0  & 3.0     & 5.7          & 3.7   & 3.3  & {\color{olive}4.8}& 4.9     \\
+		CoreGCN        & 4.3    & 6.3  & 10.3 & 7.3     & 5.3          & 5.7   & 5.3  & 6.4       & 8.1     \\
+		DSA            & 8.7    & 7.3  & 7.3  & 6.0     & 4.3          & 5.3   & 6.0  & 6.4       & 6.5     \\
+		{\color{red}BALD}& 4.7    & 4.0  & 4.7  & 12.0    & 7.3          & 6.7   & 6.7  & 6.6       & 7.5     \\
+		Entropy        & 6.7    & 4.7  & 7.7  & 5.3     & 5.0          & 7.3   & 9.3  & 6.6       & 6.8     \\
+		LeastConf      & 7.7    & 10.0 & 8.3  & 3.3     & 6.0          & 8.7   & 3.0  & 6.7       & 7.3     \\
+		LSA            & 7.7    & 5.3  & 6.0  & 9.0     & 11.0         & 9.0   & 7.3  & 7.9       & 7.5     \\
+		Random         & 9.3    & 8.0  & 5.0  & 8.7     & 11.7         & 8.3   & 8.7  & 8.5       & 7.6     \\
+		Coreset        & 6.0    & 10.7 & 10.7 & 8.0     & 8.3          & 8.3   & 11.0 & 9.0       & 6.3     \\
+		TypiClust      & {\color{purple}10.0}& 8.3  & 12.0 & 8.7     & 10.3         & 12.0  & 12.0 & 10.5      & 9.4    
+	\end{tabular}
+	}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Ranks by Query Size}\label{app:ranks_by_query_size}
+%TODO
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Individual Results}\label{app:all_results}
 % Splice

### 2024-06-02 agents/__init__.py

@@ -5,7 +5,7 @@ from agents.shannon_entropy import ShannonEntropy
 from agents.least_confident import LeastConfident
 from agents.bald import BALD
 from agents.badge import Badge
-from agents.typiclust import TypiClust
+from agents.typiclust import TypiClust, TypiClust_Raw
 from agents.core_gcn import CoreGCN
 from agents.surprise_adequacy import DSA, LSA
 

### 2024-06-02 core/evaluation.py

@@ -62,6 +62,12 @@ def _find_missing_runs():
                     else:
                         print(f"Accuracy file missing {acc_file}")
 
+def sort_according_to_reference(unsorted_list:list, reference:list):
+    result = []
+    for entry in reference:
+        if entry in unsorted_list:
+            result.append(entry)
+    return result
 
 def _t_value_for_samplesize(n_samples, sig_level= 0.95):
     return t.ppf(sig_level, n_samples)
@@ -156,18 +162,19 @@ def _query_to_list(query, current_folder):
         raise ValueError(f"Query not recognized: {query}")
     return result_list
 
-def generate_full_overview(precision=1, add_std=True):
+def generate_rank_leaderboard(precision=1, add_std=True, subsample_runs=None):
     datasets_raw = ["Splice", "DNA", "USPS", "Cifar10", "FashionMnist", "TopV2", "News",]
                     #"DivergingSin", "ThreeClust"]
     datasets_encoded = ["SpliceEncoded", "DNAEncoded", "USPSEncoded",
                         "Cifar10Encoded", "FashionMnistEncoded"]
-    df_raw = combine_agents_into_df(dataset=datasets_raw, include_oracle=True)
-    df_raw = _insert_oracle_forecast(df_raw)
-    df_raw = average_out_columns(df_raw, ["iteration", "query_size"])
+    df_raw = combine_agents_into_df(dataset=datasets_raw, include_oracle=True, subsample_runs=subsample_runs)
+    # df_raw = _insert_oracle_forecast(df_raw)
+    df_raw = average_out_columns(df_raw, ["iteration"])
+    df_raw = average_out_columns(df_raw, ["query_size"])
     df_raw = compute_ranks_over_trials(df_raw)
     print("range of stds", df_raw["rank_std"].min(), "-", df_raw["rank_std"].max())
 
-    df_enc = combine_agents_into_df(dataset=datasets_encoded, include_oracle=True)
+    df_enc = combine_agents_into_df(dataset=datasets_encoded, include_oracle=True, subsample_runs=subsample_runs)
     df_enc = average_out_columns(df_enc, ["iteration", "query_size"])
     df_enc = compute_ranks_over_trials(df_enc)
 
@@ -188,7 +195,7 @@ def generate_full_overview(precision=1, add_std=True):
                 print(f"No runs found for {index} on {dataset}")
                 continue
             if add_std:
-                values.append(f"{round(r.item(), precision)}+-{round(r_std.item(), 3)}")
[…]

### 2024-06-02 core/helper_functions.py

@@ -112,7 +112,7 @@ def plot_upper_bound(ax, dataset, x_values, color, alpha=0.8, percentile=0.99, l
 
 _all_query_sizes = [1, 5, 20, 50, 100, 500]
 _all_agents = ["Badge", "BALD", "CoreGCN", "Coreset_Greedy", "DSA", "LSA", "MarginScore", "RandomAgent",
-               "ShannonEntropy", "TypiClust", "Coreset_Raw", "Oracle"]
+               "ShannonEntropy", "LeastConfident", "TypiClust", "TypiClust_Raw", "Coreset_Raw", "Oracle"]
 _agent_colors = {
     "Oracle": "red",
     "UpperBound": "black",
@@ -123,9 +123,11 @@ _agent_colors = {
     "DSA": "olive",
     "LSA": "brown",
     "MarginScore": "blue",
+    "LeastConfident": "cyan",
     "RandomAgent": "grey",
     "ShannonEntropy": "green",
     "TypiClust": "pink",
+    "TypiClust_Raw": "hotpink",
     "Coreset_Raw": "magenta",
 }
 _agent_names = { # only corrected names
@@ -216,7 +218,7 @@ def _get_oracle_regression(x, y, x_test, upper_bound):
     multipliers = np.exp(-space / 0.5) # exponential
     for i in range(0, len(y_test)):
         multiplicator = multipliers[i]
-        y_test[i] = y_test[i] * multiplicator + upper_bound * 0.99 * (1-multiplicator)
+        y_test[i] = min(y_test[i], y_test[i] * multiplicator + upper_bound * 0.99 * (1-multiplicator))
     idx = [i-max(x) for i in x_test]
     return x_test, y_test[idx]
 
@@ -353,17 +355,20 @@ def collect_results(base_path, folder_prefix):
     result_acc = pd.DataFrame()
     result_loss = pd.DataFrame()
     runs = sorted(os.listdir(base_path), key=functools.cmp_to_key(sort_by_run_id))
+    column_names = []
     for run_folder in runs:
         if run_folder.startswith(folder_prefix):
             acc_file_path = join(base_path, run_folder, "accuracies.csv")
-            if exists(acc_file_path):
-                accuracies = pd.read_csv(acc_file_path, header=0, index_col=0)
-                result_acc = pd.concat([result_acc, accuracies], axis=1, ignore_index=True)
-
             loss_file_path = join(base_path, run_folder, "losses.csv")
+            assert exists(acc_file_path)
[…]

### 2024-06-02 results/cd_diagram.py

@@ -385,24 +385,28 @@ def wilcoxon_holm(alpha=0.05, df_perf=None):
 
 def prepare_df(df):
     df = df[df["trial"] <= 50]
-    df = average_out_columns(df, ["iteration", "trial"])
+    df = average_out_columns(df, ["iteration"])
+    # df = average_out_columns(df, ["iteration", "trial"])
     oracle_data = df[df["agent"]=="Oracle"]
     df_res = df[df["agent"]!="Oracle"]
     # non-oracle results
     df_res = pd.DataFrame({
         "classifier_name": df_res["agent"],
-        "dataset_name": df_res["dataset"].astype(str) + df_res["query_size"].astype(str),
+        "dataset_name": df_res["dataset"].astype(str) + df_res["query_size"].astype(str) + df_res["trial"].astype(str),
         "accuracy": df_res["auc"]
     })
     # copy oracle results for every present query size of a dataset
     for d in pd.unique(df["dataset"]):
         sub_df = df[(df["dataset"] == d).astype(bool) & (df["agent"] != "Oracle").astype(bool)]
         query_sizes = pd.unique(sub_df["query_size"].astype(int))
+        sample_agent_name = pd.unique(df_res["classifier_name"])[0]
+        num_trials = len(pd.unique(df[df["agent"]==sample_agent_name]["trial"]))
         oracle_acc = oracle_data[oracle_data["dataset"]==d]["auc"].values[0]
         for q in query_sizes:
-            df_res = df_res._append(pd.DataFrame({"classifier_name":["Oracle"],
-                                          "dataset_name":str(d)+str(q),
-                                          "accuracy":oracle_acc}))
+            for t in range(num_trials):
+                df_res = df_res._append(pd.DataFrame({"classifier_name":["Oracle"],
+                                              "dataset_name":str(d)+str(q)+str(t),
+                                              "accuracy":oracle_acc}))
 
     # replace names for agents
     for wrong, right in name_corrections.items():
@@ -427,35 +431,26 @@ if __name__ == '__main__':
 
     # df = combine_agents_into_df(["Splice", "DNA", "USPS"], include_oracle=True)
     # df = prepare_df(df)
-    # draw_cd_diagram(df, title="Tabular", alpha=10000.0, file="doc/img/macro_vector.jpg")
[…]

### 2024-06-03 doc/al_benchmark_paper.tex
paper und ergebnisse

@@
 \usepackage{makecell}
+\usepackage{arydshln}
 \usepackage{neurips_data_2024}
@@
 % - Ergebnisse schwierig zu reproduzieren
-% - Vorherige Benchmarks haben dem Versuchsaufbau zu wenig Beachtung geschenkt - erst Ji und Luth haben verlässliche Benchmarks beschrieben
+% - Vorherige Benchmarks haben dem Versuchsaufbau zu wenig Beachtung geschenkt - erst Ji und Luth haben Methoden für verlässliche AL Experimente beschrieben
 % - keiner (!) der Benchmarks hat beide unserer Top-2 Methoden (Margin und Badge)
@@
 % - Unser Oracle ist viel leichter zu berechnen
+% - Wir erhöhen die Anzahl der Experimente exponentiell: Ji und Lueth: 300-400; wir: 24200+Oracle; Größtes Benchmark bis jetzt ist Zhan (2021): ~31500 (nur tabular)
 % Why is that important?
@@
 % !! Sec 4 Methodology ist sehr chaotisch. Wir müssen neu sortieren und unwichtige Dinge streichen
+% !! Sec 3 Related Work sollte Ji et al. im Detail einführen, weil wir uns an deren Guidelines halten. Alles andere sollte dort gekürzt werden
 % Contributions neu sortieren und ausführlicher beschreiben
@@
 % Advantage-Matrix von Ji et. al. neu aufarbeiten. Wo sind die Unterschiede? Ist das CD-Diagramm besser/schlechter?
+% (Un-)Encoded durch Normal uns Semi-Supervised ersetzen, um die Verbindung klar zu machen
 % Gibt es eine Conclusion für die Encoded-Datasets? Im Moment werden die nicht besprochen
@@
 % Conclusion: Es gibt Unterschiede bei der Top 1 zwischen den Domänen, aber die Top 3 ist konstant
+
+%%%%%%%%%%%%%%%%
+% Anzahl der Runs (Ohne Oracle und Upper Bound)
+% Runs pro query size: 11alg * 50runs = 550
+% Cifar10   x2 = 1100
+% Cfr10Enc  x4 = 2200
+% DivSin    x2 = 1100
+% DNA       x4 = 2200
+% DNAEnc    x2 = 1100
+% FMnist    x2 = 1100
+% FMnistEnc x4 = 2200
+% News      x3 = 1650
+% Splice    x4 = 2200
+% SpliceEnc x3 = 1650
+% Honeypot  x2 = 1100
+% TopV2     x4 = 2200
+% USPS      x4 = 2200
+% USPSEnc   x4 = 2200
+% = 24200 runs
+% Lueth: ~315
+% Ji: ~432
+% Zhan: ~31500 (nur tabular data)
 
@@
 \begin{tabular}{lccccllllcl}
-\hline
 Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
@@
 \multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\
+
+\multicolumn{1}{l|}{Rauch et al. \cite{rauch2023activeglae}} & batch & 11 & 5 & - & \checkmark & - & - & - & - & 5 \\
+\hdashline
 \multicolumn{1}{l|}{Ji et al. \cite{ji2023randomness}} & batch & 3 & 8 & \checkmark & - & - & - & - & - & - \\
-\multicolumn{1}{l|}{Luth et al. \cite{luth2024navigating}} & batch & 4 & 5 & \checkmark & - & - & - & \checkmark & - & 3 \\
+\multicolumn{1}{l|}{Lueth et al. \cite{luth2024navigating}} & batch & 4 & 5 & \checkmark & - & - & - & \checkmark & - & 3 \\
 \multicolumn{1}{l|}{\textbf{Ours}} & sngl+batch & 9(14) & 11 & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & 50 
@@
 Please note, that for Tab \ref{tab:results} we are averaging not only over runs, but also over query sizes per dataset. 
-For the results per query size, please refer to App. \ref{app:ranks_by_query_size}. \\
+For the results per query size, please refer to App. \ref{app:auc_by_query_size}. \\
 Additionally, we provide aggregated ranks per data domain in Fig. \ref{fig:ranks_by_domain}, revealing varying performances of acquisition functions like DSA/LSA, BALD and CoreGCN.
@@
 		LSA            & 5.7    & 6.7  & 5.3  & 6.7     & 10.7         & 7.7   & 7.0  & 7.1       & 6.3     \\
-		Entropy        & 11.0   & 3.3  & 7.3  & 4.0     & 6.7          & 8.3   & 9.7  & 7.2       & 7.0     \\
+		{\color{red}Entropy}        & 11.0   & 3.3  & 7.3  & 4.0     & 6.7          & 8.3   & 9.7  & 7.2       & 7.0     \\
 		Random         & 7.7    & 8.7  & 5.3  & 8.0     & 11.0         & 8.0   & 9.0  & 8.2       & 6.3     \\
@@
 		{\color{red}BALD}& 4.7    & 4.0  & 4.7  & 12.0    & 7.3          & 6.7   & 6.7  & 6.6       & 7.5     \\
-		Entropy        & 6.7    & 4.7  & 7.7  & 5.3     & 5.0          & 7.3   & 9.3  & 6.6       & 6.8     \\
+		{\color{red}Entropy}        & 6.7    & 4.7  & 7.7  & 5.3     & 5.0          & 7.3   & 9.3  & 6.6       & 6.8     \\
 		LeastConf      & 7.7    & 10.0 & 8.3  & 3.3     & 6.0          & 8.7   & 3.0  & 6.7       & 7.3     \\
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Ranks by Query Size}\label{app:ranks_by_query_size}
-%TODO
+\section{AUCs by Query Size}\label{app:auc_by_query_size}
+% QS 1
+\begin{table}[H]
+	\caption{AUC values for each dataset that supports query size 1.}
+	\resizebox{\columnwidth}{!}{
+	\begin{tabular}{l|lllllllllll}
+		       & Splice       & SpliceEncoded& DNA          & DNAEncoded  & USPS          & USPSEncoded  & Cifar10Encoded& FashionMnistEnc& TopV2     & DivergingSin & ThreeClust \\
+		\hline
+		Oracle & 0.803+-0.012 & 0.678+-0.021 & 0.825+-0.009 & 0.721+-0.013 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.884+-0.006 & 0.957+-0.009 & 0.783+-0.03 \\
+		Margin & 0.769+-0.021 & 0.678+-0.032 & 0.806+-0.013 & 0.642+-0.047 & 0.858+-0.006 & 0.426+-0.038 &  0.653+-0.013 & 0.68+-0.012 & 0.861+-0.009 & 0.941+-0.018 & 0.704+-0.074 \\
+		Badge & 0.767+-0.02 & 0.661+-0.026 & 0.78+-0.014 & 0.642+-0.046 & 0.83+-0.008 & 0.371+-0.035 &  0.656+-0.013 & 0.68+-0.009 & 0.826+-0.024 & 0.941+-0.017 & 0.69+-0.083 \\
+		LeastConfident & 0.779+-0.019 & 0.68+-0.032 & 0.809+-0.01 & 0.629+-0.05 & 0.846+-0.009 & 0.421+-0.039 &  0.668+-0.014 & 0.685+-0.009 & 0.843+-0.013 & 0.94+-0.016 & 0.692+-0.094 \\
+		DSA & 0.766+-0.021 & 0.691+-0.022 & 0.803+-0.01 & 0.646+-0.032 & 0.829+-0.01 & 0.431+-0.05 & 0.663+-0.014 & 0.679+-0.01 & 0.844+-0.017 & 0.941+-0.014 & 0.731+-0.032 \\
+		BALD & 0.78+-0.014 & 0.649+-0.04 & 0.784+-0.01 & 0.632+-0.042 & 0.819+-0.01 & 0.242+-0.046 & 0.666+-0.014 & 0.644+-0.018 & 0.815+-0.024 & 0.928+-0.014 & 0.698+-0.043 \\
+		CoreGCN & 0.765+-0.021 & 0.686+-0.023 & 0.804+-0.012 & 0.646+-0.03 & 0.753+-0.016 & 0.39+-0.044 & 0.623+-0.018 & 0.647+-0.012 & 0.85+-0.01 & 0.938+-0.014 & 0.731+-0.028 \\
+		Entropy & 0.768+-0.022 & 0.678+-0.035 & 0.812+-0.013 & 0.635+-0.045 & 0.83+-0.011 & 0.399+-0.035 & 0.663+-0.013 & 0.681+-0.011 & 0.815+-0.021 & 0.942+-0.017 & 0.696+-0.083 \\
+		LSA & 0.772+-0.016 & 0.68+-0.026 & 0.787+-0.012 & 0.618+-0.036 & 0.821+-0.009 & 0.422+-0.037 & 0.613+-0.014 & 0.642+-0.012 & 0.816+-0.013 & 0.932+-0.016 & 0.727+-0.033 \\
+		Random & 0.76+-0.016 & 0.674+-0.027 & 0.774+-0.013 & 0.63+-0.035 & 0.823+-0.009 & 0.404+-0.036 & 0.613+-0.014 & 0.639+-0.013 & 0.815+-0.012 & 0.933+-0.017 & 0.721+-0.036 \\
+		Coreset & 0.772+-0.016 & 0.69+-0.017 & 0.79+-0.012 & 0.638+-0.041 & 0.767+-0.016 & 0.404+-0.046 & 0.659+-0.011 & 0.684+-0.009 & 0.826+-0.022 & 0.937+-0.014 & 0.73+-0.031 \\
+		TypiClust & 0.762+-0.016 & 0.685+-0.025 & 0.778+-0.01 & 0.663+-0.028 & 0.828+-0.007 & 0.396+-0.046 & 0.653+-0.013 & 0.649+-0.007 & 0.831+-0.011 & 0.934+-0.018 & 0.727+-0.033
+	\end{tabular}
+}
+\end{table}
+% QS 5
+\begin{table}[H]
+	\caption{AUC values for each dataset that supports query size 5.}
+	\resizebox{\columnwidth}{!}{
+	\begin{tabular}{l|lllllllllll}
+		& Splice & SpliceEncoded & DNA & DNAEncoded & USPS & USPSEncoded & Cifar10Encoded & FashionMnistEncoded & TopV2 & DivergingSin & ThreeClust \\
+		\hline
+		Oracle & 0.803+-0.012 & 0.678+-0.021 & 0.825+-0.009 & 0.721+-0.013 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.884+-0.006 & 0.957+-0.009 & 0.783+-0.03\\
+		Margin & 0.765+-0.021 & 0.662+-0.032 & 0.794+-0.011 & 0.611+-0.05 & 0.855+-0.006 & 0.508+-0.02 & 0.656+-0.014 & 0.678+-0.009 & 0.848+-0.013 & 0.923+-0.019 & 0.697+-0.055 \\
+		Badge & 0.768+-0.014 & 0.646+-0.035 & 0.785+-0.011 & 0.624+-0.036 & 0.846+-0.007 & 0.48+-0.021 & 0.647+-0.012 & 0.67+-0.009 & 0.847+-0.01 & 0.924+-0.019 & 0.72+-0.036 \\
+		LeastConfident & 0.763+-0.023 & 0.643+-0.034 & 0.798+-0.013 & 0.585+-0.065 & 0.831+-0.014 & 0.478+-0.028 & 0.67+-0.01 & 0.681+-0.009 & 0.819+-0.023 & 0.921+-0.019 & 0.675+-0.072 \\
+		DSA & 0.765+-0.023 & 0.653+-0.029 & 0.793+-0.009 & 0.613+-0.034 & 0.822+-0.01 & 0.489+-0.024 & 0.661+-0.013 & 0.662+-0.012 & 0.833+-0.02 & 0.924+-0.018 & 0.718+-0.033 \\
+		BALD & 0.775+-0.018 & 0.641+-0.034 & 0.801+-0.013 & 0.592+-0.054 & 0.84+-0.008 & 0.332+-0.054 & 0.681+-0.011 & 0.681+-0.013 & 0.824+-0.023 & 0.893+-0.035 & 0.673+-0.041 \\
+		CoreGCN & 0.759+-0.018 & 0.662+-0.027 & 0.79+-0.011 & 0.62+-0.03 & 0.755+-0.011 & 0.45+-0.03 & 0.604+-0.016 & 0.609+-0.013 & 0.837+-0.014 & 0.922+-0.018 & 0.723+-0.034 \\
+		Entropy & 0.765+-0.022 & 0.66+-0.03 & 0.798+-0.011 & 0.611+-0.054 & 0.823+-0.013 & 0.464+-0.024 & 0.663+-0.013 & 0.672+-0.011 & 0.801+-0.025 & 0.924+-0.02 & 0.689+-0.066 \\
+		LSA & 0.769+-0.016 & 0.654+-0.032 & 0.781+-0.013 & 0.61+-0.041 & 0.82+-0.009 & 0.484+-0.022 & 0.617+-0.012 & 0.641+-0.011 & 0.816+-0.012 & 0.915+-0.018 & 0.718+-0.038 \\
+		Random & 0.758+-0.015 & 0.655+-0.026 & 0.771+-0.013 & 0.623+-0.031 & 0.82+-0.009 & 0.476+-0.024 & 0.616+-0.016 & 0.637+-0.012 & 0.812+-0.014 & 0.921+-0.018 & 0.713+-0.034 \\
+		Coreset & 0.765+-0.017 & 0.663+-0.023 & 0.784+-0.014 & 0.603+-0.034 & 0.765+-0.015 & 0.449+-0.022 & 0.657+-0.009 & 0.674+-0.009 & 0.817+-0.017 & 0.92+-0.017 & 0.713+-0.035 \\
+		TypiClust & 0.759+-0.014 & 0.641+-0.028 & 0.775+-0.01 & 0.603+-0.04 & 0.757+-0.02 & 0.465+-0.027 & 0.596+-0.014 & 0.567+-0.012 & 0.727+-0.026 & 0.916+-0.02 & 0.693+-0.045
+	\end{tabular}
+}
+\end{table}
+% QS 20
+\begin{table}[H]
+	\caption{AUC values for each dataset that supports query size 20.}
+	\resizebox{\columnwidth}{!}{
+	\begin{tabular}{l|lllllllll}
+		       & Splice       & SpliceEncoded& DNA          & USPS         & USPSEncoded  & Cifar10Encoded& FashionMnistEnc & TopV2 & News \\
+		\hline
+		Oracle & 0.803+-0.012 & 0.678+-0.021 & 0.825+-0.009 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.884+-0.006 & 0.49+-0.003 \\
+		Margin & 0.759+-0.027 & 0.618+-0.04 & 0.779+-0.013 & 0.847+-0.008 & 0.439+-0.027 & 0.656+-0.01 & 0.67+-0.011 & 0.823+-0.014 & 0.464+-0.007 \\
+		Badge & 0.767+-0.013 & 0.619+-0.033 & 0.776+-0.013 & 0.845+-0.006 & 0.44+-0.019 & 0.647+-0.013 & 0.665+-0.007 & 0.827+-0.016 & 0.463+-0.007 \\
+		LeastConfident & 0.751+-0.02 & 0.597+-0.05 & 0.748+-0.025 & 0.798+-0.027 & 0.391+-0.024 & 0.665+-0.013 & 0.669+-0.011 & 0.775+-0.035 & 0.467+-0.008 \\
+		DSA & 0.759+-0.02 & 0.599+-0.034 & 0.769+-0.013 & 0.809+-0.012 & 0.421+-0.023 & 0.647+-0.014 & 0.63+-0.013 & 0.793+-0.026 & 0.459+-0.01 \\
+		BALD & 0.768+-0.022 & 0.57+-0.037 & 0.784+-0.015 & 0.822+-0.009 & 0.298+-0.039 & 0.675+-0.008 & 0.673+-0.01 & 0.789+-0.024 & 0.468+-0.009 \\
+		CoreGCN & 0.759+-0.018 & 0.612+-0.039 & 0.774+-0.012 & 0.754+-0.016 & 0.397+-0.026 & 0.587+-0.015 & 0.583+-0.015 & 0.807+-0.018 & 0.453+-0.006 \\
+		Entropy & 0.759+-0.027 & 0.618+-0.038 & 0.773+-0.015 & 0.803+-0.019 & 0.372+-0.022 & 0.656+-0.011 & 0.65+-0.012 & 0.773+-0.031 & 0.451+-0.007 \\
+		LSA & 0.761+-0.014 & 0.611+-0.039 & 0.768+-0.015 & 0.816+-0.009 & 0.411+-0.022 & 0.621+-0.01 & 0.635+-0.011 & 0.796+-0.016 & 0.452+-0.007 \\
+		Random & 0.755+-0.014 & 0.612+-0.039 & 0.763+-0.012 & 0.818+-0.009 & 0.439+-0.019 & 0.622+-0.013 & 0.633+-0.012 & 0.795+-0.016 & 0.45+-0.006 \\
+		Coreset & 0.759+-0.016 & 0.601+-0.034 & 0.764+-0.015 & 0.757+-0.015 & 0.39+-0.029 & 0.647+-0.009 & 0.651+-0.011 & 0.784+-0.026 & 0.435+-0.012 \\
+		TypiClust & 0.751+-0.012 & 0.551+-0.036 & 0.76+-0.016 & 0.643+-0.026 & 0.411+-0.024 & 0.488+-0.02 & 0.449+-0.017 & 0.652+-0.035 & 0.406+-0.011
+	\end{tabular}
+}
+\end{table}
+% QS 50
+\begin{table}[H]
+	\caption{AUC values for each dataset that supports query size 50.}
+	\resizebox{\columnwidth}{!}{
+	\begin{tabular}{l|llllllll}
+		       & Splice       & DNA          & USPS         & USPSEncoded  & Cifar10Encoded& FashionMnistEnc& TopV2     & News \\
+		\hline
+		Oracle & 0.803+-0.012 & 0.825+-0.009 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.884+-0.006 & 0.49+-0.003 \\
+		Margin & 0.747+-0.023 & 0.751+-0.019 & 0.828+-0.009 & 0.363+-0.031 & 0.64+-0.013 & 0.653+-0.01 & 0.774+-0.029 & 0.46+-0.006 \\
+		Badge & 0.758+-0.017 & 0.754+-0.018 & 0.831+-0.008 & 0.376+-0.028 & 0.632+-0.013 & 0.649+-0.011 & 0.781+-0.026 & 0.462+-0.007 \\
+		LeastConfident & 0.731+-0.025 & 0.688+-0.041 & 0.761+-0.037 & 0.291+-0.03 & 0.644+-0.013 & 0.65+-0.011 & 0.73+-0.049 & 0.462+-0.009 \\
+		DSA & 0.748+-0.021 & 0.738+-0.018 & 0.783+-0.016 & 0.346+-0.027 & 0.624+-0.014 & 0.588+-0.016 & 0.748+-0.041 & 0.45+-0.011 \\
+		BALD & 0.76+-0.017 & 0.756+-0.018 & 0.796+-0.016 & 0.241+-0.026 & 0.65+-0.009 & 0.645+-0.01 & 0.746+-0.038 & 0.455+-0.007 \\
+		CoreGCN & 0.755+-0.016 & 0.745+-0.018 & 0.752+-0.019 & 0.328+-0.027 & 0.581+-0.015 & 0.568+-0.018 & 0.771+-0.025 & 0.453+-0.007 \\
+		Entropy & 0.747+-0.024 & 0.748+-0.018 & 0.778+-0.024 & 0.275+-0.026 & 0.633+-0.011 & 0.625+-0.012 & 0.734+-0.036 & 0.442+-0.007 \\
+		LSA & 0.754+-0.013 & 0.749+-0.019 & 0.807+-0.01 & 0.341+-0.029 & 0.613+-0.012 & 0.625+-0.01 & 0.763+-0.025 & 0.45+-0.006 \\
+		Random & 0.746+-0.012 & 0.745+-0.015 & 0.806+-0.008 & 0.379+-0.028 & 0.615+-0.014 & 0.621+-0.01 & 0.759+-0.026 & 0.448+-0.006 \\
+		Coreset & 0.751+-0.016 & 0.733+-0.019 & 0.74+-0.017 & 0.325+-0.034 & 0.624+-0.012 & 0.608+-0.013 & 0.731+-0.045 & 0.432+-0.012 \\
+		TypiClust & 0.749+-0.016 & 0.736+-0.016 & 0.586+-0.038 & 0.348+-0.027 & 0.451+-0.024 & 0.375+-0.022 & 0.614+-0.046 & 0.397+-0.012
+	\end{tabular}
+}
+\end{table}
+% QS 100
+\begin{table}[H]
+	\caption{AUC values for each dataset that supports query size 100.}
+	\resizebox{\columnwidth}{!}{
+	\begin{tabular}{l|lllllll}
+		       & Splice       & DNA          & USPS         & USPSEncoded & Cifar10Encoded& FashionMnistEnc& News \\
+		\hline
+		Oracle & 0.803+-0.012 & 0.825+-0.009 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.49+-0.003 \\
+		Margin & 0.733+-0.024 & 0.711+-0.027 & 0.799+-0.013 & 0.473+-0.026 & 0.629+-0.012 & 0.628+-0.009 & 0.455+-0.006 \\
+		Badge & 0.743+-0.014 & 0.714+-0.032 & 0.804+-0.013 & 0.472+-0.029 & 0.623+-0.01 & 0.621+-0.01 & 0.456+-0.006 \\
+		LeastConfident & 0.715+-0.033 & 0.639+-0.05 & 0.708+-0.034 & 0.23+-0.034 & 0.631+-0.013 & 0.62+-0.012 & 0.457+-0.008 \\
+		DSA & 0.729+-0.021 & 0.697+-0.031 & 0.753+-0.021 & 0.427+-0.028 & 0.609+-0.013 & 0.546+-0.017 & 0.442+-0.01 \\
+		BALD & 0.744+-0.015 & 0.718+-0.024 & 0.765+-0.021 & 0.285+-0.046 & 0.632+-0.009 & 0.609+-0.01 & 0.444+-0.007 \\
+		CoreGCN & 0.742+-0.015 & 0.713+-0.025 & 0.744+-0.019 & 0.433+-0.032 & 0.583+-0.013 & 0.554+-0.015 & 0.448+-0.007 \\
+		Entropy & 0.733+-0.023 & 0.713+-0.031 & 0.743+-0.026 & 0.395+-0.037 & 0.618+-0.012 & 0.59+-0.012 & 0.432+-0.007 \\
+		LSA & 0.738+-0.017 & 0.716+-0.027 & 0.789+-0.011 & 0.439+-0.03 & 0.609+-0.013 & 0.608+-0.01 & 0.447+-0.006 \\
+		Random & 0.733+-0.013 & 0.713+-0.023 & 0.789+-0.012 & 0.468+-0.024 & 0.611+-0.01 & 0.606+-0.01 & 0.446+-0.005 \\
+		Coreset & 0.735+-0.019 & 0.698+-0.026 & 0.721+-0.021 & 0.396+-0.024 & 0.608+-0.012 & 0.562+-0.016 & 0.426+-0.012 \\
+		TypiClust & 0.733+-0.016 & 0.704+-0.025 & 0.592+-0.042 & 0.427+-0.027 & 0.501+-0.02 & 0.338+-0.02 & 0.383+-0.012
+	\end{tabular}
+}
+\end{table}
+\begin{minipage}{0.47\linewidth}
+	% QS 500
+	\begin{table}[H]
+		\caption{AUC values for each dataset that supports query size 500.}
+		\begin{tabular}{l|ll}
+			& Cifar10 & FashionMnist \\
+			\hline
+			Oracle & 0.689+-0.001 & 0.905+-0.001 \\
+			Margin & 0.556+-0.008 & 0.882+-0.004 \\
+			Badge & 0.56+-0.008 & 0.883+-0.005 \\
+			LeastConfident & 0.591+-0.01 & 0.884+-0.005 \\
+			DSA & 0.56+-0.009 & 0.882+-0.004 \\
+			BALD & 0.478+-0.014 & 0.878+-0.003 \\
+			CoreGCN & 0.553+-0.01 & 0.88+-0.007 \\
+			Entropy & 0.553+-0.009 & 0.882+-0.006 \\
+			LSA & 0.558+-0.01 & 0.866+-0.005 \\
+			Random & 0.557+-0.01 & 0.863+-0.005 \\
+			Coreset & 0.553+-0.007 & 0.878+-0.006 \\
+			TypiClust & 0.557+-0.009 & 0.864+-0.004
+		\end{tabular}
+	\end{table}
+\end{minipage}
+\hspace{10mm}
+\begin{minipage}{0.47\linewidth}
+	% QS 1000
+	\begin{table}[H]
+		\caption{AUC values for each dataset that supports query size 1000.}
+		\begin{tabular}{l|ll}
+			& Cifar10 & FashionMnist \\
+			\hline
+			Oracle & 0.689+-0.001 & 0.905+-0.001 \\
+			Margin & 0.56+-0.011 & 0.872+-0.007 \\
+			Badge & 0.562+-0.013 & 0.871+-0.007 \\
+			LeastConfident & 0.561+-0.012 & 0.873+-0.006 \\
+			DSA & 0.56+-0.011 & 0.87+-0.008 \\
+			BALD & 0.535+-0.011 & 0.866+-0.003 \\
+			CoreGCN & 0.557+-0.011 & 0.867+-0.012 \\
+			Entropy & 0.557+-0.014 & 0.871+-0.009 \\
+			LSA & 0.551+-0.012 & 0.854+-0.009 \\
+			Random & 0.55+-0.01 & 0.855+-0.006 \\
+			Coreset & 0.562+-0.012 & 0.869+-0.004 \\
+			TypiClust & 0.552+-0.011 & 0.854+-0.009
+		\end{tabular}
+	\end{table}
+\end{minipage}
 
@@
 However, the hypothetical \textbf{tree} of all possible sequences of samples from $\mathcal{U}$ remains the same, granting every acquisition function equal possibilities.
-
-\section{Number of Restarts Ablation for "Training from Scratch"}\label{app:restarts_from_scratch}
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\linewidth]{img/ablation_cifar_restarts}
-    \caption{Created by the protocol described in Sec.
-\ref{sec:evaluation} for the Encoded Cifar10 dataset, with the classifier being trained from scratch.
-We observe similar characteristics, where a low number of restarts leads to extreme variance in the reported results.}
-\end{figure}
 

### 2024-06-03 core/evaluation.py

@@ -234,8 +234,10 @@ def generate_auc_leaderboard(sorted_agents:list, precision=3, add_std=True, subs
                 result_df = pd.concat([result_df, auc_values])
             result_df.to_csv(f"results/auc{postfix}_qs{q}.csv")
 
-    datasets_raw = ["Splice", "DNA", "USPS", "Cifar10", "FashionMnist", "TopV2", "News",]
-                    #"DivergingSin", "ThreeClust"]
+    datasets_raw = ["Splice", "SpliceEncoded", "DNA", "DNAEncoded", "USPS", "USPSEncoded",
+                    "Cifar10", "Cifar10Encoded", "FashionMnist", "FashionMnistEncoded",
+                    "TopV2", "News",
+                    "DivergingSin", "ThreeClust"]
     df_raw = combine_agents_into_df(dataset=datasets_raw, include_oracle=True, subsample_runs=subsample_runs)
     df_raw = average_out_columns(df_raw, ["iteration"])
     df_raw = std_for_column(df_raw, "auc")
@@ -244,15 +246,15 @@ def generate_auc_leaderboard(sorted_agents:list, precision=3, add_std=True, subs
     df_raw = df_raw.drop("auc_std", axis=1)
     save_leaderboard(df_raw, datasets_raw, postfix="")
 
-    datasets_encoded = ["SpliceEncoded", "DNAEncoded", "USPSEncoded",
-                        "Cifar10Encoded", "FashionMnistEncoded"]
-    df_enc = combine_agents_into_df(dataset=datasets_encoded, include_oracle=True, subsample_runs=subsample_runs)
-    df_enc = average_out_columns(df_enc, ["iteration"])
-    df_enc = std_for_column(df_enc, "auc")
-    df_enc = average_out_columns(df_enc, ["trial"])
-    df_enc["auc"] = df_enc["auc"].round(precision).astype(str) + "+-" + df_enc["auc_std"].round(precision).astype(str)
-    df_enc = df_enc.drop("auc_std", axis=1)
-    save_leaderboard(df_enc, datasets_encoded, postfix="_enc")
+    # datasets_encoded = [
+    #                     ]
+    # df_enc = combine_agents_into_df(dataset=datasets_encoded, include_oracle=True, subsample_runs=subsample_runs)
+    # df_enc = average_out_columns(df_enc, ["iteration"])
+    # df_enc = std_for_column(df_enc, "auc")
[…]

### 2024-06-03 evaluate_oracle.py
paper und ergebnisse

@@ -12,8 +12,8 @@ parser.add_argument("--data_folder", type=str, required=True)
 parser.add_argument("--run_id", type=int, default=1)
 parser.add_argument("--pool_seed", type=int, default=1)
 parser.add_argument("--model_seed", type=int, default=1)
-parser.add_argument("--dataset", type=str, default="Cifar10")
-parser.add_argument("--encoded", type=int, default=1)
+parser.add_argument("--dataset", type=str, default="TopV2")
+parser.add_argument("--encoded", type=int, default=0)
 parser.add_argument("--sample_size", type=int, default=20)
 parser.add_argument("--restarts", type=int, default=50)
 parser.add_argument("--store_dataset", type=bool, default=False)
@@ -60,13 +60,14 @@ while run_id < max_run_id:
         done = False
         dataset.reset()
         state = env.reset()
-        iterations = math.ceil(env.env.budget / args.query_size)
+        iterations = math.ceil(env.env.budget / args.points_per_iter)
         iterator = tqdm(range(iterations), miniters=2)
         for i in iterator:
             state, reward, done, truncated, info = env.step()
             if done or truncated:
                 break # fail save; should not happen
 
+    print("Times the oracle needed the fallback", env.env.fall_back_counter)
     save_meta_data(log_path, None, env, dataset, config)
     if args.store_dataset:
         # store dataset for later HP optimization

### 2024-06-03 doc/al_benchmark_paper.tex
paper und ergebnisse
Moved the Introduction of the synthetic datasets and related experiments to a seperate section.

@@
 The checkpoint of each encoder model will be provided together with the framework. \\ [1mm]
-\textbf{Synthetic Data}
-Existing AL algorithms can broadly be categorized into two types, uncertainty \cite{wang2014new, gal2017deep} - and geometric-approaches \cite{sener2017active, hacohen2022active, ashdeep}. 
-Both types have principled shortcomings in terms of the utilized information that makes them unsuitable for certain data distributions.
-To test for these shortcomings, we created two synthetic datasets, namely "Honeypot" and "Diverging Sine", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively.
-To avoid algorithms memorizing these datasets they are generated during the experiment, depending on $s_\D$. \\
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.4\linewidth]{img/diverging_sin.jpg}
-	\includegraphics[width=0.4\linewidth]{img/honeypot.jpg}
-	\caption{Synthetic "Honeypot" and "Diverging Sine" datasets. The decision boundary is not part of the dataset and serves only as a visual guide.}
-        \label{fig:synthDataAppendix}
-\end{figure}
-Honeypot creates to two easy to distinguish clusters with 150 samples each and one overlapping "honeypot" that represents a noisy region of the dataset with potentially miss-labeled, miss-measured or generally adverse samples.
-This honeypot contains 150 samples of each class, creating a balance of 50\% beneficial samples and 50\% adverse samples in the dataset.
-The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its negative impact on purely uncertainty based acquisition functions.
-Diverging Sine samples the datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.
-This creates a challenging region one the left hand side, where a lot of datapoints need to be sampled and an easy region on the right hand side, where very few datapoints are enough. 
-The repeating nature of a sin function encourages diversity based acquisition functions to equally sample the entire length, drastically oversampling the right hand side of the dataset.
-Each class has 500 datapoints. \\ [1mm]
 Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.
@@
         Enc. DNA    & 40 & o & o &&&&& \\
-        Honeypot    & 60 & o & o &&&&& \\
-        Diverging Sine& 60 & o & o &&&&& \\
         Enc. Splice & 100 & o & o & o & o &&&\\
@@
 %
-\begin{figure}
-    \centering
-    \caption{Results for all acquisition functions on both synthetic datasets.}
-    \label{fig:main_body_result}
-    \includegraphics[width=0.49\linewidth]{img/micro_honeypot.jpg}
-    \includegraphics[width=0.49\linewidth]{img/micro_diverging_sin.jpg}
-\end{figure}
+
 Storing the results of every run in a separate file allows us to create various aggregates.
@@
 The leading algorithms Margin-Sampling and BADGE show very stable performance across all tested data domains. \\ [1mm]
-Results for the Honeypot dataset reveal expected shortcomings of uncertainty sampling algorithms like margin, entropy and least confident sampling as well as BALD.
-In addition, BADGE is underperforming for this dataset compared to real-world data. 
-Results for Diverging Sine also confirm expected behavior, as clustering algorithms (Coreset, TypiClust) fall behind uncertainty algorithms (Entropy-, Margin-Sampling), with the exception of BADGE. \\
-We provide a very small ablation study on the importance of the embeddings by testing a version of Coreset and TypiClust on this dataset that does not use the embeddings produced by the classification model, but rather clusters the data directly.
-"Coreset Raw" and "TypiClust Raw" both perform worse than their embedding-based counterpart.
+
+\include{synthetic_datasets}
 

### 2024-06-03 doc/synthetic_datasets.tex
Moved the Introduction of the synthetic datasets and related experiments to a seperate section.

\section{Synthetic Datasets for AL}

AL approaches can be categorized into two types, uncertainty \cite{wang2014new, gal2017deep} - and geometric-approaches \cite{sener2017active, hacohen2022active, ashdeep}. Both types have principled shortcomings in terms of the utilized information that makes them unsuitable for certain data distributions. To test for these specific shortcomings, we created two synthetic datasets, namely "Honeypot" and "Diverging Sine", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively. 
To avoid algorithms memorizing these datasets they are generated from scratch for each experiment, depending on $s_\D$. \\
\begin{figure}[H]
	\centering
	\includegraphics[width=0.4\linewidth]{img/diverging_sin.jpg}
	\includegraphics[width=0.4\linewidth]{img/honeypot.jpg}
	\caption{Synthetic "Honeypot" and "Diverging Sine" datasets. The optimal decision boundary is not part of the dataset and serves only as a visual guide.}
	\label{fig:synthDataAppendix}
\end{figure}

\begin{wraptable}{R}{0.65\textwidth}
	\vspace{-0.7cm}
	\caption{Overview of available batch sizes for our synthetic datasets}
	\vspace{0.1cm}
	\label{tab:batch_sizes_syn}
	\begin{tabular}{l|c|c|c|c|c|c|c|c}
		& B & 1 & 5 & 20 & 50 & 100 & 500 & 1000 \\
		\hline
		Honeypot    & 60 & o & o &&&&& \\
		Diverging Sine& 60 & o & o &&&&& \\
	\end{tabular}
	\vspace{-0.35cm}
\end{wraptable}
Honeypot creates to two easy to distinguish clusters with 150 samples each and one overlapping "honeypot" that represents a noisy region of the dataset with potentially miss-labeled, miss-measured or generally adverse samples.
This honeypot contains 150 samples of each class, creating a balance of 50\% beneficial samples and 50\% adverse samples in the dataset.
The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its negative impact on purely uncertainty based acquisition functions.
Diverging Sine samples the datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.
This creates a challenging region one the left hand side, where a lot of datapoints need to be sampled and an easy region on the right hand side, where very few datapoints are enough. 
The repeating nature of a sin function encourages diversity based acquisition functions to equally sample the entire length, drastically oversampling the right hand side of the dataset.
Each class has 500 datapoints. The available batch sizes for our proposed datasets are shown in Table \ref{tab:batch_sizes_syn}.\\

Results for the Honeypot dataset reveal expected shortcomings of uncertainty sampling algorithms like margin, entropy and least confident sampling as well as BALD.
In addition, BADGE is underperforming for this dataset compared to real-world data. 
Results for Diverging Sine also confirm expected behavior, as clustering algorithms (Coreset, TypiClust) fall behind uncertainty algorithms (Entropy-, Margin-Sampling), with the exception of BADGE. \\
We provide a very small ablation study on the importance of the embeddings by testing a version of Coreset and TypiClust on this dataset that does not use the embeddings produced by the classification model, but rather clusters the data directly.
"Coreset Raw" and "TypiClust Raw" both perform worse than their embedding-based counterpart.

\begin{figure}
	\centering
	\caption{Results for all acquisition functions on both synthetic datasets.}
	\label{fig:main_body_result}
	\includegraphics[width=0.49\linewidth]{img/micro_diverging_sin.jpg}
	\includegraphics[width=0.49\linewidth]{img/micro_honeypot.jpg}
\end{figure}

### 2024-06-03 doc/al_benchmark_paper.tex
paper

--- 2024-06-03 doc/al_benchmark_paper.tex
@@
 % - Wir erhöhen die Anzahl der Experimente exponentiell: Ji und Lueth: 300-400; wir: 24200+Oracle; Größtes Benchmark bis jetzt ist Zhan (2021): ~31500 (nur tabular)
+% 3 runs vs 50 runs -> statistik
 % Why is that important?
@@
 % !! Sec 4 Methodology ist sehr chaotisch. Wir müssen neu sortieren und unwichtige Dinge streichen
-% !! Sec 3 Related Work sollte Ji et al. im Detail einführen, weil wir uns an deren Guidelines halten. Alles andere sollte dort gekürzt werden
 % Contributions neu sortieren und ausführlicher beschreiben
@@
 % Eventuell Batchsize 1 von den anderen Ergebnissen separieren. Der Vergleich macht mehr Sinn, als einfach die Batchsizes zu mitteln
-% Infos und Ergebnisse für die syntetischen Datensätze in ein Kapitel bündeln und direkt auf die Contribution beziehen
 % Tabelle 1 für Algorithmen statt Domänen. Das zeigt, dass immer entweder BADGE oder Margin (unsere Top 2) fehlt
@@
 % Conclusion: Es gibt Unterschiede bei der Top 1 zwischen den Domänen, aber die Top 3 ist konstant
+% !! Sec 3 Related Work sollte Ji et al. im Detail einführen, weil wir uns an deren Guidelines halten. Alles andere sollte dort gekürzt werden
+% Infos und Ergebnisse für die syntetischen Datensätze in ein Kapitel bündeln und direkt auf die Contribution beziehen
 
@@
 To the best of our knowledge, we propose the first AL benchmark that applies state-of-the-art evaluation on algorithms in 3 major domains: Tabular, Image, and Text as well as synthetic data.
-We report empirical results for 10 widely used algorithms on 7 real-world and 2 synthetic datasets and aggregate them into domain-specific and overall rankings of AL algorithms.
+We report empirical results for 11 widely used algorithms on 7 real-world and 2 synthetic datasets and aggregate them into domain-specific and overall rankings of AL algorithms.
 \end{abstract}
@@
 	\item[C2] Efficient and performant algorithm for an oracle that can be constructed greedily and does not rely on search
-	\item[C3] Two novel synthetic datasets that highlight principled shortcomings of of the top-performing AL algorithms
+	\item[C3] Two novel synthetic datasets named "Honeypot" and "Diverging Sine" that highlight two principled shortcomings AL algorithms.
+Firstly, a susceptibility to noisy or adverse samples and secondly, the oversampling of easy to distinguish regions of the dataset
     \item[C4] Extending the evaluation of Active Learning algorithms to 5 data domains and revealing significant differences in algorithm performance between them
@@
 \section{Related Work}
-Multiple benchmark suites have already been proposed for Active Learning:
-The authors of \cite{beck2021effective}, \cite{munjal2022towards}, \cite{li2022empirical}, \cite{ji2023randomness} and \cite{luth2024navigating} focus exclusively on batch AL in the image domain.
-While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training, and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
-The authors of \cite{hu2021towards} study models that are trained on actively learned datasets in the image and text domain.
-They test for several different properties of the models including robustness, response to compression techniques and final performance.
-\cite{zhou2021towards} proposed an oracle algorithm for AL that uses Simulated Annealing search to approximate a solution for the optimal subset of labeled data.
-Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
-The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also discussed the problems of evaluating AL algorithms under many forms of variance.
-We largely adapt the proposed guidelines of \cite{ji2023randomness} and extend their work to multiple domains, batch sizes and comparisons. 
-This work also pays attention to the so-called "pitfalls" of AL evaluation mentioned in \cite{luth2024navigating} with the exception of P1 - Controlling the data distribution of the tested datasets.
-We leave this aspect for future versions of this benchmark. \\
-Two preceding benchmarks have also proposed an oracle algorithm to find a near-optimal labeled set $\LL$, however, both algorithms rely on search and are for that reason considerably slower than our proposed greedy oracle.
+While multiple benchmark suites have been proposed for Active Learning, none of them provide experiments for more than two domains.
+The authors of \cite{beck2021effective}, \cite{munjal2022towards}, \cite{li2022empirical}, \cite{ji2023randomness} and \cite{luth2024navigating} even focus exclusively on the image domain.
+Experiments on the interplay between AL and semi-supervised learning have only been provided by two works so far \cite{li2022empirical, luth2024navigating}, both of them only for images.
+An oracle algorithm has so far been proposed by only two works \cite{zhou2021towards, zhan2022comparative}. 
+Both of these algorithms rely on search, while our proposed method can be constructed sequentially.
+%While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training, and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
+%The authors of \cite{hu2021towards} study models that are trained on actively learned datasets in the image and text domain.
+%They test for several different properties of the models including robustness, response to compression techniques and final performance.
+%\cite{zhou2021towards} proposed an oracle algorithm for AL that uses Simulated Annealing search to approximate a solution for the optimal subset of labeled data.
+%Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
+The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also place a much higher emphasis on the problem of evaluating AL algorithms under many forms of variance than their predecessors (indicated in Tab.
+\ref{tab:benchmark_comparison} by a dashed line).
+The authors of \cite{ji2023randomness} posed a total of 12 "recommendations" for reliable evaluation of AL algorithms.
+We largely adapt the proposed recommendations of \cite{ji2023randomness} and extend their work to multiple domains, batch sizes and comparisons.
+For a complete list of the recommendations and our implementation of them, please refer to App. \ref{app:recommendations}.
+This work also pays attention to the so-called "pitfalls" of AL evaluation proposed in \cite{luth2024navigating}.
+For a complete list of the pitfalls and our implementation of them, please refer to App. \ref{app:pitfalls}.
+To the best of our knowledge, we are the first to extend reliable SOTA (based on \cite{ji2023randomness, luth2024navigating}) experimentation to a total of 5 data domains.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation Protocol}\label{sec:evaluation}
-Following \cite{zhou2021towards}, the quality of an AL algorithm is evaluated by an ``anytime protocol" that incorporates classification performance at every iteration, as opposed to evaluating final performance after the budget is exhausted.
-We employ the normalized area under the accuracy curve (AUC):
-\begin{equation}\label{eq:auc}
-	\operatorname{AUC}(\D_\test, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\D_\test, \hat y^{(i)})
-\end{equation}
-The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
-AL algorithms have to perform well in both scenarios. 
-\\ [1mm]
-Since AUC is still influenced by the budget, we define a set of rules to set this hyperparameter upfront, so that we are not favoring a subset of algorithms by handcrafting a budget.
-In this work, we choose the budget per dataset to be the first point at which one of 2 stopping conditions apply: (i) an algorithm (except Oracle) manages to reach 99\% of the full-dataset-performance (using the smallest query size) or (ii) the best algorithm (except oracle) did not improve the classifier's accuracy by at least 2\% in the last 20\% of iterations.
-\\ [1mm]
-%\footnote{Since we only consider single-sample AL, increasing the budget comes at a high computational cost. Therefore we could not freely increase the budget above 2000.} 
-To mimic the leave-one-out protocol for cross-validation, we will restart each experiment multiple times.
-Each restart retains the train/test split (often given by the dataset itself), but introduces a new validation split that is sampled from the entire dataset (not just $\LL$).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Why we need 50 restarts}\label{sec:restarts}
-To evaluate how many restarts are necessary to obtain conclusive and reproducible results in an AL experiment, we set the number of restarts fore margin sampling on the Splice dataset to 100.
+To evaluate how many restarts are necessary to obtain conclusive and reproducible results in an AL experiment, we computed 100 runs of our top-performing algorithm on one dataset.
+Our best algorithm is margin sampling and we chose the Splice dataset for its average size and complexity. \\
 This allows us firstly, to obtain a very strong estimation of the "true" average performance of margin sampling on this dataset and secondly, to draw subsets from this pool of 100 runs.
 Setting the size of our draws to $\alpha$ and sampling uniformly, we can approximate a cross-validation process with $\alpha$ restarts.
-Each of these draws can be interpreted as a reported result in AL literature where the authors employed $\alpha$ restarts.
+Each of these draws can be interpreted as a \textbf{reported result in AL literature} where the authors employed $\alpha$ restarts.
 %We provide evidence in Fig. \ref{fig:restarts} that previous works might not have evaluated their experiments with a sufficient number of restarts.
@@
 This details can be found in App. \ref{app:rank_difference}.
-For this benchmark we always employ 50 restarts of the same experiment.
+For this benchmark we employ 50 restarts of every experiment.
 
@@
 The checkpoint of each encoder model will be provided together with the framework. \\ [1mm]
-\textbf{Synthetic Data}
-Existing AL algorithms can broadly be categorized into two types, uncertainty \cite{wang2014new, gal2017deep} - and geometric-approaches \cite{sener2017active, hacohen2022active, ashdeep}. 
-Both types have principled shortcomings in terms of the utilized information that makes them unsuitable for certain data distributions.
-To test for these shortcomings, we created two synthetic datasets, namely "Honeypot" and "Diverging Sine", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively.
-To avoid algorithms memorizing these datasets they are generated during the experiment, depending on $s_\D$. \\
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.4\linewidth]{img/diverging_sin.jpg}
-	\includegraphics[width=0.4\linewidth]{img/honeypot.jpg}
-	\caption{Synthetic "Honeypot" and "Diverging Sine" datasets. The decision boundary is not part of the dataset and serves only as a visual guide.}
-        \label{fig:synthDataAppendix}
-\end{figure}
-Honeypot creates to two easy to distinguish clusters with 150 samples each and one overlapping "honeypot" that represents a noisy region of the dataset with potentially miss-labeled, miss-measured or generally adverse samples.
-This honeypot contains 150 samples of each class, creating a balance of 50\% beneficial samples and 50\% adverse samples in the dataset.
-The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its negative impact on purely uncertainty based acquisition functions.
-Diverging Sine samples the datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.
-This creates a challenging region one the left hand side, where a lot of datapoints need to be sampled and an easy region on the right hand side, where very few datapoints are enough. 
-The repeating nature of a sin function encourages diversity based acquisition functions to equally sample the entire length, drastically oversampling the right hand side of the dataset.
-Each class has 500 datapoints. \\ [1mm]
 Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Classification Model}\label{sec:choosing_the_classifier}
 \begin{wraptable}{R}{0.65\textwidth}
     \vspace{-0.7cm}
-    \caption{Overview of available batch sizes for each dataset}
+    \caption{Employed model, chosen budget and available batch sizes for each dataset}
     \vspace{0.1cm}
     \label{tab:batch_sizes}
-    \begin{tabular}{l|c|c|c|c|c|c|c|c}
-                    & B & 1 & 5 & 20 & 50 & 100 & 500 & 1000 \\
-         \hline
-        Enc. DNA    & 40 & o & o &&&&& \\
-        Honeypot    & 60 & o & o &&&&& \\
-        Diverging Sine& 60 & o & o &&&&& \\
-        Enc. Splice & 100 & o & o & o & o &&&\\
-        TopV2       & 200 & o & o & o & o &&& \\
-        Splice      & 400 & o & o & o & o & o && \\
-        DNA         & 300 & o & o & o & o & o && \\
-        USPS        & 400 & o & o & o & o & o && \\
-        Enc. Cifar10& 450 & o & o & o & o & o && \\
-        Enc. FMnist & 500 & o & o & o & o & o && \\
-        Enc. USPS   & 600 & o & o & o & o & o && \\
-        News        & 3K &&& o & o & o & o &\\
-        FMnist      & 10K &&&&&& o & o\\
-        Cifar10     & 10K &&&&&& o & o \\
-    \end{tabular}
-    \vspace{-0.35cm}
+    {\scriptsize
+    \begin{tabular}{l|l|c|c|c|c|c|c|c|c}
+                    & Model  & B & 1 & 5 & 20 & 50 & 100 & 500 & 1K \\
+        \hline
+        Enc. DNA    & Linear & 40 & o & o &&&&& \\
+        Enc. Splice & Linear & 100 & o & o & o & o &&&\\
+        TopV2       & BiLSTM & 200 & o & o & o & o &&& \\
+        Splice      & MLP    & 400 & o & o & o & o & o && \\
+        DNA         & MLP    & 300 & o & o & o & o & o && \\
+        USPS        & MLP    & 400 & o & o & o & o & o && \\
+        Enc. Cifar10& Linear & 450 & o & o & o & o & o && \\
+        Enc. FMnist & Linear & 500 & o & o & o & o & o && \\
+        Enc. USPS   & Linear & 600 & o & o & o & o & o && \\
+        News        & BiLSTM & 3K &&& o & o & o & o &\\
+        FMnist      & ResNet18& 10K &&&&&& o & o\\
+        Cifar10     & ResNet18& 10K &&&&&& o & o \\
+    \end{tabular}
+	}
+    \vspace{-0.85cm}
 \end{wraptable}
-The model $\hat y$ can be trained in two ways.
-Either the parameters of the model are reset to a fixed initial setting $\hat y^{(0)}$ after each AL iteration and the classifier is trained from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or the previous state $\hat y^{(i-1)}$ is retained and the classifier is fine-tuned on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
-In this work, we use the fine-tuning method for un-encoded datasets to save computational time, while we use the from-scratch training for embedded datasets since they have very small classifiers and this approach generally produces better results.
-Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping with a patience of 2 afterwards.
 
@@
 Every encoded dataset is classified by a single linear layer with softmax activation.
-For an overview of architectures and hyperparameters please refer to Appendix \ref{app:hyperparameters}.
+The used model for each dataset can be found in Tab. \ref{tab:batch_sizes}.
+For a detailed description and employed hyperparameters please refer to Appendix \ref{app:hyperparameters}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evaluation Protocol}\label{sec:evaluation}
+Following \cite{zhou2021towards}, the quality of an AL algorithm is evaluated by an ``anytime protocol" that incorporates classification performance at every iteration, as opposed to evaluating final performance after the budget is exhausted.
+We employ the normalized area under the accuracy curve (AUC):
+\begin{equation}\label{eq:auc}
+	\operatorname{AUC}(\D_\test, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\D_\test, \hat y^{(i)})
+\end{equation}
+The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
+AL algorithms have to perform well in both scenarios. 
+\\ [1mm]
+Since AUC is still influenced by the budget, we define a set of rules to set this hyperparameter upfront, so that we are not favoring a subset of algorithms by handcrafting a budget.
+In this work, we choose the budget per dataset to be the first point at which one of 2 stopping conditions apply: (i) an algorithm (except Oracle) manages to reach 99\% of the full-dataset-performance (using the smallest query size) or (ii) the best algorithm (except oracle) did not improve the classifier's accuracy by at least 2\% in the last 20\% of iterations.
+\\ [1mm]
+%\footnote{Since we only consider single-sample AL, increasing the budget comes at a high computational cost. Therefore we could not freely increase the budget above 2000.} 
+As described in Sec. \ref{sec:restarts}, we will restart each experiment multiple times.
+Each restart retains the train/test split (often given by the dataset itself), but creates a new validation split that is sampled (based on $s_\D$) from the entire dataset (not just the seed set $\LL_0$).
+\\ [1mm]
+%
+Apart from plotting standard performance curves and reporting their AUC values per dataset in App.
+\ref{app:all_results}, we primarily rely on ranks to aggregate the performance of an acquisition function across datasets.
+For each dataset and query size, the AUC values of all acquisition functions are sorted and assigned a rank based on position, with the best rank being 1.
+These ranks can safely be averages across datasets as they are no longer subjected to scaling differences of each dataset.
+Additionally, we employ Critical Difference (CD) diagrams (like Fig. \ref{fig:ranks_by_domain}) for statistical testing.
+CD diagrams use the Wilcoxon signed-rank test, which is a variant of the paired T-test, to find significant differences of ranks between acquisition functions.
+For these diagrams, each combination of dataset, query size and run is considered a separate experiment, i.e.
+the results of \verb|Dataset1-QuerySize1-run5| of an acquisition function \verb|x| is only compared to the results of \verb|Dataset1-QuerySize1-run5| of acquisition function \verb|y|.
+Due to the large number of restarts and the wide range of datasets and query sizes, we can provide very accurate significance tests.
+For a detailed description of how every CD diagram is created, please refer to App. \ref{app:cd_diagrams}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
+The model $\hat y$ can be trained in two ways.
+Either the parameters of the model are reset to a fixed initial setting $\hat y^{(0)}$ after each AL iteration and the classifier is trained from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or the previous state $\hat y^{(i-1)}$ is retained and the classifier is fine-tuned on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
+In this work, we use the fine-tuning method for un-encoded datasets to save computational time, while we use the from-scratch training for embedded datasets since they have very small classifiers and this approach generally produces better results.
+Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping with a patience of 2 afterwards.
 %This work employs a validation split that is produced from the source dataset itself before active learning instead of splitting it from the seed set or omitting it altogether.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Results}
-%
+\subsection{Results on Real-world Data}
+In Table \ref{tab:results} we provide the rank of each acquisition function per dataset and averaged for each (un-)encoded dataset. 
+Please note, that for Tab \ref{tab:results} we are averaging not only over runs, but also over query sizes per dataset. 
+For the results per query size, please refer to App. \ref{app:auc_by_query_size}. \\
+As stated in contribution C4, our results on real-world data shows significant differences in the performance of the tested algorithms between data domains.
+Not only do some algorithms overperform on some domains (like least confidence sampling on Images), but the Top-3 of algorithms (except Oracle) does not contain the same three algorithms for any two domains.
+Most interestingly, the image domain, which received the most attention in benchmarking so far could even be considered an outlier, as this is the only domain where the Top-1 algorithm changes.
+This highlights the dire need for diverse data domains in AL benchmarking. \\ [1mm]
 \begin{table}[]
@@
 %
-\begin{figure}
-    \centering
-    \caption{Results for all acquisition functions on both synthetic datasets.}
-    \label{fig:main_body_result}
-    \includegraphics[width=0.49\linewidth]{img/micro_honeypot.jpg}
-    \includegraphics[width=0.49\linewidth]{img/micro_diverging_sin.jpg}
-\end{figure}
-Storing the results of every run in a separate file allows us to create various aggregates.
-Apart from plotting standard performance curves and reporting their AUC values per dataset (Fig.
-\ref{fig:main_body_result}), we primarily rely on ranks to aggregate the performance of an acquisition function across datasets.
-For each dataset, the AUC values of all acquisition functions are sorted and assigned a rank based on position, with the best rank being 1.
-These ranks can safely be averages across datasets as they are no longer subjected to scaling differences of each dataset.
-In Table \ref{tab:results} we provide the rank of each acquisition function per dataset and averaged for each (un-)encoded dataset. 
-Please note, that for Tab \ref{tab:results} we are averaging not only over runs, but also over query sizes per dataset. 
-For the results per query size, please refer to App. \ref{app:auc_by_query_size}. \\
-Additionally, we provide aggregated ranks per data domain in Fig. \ref{fig:ranks_by_domain}, revealing varying performances of acquisition functions like DSA/LSA, BALD and CoreGCN.
-The leading algorithms Margin-Sampling and BADGE show very stable performance across all tested data domains. \\ [1mm]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Synthetic Datasets for AL}
+AL approaches can be categorized into two types, uncertainty and geometric approaches.
+Typical members of the first category are variants of uncertainty sampling like entropy-, margin and least-confident-sampling \cite{wang2014new} as well as BALD \cite{gal2017deep}.
+Typical members of the second category are clustering approaches like Coreset \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
+Both types of algorithms have principled shortcomings in terms of the utilized information that makes them unsuitable for certain data distributions. 
+To test for these specific shortcomings, we created two synthetic datasets, namely "Honeypot" and "Diverging Sine", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively.
+To avoid algorithms memorizing these datasets they are generated from scratch for each experiment, depending on $s_\D$. \\
+\begin{figure}[]
+	\centering
+	\includegraphics[width=0.4\linewidth]{img/diverging_sin.jpg}
+	\includegraphics[width=0.4\linewidth]{img/honeypot.jpg}
+	\caption{Synthetic "Honeypot" and "Diverging Sine" datasets. The optimal decision boundary is not part of the dataset and serves only as a visual guide.}
+	\label{fig:synthDataAppendix}
+\end{figure}
+%
+Honeypot creates to two easy to distinguish clusters with 150 samples each and one overlapping "honeypot" that represents a noisy region of the dataset with potentially miss-labeled, miss-measured or generally adverse samples.
+This honeypot contains 150 samples of each class, creating a balance of 50\% beneficial samples and 50\% adverse samples in the dataset.
+The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its negative impact on purely uncertainty based acquisition functions.
+Diverging Sine samples the datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.
+This creates a challenging region one the left hand side, where a lot of datapoints need to be sampled and an easy region on the right hand side, where very few datapoints are enough. 
+The repeating nature of a sin function encourages diversity based acquisition functions to equally sample the entire length, drastically oversampling the right hand side of the dataset.
+Each class has 500 datapoints. 
+Both datasets have a budget of $B=60$ and are tested with query sizes 1 and 5.\\
+%
 Results for the Honeypot dataset reveal expected shortcomings of uncertainty sampling algorithms like margin, entropy and least confident sampling as well as BALD.
@@
 "Coreset Raw" and "TypiClust Raw" both perform worse than their embedding-based counterpart.
-
-\section{Conclusion}
-As stated in contribution C4, our results on real-world data shows significant differences in the performance of the tested algorithms between data domains.
-Not only do some algorithms overperform on some domains (like least confidence sampling on Images), but the Top-3 of algorithms (except Oracle) does not contain the same three algorithms for any two domains.
-This highlights the dire need for diverse data domains in AL benchmarking.
-Furthermore, our results on Honeypot reveal principled shortcomings for the two best algorithms in BADGE and margin sampling.
+\begin{figure}
+	\centering
+	\caption{Results for all acquisition functions on both synthetic datasets.}
+	\label{fig:main_body_result}
+	\includegraphics[width=0.49\linewidth]{img/micro_diverging_sin.jpg}
+	\includegraphics[width=0.49\linewidth]{img/micro_honeypot.jpg}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Results on Synthetic Data}
+Our results on Honeypot reveal principled shortcomings for the two best algorithms in BADGE and margin sampling.
 Both are vulnerable to adverse samples or simply measurement noise, which highlights the need for further research in this area. \\
@@
 In conclusion, we strongly advocate to test newly proposed AL algorithms not only on a wide variety of real data domains, but also to pay close attention to the Honeypot and Diverging Sine datasets to reveal principled shortcomings of the algorithm in question.
+
+\section{Conclusion}
+TODO
 
@@
 %\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{AL Recommendations from Ji et al.}\label{app:recommendations}
+TODO
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{AL Pitfalls from Lueth et al.}\label{app:pitfalls}
+TODO
 
@@
 % QS 1
+%TODO highlight best results
 \begin{table}[H]
@@
 		Oracle & 0.803+-0.012 & 0.678+-0.021 & 0.825+-0.009 & 0.721+-0.013 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.884+-0.006 & 0.957+-0.009 & 0.783+-0.03 \\
-		Margin & 0.769+-0.021 & 0.678+-0.032 & 0.806+-0.013 & 0.642+-0.047 & 0.858+-0.006 & 0.426+-0.038 &  0.653+-0.013 & 0.68+-0.012 & 0.861+-0.009 & 0.941+-0.018 & 0.704+-0.074 \\
+		Margin & 0.769+-0.021 & \textbf{0.678+-0.032} & 0.806+-0.013 & 0.642+-0.047 & \textbf{0.858+-0.006} & 0.426+-0.038 &  0.653+-0.013 & 0.68+-0.012 & \textbf{0.861+-0.009} & 0.941+-0.018 & 0.704+-0.074 \\
 		Badge & 0.767+-0.02 & 0.661+-0.026 & 0.78+-0.014 & 0.642+-0.046 & 0.83+-0.008 & 0.371+-0.035 &  0.656+-0.013 & 0.68+-0.009 & 0.826+-0.024 & 0.941+-0.017 & 0.69+-0.083 \\
-		LeastConfident & 0.779+-0.019 & 0.68+-0.032 & 0.809+-0.01 & 0.629+-0.05 & 0.846+-0.009 & 0.421+-0.039 &  0.668+-0.014 & 0.685+-0.009 & 0.843+-0.013 & 0.94+-0.016 & 0.692+-0.094 \\
-		DSA & 0.766+-0.021 & 0.691+-0.022 & 0.803+-0.01 & 0.646+-0.032 & 0.829+-0.01 & 0.431+-0.05 & 0.663+-0.014 & 0.679+-0.01 & 0.844+-0.017 & 0.941+-0.014 & 0.731+-0.032 \\
-		BALD & 0.78+-0.014 & 0.649+-0.04 & 0.784+-0.01 & 0.632+-0.042 & 0.819+-0.01 & 0.242+-0.046 & 0.666+-0.014 & 0.644+-0.018 & 0.815+-0.024 & 0.928+-0.014 & 0.698+-0.043 \\
-		CoreGCN & 0.765+-0.021 & 0.686+-0.023 & 0.804+-0.012 & 0.646+-0.03 & 0.753+-0.016 & 0.39+-0.044 & 0.623+-0.018 & 0.647+-0.012 & 0.85+-0.01 & 0.938+-0.014 & 0.731+-0.028 \\
-		Entropy & 0.768+-0.022 & 0.678+-0.035 & 0.812+-0.013 & 0.635+-0.045 & 0.83+-0.011 & 0.399+-0.035 & 0.663+-0.013 & 0.681+-0.011 & 0.815+-0.021 & 0.942+-0.017 & 0.696+-0.083 \\
+		LeastConfident & 0.779+-0.019 & 0.68+-0.032 & 0.809+-0.01 & 0.629+-0.05 & 0.846+-0.009 & 0.421+-0.039 &  \textbf{0.668+-0.014} & \textbf{0.685+-0.009} & 0.843+-0.013 & 0.94+-0.016 & 0.692+-0.094 \\
+		DSA & 0.766+-0.021 & 0.691+-0.022 & 0.803+-0.01 & \textbf{0.646+-0.032} & 0.829+-0.01 & \textbf{0.431+-0.05} & 0.663+-0.014 & 0.679+-0.01 & 0.844+-0.017 & 0.941+-0.014 & \textbf{0.731+-0.032} \\
+		BALD & \textbf{0.78+-0.014} & 0.649+-0.04 & 0.784+-0.01 & 0.632+-0.042 & 0.819+-0.01 & 0.242+-0.046 & 0.666+-0.014 & 0.644+-0.018 & 0.815+-0.024 & 0.928+-0.014 & 0.698+-0.043 \\
+		CoreGCN & 0.765+-0.021 & 0.686+-0.023 & 0.804+-0.012 & \textbf{0.646+-0.03} & 0.753+-0.016 & 0.39+-0.044 & 0.623+-0.018 & 0.647+-0.012 & 0.85+-0.01 & 0.938+-0.014 & \textbf{0.731+-0.028} \\
+		Entropy & 0.768+-0.022 & \textbf{0.678+-0.035} & \textbf{0.812+-0.013} & 0.635+-0.045 & 0.83+-0.011 & 0.399+-0.035 & 0.663+-0.013 & 0.681+-0.011 & 0.815+-0.021 & \textbf{0.942+-0.017} & 0.696+-0.083 \\
 		LSA & 0.772+-0.016 & 0.68+-0.026 & 0.787+-0.012 & 0.618+-0.036 & 0.821+-0.009 & 0.422+-0.037 & 0.613+-0.014 & 0.642+-0.012 & 0.816+-0.013 & 0.932+-0.016 & 0.727+-0.033 \\
@@
 \end{minipage}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Critical Difference Diagrams}\label{app:cd_diagrams}
+%TODO
 

### 2024-06-03 doc/al_benchmark_paper.tex
paper

@@
 \section{Problem Description}
-Given two spaces $\X, \Y$, $n=l+u$ data points with $l \in \mathbb{N}$ labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a model $\hat y: \X \to \Y$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathcal{X} \to \mathcal{Y}$ that can label $x$ (we consider only hard labels in the one-hot format).
+Given two spaces $\X, \Y$, $n=l+u$ data points with $l \in \mathbb{N}$ labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a model $\hat y: \X \to \Y$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathcal{X} \to \mathcal{Y}$ that can label $x$.
+% (we consider only hard labels in the one-hot format).
 We call $x \in \mathcal{X}$, $y \in \mathcal{Y}$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$. 
@@
 	\LL^{(i+1)} &\gets \LL^{(i)} \cup \{\left(x^{(i)}, A(x^{(i)})\right)\} \\
-	\U^{(i+1)} &\gets \U^{(i)} \setminus x^{(i)} %, i \in 1{:}B
+	\U^{(i+1)} &\gets \U^{(i)} \setminus \left\{x^{(i)}\right\} %, i \in 1{:}B
 \end{align*}
-with $\U^{(0)} = \operatorname{seed}(\U, s)$ and $\LL^{(0)} = \{\left(\U^{(0)}, A(\U^{(0)})\right)\}$, where $\operatorname{seed}(\U, s)$ selects $s$ points per class for the initial labeled set. \\
+with $\U^{(0)} = \operatorname{seed}(\U, s)$ and $\LL^{(0)} = \left(\U^{(0)}_i, A(\U^{(0)}_i)\right) \hspace{1mm} i \in [1, \ldots, s]$, where $\operatorname{seed}(\U, s)$ selects $s$ points per class for the initial labeled set.
+\\
 So that the average expected loss $\ell: \mathcal{Y} \times \mathcal{Y} \to \mathbb{R}$ of a machine learning algorithm fitting $\hat y^{(i)}$ on the respective labeled set $\LL^{(i)}$ is minimal: 
@@
 \subsection{Why we need 50 restarts}\label{sec:restarts}
-To evaluate how many restarts are necessary to obtain conclusive and reproducible results in an AL experiment, we computed 100 runs of our top-performing algorithm on one dataset.
+To evaluate how many restarts are necessary to obtain conclusive results in an AL experiment, we computed 100 runs of our top-performing algorithm on one dataset.
 Our best algorithm is margin sampling and we chose the Splice dataset for its average size and complexity. \\
@@
 In this work, we choose the budget per dataset to be the first point at which one of 2 stopping conditions apply: (i) an algorithm (except Oracle) manages to reach 99\% of the full-dataset-performance (using the smallest query size) or (ii) the best algorithm (except oracle) did not improve the classifier's accuracy by at least 2\% in the last 20\% of iterations.
-\\ [1mm]
-%\footnote{Since we only consider single-sample AL, increasing the budget comes at a high computational cost. Therefore we could not freely increase the budget above 2000.} 
-As described in Sec. \ref{sec:restarts}, we will restart each experiment multiple times.
-Each restart retains the train/test split (often given by the dataset itself), but creates a new validation split that is sampled (based on $s_\D$) from the entire dataset (not just the seed set $\LL_0$).
+The first rule follows \cite{ji2023randomness}, while the second rule prevents excessive budgets for cases with diminishing returns in the budget.
+The resulting budgets can be found in Tab. \ref{tab:batch_sizes}. \\ [1mm]
+As described in Sec. \ref{sec:restarts}, we restart each experiment multiple times.
+Each restart retains the train/test split (often given by the dataset itself), but creates a new validation split that is sampled (based on $s_\D$) from the entire dataset (not just the seed set $\LL^{(0)}$).
 \\ [1mm]
@@
 	\centering
+	\includegraphics[width=0.4\linewidth]{img/honeypot.jpg}
 	\includegraphics[width=0.4\linewidth]{img/diverging_sin.jpg}
-	\includegraphics[width=0.4\linewidth]{img/honeypot.jpg}
 	\caption{Synthetic "Honeypot" and "Diverging Sine" datasets. The optimal decision boundary is not part of the dataset and serves only as a visual guide.}
@@
 	\label{fig:main_body_result}
+	\includegraphics[width=0.49\linewidth]{img/micro_honeypot.jpg}
 	\includegraphics[width=0.49\linewidth]{img/micro_diverging_sin.jpg}
-	\includegraphics[width=0.49\linewidth]{img/micro_honeypot.jpg}
 \end{figure}
@@
 Finally, the fact that BADGE is able to perform well on Diverging Sine highlights the importance of embeddings for the clustering algorithms, as the so-called gradient embedding from BADGE seems to be able to encode uncertainty information, guiding the selection into the left hand regions of the dataset.
-We also show that embeddings are generally useful for this dataset, by providing results for "Coreset Raw" and "TypiClust Raw". \\ [1mm]
-In conclusion, we strongly advocate to test newly proposed AL algorithms not only on a wide variety of real data domains, but also to pay close attention to the Honeypot and Diverging Sine datasets to reveal principled shortcomings of the algorithm in question.
+We also show that embeddings are generally useful for this dataset, by providing results for "Coreset Raw" and "TypiClust Raw".
 
 \section{Conclusion}
-TODO
+We strongly advocate to test newly proposed AL algorithms not only on a wide variety of real data domains, but also to pay close attention to the Honeypot and Diverging Sine datasets to reveal principled shortcomings of the algorithm in question.
+Both tasks can be easily carried out by implementing the new acquisition function into our code base.
 

### 2024-06-03 doc/al_benchmark_paper.tex
Max: Intro und Abstract

@@
 \usepackage{neurips_data_2024}
+\usepackage{paralist}
+\usepackage{cleveref}
+\usepackage{todonotes}
 
@@
 
-\title{Towards Comparable Active Learning}
+%\title{Towards Comparable Active Learning}
+\title{A Cross-Domain Benchmark for Active Learning}
+
 \author{%
@@
 \begin{abstract}
-Active Learning has received significant attention in the field of machine learning. Its goal is to select the most informative samples for labeling, thereby reducing data annotation costs. 
-However, it has been brought to attention multiple times that reported lifts from literature generalize poorly and display high variance, leading to an inconclusive landscape in Active Learning research.
-Based on recent insights for reliable evaluation for Active Learning, this work extends experimentation from the commonly used image domain to a wide spectrum of domains.
-%This work follows up on recent benchmarking suites for Active Learning, applying the proposed solutions and evaluation protocols to a wide spectrum of domains.
-Additionally, we provide an analysis of how many repetitions an Active Learning experiment needs in order to derive conclusive results and propose that previous benchmarks have not met the necessary number of repetitions.
-%evidence that existing benchmarks might not have repeated their experiments often enough.
-To the best of our knowledge, we propose the first AL benchmark that applies state-of-the-art evaluation on algorithms in 3 major domains: Tabular, Image, and Text as well as synthetic data.
-We report empirical results for 11 widely used algorithms on 7 real-world and 2 synthetic datasets and aggregate them into domain-specific and overall rankings of AL algorithms.
+
+Active Learning (AL) deals with identifying the most informative samples for
+labeling to reduce data annotation costs for supervised learning tasks. AL
+research suffers from the fact that lifts from literature generalize poorly and
+are only conducted on a small number of repetitions of experiments. To overcome
+this obstacles, we propose \emph{ALBench}, the first active learning benchmark
+which includes tasks in computer vision, natural language processing and tabular
+learning. Furthermore, by providing an efficient greedy oracle, \emph{ALBench}
+can be evaluated with 50 runs for each experiment. We will show, that both the
+cross-domain character and the large amount of repetitions are crucial for
+sophisticated evaluation of AL research. Concretely, we will show that the
+superiority of specific methods varies over the different domains, making it
+important to evaluate Active Learning with a cross-domain benchmark.
+Additionally, we show, that having a large amount of runs is crucial. With only
+conducting five runs as often done in the literature, the superiority of
+specific methods can strongly vary with the specific runs. This effect goes so
+far, that it may even happen that dedicated methods like \emph{margin-sampling}
+are not able to outperform the random baseline.
 \end{abstract}
 
+% \begin{abstract}
+
+% 	Active Learning has received significant attention in the field of machine learning. Its goal is to select the most informative samples for labeling, thereby reducing data annotation costs. 
+% 	However, it has been brought to attention multiple times that reported lifts from literature generalize poorly and display high variance, leading to an inconclusive landscape in Active Learning research.
+% 	Based on recent insights for reliable evaluation for Active Learning, this work extends experimentation from the commonly used image domain to a wide spectrum of domains.
+% 	%This work follows up on recent benchmarking suites for Active Learning, applying the proposed solutions and evaluation protocols to a wide spectrum of domains.
+% 	Additionally, we provide an analysis of how many repetitions an Active Learning experiment needs in order to derive conclusive results and propose that previous benchmarks have not met the necessary number of repetitions.
+% 	%evidence that existing benchmarks might not have repeated their experiments often enough.
+% 	To the best of our knowledge, we propose the first AL benchmark that applies state-of-the-art evaluation on algorithms in 3 major domains: Tabular, Image, and Text as well as synthetic data.
+% 	We report empirical results for 11 widely used algorithms on 7 real-world and 2 synthetic datasets and aggregate them into domain-specific and overall rankings of AL algorithms.
+% 	\end{abstract}
+
+% \begin{abstract}
+% Active Learning has received significant attention in the field of machine
+% learning. Its goal is to select the most informative samples for labeling,
+% thereby reducing data annotation costs. However, it has been brought to
+% attention multiple times that reported lifts from literature generalize poorly
+% and display high variance, leading to an inconclusive landscape in Active
+% Learning research. Based on recent insights for reliable evaluation for Active
+% Learning, this work extends experimentation from the commonly used image domain
+% to a wide spectrum of domains.
+% %This work follows up on recent benchmarking suites for Active Learning, applying the proposed solutions and evaluation protocols to a wide spectrum of domains.
+% Additionally, we provide an analysis of how many repetitions an Active Learning
+% experiment needs in order to derive conclusive results and propose that previous
+% benchmarks have not met the necessary number of repetitions.
+% %evidence that existing benchmarks might not have repeated their experiments often enough.
+% To the best of our knowledge, we propose the first AL benchmark that applies
+% state-of-the-art evaluation on algorithms in 3 major domains: Tabular, Image,
+% and Text as well as synthetic data. We report empirical results for 10 widely
+% used algorithms on 7 real-world and 2 synthetic datasets and aggregate them into
+% domain-specific and overall rankings of AL algorithms.
+% \end{abstract}
+
+% \section{Introduction-Old Backup}
+% Deep neural networks (NN) have produced state-of-the-art results on many important supervised learning tasks.
+% Since Deep NNs usually require large amounts of labeled training data, Active Learning (AL) can be used to instead select the most informative samples out of a large pool of unlabeled data, so that only these samples need to be labeled.
+% It has been shown that a small labeled set of this nature can be used to train well-performing models \cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}. \\
+% On top of providing a principled way to label unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
+% %Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
+% In the last decade, many different algorithms for AL have been proposed.
+% Even though, almost every method has reported lifts over all its predecessors,
+% \footnote{Out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not claim a new SOTA performance in their result section.}
+% AL research faces four central difficulties:
+% {\let\thefootnote\relax\footnotetext{Code available at: anonymous}}
+% %\let\thefootnote\relax\footnotetext{Code available at: \url{https://github.com/wernerth94/comparable-active-learning}}
+% (i) The experiments are often carried out on different datasets and model architectures, hindering direct comparison,
+% (ii) generalize poorly across different domains,
+% (iii) the reported results can be subject to very high variance across restarts and
+% (iv) are not always compared against important baselines like margin sampling \cite{wang2014new}.
+% While multiple benchmark suites have been proposed to solve (i), to the best of our knowledge, we are the first to report results on all 3 data domains of tabular, image and text.
+% Additionally, we provide synthetic datasets to highlight principled shortcoming of existing AL algorithms.
+% Regarding (ii) and (iii), \cite{zhou2021towards} has pointed out severe inconsistencies in results of AL papers in recent years. 
+% They conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm (\cite{zhou2021towards} Appendix A).
+% These outlined issues lead to an inconclusive landscape of AL algorithms, where the vast majority of reported lifts are neither statistically significant, nor prove to be generalizable.
+% This makes it very hard to identify the best AL algorithm, or even identifying state-of-the-art algorithms.
+% In this work we propose an evaluation protocol that was designed to handle the high variance in the performances of AL algorithms as well as being fully controllable regardless of the combination of dataset, model and AL algorithm.
+% We base our work largely on \cite{ji2023randomness}, following their guidelines for a reliable evaluation of AL algorithms, while extending the number of evaluated data domains from 1 to 5. \\
+% We focus on pool-based AL where a pool of unlabeled samples is fixed at the start of each experiment and one or more samples are chosen sequentially.
+% In addition to the default scenario of selecting a batch of samples every iteration we incorporate the single sample case into our benchmark.
+% Batched algorithms (and benchmarks) do not have a principled advantage over single-sample AL except for speed of computation.
+% The problem of optimizing a portfolio of unlabeled samples in each iteration is more complicated to solve and the algorithms have systematically less information per sample to work with leading to a generally worse performance, that impacts some algorithms more than others.
+% %In Fig. \ref{fig:batch_vs_single} we present a comparison of Badge \cite{ashdeep} and Coreset\cite{sener2017active}, in the batch and single-sample setting.
+% %Even though both algorithms were designed for batch sampling, both perform better in the single-sample setting.
+% We propose single-sample AL as an important tool to identify the best acquisition function, rather than the best combination of acquisition function and diversity regularization. \\
+% %\begin{figure}
+% %    \centering
+% %    \includegraphics[width=\linewidth]{img/ablation_batch_single}
+% %    \caption{Comparison of single-sample vs.
+batch selection for two acquisition functions that were specifically designed for batch selection.
+Even when designed for batch selection, acquisition functions tend to perform better in the single sample setting.}
+% %    \label{fig:batch_vs_single}
+% %\end{figure}
+% Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature. 
+% We offer our datasets in two versions - normal and encoded.
+% An encoded dataset was pre-encoded by an self-supervised encoder model, providing a different use case for active learning.
+% Our synthetic and text datasets (which use word embeddings in the normal setting) are exempt from this, bringing our total dataset count to 14 across 5 domains (Tabular, Image, Text, Synthetic, Encoded).
+% \begin{table}
+% 	\centering
+%         \caption{Comparison of our benchmark with the existing literature.
+Oracle curves serve as an approximation of the best possible AL algorithm.
+Including the encoded versions of our datasets we reach 14 datasets.
+"Semi" indicates whether the paper is employing any form of self- or semi-supervised learning.
+A "-" for repetitions means that we could not determine how often each experiment is repeated in the respective framework.}
+% 	\label{tab:benchmark_comparison}
+% 	\resizebox{\columnwidth}{!}{
+% \begin{tabular}{lccccllllcl}
+% Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
+% \multicolumn{1}{l|}{Beck et al. \cite{beck2021effective}} & batch & 4 & 7 & \checkmark & - & - & - & - & - & - \\
+% \multicolumn{1}{l|}{Hu et al. \cite{hu2021towards}} & batch & 5 & 13 & \checkmark & \checkmark & - & - & - & - & 3 \\
+% \multicolumn{1}{l|}{Zhou et al. \cite{zhou2021towards}} & batch & 3 & 2 & \checkmark & \checkmark & - & - & - & \checkmark & 5 \\
+% \multicolumn{1}{l|}{Zhan et al. \cite{zhan2021comparative}} & sngl+batch & 35 & 18 & - & - & \checkmark & \checkmark & - & \checkmark & 10-100 \\
+% \multicolumn{1}{l|}{Munjal et al. \cite{munjal2022towards}} & batch & 2 & 8 & \checkmark & - & - & - & - & - & 3 \\
+% \multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\
+
+% \multicolumn{1}{l|}{Rauch et al. \cite{rauch2023activeglae}} & batch & 11 & 5 & - & \checkmark & - & - & - & - & 5 \\
+% \hdashline
+% \multicolumn{1}{l|}{Ji et al. \cite{ji2023randomness}} & batch & 3 & 8 & \checkmark & - & - & - & - & - & - \\
+% \multicolumn{1}{l|}{Lueth et al. \cite{luth2024navigating}} & batch & 4 & 5 & \checkmark & - & - & - & \checkmark & - & 3 \\
+% \multicolumn{1}{l|}{\textbf{Ours}} & sngl+batch & 9(14) & 11 & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & 50 
+% \end{tabular}
+% }
+% \end{table}
+
+% \subsection*{Contributions}
+% %Maybe absorb this into the introduction text
+% \begin{enumerate}
+% 	\item[C1] Study on the reproducibility of AL experiments, providing evidence that previous works might have not repeated their experiments often enough to provide reliable results
+% 	\item[C2] Efficient and performant algorithm for an oracle that can be constructed greedily and does not rely on search
+% 	\item[C3] Two novel synthetic datasets that highlight principled shortcomings of of the top-performing AL algorithms
+%     \item[C4] Extending the evaluation of Active Learning algorithms to 5 data domains and revealing significant differences in algorithm performance between them
+% \end{enumerate}
+
 \section{Introduction}
-Deep neural networks (NN) have produced state-of-the-art results on many important supervised learning tasks.
-Since Deep NNs usually require large amounts of labeled training data, Active Learning (AL) can be used to instead select the most informative samples out of a large pool of unlabeled data, so that only these samples need to be labeled.
-It has been shown that a small labeled set of this nature can be used to train well-performing models \cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}. \\
-On top of providing a principled way to label unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
+Deep neural networks (NN) have produced state-of-the-art results on many
+important supervised learning tasks. Since Deep NNs usually require large
+amounts of labeled training data, Active Learning (AL) deals with
+selecting the most informative samples out of a large pool of unlabeled data, so
+that only these samples need to be labeled. It has been shown that a small
+labeled set of this nature can be used to train well-performing models
+%\cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}.
+% On top of providing a principled way to label unlabeled datasets, active
+% learning is one of the two major approaches besides semi-supervised learning to
+% make deep learning models more data efficient by requiring only a limited set of
+% manually labeled data.
 %Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
-In the last decade, many different algorithms for AL have been proposed.
-Even though, almost every method has reported lifts over all its predecessors,
-\footnote{Out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not claim a new SOTA performance in their result section.}
-AL research faces four central difficulties:
-{\let\thefootnote\relax\footnotetext{Code available at: anonymous}}
-%\let\thefootnote\relax\footnotetext{Code available at: \url{https://github.com/wernerth94/comparable-active-learning}}
-(i) The experiments are often carried out on different datasets and model architectures, hindering direct comparison,
-(ii) generalize poorly across different domains,
-(iii) the reported results can be subject to very high variance across restarts and
-(iv) are not always compared against important baselines like margin sampling \cite{wang2014new}.
-While multiple benchmark suites have been proposed to solve (i), to the best of our knowledge, we are the first to report results on all 3 data domains of tabular, image and text.
-Additionally, we provide synthetic datasets to highlight principled shortcoming of existing AL algorithms.
-Regarding (ii) and (iii), \cite{zhou2021towards} has pointed out severe inconsistencies in results of AL papers in recent years. 
-They conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm (\cite{zhou2021towards} Appendix A).
-These outlined issues lead to an inconclusive landscape of AL algorithms, where the vast majority of reported lifts are neither statistically significant, nor prove to be generalizable.
-This makes it very hard to identify the best AL algorithm, or even identifying state-of-the-art algorithms.
-In this work we propose an evaluation protocol that was designed to handle the high variance in the performances of AL algorithms as well as being fully controllable regardless of the combination of dataset, model and AL algorithm.
-We base our work largely on \cite{ji2023randomness}, following their guidelines for a reliable evaluation of AL algorithms, while extending the number of evaluated data domains from 1 to 5. \\
-We focus on pool-based AL where a pool of unlabeled samples is fixed at the start of each experiment and one or more samples are chosen sequentially.
-In addition to the default scenario of selecting a batch of samples every iteration we incorporate the single sample case into our benchmark.
-Batched algorithms (and benchmarks) do not have a principled advantage over single-sample AL except for speed of computation.
-The problem of optimizing a portfolio of unlabeled samples in each iteration is more complicated to solve and the algorithms have systematically less information per sample to work with leading to a generally worse performance, that impacts some algorithms more than others.
-%In Fig. \ref{fig:batch_vs_single} we present a comparison of Badge \cite{ashdeep} and Coreset\cite{sener2017active}, in the batch and single-sample setting.
-%Even though both algorithms were designed for batch sampling, both perform better in the single-sample setting.
-We propose single-sample AL as an important tool to identify the best acquisition function, rather than the best combination of acquisition function and diversity regularization. \\
+In the last decade, many different algorithms for AL have been proposed and
+almost every method has reported lifts over all its predecessors, \footnote{Out
+of all considered algorithms for this paper, only BALD \cite{gal2017deep} did
+not claim a new SOTA performance in their result section.} However, real
+insights into the current state of AL are hard to draw from these works, due to
+the following reasons:
+\begin{inparaenum}
+\item\label{prob1al} These works do not use a standardized evaluation setting with fixed
+datasets and baseline approaches.
+\item Due to cost reasons, a lot of works do only a small amount of experimental
+runs, hence it is questionable wether the superiority of a specific approach can
+be concluded from the conducted experiments.
+\item The works are only evaluated in a specific domain, such as computer vision
+or language processing. However, AL is a general principle of supervised
+learning, and thus methods should be evaluated in multiple domains to assess
+their capabilities.
+\end{inparaenum}
+% %{\let\thefootnote\relax\footnotetext{Code available at: anonymous}}
+% %\let\thefootnote\relax\footnotetext{Code available at: \url{https://github.com/wernerth94/comparable-active-learning}}
+% % (ii) Most of the research is focused on a specific application domain like
+% % computer vision, even though AL is a general principle for supervised learning
+% % (iii) the reported results can be subject to very high variances across restarts,
+% % still A lot of papers do only a hand full of runs, which makes questions the
+% % validity of the drawn conclusions.
+% % (iv) are not always compared against important baselines like margin
+% sampling \cite{wang2014new}.
+
+% These outlined issues lead to an
+% inconclusive landscape of AL algorithms, where the vast majority of reported
+% lifts are neither statistically significant, nor prove to be generalizable. This
+% makes it very hard to identify the best AL algorithm, or even identifying
+% state-of-the-art algorithms. 
+While multiple benchmark suites have been proposed to solve
+problem~\ref{prob1al}, to the best of
+our knowledge, all of them are either limited in the domains they consider or do
+not do enough runs for conclusive results 
+%\footnote{See
+% \Cref{tab:benchmark_comparison} for a comparison of existing benchmarks}. One
+% reason for that, may be that in AL research it is common to additionally include
+% a so-called \emph{oracle} baseline, i.e., a time-costly search-based approach
+% which gives a lower bound for the best possible solution. The need for an oracle
+% makes it challenging to use multiple datasets and enough runs for the
+% experimental evaluation. \todo{\@ Thorben: Ist das was ich hier zum Oracle
+% schreibe richtg? Oder ist das Quatsch?} 
+Hence, the current
+state of the art in Active Learning is still not well-understood and principle
+shortcomings of different algorithms and wether they are present
+domain-independent, are currently not identified.
+
+Here we step in with \emph{ALBench}, an active learning benchmark wich covers
+multiple application domains and which reports a large amount of runs per
+experiment, so that the significance of performance differences can can be
+estimated. To be more detailed, \emph{ALBench} consists of datasets from
+computer vision, natural language processing and the tabular domain. We provide
+our datasets both in raw format (i.e. consisting of text, images etc) or already
+``embedded'' where one can evaluate AL approaches on already learned representations.
+Furthermore, we provide further synthetic datasets to identify general
+challenges for AL methods.
+
+The evaluation protocol on which we evaluate in \emph{ALBench} uses 50 runs for
+each experiment. By having such a large amount of runs, we can evaluate the
+significance of performance gaps and identify the in-fact best performing
+approaches for different runs. Furthermore, we will show, that the small amounts
+of runs other works do can in fact lead to misleading results. To be more
+specific, we show that if only 5 runs are randomly sampled, the performance of
+specific methods strongly varies. As we will see, the ranking of
+the different method heavily fluctuates with the specific set of randomly
+selected runs. This effect goes so far, that for a well-established method such
+as margin-sampling, the question wether this method is significantly better or
+worse then the random baselines, will be answered differently depending on the 5
+runs which were chosen.
+
+To enable a large amount of runs, we propose a \emph{greedy oracle
+function}, which uses only a small amount of search steps to estimate the
+optimal solution. While being more time-efficient than established oracle
+functions, it can be a large underestimate of the real performance of the
+optimal solution. However, as our experiments will show, it is still far ahead
+of all current AL methods and thus is perfectly suitable for having a lower
+bound on what AL methods could be able to achieve.
+
+Our experimental evaluation shows, that there exists no clear SoTa
+method for Active Learning. The superiority of specific datasets is strongly
+domain-dependent with the outstanding observation, that the image domain works
+fundamentally different then the tabular and text domain. Here, the best
+performing approach in text and tabular data, namely \emph{margin sampling}, is
+significantly outperformed by \emph{Least Confident}, which does not belong to
+the top performing approaches in text and tabular. Thus, using the performance
+of Active Learning approaches on the image domain as a proxy of Active Learning
+in general, as often done, is questionable. To further understand this
+performance gaps, we propose \emph{Honeypot} and \emph{Diverging Sin}, two
+synthetical datasets, which are designed to be challenging for naive clustering
+or decision-boundary based approaches. Hence, they can be used as a lower bar
+for assessing a AL approach as sophisticated.
+
+To sum up, \emph{ALBench} consists of an experimental framework which includes a
+fast oracle approximation, multiple application domains, enough repetitions to
+draw valid conclusions and two synthetic tasks to estimate if a specific method
+reaches a minimum amount of complexity. By being the first benchmark to proving
+these things together, we believe that \emph{ALBench} is a major step forward of
+assessing the overall state of Active Learning research, independent of a
+specific application domain. \emph{ALBench} is publicly available.\todo{Link etc}
+Our contributions include the following:
+
+\begin{enumerate}
+\item We show that the small of runs that previous work has conducted is not
+sufficient for meaningful conclusions. This can go so far, that it may not even
+be possible to access wether margin sampling is better then random.
+\item We propose an efficient and performant oracle which is computed greedily.
+With that, we overcome one of the major obstacles for sufficient repetitions of
+AL experiments. 
+\todo{Sagt man ``is computed greedily''? Ist das wirklich NICHT search based? Weil eine greedy-search ist ja auch search based}
+\item We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
+designed to hinder active learning by simple decision-boundary or
+clustering-based approaches. Thus, they provide as a lower bar on the complexity
+of AL methods.
+\item We propose \emph{IMTSBench}, the first general benchmark providing tasks
+in the domains of images, text and tabular learning. It further contains
+synthetic and already encoded data to allow for a sophisticated evaluation of AL
+methods. Our experiments show, that there is no clear state-of-the art method
+for active learning over different domains.
+\end{enumerate}
+
+\todo{Ich bin mir nicht ganz sicher, wie man Honeypot und Diverging Sin verkauft. Macht as mit der Lower bar sinn?}
+% \begin{enumerate}
+% \item[C1] Study on the reproducibility of AL experiments, providing evidence
+% that previous works might have not repeated their experiments often enough to
+% provide reliable results
+% \item[C2] Efficient and performant algorithm for an oracle that can be
+% constructed greedily and does not rely on search
+% \item[C3] Two novel synthetic datasets that highlight principled shortcomings of
+% of the top-performing AL algorithms
+% \item[C4] Extending the evaluation of Active Learning algorithms to 5 data
+% domains and revealing significant differences in algorithm performance between
+% them to overcome one of the major obstacles for sufficient 
+% \end{enumerate}
+
+% we are the first to report results on all 3 data domains of
+% tabular, image and text. Additionally, we provide synthetic datasets to
+% highlight principled shortcoming of existing AL algorithms. Regarding (ii) and
+% (iii), \cite{zhou2021towards} has pointed out severe inconsistencies in results
+% of AL papers in recent years. They conducted a meta analysis of reported results
+% of several different AL algorithms and found that all considered algorithms only
+% provided significant lifts in their own original papers, while following
+% literature reported performances no better that uncertainty sampling, or in some
+% cases no better than random sampling for the same algorithm
+% (\cite{zhou2021towards} Appendix A). These outlined issues lead to an
+% inconclusive landscape of AL algorithms, where the vast majority of reported
+% lifts are neither statistically significant, nor prove to be generalizable. This
+% makes it very hard to identify the best AL algorithm, or even identifying
+% state-of-the-art algorithms. In this work we propose an evaluation protocol that
+% was designed to handle the high variance in the performances of AL algorithms as
+% well as being fully controllable regardless of the combination of dataset, model
+% and AL algorithm. We base our work largely on \cite{ji2023randomness}, following
+% their guidelines for a reliable evaluation of AL algorithms, while extending the
+% number of evaluated data domains from 1 to 5. \\
+% We focus on pool-based AL where a pool of unlabeled samples is fixed at the
+% start of each experiment and one or more samples are chosen sequentially. In
+% addition to the default scenario of selecting a batch of samples every iteration
+% we incorporate the single sample case into our benchmark. Batched algorithms
+% (and benchmarks) do not have a principled advantage over single-sample AL except
+% for speed of computation. The problem of optimizing a portfolio of unlabeled
+% samples in each iteration is more complicated to solve and the algorithms have
+% systematically less information per sample to work with leading to a generally
+% worse performance, that impacts some algorithms more than others. 
+% %In Fig. \ref{fig:batch_vs_single} we present a comparison of Badge \cite{ashdeep} and Coreset\cite{sener2017active}, in the batch and single-sample setting.
+% %Even though both algorithms were designed for batch sampling, both perform better in the single-sample setting.
+% We propose single-sample AL as an important tool to identify the best
+% acquisition function, rather than the best combination of acquisition function
+% and diversity regularization. \\
 %\begin{figure}
@@
 %\end{figure}
-Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature. 
-We offer our datasets in two versions - normal and encoded.
-An encoded dataset was pre-encoded by an self-supervised encoder model, providing a different use case for active learning.
-Our synthetic and text datasets (which use word embeddings in the normal setting) are exempt from this, bringing our total dataset count to 14 across 5 domains (Tabular, Image, Text, Synthetic, Encoded).
+% Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature. 
+% We offer our datasets in two versions - normal and encoded.
+% An encoded dataset was pre-encoded by an self-supervised encoder model, providing a different use case for active learning.
+% Our synthetic and text datasets (which use word embeddings in the normal
+% setting) are exempt from this, bringing our total dataset count to 14 across 5
+% domains (Tabular, Image, Text, Synthetic, Encoded). 
 \begin{table}
 	\centering
-        \caption{Comparison of our benchmark with the existing literature.
-Oracle curves serve as an approximation of the best possible AL algorithm.
-Including the encoded versions of our datasets we reach 14 datasets.
-"Semi" indicates whether the paper is employing any form of self- or semi-supervised learning.
-A "-" for repetitions means that we could not determine how often each experiment is repeated in the respective framework.}
+\caption{Comparison of our benchmark with the existing literature. Oracle curves
+serve as an approximation of the best possible AL algorithm. Including the
+encoded versions of our datasets we reach 14 datasets, without we have 9. "Semi"
+indicates whether the paper is employing any form of self- or semi-supervised
+learning. A "-" for repetitions means that we could not determine how often each
+experiment is repeated in the respective framework. \emph{ALBench} is the only
+benchmark which reports over enough runs for meaningful results and considers
+all 5 domains.}
 	\label{tab:benchmark_comparison}
 	\resizebox{\columnwidth}{!}{
-\begin{tabular}{lccccllllcl}
+\begin{tabular}{lccc|cllll|cl}
 Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
@@
 \end{table}
-
-\subsection*{Contributions}
-%Maybe absorb this into the introduction text
-\begin{enumerate}
-	\item[C1] Study on the reproducibility of AL experiments, providing evidence that previous works might have not repeated their experiments often enough to provide reliable results
-	\item[C2] Efficient and performant algorithm for an oracle that can be constructed greedily and does not rely on search
-	\item[C3] Two novel synthetic datasets named "Honeypot" and "Diverging Sine" that highlight two principled shortcomings AL algorithms.
-Firstly, a susceptibility to noisy or adverse samples and secondly, the oversampling of easy to distinguish regions of the dataset
-    \item[C4] Extending the evaluation of Active Learning algorithms to 5 data domains and revealing significant differences in algorithm performance between them
-\end{enumerate}
 

### 2024-06-03 doc/al_benchmark_paper.tex
Some english stuff

@@
 research suffers from the fact that lifts from literature generalize poorly and
-are only conducted on a small number of repetitions of experiments. To overcome
+that only a small number of repetitions of experiments are conducted. To overcome
 this obstacles, we propose \emph{ALBench}, the first active learning benchmark

### 2024-06-03 doc/al_benchmark_paper mst.tex
Max Proposal Structure

--- 2024-06-03 doc/al_benchmark_paper.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methodology}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Why we need 50 restarts}\label{sec:restarts}
+\section{A Small Amount of Repetitions is not Sufficient for Meaningful Results}
+\label{sec:restarts}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{Why we need 50 restarts}\label{sec:restarts}
 To evaluate how many restarts are necessary to obtain conclusive results in an AL experiment, we computed 100 runs of our top-performing algorithm on one dataset.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}\label{sec:datasets}
+\section{\emph{ALBench:} A Cross-Domain Active Learning Benchmark}\label{sec:datasets}
 A detailed description of the preprocessing of each dataset can be found in Appendix \ref{app:hyperparameters}. \\ [1mm]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Greedy Oracle Algorithm}\label{sec:oracle}
-%\begin{wrapfigure}{R}{0.36\textwidth}
-%    \vspace{-0.83cm}
-%    \begin{minipage}{0.36\textwidth}
-%       Space for the Oracle Algorithm
-%    \vspace{-0.83cm}
-%\end{minipage}
-%\end{wrapfigure}
-
-%an oracle set was so far always described as the optimal set and therefore
-Posing Active Learning as a combinatorial problem, the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure is the set that induces the highest AUC score for a given budget.
-However, since this problem is computationally infeasible for realistic datasets, previous works have proposed approximations to this oracle sequence. 
-\cite{zhou2021towards} used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
-Even though their reported performance curves display a significant lift over all other acquisition functions, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
-In this paper, we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
-Our oracle algorithm evaluates every data point $u_k = \operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$ in a subsample of unlabeled points by fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test performance.
-The data point with the best test performance is selected and added to the labeled pool for that iteration.
-We noticed that this oracle is over-specializing on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
-This can happen, for example, if the oracle picked a labeled set that enables the classifier to correctly classify a big portion of easy samples in the test set, but now fails to find the next \textbf{single} unlabeled point that would enable the classifier to succeed on one of the hard samples.
-This leads to a situation, where no point can immediately incur an increase in test performance and therefore the selected data point can be considered random.
-To circumvent this problem, we use margin sampling \cite{wang2014new} as a fallback option for the oracle.
-Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
-The resulting greedy algorithm constructs an approximation of the optimal labeled set that consistently outperforms all other algorithms by a significant margin, while requiring relatively low computational cost ($\mathcal{O}(B\tau)$).
-We fix $\tau = 20$ in this work, as this gave us already a significant lift and we expect diminishing returns for larger $\tau$.
-The pseudocode for our oracle can be found in App. \ref{app:pseudocode}.
-%In the algorithm $\operatorname{Train}(\LL, \hat y_\theta)$ trains the classification model $\hat y_\theta$ on $\LL$. \\
-%Alg. \ref{alg:oracle} replaces the acquisition function $\Omega$ in the AL loop (Appendix \ref{app:pseudocode} Alg. \ref{alg:active_learning}).
-Even though our proposed algorithm is more efficient than other approaches, the computational costs for high budget datasets like Cifa10 and FashionMnist meant that we could not compute the oracle for all 10000 datapoints.
-To still provide an oracle for these two datasets, we select two points per iteration instead of one and stop the oracle computation at a budget of 5000.
-The rest of the curve is forecast with a simple linear regression that asymptotically approaches the upper bound performance of the dataset. 
-A detailed description can be found in App. \ref{app:oracle_forecasting}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Evaluation Protocol}\label{sec:evaluation}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{A Greedy Oracle Algorithm}\label{sec:oracle}
+%\begin{wrapfigure}{R}{0.36\textwidth}
+%    \vspace{-0.83cm}
+%    \begin{minipage}{0.36\textwidth}
+%       Space for the Oracle Algorithm
+%    \vspace{-0.83cm}
+%\end{minipage}
+%\end{wrapfigure}
+
+%an oracle set was so far always described as the optimal set and therefore
+Posing Active Learning as a combinatorial problem, the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure is the set that induces the highest AUC score for a given budget.
+However, since this problem is computationally infeasible for realistic datasets, previous works have proposed approximations to this oracle sequence. 
+\cite{zhou2021towards} used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
+Even though their reported performance curves display a significant lift over all other acquisition functions, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
+In this paper, we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
+Our oracle algorithm evaluates every data point $u_k = \operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$ in a subsample of unlabeled points by fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test performance.
+The data point with the best test performance is selected and added to the labeled pool for that iteration.
+We noticed that this oracle is over-specializing on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
+This can happen, for example, if the oracle picked a labeled set that enables the classifier to correctly classify a big portion of easy samples in the test set, but now fails to find the next \textbf{single} unlabeled point that would enable the classifier to succeed on one of the hard samples.
+This leads to a situation, where no point can immediately incur an increase in test performance and therefore the selected data point can be considered random.
+To circumvent this problem, we use margin sampling \cite{wang2014new} as a fallback option for the oracle.
+Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
+The resulting greedy algorithm constructs an approximation of the optimal labeled set that consistently outperforms all other algorithms by a significant margin, while requiring relatively low computational cost ($\mathcal{O}(B\tau)$).
+We fix $\tau = 20$ in this work, as this gave us already a significant lift and we expect diminishing returns for larger $\tau$.
+The pseudocode for our oracle can be found in App. \ref{app:pseudocode}.
+%In the algorithm $\operatorname{Train}(\LL, \hat y_\theta)$ trains the classification model $\hat y_\theta$ on $\LL$. \\
+%Alg. \ref{alg:oracle} replaces the acquisition function $\Omega$ in the AL loop (Appendix \ref{app:pseudocode} Alg. \ref{alg:active_learning}).
+Even though our proposed algorithm is more efficient than other approaches, the computational costs for high budget datasets like Cifa10 and FashionMnist meant that we could not compute the oracle for all 10000 datapoints.
+To still provide an oracle for these two datasets, we select two points per iteration instead of one and stop the oracle computation at a budget of 5000.
+The rest of the curve is forecast with a simple linear regression that asymptotically approaches the upper bound performance of the dataset. 
+A detailed description can be found in App. \ref{app:oracle_forecasting}.
+\todo{Extend this section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Experiments}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Synthetic Datasets for AL}
+\section{Assessing the Complexity of AL Methods with \emph{Honeypot} and
+\emph{Diverding Sin}}
 AL approaches can be categorized into two types, uncertainty and geometric approaches.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Results on Synthetic Data}
+\subsection{Results}
 Our results on Honeypot reveal principled shortcomings for the two best algorithms in BADGE and margin sampling.

### 2024-06-03 doc/al_benchmark_paper.tex
paper

@@
 Hence, the current
-state of the art in Active Learning is still not well-understood and principle
-shortcomings of different algorithms and wether they are present
+state of the art in Active Learning is still not well-understood and principled
+shortcomings of different algorithms and wether they are 
 domain-independent, are currently not identified.
 
-Here we step in with \emph{ALBench}, an active learning benchmark wich covers
-multiple application domains and which reports a large amount of runs per
+Here we step in with \emph{ALBench}, an active learning benchmark which covers
+multiple application domains and reports a large amount of runs per
 experiment, so that the significance of performance differences can can be
-estimated. To be more detailed, \emph{ALBench} consists of datasets from
+estimated. Specifically, \emph{ALBench} consists of datasets from
 computer vision, natural language processing and the tabular domain. We provide
-our datasets both in raw format (i.e. consisting of text, images etc) or already
-``embedded'' where one can evaluate AL approaches on already learned representations.
-Furthermore, we provide further synthetic datasets to identify general
+our datasets both in raw format %(i.e. consisting of text, images etc) 
+as well as ``embedded'' by a fixed embedding model, enabling evaluation of AL methods in
+this semi-supervised setting.
+Furthermore, we propose two novel synthetic datasets to highlight general
 challenges for AL methods.
 
-The evaluation protocol on which we evaluate in \emph{ALBench} uses 50 runs for
+The applied evaluation protocol in \emph{ALBench} uses 50 runs for
 each experiment. By having such a large amount of runs, we can evaluate the
-significance of performance gaps and identify the in-fact best performing
-approaches for different runs. Furthermore, we will show, that the small amounts
-of runs other works do can in fact lead to misleading results. To be more
-specific, we show that if only 5 runs are randomly sampled, the performance of
-specific methods strongly varies. As we will see, the ranking of
-the different method heavily fluctuates with the specific set of randomly
-selected runs. This effect goes so far, that for a well-established method such
-as margin-sampling, the question wether this method is significantly better or
-worse then the random baselines, will be answered differently depending on the 5
-runs which were chosen.
-
-To enable a large amount of runs, we propose a \emph{greedy oracle
-function}, which uses only a small amount of search steps to estimate the
+significance of performance gaps and identify the best performing
+approaches for each dataset as well as whole domains. 
+Furthermore, we show that the small amount of runs other works do in fact 
+produce to misleading results. 
+To be more specific, we show that if only 3 restarts are employed for each experiment, 
+the performance of specific methods strongly varies. As we will see, even the ranking of
+the different methods averaged over many datasets fluctuates with the specific set of runs. 
+This effect is so strong, that even for a well-established method such
+as margin-sampling, it's performance can be significantly better and significantly
+worse than random for the same dataset, depending on the seed.
+
+To enable the computation of an oracle performance for a protocol with large amounts of restarts
+we propose a \emph{greedy oracle function} 
+which uses only a small amount of search steps to estimate the
 optimal solution. While being more time-efficient than established oracle
-functions, it can be a large underestimate of the real performance of the
-optimal solution. However, as our experiments will show, it is still far ahead
-of all current AL methods and thus is perfectly suitable for having a lower
-bound on what AL methods could be able to achieve.
-
-Our experimental evaluation shows, that there exists no clear SoTa
-method for Active Learning. The superiority of specific datasets is strongly
+functions, it possibly underestimates the real upper bound performance. 
+However, as our experiments will show, it is still outperforming all current AL methods
+by at least 5\% and thus is suitable as an upper bound.
+
+Our experimental evaluation shows, that there exists no clear SOTA
+method for Active Learning. The superiority of methods is strongly dataset- and
 domain-dependent with the outstanding observation, that the image domain works
-fundamentally different then the tabular and text domain. Here, the best
-performing approach in text and tabular data, namely \emph{margin sampling}, is
-significantly outperformed by \emph{Least Confident}, which does not belong to
-the top performing approaches in text and tabular. Thus, using the performance
+fundamentally different than the tabular and text domain. Here, the best
+performing approach for text and tabular data, namely \emph{margin sampling}, is
+significantly outperformed by \emph{least confident sampling}, which does not belong to
+the top performing approaches in any other domain. Thus, using the performance
 of Active Learning approaches on the image domain as a proxy of Active Learning
-in general, as often done, is questionable. To further understand this
-performance gaps, we propose \emph{Honeypot} and \emph{Diverging Sin}, two
-synthetical datasets, which are designed to be challenging for naive clustering
-or decision-boundary based approaches. Hence, they can be used as a lower bar
-for assessing a AL approach as sophisticated.
-
-To sum up, \emph{ALBench} consists of an experimental framework which includes a
-fast oracle approximation, multiple application domains, enough repetitions to
-draw valid conclusions and two synthetic tasks to estimate if a specific method
-reaches a minimum amount of complexity. By being the first benchmark to proving
-these things together, we believe that \emph{ALBench} is a major step forward of
-assessing the overall state of Active Learning research, independent of a
-specific application domain. \emph{ALBench} is publicly available.\todo{Link etc}
+in general, as it is often done \cite{beck2021effective, munjal2022towards, li2022empirical, ji2023randomness, luth2024navigating}, is questionable. To further analyze
+performance of common methods, we propose \emph{Honeypot} and \emph{Diverging Sin}, two
+synthetic datasets, designed to be challenging for naive decision-boundary- and clustering-based approaches respectively. Hence, they provide insights in principled shortcomings of AL methods.
+
+To sum up, \emph{ALBench} is an experimental framework which includes an
+efficient oracle approximation, multiple application domains, enough repetitions to
+draw valid conclusions and two synthetic tasks to highlight shortcomings of AL methods. 
+By being the first benchmark to proving these things in one code-base, 
+we believe that \emph{ALBench} is a major step forward of assessing the overall state of Active Learning research, independent of
+specific application domains. \emph{ALBench} is publicly available under \emph{anonymous}. \\ [1mm]
 Our contributions include the following:
@@
 \begin{enumerate}
-\item We show that the small of runs that previous work has conducted is not
-sufficient for meaningful conclusions. This can go so far, that it may not even
-be possible to access wether margin sampling is better then random.
-\item We propose an efficient and performant oracle which is computed greedily.
-With that, we overcome one of the major obstacles for sufficient repetitions of
-AL experiments. 
-\todo{Sagt man ``is computed greedily''? Ist das wirklich NICHT search based? Weil eine greedy-search ist ja auch search based}
-\item We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
-designed to hinder active learning by simple decision-boundary or
-clustering-based approaches. Thus, they provide as a lower bar on the complexity
-of AL methods.
+\item We show that the small of repetitions that previous works have employed is not
+sufficient for meaningful conclusions. Sometimes even making it impossible to assess if a performance is above or below random.
+\item We propose an efficient and performant oracle which is constructed iteratively in a greedy fashion,
+overcoming major computational hurdles.
+% one of the major obstacles for sufficient repetitions of AL experiments. 
+% [Thorben] Das können wir so nicht sagen, glaube ich. Niemand verlangt von uns, dass wir das Oracle auch 50 mal wiederholen
 \item We propose \emph{IMTSBench}, the first general benchmark providing tasks
 in the domains of images, text and tabular learning. It further contains
-synthetic and already encoded data to allow for a sophisticated evaluation of AL
+synthetic and pre-encoded data to allow for a sophisticated evaluation of AL
 methods. Our experiments show, that there is no clear state-of-the art method
-for active learning over different domains.
+for active learning across different domains.
+\item We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
+designed to hinder active learning by naive decision-boundary- or
+clustering-based approaches respectively. Thus, they provide an important tool to identify 
+shortcomings of existing AL methods.
 \end{enumerate}
 
-\todo{Ich bin mir nicht ganz sicher, wie man Honeypot und Diverging Sin verkauft. Macht as mit der Lower bar sinn?}
+%\todo{Ich bin mir nicht ganz sicher, wie man Honeypot und Diverging Sin verkauft. Macht as mit der Lower bar sinn?}
 % \begin{enumerate}
@@
 The authors of \cite{beck2021effective}, \cite{munjal2022towards}, \cite{li2022empirical}, \cite{ji2023randomness} and \cite{luth2024navigating} even focus exclusively on the image domain.
-Experiments on the interplay between AL and semi-supervised learning have only been provided by two works so far \cite{li2022empirical, luth2024navigating}, both of them only for images.
-An oracle algorithm has so far been proposed by only two works \cite{zhou2021towards, zhan2022comparative}. 
-Both of these algorithms rely on search, while our proposed method can be constructed sequentially.
+Especially the tabular domain is underrepresented in preceding benchmarks, as only \cite{zhan2022comparative} provides experiments for it.
+Experiments on the interplay between AL and semi-supervised learning are similarly under-researched, as only two works exist \cite{li2022empirical, luth2024navigating}, both of them only using images.
+An oracle algorithm has been proposed by two works \cite{zhou2021towards, zhan2022comparative}. 
+Both of these algorithms rely on search and are computationally very expensive, while our proposed method efficiently can be constructed sequentially.
 %While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training, and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
@@
 For a complete list of the pitfalls and our implementation of them, please refer to App. \ref{app:pitfalls}.
-To the best of our knowledge, we are the first to extend reliable SOTA (based on \cite{ji2023randomness, luth2024navigating}) experimentation to a total of 5 data domains.
+To the best of our knowledge, we are the first to extend reliable SOTA (based on \cite{ji2023randomness, luth2024navigating}) experimentation to a total of 5 data domains and a high number of restarts per experiment.
 
@@
 \end{figure}
-In addition to this motivating example, we carried out our main evaluation (Tab. \ref{tab:results}) multiple times by uniformly sampling 3 random from our 50 available runs and comparing the results.
+In addition to this motivating example, we carried out our main evaluation (Tab. \ref{tab:results}) multiple times by uniformly sampling 3 random from our available runs and comparing the results.
 We found significant differences in the performance of acquisition functions on individual datasets, as well as permutations in the final ranking.
 This partly explains the ongoing difficulties in reproducing results for AL experiments and benchmarks.
-This details can be found in App. \ref{app:rank_difference}.
+The details can be found in App. \ref{app:rank_difference}.
 For this benchmark we employ 50 restarts of every experiment.
@@
 \subsection{Seeding vs. Restarts}\label{sec:reproducibility}
-Considering the high computational cost of 50 repetitions, another approach to ensure reproducibility would be to reduce the amount of variance in the experiment by keeping as many subsystems (weight initialization, data splits, etc.) as possible fixed with specialized seeding.
+Considering the high computational cost of 50 repetitions, another approach to ensure consistency between experiments would be to reduce the amount of variance in the experiment by keeping as many subsystems (weight initialization, data splits, etc.) as possible fixed with specialized seeding.
 \\
@@
 Unless stated otherwise, we will keep $s_\Omega$ fixed, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
-While this seeding strategy is capable of controlling the amount variance in the experiment, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), reducing its value in practice and providing a bad approximation of the quality of an AL algorithm.
+While this seeding strategy is capable of controlling the amount variance in the experiment, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), providing a bad approximation of the quality of an AL algorithm (i.e.
+measured performances for an algorithm might not even transfer to a different model initialization).
 Hence, we opt for letting the subsystems vary (by increasing $s_\mathcal{D}$ and $s_\theta$) and combine that with a high number of restarts to obtain a good average of the generalization performance of each AL algorithm.
@@
 %
-We would like to point out that these datasets are selected for speed of computation (both in terms of number of features and necessary budget to solve the dataset). 
+We would like to point out that these datasets are selected for speed of computation (both in terms of the required classifier and the necessary budget to solve the dataset). 
 However, similar to our argumentation for picking smaller classifiers, we are solely focused on comparing different AL algorithms in this paper and do not aim to develop novel classification models on these datasets.
@@
 \\ [1mm]
-Adapting the experimental setting from \cite{hacohen2022active}, we offer all our datasets in the un-encoded (normal) setting as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning.
+Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning.
 The text datasets are an exception to this, as they are only offered in their encoded form.
-The pre-encoded datasets enable us to test single-sample algorithms on more complex datasets like Cifar10 and FashionMnist.
+Pre-encoded datasets enable us to test small query sizes on more complex datasets like Cifar10 and FashionMnist.
 They also serve the purpose of investigating the interplay between self-supervised learning techniques and AL, as well as alleviating the cold-start problem described in \cite{luth2024navigating} as they require a way smaller seed set.
@@
 To ensure that enough information from the data is encoded by our embedding model, the quality of embeddings during pretext training was measured after each epoch.
-We attached a linear classification head to the encoder, fine-tuned it to the data and evaluated this classifier for test accuracy, mirroring our AL setup for embedded datasets. 
+To this end, we attached a linear classification head to the encoder, fine-tuned it to the data and evaluated this classifier for test accuracy. 
 The checkpoint of each encoder model will be provided together with the framework. \\ [1mm]
@@
 (i) Creating test and validation splits from the full dataset rather than only the labeled seed set.
-Fully fledged test and validation splits are unobtainable in practice, but they provide not only a better approximation of algorithm performance, but also a better foundation for hyperparameter tuning, which is bound to reduce variance in the experiment.
+Fully fledged test and validation splits are unobtainable in practice, but they not only provide a better approximation of the algorithms generalization performance, but also a better foundation for hyperparameter tuning, which is bound to reduce variance in the experiment.
 \\
@@
 \subsection{Greedy Oracle Algorithm}\label{sec:oracle}
-%\begin{wrapfigure}{R}{0.36\textwidth}
-%    \vspace{-0.83cm}
-%    \begin{minipage}{0.36\textwidth}
-%       Space for the Oracle Algorithm
-%    \vspace{-0.83cm}
-%\end{minipage}
-%\end{wrapfigure}
-
 %an oracle set was so far always described as the optimal set and therefore
-Posing Active Learning as a combinatorial problem, the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure is the set that induces the highest AUC score for a given budget.
-However, since this problem is computationally infeasible for realistic datasets, previous works have proposed approximations to this oracle sequence. 
+A oracle method for AL tries to find the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure that induces the highest AUC score for a given budget.
+However, since this combinatorial problem is computationally infeasible for realistic datasets, previous works have proposed approximations to this oracle sequence. 
 \cite{zhou2021towards} used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
@@
 In this paper, we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
-Our oracle algorithm evaluates every data point $u_k = \operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$ in a subsample of unlabeled points by fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test performance.
-The data point with the best test performance is selected and added to the labeled pool for that iteration.
-We noticed that this oracle is over-specializing on the test set, resulting in stagnating or even decreasing performance curves in later AL iterations.
+Our oracle algorithm evaluates every data point $u_k = \operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$ in a subsample of unlabeled points by recovering the label, fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test performance.
+The data point that incurs the largest lift in test performance is selected and added to the labeled pool for that iteration.
+\todo{[Thorben] Ich habe die Erklärung für den Fallback im Oracle neu gemacht. Ist das verständlich?}
+Due to the algorithms greedy nature (just considering the next point to pick), our oracle frequently encounters situations where every point in $u_k$ would incur a negative lift (worsening the test performance).
 This can happen, for example, if the oracle picked a labeled set that enables the classifier to correctly classify a big portion of easy samples in the test set, but now fails to find the next \textbf{single} unlabeled point that would enable the classifier to succeed on one of the hard samples.
 This leads to a situation, where no point can immediately incur an increase in test performance and therefore the selected data point can be considered random.
-To circumvent this problem, we use margin sampling \cite{wang2014new} as a fallback option for the oracle.
+To circumvent this problem, we use our best-performing acquisition function (margin sampling \cite{wang2014new}) as a fallback option for the oracle.
 Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
 The resulting greedy algorithm constructs an approximation of the optimal labeled set that consistently outperforms all other algorithms by a significant margin, while requiring relatively low computational cost ($\mathcal{O}(B\tau)$).
-We fix $\tau = 20$ in this work, as this gave us already a significant lift and we expect diminishing returns for larger $\tau$.
+\todo{Da wir den Fallback haben, sollten wir das Oracle vielleicht einfach als Erweiterung von MarginSampling verkaufen?}
+We fix $\tau = 20$ in this work, as this gave us an average lift of 5\% over the best performing acquisition function per dataset (which is significant for AL settings) and we expect diminishing returns for larger $\tau$.
 The pseudocode for our oracle can be found in App. \ref{app:pseudocode}.
@@
 %Alg. \ref{alg:oracle} replaces the acquisition function $\Omega$ in the AL loop (Appendix \ref{app:pseudocode} Alg. \ref{alg:active_learning}).
-Even though our proposed algorithm is more efficient than other approaches, the computational costs for high budget datasets like Cifa10 and FashionMnist meant that we could not compute the oracle for all 10000 datapoints.
-To still provide an oracle for these two datasets, we select two points per iteration instead of one and stop the oracle computation at a budget of 5000.
+Even though our proposed algorithm is more efficient than other approaches, the computational costs for high budget datasets like Cifar10 and FashionMnist meant that we could not compute the oracle for all 10000 datapoints.
+To still provide an oracle for these two datasets, we select two points per iteration instead of one and stop the oracle computation at a budget of 2000.
 The rest of the curve is forecast with a simple linear regression that asymptotically approaches the upper bound performance of the dataset. 
@@
 Table \ref{tab:rank_diff_1} and Table \ref{tab:rank_diff_2} follow the exact same computation of ranks that created the main result (Table \ref{tab:results}) with the only difference being a reduced number of runs per acquisition function.
-For each table we uniformly sampled 3 runs from the available 50 per acquisition function. \\
+For each table we sampled 3 runs uniformly at random from the available 50 per acquisition function. \\
 We can observe significant differences between the two tables: \\

### 2024-06-03 doc/al_benchmark_paper.tex
paper

@@
 datasets and baseline approaches.
-\item Due to cost reasons, a lot of works do only a small amount of experimental
+\item Due to computational constraints, a lot of works do only a small amount of experimental
 runs, hence it is questionable wether the superiority of a specific approach can
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methodology}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Why we need 50 restarts}\label{sec:restarts}
+\section{Few Repetitions are not Sufficient for Meaningful Results}\label{sec:restarts}
 To evaluate how many restarts are necessary to obtain conclusive results in an AL experiment, we computed 100 runs of our top-performing algorithm on one dataset.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}\label{sec:datasets}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{ALBench: A Cross-Domain Active Learning Benchmark}\label{sec:datasets}
 A detailed description of the preprocessing of each dataset can be found in Appendix \ref{app:hyperparameters}. \\ [1mm]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Greedy Oracle Algorithm}\label{sec:oracle}
-%an oracle set was so far always described as the optimal set and therefore
-A oracle method for AL tries to find the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure that induces the highest AUC score for a given budget.
-However, since this combinatorial problem is computationally infeasible for realistic datasets, previous works have proposed approximations to this oracle sequence. 
-\cite{zhou2021towards} used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
-Even though their reported performance curves display a significant lift over all other acquisition functions, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
-In this paper, we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
-Our oracle algorithm evaluates every data point $u_k = \operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$ in a subsample of unlabeled points by recovering the label, fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test performance.
-The data point that incurs the largest lift in test performance is selected and added to the labeled pool for that iteration.
-\todo{[Thorben] Ich habe die Erklärung für den Fallback im Oracle neu gemacht. Ist das verständlich?}
-Due to the algorithms greedy nature (just considering the next point to pick), our oracle frequently encounters situations where every point in $u_k$ would incur a negative lift (worsening the test performance).
-This can happen, for example, if the oracle picked a labeled set that enables the classifier to correctly classify a big portion of easy samples in the test set, but now fails to find the next \textbf{single} unlabeled point that would enable the classifier to succeed on one of the hard samples.
-This leads to a situation, where no point can immediately incur an increase in test performance and therefore the selected data point can be considered random.
-To circumvent this problem, we use our best-performing acquisition function (margin sampling \cite{wang2014new}) as a fallback option for the oracle.
-Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
-The resulting greedy algorithm constructs an approximation of the optimal labeled set that consistently outperforms all other algorithms by a significant margin, while requiring relatively low computational cost ($\mathcal{O}(B\tau)$).
-\todo{Da wir den Fallback haben, sollten wir das Oracle vielleicht einfach als Erweiterung von MarginSampling verkaufen?}
-We fix $\tau = 20$ in this work, as this gave us an average lift of 5\% over the best performing acquisition function per dataset (which is significant for AL settings) and we expect diminishing returns for larger $\tau$.
-The pseudocode for our oracle can be found in App. \ref{app:pseudocode}.
-%In the algorithm $\operatorname{Train}(\LL, \hat y_\theta)$ trains the classification model $\hat y_\theta$ on $\LL$. \\
-%Alg. \ref{alg:oracle} replaces the acquisition function $\Omega$ in the AL loop (Appendix \ref{app:pseudocode} Alg. \ref{alg:active_learning}).
-Even though our proposed algorithm is more efficient than other approaches, the computational costs for high budget datasets like Cifar10 and FashionMnist meant that we could not compute the oracle for all 10000 datapoints.
-To still provide an oracle for these two datasets, we select two points per iteration instead of one and stop the oracle computation at a budget of 2000.
-The rest of the curve is forecast with a simple linear regression that asymptotically approaches the upper bound performance of the dataset. 
-A detailed description can be found in App. \ref{app:oracle_forecasting}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Evaluation Protocol}\label{sec:evaluation}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{A Greedy Oracle Algorithm}\label{sec:oracle}
+%an oracle set was so far always described as the optimal set and therefore
+A oracle method for AL tries to find the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure that induces the highest AUC score for a given budget.
+However, since this combinatorial problem is computationally infeasible for realistic datasets, previous works have proposed approximations to this oracle sequence. 
+\cite{zhou2021towards} used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
+Even though their reported performance curves display a significant lift over all other acquisition functions, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
+In this paper, we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
+Our oracle algorithm evaluates every data point $u_k = \operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$ in a subsample of unlabeled points by recovering the label, fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test performance.
+The data point that incurs the largest lift in test performance is selected and added to the labeled pool for that iteration.
+\todo{[Thorben] Ich habe die Erklärung für den Fallback im Oracle neu gemacht. Ist das verständlich?}
+Due to the algorithms greedy nature (just considering the next point to pick), our oracle frequently encounters situations where every point in $u_k$ would incur a negative lift (worsening the test performance).
+This can happen, for example, if the oracle picked a labeled set that enables the classifier to correctly classify a big portion of easy samples in the test set, but now fails to find the next \textbf{single} unlabeled point that would enable the classifier to succeed on one of the hard samples.
+This leads to a situation, where no point can immediately incur an increase in test performance and therefore the selected data point can be considered random.
+To circumvent this problem, we use our best-performing acquisition function (margin sampling \cite{wang2014new}) as a fallback option for the oracle.
+Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
+The resulting greedy algorithm constructs an approximation of the optimal labeled set that consistently outperforms all other algorithms by a significant margin, while requiring relatively low computational cost ($\mathcal{O}(B\tau)$).
+\todo{Da wir den Fallback haben, sollten wir das Oracle vielleicht einfach als Erweiterung von MarginSampling verkaufen?}
+We fix $\tau = 20$ in this work, as this gave us an average lift of 5\% over the best performing acquisition function per dataset (which is significant for AL settings) and we expect diminishing returns for larger $\tau$.
+The pseudocode for our oracle can be found in App. \ref{app:pseudocode}.
+%In the algorithm $\operatorname{Train}(\LL, \hat y_\theta)$ trains the classification model $\hat y_\theta$ on $\LL$. \\
+%Alg. \ref{alg:oracle} replaces the acquisition function $\Omega$ in the AL loop (Appendix \ref{app:pseudocode} Alg. \ref{alg:active_learning}).
+Even though our proposed algorithm is more efficient than other approaches, the computational costs for high budget datasets like Cifar10 and FashionMnist meant that we could not compute the oracle for all 10000 datapoints.
+To still provide an oracle for these two datasets, we select two points per iteration instead of one and stop the oracle computation at a budget of 2000.
+The rest of the curve is forecast with a simple linear regression that asymptotically approaches the upper bound performance of the dataset. 
+A detailed description can be found in App. \ref{app:oracle_forecasting}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Experiments}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Synthetic Datasets for AL}
+\section{Honeypot and Diverging Sine}
 AL approaches can be categorized into two types, uncertainty and geometric approaches.

### 2024-06-04 doc/al_benchmark_paper.tex
paper

@@
 % TODOs nach Priorität
-% !! "Overfitting on test" für Oracle: Sanity-Check durchführen und eventuell komplett streichen
-% !! Introduction neu schreiben: Wir brauchen eine klare Benennung der Ziele und bessere Abgrenzung zu anderen papern 
-% !! Sec 4 Methodology ist sehr chaotisch. Wir müssen neu sortieren und unwichtige Dinge streichen
-% Contributions neu sortieren und ausführlicher beschreiben
+% Terminologie aufräumen:
+% - Query size vs batch size
+% - Acq. function vs. AL Alg. vs. AL Method
+% - (Un-)encoded vs. semi-supervised/normal
+% Gibt es eine Conclusion für die Encoded-Datasets? Im Moment werden die nicht besprochen
 % Methode zum CD-Diagramm genau erklären (paired-t-test etc.) und die Relevanz fürs paper zeigen
@@
 % (Un-)Encoded durch Normal uns Semi-Supervised ersetzen, um die Verbindung klar zu machen
-% Gibt es eine Conclusion für die Encoded-Datasets? Im Moment werden die nicht besprochen
 % Eventuell Batchsize 1 von den anderen Ergebnissen separieren. Der Vergleich macht mehr Sinn, als einfach die Batchsizes zu mitteln
 % Tabelle 1 für Algorithmen statt Domänen. Das zeigt, dass immer entweder BADGE oder Margin (unsere Top 2) fehlt
-% Oracle Performance mit einer Zahl benennen: z.B. Da wird ~20% lift zu allen Baselines haben, brauchen wir kein aufwendigeres Oracle
 % Unsere Rankings mit Literatur vergleichen. Wie anders sind wir? -> Evtl. besten Alg. pro Benchmark sammeln und vergleichen
 % Causalitäten checken: Viele Dinge werden einfach genannt, ohne das der Grund richtig eingeführt wird
-% "reproducible" als Aufhänger für 50 repeats ist nicht intuitiv
 % Fig. 1: Schlechte Performance am Anfang verwirrt nur. Entweder BADGE benutzen, anders plotten, oder im Text in Kontext setzen
-% Eventuell Honeypot und Diverging Sine direkt in den Contributions nennen
 % Oracle Forecast im Appendix erklären
-% Tabelle mit AUCs in den Appendix
 
@@
 % !! Sec 3 Related Work sollte Ji et al. im Detail einführen, weil wir uns an deren Guidelines halten. Alles andere sollte dort gekürzt werden
+% !! "Overfitting on test" für Oracle: Sanity-Check durchführen und eventuell komplett streichen
+% !! Introduction neu schreiben: Wir brauchen eine klare Benennung der Ziele und bessere Abgrenzung zu anderen papern 
 % Infos und Ergebnisse für die syntetischen Datensätze in ein Kapitel bündeln und direkt auf die Contribution beziehen
+% !! Sec 4 Methodology ist sehr chaotisch. Wir müssen neu sortieren und unwichtige Dinge streichen
+% Contributions neu sortieren und ausführlicher beschreiben
+% Oracle Performance mit einer Zahl benennen: z.B. Da wird ~20% lift zu allen Baselines haben, brauchen wir kein aufwendigeres Oracle
+% "reproducible" als Aufhänger für 50 repeats ist nicht intuitiv
+% Eventuell Honeypot und Diverging Sine direkt in den Contributions nennen
+% Tabelle mit AUCs in den Appendix
 
@@
 % \end{abstract}
-
-% \section{Introduction-Old Backup}
-% Deep neural networks (NN) have produced state-of-the-art results on many important supervised learning tasks.
-% Since Deep NNs usually require large amounts of labeled training data, Active Learning (AL) can be used to instead select the most informative samples out of a large pool of unlabeled data, so that only these samples need to be labeled.
-% It has been shown that a small labeled set of this nature can be used to train well-performing models \cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}. \\
-% On top of providing a principled way to label unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
-% %Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
-% In the last decade, many different algorithms for AL have been proposed.
-% Even though, almost every method has reported lifts over all its predecessors,
-% \footnote{Out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not claim a new SOTA performance in their result section.}
-% AL research faces four central difficulties:
-% {\let\thefootnote\relax\footnotetext{Code available at: anonymous}}
-% %\let\thefootnote\relax\footnotetext{Code available at: \url{https://github.com/wernerth94/comparable-active-learning}}
-% (i) The experiments are often carried out on different datasets and model architectures, hindering direct comparison,
-% (ii) generalize poorly across different domains,
-% (iii) the reported results can be subject to very high variance across restarts and
-% (iv) are not always compared against important baselines like margin sampling \cite{wang2014new}.
-% While multiple benchmark suites have been proposed to solve (i), to the best of our knowledge, we are the first to report results on all 3 data domains of tabular, image and text.
-% Additionally, we provide synthetic datasets to highlight principled shortcoming of existing AL algorithms.
-% Regarding (ii) and (iii), \cite{zhou2021towards} has pointed out severe inconsistencies in results of AL papers in recent years. 
-% They conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm (\cite{zhou2021towards} Appendix A).
-% These outlined issues lead to an inconclusive landscape of AL algorithms, where the vast majority of reported lifts are neither statistically significant, nor prove to be generalizable.
-% This makes it very hard to identify the best AL algorithm, or even identifying state-of-the-art algorithms.
-% In this work we propose an evaluation protocol that was designed to handle the high variance in the performances of AL algorithms as well as being fully controllable regardless of the combination of dataset, model and AL algorithm.
-% We base our work largely on \cite{ji2023randomness}, following their guidelines for a reliable evaluation of AL algorithms, while extending the number of evaluated data domains from 1 to 5. \\
-% We focus on pool-based AL where a pool of unlabeled samples is fixed at the start of each experiment and one or more samples are chosen sequentially.
-% In addition to the default scenario of selecting a batch of samples every iteration we incorporate the single sample case into our benchmark.
-% Batched algorithms (and benchmarks) do not have a principled advantage over single-sample AL except for speed of computation.
-% The problem of optimizing a portfolio of unlabeled samples in each iteration is more complicated to solve and the algorithms have systematically less information per sample to work with leading to a generally worse performance, that impacts some algorithms more than others.
-% %In Fig. \ref{fig:batch_vs_single} we present a comparison of Badge \cite{ashdeep} and Coreset\cite{sener2017active}, in the batch and single-sample setting.
-% %Even though both algorithms were designed for batch sampling, both perform better in the single-sample setting.
-% We propose single-sample AL as an important tool to identify the best acquisition function, rather than the best combination of acquisition function and diversity regularization. \\
-% %\begin{figure}
-% %    \centering
-% %    \includegraphics[width=\linewidth]{img/ablation_batch_single}
-% %    \caption{Comparison of single-sample vs.
-batch selection for two acquisition functions that were specifically designed for batch selection.
-Even when designed for batch selection, acquisition functions tend to perform better in the single sample setting.}
-% %    \label{fig:batch_vs_single}
-% %\end{figure}
-% Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature. 
-% We offer our datasets in two versions - normal and encoded.
-% An encoded dataset was pre-encoded by an self-supervised encoder model, providing a different use case for active learning.
-% Our synthetic and text datasets (which use word embeddings in the normal setting) are exempt from this, bringing our total dataset count to 14 across 5 domains (Tabular, Image, Text, Synthetic, Encoded).
-% \begin{table}
-% 	\centering
-%         \caption{Comparison of our benchmark with the existing literature.
-Oracle curves serve as an approximation of the best possible AL algorithm.
-Including the encoded versions of our datasets we reach 14 datasets.
-"Semi" indicates whether the paper is employing any form of self- or semi-supervised learning.
-A "-" for repetitions means that we could not determine how often each experiment is repeated in the respective framework.}
-% 	\label{tab:benchmark_comparison}
-% 	\resizebox{\columnwidth}{!}{
-% \begin{tabular}{lccccllllcl}
-% Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
-% \multicolumn{1}{l|}{Beck et al. \cite{beck2021effective}} & batch & 4 & 7 & \checkmark & - & - & - & - & - & - \\
-% \multicolumn{1}{l|}{Hu et al. \cite{hu2021towards}} & batch & 5 & 13 & \checkmark & \checkmark & - & - & - & - & 3 \\
-% \multicolumn{1}{l|}{Zhou et al. \cite{zhou2021towards}} & batch & 3 & 2 & \checkmark & \checkmark & - & - & - & \checkmark & 5 \\
-% \multicolumn{1}{l|}{Zhan et al. \cite{zhan2021comparative}} & sngl+batch & 35 & 18 & - & - & \checkmark & \checkmark & - & \checkmark & 10-100 \\
-% \multicolumn{1}{l|}{Munjal et al. \cite{munjal2022towards}} & batch & 2 & 8 & \checkmark & - & - & - & - & - & 3 \\
-% \multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\
-
-% \multicolumn{1}{l|}{Rauch et al. \cite{rauch2023activeglae}} & batch & 11 & 5 & - & \checkmark & - & - & - & - & 5 \\
-% \hdashline
-% \multicolumn{1}{l|}{Ji et al. \cite{ji2023randomness}} & batch & 3 & 8 & \checkmark & - & - & - & - & - & - \\
-% \multicolumn{1}{l|}{Lueth et al. \cite{luth2024navigating}} & batch & 4 & 5 & \checkmark & - & - & - & \checkmark & - & 3 \\
-% \multicolumn{1}{l|}{\textbf{Ours}} & sngl+batch & 9(14) & 11 & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & 50 
-% \end{tabular}
-% }
-% \end{table}
-
-% \subsection*{Contributions}
-% %Maybe absorb this into the introduction text
-% \begin{enumerate}
-% 	\item[C1] Study on the reproducibility of AL experiments, providing evidence that previous works might have not repeated their experiments often enough to provide reliable results
-% 	\item[C2] Efficient and performant algorithm for an oracle that can be constructed greedily and does not rely on search
-% 	\item[C3] Two novel synthetic datasets that highlight principled shortcomings of of the top-performing AL algorithms
-%     \item[C4] Extending the evaluation of Active Learning algorithms to 5 data domains and revealing significant differences in algorithm performance between them
-% \end{enumerate}
 
@@
 Our contributions include the following:
-
+%
 \begin{enumerate}
-\item We show that the small of repetitions that previous works have employed is not
+\item\label{cont:repetitions} We show that the small of repetitions that previous works have employed is not
 sufficient for meaningful conclusions. Sometimes even making it impossible to assess if a performance is above or below random.
-\item We propose an efficient and performant oracle which is constructed iteratively in a greedy fashion,
+\item\label{cont:oracle} We propose an efficient and performant oracle which is constructed iteratively in a greedy fashion,
 overcoming major computational hurdles.
@@
 % [Thorben] Das können wir so nicht sagen, glaube ich. Niemand verlangt von uns, dass wir das Oracle auch 50 mal wiederholen
-\item We propose \emph{IMTSBench}, the first general benchmark providing tasks
+\item\label{cont:domains} We propose \emph{ALBench}, the first general benchmark providing tasks
 in the domains of images, text and tabular learning. It further contains
@@
 for active learning across different domains.
-\item We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
+\item\label{cont:synthdata} We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
 designed to hinder active learning by naive decision-boundary- or
@@
 \end{enumerate}
-
-%\todo{Ich bin mir nicht ganz sicher, wie man Honeypot und Diverging Sin verkauft. Macht as mit der Lower bar sinn?}
-% \begin{enumerate}
-% \item[C1] Study on the reproducibility of AL experiments, providing evidence
-% that previous works might have not repeated their experiments often enough to
-% provide reliable results
-% \item[C2] Efficient and performant algorithm for an oracle that can be
-% constructed greedily and does not rely on search
-% \item[C3] Two novel synthetic datasets that highlight principled shortcomings of
-% of the top-performing AL algorithms
-% \item[C4] Extending the evaluation of Active Learning algorithms to 5 data
-% domains and revealing significant differences in algorithm performance between
-% them to overcome one of the major obstacles for sufficient 
-% \end{enumerate}
-
-% we are the first to report results on all 3 data domains of
-% tabular, image and text. Additionally, we provide synthetic datasets to
-% highlight principled shortcoming of existing AL algorithms. Regarding (ii) and
-% (iii), \cite{zhou2021towards} has pointed out severe inconsistencies in results
-% of AL papers in recent years. They conducted a meta analysis of reported results
-% of several different AL algorithms and found that all considered algorithms only
-% provided significant lifts in their own original papers, while following
-% literature reported performances no better that uncertainty sampling, or in some
-% cases no better than random sampling for the same algorithm
-% (\cite{zhou2021towards} Appendix A). These outlined issues lead to an
-% inconclusive landscape of AL algorithms, where the vast majority of reported
-% lifts are neither statistically significant, nor prove to be generalizable. This
-% makes it very hard to identify the best AL algorithm, or even identifying
-% state-of-the-art algorithms. In this work we propose an evaluation protocol that
-% was designed to handle the high variance in the performances of AL algorithms as
-% well as being fully controllable regardless of the combination of dataset, model
-% and AL algorithm. We base our work largely on \cite{ji2023randomness}, following
-% their guidelines for a reliable evaluation of AL algorithms, while extending the
-% number of evaluated data domains from 1 to 5. \\
-% We focus on pool-based AL where a pool of unlabeled samples is fixed at the
-% start of each experiment and one or more samples are chosen sequentially. In
-% addition to the default scenario of selecting a batch of samples every iteration
-% we incorporate the single sample case into our benchmark. Batched algorithms
-% (and benchmarks) do not have a principled advantage over single-sample AL except
-% for speed of computation. The problem of optimizing a portfolio of unlabeled
-% samples in each iteration is more complicated to solve and the algorithms have
-% systematically less information per sample to work with leading to a generally
-% worse performance, that impacts some algorithms more than others. 
-% %In Fig. \ref{fig:batch_vs_single} we present a comparison of Badge \cite{ashdeep} and Coreset\cite{sener2017active}, in the batch and single-sample setting.
-% %Even though both algorithms were designed for batch sampling, both perform better in the single-sample setting.
-% We propose single-sample AL as an important tool to identify the best
-% acquisition function, rather than the best combination of acquisition function
-% and diversity regularization. \\
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=\linewidth]{img/ablation_batch_single}
-%    \caption{Comparison of single-sample vs.
-batch selection for two acquisition functions that were specifically designed for batch selection.
-Even when designed for batch selection, acquisition functions tend to perform better in the single sample setting.}
-%    \label{fig:batch_vs_single}
-%\end{figure}
-% Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature. 
-% We offer our datasets in two versions - normal and encoded.
-% An encoded dataset was pre-encoded by an self-supervised encoder model, providing a different use case for active learning.
-% Our synthetic and text datasets (which use word embeddings in the normal
-% setting) are exempt from this, bringing our total dataset count to 14 across 5
-% domains (Tabular, Image, Text, Synthetic, Encoded). 
+%
 \begin{table}
@@
 %Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
-The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also place a much higher emphasis on the problem of evaluating AL algorithms under many forms of variance than their predecessors (indicated in Tab.
+The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also place a much higher emphasis on the problem of evaluating AL algorithms under high variance than their predecessors (indicated in Tab.
 \ref{tab:benchmark_comparison} by a dashed line).
 The authors of \cite{ji2023randomness} posed a total of 12 "recommendations" for reliable evaluation of AL algorithms.
-We largely adapt the proposed recommendations of \cite{ji2023randomness} and extend their work to multiple domains, batch sizes and comparisons.
+We largely adapt the proposed recommendations and extend their work to multiple domains and batch sizes.
 For a complete list of the recommendations and our implementation of them, please refer to App. \ref{app:recommendations}.
 This work also pays attention to the so-called "pitfalls" of AL evaluation proposed in \cite{luth2024navigating}.
-For a complete list of the pitfalls and our implementation of them, please refer to App. \ref{app:pitfalls}.
+For a complete list of the pitfalls and our considerations regarding them, please refer to App. \ref{app:pitfalls}.
 To the best of our knowledge, we are the first to extend reliable SOTA (based on \cite{ji2023randomness, luth2024navigating}) experimentation to a total of 5 data domains and a high number of restarts per experiment.
@@
 For this benchmark we employ 50 restarts of every experiment.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Seeding vs. Restarts}\label{sec:reproducibility}
-Considering the high computational cost of 50 repetitions, another approach to ensure consistency between experiments would be to reduce the amount of variance in the experiment by keeping as many subsystems (weight initialization, data splits, etc.) as possible fixed with specialized seeding.
-\\
-We describe a novel seeding strategy in Appendix \ref{app:seeding_strategy} that creates 3 separate Random Number Generators (RNG) based on 3 different seeds.
-In short, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini-batch sampling, and $s_\theta$ for model initialization and sampling of dropout masks.
-Unless stated otherwise, we will keep $s_\Omega$ fixed, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
-While this seeding strategy is capable of controlling the amount variance in the experiment, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), providing a bad approximation of the quality of an AL algorithm (i.e.
-measured performances for an algorithm might not even transfer to a different model initialization).
-Hence, we opt for letting the subsystems vary (by increasing $s_\mathcal{D}$ and $s_\theta$) and combine that with a high number of restarts to obtain a good average of the generalization performance of each AL algorithm.
-\\
-Where a high number of restarts is computationally not feasible, we advise to additionally keep either $s_\mathcal{D}$ or $s_\theta$ (or both) fixed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{ALBench: A Cross-Domain Active Learning Benchmark}\label{sec:datasets}
-A detailed description of the preprocessing of each dataset can be found in Appendix \ref{app:hyperparameters}. \\ [1mm]
-\textbf{Tabular:}
-AL research conducted on tabular data is sparse (only \cite{ashdeep} from the considered baseline papers). 
-We, therefore, introduce a set of tabular datasets that we selected according to the following criteria:
-(i) They should be solvable by medium-sized models in under 1000 samples, (ii) the gap between most AL algorithms and random sampling should be significant (potential for AL is present) and (iii) the gap between the AL algorithms and our oracle should also be significant (research on these datasets can produce further lifts).
-We use \textbf{Splice}, \textbf{DNA} and \textbf{USPS} from LibSVMTools \cite{libsvmtools}.\\
-\textbf{Image:}
-We use \textbf{FashionMNIST} \cite{xiao2017fashion} and \textbf{Cifar10} \cite{krizhevsky2009learning}, since both are widely used in AL literature.\\
-\textbf{Text:}
-We use \textbf{News Category} \cite{misra2022news} and \textbf{TopV2} \cite{chen-etal-2020-low-resource}.
-Text datasets have seen less attention in AL research, but most of the papers that evaluate on text (\cite{hu2021towards}, \cite{zhou2021towards}) use at least one of these datasets.\\ [1mm]
-%
-We would like to point out that these datasets are selected for speed of computation (both in terms of the required classifier and the necessary budget to solve the dataset). 
-However, similar to our argumentation for picking smaller classifiers, we are solely focused on comparing different AL algorithms in this paper and do not aim to develop novel classification models on these datasets.
-Our assumption is that a well-performing algorithm in our benchmark will also generalize well to larger real-world datasets, because we included multiple different data domains and classifier sizes in our experiments.
-\\ [1mm]
-Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning.
-The text datasets are an exception to this, as they are only offered in their encoded form.
-Pre-encoded datasets enable us to test small query sizes on more complex datasets like Cifar10 and FashionMnist.
-They also serve the purpose of investigating the interplay between self-supervised learning techniques and AL, as well as alleviating the cold-start problem described in \cite{luth2024navigating} as they require a way smaller seed set.
-The classification model for every encoded dataset is a single linear layer with softmax activation.
-The embedding model was trained with the SimCLR \cite{chen2020simple} algorithm adopting the protocol from \cite{hacohen2022active}. 
-%For Cifar10 and FashionMnist we adapt the reported hyperparameters from \cite{hacohen2022active} and for the tabular datasets we use random search to optimize the hyperparameters.
-To ensure that enough information from the data is encoded by our embedding model, the quality of embeddings during pretext training was measured after each epoch.
-To this end, we attached a linear classification head to the encoder, fine-tuned it to the data and evaluated this classifier for test accuracy. 
-The checkpoint of each encoder model will be provided together with the framework. \\ [1mm]
-Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.
-
 \begin{wraptable}{R}{0.65\textwidth}
@@
 	}
-    \vspace{-0.85cm}
+	\vspace{-0.55cm}
 \end{wraptable}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Seeding vs. Restarts}\label{sec:reproducibility}
+Considering the high computational cost of 50 repetitions, another approach to ensure consistency between experiments would be to reduce the amount of variance in the experiment by keeping as many subsystems (weight initialization, data splits, etc.) as possible fixed with specialized seeding.
+\\
+We describe a novel seeding strategy in Appendix \ref{app:seeding_strategy} that creates 3 separate Random Number Generators (RNG) based on 3 different seeds.
+\todo{Wahrscheinlich zu wenig Erklärung für den main body. Eventuell die gesamten Erklärungen zum seeding in den appendix?}
+In short, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini-batch sampling, and $s_\theta$ for model initialization and sampling of dropout masks.
+Unless stated otherwise, we will keep $s_\Omega$ fixed, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
+While this seeding strategy is capable of controlling the amount variance in the experiment, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), providing a bad approximation of the quality of an AL algorithm (i.e.
+measured performances for an algorithm might not even transfer to a different model initialization).
+Hence, we opt for letting the subsystems vary (by increasing $s_\mathcal{D}$ and $s_\theta$) and combine that with a high number of restarts to obtain a good average of the generalization performance of each AL algorithm.
+\\
+Where a high number of restarts is computationally not feasible, we advise to additionally keep either $s_\mathcal{D}$ or $s_\theta$ (or both) fixed.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{ALBench: A Cross-Domain Active Learning Benchmark}\label{sec:datasets}
+A detailed description of the preprocessing of each dataset can be found in Appendix \ref{app:hyperparameters}. \\ [1mm]
+\textbf{Tabular:}
+AL research conducted on tabular data is sparse (only \cite{ashdeep} from the considered baseline papers). 
+We, therefore, introduce a set of tabular datasets that we selected according to the following criteria:
+(i) They should be solvable by medium-sized models in under 1000 samples, (ii) the gap between most AL algorithms and random sampling should be significant (potential for AL is present) and (iii) the gap between the AL algorithms and our oracle should also be significant (research on these datasets can produce further lifts).
+We use \textbf{Splice}, \textbf{DNA} and \textbf{USPS} from LibSVMTools \cite{libsvmtools}.\\
+\textbf{Image:}
+We use \textbf{FashionMNIST} \cite{xiao2017fashion} and \textbf{Cifar10} \cite{krizhevsky2009learning}, since both are widely used in AL literature.\\
+\textbf{Text:}
+\todo{Schwache Erklärung}
+We use \textbf{News Category} \cite{misra2022news} and \textbf{TopV2} \cite{chen-etal-2020-low-resource}.
+Text datasets have seen less attention in AL research, but most of the papers that evaluate on text (\cite{hu2021towards}, \cite{zhou2021towards}) use at least one of these datasets.\\ [1mm]
+%
+We would like to point out that these datasets are selected for speed of computation (both in terms of the required classifier and the necessary budget to solve the dataset). 
+However, we are solely focused on comparing different AL algorithms in this paper and do not aim to develop novel classification models on these datasets.
+Our assumption is that a well-performing algorithm in our benchmark will also generalize well to larger real-world datasets, because we included multiple different data domains and classifier types and sizes in our experiments.
+\\ [1mm]
+\todo{Muss die Verbindung zwischen encodings und semi-supervised learning klarer?}
+Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning.
+The text datasets are an exception to this, as they are only offered in their encoded form.
+Pre-encoded datasets enable us to test small query sizes on more complex datasets like Cifar10 and FashionMnist.
+They also serve the purpose of investigating the interplay between self-supervised learning techniques and AL, as well as alleviating the cold-start problem described in \cite{luth2024navigating} as they require a way smaller seed set.
+The classification model for every encoded dataset is a single linear layer with softmax activation.
+The embedding model was trained with the SimCLR \cite{chen2020simple} algorithm adopting the protocol from \cite{hacohen2022active}. 
+%For Cifar10 and FashionMnist we adapt the reported hyperparameters from \cite{hacohen2022active} and for the tabular datasets we use random search to optimize the hyperparameters.
+To ensure that enough information from the data is encoded by our embedding model, the quality of embeddings during pretext training was measured after each epoch.
+To this end, we attached a linear classification head to the encoder, fine-tuned it to the data and evaluated this classifier for test accuracy. 
+The checkpoint of each encoder model will be provided together with the framework. \\ [1mm]
+Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.
 
@@
 \end{equation}
-The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
-AL algorithms have to perform well in both scenarios. 
-\\ [1mm]
+%The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
+%AL algorithms have to perform well in both scenarios. \\ [1mm]
 Since AUC is still influenced by the budget, we define a set of rules to set this hyperparameter upfront, so that we are not favoring a subset of algorithms by handcrafting a budget.
@@
 These ranks can safely be averages across datasets as they are no longer subjected to scaling differences of each dataset.
+\todo{Erklärung zum paired-t-test ist schlecht geschrieben. Wie machen wir es besser?}
 Additionally, we employ Critical Difference (CD) diagrams (like Fig. \ref{fig:ranks_by_domain}) for statistical testing.
@@
 \section{A Greedy Oracle Algorithm}\label{sec:oracle}
-%an oracle set was so far always described as the optimal set and therefore
-A oracle method for AL tries to find the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure that induces the highest AUC score for a given budget.
-However, since this combinatorial problem is computationally infeasible for realistic datasets, previous works have proposed approximations to this oracle sequence. 
+An oracle method for AL tries to find the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure that induces the highest AUC score for a given budget.
+However, due to the combinatorial nature of the problem, it is computationally infeasible for realistic datasets.
+Hence, previous works have proposed approximations to this oracle sequence. 
 \cite{zhou2021towards} used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
@@
 \subsection{Implementation Details}\label{sec:implementation_details}
+\todo{[Thorben] Sollten wir den ganzen ersten Teil steichen? Wir haben uns von technischen Beschreibungen weitesgehend wegbewegt.}
 At each iteration $i$ the acquisition function $\Omega$ picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \hat y^{(i)}, \text{opt}_{\hat y}\}$, where $\text{opt}_{\hat y}$ is the optimizer used to fit $\hat y^{(i)}$.
@@
 In Table \ref{tab:results} we provide the rank of each acquisition function per dataset and averaged for each (un-)encoded dataset. 
-Please note, that for Tab \ref{tab:results} we are averaging not only over runs, but also over query sizes per dataset. 
+Please note, that for Tab. \ref{tab:results} we are averaging not only over runs, but also over query sizes per dataset. 
+This leads to a negative impact for AL methods that do not adapt well to larger batch sizes.
 For the results per query size, please refer to App. \ref{app:auc_by_query_size}. \\
-As stated in contribution C4, our results on real-world data shows significant differences in the performance of the tested algorithms between data domains.
+As mentioned in contribution \ref{cont:domains}, our results on real-world data shows significant differences in the performance algorithms between data domains.
 Not only do some algorithms overperform on some domains (like least confidence sampling on Images), but the Top-3 of algorithms (except Oracle) does not contain the same three algorithms for any two domains.
-Most interestingly, the image domain, which received the most attention in benchmarking so far could even be considered an outlier, as this is the only domain where the Top-1 algorithm changes.
+Most interestingly, the image domain, which received most of the attention in benchmarking so far could even be considered an outlier, as this is the only domain where the Top-1 algorithm changes.
 This highlights the dire need for diverse data domains in AL benchmarking. \\ [1mm]
@@
 \begin{tabular}{l|lllllll|ll}
-          & Splice & DNA  & USPS 		 	& Cfr10 & FMnist & TopV2 & News  & Un-enc. & Enc. \\
-          \hline
-Oracle  & 1.0 $\pm$ 0.01 & 1.0 $\pm$ 0.01 & 1.0 $\pm$ 0.0   & 1.0 $\pm$ 0.0   & 1.0 $\pm$ 0.0   & 1.0 $\pm$ 0.01 & 1.0 $\pm$ 0.0  & 1.0 & 2.0 \\
-Margin  & 6.6 $\pm$ 0.02 & 4.3 $\pm$ 0.01 & 2.1 $\pm$ 0.01  & 6.3 $\pm$ 0.01  & 4.4 $\pm$ 0.0   & 2.4 $\pm$ 0.01 & 3.7 $\pm$ 0.0  & 4.3 & 4.2 \\
-Badge   & 5.2 $\pm$ 0.01 & 6.3 $\pm$ 0.01 & 2.9 $\pm$ 0.01  & 5.2 $\pm$ 0.01  & 4.7 $\pm$ 0.0   & 3.3 $\pm$ 0.01 & 3.5 $\pm$ 0.0  & 4.5 & 5.4 \\
-LeastConf & 9.2 $\pm$ 0.02 & 10.3 $\pm$ 0.02 & 8.1 $\pm$ 0.02  & 2.1 $\pm$ 0.01 & 4.0 $\pm$ 0.0   & 7.9 $\pm$ 0.02  & 3.0 $\pm$ 0.01  & 6.4  & 6.5 \\
-DSA     & 7.4 $\pm$ 0.02 & 7.3 $\pm$ 0.01 & 7.5 $\pm$ 0.01  & 5.4 $\pm$ 0.01  & 5.1 $\pm$ 0.0   & 6.0 $\pm$ 0.02 & 7.3 $\pm$ 0.01 & 6.6 & 6.7 \\
-BALD    & 4.0 $\pm$ 0.01 & 4.7 $\pm$ 0.01 & 5.4 $\pm$ 0.01  & 12.0 $\pm$ 0.01 & 7.6 $\pm$ 0.0   & 7.6 $\pm$ 0.02 & 5.0 $\pm$ 0.0  & 6.6 & 7.6 \\
-CoreGCN & 6.9 $\pm$ 0.01 & 4.9 $\pm$ 0.01 & 10.4 $\pm$ 0.01 & 7.6 $\pm$ 0.01  & 6.5 $\pm$ 0.01  & 4.0 $\pm$ 0.01 & 6.8 $\pm$ 0.0  & 6.7 & 8.2 \\
-Entropy & 6.6 $\pm$ 0.02 & 3.9 $\pm$ 0.01 & 7.6 $\pm$ 0.01  & 7.6 $\pm$ 0.01  & 4.9 $\pm$ 0.01  & 9.8 $\pm$ 0.02 & 9.6 $\pm$ 0.0  & 7.1 & 6.5 \\
-LSA     & 6.1 $\pm$ 0.01 & 6.8 $\pm$ 0.01 & 5.3 $\pm$ 0.01  & 7.7 $\pm$ 0.01  & 10.6 $\pm$ 0.01 & 7.5 $\pm$ 0.01 & 7.3 $\pm$ 0.01 & 7.3 & 7.5 \\
-Random  & 9.0 $\pm$ 0.01 & 9.3 $\pm$ 0.01 & 5.3 $\pm$ 0.01  & 8.4 $\pm$ 0.01  & 11.1 $\pm$ 0.0  & 7.9 $\pm$ 0.01 & 8.0 $\pm$ 0.0  & 8.4 & 6.9 \\
-Coreset        & 7.1 $\pm$ 0.01 & 9.0 $\pm$ 0.01  & 10.5 $\pm$ 0.01 & 6.8 $\pm$ 0.01 & 7.1 $\pm$ 0.0   & 8.5 $\pm$ 0.02  & 10.8 $\pm$ 0.01 & 8.5  & 7.2 \\
-TypiClust      & 8.8 $\pm$ 0.01 & 10.2 $\pm$ 0.01 & 12.0 $\pm$ 0.02 & 7.9 $\pm$ 0.01 & 11.0 $\pm$ 0.01 & 12.0 $\pm$ 0.02 & 12.0 $\pm$ 0.01 & 10.5 & 9.2
+          & Splice & DNA  & USPS 		 	& Cifar10 & FMnist & TopV2 & News  & Un-enc. & Enc. \\
+          \hline
+Oracle   & 1.0 $\pm$ {\small 0.01} & 1.0 $\pm$ {\small 0.01} & 1.0 $\pm$ {\small 0.0}  & 1.0 $\pm$ {\small 0.0}  & 1.0 $\pm$ {\small 0.0}  & 1.0 $\pm$ {\small 0.01} & 1.0 $\pm$ {\small 0.0} & 1.0 & 2.0 \\
+Margin   & 6.6 $\pm$ {\small 0.02} & 4.3 $\pm$ {\small 0.01} & 2.1 $\pm$ {\small 0.01} & 6.3 $\pm$ {\small 0.01} & 4.4 $\pm$ {\small 0.0}  & 2.4 $\pm$ {\small 0.01} & 3.7 $\pm$ {\small 0.0} & 4.3 & 4.2 \\
+Badge    & 5.2 $\pm$ {\small 0.01} & 6.3 $\pm$ {\small 0.01} & 2.9 $\pm$ {\small 0.01} & 5.2 $\pm$ {\small 0.01} & 4.7 $\pm$ {\small 0.0}  & 3.3 $\pm$ {\small 0.01} & 3.5 $\pm$ {\small 0.0} & 4.5 & 5.4 \\
+LeastConf & 9.2 $\pm$ {\small 0.02} & 10.3 $\pm$ {\small 0.02} & 8.1 $\pm$ {\small 0.02} & 2.1 $\pm$ {\small 0.01} & 4.0 $\pm$ {\small 0.0}  & 7.9 $\pm$ {\small 0.02} & 3.0 $\pm$ {\small 0.01} & 6.4 & 6.5 \\
+DSA      & 7.4 $\pm$ {\small 0.02} & 7.3 $\pm$ {\small 0.01} & 7.5 $\pm$ {\small 0.01} & 5.4 $\pm$ {\small 0.01} & 5.1 $\pm$ {\small 0.0}  & 6.0 $\pm$ {\small 0.02} & 7.3 $\pm$ {\small 0.01} & 6.6 & 6.7 \\
+BALD     & 4.0 $\pm$ {\small 0.01} & 4.7 $\pm$ {\small 0.01} & 5.4 $\pm$ {\small 0.01} & 12.0 $\pm$ {\small 0.01} & 7.6 $\pm$ {\small 0.0}  & 7.6 $\pm$ {\small 0.02} & 5.0 $\pm$ {\small 0.0} & 6.6 & 7.6 \\
+CoreGCN  & 6.9 $\pm$ {\small 0.01} & 4.9 $\pm$ {\small 0.01} & 10.4 $\pm$ {\small 0.01} & 7.6 $\pm$ {\small 0.01} & 6.5 $\pm$ {\small 0.01} & 4.0 $\pm$ {\small 0.01} & 6.8 $\pm$ {\small 0.0} & 6.7 & 8.2 \\
+Entropy  & 6.6 $\pm$ {\small 0.02} & 3.9 $\pm$ {\small 0.01} & 7.6 $\pm$ {\small 0.01} & 7.6 $\pm$ {\small 0.01} & 4.9 $\pm$ {\small 0.01} & 9.8 $\pm$ {\small 0.02} & 9.6 $\pm$ {\small 0.0} & 7.1 & 6.5 \\
+LSA      & 6.1 $\pm$ {\small 0.01} & 6.8 $\pm$ {\small 0.01} & 5.3 $\pm$ {\small 0.01} & 7.7 $\pm$ {\small 0.01} & 10.6 $\pm$ {\small 0.01} & 7.5 $\pm$ {\small 0.01} & 7.3 $\pm$ {\small 0.01} & 7.3 & 7.5 \\
+Random   & 9.0 $\pm$ {\small 0.01} & 9.3 $\pm$ {\small 0.01} & 5.3 $\pm$ {\small 0.01} & 8.4 $\pm$ {\small 0.01} & 11.1 $\pm$ {\small 0.0}  & 7.9 $\pm$ {\small 0.01} & 8.0 $\pm$ {\small 0.0} & 8.4 & 6.9 \\
+Coreset  & 7.1 $\pm$ {\small 0.01} & 9.0 $\pm$ {\small 0.01} & 10.5 $\pm$ {\small 0.01} & 6.8 $\pm$ {\small 0.01} & 7.1 $\pm$ {\small 0.0}  & 8.5 $\pm$ {\small 0.02} & 10.8 $\pm$ {\small 0.01} & 8.5 & 7.2 \\
+TypiClust & 8.8 $\pm$ {\small 0.01} & 10.2 $\pm$ {\small 0.01} & 12.0 $\pm$ {\small 0.02} & 7.9 $\pm$ {\small 0.01} & 11.0 $\pm$ {\small 0.01} & 12.0 $\pm$ {\small 0.02} & 12.0 $\pm$ {\small 0.01} & 10.5 & 9.2 \\
 \end{tabular}
@@
     \includegraphics[width=0.49\linewidth]{img/macro_enc.jpg}
+    \vspace{-7mm}
 \end{figure}
@@
 	\centering
+	\caption{Synthetic "Honeypot" and "Diverging Sine" datasets. The optimal decision boundary is not part of the dataset and serves only as a visual guide.}
 	\includegraphics[width=0.4\linewidth]{img/honeypot.jpg}
 	\includegraphics[width=0.4\linewidth]{img/diverging_sin.jpg}
-	\caption{Synthetic "Honeypot" and "Diverging Sine" datasets. The optimal decision boundary is not part of the dataset and serves only as a visual guide.}
 	\label{fig:synthDataAppendix}
@@
 %
+\todo{Hier könnte gekürzt werden. Wir brauchen wohl kaum die genaue Anzahl der Punkte}
 Honeypot creates to two easy to distinguish clusters with 150 samples each and one overlapping "honeypot" that represents a noisy region of the dataset with potentially miss-labeled, miss-measured or generally adverse samples.
@@
 Results for Diverging Sine also confirm expected behavior, as clustering algorithms (Coreset, TypiClust) fall behind uncertainty algorithms (Entropy-, Margin-Sampling), with the exception of BADGE. \\
-We provide a very small ablation study on the importance of the embeddings by testing a version of Coreset and TypiClust on this dataset that does not use the embeddings produced by the classification model, but rather clusters the data directly.
+We provide a small ablation study on the importance of the embeddings by testing a version of Coreset and TypiClust on this dataset that does not use the embeddings produced by the classification model, but rather clusters the data directly.
 "Coreset Raw" and "TypiClust Raw" both perform worse than their embedding-based counterpart.

### 2024-06-04 doc/al_benchmark_paper.tex
Clarified semi-supervised setting

@@
 computer vision, natural language processing and the tabular domain. We provide
-our datasets both in raw format %(i.e. consisting of text, images etc) 
+our datasets both in normal format %(i.e. consisting of text, images etc)
 as well as ``embedded'' by a fixed embedding model, enabling evaluation of AL methods in
@@
 \todo{Muss die Verbindung zwischen encodings und semi-supervised learning klarer?}
-Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning.
+Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning (semi-supervised).
 The text datasets are an exception to this, as they are only offered in their encoded form.
 Pre-encoded datasets enable us to test small query sizes on more complex datasets like Cifar10 and FashionMnist.
-They also serve the purpose of investigating the interplay between self-supervised learning techniques and AL, as well as alleviating the cold-start problem described in \cite{luth2024navigating} as they require a way smaller seed set.
+They also serve the purpose of investigating the interplay between semi-supervised learning techniques and AL, as well as alleviating the cold-start problem described in \cite{luth2024navigating} as they require a way smaller seed set.
 The classification model for every encoded dataset is a single linear layer with softmax activation.

### 2024-06-04 doc/al_benchmark_paper.tex
recommendations and pitfalls from baselines papers

@@
 
+Recommendation. Use the backbone architecture with
+the community-accepted definition that is best suited
+for the dataset at hand and consistently use it across all
+experiments. In the image classification domain, we
+suggest using ResNet18 for CIFAR-10 and CIFAR-100.
+
+Recommendation. Control the type of optimizer across
+methods for comparative evaluations to ensure that the
+yield performance difference stems from an active learning method itself. As SGD often generalizes better, we
+encourage its use for deep active learning.
+
+Recommendation. Pragmatically fix the learning rate
+to 0.1 for SGD on image datasets. While continuous
+hyperparameter tuning can improve overall performance,
+a fixed learning rate does not change the ranking of AL
+methods from a comparative evaluation’s point of view.
+
+Recommendation. One may use data augmentation if
+applied consistently across methods, such that it does
+not affect the overall ranking. However, a commonly
+accepted baseline is needed, e.g., random horizontal flipping and random cropping for image classification.
+
+Recommendation. Refine model parameters across AL
+batch (“warm starts”) to prevent exhaustive reinitialization and feed initialization of the backbone model’s
+weights and the “init sets” with fixed inputs over multiple runs to average out the randomness. Moreover, use
+identical seeds for all methods under investigation.
+
+Recommendation. Run experiments multiple times to
+compensate for non-deterministic operations. If the resulting variance is larger than the gained improvement,
+use deterministic operations stringently
+
+Recommendation SW. Configure and verify influence parameter in active learning implementations thoroughly. To foster future research, we
+provide implementations as part of our framework at:
+https://intellisec.de/research/eval-al
+
+Recommendation HW. Ensure that comparative evaluations are run on identical hardware. While it is not
+necessary to execute all experiments on the same physical device, the GPU model, for instance, should be the
+same. Do not mix hardware and list hardware details.
+
+Recommendation. Consider multiple query-batch sizes
+in the evaluation. The choice of the sizes needs to be
+appropriate for the total number of unlabeled samples.
+
+Recommendation. Compare active learning strategies
+without sub-sampling, unless one of the approaches uses
+it as a fundamental building block. In this case a detailed
+analysis of the influence of sub-sampling is necessary.
+
+Recommendation. Evaluate active learning strategies
+on multiple benchmark datasets, that comprise balanced,
+imbalanced, small-scale, and large-scale datasets to cover
+most relevant cases in practice.
+
+Recommendation. For a comprehensive analysis of
+AL strategies, the overall comparative evaluation should
+incorporate as many variables from Section 3 to yield a
+summarized PPM that is as expressive as possible.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 TODO
+
+P1 Data distribution. The proposed evaluation over a diverse selection of dataset distributions
+including specific roll-out datasets proved essential for realistic evaluation of QMs as well as the
+different training strategies. One main insight is the fact that class distribution is a crucial predictor
+for the potential performance gains of AL on a dataset: Performance gains of AL are generally higher
+on imbalanced datasets and occur consistently even for ST models with a small starting budget,
+which are typically prone to experience cold start problems. This observation is consistent with a few
+previous studies [30, 35, 54]. Further, our results underpin the importance of the roll-out datasets e.g.
+when looking at the sub-random performance of BALD (with Self-SL) and Entropy (with ST) on
+MIO-TCD. Such worst-case failures of AL application (increased compute and labeling effort due to
+AL) could not have been predicted based on development data where all AL-parameters are optimized.
+Another example is the lack of generalizability of Semi-SL, where performance in relation to other
+Self-SL and ST decreases gradually with data complexity (going from CIFAR-10 to CIFAR-100/-10
+LT to the two roll-out datasets MIO-TCD and ISIC-2019.)
+
+P2 Starting budget. The comprehensive study of various starting budgets on all datasets reveals
+that AL methods are more robust with regard to small starting budgets than previously reported
+[6, 20, 43]. With the exception of Entropy we did not observe cold start problems even for any QM
+even in combination with notoriously prone ST models. The described robustness is presumably
+enabled by our thorough classifier configuration (P4) and heuristically adapted query sizes (P3). This
+finding has great impact potential suggesting that AL can be applied at earlier points in the annotation
+process thereby further reducing the labeling cost, especially in combination with BADGE which
+performed consistently better or on par with Random. Similarly for Self-SL models, AL performs
+well on small starting budgets with the exceptions of CIFAR-100 (BALD, Entropy) and MIO-TCD
+(BALD, Core-Set, Entropy).
+
+P3 Query size. Based on our evaluation of the query size we can empirically confirm its importance
+with regard to 1) general AL performance and 2) counteracting the cold start problem. The are,
+however, surprising findings indicating that the exact interaction between query size and performance
+remains an open research question. For instance, we observe the cold start problem for BALD with
+Self-SL on CIFAR-100 (∼ 50% accuracy at 2k labeled samples for query sizes of 500 (low-label
+regime) and 1k (mid-label regime). On the other hand, in the high label regime (budget of 5k and
+query size of 5k) ST and Self-SL models with similar accuracies of 50% and 60%, respectively,
+benefit from BALD. Since cold start problems are commonly associated with large query sizes, this
+finding seems counter-intuitive, but has been reported before although without further investigations
+[6, 20, 43]. To gain a better understanding of this phenomenon and how QMs interact with query
+size in the low-label regime, we performed a dedicated experiment series for Self-SL training on
+CIFAR-100 and ISIC-2019 (see Figure 3 and Appendix F.2). Due to a significant improvement
+of BALD on CIFAR-100 for even smaller query sizes from sub-Random to Random performance
+and no worsening on ISIC-2019, we can conclude that small query sizes represent an effective
+countermeasure against the cold-start problem for BALD contrary to the findings of [44, 58] and
+currently not considered as a solution [6, 20, 43]. This potentially explains the gap between these
+findings and more theoretical works which advertise using the smallest query size possible [19, 33].
+Importantly, while smaller query sizes appear as a potential solution for the observed instabilities of
+BALD, they seem to have no considerable effect on the other QMs. Moreover, our ablation reveals
+that even in the low-label settings, BADGE is the overall most reliable of all compared QMs and
+exhibits no sub-Random performance. This adds to the existing reports of the robustness of BADGE
+for higher label regimes [4, 58].
+
+P4 Classifer configuration. Our results show that method configuration on a properly sized validation
+set is essential for realistic evaluation in AL. For instance, our method configuration has the same
+effect on the classifier performance as increasing the number of labeled training samples by a factor
+of ∼ 5 (e.g. our ST model reach approximately the same accuracy of ∼ 44% trained on 200 samples
+compared to models in the AL study by Bengar et al. [6] trained on 1k samples). The effectiveness of
+our proposed lightweight HP selection on the starting budget including only three parameters (Sec. 3)
+is demonstrated by the fact that all our ST models substantially outperform respective models found
+in relevant literature [6, 20, 30, 35, 43, 54, 55] where HP optimization is generally neglected. Details
+for this comparison are provided in Appendix H. This raises the question of to which extent reported
+AL advantages could have been achieved by simple classifier configurations. Further, our models
+also generally outperform expensively configured models by Munjal et al. [44]. Thus, we conclude
+that manually constraining the search space renders HP optimization feasible in practice without
+decreasing performance and ensures performance gains by Active Learning are not overstated. The
+importance of the proposed strategy to optimize HPs on the starting budget for each new dataset is
+supported by the fact that the resulting configurations change across datasets.
+
+P5 Alternative training paradigms. Based on our study benchmarking AL in the context of
+both Self-SL and Semi-SL, we see that while Self-SL generally leads to improvements across all
+experiments, Semi-SL only leads to considerably improved performance on the simpler datasets
+CIFAR-10/100, on which Semi-SL methods are typically developed. Generally, models trained
+with either of the two training paradigms receive a lower performance gain from AL (over random
+querying) compared to ST. Crucially, Self-SL models converge around 2.5 times faster than ST
+models, while the training time of Semi-SL models is around 80 times longer than ST and often yields
+only small benefits over Self-SL or ST models. The fact that AL entails multiple training iterations
+amplifies the computational burden of Semi-SL, rendering their combination prohibitively expensive
+in most practical scenarios. Further, the fact that our Semi-SL models based on Fixmatch do not seem
+to generalize to more complex datasets in our setting stands in stark contrast to conclusions drawn by
+[20, 43] as to which the emergence of Semi-SL renders AL redundant. Interestingly, the exact settings
+where Semi-SL does not provide benefits in our study are the ones where AL proved advantageous.
+The described contradiction with the literature underlines the importance of our proposed protocol
+testing for a method’s generalizability to unseen datasets. This is especially critical for Semi-SL,
+which is known for unstable performance on noisy and class imbalanced datasets as noted by studies
+focusing on Semi-SL [7, 22, 23, 29, 45, 56].
+
+OPTIONAL - aus deren conclusion
+Main empirical insights revealed by our study:
+\begin{itemize}
+\item Assessment of AL methods in the literature is substantially hampered by subpar classifier configurations, but meaningful assessment can be restored by our protocol for lightweight hyperparameter
+tuning on the starting budget.
+\item AL generally provides substantial gains in class-imbalanced settings.
+\item BADGE is the best-performing QM across a realistic range of datasets, starting budgets, and query
+sizes and exhibits no failure modes (i.e. sub-Random performance).
+\item Combining AL with Self-SL considerably improves performance, shortens the training time, and
+stabilizes optimization, especially in low-label settings.
+\item FixMatch results indicate that Semi-SL methods perform well on datasets where they have been
+developed, but may struggle to generalize to more realistic scenarios such as class imbalanced data.
+Combining the paradigm with AL additionally suffers from extensive training times.
+\item BALD with Self-SL pre-trained models benefits from smaller query sizes on small starting budgets
+possibly circumventing the "cold start problem".
+\end{itemize}
+
+Take-aways for developing and proposing new Active Learning algorithms:
+AL methods should be tested for generalizability on roll-out datasets and come with a clear recipe
+for real-world application including how to adapt all design choices to new settings. Since the
+expected benefit of AL on a new setting increases with lower application costs, we believe there is a
+high potential for wide-spread real-world use of AL by reducing the two prevalent cost factors: 1)
+Engineering costs, which are reduced by building easy-to-use AL tools. 2) Computational costs,
+which are reduced by explicitly including methods that shorten the training time in AL.
+
+Take-aways for the cost-benefit analysis of deploying Active Learning:
+\begin{enumerate}
+\item Since we identify BADGE as a robust query method exhibiting no sub-random-sampling performance across all tested settings, the potential harm of AL is minimized allowing to base decisions
+around deploying AL on the described cost-benefit analysis.
+\item The expected benefit is high in settings, where there is a mismatch between the task-specific
+importance of individual classes and their frequency in the dataset (e.g. class-imbalanced datasets
+in combination with tasks requiring a balanced classifier).
+\item The expected benefit further increases with the labeling cost that will be reduced by AL. AL is thus
+likely to yield a net benefit in settings with high label cost and low computational and engineering
+costs.
+\end{enumerate}
 

### 2024-06-04 doc/al_benchmark_paper.tex
unicode fixes

@@
 % - (Un-)encoded vs. semi-supervised/normal
+% - ThreeClust vs. Honeypot
 % Gibt es eine Conclusion für die Encoded-Datasets? Im Moment werden die nicht besprochen
@@
 learning. Furthermore, by providing an efficient greedy oracle, \emph{ALBench}
-can be evaluated with 50 runs for each experiment. We will show, that both the
+can be evaluated with 50 runs for each experiment. We show, that both the
 cross-domain character and the large amount of repetitions are crucial for
-sophisticated evaluation of AL research. Concretely, we will show that the
+sophisticated evaluation of AL research. Concretely, we show that the
 superiority of specific methods varies over the different domains, making it
@@
 conducting five runs as often done in the literature, the superiority of
-specific methods can strongly vary with the specific runs. This effect goes so
-far, that it may even happen that dedicated methods like \emph{margin-sampling}
-are not able to outperform the random baseline.
+specific methods can strongly vary with the specific runs.
+This effect is so strong, that, depending on the seed, even a well-established method's performance can be significantly better and significantly
+worse than random for the same dataset. 
 \end{abstract}
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 Deep neural networks (NN) have produced state-of-the-art results on many
@@
 the different methods averaged over many datasets fluctuates with the specific set of runs. 
-This effect is so strong, that even for a well-established method such
-as margin-sampling, it's performance can be significantly better and significantly
-worse than random for the same dataset, depending on the seed.
+This effect is so strong, that, depending on the seed, even a well-established method's performance can be significantly better and significantly worse than random for the same dataset.
 
@@
 we believe that \emph{ALBench} is a major step forward of assessing the overall state of Active Learning research, independent of
-specific application domains. \emph{ALBench} is publicly available under \emph{anonymous}. \\ [1mm]
+specific application domains. \emph{ALBench} is publicly available under \emph{anonymous} \todo{Github}. \\ [1mm]
 Our contributions include the following:
@@
 Each of these draws can be interpreted as a \textbf{reported result in AL literature} where the authors employed $\alpha$ restarts.
-%We provide evidence in Fig. \ref{fig:restarts} that previous works might not have evaluated their experiments with a sufficient number of restarts.
 Figure \ref{fig:restarts} shows the "true" mean performance of margin sampling (green) in relation to random sampling (black) and the oracle performance (red).
@@
 	\includegraphics[width=1.24\linewidth]{img/ablation_restarts}
-	\caption{Random draws from a pool of 100 runs for margin sampling on the Splice dataset with different numbers of repetitions ($\alpha=\{5,10,50\}$).
+	\caption{Random draws from a pool of 100 runs for margin sampling on the Splice dataset with different numbers of repetitions ($\alpha=\{3,5,50\}$).
 Green curves are the mean performance of all 100 runs, while the samples are blue.
-Even with 5 or 10 repetitions, we can observe that single draws for margin sampling display below-random performance (black), while the true mean should be above random.}
+Even with 3 or 5 repetitions, we can observe that single draws for margin sampling display below-random performance (black), while the true mean should be above random.}
 	\label{fig:restarts}
@@
 \bibliography{al_benchmark_paper.bib} 
+\begin{itemize}
+	\item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
+	\item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
+	\item Did you include the license to the code and datasets? \answerNA{}
+\end{itemize}
+
+\begin{enumerate}
+
+	\item For all authors...
+	\begin{enumerate}
+		\item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
+		\answerYes{See list of contributions in the Sec. \ref{sec:introduction}.}
+		\item Did you describe the limitations of your work?
+		\answerTODO{}
+		\item Did you discuss any potential negative societal impacts of your work?
+		\answerTODO{}
+		\item Have you read the ethics review guidelines and ensured that your paper conforms to them?
+		\answerTODO{}
+	\end{enumerate}
+
+	\item If you are including theoretical results...
+	\begin{enumerate}
+		\item Did you state the full set of assumptions of all theoretical results?
+		\answerTODO{}
+		\item Did you include complete proofs of all theoretical results?
+		\answerTODO{}
+	\end{enumerate}
+
+	\item If you ran experiments (e.g. for benchmarks)...
+	\begin{enumerate}
+		\item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
+		\answerTODO{}
+		\item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
+		\answerTODO{}
+		\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
+		\answerTODO{}
+		\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
+		\answerTODO{}
+	\end{enumerate}
+
+	\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
+	\begin{enumerate}
+		\item If your work uses existing assets, did you cite the creators?
+		\answerTODO{}
+		\item Did you mention the license of the assets?
+		\answerTODO{}
+		\item Did you include any new assets either in the supplemental material or as a URL?
+		\answerTODO{}
+		\item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
+		\answerTODO{}
+		\item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
+		\answerTODO{}
+	\end{enumerate}
+
+	\item If you used crowdsourcing or conducted research with human subjects...
+	\begin{enumerate}
+		\item Did you include the full text of instructions given to participants and screenshots, if applicable?
+		\answerTODO{}
+		\item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
+		\answerTODO{}
+		\item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
+		\answerTODO{}
+	\end{enumerate}
+
+\end{enumerate}
 
@@
 remains an open research question. For instance, we observe the cold start problem for BALD with
-Self-SL on CIFAR-100 (∼ 50% accuracy at 2k labeled samples for query sizes of 500 (low-label
+Self-SL on CIFAR-100 (~50\% accuracy at 2k labeled samples for query sizes of 500 (low-label
 regime) and 1k (mid-label regime). On the other hand, in the high label regime (budget of 5k and
-query size of 5k) ST and Self-SL models with similar accuracies of 50% and 60%, respectively,
+query size of 5k) ST and Self-SL models with similar accuracies of 50\% and 60\%, respectively,
 benefit from BALD. Since cold start problems are commonly associated with large query sizes, this
@@
 effect on the classifier performance as increasing the number of labeled training samples by a factor
-of ∼ 5 (e.g. our ST model reach approximately the same accuracy of ∼ 44% trained on 200 samples
+of ~5 (e.g. our ST model reach approximately the same accuracy of ~44\% trained on 200 samples
 compared to models in the AL study by Bengar et al. [6] trained on 1k samples). The effectiveness of
@@
 % Splice
-\begin{minipage}{0.65\linewidth}
 \begin{figure}[H]
     \centering
-    \includegraphics[width=\linewidth]{img/eval_splice}\\ [2mm]
-    \includegraphics[width=\linewidth]{img/micro_splice.jpg}
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-    &Splice \\
-    \hline
-    Oracle&0.811 $\pm$ 0.010\\
-    BALD&0.785 $\pm$ 0.013\\
-    Coreset&0.778 $\pm$ 0.014\\
-    Entropy&0.774 $\pm$ 0.016\\
-    Margin&0.773 $\pm$ 0.016\\
-    Badge&0.770 $\pm$ 0.016\\
-    Random&0.768 $\pm$ 0.014\\
-    TypiClust&0.766 $\pm$ 0.014\\
-\end{tabular}
-\end{minipage}
+    \caption{Performance curves per query size for normal (un-encoded) Splice}
+    \includegraphics[width=\linewidth]{img/eval_splice}
+\end{figure}
+%Splice Enc
+\begin{figure}[H]
+\centering
+\caption{Performance curves per query size for semi-supervised (encoded) Splice}
+\includegraphics[width=\linewidth]{img/eval_splice_enc}
+\end{figure}
 %DNA
-\begin{minipage}{0.65\linewidth}
 \begin{figure}[H]
     \centering
-	\includegraphics[width=\linewidth]{img/eval_dna} \\ [2mm]
-        \includegraphics[width=\linewidth]{img/micro_dna.jpg}
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&DNA \\
-\hline
-Oracle&0.842 $\pm$ 0.021\\
-Margin&0.807 $\pm$ 0.035\\
-Entropy&0.805 $\pm$ 0.038\\
-Coreset&0.796 $\pm$ 0.028\\
-Badge&0.789 $\pm$ 0.056\\
-TypiClust&0.788 $\pm$ 0.036\\
-Random&0.768 $\pm$ 0.024\\
-BALD&0.749 $\pm$ 0.044\\
-\end{tabular}
-\end{minipage}
+    \caption{Performance curves per query size for normal (un-encoded) DNA}
+	\includegraphics[width=\linewidth]{img/eval_dna}
+\end{figure}
+%DNA Enc
+\begin{figure}[H]
+\centering
+\caption{Performance curves per query size for semi-supervised (encoded) DNA}
+\includegraphics[width=\linewidth]{img/eval_dna_enc} 
+\end{figure}
 %USPS
-\begin{minipage}{0.65\linewidth}
 \begin{figure}[H]
     \centering
-	\includegraphics[width=\linewidth]{img/eval_usps} \\ [2mm]
-	\includegraphics[width=\linewidth]{img/micro_usps.jpg}
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&USPS\\
-\hline
-TypiClust&0.830 $\pm$ 0.007\\
-Oracle&0.823 $\pm$ 0.011\\
-Margin&0.809 $\pm$ 0.013\\
-Entropy&0.807 $\pm$ 0.013\\
-Badge&0.795 $\pm$ 0.018\\
-Coreset&0.788 $\pm$ 0.017\\
-Random&0.774 $\pm$ 0.012\\
-BALD&0.725 $\pm$ 0.050\\
-\end{tabular}
-\end{minipage}
-%Splice Enc
-\begin{minipage}{0.65\linewidth}
+    \caption{Performance curves per query size for normal (un-encoded) USPS}
+	\includegraphics[width=\linewidth]{img/eval_usps}
+\end{figure}
+%USPS Enc
 \begin{figure}[H]
     \centering
-    \includegraphics[width=\linewidth]{img/eval_splice_enc} \\ [2mm]
-    \includegraphics[width=\linewidth]{img/micro_splice_enc.jpg}
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&SpliceEncoded \\
-\hline
-Oracle&0.728 $\pm$ 0.022\\
-Coreset&0.648 $\pm$ 0.027\\
-TypiClust&0.645 $\pm$ 0.042\\
-Random&0.643 $\pm$ 0.036\\
-Entropy&0.636 $\pm$ 0.033\\
-Margin&0.636 $\pm$ 0.033\\
-Badge&0.627 $\pm$ 0.040\\
-BALD&0.565 $\pm$ 0.049\\
-\end{tabular}
-\end{minipage}
-%DNA Enc
-\begin{minipage}{0.65\linewidth}
+    \caption{Performance curves per query size for semi-supervised (encoded) USPS}
+    \includegraphics[width=\linewidth]{img/eval_usps_enc}
+\end{figure}
+%Cifar10
 \begin{figure}[H]
     \centering
-    \includegraphics[width=\linewidth]{img/eval_dna_enc}\\ [2mm]
-    \includegraphics[width=\linewidth]{img/micro_dna_enc.jpg} 
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&DNAEncoded\\
-\hline
-Oracle&0.709 $\pm$ 0.023\\
-TypiClust&0.672 $\pm$ 0.029\\
-Margin&0.648 $\pm$ 0.047\\
-Badge&0.647 $\pm$ 0.037\\
-Coreset&0.640 $\pm$ 0.041\\
-Entropy&0.629 $\pm$ 0.062\\
-Random&0.626 $\pm$ 0.035\\
-BALD&0.594 $\pm$ 0.039\\
-\end{tabular}
-\end{minipage}
-%USPS Enc
-\begin{minipage}{0.65\linewidth}
+    \caption{Performance curves per query size for normal (un-encoded) Cifar10}
+    \includegraphics[width=\linewidth]{img/eval_cifar10}
+\end{figure}
+%Cifar10 Encoded
 \begin{figure}[H]
 \centering
-    \includegraphics[width=\linewidth]{img/eval_usps_enc}\\ [2mm]
-    \includegraphics[width=\linewidth]{img/micro_usps_enc.jpg} 
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&USPSEncoded\\
-\hline
-Oracle&0.522 $\pm$ 0.021\\
-TypiClust&0.507 $\pm$ 0.025\\
-Margin&0.496 $\pm$ 0.030\\
-Random&0.468 $\pm$ 0.025\\
-Entropy&0.459 $\pm$ 0.021\\
-Badge&0.440 $\pm$ 0.026\\
-Coreset&0.435 $\pm$ 0.027\\
-BALD&0.402 $\pm$ 0.052\\
-\end{tabular}
-\end{minipage}
-%Cifar10
-\begin{minipage}{0.65\linewidth}
+\caption{Performance curves per query size for semi-supervised (encoded) Cifar10}
+\includegraphics[width=\linewidth]{img/eval_cifar10_enc}
+\end{figure}
+%FMnist
 \begin{figure}[H]
     \centering
-    \includegraphics[width=\linewidth]{img/eval_cifar10}\\ [2mm]
-    \includegraphics[width=\linewidth]{img/micro_cifar10.jpg}
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&Cifar10\\
-\hline
-Oracle&0.500 $\pm$ 0.010\\
-Badge&0.453 $\pm$ 0.012\\
-Coreset&0.453 $\pm$ 0.009\\
-Margin&0.451 $\pm$ 0.010\\
-Random&0.450 $\pm$ 0.012\\
-Entropy&0.449 $\pm$ 0.010\\
-TypiClust&0.436 $\pm$ 0.010\\
-BALD&0.436 $\pm$ 0.010\\
-\end{tabular}
-\end{minipage}
-%FMnist
-\begin{minipage}{0.65\linewidth}
+    \caption{Performance curves per query size for normal (un-encoded) FashionMnist}
+    \includegraphics[width=\linewidth]{img/eval_fmnist}
+\end{figure}
+%FashionMnist Encoded
 \begin{figure}[H]
     \centering
-    \includegraphics[width=\linewidth]{img/eval_fmnist}\\ [2mm]
-    \includegraphics[width=\linewidth]{img/micro_fmnist.jpg}
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&FashionMnist\\
-\hline
-Oracle&0.862 $\pm$ 0.003\\
-BALD&0.854 $\pm$ 0.003\\
-Margin&0.851 $\pm$ 0.003\\
-Badge&0.851 $\pm$ 0.003\\
-Entropy&0.847 $\pm$ 0.004\\
-TypiClust&0.846 $\pm$ 0.004\\
-Coreset&0.840 $\pm$ 0.004\\
-Random&0.837 $\pm$ 0.004\\
-\end{tabular}
-\end{minipage}
-%Cifar10 Encoded
-\begin{minipage}{0.65\linewidth}
+    \caption{Performance curves per query size for semi-supervised (encoded) FashionMnist}
+    \includegraphics[width=\linewidth]{img/eval_fmnist_enc}
+\end{figure}
+%TopV2
 \begin{figure}[H]
     \centering
-    \includegraphics[width=\linewidth]{img/eval_cifar10_enc} \\ [2mm]
-    \includegraphics[width=\linewidth]{img/micro_cifar10_enc.jpg} 
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&Cifar10Encoded\\
-\hline
-Oracle&0.714 $\pm$ 0.007\\
-Entropy&0.654 $\pm$ 0.013\\
-Coreset&0.653 $\pm$ 0.012\\
-Badge&0.653 $\pm$ 0.012\\
-BALD&0.650 $\pm$ 0.016\\
-Margin&0.647 $\pm$ 0.012\\
-TypiClust&0.636 $\pm$ 0.009\\
-Random&0.607 $\pm$ 0.013\\
-\end{tabular}
-\end{minipage}
-%FashionMnist Encoded
-\begin{minipage}{0.65\linewidth}
+    \caption{Performance curves per query size for normal (GloVe) TopV2}
+\includegraphics[width=\linewidth]{img/eval_topv2}
+\end{figure}
+%News
 \begin{figure}[H]
     \centering
-    \includegraphics[width=\linewidth]{img/eval_fmnist_enc} \\ [2mm]
-\includegraphics[width=\linewidth]{img/micro_fmnist_enc.jpg} 
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&FashionMnistEncoded\\
-\hline
-Oracle&0.732 $\pm$ 0.006\\
-Coreset&0.686 $\pm$ 0.008\\
-Badge&0.685 $\pm$ 0.008\\
-Entropy&0.684 $\pm$ 0.009\\
-Margin&0.682 $\pm$ 0.011\\
-BALD&0.668 $\pm$ 0.009\\
-TypiClust&0.652 $\pm$ 0.009\\
-Random&0.640 $\pm$ 0.011\\
-\end{tabular}
-\end{minipage}
-%TopV2
-\begin{minipage}{0.65\linewidth}
+    \caption{Performance curves per query size for normal (GloVe) News}
+\includegraphics[width=\linewidth]{img/eval_news}
+\end{figure}
+%ThreeClust
 \begin{figure}[H]
     \centering
-\includegraphics[width=\linewidth]{img/eval_topv2} \\[2mm]
-\includegraphics[width=\linewidth]{img/micro_topv2.jpg}
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&TopV2\\
-\hline
-Oracle&0.862 $\pm$ 0.006\\
-BALD&0.831 $\pm$ 0.013\\
-Badge&0.826 $\pm$ 0.015\\
-Coreset&0.823 $\pm$ 0.016\\
-Margin&0.822 $\pm$ 0.015\\
-TypiClust&0.805 $\pm$ 0.015\\
-Entropy&0.801 $\pm$ 0.025\\
-Random&0.787 $\pm$ 0.015\\
-\end{tabular}
-\end{minipage}
-%News
-\begin{minipage}{0.65\linewidth}
+    \caption{Performance curves per query size for normal (un-encoded) Honeypot}
+\includegraphics[width=\linewidth]{img/eval_honeypot}
+\end{figure}
+%DivergingSin
 \begin{figure}[H]
     \centering
-\includegraphics[width=\linewidth]{img/eval_news} \\[2mm]
-\includegraphics[width=\linewidth]{img/micro_news.jpg}
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&News\\
-\hline
-Oracle&0.502 $\pm$ 0.005\\
-Margin&0.427 $\pm$ 0.007\\
-BALD&0.421 $\pm$ 0.008\\
-Badge&0.420 $\pm$ 0.011\\
-Entropy&0.416 $\pm$ 0.010\\
-Coreset&0.415 $\pm$ 0.011\\
-Random&0.409 $\pm$ 0.008\\
-TypiClust&0.385 $\pm$ 0.010\\
-\end{tabular}
-\end{minipage}
-%ThreeClust
-\begin{minipage}{0.65\linewidth}
-\begin{figure}[H]
-    \centering
-\includegraphics[width=\linewidth]{img/eval_threeclust.pdf} \\[2mm]
-\includegraphics[width=\linewidth]{img/micro_threeclust.jpg}
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&ThreeClust \\
-\hline
-Oracle&0.722 $\pm$ 0.097\\
-Coreset&0.698 $\pm$ 0.058\\
-TypiClust&0.697 $\pm$ 0.055\\
-Entropy&0.682 $\pm$ 0.098\\
-Random&0.672 $\pm$ 0.067\\
-Margin&0.669 $\pm$ 0.095\\
-Badge&0.524 $\pm$ 0.086\\
-BALD&0.507 $\pm$ 0.050\\
-\end{tabular}
-\end{minipage}
-%DivergingSin
-\begin{minipage}{0.65\linewidth}
-\begin{figure}[H]
-    \centering
-\includegraphics[width=\linewidth]{img/eval_divergingsin.pdf} \\[2mm]
-\includegraphics[width=\linewidth]{img/micro_divsin.jpg}
-\end{figure}
-\end{minipage}
-\begin{minipage}{0.29\linewidth}
-\begin{tabular}{c|c}
-&DivergingSin\\
-\hline
-Oracle&0.948 $\pm$ 0.198\\
-Entropy&0.936 $\pm$ 0.202\\
-TypiClust&0.930 $\pm$ 0.196\\
-Margin&0.929 $\pm$ 0.201\\
-Random&0.919 $\pm$ 0.191\\
-Badge&0.914 $\pm$ 0.202\\
-Coreset&0.914 $\pm$ 0.197\\
-BALD&0.661 $\pm$ 0.167\\
-\end{tabular}
-\end{minipage}
+    \caption{Performance curves per query size for normal (un-encoded) Diverging Sine}
+\includegraphics[width=\linewidth]{img/eval_diverging_sine}
+\end{figure}
 

### 2024-06-04 core/helper_functions.py

@@ -263,8 +263,10 @@ def _create_plot_for_query_size(ax, dataset, query_size, y_label, title, smoothi
         ax.set_ylim(*y_lim)
     ax.grid(True)
 
-def full_plot(dataset, query_size=None, show_auc=True, radjust=0.8, forecast_oracle=True,
-              smoothing_weight=0.0, y_label="Accuracy", y_lim=None):
+def full_plot(dataset:str, query_size=None, show_auc=True, radjust=0.8, forecast_oracle=True,
+              smoothing_weight=0.0, y_label="Accuracy", y_lim=None, plot_title=None):
+    if plot_title is None:
+        plot_title = dataset
     if query_size is None:
         base_path = os.path.join("runs", dataset)
         query_size = list(os.listdir(base_path))
@@ -281,7 +283,7 @@ def full_plot(dataset, query_size=None, show_auc=True, radjust=0.8, forecast_ora
         legend_created = False
         for i in range(len(query_size)):
             _create_plot_for_query_size(ax[i], dataset, query_size[i], y_label if i%2==0 else "",
-                                        f"{dataset} - {query_size[i]}", smoothing_weight, show_auc,
+                                        f"{plot_title} - {query_size[i]}", smoothing_weight, show_auc,
                                         forecast_oracle, y_lim)
             if not legend_created:
                 fig.legend(loc=7)

### 2024-06-04 doc/al_benchmark_paper.tex
Notes on Orcale

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A Greedy Oracle Algorithm}\label{sec:oracle}
-An oracle method for AL tries to find the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure that induces the highest AUC score for a given budget.
-However, due to the combinatorial nature of the problem, it is computationally infeasible for realistic datasets.
-Hence, previous works have proposed approximations to this oracle sequence. 
-\cite{zhou2021towards} used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
-Even though their reported performance curves display a significant lift over all other acquisition functions, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
-In this paper, we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
-Our oracle algorithm evaluates every data point $u_k = \operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$ in a subsample of unlabeled points by recovering the label, fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test performance.
-The data point that incurs the largest lift in test performance is selected and added to the labeled pool for that iteration.
+\section{A Greedy Oracle Algorithm}\label{sec:oracle} An oracle method for AL
+tries to find the oracle set $\mathcal{O}_b$ for a given dataset,\todo{Versuchen
+die Beste zu finden tut doch jede AL-Methode! Vlt sowas wie: Ist ein Verfahren,
+dass zwar die Optimale Lösung findet, aber dabei keine Rücksicht auf Laufzeiten nimmt} model, and
+training procedure that induces the probably highest AUC score for a given budget.
+However, due to the combinatorial nature of the problem, it is computationally
+infeasible for realistic datasets. Hence, previous works have proposed
+approximations to this oracle sequence. \cite{zhou2021towards} used simulated
+annealing to search for the optimal subset and used the best  solution found
+after a fixed time budget. \todo{Okay und warum ist das dann keine Aquisition funnktion}Even though their reported performance curves display
+a significant lift over all other acquisition functions, we found the
+computational cost of reproducing this oracle for all our datasets to be
+prohibitive (The authors reported the search to take several days per dataset on
+8 V100 GPUs). In this paper, we propose a greedy oracle algorithm that
+constructs an approximation of the optimal set in an iterative fashion. Our
+oracle algorithm evaluates every data point $u_k =
+\operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$\todo[inline]{Die Notation
+stimmt so nicht, willst du sagen, dass du hier ein Subset der größe $\tau$
+gesamplet hast uniform gesamplet hast? Dann schreibe das: "Our oracle algorithm
+uniformly samples at iteration $i$ a subset $\mathcal{U}_S$ of the not already
+labeled data points $\mathcal{U}^{(i)}$.
+Then it recovers the label $y$ for each of the sampled $u \in \mathcal{U}_S$ and selects the
+point $u \in \mathcal{U}_S$ for which the classifier $\hat{y}$ trained on
+$\mathcal{L}^{(i)} \cup \{u\}$ has the maximal performance. } in a subsample of
+unlabeled points by recovering the label, fitting the classifier $\hat y$ on
+$\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test
+performance. \todo{Okay, hier verstehe ich was ein Orcale ist. Eine Aquisition
+funktion die schon das wissen über die Labels beim aquirieren nutzen kann?
+Stimmt das?}The data point that incurs the largest lift in test performance is
+selected and added to the labeled pool for that iteration.
 \todo{[Thorben] Ich habe die Erklärung für den Fallback im Oracle neu gemacht. Ist das verständlich?}
-Due to the algorithms greedy nature (just considering the next point to pick), our oracle frequently encounters situations where every point in $u_k$ would incur a negative lift (worsening the test performance).
-This can happen, for example, if the oracle picked a labeled set that enables the classifier to correctly classify a big portion of easy samples in the test set, but now fails to find the next \textbf{single} unlabeled point that would enable the classifier to succeed on one of the hard samples.
-This leads to a situation, where no point can immediately incur an increase in test performance and therefore the selected data point can be considered random.
-To circumvent this problem, we use our best-performing acquisition function (margin sampling \cite{wang2014new}) as a fallback option for the oracle.
-Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
-The resulting greedy algorithm constructs an approximation of the optimal labeled set that consistently outperforms all other algorithms by a significant margin, while requiring relatively low computational cost ($\mathcal{O}(B\tau)$).
+Due to the algorithms greedy nature (just considering the next point to pick),
+our oracle frequently encounters situations where every point in $u_k$ would
+incur a negative lift (worsening the test performance). This can happen, for
+example, if the oracle picked a labeled set that enables the classifier to
+correctly classify a big portion of easy samples in the test set, but now fails
+to find the next \textbf{single} unlabeled point that would enable the
+classifier to succeed on one of the hard samples. This leads to a situation,
+where no point can immediately incur an increase in test performance and
+therefore the selected data point can be considered random. To circumvent this
+problem, we use our best-performing acquisition function (margin sampling
+\cite{wang2014new}) as a fallback option for the oracle. Whenever the oracle
+does not find an unlabeled point that results in an increase in performance, it
+defaults to margin sampling in that iteration. The resulting greedy algorithm
+constructs an approximation of the optimal labeled set that consistently
+outperforms all other algorithms by a significant margin, while requiring
+relatively low computational cost ($\mathcal{O}(B\tau)$).
 \todo{Da wir den Fallback haben, sollten wir das Oracle vielleicht einfach als Erweiterung von MarginSampling verkaufen?}
-We fix $\tau = 20$ in this work, as this gave us an average lift of 5\% over the best performing acquisition function per dataset (which is significant for AL settings) and we expect diminishing returns for larger $\tau$.
-The pseudocode for our oracle can be found in App. \ref{app:pseudocode}.
+We fix $\tau = 20$ in this work, as this gave us an average lift of 5\% over the
+best performing acquisition function per dataset (which is significant for AL
+settings) and we expect diminishing returns for larger $\tau$. The pseudocode
+for our oracle can be found in App. \ref{app:pseudocode}.
 %In the algorithm $\operatorname{Train}(\LL, \hat y_\theta)$ trains the classification model $\hat y_\theta$ on $\LL$. \\
 %Alg. \ref{alg:oracle} replaces the acquisition function $\Omega$ in the AL loop (Appendix \ref{app:pseudocode} Alg. \ref{alg:active_learning}).
-Even though our proposed algorithm is more efficient than other approaches, the computational costs for high budget datasets like Cifar10 and FashionMnist meant that we could not compute the oracle for all 10000 datapoints.
-To still provide an oracle for these two datasets, we select two points per iteration instead of one and stop the oracle computation at a budget of 2000.
-The rest of the curve is forecast with a simple linear regression that asymptotically approaches the upper bound performance of the dataset. 
-A detailed description can be found in App. \ref{app:oracle_forecasting}.
+Even though our proposed algorithm is more efficient than other approaches, the
+computational costs for high budget datasets like Cifar10 and FashionMnist meant
+that we could not compute the oracle for all 10000 datapoints. To still provide
+an oracle for these two datasets, we select two points per iteration instead of
+one and stop the oracle computation at a budget of 2000. The rest of the curve
+is forecast with a simple linear regression that asymptotically approaches the
+upper bound performance of the dataset. A detailed description can be found in
+App. \ref{app:oracle_forecasting}.\todo[inline]{Approached heißt konvergiert,,
+oder? Dann muss das doch nicht stimmen. Ich würde nach "linear regression" einen
+Punkt machen und den Teil danach löschen.}
 

### 2024-06-05 doc/al_benchmark_paper.tex
paper version 1.0

@@
 
-% Introduction:
-% What is the problem?
-% - Ergebnisse schwierig zu reproduzieren
-% - Vorherige Benchmarks haben dem Versuchsaufbau zu wenig Beachtung geschenkt - erst Ji und Luth haben Methoden für verlässliche AL Experimente beschrieben
-% - keiner (!) der Benchmarks hat beide unserer Top-2 Methoden (Margin und Badge)
-% - Es werden maximal zwei Domänen abgedeckt
-% - Bisherige Oracle-Algorithmen sind extrem rechenaufwenig
-% What are we adding to it?
-% - Wir implementieren die Guidelines von Ji et al
-% - Wir erweitern die Anzahl der Domänen von 1 auf 5 und führen mehr Wiederholungen ein
-% - Bessere Abdeckung der relevanten Algorithmen (von 8 (Ji) auf 11)
-% - Unser Oracle ist viel leichter zu berechnen
-% - Wir erhöhen die Anzahl der Experimente exponentiell: Ji und Lueth: 300-400; wir: 24200+Oracle; Größtes Benchmark bis jetzt ist Zhan (2021): ~31500 (nur tabular)
-% 3 runs vs 50 runs -> statistik
-% Why is that important?
-% - Wir zeigen, dass die Wiederholungen wichtig sind, um reproduzierbare Ergebnisse zu liefern
-% - Wir zeigen, dass es signifikante Unterschiede zwischen den Domänen gibt
-% - Wir zeigen, dass Margin und Badge zwei starke Baselines sind, die immer dabei sein müssen
-% - Unser Oracle hat x% lift über alle Algorithmen und ist somit ausreichend
-
 % TODOs nach Priorität
-% Terminologie aufräumen:
-% - Query size vs batch size
-% - Acq. function vs. AL Alg. vs. AL Method
-% - (Un-)encoded vs. semi-supervised/normal
-% - ThreeClust vs. Honeypot
-% Gibt es eine Conclusion für die Encoded-Datasets? Im Moment werden die nicht besprochen
-% Methode zum CD-Diagramm genau erklären (paired-t-test etc.) und die Relevanz fürs paper zeigen
 % Advantage-Matrix von Ji et. al. neu aufarbeiten. Wo sind die Unterschiede? Ist das CD-Diagramm besser/schlechter?
-% (Un-)Encoded durch Normal uns Semi-Supervised ersetzen, um die Verbindung klar zu machen
 % Eventuell Batchsize 1 von den anderen Ergebnissen separieren. Der Vergleich macht mehr Sinn, als einfach die Batchsizes zu mitteln
@@
 % Unsere Rankings mit Literatur vergleichen. Wie anders sind wir? -> Evtl. besten Alg. pro Benchmark sammeln und vergleichen
-% Causalitäten checken: Viele Dinge werden einfach genannt, ohne das der Grund richtig eingeführt wird
 % Fig. 1: Schlechte Performance am Anfang verwirrt nur. Entweder BADGE benutzen, anders plotten, oder im Text in Kontext setzen
 % Oracle Forecast im Appendix erklären
-
-%%%%%%%%%%%%%%%%%%
-% DONE
-%%%%%%%%%%%%%%%%%
-% !! Ergebnis-Tabelle mit nur 3 runs neu berechnen und Unterschiede zeigen
-% !! Least-Confident Baseline
-% Conclusion: Es gibt Unterschiede bei der Top 1 zwischen den Domänen, aber die Top 3 ist konstant
-% !! Sec 3 Related Work sollte Ji et al. im Detail einführen, weil wir uns an deren Guidelines halten. Alles andere sollte dort gekürzt werden
-% !! "Overfitting on test" für Oracle: Sanity-Check durchführen und eventuell komplett streichen
-% !! Introduction neu schreiben: Wir brauchen eine klare Benennung der Ziele und bessere Abgrenzung zu anderen papern 
-% Infos und Ergebnisse für die syntetischen Datensätze in ein Kapitel bündeln und direkt auf die Contribution beziehen
-% !! Sec 4 Methodology ist sehr chaotisch. Wir müssen neu sortieren und unwichtige Dinge streichen
-% Contributions neu sortieren und ausführlicher beschreiben
-% Oracle Performance mit einer Zahl benennen: z.B. Da wird ~20% lift zu allen Baselines haben, brauchen wir kein aufwendigeres Oracle
-% "reproducible" als Aufhänger für 50 repeats ist nicht intuitiv
-% Eventuell Honeypot und Diverging Sine direkt in den Contributions nennen
-% Tabelle mit AUCs in den Appendix
-
-%%%%%%%%%%%%%%%%
-% Anzahl der Runs (Ohne Oracle und Upper Bound)
-% Runs pro query size: 11alg * 50runs = 550
-% Cifar10   x2 = 1100
-% Cfr10Enc  x4 = 2200
-% DivSin    x2 = 1100
-% DNA       x4 = 2200
-% DNAEnc    x2 = 1100
-% FMnist    x2 = 1100
-% FMnistEnc x4 = 2200
-% News      x3 = 1650
-% Splice    x4 = 2200
-% SpliceEnc x3 = 1650
-% Honeypot  x2 = 1100
-% TopV2     x4 = 2200
-% USPS      x4 = 2200
-% USPSEnc   x4 = 2200
-% = 24200 runs
-% Lueth: ~315
-% Ji: ~432
-% Zhan: ~31500 (nur tabular data)
 
@@
 that only a small number of repetitions of experiments are conducted. To overcome
-this obstacles, we propose \emph{ALBench}, the first active learning benchmark
+these obstacles, we propose \emph{CDALBench}, the first active learning benchmark
 which includes tasks in computer vision, natural language processing and tabular
-learning. Furthermore, by providing an efficient greedy oracle, \emph{ALBench}
+learning. Furthermore, by providing an efficient, greedy oracle, \emph{CDALBench}
 can be evaluated with 50 runs for each experiment. We show, that both the
-cross-domain character and the large amount of repetitions are crucial for
+cross-domain character and a large amount of repetitions are crucial for
 sophisticated evaluation of AL research. Concretely, we show that the
@@
 important to evaluate Active Learning with a cross-domain benchmark.
-Additionally, we show, that having a large amount of runs is crucial. With only
+Additionally, we show that having a large amount of runs is crucial. With only
 conducting five runs as often done in the literature, the superiority of
@@
 that only these samples need to be labeled. It has been shown that a small
-labeled set of this nature can be used to train well-performing models
+labeled set of this nature can be used to train well-performing models.
 %\cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}.
@@
 In the last decade, many different algorithms for AL have been proposed and
-almost every method has reported lifts over all its predecessors, \footnote{Out
+almost every method has reported lifts over all its predecessors. \footnote{Out
 of all considered algorithms for this paper, only BALD \cite{gal2017deep} did
-not claim a new SOTA performance in their result section.} However, real
+not claim a new state-of-the-art (SOTA) performance in their result section.} However, real
 insights into the current state of AL are hard to draw from these works, due to
@@
 datasets and baseline approaches.
-\item Due to computational constraints, a lot of works do only a small amount of experimental
+\item Due to computational constraints, a lot of works perform only a small amount of experimental
 runs, hence it is questionable wether the superiority of a specific approach can
@@
 our knowledge, all of them are either limited in the domains they consider or do
-not do enough runs for conclusive results 
-%\footnote{See
-% \Cref{tab:benchmark_comparison} for a comparison of existing benchmarks}. One
-% reason for that, may be that in AL research it is common to additionally include
-% a so-called \emph{oracle} baseline, i.e., a time-costly search-based approach
-% which gives a lower bound for the best possible solution. The need for an oracle
-% makes it challenging to use multiple datasets and enough runs for the
-% experimental evaluation. \todo{\@ Thorben: Ist das was ich hier zum Oracle
-% schreibe richtg? Oder ist das Quatsch?} 
-Hence, the current
-state of the art in Active Learning is still not well-understood and principled
-shortcomings of different algorithms and wether they are 
-domain-independent, are currently not identified.
-
-Here we step in with \emph{ALBench}, an active learning benchmark which covers
+not contain enough runs to generate conclusive results.
+Hence, the current SOTA in AL is still not well-understood and principled
+shortcomings of different algorithms and wether they are domain-independent are currently not identified.
+
+Here we step in with \emph{CDALBench}, an AL benchmark which covers
 multiple application domains and reports a large amount of runs per
 experiment, so that the significance of performance differences can can be
-estimated. Specifically, \emph{ALBench} consists of datasets from
+estimated. Specifically, \emph{CDALBench} consists of datasets from
 computer vision, natural language processing and the tabular domain. We provide
@@
 challenges for AL methods.
-
-The applied evaluation protocol in \emph{ALBench} uses 50 runs for
-each experiment. By having such a large amount of runs, we can evaluate the
-significance of performance gaps and identify the best performing
-approaches for each dataset as well as whole domains. 
-Furthermore, we show that the small amount of runs other works do in fact 
-produce to misleading results. 
-To be more specific, we show that if only 3 restarts are employed for each experiment, 
-the performance of specific methods strongly varies. As we will see, even the ranking of
-the different methods averaged over many datasets fluctuates with the specific set of runs. 
-This effect is so strong, that, depending on the seed, even a well-established method's performance can be significantly better and significantly worse than random for the same dataset.
-
-To enable the computation of an oracle performance for a protocol with large amounts of restarts
-we propose a \emph{greedy oracle function} 
-which uses only a small amount of search steps to estimate the
-optimal solution. While being more time-efficient than established oracle
-functions, it possibly underestimates the real upper bound performance. 
-However, as our experiments will show, it is still outperforming all current AL methods
-by at least 5\% and thus is suitable as an upper bound.
-
-Our experimental evaluation shows, that there exists no clear SOTA
-method for Active Learning. The superiority of methods is strongly dataset- and
-domain-dependent with the outstanding observation, that the image domain works
-fundamentally different than the tabular and text domain. Here, the best
-performing approach for text and tabular data, namely \emph{margin sampling}, is
-significantly outperformed by \emph{least confident sampling}, which does not belong to
-the top performing approaches in any other domain. Thus, using the performance
-of Active Learning approaches on the image domain as a proxy of Active Learning
-in general, as it is often done \cite{beck2021effective, munjal2022towards, li2022empirical, ji2023randomness, luth2024navigating}, is questionable. To further analyze
-performance of common methods, we propose \emph{Honeypot} and \emph{Diverging Sin}, two
-synthetic datasets, designed to be challenging for naive decision-boundary- and clustering-based approaches respectively. Hence, they provide insights in principled shortcomings of AL methods.
-
-To sum up, \emph{ALBench} is an experimental framework which includes an
-efficient oracle approximation, multiple application domains, enough repetitions to
-draw valid conclusions and two synthetic tasks to highlight shortcomings of AL methods. 
-By being the first benchmark to proving these things in one code-base, 
-we believe that \emph{ALBench} is a major step forward of assessing the overall state of Active Learning research, independent of
-specific application domains. \emph{ALBench} is publicly available under \emph{anonymous} \todo{Github}. \\ [1mm]
-Our contributions include the following:
-%
-\begin{enumerate}
-\item\label{cont:repetitions} We show that the small of repetitions that previous works have employed is not
-sufficient for meaningful conclusions. Sometimes even making it impossible to assess if a performance is above or below random.
-\item\label{cont:oracle} We propose an efficient and performant oracle which is constructed iteratively in a greedy fashion,
-overcoming major computational hurdles.
-% one of the major obstacles for sufficient repetitions of AL experiments. 
-% [Thorben] Das können wir so nicht sagen, glaube ich. Niemand verlangt von uns, dass wir das Oracle auch 50 mal wiederholen
-\item\label{cont:domains} We propose \emph{ALBench}, the first general benchmark providing tasks
-in the domains of images, text and tabular learning. It further contains
-synthetic and pre-encoded data to allow for a sophisticated evaluation of AL
-methods. Our experiments show, that there is no clear state-of-the art method
-for active learning across different domains.
-\item\label{cont:synthdata} We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
-designed to hinder active learning by naive decision-boundary- or
-clustering-based approaches respectively. Thus, they provide an important tool to identify 
-shortcomings of existing AL methods.
-\end{enumerate}
-%
 \begin{table}
@@
 	\caption{Comparison of our benchmark with the existing literature. Oracle curves
-	serve as an approximation of the best possible AL algorithm. Including the
-	encoded versions of our datasets we reach 14 datasets, without we have 9. "Semi"
+		serve as an approximation of the best possible AL algorithm. Our benchmark contains 9 
+		datasets (14 including the encoded versions). ``Semi"
 		indicates whether the paper is employing any form of self- or semi-supervised
-	learning. A "-" for repetitions means that we could not determine how often each
-	experiment is repeated in the respective framework. \emph{ALBench} is the only
-	benchmark which reports over enough runs for meaningful results and considers
+		learning. A ``-" for repetitions means that we could not determine how often each
+		experiment was repeated in the respective framework. \emph{CDALBench} is the only
+		benchmark which evaluates a high number of runs and considers
 		all 5 domains.}
@@
 \end{table}
+The applied evaluation protocol in \emph{CDALBench} uses 50 runs for
+each experiment. By having such a large amount of runs, we can evaluate the
+significance of performance gaps and identify the best performing
+approaches for each dataset as well as whole domains. 
+Furthermore, we show that the small amount of runs other works do, in fact, 
+produce misleading results. 
+To be more specific, we show that if only 3 restarts are employed for each experiment, 
+the performance of specific methods strongly varies. As we will see, even the ranking of
+the different methods averaged over many datasets fluctuates with the specific set of runs. 
+This effect is so strong, that, depending on the seed, even a well-established method's performance can be significantly better and significantly worse than random for the same dataset.
+
+To enable the computation of an oracle performance for a protocol with large amounts of restarts,
+we propose a \emph{greedy oracle algorithm} 
+which uses only a small amount of search steps to estimate the
+optimal solution. While being more time-efficient than established oracle
+functions, it possibly underestimates the real upper bound performance. 
+However, as our experiments will show, it is still outperforming all current AL methods
+by at least 5\% and thus is suitable as an upper bound.
+
+Our experimental evaluation shows that there exists no clear SOTA
+method for AL. The superiority of methods is strongly dataset- and
+domain-dependent with the outstanding observation, that the image domain works
+fundamentally different than the tabular and text domain. Here, the best
+performing approach for text and tabular data, namely \emph{margin sampling}, is
+significantly outperformed by \emph{least confident sampling}, which does not belong to
+the top performing approaches in any other domain. Thus, using the performance
+of AL approaches on the image domain as a proxy of AL
+in general, as it is often done \cite{beck2021effective, munjal2022towards, li2022empirical, ji2023randomness, luth2024navigating}, is questionable. To further analyze
+performance of common methods, we propose \emph{Honeypot} and \emph{Diverging Sine}, two
+synthetic datasets, designed to be challenging for naive decision-boundary- and clustering-based approaches respectively. Hence, they provide insights in principled shortcomings of AL methods.
+
+In summary, \emph{CDALBench} is an experimental framework which includes an
+efficient oracle approximation, multiple application domains, enough repetitions to
+draw valid conclusions and two synthetic tasks to highlight shortcomings of AL methods. 
+By being the first benchmark to providing these points in one code-base, 
+we believe that \emph{CDALBench} is a major step forward of assessing the overall state of AL research, independent of
+specific application domains. \emph{CDALBench} is publicly available under \url{https://github.com/wernerth94/A-Cross-Domain-Benchmark-for-Active-Learning/}. \\ [1mm]
+Our contributions include the following:
+%
+\begin{enumerate}
+\item\label{cont:repetitions} We show that the small number of repetitions that previous works have employed is not
+sufficient for meaningful conclusions. Sometimes even making it impossible to assess if a performance is above or below random.
+\item\label{cont:oracle} We propose an efficient and performant oracle which is constructed iteratively in a greedy fashion,
+overcoming major computational hurdles.
+% one of the major obstacles for sufficient repetitions of AL experiments. 
+% [Thorben] Das können wir so nicht sagen, glaube ich. Niemand verlangt von uns, dass wir das Oracle auch 50 mal wiederholen
+\item\label{cont:domains} We propose \emph{CDALBench}, the first general benchmark providing tasks
+in the domains of image, text and tabular learning. It further contains
+synthetic and pre-encoded data to allow for a sophisticated evaluation of AL
+methods. Our experiments show, that there is no clear SOTA method
+for AL across different domains.
+\item\label{cont:synthdata} We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
+designed to hinder AL by naive decision-boundary- or
+clustering-based approaches respectively. Thus, they provide an important tool to identify 
+shortcomings of existing AL methods.
+\end{enumerate}
 
@@
 We call $x \in \mathcal{X}$, $y \in \mathcal{Y}$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$. 
-Find an acquisition function $\Omega: \U^{(i)},\LL^{(i)} \mapsto x^{(i)} \in \U^{(i)}$ that iteratively selects the next unlabeled point $x^{(i)}$ for labeling
+Find an AL method $\Omega: \U^{(i)},\LL^{(i)} \mapsto x^{(i)} \in \U^{(i)}$ that iteratively selects the next unlabeled point $x^{(i)}$ for labeling
 \begin{align*}
@@
 \end{align*}
-with $\U^{(0)} = \operatorname{seed}(\U, s)$ and $\LL^{(0)} = \left(\U^{(0)}_i, A(\U^{(0)}_i)\right) \hspace{1mm} i \in [1, \ldots, s]$, where $\operatorname{seed}(\U, s)$ selects $s$ points per class for the initial labeled set.
+with $\U^{(0)} = \operatorname{seed}(\U, s)$ and $\LL^{(0)} = \left(\U^{(0)}_i, A(\U^{(0)}_i)\right) \hspace{1mm} i \in [1, \ldots, s]$, where $\operatorname{seed}(\U, s)$ selects $s$ points per class for the initial labeled set $\LL^{(0)}$.
 \\
@@
 \section{Related Work}
-While multiple benchmark suites have been proposed for Active Learning, none of them provide experiments for more than two domains.
+While multiple benchmark suites have been proposed for AL, none of them provide experiments for more than two domains.
 The authors of \cite{beck2021effective}, \cite{munjal2022towards}, \cite{li2022empirical}, \cite{ji2023randomness} and \cite{luth2024navigating} even focus exclusively on the image domain.
@@
 %Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
-The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also place a much higher emphasis on the problem of evaluating AL algorithms under high variance than their predecessors (indicated in Tab.
+The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also place a much higher emphasis on the problem of evaluating AL methods under high variance than their predecessors (indicated in Tab.
 \ref{tab:benchmark_comparison} by a dashed line).
-The authors of \cite{ji2023randomness} posed a total of 12 "recommendations" for reliable evaluation of AL algorithms.
-We largely adapt the proposed recommendations and extend their work to multiple domains and batch sizes.
+The authors of \cite{ji2023randomness} posed a total of 11 ``recommendations" for reliable evaluation of AL methods.
+We largely adapt the proposed recommendations and extend their work to multiple domains and query sizes.
 For a complete list of the recommendations and our implementation of them, please refer to App. \ref{app:recommendations}.
-This work also pays attention to the so-called "pitfalls" of AL evaluation proposed in \cite{luth2024navigating}.
+This work also pays attention to the so-called ``pitfalls" of AL evaluation proposed in \cite{luth2024navigating}.
 For a complete list of the pitfalls and our considerations regarding them, please refer to App. \ref{app:pitfalls}.
-To the best of our knowledge, we are the first to extend reliable SOTA (based on \cite{ji2023randomness, luth2024navigating}) experimentation to a total of 5 data domains and a high number of restarts per experiment.
+To the best of our knowledge, we are the first to extend reliable SOTA (based on \cite{ji2023randomness, luth2024navigating}) experimentation to a total of 5 data domains and a high number of repetitions per experiment.
 
@@
 \section{Few Repetitions are not Sufficient for Meaningful Results}\label{sec:restarts}
-To evaluate how many restarts are necessary to obtain conclusive results in an AL experiment, we computed 100 runs of our top-performing algorithm on one dataset.
-Our best algorithm is margin sampling and we chose the Splice dataset for its average size and complexity. \\
-This allows us firstly, to obtain a very strong estimation of the "true" average performance of margin sampling on this dataset and secondly, to draw subsets from this pool of 100 runs.
-Setting the size of our draws to $\alpha$ and sampling uniformly, we can approximate a cross-validation process with $\alpha$ restarts.
-Each of these draws can be interpreted as a \textbf{reported result in AL literature} where the authors employed $\alpha$ restarts.
-Figure \ref{fig:restarts} shows the "true" mean performance of margin sampling (green) in relation to random sampling (black) and the oracle performance (red).
+To evaluate how many repetitions are necessary to obtain conclusive results in an AL experiment, we computed 100 runs of our top-performing AL method on one dataset.
+Our best method is margin sampling and we chose the Splice dataset for its average size and complexity. \\
+This allows us firstly, to obtain a very strong estimation of the ``true" average performance of margin sampling on this dataset and secondly, to draw subsets from this pool of 100 runs.
+Setting the size of our draws to $\alpha$ and sampling uniformly, we can approximate a cross-validation process with $\alpha$ repetitions.
+Each of these draws can be interpreted as a \textbf{reported result in AL literature} where the authors employed $\alpha$ repetitions.
+Figure \ref{fig:restarts} shows the ``true" mean performance of margin sampling (green) in relation to random sampling (black) and the oracle performance (red).
 We display 5 random draws of size $\alpha$ in blue.
-We can observe that even for a relatively high number of restarts the variance between the samples is extremely high, resulting in some performance curves being worse that random and some being significantly better.
+We can observe that even for a relatively high number of repetitions the variance between the samples is extremely high, resulting in some performance curves being worse than random and some being significantly better.
 When setting $\alpha = 50$ we observe all samples to converge close to the true mean performance. 
 \begin{figure}
-        \hspace{-20mm}
-	\includegraphics[width=1.24\linewidth]{img/ablation_restarts}
+	\centering
 	\caption{Random draws from a pool of 100 runs for margin sampling on the Splice dataset with different numbers of repetitions ($\alpha=\{3,5,50\}$).
@@
 	\label{fig:restarts}
-\end{figure}
-In addition to this motivating example, we carried out our main evaluation (Tab. \ref{tab:results}) multiple times by uniformly sampling 3 random from our available runs and comparing the results.
-We found significant differences in the performance of acquisition functions on individual datasets, as well as permutations in the final ranking.
+	\includegraphics[width=\linewidth]{img/ablation_restarts}
+\end{figure}
+In addition to this motivating example, we carried out our main evaluation (Tab. \ref{tab:results}) multiple times by sampling 3 from our available runs uniformly at random and comparing the results.
+We found significant differences in the performance of AL methods on individual datasets, as well as permutations in the final ranking.
 This partly explains the ongoing difficulties in reproducing results for AL experiments and benchmarks.
 The details can be found in App. \ref{app:rank_difference}.
-For this benchmark we employ 50 restarts of every experiment.
+For this benchmark we employ 50 repetitions of every experiment.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Seeding vs. Repetitions}\label{sec:reproducibility}
+Considering the high computational cost of 50 repetitions, another approach to ensure consistency between experiments would be to reduce the amount of variance in the experiment by keeping as many subsystems (weight initialization, data splits, etc.) as possible fixed with specialized seeding.
+\\
+We describe a novel seeding strategy in Appendix \ref{app:seeding_strategy} that is capable of tightly controlling the amount variance in the experiment.
+However, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), providing a bad approximation of the quality of an AL method (i.e.
+measured performances for an method might not even transfer to a different model initialization).
+Hence, we opt for letting the subsystems vary in controlled way (For details, please refer to App.
+\ref{app:seeding_strategy}) and combine that with a high number of repetitions to obtain a good average of the generalization performance of each AL method.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{CDALBench: A Cross-Domain Active Learning Benchmark}\label{sec:datasets}
 \begin{wraptable}{R}{0.65\textwidth}
 	\vspace{-0.7cm}
-	\caption{Employed model, chosen budget and available batch sizes for each dataset}
+	\caption{Employed model, chosen budget and available query sizes for each dataset}
 	\vspace{0.1cm}
@@
 \end{wraptable}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Seeding vs. Restarts}\label{sec:reproducibility}
-Considering the high computational cost of 50 repetitions, another approach to ensure consistency between experiments would be to reduce the amount of variance in the experiment by keeping as many subsystems (weight initialization, data splits, etc.) as possible fixed with specialized seeding.
-\\
-We describe a novel seeding strategy in Appendix \ref{app:seeding_strategy} that creates 3 separate Random Number Generators (RNG) based on 3 different seeds.
-\todo{Wahrscheinlich zu wenig Erklärung für den main body. Eventuell die gesamten Erklärungen zum seeding in den appendix?}
-In short, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini-batch sampling, and $s_\theta$ for model initialization and sampling of dropout masks.
-Unless stated otherwise, we will keep $s_\Omega$ fixed, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
-While this seeding strategy is capable of controlling the amount variance in the experiment, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), providing a bad approximation of the quality of an AL algorithm (i.e.
-measured performances for an algorithm might not even transfer to a different model initialization).
-Hence, we opt for letting the subsystems vary (by increasing $s_\mathcal{D}$ and $s_\theta$) and combine that with a high number of restarts to obtain a good average of the generalization performance of each AL algorithm.
-\\
-Where a high number of restarts is computationally not feasible, we advise to additionally keep either $s_\mathcal{D}$ or $s_\theta$ (or both) fixed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{ALBench: A Cross-Domain Active Learning Benchmark}\label{sec:datasets}
 A detailed description of the preprocessing of each dataset can be found in Appendix \ref{app:hyperparameters}. \\ [1mm]
@@
 We, therefore, introduce a set of tabular datasets that we selected according to the following criteria:
-(i) They should be solvable by medium-sized models in under 1000 samples, (ii) the gap between most AL algorithms and random sampling should be significant (potential for AL is present) and (iii) the gap between the AL algorithms and our oracle should also be significant (research on these datasets can produce further lifts).
+(i) They should be solvable by medium-sized models in under 1000 samples, (ii) the gap between most AL methods and random sampling should be significant (potential for AL is present) and (iii) the gap between the AL methods and our oracle should also be significant (research on these datasets can produce further lifts).
 We use \textbf{Splice}, \textbf{DNA} and \textbf{USPS} from LibSVMTools \cite{libsvmtools}.\\
@@
 \textbf{Text:}
-\todo{Schwache Erklärung}
 We use \textbf{News Category} \cite{misra2022news} and \textbf{TopV2} \cite{chen-etal-2020-low-resource}.
-Text datasets have seen less attention in AL research, but most of the papers that evaluate on text (\cite{hu2021towards}, \cite{zhou2021towards}) use at least one of these datasets.\\ [1mm]
+Text datasets have seen less attention in AL research, but most of the papers that evaluate on text (\cite{hu2021towards}, \cite{zhou2021towards}) use at least one of these datasets. 
+We use both, as they complement each other in size and complexity. \\ [1mm]
 %
 We would like to point out that these datasets are selected for speed of computation (both in terms of the required classifier and the necessary budget to solve the dataset). 
-However, we are solely focused on comparing different AL algorithms in this paper and do not aim to develop novel classification models on these datasets.
-Our assumption is that a well-performing algorithm in our benchmark will also generalize well to larger real-world datasets, because we included multiple different data domains and classifier types and sizes in our experiments.
+However, we are solely focused on comparing different AL methods in this paper and do not aim to develop novel classification models on these datasets.
+Our assumption is that a well-performing method in our benchmark will also generalize well to larger real-world datasets, because we included multiple different data domains, classifier types and sizes in our experiments.
 \\ [1mm]
-\todo{Muss die Verbindung zwischen encodings und semi-supervised learning klarer?}
 Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning (semi-supervised).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Batch Sizes}\label{sec:batch_sizes}
-We selected batch sizes for each dataset to accommodate the widest range possible that results in a reasonable runtime for low batch sizes and allows for at least 4 round of data acquisition for high batch sizes.
-The available batch sizes per dataset can be found in Table \ref{tab:batch_sizes}.
+\subsection{Query Sizes}\label{sec:batch_sizes}
+We selected query sizes for each dataset to accommodate the widest range possible that results in a reasonable runtime for low query sizes and allows for at least 4 round of data acquisition for high query sizes.
+The available query sizes per dataset can be found in Table \ref{tab:batch_sizes}.
 
@@
 We would like to highlight the following design decisions: \\
-(i) Creating test and validation splits from the full dataset rather than only the labeled seed set.
-Fully fledged test and validation splits are unobtainable in practice, but they not only provide a better approximation of the algorithms generalization performance, but also a better foundation for hyperparameter tuning, which is bound to reduce variance in the experiment.
+(i) Creating test and validation splits from the full dataset rather than only the labeled seed set (following \cite{luth2024navigating}).
+Fully fledged test and validation splits are unobtainable in practice, but they provide not only a better approximation of the methods generalization performance, but also a better foundation for hyperparameter tuning, which is bound to reduce variance in the experiment.
 \\
@@
 \subsection{Evaluation Protocol}\label{sec:evaluation}
-Following \cite{zhou2021towards}, the quality of an AL algorithm is evaluated by an ``anytime protocol" that incorporates classification performance at every iteration, as opposed to evaluating final performance after the budget is exhausted.
+Following \cite{zhou2021towards}, the quality of an AL method is evaluated by an ``anytime protocol" that incorporates classification performance at every iteration, as opposed to evaluating final performance after the budget is exhausted.
 We employ the normalized area under the accuracy curve (AUC):
@@
 %AL algorithms have to perform well in both scenarios. \\ [1mm]
-Since AUC is still influenced by the budget, we define a set of rules to set this hyperparameter upfront, so that we are not favoring a subset of algorithms by handcrafting a budget.
-In this work, we choose the budget per dataset to be the first point at which one of 2 stopping conditions apply: (i) an algorithm (except Oracle) manages to reach 99\% of the full-dataset-performance (using the smallest query size) or (ii) the best algorithm (except oracle) did not improve the classifier's accuracy by at least 2\% in the last 20\% of iterations.
+Since AUC is still influenced by the budget, we define a set of rules to set this hyperparameter upfront, so that we are not favoring a subset of methods by handcrafting a budget.
+In this work, we choose the budget per dataset to be the first point at which one of 2 stopping conditions apply: (i) an method (except oracle) manages to reach 99\% of the full-dataset-performance (using the smallest query size) or (ii) the best method (except oracle) did not improve the classifier's accuracy by at least 2\% in the last 20\% of iterations.
 The first rule follows \cite{ji2023randomness}, while the second rule prevents excessive budgets for cases with diminishing returns in the budget.
 The resulting budgets can be found in Tab. \ref{tab:batch_sizes}. \\ [1mm]
-As described in Sec. \ref{sec:restarts}, we restart each experiment multiple times.
-Each restart retains the train/test split (often given by the dataset itself), but creates a new validation split that is sampled (based on $s_\D$) from the entire dataset (not just the seed set $\LL^{(0)}$).
+As described in Sec. \ref{sec:restarts}, we repeat each experiment multiple times.
+Each repetition retains the train/test split (often given by the dataset itself), but creates a new validation split that is sampled (based on $s_\D$) from the entire dataset (not just the seed set $\LL^{(0)}$).
 \\ [1mm]
@@
 Apart from plotting standard performance curves and reporting their AUC values per dataset in App.
-\ref{app:all_results}, we primarily rely on ranks to aggregate the performance of an acquisition function across datasets.
-For each dataset and query size, the AUC values of all acquisition functions are sorted and assigned a rank based on position, with the best rank being 1.
+\ref{app:all_results}, we primarily rely on ranks to aggregate the performance of an AL method across datasets.
+For each dataset and query size, the AUC values of all AL methods are sorted and assigned a rank based on position, with the best rank being 1.
 These ranks can safely be averages across datasets as they are no longer subjected to scaling differences of each dataset.
-\todo{Erklärung zum paired-t-test ist schlecht geschrieben. Wie machen wir es besser?}
 Additionally, we employ Critical Difference (CD) diagrams (like Fig. \ref{fig:ranks_by_domain}) for statistical testing.
-CD diagrams use the Wilcoxon signed-rank test, which is a variant of the paired T-test, to find significant differences of ranks between acquisition functions.
-For these diagrams, each combination of dataset, query size and run is considered a separate experiment, i.e.
-the results of \verb|Dataset1-QuerySize1-run5| of an acquisition function \verb|x| is only compared to the results of \verb|Dataset1-QuerySize1-run5| of acquisition function \verb|y|.
-Due to the large number of restarts and the wide range of datasets and query sizes, we can provide very accurate significance tests.
+CD diagrams \cite{IsmailFawaz2018deep} use the Wilcoxon signed-rank test, which is a variant of the paired T-test, to find significant differences of ranks between AL methods.
 For a detailed description of how every CD diagram is created, please refer to App. \ref{app:cd_diagrams}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A Greedy Oracle Algorithm}\label{sec:oracle} An oracle method for AL
-tries to find the oracle set $\mathcal{O}_b$ for a given dataset,\todo{Versuchen
-die Beste zu finden tut doch jede AL-Methode! Vlt sowas wie: Ist ein Verfahren,
-dass zwar die Optimale Lösung findet, aber dabei keine Rücksicht auf Laufzeiten nimmt} model, and
-training procedure that induces the probably highest AUC score for a given budget.
+\section{A Greedy Oracle Algorithm}\label{sec:oracle} 
+Using additional resources, like excessive training time, or direct access to the labeled test set, an oracle method for AL finds the oracle set $\mathcal{O}_b$ for a given dataset, model, and
+training procedure that induces the highest AUC score for a given budget.
 However, due to the combinatorial nature of the problem, it is computationally
@@
 approximations to this oracle sequence. \cite{zhou2021towards} used simulated
-annealing to search for the optimal subset and used the best  solution found
-after a fixed time budget. \todo{Okay und warum ist das dann keine Aquisition funnktion}Even though their reported performance curves display
-a significant lift over all other acquisition functions, we found the
+annealing to search for a subset with maximal test accuracy and used the best solution found
+after a fixed time budget. 
+Even though their reported performance curves display
+a significant lift over all other AL methods, we found the
 computational cost of reproducing this oracle for all our datasets to be
@@
 8 V100 GPUs). In this paper, we propose a greedy oracle algorithm that
-constructs an approximation of the optimal set in an iterative fashion. Our
-oracle algorithm evaluates every data point $u_k =
-\operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$\todo[inline]{Die Notation
-stimmt so nicht, willst du sagen, dass du hier ein Subset der größe $\tau$
-gesamplet hast uniform gesamplet hast? Dann schreibe das: "Our oracle algorithm
+constructs an approximation of the optimal set in an iterative fashion. 
+Our oracle algorithm
 uniformly samples at iteration $i$ a subset $\mathcal{U}_S$ of the not already
@@
 Then it recovers the label $y$ for each of the sampled $u \in \mathcal{U}_S$ and selects the
-point $u \in \mathcal{U}_S$ for which the classifier $\hat{y}$ trained on
-$\mathcal{L}^{(i)} \cup \{u\}$ has the maximal performance. } in a subsample of
-unlabeled points by recovering the label, fitting the classifier $\hat y$ on
-$\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test
-performance. \todo{Okay, hier verstehe ich was ein Orcale ist. Eine Aquisition
-funktion die schon das wissen über die Labels beim aquirieren nutzen kann?
-Stimmt das?}The data point that incurs the largest lift in test performance is
-selected and added to the labeled pool for that iteration.
-\todo{[Thorben] Ich habe die Erklärung für den Fallback im Oracle neu gemacht. Ist das verständlich?}
-Due to the algorithms greedy nature (just considering the next point to pick),
+point $u$ for which the classifier $\hat{y}$ trained on
+$\mathcal{L}^{(i)} \cup \{u\}$ has maximal performance.
+%Our oracle algorithm evaluates every data point $u_k =
+%\operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$ in a subsample of
+%unlabeled points by recovering the label, fitting the classifier $\hat y$ on
+%$\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test
+%performance.
+%The data point that incurs the largest lift in test performance is
+%selected and added to the labeled pool for that iteration.
+Due to the algorithms greedy nature (considering only the next point to pick),
 our oracle frequently encounters situations where every point in $u_k$ would
@@
 therefore the selected data point can be considered random. To circumvent this
-problem, we use our best-performing acquisition function (margin sampling
+problem, we use our best-performing AL method (margin sampling
 \cite{wang2014new}) as a fallback option for the oracle. Whenever the oracle
@@
 relatively low computational cost ($\mathcal{O}(B\tau)$).
-\todo{Da wir den Fallback haben, sollten wir das Oracle vielleicht einfach als Erweiterung von MarginSampling verkaufen?}
-We fix $\tau = 20$ in this work, as this gave us an average lift of 5\% over the
-best performing acquisition function per dataset (which is significant for AL
+We fix $\tau = 20$ in this work, as this gives us an average lift of 5\% over the
+best performing AL method per dataset (which is significant for AL
 settings) and we expect diminishing returns for larger $\tau$. The pseudocode
@@
 one and stop the oracle computation at a budget of 2000. The rest of the curve
-is forecast with a simple linear regression that asymptotically approaches the
+is forecast with a 2-stage linear regression that asymptotically approaches the
 upper bound performance of the dataset. A detailed description can be found in
-App. \ref{app:oracle_forecasting}.\todo[inline]{Approached heißt konvergiert,,
-oder? Dann muss das doch nicht stimmen. Ich würde nach "linear regression" einen
-Punkt machen und den Teil danach löschen.}
+App. \ref{app:oracle_forecasting}.
 
@@
 \subsection{Implementation Details}\label{sec:implementation_details}
-\todo{[Thorben] Sollten wir den ganzen ersten Teil steichen? Wir haben uns von technischen Beschreibungen weitesgehend wegbewegt.}
-At each iteration $i$ the acquisition function $\Omega$ picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \hat y^{(i)}, \text{opt}_{\hat y}\}$, where $\text{opt}_{\hat y}$ is the optimizer used to fit $\hat y^{(i)}$.
+At each iteration $i$ the AL method $\Omega$ picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \hat y^{(i)}, \text{opt}_{\hat y}\}$, where $\text{opt}_{\hat y}$ is the optimizer used to fit $\hat y^{(i)}$.
 This set grants full access to the labeled and unlabeled set, as well as all parameters of the classifier and the optimizer.
-Additionally, we provide meta-information, like the size of the seed set through $|\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|$, the remaining budget though the addition of $B$ and the classifiers potential though $\text{acc}^{(1)}$ and $\text{acc}^{(i)}$.
-We allow acquisition functions to derive information from this set, e.g.
+Additionally, we provide meta-information, like the size of the seed set through $|\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|$, the remaining budget though the addition of $B$ and the classifiers potential through $\text{acc}^{(1)}$ and $\text{acc}^{(i)}$.
+We allow AL methods to derive information from this set, e.g.
 predictions of the classifier $\hat y^{(i)}(x); \hspace{2mm} x \in \mathcal{U}^{(i)} \cup \mathcal{L}^{(i)}$, clustering, or even training additional models.
-However, the algorithm may not incorporate external information e.g. other datasets, queries to recover additional labels, additional training steps for $\hat y$, or the test/validation set. \\
+However, the method may not incorporate external information e.g. other datasets, queries to recover additional labels, additional training steps for $\hat y$, or the test/validation set. \\
 For our study we selected acquisition functions with good performances reported by multiple different sources that can work with the set of information stated above.
-For a list of all acquisition functions, please refer to Table \ref{tab:results}, with detailed descriptions being found in Appendix \ref{app:acquisition_functions}. \\ [1mm]
+For a list of all AL methods, please refer to Table \ref{tab:results}, with detailed descriptions being found in Appendix \ref{app:acquisition_functions}. \\ [1mm]
 %
@@
 \subsection{Results on Real-world Data}
-In Table \ref{tab:results} we provide the rank of each acquisition function per dataset and averaged for each (un-)encoded dataset. 
-Please note, that for Tab. \ref{tab:results} we are averaging not only over runs, but also over query sizes per dataset. 
-This leads to a negative impact for AL methods that do not adapt well to larger batch sizes.
-For the results per query size, please refer to App. \ref{app:auc_by_query_size}. \\
-As mentioned in contribution \ref{cont:domains}, our results on real-world data shows significant differences in the performance algorithms between data domains.
-Not only do some algorithms overperform on some domains (like least confidence sampling on Images), but the Top-3 of algorithms (except Oracle) does not contain the same three algorithms for any two domains.
-Most interestingly, the image domain, which received most of the attention in benchmarking so far could even be considered an outlier, as this is the only domain where the Top-1 algorithm changes.
-This highlights the dire need for diverse data domains in AL benchmarking. \\ [1mm]
+In Table \ref{tab:results} we provide the rank of each AL method per dataset. 
+Please note, that we are averaging not only over runs, but also over query sizes per dataset. 
+This leads to a negative impact for AL methods that do not adapt well to larger query sizes.
+For the results per query size, please refer to App. \ref{app:auc_by_query_size}. 
+As mentioned in contribution \ref{cont:domains}, our results on real-world data show significant differences in the performance of methods between data domains.
+Not only do some methods overperform on some domains (like least confidence (LC) sampling on images), but the Top-3 of methods (except oracle) does not contain the same three methods for \textbf{any} two domains.
+Most interestingly, the image domain, which received most of the attention in benchmarking so far could even be considered an outlier, as this is the only domain where the Top-1 method changes.
+This highlights the dire need for diverse data domains in AL benchmarking. \\
+Results for the semi-supervised domain appear mostly in line with the other 3 domains, but a closer analysis of performances split into semi-sup.
+images and semi-sup.
+tabular reveals the need for further research.
+For details, please refer to App. \ref{app:semi_sup_analysis}.
 \begin{table}[]
-	\caption{Performances for acquisition functions on real-world datasets, aggregated for un-encoded and encoded datasets.
-Performance is shown as average ranks over restarts (1.0 is the best rank).
-Algorithms are sorted by aggregated performance on un-encoded datasets.
+	\caption{Performances for AL methods on real-world datasets, aggregated for un-encoded (normal) and encoded (semi-supervised) datasets.
+Performance is shown as average ranks over repetitions (1.0 is the best rank).
+Methods are sorted by aggregated performance on un-encoded (normal) datasets.
 }
@@
 \begin{tabular}{l|lllllll|ll}
-          & Splice & DNA  & USPS 		 	& Cifar10 & FMnist & TopV2 & News  & Un-enc. & Enc. \\
+          & Splice & DNA  & USPS 		 	& Cifar10 & FMnist & TopV2 & News  & Normal & Semi \\
           \hline
@@
     \centering
-    \caption{Ranks of each acquisition function aggregated by domain.
-Horizontal bars indicate a \textbf{non}-significant rank difference.
-The significance is tested via a paired-t-test with $\alpha=0.05$.}
+    \caption{Ranks of each AL method aggregated by domain. Horizontal bars indicate a \textbf{non}-significant rank difference. The significance is tested via a paired-t-test with $\alpha=0.05$.}
     \label{fig:ranks_by_domain}
@@
     \includegraphics[width=0.49\linewidth]{img/macro_enc.jpg}
-    \vspace{-7mm}
+    \vspace{-5mm}
 \end{figure}
@@
 \section{Honeypot and Diverging Sine}
-AL approaches can be categorized into two types, uncertainty and geometric approaches.
-Typical members of the first category are variants of uncertainty sampling like entropy-, margin and least-confident-sampling \cite{wang2014new} as well as BALD \cite{gal2017deep}.
+AL approaches can be categorized into two types: uncertainty and geometric approaches.
+Typical members of the first category are variants of uncertainty sampling like entropy, margin and LC sampling \cite{wang2014new} as well as BALD \cite{gal2017deep}.
 Typical members of the second category are clustering approaches like Coreset \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
-Both types of algorithms have principled shortcomings in terms of the utilized information that makes them unsuitable for certain data distributions. 
-To test for these specific shortcomings, we created two synthetic datasets, namely "Honeypot" and "Diverging Sine", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively.
-To avoid algorithms memorizing these datasets they are generated from scratch for each experiment, depending on $s_\D$. \\
+Both types of methods have principled shortcomings in terms of the utilized information that makes them unsuitable for certain data distributions. 
+To test for these specific shortcomings, we created two synthetic datasets, namely ``Honeypot" and ``Diverging Sine", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively.
+To avoid methods memorizing these datasets they are generated from scratch for each experiment, depending on $s_\D$. \\
 \begin{figure}[]
 	\centering
-	\caption{Synthetic "Honeypot" and "Diverging Sine" datasets. The optimal decision boundary is not part of the dataset and serves only as a visual guide.}
+	\caption{Synthetic ``Honeypot" and ``Diverging Sine" datasets. The optimal decision boundary is not part of the dataset and serves only as a visual guide.}
+	\label{fig:synth_results}
 	\includegraphics[width=0.4\linewidth]{img/honeypot.jpg}
+	\hspace{10mm}
 	\includegraphics[width=0.4\linewidth]{img/diverging_sin.jpg}
-	\label{fig:synthDataAppendix}
-\end{figure}
-%
-\todo{Hier könnte gekürzt werden. Wir brauchen wohl kaum die genaue Anzahl der Punkte}
-Honeypot creates to two easy to distinguish clusters with 150 samples each and one overlapping "honeypot" that represents a noisy region of the dataset with potentially miss-labeled, miss-measured or generally adverse samples.
-This honeypot contains 150 samples of each class, creating a balance of 50\% beneficial samples and 50\% adverse samples in the dataset.
-The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its negative impact on purely uncertainty based acquisition functions.
-Diverging Sine samples the datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.
-This creates a challenging region one the left hand side, where a lot of datapoints need to be sampled and an easy region on the right hand side, where very few datapoints are enough. 
-The repeating nature of a sin function encourages diversity based acquisition functions to equally sample the entire length, drastically oversampling the right hand side of the dataset.
-Each class has 500 datapoints. 
-Both datasets have a budget of $B=60$ and are tested with query sizes 1 and 5.\\
-%
-Results for the Honeypot dataset reveal expected shortcomings of uncertainty sampling algorithms like margin, entropy and least confident sampling as well as BALD.
-In addition, BADGE is underperforming for this dataset compared to real-world data. 
-Results for Diverging Sine also confirm expected behavior, as clustering algorithms (Coreset, TypiClust) fall behind uncertainty algorithms (Entropy-, Margin-Sampling), with the exception of BADGE. \\
-We provide a small ablation study on the importance of the embeddings by testing a version of Coreset and TypiClust on this dataset that does not use the embeddings produced by the classification model, but rather clusters the data directly.
-"Coreset Raw" and "TypiClust Raw" both perform worse than their embedding-based counterpart.
-\begin{figure}
-	\centering
-	\caption{Results for all acquisition functions on both synthetic datasets.}
-	\label{fig:main_body_result}
 	\includegraphics[width=0.49\linewidth]{img/micro_honeypot.jpg}
 	\includegraphics[width=0.49\linewidth]{img/micro_diverging_sin.jpg}
-\end{figure}
-
+	\vspace{-5mm}
+\end{figure}
+%
+Honeypot creates two easy to distinguish clusters and one ``honeypot" that represents a noisy region of the dataset with potentially miss-labeled, miss-measured or generally adverse samples.
+%This honeypot contains 150 samples of each class, creating a balance of 50\% beneficial samples and 50\% adverse samples in the dataset.
+The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its negative impact on purely uncertainty-based AL methods.
+Diverging Sine samples the datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.
+This creates a challenging region on the left hand side, where a lot of datapoints need to be sampled, and an easy region on the right hand side, where very few datapoints are sufficient. 
+The repeating nature of a sine function encourages diversity-based AL methods to equally sample the entire length, drastically oversampling the right hand side of the dataset. \\
+Both datasets have a budget of $B=60$ and are tested with query sizes 1 and 5. \\[1mm]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Results}
-Our results on Honeypot reveal principled shortcomings for the two best algorithms in BADGE and margin sampling.
-Both are vulnerable to adverse samples or simply measurement noise, which highlights the need for further research in this area. \\
-Finally, the fact that BADGE is able to perform well on Diverging Sine highlights the importance of embeddings for the clustering algorithms, as the so-called gradient embedding from BADGE seems to be able to encode uncertainty information, guiding the selection into the left hand regions of the dataset.
-We also show that embeddings are generally useful for this dataset, by providing results for "Coreset Raw" and "TypiClust Raw".
+%\subsection{Results}
+We provide the rank of all AL Methods on Honeypot and Diverging Sine in Fig. \ref{fig:synth_results}.
+Results for the Honeypot dataset reveal expected shortcomings of uncertainty sampling methods like margin, entropy and LC sampling as well as BALD.
+In addition, BADGE is underperforming for this dataset compared to real-world data. 
+Both margin sampling and BADGE (the two best methods) being vulnerable to adverse samples or simply measurement noise, highlights the need for further research in this area. \\
+%
+Results for Diverging Sine also confirm expected behavior, as clustering methods (Coreset, TypiClust) fall behind uncertainty methods (entropy, margin, LC sampling), with the exception of BADGE. 
+The fact that BADGE is able to perform well on Diverging Sine highlights the importance of embeddings for the clustering methods, as the gradient embedding from BADGE seems to be able to encode uncertainty information, guiding the selection into the left hand regions of the dataset.
+We provide a small ablation study on the importance of the embeddings by testing a version of Coreset and TypiClust on this dataset that does not use the embeddings produced by the classification model, but rather clusters the data directly.
+``Coreset Raw" and ``TypiClust Raw" both perform worse than their embedding-based counterpart.
 
 \section{Conclusion}
-We strongly advocate to test newly proposed AL algorithms not only on a wide variety of real data domains, but also to pay close attention to the Honeypot and Diverging Sine datasets to reveal principled shortcomings of the algorithm in question.
-Both tasks can be easily carried out by implementing the new acquisition function into our code base.
-
-%\section{Future Work}
-% Wahrscheinlich werden wir keinen Platz dafür haben
-% Wie sehr hängen die Unterschiede am Classifier und nicht an den Daten? -> Mehrere Classifier pro Domäne
-% Mehr Datensätze pro Domäne
-% Neuere Algorithmen. (Neuster von 2022)
+We strongly advocate to test newly proposed AL methods not only on a wide variety of real data domains, but also to pay close attention to the Honeypot and Diverging Sine datasets to reveal principled shortcomings of the method in question.
+Both tasks can be easily carried out by implementing the new AL method into our code base.
+For Limitations and Future Work, please refer to App. \ref{app:future_work}.
 
@@
 \bibliography{al_benchmark_paper.bib}
-\begin{itemize}
-	\item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
-	\item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-	\item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-
+
+\newpage
+\section*{NeurIPS Checklist}
 \begin{enumerate}
@@
 		\item Did you describe the limitations of your work?
-		\answerTODO{}
+		\answerYes{We have prepared a Limitations and Future Work section in App. \ref{app:future_work}}
 		\item Did you discuss any potential negative societal impacts of your work?
-		\answerTODO{}
+		\answerNo{We strongly believe that there will be no negative impact of our work, as we have only used publicly available datasets, models and methods}
 		\item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-		\answerTODO{}
+		\answerYes{}
 	\end{enumerate}
@@
 		\item Did you state the full set of assumptions of all theoretical results?
-		\answerTODO{}
+		\answerNo{We do not provide theoretical results.}
 		\item Did you include complete proofs of all theoretical results?
-		\answerTODO{}
+		\answerNo{We do not provide theoretical results.}
 	\end{enumerate}
@@
 		\item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-		\answerTODO{}
+		\answerYes{We provide a GitHub link in Sec. \ref{sec:introduction}.}
 		\item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-		\answerTODO{}
+		\answerYes{All Hyperparameters can be found in App. \ref{app:hyperparameters}}
 		\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-		\answerTODO{}
+		\answerYes{All tables provide standard deviations}
 		\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-		\answerTODO{}
+		\answerYes{Please refer to App. \ref{app:compute}}
 	\end{enumerate}
@@
 		\item If your work uses existing assets, did you cite the creators?
-		\answerTODO{}
+		\answerYes{}
 		\item Did you mention the license of the assets?
-		\answerTODO{}
+		\answerNA{Licences are public.}
 		\item Did you include any new assets either in the supplemental material or as a URL?
-		\answerTODO{}
+		\answerNo{}
 		\item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-		\answerTODO{}
+		\answerNA{All datasets are public.}
 		\item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-		\answerTODO{}
+		\answerNo{The data does not contain personal information.}
 	\end{enumerate}
@@
 		\item Did you include the full text of instructions given to participants and screenshots, if applicable?
-		\answerTODO{}
+		\answerNo{}
 		\item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-		\answerTODO{}
+		\answerNo{}
 		\item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-		\answerTODO{}
+		\answerNo{}
 	\end{enumerate}
@@
 \section{AL Recommendations from Ji et al.}\label{app:recommendations}
-TODO
-
-Recommendation. Use the backbone architecture with
+\paragraph{Recommendation 1} Use the backbone architecture with
 the community-accepted definition that is best suited
@@
 experiments. In the image classification domain, we
-suggest using ResNet18 for CIFAR-10 and CIFAR-100.
-
-Recommendation. Control the type of optimizer across
+suggest using ResNet18 for CIFAR-10 and CIFAR-100. \\
+$\rightarrow$ We are using Resnet18 for our image datasets and re-purposed the LSTM model from \cite{zhou2021towards} for the text datasets.
+For the tabular data, we ran a grid-search (full dataset test accuracy) over MLP architectures.
+
+\paragraph{Recommendation 2} Control the type of optimizer across
 methods for comparative evaluations to ensure that the
 yield performance difference stems from an active learning method itself. As SGD often generalizes better, we
-encourage its use for deep active learning.
-
-Recommendation. Pragmatically fix the learning rate
+encourage its use for deep active learning.\\
+$\rightarrow$ We searched for the best optimizer per dataset via generalization performance (test accuracy).
+
+\paragraph{Recommendation 3} Pragmatically fix the learning rate
 to 0.1 for SGD on image datasets. While continuous
@@
 a fixed learning rate does not change the ranking of AL
-methods from a comparative evaluation’s point of view.
-
-Recommendation. One may use data augmentation if
+methods from a comparative evaluation’s point of view.\\
+$\rightarrow$ We found that for smaller datasets, a learning rate of 0.1 was unsuitable.
+Ji et al.
+only used large image datasets, so a learning rate of 0.1 was sufficient for them.
+The learning rate is part of our hyperparameter grid-search.
+
+\paragraph{Recommendation 4} One may use data augmentation if
 applied consistently across methods, such that it does
 not affect the overall ranking. However, a commonly
-accepted baseline is needed, e.g., random horizontal flipping and random cropping for image classification.
-
-Recommendation. Refine model parameters across AL
+accepted baseline is needed, e.g., random horizontal flipping and random cropping for image classification.\\
+$\rightarrow$ We did not find a data augmentation technique that could be applied equally on all datasets, so we refrained from it.
+The only possibility would be Gaussian noise, but the impact of gaussian noise on pre-encoded is not well-understood.
+
+\paragraph{Recommendation 5} Refine model parameters across AL
 batch (“warm starts”) to prevent exhaustive reinitialization and feed initialization of the backbone model’s
 weights and the “init sets” with fixed inputs over multiple runs to average out the randomness. Moreover, use
-identical seeds for all methods under investigation.
-
-Recommendation. Run experiments multiple times to
+identical seeds for all methods under investigation.\\
+$\rightarrow$ We employed a novel seeding strategy to closely control the seeding of our experiments (Details in App.
+\ref{app:seeding_strategy}).
+We applied warm-starts for most datasets, except pre-encoded ones, because we found a generally better performance, if we train the classifier from scratch.
+
+\paragraph{Recommendation 6} Run experiments multiple times to
 compensate for non-deterministic operations. If the resulting variance is larger than the gained improvement,
-use deterministic operations stringently
-
-Recommendation SW. Configure and verify influence parameter in active learning implementations thoroughly. To foster future research, we
+use deterministic operations stringently\\
+$\rightarrow$ A study on (non-)deterministic operations has not been conducted in this work, but our large number of repetitions compensate for that.
+
+\paragraph{Recommendation 7SW} Configure and verify influence parameter in active learning implementations thoroughly. To foster future research, we
 provide implementations as part of our framework at:
-https://intellisec.de/research/eval-al
-
-Recommendation HW. Ensure that comparative evaluations are run on identical hardware. While it is not
+https://intellisec.de/research/eval-al\\
+$\rightarrow$ We provide our own baseline code, since we implemented a novel seeding strategy and unify many additional data domains in the code.
+
+\paragraph{Recommendation 7HW} Ensure that comparative evaluations are run on identical hardware. While it is not
 necessary to execute all experiments on the same physical device, the GPU model, for instance, should be the
-same. Do not mix hardware and list hardware details.
-
-Recommendation. Consider multiple query-batch sizes
+same. Do not mix hardware and list hardware details.\\
+$\rightarrow$ The large computational cost of our benchmark did not allow us to compute on only one type of hardware (Which would mean to only use part of our cluster).
+However, our large number of repetitions compensate for that.
+
+\paragraph{Recommendation 8} Consider multiple query-batch sizes
 in the evaluation. The choice of the sizes needs to be
-appropriate for the total number of unlabeled samples.
-
-Recommendation. Compare active learning strategies
+appropriate for the total number of unlabeled samples.\\
+$\rightarrow$ We employed a wide range of batch sizes. For details, please refer to Table \ref{tab:batch_sizes}.
+
+\paragraph{Recommendation 9} Compare active learning strategies
 without sub-sampling, unless one of the approaches uses
 it as a fundamental building block. In this case a detailed
-analysis of the influence of sub-sampling is necessary.
-
-Recommendation. Evaluate active learning strategies
+analysis of the influence of sub-sampling is necessary.\\
+$\rightarrow$ We only carefully employed sub-sampling, when it was absolutely necessary to keep the computation times feasible.
+
+\paragraph{Recommendation 10} Evaluate active learning strategies
 on multiple benchmark datasets, that comprise balanced,
 imbalanced, small-scale, and large-scale datasets to cover
-most relevant cases in practice.
-
-Recommendation. For a comprehensive analysis of
+most relevant cases in practice.\\
+$\rightarrow$ We deferred the study on (im-)balanced datasets to future work. However, our benchmark contains datasets of many different sizes and we extend this argument to domains as well.
+
+\paragraph{Recommendation 11} For a comprehensive analysis of
 AL strategies, the overall comparative evaluation should
 incorporate as many variables from Section 3 to yield a
-summarized PPM that is as expressive as possible.
+summarized PPM that is as expressive as possible. \\
+$\rightarrow$ In our main result (Tab. \ref{tab:results}) we average the performance over runs, query sizes and datasets.
 
@@
 \section{AL Pitfalls from Lueth et al.}\label{app:pitfalls}
-TODO
-
-P1 Data distribution. The proposed evaluation over a diverse selection of dataset distributions
+\paragraph{P1 Data distribution} The proposed evaluation over a diverse selection of dataset distributions
 including specific roll-out datasets proved essential for realistic evaluation of QMs as well as the
@@
 which are typically prone to experience cold start problems. This observation is consistent with a few
-previous studies [30, 35, 54]. Further, our results underpin the importance of the roll-out datasets e.g.
-when looking at the sub-random performance of BALD (with Self-SL) and Entropy (with ST) on
-MIO-TCD. Such worst-case failures of AL application (increased compute and labeling effort due to
-AL) could not have been predicted based on development data where all AL-parameters are optimized.
-Another example is the lack of generalizability of Semi-SL, where performance in relation to other
-Self-SL and ST decreases gradually with data complexity (going from CIFAR-10 to CIFAR-100/-10
-LT to the two roll-out datasets MIO-TCD and ISIC-2019.)
-
-P2 Starting budget. The comprehensive study of various starting budgets on all datasets reveals
+previous studies [...].\\
+$\rightarrow$ We selected our datasets according to their ``potential" for AL.
+We measured this potential by the distance of most AL methods to random and the distance of the best AL method to our oracle.
+If both distances are >0, we consider the dataset useful.
+
+\paragraph{P2 Starting budget} The comprehensive study of various starting budgets on all datasets reveals
 that AL methods are more robust with regard to small starting budgets than previously reported
@@
 finding has great impact potential suggesting that AL can be applied at earlier points in the annotation
-process thereby further reducing the labeling cost, especially in combination with BADGE which
-performed consistently better or on par with Random. Similarly for Self-SL models, AL performs
-well on small starting budgets with the exceptions of CIFAR-100 (BALD, Entropy) and MIO-TCD
-(BALD, Core-Set, Entropy).
-
-P3 Query size. Based on our evaluation of the query size we can empirically confirm its importance
+process thereby further reducing the labeling cost [...].\\
+$\rightarrow$ Our experiments also showed a high resistance against the cold-start problem, which prompted us to use the smallest possible seed set for most datasets (1 point per class), with the only exception being (un-)encoded Cifar10 and FashionMnist.
+Here we employ a seed set of 100 points per class to avoid a cold-start.
+
+\paragraph{P3 Query size} Based on our evaluation of the query size we can empirically confirm its importance
 with regard to 1) general AL performance and 2) counteracting the cold start problem. The are,
 however, surprising findings indicating that the exact interaction between query size and performance
-remains an open research question. For instance, we observe the cold start problem for BALD with
-Self-SL on CIFAR-100 (~50\% accuracy at 2k labeled samples for query sizes of 500 (low-label
-regime) and 1k (mid-label regime). On the other hand, in the high label regime (budget of 5k and
-query size of 5k) ST and Self-SL models with similar accuracies of 50\% and 60\%, respectively,
-benefit from BALD. Since cold start problems are commonly associated with large query sizes, this
-finding seems counter-intuitive, but has been reported before although without further investigations
-[6, 20, 43]. To gain a better understanding of this phenomenon and how QMs interact with query
-size in the low-label regime, we performed a dedicated experiment series for Self-SL training on
-CIFAR-100 and ISIC-2019 (see Figure 3 and Appendix F.2). Due to a significant improvement
-of BALD on CIFAR-100 for even smaller query sizes from sub-Random to Random performance
-and no worsening on ISIC-2019, we can conclude that small query sizes represent an effective
-countermeasure against the cold-start problem for BALD contrary to the findings of [44, 58] and
-currently not considered as a solution [6, 20, 43]. This potentially explains the gap between these
-findings and more theoretical works which advertise using the smallest query size possible [19, 33].
-Importantly, while smaller query sizes appear as a potential solution for the observed instabilities of
-BALD, they seem to have no considerable effect on the other QMs. Moreover, our ablation reveals
-that even in the low-label settings, BADGE is the overall most reliable of all compared QMs and
-exhibits no sub-Random performance. This adds to the existing reports of the robustness of BADGE
-for higher label regimes [4, 58].
-
-P4 Classifer configuration. Our results show that method configuration on a properly sized validation
-set is essential for realistic evaluation in AL. For instance, our method configuration has the same
-effect on the classifier performance as increasing the number of labeled training samples by a factor
-of ~5 (e.g. our ST model reach approximately the same accuracy of ~44\% trained on 200 samples
-compared to models in the AL study by Bengar et al. [6] trained on 1k samples). The effectiveness of
-our proposed lightweight HP selection on the starting budget including only three parameters (Sec. 3)
-is demonstrated by the fact that all our ST models substantially outperform respective models found
-in relevant literature [6, 20, 30, 35, 43, 54, 55] where HP optimization is generally neglected. Details
-for this comparison are provided in Appendix H. This raises the question of to which extent reported
+remains an open research question. [...].\\
+$\rightarrow$ We generally observed a decreasing performance for larger query sizes.
+We therefore made sure, that we include the smallest possible query sizes that result in feasible computation times.
+
+\paragraph{P4 Classifer configuration} Our results show that method configuration on a properly sized validation
+set is essential for realistic evaluation in AL. [...] This raises the question of to which extent reported
 AL advantages could have been achieved by simple classifier configurations. Further, our models
-also generally outperform expensively configured models by Munjal et al. [44]. Thus, we conclude
+also generally outperform expensively configured models by Munjal et al.. Thus, we conclude
 that manually constraining the search space renders HP optimization feasible in practice without
@@
 importance of the proposed strategy to optimize HPs on the starting budget for each new dataset is
-supported by the fact that the resulting configurations change across datasets.
-
-P5 Alternative training paradigms. Based on our study benchmarking AL in the context of
+supported by the fact that the resulting configurations change across datasets.\\
+$\rightarrow$ We also strongly advocate the use of a fully fledged validation set for HP tuning, as this allows for a higher quality of HPs, which in turn reduces the variance of the experiments.
+
+\paragraph{P5 Alternative training paradigms} Based on our study benchmarking AL in the context of
 both Self-SL and Semi-SL, we see that while Self-SL generally leads to improvements across all
@@
 with either of the two training paradigms receive a lower performance gain from AL (over random
-querying) compared to ST. Crucially, Self-SL models converge around 2.5 times faster than ST
-models, while the training time of Semi-SL models is around 80 times longer than ST and often yields
-only small benefits over Self-SL or ST models. The fact that AL entails multiple training iterations
+querying) compared to ST. [...] The fact that AL entails multiple training iterations
 amplifies the computational burden of Semi-SL, rendering their combination prohibitively expensive
@@
 to generalize to more complex datasets in our setting stands in stark contrast to conclusions drawn by
-[20, 43] as to which the emergence of Semi-SL renders AL redundant. Interestingly, the exact settings
+[...] as to which the emergence of Semi-SL renders AL redundant. Interestingly, the exact settings
 where Semi-SL does not provide benefits in our study are the ones where AL proved advantageous.
 The described contradiction with the literature underlines the importance of our proposed protocol
-testing for a method’s generalizability to unseen datasets. This is especially critical for Semi-SL,
-which is known for unstable performance on noisy and class imbalanced datasets as noted by studies
-focusing on Semi-SL [7, 22, 23, 29, 45, 56].
-
-OPTIONAL - aus deren conclusion
-Main empirical insights revealed by our study:
-\begin{itemize}
-\item Assessment of AL methods in the literature is substantially hampered by subpar classifier configurations, but meaningful assessment can be restored by our protocol for lightweight hyperparameter
-tuning on the starting budget.
-\item AL generally provides substantial gains in class-imbalanced settings.
-\item BADGE is the best-performing QM across a realistic range of datasets, starting budgets, and query
-sizes and exhibits no failure modes (i.e. sub-Random performance).
-\item Combining AL with Self-SL considerably improves performance, shortens the training time, and
-stabilizes optimization, especially in low-label settings.
-\item FixMatch results indicate that Semi-SL methods perform well on datasets where they have been
-developed, but may struggle to generalize to more realistic scenarios such as class imbalanced data.
-Combining the paradigm with AL additionally suffers from extensive training times.
-\item BALD with Self-SL pre-trained models benefits from smaller query sizes on small starting budgets
-possibly circumventing the "cold start problem".
-\end{itemize}
-
-Take-aways for developing and proposing new Active Learning algorithms:
-AL methods should be tested for generalizability on roll-out datasets and come with a clear recipe
-for real-world application including how to adapt all design choices to new settings. Since the
-expected benefit of AL on a new setting increases with lower application costs, we believe there is a
-high potential for wide-spread real-world use of AL by reducing the two prevalent cost factors: 1)
-Engineering costs, which are reduced by building easy-to-use AL tools. 2) Computational costs,
-which are reduced by explicitly including methods that shorten the training time in AL.
-
-Take-aways for the cost-benefit analysis of deploying Active Learning:
-\begin{enumerate}
-\item Since we identify BADGE as a robust query method exhibiting no sub-random-sampling performance across all tested settings, the potential harm of AL is minimized allowing to base decisions
-around deploying AL on the described cost-benefit analysis.
-\item The expected benefit is high in settings, where there is a mismatch between the task-specific
-importance of individual classes and their frequency in the dataset (e.g. class-imbalanced datasets
-in combination with tasks requiring a balanced classifier).
-\item The expected benefit further increases with the labeling cost that will be reduced by AL. AL is thus
-likely to yield a net benefit in settings with high label cost and low computational and engineering
-costs.
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acquistion Functions}\label{app:acquisition_functions}
-\textbf{Uncertainty Sampling} 
-tries to find the sample that the classifier is most uncertain about by computing heuristics of the class probabilities.
-For our benchmark, we use entropy and margin (a.k.a.
-best-vs-second-best) sampling.\\
-\textbf{BALD \cite{kirsch2019batchbald}}
-applies the query-by-committee strategy of model ensembles to a single model by interpreting the classifier's parameters as distributions and then sample multiple outputs from them via Monte-Carlo dropout.\\
-\textbf{BADGE \cite{ashdeep}} uses gradient embeddings of unlabeled points to select samples where the classifier is expected to change a lot.
-The higher the magnitude of the gradient the higher the expected improvement in model performance.\\
-\textbf{Coreset \cite{sener2017active}}
-employs K-Means clustering trying to cover the whole data distribution.
-Selects the unlabeled sample that is the furthest away from all cluster centers.
-Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\hat y$.
-In this work, we use the greedy variant of Coreset.\\
-\textbf{TypiClust \cite{hacohen2022active}}
-relies on clustering similar to Coreset, but proposes a new measure called ``Typicality" to select unlabeled samples.
-It selects points that are in the densest regions of clusters that do not contain labeled samples yet.
-Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\hat y$.
-It has to be pointed out that TypiClust was designed for low-budget scenarios, but we think it is still worthwhile to test and compare this algorithm with higher budgets. \\
-\textbf{Core-GCN \cite{caramalau2021sequential}} TODO \\
-\textbf{DSA/LSA \cite{kim2019guiding}} TODO \\ [2mm]
-%
-\textbf{Excluded Algorithms}\\
-\textbf{Learning Loss for AL \cite{yoo2019learning}}
-Introduces an updated training of the classification model with an auxiliary loss and therefore cannot be compared fairly against classification models without this boosted training regime.\\ [1mm]
-%
-\textbf{Reinforcement Learning Algorithms} \\
-%TODO \\
+testing for a method’s generalizability to unseen datasets. \\
+$\rightarrow$ Due to the high computational costs described by Lueth et al., we opted for the most efficient form of semi-supervised learning, which is to train a fixed encoder-model, pre-encode the datasets and then only train a single linear layer as classifier.
 
@@
 \section{Difference of Ranks with 3 Repetitions}\label{app:rank_difference}
-Table \ref{tab:rank_diff_1} and Table \ref{tab:rank_diff_2} follow the exact same computation of ranks that created the main result (Table \ref{tab:results}) with the only difference being a reduced number of runs per acquisition function.
-For each table we sampled 3 runs uniformly at random from the available 50 per acquisition function. \\
+Table \ref{tab:rank_diff_1} and Table \ref{tab:rank_diff_2} follow the exact same computation of ranks that created the main result (Table \ref{tab:results}) with the only difference being a reduced number of runs per AL method.
+For each table we sampled 3 runs uniformly at random from the available 50 per AL method. \\
 We can observe significant differences between the two tables: \\
-{\color{purple}Purple}: A multitude of rank differences of acquisition functions for specific datasets, some as high as 4.7 ranks for TypiClust on the Splice dataset \\
-{\color{olive}Olive}: Well separated acquisition functions in Tab. \ref{tab:rank_diff_2} (Margin and BADGE) are almost indistinguishable in Tab \ref{tab:rank_diff_1} \\
+{\color{purple}Purple}: A multitude of rank differences of AL methods for specific datasets, some as high as 4.7 ranks for TypiClust on the Splice dataset \\
+{\color{olive}Olive}: Well separated AL methods in Tab. \ref{tab:rank_diff_2} (Margin and BADGE) are almost indistinguishable in Tab \ref{tab:rank_diff_1} \\
 {\color{red}Red}: BALD lost 2 places in the overall ranking and Entropy gained 2 \\ [1mm]
-Even though the overall ordering of acquisition functions stayed relatively unchanged due to the averaging across many datasets, each individual dataset was subject to drastic permutations.
+Even though the overall ordering of AL methods stayed relatively unchanged due to the averaging across many datasets, each individual dataset was subject to drastic permutations.
 This highlights the need for many repetitions in AL experiments.
@@
 	\centering
-	\caption{Ranks of all acquisition functions per dataset. First random draw of 3 runs from the overall pool of 50.}
+	\caption{Ranks of all AL methods per dataset. First random draw of 3 runs from the overall pool of 50.}
 	\label{tab:rank_diff_1}
@@
 	\centering
-	\caption{Ranks of all acquisition functions per dataset. Second random draw of 3 runs from the overall pool of 50.}
+	\caption{Ranks of all AL methods per dataset. Second random draw of 3 runs from the overall pool of 50.}
 	\label{tab:rank_diff_2}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{AUCs by Query Size}\label{app:auc_by_query_size}
-% QS 1
-%TODO highlight best results
-\begin{table}[H]
-	\caption{AUC values for each dataset that supports query size 1.}
-	\resizebox{\columnwidth}{!}{
-	\begin{tabular}{l|lllllllllll}
-		       & Splice       & SpliceEncoded& DNA          & DNAEncoded  & USPS          & USPSEncoded  & Cifar10Encoded& FashionMnistEnc& TopV2     & DivergingSin & ThreeClust \\
-		\hline
-		Oracle & 0.803+-0.012 & 0.678+-0.021 & 0.825+-0.009 & 0.721+-0.013 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.884+-0.006 & 0.957+-0.009 & 0.783+-0.03 \\
-		Margin & 0.769+-0.021 & \textbf{0.678+-0.032} & 0.806+-0.013 & 0.642+-0.047 & \textbf{0.858+-0.006} & 0.426+-0.038 &  0.653+-0.013 & 0.68+-0.012 & \textbf{0.861+-0.009} & 0.941+-0.018 & 0.704+-0.074 \\
-		Badge & 0.767+-0.02 & 0.661+-0.026 & 0.78+-0.014 & 0.642+-0.046 & 0.83+-0.008 & 0.371+-0.035 &  0.656+-0.013 & 0.68+-0.009 & 0.826+-0.024 & 0.941+-0.017 & 0.69+-0.083 \\
-		LeastConfident & 0.779+-0.019 & 0.68+-0.032 & 0.809+-0.01 & 0.629+-0.05 & 0.846+-0.009 & 0.421+-0.039 &  \textbf{0.668+-0.014} & \textbf{0.685+-0.009} & 0.843+-0.013 & 0.94+-0.016 & 0.692+-0.094 \\
-		DSA & 0.766+-0.021 & 0.691+-0.022 & 0.803+-0.01 & \textbf{0.646+-0.032} & 0.829+-0.01 & \textbf{0.431+-0.05} & 0.663+-0.014 & 0.679+-0.01 & 0.844+-0.017 & 0.941+-0.014 & \textbf{0.731+-0.032} \\
-		BALD & \textbf{0.78+-0.014} & 0.649+-0.04 & 0.784+-0.01 & 0.632+-0.042 & 0.819+-0.01 & 0.242+-0.046 & 0.666+-0.014 & 0.644+-0.018 & 0.815+-0.024 & 0.928+-0.014 & 0.698+-0.043 \\
-		CoreGCN & 0.765+-0.021 & 0.686+-0.023 & 0.804+-0.012 & \textbf{0.646+-0.03} & 0.753+-0.016 & 0.39+-0.044 & 0.623+-0.018 & 0.647+-0.012 & 0.85+-0.01 & 0.938+-0.014 & \textbf{0.731+-0.028} \\
-		Entropy & 0.768+-0.022 & \textbf{0.678+-0.035} & \textbf{0.812+-0.013} & 0.635+-0.045 & 0.83+-0.011 & 0.399+-0.035 & 0.663+-0.013 & 0.681+-0.011 & 0.815+-0.021 & \textbf{0.942+-0.017} & 0.696+-0.083 \\
-		LSA & 0.772+-0.016 & 0.68+-0.026 & 0.787+-0.012 & 0.618+-0.036 & 0.821+-0.009 & 0.422+-0.037 & 0.613+-0.014 & 0.642+-0.012 & 0.816+-0.013 & 0.932+-0.016 & 0.727+-0.033 \\
-		Random & 0.76+-0.016 & 0.674+-0.027 & 0.774+-0.013 & 0.63+-0.035 & 0.823+-0.009 & 0.404+-0.036 & 0.613+-0.014 & 0.639+-0.013 & 0.815+-0.012 & 0.933+-0.017 & 0.721+-0.036 \\
-		Coreset & 0.772+-0.016 & 0.69+-0.017 & 0.79+-0.012 & 0.638+-0.041 & 0.767+-0.016 & 0.404+-0.046 & 0.659+-0.011 & 0.684+-0.009 & 0.826+-0.022 & 0.937+-0.014 & 0.73+-0.031 \\
-		TypiClust & 0.762+-0.016 & 0.685+-0.025 & 0.778+-0.01 & 0.663+-0.028 & 0.828+-0.007 & 0.396+-0.046 & 0.653+-0.013 & 0.649+-0.007 & 0.831+-0.011 & 0.934+-0.018 & 0.727+-0.033
-	\end{tabular}
-}
-\end{table}
-% QS 5
-\begin{table}[H]
-	\caption{AUC values for each dataset that supports query size 5.}
-	\resizebox{\columnwidth}{!}{
-	\begin{tabular}{l|lllllllllll}
-		& Splice & SpliceEncoded & DNA & DNAEncoded & USPS & USPSEncoded & Cifar10Encoded & FashionMnistEncoded & TopV2 & DivergingSin & ThreeClust \\
-		\hline
-		Oracle & 0.803+-0.012 & 0.678+-0.021 & 0.825+-0.009 & 0.721+-0.013 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.884+-0.006 & 0.957+-0.009 & 0.783+-0.03\\
-		Margin & 0.765+-0.021 & 0.662+-0.032 & 0.794+-0.011 & 0.611+-0.05 & 0.855+-0.006 & 0.508+-0.02 & 0.656+-0.014 & 0.678+-0.009 & 0.848+-0.013 & 0.923+-0.019 & 0.697+-0.055 \\
-		Badge & 0.768+-0.014 & 0.646+-0.035 & 0.785+-0.011 & 0.624+-0.036 & 0.846+-0.007 & 0.48+-0.021 & 0.647+-0.012 & 0.67+-0.009 & 0.847+-0.01 & 0.924+-0.019 & 0.72+-0.036 \\
-		LeastConfident & 0.763+-0.023 & 0.643+-0.034 & 0.798+-0.013 & 0.585+-0.065 & 0.831+-0.014 & 0.478+-0.028 & 0.67+-0.01 & 0.681+-0.009 & 0.819+-0.023 & 0.921+-0.019 & 0.675+-0.072 \\
-		DSA & 0.765+-0.023 & 0.653+-0.029 & 0.793+-0.009 & 0.613+-0.034 & 0.822+-0.01 & 0.489+-0.024 & 0.661+-0.013 & 0.662+-0.012 & 0.833+-0.02 & 0.924+-0.018 & 0.718+-0.033 \\
-		BALD & 0.775+-0.018 & 0.641+-0.034 & 0.801+-0.013 & 0.592+-0.054 & 0.84+-0.008 & 0.332+-0.054 & 0.681+-0.011 & 0.681+-0.013 & 0.824+-0.023 & 0.893+-0.035 & 0.673+-0.041 \\
-		CoreGCN & 0.759+-0.018 & 0.662+-0.027 & 0.79+-0.011 & 0.62+-0.03 & 0.755+-0.011 & 0.45+-0.03 & 0.604+-0.016 & 0.609+-0.013 & 0.837+-0.014 & 0.922+-0.018 & 0.723+-0.034 \\
-		Entropy & 0.765+-0.022 & 0.66+-0.03 & 0.798+-0.011 & 0.611+-0.054 & 0.823+-0.013 & 0.464+-0.024 & 0.663+-0.013 & 0.672+-0.011 & 0.801+-0.025 & 0.924+-0.02 & 0.689+-0.066 \\
-		LSA & 0.769+-0.016 & 0.654+-0.032 & 0.781+-0.013 & 0.61+-0.041 & 0.82+-0.009 & 0.484+-0.022 & 0.617+-0.012 & 0.641+-0.011 & 0.816+-0.012 & 0.915+-0.018 & 0.718+-0.038 \\
-		Random & 0.758+-0.015 & 0.655+-0.026 & 0.771+-0.013 & 0.623+-0.031 & 0.82+-0.009 & 0.476+-0.024 & 0.616+-0.016 & 0.637+-0.012 & 0.812+-0.014 & 0.921+-0.018 & 0.713+-0.034 \\
-		Coreset & 0.765+-0.017 & 0.663+-0.023 & 0.784+-0.014 & 0.603+-0.034 & 0.765+-0.015 & 0.449+-0.022 & 0.657+-0.009 & 0.674+-0.009 & 0.817+-0.017 & 0.92+-0.017 & 0.713+-0.035 \\
-		TypiClust & 0.759+-0.014 & 0.641+-0.028 & 0.775+-0.01 & 0.603+-0.04 & 0.757+-0.02 & 0.465+-0.027 & 0.596+-0.014 & 0.567+-0.012 & 0.727+-0.026 & 0.916+-0.02 & 0.693+-0.045
-	\end{tabular}
-}
-\end{table}
-% QS 20
-\begin{table}[H]
-	\caption{AUC values for each dataset that supports query size 20.}
-	\resizebox{\columnwidth}{!}{
-	\begin{tabular}{l|lllllllll}
-		       & Splice       & SpliceEncoded& DNA          & USPS         & USPSEncoded  & Cifar10Encoded& FashionMnistEnc & TopV2 & News \\
-		\hline
-		Oracle & 0.803+-0.012 & 0.678+-0.021 & 0.825+-0.009 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.884+-0.006 & 0.49+-0.003 \\
-		Margin & 0.759+-0.027 & 0.618+-0.04 & 0.779+-0.013 & 0.847+-0.008 & 0.439+-0.027 & 0.656+-0.01 & 0.67+-0.011 & 0.823+-0.014 & 0.464+-0.007 \\
-		Badge & 0.767+-0.013 & 0.619+-0.033 & 0.776+-0.013 & 0.845+-0.006 & 0.44+-0.019 & 0.647+-0.013 & 0.665+-0.007 & 0.827+-0.016 & 0.463+-0.007 \\
-		LeastConfident & 0.751+-0.02 & 0.597+-0.05 & 0.748+-0.025 & 0.798+-0.027 & 0.391+-0.024 & 0.665+-0.013 & 0.669+-0.011 & 0.775+-0.035 & 0.467+-0.008 \\
-		DSA & 0.759+-0.02 & 0.599+-0.034 & 0.769+-0.013 & 0.809+-0.012 & 0.421+-0.023 & 0.647+-0.014 & 0.63+-0.013 & 0.793+-0.026 & 0.459+-0.01 \\
-		BALD & 0.768+-0.022 & 0.57+-0.037 & 0.784+-0.015 & 0.822+-0.009 & 0.298+-0.039 & 0.675+-0.008 & 0.673+-0.01 & 0.789+-0.024 & 0.468+-0.009 \\
-		CoreGCN & 0.759+-0.018 & 0.612+-0.039 & 0.774+-0.012 & 0.754+-0.016 & 0.397+-0.026 & 0.587+-0.015 & 0.583+-0.015 & 0.807+-0.018 & 0.453+-0.006 \\
-		Entropy & 0.759+-0.027 & 0.618+-0.038 & 0.773+-0.015 & 0.803+-0.019 & 0.372+-0.022 & 0.656+-0.011 & 0.65+-0.012 & 0.773+-0.031 & 0.451+-0.007 \\
-		LSA & 0.761+-0.014 & 0.611+-0.039 & 0.768+-0.015 & 0.816+-0.009 & 0.411+-0.022 & 0.621+-0.01 & 0.635+-0.011 & 0.796+-0.016 & 0.452+-0.007 \\
-		Random & 0.755+-0.014 & 0.612+-0.039 & 0.763+-0.012 & 0.818+-0.009 & 0.439+-0.019 & 0.622+-0.013 & 0.633+-0.012 & 0.795+-0.016 & 0.45+-0.006 \\
-		Coreset & 0.759+-0.016 & 0.601+-0.034 & 0.764+-0.015 & 0.757+-0.015 & 0.39+-0.029 & 0.647+-0.009 & 0.651+-0.011 & 0.784+-0.026 & 0.435+-0.012 \\
-		TypiClust & 0.751+-0.012 & 0.551+-0.036 & 0.76+-0.016 & 0.643+-0.026 & 0.411+-0.024 & 0.488+-0.02 & 0.449+-0.017 & 0.652+-0.035 & 0.406+-0.011
-	\end{tabular}
-}
-\end{table}
-% QS 50
-\begin{table}[H]
-	\caption{AUC values for each dataset that supports query size 50.}
-	\resizebox{\columnwidth}{!}{
-	\begin{tabular}{l|llllllll}
-		       & Splice       & DNA          & USPS         & USPSEncoded  & Cifar10Encoded& FashionMnistEnc& TopV2     & News \\
-		\hline
-		Oracle & 0.803+-0.012 & 0.825+-0.009 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.884+-0.006 & 0.49+-0.003 \\
-		Margin & 0.747+-0.023 & 0.751+-0.019 & 0.828+-0.009 & 0.363+-0.031 & 0.64+-0.013 & 0.653+-0.01 & 0.774+-0.029 & 0.46+-0.006 \\
-		Badge & 0.758+-0.017 & 0.754+-0.018 & 0.831+-0.008 & 0.376+-0.028 & 0.632+-0.013 & 0.649+-0.011 & 0.781+-0.026 & 0.462+-0.007 \\
-		LeastConfident & 0.731+-0.025 & 0.688+-0.041 & 0.761+-0.037 & 0.291+-0.03 & 0.644+-0.013 & 0.65+-0.011 & 0.73+-0.049 & 0.462+-0.009 \\
-		DSA & 0.748+-0.021 & 0.738+-0.018 & 0.783+-0.016 & 0.346+-0.027 & 0.624+-0.014 & 0.588+-0.016 & 0.748+-0.041 & 0.45+-0.011 \\
-		BALD & 0.76+-0.017 & 0.756+-0.018 & 0.796+-0.016 & 0.241+-0.026 & 0.65+-0.009 & 0.645+-0.01 & 0.746+-0.038 & 0.455+-0.007 \\
-		CoreGCN & 0.755+-0.016 & 0.745+-0.018 & 0.752+-0.019 & 0.328+-0.027 & 0.581+-0.015 & 0.568+-0.018 & 0.771+-0.025 & 0.453+-0.007 \\
-		Entropy & 0.747+-0.024 & 0.748+-0.018 & 0.778+-0.024 & 0.275+-0.026 & 0.633+-0.011 & 0.625+-0.012 & 0.734+-0.036 & 0.442+-0.007 \\
-		LSA & 0.754+-0.013 & 0.749+-0.019 & 0.807+-0.01 & 0.341+-0.029 & 0.613+-0.012 & 0.625+-0.01 & 0.763+-0.025 & 0.45+-0.006 \\
-		Random & 0.746+-0.012 & 0.745+-0.015 & 0.806+-0.008 & 0.379+-0.028 & 0.615+-0.014 & 0.621+-0.01 & 0.759+-0.026 & 0.448+-0.006 \\
-		Coreset & 0.751+-0.016 & 0.733+-0.019 & 0.74+-0.017 & 0.325+-0.034 & 0.624+-0.012 & 0.608+-0.013 & 0.731+-0.045 & 0.432+-0.012 \\
-		TypiClust & 0.749+-0.016 & 0.736+-0.016 & 0.586+-0.038 & 0.348+-0.027 & 0.451+-0.024 & 0.375+-0.022 & 0.614+-0.046 & 0.397+-0.012
-	\end{tabular}
-}
-\end{table}
-% QS 100
-\begin{table}[H]
-	\caption{AUC values for each dataset that supports query size 100.}
-	\resizebox{\columnwidth}{!}{
-	\begin{tabular}{l|lllllll}
-		       & Splice       & DNA          & USPS         & USPSEncoded & Cifar10Encoded& FashionMnistEnc& News \\
-		\hline
-		Oracle & 0.803+-0.012 & 0.825+-0.009 & 0.866+-0.004 & 0.436+-0.057 & 0.749+-0.009 & 0.755+-0.005 & 0.49+-0.003 \\
-		Margin & 0.733+-0.024 & 0.711+-0.027 & 0.799+-0.013 & 0.473+-0.026 & 0.629+-0.012 & 0.628+-0.009 & 0.455+-0.006 \\
-		Badge & 0.743+-0.014 & 0.714+-0.032 & 0.804+-0.013 & 0.472+-0.029 & 0.623+-0.01 & 0.621+-0.01 & 0.456+-0.006 \\
-		LeastConfident & 0.715+-0.033 & 0.639+-0.05 & 0.708+-0.034 & 0.23+-0.034 & 0.631+-0.013 & 0.62+-0.012 & 0.457+-0.008 \\
-		DSA & 0.729+-0.021 & 0.697+-0.031 & 0.753+-0.021 & 0.427+-0.028 & 0.609+-0.013 & 0.546+-0.017 & 0.442+-0.01 \\
-		BALD & 0.744+-0.015 & 0.718+-0.024 & 0.765+-0.021 & 0.285+-0.046 & 0.632+-0.009 & 0.609+-0.01 & 0.444+-0.007 \\
-		CoreGCN & 0.742+-0.015 & 0.713+-0.025 & 0.744+-0.019 & 0.433+-0.032 & 0.583+-0.013 & 0.554+-0.015 & 0.448+-0.007 \\
-		Entropy & 0.733+-0.023 & 0.713+-0.031 & 0.743+-0.026 & 0.395+-0.037 & 0.618+-0.012 & 0.59+-0.012 & 0.432+-0.007 \\
-		LSA & 0.738+-0.017 & 0.716+-0.027 & 0.789+-0.011 & 0.439+-0.03 & 0.609+-0.013 & 0.608+-0.01 & 0.447+-0.006 \\
-		Random & 0.733+-0.013 & 0.713+-0.023 & 0.789+-0.012 & 0.468+-0.024 & 0.611+-0.01 & 0.606+-0.01 & 0.446+-0.005 \\
-		Coreset & 0.735+-0.019 & 0.698+-0.026 & 0.721+-0.021 & 0.396+-0.024 & 0.608+-0.012 & 0.562+-0.016 & 0.426+-0.012 \\
-		TypiClust & 0.733+-0.016 & 0.704+-0.025 & 0.592+-0.042 & 0.427+-0.027 & 0.501+-0.02 & 0.338+-0.02 & 0.383+-0.012
-	\end{tabular}
-}
-\end{table}
-\begin{minipage}{0.47\linewidth}
-	% QS 500
-	\begin{table}[H]
-		\caption{AUC values for each dataset that supports query size 500.}
-		\begin{tabular}{l|ll}
-			& Cifar10 & FashionMnist \\
-			\hline
-			Oracle & 0.689+-0.001 & 0.905+-0.001 \\
-			Margin & 0.556+-0.008 & 0.882+-0.004 \\
-			Badge & 0.56+-0.008 & 0.883+-0.005 \\
-			LeastConfident & 0.591+-0.01 & 0.884+-0.005 \\
-			DSA & 0.56+-0.009 & 0.882+-0.004 \\
-			BALD & 0.478+-0.014 & 0.878+-0.003 \\
-			CoreGCN & 0.553+-0.01 & 0.88+-0.007 \\
-			Entropy & 0.553+-0.009 & 0.882+-0.006 \\
-			LSA & 0.558+-0.01 & 0.866+-0.005 \\
-			Random & 0.557+-0.01 & 0.863+-0.005 \\
-			Coreset & 0.553+-0.007 & 0.878+-0.006 \\
-			TypiClust & 0.557+-0.009 & 0.864+-0.004
-		\end{tabular}
-	\end{table}
-\end{minipage}
-\hspace{10mm}
-\begin{minipage}{0.47\linewidth}
-	% QS 1000
-	\begin{table}[H]
-		\caption{AUC values for each dataset that supports query size 1000.}
-		\begin{tabular}{l|ll}
-			& Cifar10 & FashionMnist \\
-			\hline
-			Oracle & 0.689+-0.001 & 0.905+-0.001 \\
-			Margin & 0.56+-0.011 & 0.872+-0.007 \\
-			Badge & 0.562+-0.013 & 0.871+-0.007 \\
-			LeastConfident & 0.561+-0.012 & 0.873+-0.006 \\
-			DSA & 0.56+-0.011 & 0.87+-0.008 \\
-			BALD & 0.535+-0.011 & 0.866+-0.003 \\
-			CoreGCN & 0.557+-0.011 & 0.867+-0.012 \\
-			Entropy & 0.557+-0.014 & 0.871+-0.009 \\
-			LSA & 0.551+-0.012 & 0.854+-0.009 \\
-			Random & 0.55+-0.01 & 0.855+-0.006 \\
-			Coreset & 0.562+-0.012 & 0.869+-0.004 \\
-			TypiClust & 0.552+-0.011 & 0.854+-0.009
-		\end{tabular}
-	\end{table}
-\end{minipage}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Critical Difference Diagrams}\label{app:cd_diagrams}
-%TODO
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Individual Results}\label{app:all_results}
-% Splice
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for normal (un-encoded) Splice}
-    \includegraphics[width=\linewidth]{img/eval_splice}
-\end{figure}
-%Splice Enc
-\begin{figure}[H]
-\centering
-\caption{Performance curves per query size for semi-supervised (encoded) Splice}
-\includegraphics[width=\linewidth]{img/eval_splice_enc}
-\end{figure}
-%DNA
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for normal (un-encoded) DNA}
-	\includegraphics[width=\linewidth]{img/eval_dna}
-\end{figure}
-%DNA Enc
-\begin{figure}[H]
-\centering
-\caption{Performance curves per query size for semi-supervised (encoded) DNA}
-\includegraphics[width=\linewidth]{img/eval_dna_enc} 
-\end{figure}
-%USPS
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for normal (un-encoded) USPS}
-	\includegraphics[width=\linewidth]{img/eval_usps}
-\end{figure}
-%USPS Enc
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for semi-supervised (encoded) USPS}
-    \includegraphics[width=\linewidth]{img/eval_usps_enc}
-\end{figure}
-%Cifar10
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for normal (un-encoded) Cifar10}
-    \includegraphics[width=\linewidth]{img/eval_cifar10}
-\end{figure}
-%Cifar10 Encoded
-\begin{figure}[H]
-\centering
-\caption{Performance curves per query size for semi-supervised (encoded) Cifar10}
-\includegraphics[width=\linewidth]{img/eval_cifar10_enc}
-\end{figure}
-%FMnist
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for normal (un-encoded) FashionMnist}
-    \includegraphics[width=\linewidth]{img/eval_fmnist}
-\end{figure}
-%FashionMnist Encoded
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for semi-supervised (encoded) FashionMnist}
-    \includegraphics[width=\linewidth]{img/eval_fmnist_enc}
-\end{figure}
-%TopV2
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for normal (GloVe) TopV2}
-\includegraphics[width=\linewidth]{img/eval_topv2}
-\end{figure}
-%News
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for normal (GloVe) News}
-\includegraphics[width=\linewidth]{img/eval_news}
-\end{figure}
-%ThreeClust
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for normal (un-encoded) Honeypot}
-\includegraphics[width=\linewidth]{img/eval_honeypot}
-\end{figure}
-%DivergingSin
-\begin{figure}[H]
-    \centering
-    \caption{Performance curves per query size for normal (un-encoded) Diverging Sine}
-\includegraphics[width=\linewidth]{img/eval_diverging_sine}
-\end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
 \section{Seeding Strategy}\label{app:seeding_strategy}
-We aim to provide an experimental setup that is fully reproducible independent of the dataset, classification model, or AL algorithm used.
-For a fair comparison of two AL algorithms, both algorithms need to receive equal starting conditions in terms of train/validation split, initialization of classifier, and even the state of minor systems like the optimizer or mini-batch sampler.
+We aim to provide an experimental setup that is fully reproducible independent of the dataset, classification model, or AL method used.
+For a fair comparison of two AL methods, both methods need to receive equal starting conditions in terms of train/validation split, initialization of classifier, and even the state of minor systems like the optimizer or mini-batch sampler.
 Even though different implementations might have their own solution to some of these problems, only \cite{ji2023randomness} has described and implemented a fully reproducible pipeline for AL evaluation.
@@
 Even though some workarounds exist, e.g. re-setting the seed multiple times, this problem is not limited to the initialization phase, but also extends to the AL iterations and the systems within.
-We propose an implementation that creates separate Random Number Generators (RNGs) for each of these systems to ensure equal testing conditions even when the AL algorithm, dataset, or classifier changes.
-We hypothesize that the insufficient setup with global seeds contributes to the ongoing problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
-In summary, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini-batch sampling, and $s_\theta$ for model initialization and sampling of dropout masks.
-Unless stated otherwise, we will keep $s_\Omega$ fixed, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
-Some algorithms require a subsample to be drawn from $\mathcal{U}$ in order to reduce the computational cost in each iteration, while others need access to the full unlabeled pool (e.g.
+We propose an implementation that creates separate Random Number Generators (RNGs) for each of these systems to ensure equal testing conditions even when the AL method, dataset, or classifier changes.
+We hypothesize that the insufficient setup with global seeds contributes to the ongoing problem of inconsistent results of AL methods in different papers. \\ [1mm]
+In summary, we introduce three different seeds: $s_\Omega$ for the AL method, $s_\mathcal{D}$ for dataset splitting and mini-batch sampling, and $s_\theta$ for model initialization and sampling of dropout masks.
+Unless stated otherwise, we will keep $s_\Omega$ fixed, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between repetitions to introduce stochasticity into our framework.
+Some methods require a subsample to be drawn from $\mathcal{U}$ in order to reduce the computational cost in each iteration, while others need access to the full unlabeled pool (e.g.
 for effective clustering).
 If a subsample is required, it will be drawn from $s_\Omega$ and therefore will not influence other systems in the experiments.
-For each algorithm, we decided if subsampling is required based on our available hardware, but decided against setting a fixed time limit per experiment, since this would introduce unnecessary complexity into the benchmark.
-An overview of selected hyperparameters per AL algorithm can be found in Appendix \ref{app:agent_hyperparameters}. \\
+For each method, we decided if subsampling is required based on our available hardware, but decided against setting a fixed time limit per experiment, since this would introduce unnecessary complexity into the benchmark.
+An overview of selected hyperparameters per AL method can be found in Appendix \ref{app:agent_hyperparameters}. \\
 \textbf{Note:} Even though we decoupled the subsystems via the described seeds, the subsystems can still influence each other in a practical sense. 
-For example, keeping $s_\mathcal{D}$ fixed does not mean that always the same sequence of samples from $\mathcal{U}$ (if subsamples are drawn) are shown to all acquisition functions. 
-This is practically impossible, as different acquisition functions pick different $x^{(i)}$.
-However, the hypothetical \textbf{tree} of all possible sequences of samples from $\mathcal{U}$ remains the same, granting every acquisition function equal possibilities.
-
-\section{Oracle Curve Forecasting}\label{app:oracle_forecasting}
-TODO
-
-\section{Hyperparameters per AL Algorithm}\label{app:agent_hyperparameters}
-\begin{table}[H]
-    \caption{Selected hyperparameters for all tested acquisition functions. Last column indicates the source of our implementation.}
-	\centering
-	\begin{tabular}{l || l | l | l}
-		Algorithm & \makecell{Sample\\Size} & Other & Source\\
-		\hline
-		BADGE & 100 & & Based on \cite{ashdeep, curelab}\\
-		BALD & 100 & Dropout Trials: 5 & Based on \cite{pycls} \\
-		Coreset & 8000 & & Own \\
-		TypiClust & 10000 & \makecell[tl]{Min Cluster Size: 5\\Max \# Clusters: 500} & Based on \cite{hacohen2022active} \\
-		Margin & 8000 & & Own\\
-		Entropy & 8000 &  & Own\\
-	\end{tabular}
-\end{table}
+For example, keeping $s_\mathcal{D}$ fixed does not mean that always the same sequence of samples from $\mathcal{U}$ (if subsamples are drawn) are shown to all AL methods. 
+This is practically impossible, as different AL methods pick different $x^{(i)}$.
+However, the hypothetical \textbf{tree} of all possible sequences of samples from $\mathcal{U}$ remains the same, granting every AL methods equal possibilities.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Performance Curves per Dataset}\label{app:all_results}
+% Splice
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for normal (un-encoded) Splice}
+	\includegraphics[width=\linewidth]{img/eval_splice}
+\end{figure}
+%Splice Enc
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for semi-supervised (encoded) Splice}
+	\includegraphics[width=\linewidth]{img/eval_splice_enc}
+\end{figure}
+%DNA
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for normal (un-encoded) DNA}
+	\includegraphics[width=\linewidth]{img/eval_dna}
+\end{figure}
+%DNA Enc
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for semi-supervised (encoded) DNA}
+	\includegraphics[width=\linewidth]{img/eval_dna_enc} 
+\end{figure}
+%USPS
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for normal (un-encoded) USPS}
+	\includegraphics[width=\linewidth]{img/eval_usps}
+\end{figure}
+%USPS Enc
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for semi-supervised (encoded) USPS}
+	\includegraphics[width=\linewidth]{img/eval_usps_enc}
+\end{figure}
+%Cifar10
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for normal (un-encoded) Cifar10}
+	\includegraphics[width=\linewidth]{img/eval_cifar10}
+\end{figure}
+%Cifar10 Encoded
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for semi-supervised (encoded) Cifar10}
+	\includegraphics[width=\linewidth]{img/eval_cifar10_enc}
+\end{figure}
+%FMnist
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for normal (un-encoded) FashionMnist}
+	\includegraphics[width=\linewidth]{img/eval_fmnist}
+\end{figure}
+%FashionMnist Encoded
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for semi-supervised (encoded) FashionMnist}
+	\includegraphics[width=\linewidth]{img/eval_fmnist_enc}
+\end{figure}
+%TopV2
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for normal (GloVe) TopV2}
+	\includegraphics[width=\linewidth]{img/eval_topv2}
+\end{figure}
+%News
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for normal (GloVe) News}
+	\includegraphics[width=\linewidth]{img/eval_news}
+\end{figure}
+%ThreeClust
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for normal (un-encoded) Honeypot}
+	\includegraphics[width=\linewidth]{img/eval_honeypot}
+\end{figure}
+%DivergingSin
+\begin{figure}[H]
+	\centering
+	\caption{Performance curves per query size for normal (un-encoded) Diverging Sine}
+	\includegraphics[width=\linewidth]{img/eval_diverging_sine}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Critical Difference Diagrams}\label{app:cd_diagrams}
+We adapted the code for the CD diagrams from \cite{cd_diagram_code}. \\
+To compare each AL method, we consider each combination of dataset, query size and run is considered a separate ``experiment", i.e.
+the results of \verb|Dataset1-QuerySize1-run5| of an AL Method \verb|x| is only compared to the results of \verb|Dataset1-QuerySize1-run5| of AL method \verb|y|.
+\\
+Depending on the use-case, we build the following ``experiments":
+\begin{itemize}
+	\item Single dataset - single query size: Each AL method has 50 ``experiments" in it's 50 repetitions
+	\item Single dataset - all query sizes (Fig. \ref{fig:synth_results}): Each ``experiment" is represented by a string \verb|query_size_<qs>_run_<id>|
+	\item Multiple dataset - all query sizes (Fig. \ref{fig:ranks_by_domain}): Each ``experiment" is represented by a string \verb|dataset_<dataset>_query_size_<qs>_run_<id>|
+\end{itemize}
+Due to the large number of restarts and the wide range of datasets and query sizes, we can provide very accurate significance tests.
+
 \section{AL Pseudocode}\label{app:pseudocode}
@@
 		\Require $\LL, \U, \D_\text{test}, \operatorname{Train}, \operatorname{Seed}, \hat y$
-		\Require $\Omega$ \Comment{Acquisition Function}
+		\Require $\Omega$ \Comment{AL Method}
 		\State $\LL^{(1)} \gets \operatorname{Seed}(\U)$  \Comment{Create the initial labeled set}
@@
 \end{algorithm}
-Alg. \ref{alg:oracle} replaces the acquisition function $\Omega$ in the AL loop (Alg. \ref{app:pseudocode} line 5).
+Alg. \ref{alg:oracle} replaces the AL method $\Omega$ in the AL loop (Alg. \ref{app:pseudocode} line 5).
+
+\section{Oracle Curve Forecasting}\label{app:oracle_forecasting}
+Unfortunately, the iterative nature of our oracle means that the computational effort scales in the budget $B$.
+For datasets with large budgets, like Cifar10 and FashionMnist (both 10K), we were unable to compute the oracle set for the entire 10K iterations. \\
+We compromised by (i) picking the two points with highest test accuracy, instead of only one and (ii) only computed until iteration 2000. \\
+The rest of the curve was forecast using a simple 2-step algorithm, based on linear regression:
+\begin{enumerate}
+	\item Fit a linear regression model on the second 50\% of the existing oracle curve (to accurately capture the trend of the oracle, rather than the intercept) and forecast the oracle performance for the remaining budget.
+	\item Post-process the oracle forecast by letting it asymptotically approach the upper bound performance of the dataset.
+\end{enumerate}
+\begin{align}
+	o_i &= \operatorname{min} \begin{cases} 
+		o_i \\
+		\phi(i) * o_i + (1 - \phi(i)) * \text{upper bound} 
+	\end{cases} \\
+	\phi(i) &= e^{-i / 0.5}
+\end{align}
+\begin{figure}[H]
+	\centering
+	\caption{(left) Oracle forecast (dotted line) for FashionMnist with query size 500; (right) function $\phi$ that governs the approach towards the upper bound performance.}
+	\includegraphics[width=0.55\linewidth]{img/eval_fmnist_500.pdf}
+	\includegraphics[width=0.43\linewidth]{img/phi_example.pdf}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{AL Methods}\label{app:acquisition_functions}
+\textbf{Uncertainty Sampling} 
+tries to find the sample that the classifier is most uncertain about by computing heuristics of the class probabilities.
+For our benchmark, we use entropy and margin (a.k.a.
+best-vs-second-best) sampling.\\
+\textbf{BALD \cite{kirsch2019batchbald}}
+applies the query-by-committee strategy of model ensembles to a single model by interpreting the classifier's parameters as distributions and then sample multiple outputs from them via Monte-Carlo dropout.\\
+\textbf{BADGE \cite{ashdeep}} uses gradient embeddings of unlabeled points to select samples where the classifier is expected to change a lot.
+The higher the magnitude of the gradient the higher the expected improvement in model performance.
+BADGE employs a variant to the KMeans++ initialization technique to select batches of points.
+Even though \cite{ashdeep} provided pseudocode for this procedure that selects the first point at random, all found implementations of BADGE select the first points instead by maximum gradient magnitude.
+\\
+\textbf{Coreset \cite{sener2017active}}
+employs K-Means clustering trying to cover the whole data distribution.
+Selects the unlabeled sample that is the furthest away from all cluster centers.
+Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\hat y$.
+In this work, we use the greedy variant of Coreset.\\
+\textbf{TypiClust \cite{hacohen2022active}}
+relies on clustering similar to Coreset, but proposes a new measure called ``Typicality" to select unlabeled samples.
+It selects points that are in the densest regions of clusters that do not contain labeled samples yet.
+Clustering is done in a semantically meaningful space by encoding the data with the current classifier $\hat y$.
+It has to be pointed out that TypiClust was designed for low-budget scenarios, but we think it is still worthwhile to test and compare this method with higher budgets. \\
+\textbf{Core-GCN \cite{caramalau2021sequential}} trains a Graph-Convolutional-Network (GCN) on embeddings of the unlabeled pool, obtained from the classifier (Similar to Coreset and TypiClust). 
+This GCN model propagates uncertainty information through the graph and therefore enhances the nodes uncertainty quantification.
+Lastly, the node that displays the highest amount of uncertainty is selected for labeling.
+\\
+\textbf{DSA/LSA \cite{kim2019guiding}} use the metric of test adequacy to construct a set of points that is diverse, ranging from points that are close to points in $\LL$ and points that are significantly different.
+DSA and LSA measure the diversity of points by distance in embedding space or likelihood estimation under the given classifier respectively.
+\\ [2mm]
+%
+\textbf{Excluded Methods}\\
+\textbf{Learning Loss for AL \cite{yoo2019learning}}
+Introduces an updated training of the classification model with an auxiliary loss and therefore cannot be compared fairly against classification models without this boosted training regime.\\ [1mm]
+%
+\textbf{Reinforcement Learning Methods} \\
+We postpone the study of learned AL methods to future versions of this benchmark, as reinforcement learning is infamous for being extremely time consuming and itself hard to reproduce .
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{AUCs by Query Size}\label{app:auc_by_query_size}
+All tables are sorted according to the main result in Table \ref{tab:results}.
+% QS 1
+\begin{table}[H]
+	\caption{AUC values for each dataset that supports query size 1.}
+	\resizebox{\columnwidth}{!}{
+		\begin{tabular}{l|llllllllllll}
+			& Wins & Splice       & SpliceEnc& DNA          & DNAEnc  & USPS          & USPSEnc  & Cifar10Enc& FMnistEnc& TopV2     & Diverging Sine & ThreeClust \\
+			\hline
+			Oracle &  & 0.803 $\pm$ {\small 0.012} & 0.678 $\pm$ {\small 0.021} & 0.825 $\pm$ {\small 0.009} & 0.721 $\pm$ {\small 0.013} & 0.866 $\pm$ {\small 0.004} & 0.436 $\pm$ {\small 0.057} & 0.749 $\pm$ {\small 0.009} & 0.755 $\pm$ {\small 0.005} & 0.884 $\pm$ {\small 0.006} & 0.957 $\pm$ {\small 0.009} & 0.783 $\pm$ {\small 0.03} \\
+
+			Margin & 3 & 0.769 $\pm$ {\small 0.021} & \textbf{0.678 $\pm$ {\small 0.032}} & 0.806 $\pm$ {\small 0.013} & 0.642 $\pm$ {\small 0.047} & \textbf{0.858 $\pm$ {\small 0.006}} & 0.426 $\pm$ {\small 0.038} & 0.653 $\pm$ {\small 0.013} & 0.68 $\pm$ {\small 0.012} & \textbf{0.861 $\pm$ {\small 0.009}} & 0.941 $\pm$ {\small 0.018} & 0.704 $\pm$ {\small 0.074} \\
+
+			Badge & 0 & 0.767 $\pm$ {\small 0.02} & 0.661 $\pm$ {\small 0.026} & 0.78 $\pm$ {\small 0.014} & 0.642 $\pm$ {\small 0.046} & 0.83 $\pm$ {\small 0.008} & 0.371 $\pm$ {\small 0.035} & 0.656 $\pm$ {\small 0.013} & 0.68 $\pm$ {\small 0.009} & 0.826 $\pm$ {\small 0.024} & 0.941 $\pm$ {\small 0.017} & 0.69 $\pm$ {\small 0.083} \\
+
+			LeastConf & 2 & 0.779 $\pm$ {\small 0.019} & 0.68 $\pm$ {\small 0.032} & 0.809 $\pm$ {\small 0.01} & 0.629 $\pm$ {\small 0.05} & 0.846 $\pm$ {\small 0.009} & 0.421 $\pm$ {\small 0.039} & \textbf{0.668 $\pm$ {\small 0.014}} & \textbf{0.685 $\pm$ {\small 0.009}} & 0.843 $\pm$ {\small 0.013} & 0.94 $\pm$ {\small 0.016} & 0.692 $\pm$ {\small 0.094}\\
+
+			DSA & 3 & 0.766 $\pm$ {\small 0.021} & 0.691 $\pm$ {\small 0.022} & 0.803 $\pm$ {\small 0.01} & \textbf{0.646 $\pm$ {\small 0.032}} & 0.829 $\pm$ {\small 0.01} & \textbf{0.431 $\pm$ {\small 0.05}} & 0.663 $\pm$ {\small 0.014} & 0.679 $\pm$ {\small 0.01} & 0.844 $\pm$ {\small 0.017} & 0.941 $\pm$ {\small 0.014} & \textbf{0.731 $\pm$ {\small 0.032}}\\
+
+			BALD & 1 & \textbf{0.78 $\pm$ {\small 0.014}} & 0.649 $\pm$ {\small 0.04} & 0.784 $\pm$ {\small 0.01} & 0.632 $\pm$ {\small 0.042} & 0.819 $\pm$ {\small 0.01} & 0.242 $\pm$ {\small 0.046} & 0.666 $\pm$ {\small 0.014} & 0.644 $\pm$ {\small 0.018} & 0.815 $\pm$ {\small 0.024} & 0.928 $\pm$ {\small 0.014} & 0.698 $\pm$ {\small 0.043}\\
+
+			CoreGCN & 2 & 0.765 $\pm$ {\small 0.021} & 0.686 $\pm$ {\small 0.023} & 0.804 $\pm$ {\small 0.012} & \textbf{0.646 $\pm$ {\small 0.03}} & 0.753 $\pm$ {\small 0.016} & 0.39 $\pm$ {\small 0.044} & 0.623 $\pm$ {\small 0.018} & 0.647 $\pm$ {\small 0.012} & 0.85 $\pm$ {\small 0.01} & 0.938 $\pm$ {\small 0.014} & \textbf{0.731 $\pm$ {\small 0.028}}\\
+
+			Entropy & 3 & 0.768 $\pm$ {\small 0.022} & \textbf{0.678 $\pm$ {\small 0.035}} & \textbf{0.812 $\pm$ {\small 0.013}} & 0.635 $\pm$ {\small 0.045} & 0.83 $\pm$ {\small 0.011} & 0.399 $\pm$ {\small 0.035} & 0.663 $\pm$ {\small 0.013} & 0.681 $\pm$ {\small 0.011} & 0.815 $\pm$ {\small 0.021} & \textbf{0.942 $\pm$ {\small 0.017}} & 0.696 $\pm$ {\small 0.083}\\
+
+			LSA & 0 & 0.772 $\pm$ {\small 0.016} & 0.68 $\pm$ {\small 0.026} & 0.787 $\pm$ {\small 0.012} & 0.618 $\pm$ {\small 0.036} & 0.821 $\pm$ {\small 0.009} & 0.422 $\pm$ {\small 0.037} & 0.613 $\pm$ {\small 0.014} & 0.642 $\pm$ {\small 0.012} & 0.816 $\pm$ {\small 0.013} & 0.932 $\pm$ {\small 0.016} & 0.727 $\pm$ {\small 0.033}\\
+
+			Random & 0 & 0.76 $\pm$ {\small 0.016} & 0.674 $\pm$ {\small 0.027} & 0.774 $\pm$ {\small 0.013} & 0.63 $\pm$ {\small 0.035} & 0.823 $\pm$ {\small 0.009} & 0.404 $\pm$ {\small 0.036} & 0.613 $\pm$ {\small 0.014} & 0.639 $\pm$ {\small 0.013} & 0.815 $\pm$ {\small 0.012} & 0.933 $\pm$ {\small 0.017} & 0.721 $\pm$ {\small 0.036}\\
+
+			Coreset & 0 & 0.772 $\pm$ {\small 0.016} & 0.69 $\pm$ {\small 0.017} & 0.79 $\pm$ {\small 0.012} & 0.638 $\pm$ {\small 0.041} & 0.767 $\pm$ {\small 0.016} & 0.404 $\pm$ {\small 0.046} & 0.659 $\pm$ {\small 0.011} & 0.684 $\pm$ {\small 0.009} & 0.826 $\pm$ {\small 0.022} & 0.937 $\pm$ {\small 0.014} & 0.73 $\pm$ {\small 0.031}\\
+
+			TypiClust & 0 & 0.762 $\pm$ {\small 0.016} & 0.685 $\pm$ {\small 0.025} & 0.778 $\pm$ {\small 0.01} & 0.663 $\pm$ {\small 0.028} & 0.828 $\pm$ {\small 0.007} & 0.396 $\pm$ {\small 0.046} & 0.653 $\pm$ {\small 0.013} & 0.649 $\pm$ {\small 0.007} & 0.831 $\pm$ {\small 0.011} & 0.934 $\pm$ {\small 0.018} & 0.727 $\pm$ {\small 0.033}\\
+		\end{tabular}
+	}
+\end{table}
+% QS 5
+\begin{table}[H]
+	\caption{AUC values for each dataset that supports query size 5.}
+	\resizebox{\columnwidth}{!}{
+		\begin{tabular}{l|llllllllllll}
+			& Wins& Splice & SpliceEnc & DNA & DNAEnc & USPS & USPSEnc & Cifar10Enc & FMnistEnc & TopV2 & DivergingSin & ThreeClust \\
+			\hline
+			Oracle &  & 0.803 $\pm$ {\small 0.012} & 0.678 $\pm$ {\small 0.021} & 0.825 $\pm$ {\small 0.009} & 0.721 $\pm$ {\small 0.013} & 0.866 $\pm$ {\small 0.004} & 0.436 $\pm$ {\small 0.057} & 0.749 $\pm$ {\small 0.009} & 0.755 $\pm$ {\small 0.005} & 0.884 $\pm$ {\small 0.006} & 0.957 $\pm$ {\small 0.009} & 0.783 $\pm$ {\small 0.03} \\
+
+			Margin & 3 & 0.765 $\pm$ {\small 0.021} & 0.662 $\pm$ {\small 0.032} & 0.794 $\pm$ {\small 0.011} & 0.611 $\pm$ {\small 0.05} & \textbf{0.855 $\pm$ {\small 0.006}} & \textbf{0.508 $\pm$ {\small 0.02}} & 0.656 $\pm$ {\small 0.014} & 0.678 $\pm$ {\small 0.009} & \textbf{0.848 $\pm$ {\small 0.013}} & 0.923 $\pm$ {\small 0.019} & 0.697 $\pm$ {\small 0.055} \\
+
+			Badge & 2 & 0.768 $\pm$ {\small 0.014} & 0.646 $\pm$ {\small 0.035} & 0.785 $\pm$ {\small 0.011} & \textbf{0.624 $\pm$ {\small 0.036}} & 0.846 $\pm$ {\small 0.007} & 0.48 $\pm$ {\small 0.021} & 0.647 $\pm$ {\small 0.012} & 0.67 $\pm$ {\small 0.009} & 0.847 $\pm$ {\small 0.01} & \textbf{0.924 $\pm$ {\small 0.019}} & 0.72 $\pm$ {\small 0.036} \\
+
+			LeastConf & 1 & 0.763 $\pm$ {\small 0.023} & 0.643 $\pm$ {\small 0.034} & 0.798 $\pm$ {\small 0.013} & 0.585 $\pm$ {\small 0.065} & 0.831 $\pm$ {\small 0.014} & 0.478 $\pm$ {\small 0.028} & 0.67 $\pm$ {\small 0.01} & \textbf{0.681 $\pm$ {\small 0.009}} & 0.819 $\pm$ {\small 0.023} & 0.921 $\pm$ {\small 0.019} & 0.675 $\pm$ {\small 0.072} \\
+
+			DSA & 1 & 0.765 $\pm$ {\small 0.023} & 0.653 $\pm$ {\small 0.029} & 0.793 $\pm$ {\small 0.009} & 0.613 $\pm$ {\small 0.034} & 0.822 $\pm$ {\small 0.01} & 0.489 $\pm$ {\small 0.024} & 0.661 $\pm$ {\small 0.013} & 0.662 $\pm$ {\small 0.012} & 0.833 $\pm$ {\small 0.02} & \textbf{0.924 $\pm$ {\small 0.018}} & 0.718 $\pm$ {\small 0.033} \\
+
+			BALD & 4 & \textbf{0.775 $\pm$ {\small 0.018}} & 0.641 $\pm$ {\small 0.034} & \textbf{0.801 $\pm$ {\small 0.013}} & 0.592 $\pm$ {\small 0.054} & 0.84 $\pm$ {\small 0.008} & 0.332 $\pm$ {\small 0.054} & \textbf{0.681 $\pm$ {\small 0.011}} & \textbf{0.681 $\pm$ {\small 0.013}} & 0.824 $\pm$ {\small 0.023} & 0.893 $\pm$ {\small 0.035} & 0.673 $\pm$ {\small 0.041} \\
+
+			CoreGCN & 1 & 0.759 $\pm$ {\small 0.018} & 0.662 $\pm$ {\small 0.027} & 0.79 $\pm$ {\small 0.011} & 0.62 $\pm$ {\small 0.03} & 0.755 $\pm$ {\small 0.011} & 0.45 $\pm$ {\small 0.03} & 0.604 $\pm$ {\small 0.016} & 0.609 $\pm$ {\small 0.013} & 0.837 $\pm$ {\small 0.014} & 0.922 $\pm$ {\small 0.018} & \textbf{0.723 $\pm$ {\small 0.034}} \\
+
+			Entropy & 1 & 0.765 $\pm$ {\small 0.022} & 0.66 $\pm$ {\small 0.03} & 0.798 $\pm$ {\small 0.011} & 0.611 $\pm$ {\small 0.054} & 0.823 $\pm$ {\small 0.013} & 0.464 $\pm$ {\small 0.024} & 0.663 $\pm$ {\small 0.013} & 0.672 $\pm$ {\small 0.011} & 0.801 $\pm$ {\small 0.025} & \textbf{0.924 $\pm$ {\small 0.02}} & 0.689 $\pm$ {\small 0.066} \\
+
+			LSA & 0 & 0.769 $\pm$ {\small 0.016} & 0.654 $\pm$ {\small 0.032} & 0.781 $\pm$ {\small 0.013} & 0.61 $\pm$ {\small 0.041} & 0.82 $\pm$ {\small 0.009} & 0.484 $\pm$ {\small 0.022} & 0.617 $\pm$ {\small 0.012} & 0.641 $\pm$ {\small 0.011} & 0.816 $\pm$ {\small 0.012} & 0.915 $\pm$ {\small 0.018} & 0.718 $\pm$ {\small 0.038} \\
+
+			Random & 0 & 0.758 $\pm$ {\small 0.015} & 0.655 $\pm$ {\small 0.026} & 0.771 $\pm$ {\small 0.013} & 0.623 $\pm$ {\small 0.031} & 0.82 $\pm$ {\small 0.009} & 0.476 $\pm$ {\small 0.024} & 0.616 $\pm$ {\small 0.016} & 0.637 $\pm$ {\small 0.012} & 0.812 $\pm$ {\small 0.014} & 0.921 $\pm$ {\small 0.018} & 0.713 $\pm$ {\small 0.034} \\
+
+			Coreset & 1 & 0.765 $\pm$ {\small 0.017} & \textbf{0.663 $\pm$ {\small 0.023}} & 0.784 $\pm$ {\small 0.014} & 0.603 $\pm$ {\small 0.034} & 0.765 $\pm$ {\small 0.015} & 0.449 $\pm$ {\small 0.022} & 0.657 $\pm$ {\small 0.009} & 0.674 $\pm$ {\small 0.009} & 0.817 $\pm$ {\small 0.017} & 0.92 $\pm$ {\small 0.017} & 0.713 $\pm$ {\small 0.035} \\
+
+			TypiClust & 0 & 0.759 $\pm$ {\small 0.014} & 0.641 $\pm$ {\small 0.028} & 0.775 $\pm$ {\small 0.01} & 0.603 $\pm$ {\small 0.04} & 0.757 $\pm$ {\small 0.02} & 0.465 $\pm$ {\small 0.027} & 0.596 $\pm$ {\small 0.014} & 0.567 $\pm$ {\small 0.012} & 0.727 $\pm$ {\small 0.026} & 0.916 $\pm$ {\small 0.02} & 0.693 $\pm$ {\small 0.045} \\
+		\end{tabular}
+	}
+\end{table}
+% QS 20
+\begin{table}[H]
+	\caption{AUC values for each dataset that supports query size 20.}
+	\resizebox{\columnwidth}{!}{
+		\begin{tabular}{l|llllllllll}
+			& Wins & Splice       & SpliceEnc& DNA          & USPS         & USPSEnc  & Cifar10Enc& FMnistEnc & TopV2 & News \\
+			\hline
+			Oracle &  & 0.803 $\pm$ {\small 0.012} & 0.678 $\pm$ {\small 0.021} & 0.825 $\pm$ {\small 0.009} & 0.866 $\pm$ {\small 0.004} & 0.436 $\pm$ {\small 0.057} & 0.749 $\pm$ {\small 0.009} & 0.755 $\pm$ {\small 0.005} & 0.884 $\pm$ {\small 0.006} & 0.49 $\pm$ {\small 0.003} \\
+
+			Margin & 1 & 0.759 $\pm$ {\small 0.027} & 0.618 $\pm$ {\small 0.04} & 0.779 $\pm$ {\small 0.013} & \textbf{0.847 $\pm$ {\small 0.008}} & 0.439 $\pm$ {\small 0.027} & 0.656 $\pm$ {\small 0.01} & 0.67 $\pm$ {\small 0.011} & 0.823 $\pm$ {\small 0.014} & 0.464 $\pm$ {\small 0.007} \\
+
+			Badge & 3 & 0.767 $\pm$ {\small 0.013} & \textbf{0.619 $\pm$ {\small 0.033}} & 0.776 $\pm$ {\small 0.013} & 0.845 $\pm$ {\small 0.006} & \textbf{0.44 $\pm$ {\small 0.019}} & 0.647 $\pm$ {\small 0.013} & 0.665 $\pm$ {\small 0.007} & \textbf{0.827 $\pm$ {\small 0.016}} & 0.463 $\pm$ {\small 0.007} \\
+
+			LeastConf & 1 & 0.751 $\pm$ {\small 0.02} & 0.597 $\pm$ {\small 0.05} & 0.748 $\pm$ {\small 0.025} & 0.798 $\pm$ {\small 0.027} & 0.391 $\pm$ {\small 0.024} & \textbf{0.665 $\pm$ {\small 0.013}} & 0.669 $\pm$ {\small 0.011} & 0.775 $\pm$ {\small 0.035} & 0.467 $\pm$ {\small 0.008} \\
+
+			DSA & 0 & 0.759 $\pm$ {\small 0.02} & 0.599 $\pm$ {\small 0.034} & 0.769 $\pm$ {\small 0.013} & 0.809 $\pm$ {\small 0.012} & 0.421 $\pm$ {\small 0.023} & 0.647 $\pm$ {\small 0.014} & 0.63 $\pm$ {\small 0.013} & 0.793 $\pm$ {\small 0.026} & 0.459 $\pm$ {\small 0.01} \\
+
+			BALD & 4 & \textbf{0.768 $\pm$ {\small 0.022}} & 0.57 $\pm$ {\small 0.037} & \textbf{0.784 $\pm$ {\small 0.015}} & 0.822 $\pm$ {\small 0.009} & 0.298 $\pm$ {\small 0.039} & 0.675 $\pm$ {\small 0.008} & \textbf{0.673 $\pm$ {\small 0.01}} & 0.789 $\pm$ {\small 0.024} & \textbf{0.468 $\pm$ {\small 0.009}} \\
+
+			CoreGCN & 0 & 0.759 $\pm$ {\small 0.018} & 0.612 $\pm$ {\small 0.039} & 0.774 $\pm$ {\small 0.012} & 0.754 $\pm$ {\small 0.016} & 0.397 $\pm$ {\small 0.026} & 0.587 $\pm$ {\small 0.015} & 0.583 $\pm$ {\small 0.015} & 0.807 $\pm$ {\small 0.018} & 0.453 $\pm$ {\small 0.006} \\
+
+			Entropy & 0 & 0.759 $\pm$ {\small 0.027} & 0.618 $\pm$ {\small 0.038} & 0.773 $\pm$ {\small 0.015} & 0.803 $\pm$ {\small 0.019} & 0.372 $\pm$ {\small 0.022} & 0.656 $\pm$ {\small 0.011} & 0.65 $\pm$ {\small 0.012} & 0.773 $\pm$ {\small 0.031} & 0.451 $\pm$ {\small 0.007} \\
+
+			LSA & 0 & 0.761 $\pm$ {\small 0.014} & 0.611 $\pm$ {\small 0.039} & 0.768 $\pm$ {\small 0.015} & 0.816 $\pm$ {\small 0.009} & 0.411 $\pm$ {\small 0.022} & 0.621 $\pm$ {\small 0.01} & 0.635 $\pm$ {\small 0.011} & 0.796 $\pm$ {\small 0.016} & 0.452 $\pm$ {\small 0.007} \\
+
+			Random & 0 & 0.755 $\pm$ {\small 0.014} & 0.612 $\pm$ {\small 0.039} & 0.763 $\pm$ {\small 0.012} & 0.818 $\pm$ {\small 0.009} & 0.439 $\pm$ {\small 0.019} & 0.622 $\pm$ {\small 0.013} & 0.633 $\pm$ {\small 0.012} & 0.795 $\pm$ {\small 0.016} & 0.45 $\pm$ {\small 0.006} \\
+
+			Coreset & 0 & 0.759 $\pm$ {\small 0.016} & 0.601 $\pm$ {\small 0.034} & 0.764 $\pm$ {\small 0.015} & 0.757 $\pm$ {\small 0.015} & 0.39 $\pm$ {\small 0.029} & 0.647 $\pm$ {\small 0.009} & 0.651 $\pm$ {\small 0.011} & 0.784 $\pm$ {\small 0.026} & 0.435 $\pm$ {\small 0.012} \\
+
+			TypiClust & 0 & 0.751 $\pm$ {\small 0.012} & 0.551 $\pm$ {\small 0.036} & 0.76 $\pm$ {\small 0.016} & 0.643 $\pm$ {\small 0.026} & 0.411 $\pm$ {\small 0.024} & 0.488 $\pm$ {\small 0.02} & 0.449 $\pm$ {\small 0.017} & 0.652 $\pm$ {\small 0.035} & 0.406 $\pm$ {\small 0.011} \\
+		\end{tabular}
+	}
+\end{table}
+% QS 50
+\begin{table}[H]
+	\caption{AUC values for each dataset that supports query size 50.}
+	\resizebox{\columnwidth}{!}{
+		\begin{tabular}{l|lllllllll}
+			& Wins & Splice       & DNA          & USPS         & USPSEnc  & Cifar10Enc& FMnistEnc& TopV2     & News \\
+			\hline
+			Oracle &  & 0.803 $\pm$ {\small 0.012} & 0.825 $\pm$ {\small 0.009} & 0.866 $\pm$ {\small 0.004} & 0.436 $\pm$ {\small 0.057} & 0.749 $\pm$ {\small 0.009} & 0.755 $\pm$ {\small 0.005} & 0.884 $\pm$ {\small 0.006} & 0.49 $\pm$ {\small 0.003} \\
+
+			Margin & 1 & 0.747 $\pm$ {\small 0.023} & 0.751 $\pm$ {\small 0.019} & 0.828 $\pm$ {\small 0.009} & 0.363 $\pm$ {\small 0.031} & 0.64 $\pm$ {\small 0.013} & \textbf{0.653 $\pm$ {\small 0.01}} & 0.774 $\pm$ {\small 0.029} & 0.46 $\pm$ {\small 0.006} \\
+
+			Badge & 3 & 0.758 $\pm$ {\small 0.017} & 0.754 $\pm$ {\small 0.018} & \textbf{0.831 $\pm$ {\small 0.008}} & 0.376 $\pm$ {\small 0.028} & 0.632 $\pm$ {\small 0.013} & 0.649 $\pm$ {\small 0.011} & \textbf{0.781 $\pm$ {\small 0.026}} & \textbf{0.462 $\pm$ {\small 0.007}} \\
+
+			LeastConf & 2 & 0.731 $\pm$ {\small 0.025} & 0.688 $\pm$ {\small 0.041} & 0.761 $\pm$ {\small 0.037} & 0.291 $\pm$ {\small 0.03} & \textbf{0.644 $\pm$ {\small 0.013}} & 0.65 $\pm$ {\small 0.011} & 0.73 $\pm$ {\small 0.049} & \textbf{0.462 $\pm$ {\small 0.009}} \\
+
+			DSA & 0 & 0.748 $\pm$ {\small 0.021} & 0.738 $\pm$ {\small 0.018} & 0.783 $\pm$ {\small 0.016} & 0.346 $\pm$ {\small 0.027} & 0.624 $\pm$ {\small 0.014} & 0.588 $\pm$ {\small 0.016} & 0.748 $\pm$ {\small 0.041} & 0.45 $\pm$ {\small 0.011} \\
+
+			BALD & 2 & \textbf{0.76 $\pm$ {\small 0.017}} & \textbf{0.756 $\pm$ {\small 0.018}} & 0.796 $\pm$ {\small 0.016} & 0.241 $\pm$ {\small 0.026} & 0.65 $\pm$ {\small 0.009} & 0.645 $\pm$ {\small 0.01} & 0.746 $\pm$ {\small 0.038} & 0.455 $\pm$ {\small 0.007} \\
+
+			CoreGCN & 0 & 0.755 $\pm$ {\small 0.016} & 0.745 $\pm$ {\small 0.018} & 0.752 $\pm$ {\small 0.019} & 0.328 $\pm$ {\small 0.027} & 0.581 $\pm$ {\small 0.015} & 0.568 $\pm$ {\small 0.018} & 0.771 $\pm$ {\small 0.025} & 0.453 $\pm$ {\small 0.007} \\
+
+			Entropy & 0 & 0.747 $\pm$ {\small 0.024} & 0.748 $\pm$ {\small 0.018} & 0.778 $\pm$ {\small 0.024} & 0.275 $\pm$ {\small 0.026} & 0.633 $\pm$ {\small 0.011} & 0.625 $\pm$ {\small 0.012} & 0.734 $\pm$ {\small 0.036} & 0.442 $\pm$ {\small 0.007} \\
+
+			LSA & 0 & 0.754 $\pm$ {\small 0.013} & 0.749 $\pm$ {\small 0.019} & 0.807 $\pm$ {\small 0.01} & 0.341 $\pm$ {\small 0.029} & 0.613 $\pm$ {\small 0.012} & 0.625 $\pm$ {\small 0.01} & 0.763 $\pm$ {\small 0.025} & 0.45 $\pm$ {\small 0.006} \\
+
+			Random & 1 & 0.746 $\pm$ {\small 0.012} & 0.745 $\pm$ {\small 0.015} & 0.806 $\pm$ {\small 0.008} & \textbf{0.379 $\pm$ {\small 0.028}} & 0.615 $\pm$ {\small 0.014} & 0.621 $\pm$ {\small 0.01} & 0.759 $\pm$ {\small 0.026} & 0.448 $\pm$ {\small 0.006} \\
+
+			Coreset & 0 & 0.751 $\pm$ {\small 0.016} & 0.733 $\pm$ {\small 0.019} & 0.74 $\pm$ {\small 0.017} & 0.325 $\pm$ {\small 0.034} & 0.624 $\pm$ {\small 0.012} & 0.608 $\pm$ {\small 0.013} & 0.731 $\pm$ {\small 0.045} & 0.432 $\pm$ {\small 0.012} \\
+
+			TypiClust & 0 & 0.749 $\pm$ {\small 0.016} & 0.736 $\pm$ {\small 0.016} & 0.586 $\pm$ {\small 0.038} & 0.348 $\pm$ {\small 0.027} & 0.451 $\pm$ {\small 0.024} & 0.375 $\pm$ {\small 0.022} & 0.614 $\pm$ {\small 0.046} & 0.397 $\pm$ {\small 0.012} \\
+		\end{tabular}
+	}
+\end{table}
+% QS 100
+\begin{table}[H]
+	\caption{AUC values for each dataset that supports query size 100.}
+	\resizebox{\columnwidth}{!}{
+		\begin{tabular}{l|llllllll}
+			& Wins & Splice       & DNA          & USPS         & USPSEnc & Cifar10Enc& FMnistEnc& News \\
+			\hline
+			Oracle & 0 & 0.803 $\pm$ {\small 0.012} & 0.825 $\pm$ {\small 0.009} & 0.866 $\pm$ {\small 0.004} & 0.436 $\pm$ {\small 0.057} & 0.749 $\pm$ {\small 0.009} & 0.755 $\pm$ {\small 0.005} & 0.49 $\pm$ {\small 0.003} \\
+			Margin & 2 & 0.733 $\pm$ {\small 0.024} & 0.711 $\pm$ {\small 0.027} & 0.799 $\pm$ {\small 0.013} & \textbf{0.473 $\pm$ {\small 0.026}} & 0.629 $\pm$ {\small 0.012} & \textbf{0.628 $\pm$ {\small 0.009}} & 0.455 $\pm$ {\small 0.006} \\
+			Badge & 2 & 0.743 $\pm$ {\small 0.014} & 0.714 $\pm$ {\small 0.032} & \textbf{0.804 $\pm$ {\small 0.013}} & 0.472 $\pm$ {\small 0.029} & 0.623 $\pm$ {\small 0.01} & 0.621 $\pm$ {\small 0.01} & \textbf{0.456 $\pm$ {\small 0.006}} \\
+			LeastConf & 0 & 0.715 $\pm$ {\small 0.033} & 0.639 $\pm$ {\small 0.05} & 0.708 $\pm$ {\small 0.034} & 0.23 $\pm$ {\small 0.034} & 0.631 $\pm$ {\small 0.013} & 0.62 $\pm$ {\small 0.012} & 0.457 $\pm$ {\small 0.008} \\
+			DSA & 0 & 0.729 $\pm$ {\small 0.021} & 0.697 $\pm$ {\small 0.031} & 0.753 $\pm$ {\small 0.021} & 0.427 $\pm$ {\small 0.028} & 0.609 $\pm$ {\small 0.013} & 0.546 $\pm$ {\small 0.017} & 0.442 $\pm$ {\small 0.01} \\
+			BALD & 3 & \textbf{0.744 $\pm$ {\small 0.015}} & \textbf{0.718 $\pm$ {\small 0.024}} & 0.765 $\pm$ {\small 0.021} & 0.285 $\pm$ {\small 0.046} & \textbf{0.632 $\pm$ {\small 0.009}} & 0.609 $\pm$ {\small 0.01} & 0.444 $\pm$ {\small 0.007} \\
+			CoreGCN & 0 & 0.742 $\pm$ {\small 0.015} & 0.713 $\pm$ {\small 0.025} & 0.744 $\pm$ {\small 0.019} & 0.433 $\pm$ {\small 0.032} & 0.583 $\pm$ {\small 0.013} & 0.554 $\pm$ {\small 0.015} & 0.448 $\pm$ {\small 0.007} \\
+			Entropy & 0 & 0.733 $\pm$ {\small 0.023} & 0.713 $\pm$ {\small 0.031} & 0.743 $\pm$ {\small 0.026} & 0.395 $\pm$ {\small 0.037} & 0.618 $\pm$ {\small 0.012} & 0.59 $\pm$ {\small 0.012} & 0.432 $\pm$ {\small 0.007} \\
+			LSA & 0 & 0.738 $\pm$ {\small 0.017} & 0.716 $\pm$ {\small 0.027} & 0.789 $\pm$ {\small 0.011} & 0.439 $\pm$ {\small 0.03} & 0.609 $\pm$ {\small 0.013} & 0.608 $\pm$ {\small 0.01} & 0.447 $\pm$ {\small 0.006} \\
+			Random & 0 & 0.733 $\pm$ {\small 0.013} & 0.713 $\pm$ {\small 0.023} & 0.789 $\pm$ {\small 0.012} & 0.468 $\pm$ {\small 0.024} & 0.611 $\pm$ {\small 0.01} & 0.606 $\pm$ {\small 0.01} & 0.446 $\pm$ {\small 0.005} \\
+			Coreset & 0 & 0.735 $\pm$ {\small 0.019} & 0.698 $\pm$ {\small 0.026} & 0.721 $\pm$ {\small 0.021} & 0.396 $\pm$ {\small 0.024} & 0.608 $\pm$ {\small 0.012} & 0.562 $\pm$ {\small 0.016} & 0.426 $\pm$ {\small 0.012} \\
+			TypiClust & 0 & 0.733 $\pm$ {\small 0.016} & 0.704 $\pm$ {\small 0.025} & 0.592 $\pm$ {\small 0.042} & 0.427 $\pm$ {\small 0.027} & 0.501 $\pm$ {\small 0.02} & 0.338 $\pm$ {\small 0.02} & 0.383 $\pm$ {\small 0.012} \\
+		\end{tabular}
+	}
+\end{table}
+\begin{minipage}{0.47\linewidth}
+	% QS 500
+	\begin{table}[H]
+		\caption{AUC values for each dataset that supports query size 500.}
+		\begin{tabular}{l|ll}
+			& Cifar10 & FashionMnist \\
+			\hline
+			Oracle & 0.689 $\pm$ {\small 0.001} & 0.905 $\pm$ {\small 0.001} \\
+			Margin & 0.556 $\pm$ {\small 0.008} & 0.882 $\pm$ {\small 0.004} \\
+			Badge & 0.56 $\pm$ {\small 0.008} & 0.883 $\pm$ {\small 0.005} \\
+			LeastConf & \textbf{0.591 $\pm$ {\small 0.01}} & \textbf{0.884 $\pm$ {\small 0.005}} \\
+			DSA & 0.56 $\pm$ {\small 0.009} & 0.882 $\pm$ {\small 0.004} \\
+			BALD & 0.478 $\pm$ {\small 0.014} & 0.878 $\pm$ {\small 0.003} \\
+			CoreGCN & 0.553 $\pm$ {\small 0.01} & 0.88 $\pm$ {\small 0.007} \\
+			Entropy & 0.553 $\pm$ {\small 0.009} & 0.882 $\pm$ {\small 0.006} \\
+			LSA & 0.558 $\pm$ {\small 0.01} & 0.866 $\pm$ {\small 0.005} \\
+			Random & 0.557 $\pm$ {\small 0.01} & 0.863 $\pm$ {\small 0.005} \\
+			Coreset & 0.553 $\pm$ {\small 0.007} & 0.878 $\pm$ {\small 0.006} \\
+			TypiClust & 0.557 $\pm$ {\small 0.009} & 0.864 $\pm$ {\small 0.004}
+		\end{tabular}
+	\end{table}
+\end{minipage}
+\hspace{10mm}
+\begin{minipage}{0.47\linewidth}
+	% QS 1000
+	\begin{table}[H]
+		\caption{AUC values for each dataset that supports query size 1000.}
+		\begin{tabular}{l|ll}
+			& Cifar10 & FashionMnist \\
+			\hline
+			Oracle & 0.689 $\pm$ {\small 0.001} & 0.905 $\pm$ {\small 0.001} \\
+			Margin & 0.56 $\pm$ {\small 0.011} & 0.872 $\pm$ {\small 0.007} \\
+			Badge & \textbf{0.562 $\pm$ {\small 0.013}} & 0.871 $\pm$ {\small 0.007} \\
+			LeastConf & 0.561 $\pm$ {\small 0.012} & \textbf{0.873 $\pm$ {\small 0.006}} \\
+			DSA & 0.56 $\pm$ {\small 0.011} & 0.87 $\pm$ {\small 0.008} \\
+			BALD & 0.535 $\pm$ {\small 0.011} & 0.866 $\pm$ {\small 0.003} \\
+			CoreGCN & 0.557 $\pm$ {\small 0.011} & 0.867 $\pm$ {\small 0.012} \\
+			Entropy & 0.557 $\pm$ {\small 0.014} & 0.871 $\pm$ {\small 0.009} \\
+			LSA & 0.551 $\pm$ {\small 0.012} & 0.854 $\pm$ {\small 0.009} \\
+			Random & 0.55 $\pm$ {\small 0.01} & 0.855 $\pm$ {\small 0.006} \\
+			Coreset & 0.562 $\pm$ {\small 0.012} & 0.869 $\pm$ {\small 0.004} \\
+			TypiClust & 0.552 $\pm$ {\small 0.011} & 0.854 $\pm$ {\small 0.009} 	
+		\end{tabular}
+	\end{table}
+\end{minipage}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Analysis of Results for the Semi-Supervised Domain}\label{app:semi_sup_analysis}
+Even though the results for the aggregated semi-supervised domain appear in line with our overall ranking of methods, we observe stark differences for the sub-domains of semi-supervised image and semi-supervised tabular.
+\\
+While semi-supervised images seem to mostly mirror the results from the normal image domain (with the exception of BALD), semi-supervised tabular data display highly irregular behavior, placing random sampling as second-best method behind the margin sampling.
+Our oracle method even falls behind other methods.
+Almost all methods a bunched into one region in the CD diagram with many non-significance bars indicating few, if any, significant differences between the methods. \\
+Both the reasons, for the sub-random performance of most methods, and the bad performance of our oracle are currently unknown and require further research.
+\begin{figure}[H]
+	\centering
+	\caption{Results for the semi-supervised domain, aggregated over all data types (top) and separately for images and tabular (bottom)}
+	\includegraphics[width=0.6\linewidth]{img/macro_enc.jpg}
+	\includegraphics[width=0.49\linewidth]{img/macro_enc_img.jpg}
+	\includegraphics[width=0.49\linewidth]{img/macro_enc_tabular.jpg}
+\end{figure}
+
+\section{Hyperparameters per AL Method}\label{app:agent_hyperparameters}
+\begin{table}[H]
+    \caption{Selected hyperparameters for all tested AL methods. Last column indicates the source of our implementation.}
+	\centering
+	\begin{tabular}{l || l | l | l}
+		Method & \makecell{Sample\\Size} & Other & Source\\
+		\hline
+		BADGE & 100 & & Based on \cite{ashdeep, curelab}\\
+		BALD & 100 & Dropout Trials: 5 & Based on \cite{pycls} \\
+		Coreset & 8000 & & Own \\
+		TypiClust & 10000 & \makecell[tl]{Min Cluster Size: 5\\Max \# Clusters: 500} & Based on \cite{hacohen2022active} \\
+		Margin & 8000 & & Own\\
+		Entropy & 8000 &  & Own\\
+	\end{tabular}
+\end{table}
+
+\section{Amount of Computational Resources Invested}\label{app:compute}
+For our results we computed a total of 24200 runs (without Oracle runs) over a span of ~4 months. \\
+We used our computational cluster consisting of 30-40 GPUs. \\ [1mm]
+Number of runs per dataset and query size : 11 alg. * 50 runs = 550 \\ [1mm]
+Runs per dataset: \\
+ Cifar10   x2 = 1100\\
+ Cfr10Enc  x4 = 2200\\
+ DivSin    x2 = 1100\\
+ DNA       x4 = 2200\\
+ DNAEnc    x2 = 1100\\
+ FMnist    x2 = 1100\\
+ FMnistEnc x4 = 2200\\
+ News      x3 = 1650\\
+ Splice    x4 = 2200\\
+ SpliceEnc x3 = 1650\\
+ Honeypot  x2 = 1100\\
+ TopV2     x4 = 2200\\
+ USPS      x4 = 2200\\
+ USPSEnc   x4 = 2200\\
+ = 24200 runs
+
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ \section{Limitations and Future Work}\label{app:future_work}
+ Even though our benchmark includes a wide range of data domains, the number of datasets per domain is still limited.
+ It remains untested if our selected datasets are indeed a good representation of their domain, or if additional datasets would skew the results of Fig. \ref{fig:ranks_by_domain}. \\ 
+ Additionally, since we began working on this benchmark a few new AL Methods have been published.
+ We consciously focused on only those methods for which good results have been reported by multiple sources, consequently omitting the newest methods. \\
+ Most obviously, our future work involves the implementation of more datasets per domain and the newest AL methods. \\ [1mm]
+ The choice of ranks for the main result table, like any other choice, has an impact in the interpretation of the results.
+ E.g.
+comparing the rank of BALD in Table \ref{tab:results} (rank 6.6) with the amount of wins it is able to obtain in the AUC-based tables in Appendix \ref{app:auc_by_query_size}, suggests that an evaluation purely based on mean AUC values would count BALD to the best methods.
+\\
+ We advocate for using ranks in AL evaluations for two reasons: (i) they are more robust to outlier performances in single runs and (ii) they highlight wether an method is able to consistently outperform another method, even if the difference in mean AUC is very small.
+\\
+ Nonetheless, we think that the topic of truly fair evaluations for AL needs further research. \\ [1mm]
+ Lastly, it remains untested if the differences between domains that we observe, are truly caused by the differences in data, but could also be influenced by the type of model that is common to those domains.
+\\
+ An important part of our future work is therefore to test different model archetypes per domain.
 

### 2024-06-05 doc/al_benchmark_paper.tex
paper version 1.0

@@
 Additionally, we show that having a large amount of runs is crucial. With only
-conducting five runs as often done in the literature, the superiority of
+conducting three runs as often done in the literature, the superiority of
 specific methods can strongly vary with the specific runs.
@@
 multiple application domains and reports a large amount of runs per
-experiment, so that the significance of performance differences can can be
+experiment, so that the significance of performance differences can be
 estimated. Specifically, \emph{CDALBench} consists of datasets from
@@
 Especially the tabular domain is underrepresented in preceding benchmarks, as only \cite{zhan2022comparative} provides experiments for it.
-Experiments on the interplay between AL and semi-supervised learning are similarly under-researched, as only two works exist \cite{li2022empirical, luth2024navigating}, both of them only using images.
+The interplay between AL and semi-supervised learning is similarly under-researched, as only two works exist \cite{li2022empirical, luth2024navigating}, both of them only using images.
 An oracle algorithm has been proposed by two works \cite{zhou2021towards, zhan2022comparative}. 
@@
 However, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), providing a bad approximation of the quality of an AL method (i.e.
-measured performances for an method might not even transfer to a different model initialization).
+measured performances for an AL method might not even transfer to a different model initialization).
 Hence, we opt for letting the subsystems vary in controlled way (For details, please refer to App.
@@
 			\hline
-			Enc. DNA    & Linear & 40 & o & o &&&&& \\
-			Enc. Splice & Linear & 100 & o & o & o & o &&&\\
+			Semi DNA    & Linear & 40 & o & o &&&&& \\
+			Semi Splice & Linear & 100 & o & o & o & o &&&\\
 			TopV2       & BiLSTM & 200 & o & o & o & o &&& \\
@@
 			USPS        & MLP    & 400 & o & o & o & o & o && \\
-			Enc. Cifar10& Linear & 450 & o & o & o & o & o && \\
-			Enc. FMnist & Linear & 500 & o & o & o & o & o && \\
-			Enc. USPS   & Linear & 600 & o & o & o & o & o && \\
+			Semi Cifar10& Linear & 450 & o & o & o & o & o && \\
+			Semi FMnist & Linear & 500 & o & o & o & o & o && \\
+			Semi USPS   & Linear & 600 & o & o & o & o & o && \\
 			News        & BiLSTM & 3K &&& o & o & o & o &\\
@@
 We would like to point out that these datasets are selected for speed of computation (both in terms of the required classifier and the necessary budget to solve the dataset). 
-However, we are solely focused on comparing different AL methods in this paper and do not aim to develop novel classification models on these datasets.
-Our assumption is that a well-performing method in our benchmark will also generalize well to larger real-world datasets, because we included multiple different data domains, classifier types and sizes in our experiments.
+We are solely focused on comparing different AL methods in this paper and do not aim to develop novel classification models on these datasets.
+Our assumption is that a well-performing method in our benchmark will also generalize well to larger datasets and classifiers, because we included multiple different data domains, classifier types and sizes in our experiments.
 \\ [1mm]
-Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning (semi-supervised).
+Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded (semi-supervised) by a fixed embedding model that was trained by unsupervised contrastive learning .
 The text datasets are an exception to this, as they are only offered in their encoded form.
@@
 The checkpoint of each encoder model will be provided together with the framework. \\ [1mm]
-Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.
+Every dataset has a fixed size for the seed set $\LL^{(0)}$ of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.
 
@@
 The resulting budgets can be found in Tab. \ref{tab:batch_sizes}. \\ [1mm]
-As described in Sec. \ref{sec:restarts}, we repeat each experiment multiple times.
-Each repetition retains the train/test split (often given by the dataset itself), but creates a new validation split that is sampled (based on $s_\D$) from the entire dataset (not just the seed set $\LL^{(0)}$).
+As described in Sec. \ref{sec:restarts}, we repeat each experiment 50 times.
+Each repetition retains the train/test split (often given by the dataset itself), but creates a new validation split that is sampled from the entire dataset (not just the seed set $\LL^{(0)}$).
 \\ [1mm]
@@
 \section{A Greedy Oracle Algorithm}\label{sec:oracle} 
-Using additional resources, like excessive training time, or direct access to the labeled test set, an oracle method for AL finds the oracle set $\mathcal{O}_b$ for a given dataset, model, and
+Using additional resources, like excessive training time, or direct access to a labeled test set, an oracle method for AL finds the oracle set $\mathcal{O}_b$ for a given dataset, model, and
 training procedure that induces the highest AUC score for a given budget.
-However, due to the combinatorial nature of the problem, it is computationally
+However, due to the combinatorial nature of the problem, this is computationally
 infeasible for realistic datasets. Hence, previous works have proposed
 approximations to this oracle sequence. \cite{zhou2021towards} used simulated
-annealing to search for a subset with maximal test accuracy and used the best solution found
+annealing to search for a subset with maximal test accuracy and used the best solution
 after a fixed time budget. 
@@
 Our oracle algorithm
-uniformly samples at iteration $i$ a subset $\mathcal{U}_S$ of the not already
+uniformly samples at iteration $i$ a subset $\mathcal{U}_S$ of size $\tau$ of the not already
 labeled data points $\mathcal{U}^{(i)}$.
 Then it recovers the label $y$ for each of the sampled $u \in \mathcal{U}_S$ and selects the
-point $u$ for which the classifier $\hat{y}$ trained on
+point $u$ for which the classifier $\hat{y}^{(i)}$ trained on
 $\mathcal{L}^{(i)} \cup \{u\}$ has maximal performance.
@@
 Due to the algorithms greedy nature (considering only the next point to pick),
-our oracle frequently encounters situations where every point in $u_k$ would
+our oracle frequently encounters situations where every point in $u$ would
 incur a negative lift (worsening the test performance). This can happen, for
@@
 does not find an unlabeled point that results in an increase in performance, it
-defaults to margin sampling in that iteration. The resulting greedy algorithm
+defaults to margin sampling from the entire unlabeled pool $\mathcal{U}^{(i)}$ in that iteration.
+The resulting greedy algorithm
 constructs an approximation of the optimal labeled set that consistently
@@
 \subsection{Implementation Details}\label{sec:implementation_details}
-At each iteration $i$ the AL method $\Omega$ picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \hat y^{(i)}, \text{opt}_{\hat y}\}$, where $\text{opt}_{\hat y}$ is the optimizer used to fit $\hat y^{(i)}$.
+At each iteration $i$ the AL method picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \hat y^{(i)}, \text{opt}_{\hat y}\}$, where $\text{opt}_{\hat y}$ is the optimizer used to fit $\hat y^{(i)}$.
 This set grants full access to the labeled and unlabeled set, as well as all parameters of the classifier and the optimizer.
@@
 However, the method may not incorporate external information e.g. other datasets, queries to recover additional labels, additional training steps for $\hat y$, or the test/validation set. \\
-For our study we selected acquisition functions with good performances reported by multiple different sources that can work with the set of information stated above.
+For our study we selected AL methods with good performances reported by multiple different sources that can work with the set of information stated above.
 For a list of all AL methods, please refer to Table \ref{tab:results}, with detailed descriptions being found in Appendix \ref{app:acquisition_functions}. \\ [1mm]
 %
-The model $\hat y$ can be trained in two ways.
+The model $\hat y^{(i)}$ can be trained in two ways.
 Either the parameters of the model are reset to a fixed initial setting $\hat y^{(0)}$ after each AL iteration and the classifier is trained from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or the previous state $\hat y^{(i-1)}$ is retained and the classifier is fine-tuned on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
-In this work, we use the fine-tuning method for un-encoded datasets to save computational time, while we use the from-scratch training for embedded datasets since they have very small classifiers and this approach generally produces better results.
+In this work, we use the fine-tuning method for un-encoded datasets to save computational time, while we use the from-scratch training for encoded datasets since they have very small classifiers and this approach generally produces better results.
 Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping with a patience of 2 afterwards.
@@
 In Table \ref{tab:results} we provide the rank of each AL method per dataset. 
-Please note, that we are averaging not only over runs, but also over query sizes per dataset. 
-This leads to a negative impact for AL methods that do not adapt well to larger query sizes.
+Please note, that we are averaging not only over runs, but also over query sizes per dataset, impacting AL methods that do not adapt well to a wide range of query sizes.
 For the results per query size, please refer to App. \ref{app:auc_by_query_size}. 
-As mentioned in contribution \ref{cont:domains}, our results on real-world data show significant differences in the performance of methods between data domains.
+As mentioned in contribution \ref{cont:domains}, our results on real-world data show significant differences in the performance of methods between data domains:
 Not only do some methods overperform on some domains (like least confidence (LC) sampling on images), but the Top-3 of methods (except oracle) does not contain the same three methods for \textbf{any} two domains.
@@
 This highlights the dire need for diverse data domains in AL benchmarking. \\
-Results for the semi-supervised domain appear mostly in line with the other 3 domains, but a closer analysis of performances split into semi-sup.
-images and semi-sup.
-tabular reveals the need for further research.
+Results for the semi-supervised domain appear mostly in line with the other 3 domains, but a closer analysis of performances split into encoded images and encoded tabular reveals the need for further research.
 For details, please refer to App. \ref{app:semi_sup_analysis}.
@@
 Typical members of the second category are clustering approaches like Coreset \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
-Both types of methods have principled shortcomings in terms of the utilized information that makes them unsuitable for certain data distributions. 
+Both types of methods have principled shortcomings in terms of their utilized information that makes them unsuitable for certain data distributions. 
 To test for these specific shortcomings, we created two synthetic datasets, namely ``Honeypot" and ``Diverging Sine", that are hard to solve for methods focused on the classifier's decision boundary or data clustering respectively.
-To avoid methods memorizing these datasets they are generated from scratch for each experiment, depending on $s_\D$. \\
+To avoid methods memorizing these datasets, they are generated from scratch for each experiment. \\
 \begin{figure}[]
@@
 %This honeypot contains 150 samples of each class, creating a balance of 50\% beneficial samples and 50\% adverse samples in the dataset.
-The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its negative impact on purely uncertainty-based AL methods.
-Diverging Sine samples the datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.
+The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its adverse impact on purely uncertainty-based AL methods.
+Diverging Sine samples datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.
 This creates a challenging region on the left hand side, where a lot of datapoints need to be sampled, and an easy region on the right hand side, where very few datapoints are sufficient. 
@@
 In addition, BADGE is underperforming for this dataset compared to real-world data. 
-Both margin sampling and BADGE (the two best methods) being vulnerable to adverse samples or simply measurement noise, highlights the need for further research in this area. \\
+Both margin sampling and BADGE (the two best methods) being vulnerable to adverse samples or simply measurement noise, highlights the need for further research into robust AL methods. \\
 %

### 2024-06-06 doc/al_benchmark_paper.tex
Remove acknowledgement, it breaks page limit
anonymous github link

@@
 we believe that \emph{CDALBench} is a major step forward of assessing the overall state of AL research, independent of
-specific application domains. \emph{CDALBench} is publicly available under \url{https://github.com/wernerth94/A-Cross-Domain-Benchmark-for-Active-Learning/}. \\ [1mm]
+specific application domains. \emph{CDALBench} is publicly available under %\url{https://github.com/wernerth94/A-Cross-Domain-Benchmark-for-Active-Learning/}. 
+\url{https://anonymous.4open.science/r/A-Cross-Domain-Benchmark-for-Active-Learning-CCB1/README.md}
+\\ [1mm]
 Our contributions include the following:
@@
 
-\newpage
-\paragraph{Acknowledgement}
-anonymous
+% \newpage
+% \paragraph{Acknowledgement}
+% anonymous
 %TODO

### 2024-08-13 doc/rebuttal.md
rebuttal

# Reviewer iCXF
rating: 7, confidence 3
```
Thank you for your review...
```
The results on the synthetic benchmarks are unsurprising.
The algorithm for the greedy oracle has no clear guarantees, the design choices are unclear and the description is hard to follow.
The evaluation is done with relatively small models and datasets and it is unclear whether it translates to larger ones.

# Reviewer 3W2R
rating: 7, confidence 3
```
Thank you for your review...
```
However this work should probably cite and discuss Zhang et al.
```
We will add Zhang et al to Table 1 and our Related Work
```

The "oracle" needs to be explained earlier. It is discussed and referenced a lot, but you don't actually describe what it is until section 6. I was very confused, since typically "oracle" might also refer to an annotator. Things made sense at section 6, but a preview of what exactly the oracle does should be discussed as soon as possible. Looking again, it is discussed at line 53 but I actually found this paragraph very vague.
```
The entire section explaining the oracle will be revised for the camera ready version, apart from introducing the concept earlier in the paper, we aim to improve the understanding of readers without prior knowledge.
```

what is "sngl" in Table 1? Single point sampling per round?
```
Yes, we will improve the clarity in Table 1
```

what is 9(14) in Table 1 in the last row?
```
We have 9 datasets in our benchmark, 5 of whichhave a pre-encoded version (excluding text and synthetic) , which brings the total number of experiments to 14.
```

not sure what i\in means really after line 93
```
It simply indexes all points in the unlabeled set.
We will improve the clarity here as well
```

I think Figure 1 needs better descriptions and a better legend. I think I get the main point but it's confusing. E.g., "True mean" is vague, what is that the true mean of? Purple curve doesn't even appear on the legend.
```
Similar to the explanations about the oracle, we aim to revise this section to improve the understanding of readers without prior knowledge.
```

the choice of validation on the entire dataset needs more discussion (line 194). This is a huge criticism of active learning research, to choose parameters based on a full validation set. I understand the argument for lower variance in research evaluation, but I don't think the justification here is sufficient. In particular, Figure 1 argues that the high variance in research results is a problem. Why mask it, with an unrealistic validation? Won't that make things worse?
```
@Max

```

# Reviewer x7yL
rating: 4, confidence 5

```
Thank you for your review...
```
Although the author states that the proposed benchmark is the first one that contains multiple domains, however, in each domain, many related empirical studies focus on specific domain, e.g., [r1], [r2], [r3], which weakens the contribution of this paper.
```
We will include [r1] into Table 1 and our related work ([r2] and [r3] do not provide experiments on their own).
AL benchmarks are still riddled with comparability issues, because their choice of datasets, domains and models ([r2] Sec. 4.1 and [r3] Tab. 1) often is completely disjoint.
In our opinion, we need a benchmark that spans all domains and the most commonly used datasets and models to fix this problem.
As described in our Section 5, we selected our datasets for maximum intersection with previous works. The same is true for our model selection, which either uses common models like ResNet18 or trivial to reproduce models like MLPs.
[r2]: Zhang, Zhisong, Emma Strubell, and Eduard Hovy. "A survey of active learning for natural language processing." arXiv preprint arXiv:2210.10109 (2022).
[r3] Schröder, Christopher, and Andreas Niekler. "A survey of active learning for text classification using deep neural networks." arXiv preprint arXiv:2008.07267 (2020).
```

The insights (e.g., the importance of multiple runs) build on existing knowledge in the active learning field.
```
Even though it is common knowledge that repeating an experiment enough times is crucial for ML research, in the field of Active Learning we still observe many authors not doing so. 
To the best of our knowledge, we are the first to provide a first insight into exactly how often an AL experiment should be repeated in order to generate reliable results.
The same is true for experimenting on different domains. Everyone knows that algorithms should be tested in as many domains as possible, but few actually apply that knowledge.
```

This paper takes too much space on the greedy oracle algorithm and seeding strategies, which may be complex for some readers to fully get the meaning.
```
We will revise the entire section about our oracle algorithm and work with more collegues from outside the paper to ensure that it can be understood without prior knowledge.
Based on the suggestion of another reviewer, we will also generally extend our explanations about the oracle, placing a high level introduction of the algorithm early in the paper to introduce the reader to the concept.
```

In the main manuscript, the author takes too much space to discuss why they need to have such experimental settings and how they set the experiments. The discussions of the experimental results are weak. Maybe this paper is more suitable to submit to a journal.
```
As you have mentioned in your review, there exists a fair number of benchmarks for AL already.
Each of those specialize on 1-2 domains and place high emphasis on their results.
However, apart from [1], each benchmark has the experimental setup and technical details on low priority, leading to a situation, where everyone is doing their own framework in isolation and an informed discussion about best practices is very hard.
We hope (in combination with [1]) to open the door for such discussion in the area of AL.
Nonetheless, we will extend discussion of our results. 
Based on the comment from another reviewer, we are planning to study our results from the dimension of class imbalance, which currently is missing in our work. 
[1]: Ji, Yilin, et al. "Randomness is the root of all evil: more reliable evaluation of deep active learning." Proceedings of the IEEE/CVF Winter Conference on Applications of Computer Vision. 2023.
```

"In semi-supervised learning, which is to train a fixed encoder-model, pre-encode the datasets and then only train a single linear layer as classifier." It can be computationally efficient, but it will not fully exploit the strengths of deep learning in an active learning scenario. A more typical and potentially more effective approach in deep active learning involves training the feature extractor alongside the classifier to ensure that the representations are continuously adapted to the newly labeled data.
```
As described in Section 5.2 we are not interested in beating the SOTA performance on any dataset, but to provide a fast and low-variance framework.
The same idea applies to our choice of semi-supervised learning.
For that reason we opted for pre-encoded datasets instead of more sophisticated approaches.
Finally, the authors of [4] provide good evidence, that the ranking of algorithms does not change when the semi-supervised training changes (at least for the image domain).
[4]: Zhang, Jifan, et al. "LabelBench: A Comprehensive Framework for Benchmarking Adaptive Label-Efficient Learning." Journal of Data-centric Machine Learning Research (2024).
```

# Reviewer c7pZ
rating: 5, confidence 5
```
Thank you for your review...
```

Include more related benchmarks. For example [1] studies extensively the behavior of active learning in tabular datasets. LabelBench [2] has also proposed using embedded features and semi-supervised learning as part of their benchmark. While the authors in this paper conjecture "a well-performing method in our benchmark will also generalize well to larger datasets and classifiers", LabelBench has already demonstrated this to be true. In addition, margin sampling performing well on tabular datasets is also found in [1].
```
[1] and [2] will be integrated into Table 1 and our related work section.
Based on the comment of another reviewer we aim to extend our discussion section. 
This will also include a more in-depth comparison of our results with other benchmarks like [1].
```

The proposed metric of average ranking may not be a convenient/intuitive metric for benchmarks. Specifically, whenever a new algorithm is introduced, the scores of every algorithm will change. Moreover, practitioners are generally interested in either accuracy or label-efficiency (number of labels needed to reach a certain accuracy). The adopted metric does not capture any of these quantities directly. In some cases, entropy, badge and margin may all perform very similarly in accuracy/label-efficiency but remains stable in their relative ranking. The metric in this paper would significantly penalize the algorithms that perform relatively worse, which does not give an accurate quantification in the actual performance of the algorithms themselves.
```
We fully agree with the reviewer that accuracy scores need to be reported for any classification problem that is not severely imbalanced.
We have all accuracy scores in Appendix K, including standard deviations and the amount of wins per algorithm.
We would like to point out that our benchmark aggregrates results from different datasets, from which we cannot simply average the accuracies to display them in a table, etc.
For a fair comparison, we rely on the paired-t-test, which the Critical-Difference-Diagrams conveniently provide.
Lastly, we have been unable to include a second results table besides Table 3 within the page limit.
```

An interesting finding in LabelBench is imbalanced active learning algorithms like GALAXY performs significantly better on imbalanced datasets. Since most of the language and tabular datasets are imbalanced, it would be interesting to see how such algorithms perform. The conclusion that "entropy sampling" or "margin sampling" is the best may not be entirely accurate.
```
We omitted information about class imbalance for our datasets, as we did not focus on that aspect.
However, both our text, and 2 out of 3 tabular datasets are imbalanced by nature. Implementing GALAXY and discussing results along this dimension as well is a great suggestion. 
Since testing GALAXY on our entire benchmark is a significant computational effort, we only provide preliminary experiments on the TopV2 dataset.
```

### 2024-08-13 download_all_datasets.py

@@ -5,7 +5,7 @@ import numpy as np
 
 
 parser = argparse.ArgumentParser()
-parser.add_argument("--data_folder", type=str, required=True)
+parser.add_argument("--data_folder", type=str, default="../datasets")
 parser.add_argument("--encode", type=int, default=0)
 args = parser.parse_args()
 args.encode = bool(args.encode)
@@ -29,7 +29,8 @@ for name in all_names:
         config = yaml.load(f, yaml.Loader)
     pool_rng = np.random.default_rng(1)
     DatasetClass = get_dataset_by_name(name)
-    DatasetClass(args.data_folder, config, pool_rng, encoded=False)
+    d = DatasetClass(args.data_folder, config, pool_rng, encoded=False)
+    print("Class balance", d.y_train.sum(dim=0))
     if args.encode:
         try:
             DatasetClass(args.data_folder, config, pool_rng, encoded=True)

### 2024-08-14 doc/rebuttal.md
rebuttal

@@
 ```
-The results on the synthetic benchmarks are unsurprising.
+
+The results on the synthetic benchmarks are unsurprising and the contribution is unclear since they show that methods that focus on uncertainty sampling like margin fail on distributions that are adversarially designed to make them fail.
+```
+```
+
 The algorithm for the greedy oracle has no clear guarantees, the design choices are unclear and the description is hard to follow.
+```
+```
+
 The evaluation is done with relatively small models and datasets and it is unclear whether it translates to larger ones.
+```
+Verweis auf Zhang et. al., siehe unten
+```
 
@@
 ```
-The entire section explaining the oracle will be revised for the camera ready version, apart from introducing the concept earlier in the paper, we aim to improve the understanding of readers without prior knowledge.
+The entire section explaining the oracle will be revised for the camera ready version, apart from introducing the concept earlier in the paper, we aim to improve the understanding of readers without prior knowledge (also mentioned by another reviewer).
 ```
@@
 ```
-We have 9 datasets in our benchmark, 5 of whichhave a pre-encoded version (excluding text and synthetic) , which brings the total number of experiments to 14.
+We have 9 datasets in our benchmark, 5 of which have a pre-encoded version (excluding text and synthetic) , which brings the total number of experiments to 14.
 ```
@@
 ```
-@Max
-
+It is true that, by proposing to use a fully labeled validation set, we implicitly claim that most AL research so far was done wrong.
+However, we don't want to invalidate AL literature as a whole, but rather push for a stronger separation of AL research (with validation set) and AL applications (without validation set).
+With our idea, we don't try to mask the high variance, but rather claim that **due** to the high variance we should consider fully optimizing our hyperparameters on validation to not make the effects of it worse.
+The core hypothesis is that a top-performing algorithm in an experiment with good hyperparameters also performs well in the application case with bad/worse hyperparameters.
 ```
@@
 ```
-We fully agree with the reviewer that accuracy scores need to be reported for any classification problem that is not severely imbalanced.
+We fully agree with the reviewer that accuracy scores need to be reported for any classification problem.
 We have all accuracy scores in Appendix K, including standard deviations and the amount of wins per algorithm.
-We would like to point out that our benchmark aggregrates results from different datasets, from which we cannot simply average the accuracies to display them in a table, etc.
+We would like to point out that our benchmark aggregrates results from different datasets, which we cannot be done by simply averaging the accuracies to display them in a table, etc.
 For a fair comparison, we rely on the paired-t-test, which the Critical-Difference-Diagrams conveniently provide.
-Lastly, we have been unable to include a second results table besides Table 3 within the page limit.
+Lastly, we have been unable to include a second results table besides Table 3 within the page limit and therefore located our accuracy values in the Appendix.
 ```
@@
 However, both our text, and 2 out of 3 tabular datasets are imbalanced by nature. Implementing GALAXY and discussing results along this dimension as well is a great suggestion. 
-Since testing GALAXY on our entire benchmark is a significant computational effort, we only provide preliminary experiments on the TopV2 dataset.
+Since testing GALAXY on our entire benchmark is a significant computational effort, we can only provide experiments on the TopV2 dataset for this rebuttal (doc/img/micro_dna).
+However, GALAXY is indeed the top performing algorithm on the unbalanced TopV2 dataset.
+We thank the reviewer for this valuable addition to the benchmark and will fully incorporate GALAXY for the paper ready version.
 ```

### 2024-08-26 doc/rebuttal.md

@@
 ```
+Our primary goal is to validate these approaches through a targeted ablation study, highlighting the strengths and limitations of both types of algorithms. 
+You are correct when you say that algorithms like margin sampling behave exactly as expected, but in our opinion, the behavior of a hybrid algorithm like BADGE was completely untested in these adversarial environments (in the case of BADGE, we revealed weaknesses akin to vanilla uncertainty sampling methods, even though BADGE uses an clustering-like algorithm to select its samples).
+These datasets mainly serve as a resource for benchmarking future AL methods, particularly for hybrid approaches that use ideas from uncertainty sampling and clustering/diversity methods.
 ```
@@
 ```
+Guarantees:
+Design Choices:
+The description of the Oracle was negatively pointed out by two other reviewers as well. 
+The entire section explaining the oracle will be revised for the camera ready version. 
+Apart from introducing the concept earlier in the paper, we aim to improve the understanding of readers without prior knowledge.
 ```
@@
 ```
-Verweis auf Zhang et. al., siehe unten
+We agree that no comparable benchmark has shown a systematic conversion of small-model setup to large-model setups.
+However, we have evidence in domain-specific literature that supports our hypothesis:
+[4] use larger image models like ViT-B32 and also find least confidence sampling and BADGE to be the best algorithms for images
+[5] use BERT models in the text domain and also find BADGE to be the best (they don't evaluate margin sampling)
+[6] use considerably larger MLPs for the tabular domain than we do, but also find margin sampling to be consistenly top-performing
+
+References:
+[4]: Zhang, Jifan, et al. "LabelBench: A Comprehensive Framework for Benchmarking Adaptive Label-Efficient Learning." Journal of Data-centric Machine Learning Research (2024).
+[5]: Rauch, Lukas, et al.
+"Activeglae: A benchmark for deep active learning with transformers." Joint European Conference on Machine Learning and Knowledge Discovery in Databases.
+Cham: Springer Nature Switzerland, 2023.
+[6]: Bahri, Dara, et al. "Is margin all you need? An extensive empirical study of active learning on tabular data." arXiv preprint arXiv:2210.03822 (2022).
 ```
@@
 ```
-The entire section explaining the oracle will be revised for the camera ready version, apart from introducing the concept earlier in the paper, we aim to improve the understanding of readers without prior knowledge (also mentioned by another reviewer).
+The entire section explaining the oracle will be revised for the camera ready version. 
+Apart from introducing the concept earlier in the paper, we aim to improve the understanding for readers without prior knowledge (also mentioned by another reviewer).
 ```
@@
 ```
-It is true that, by proposing to use a fully labeled validation set, we implicitly claim that most AL research so far was done wrong.
+It is true that, by proposing to use a fully labeled validation set, we implicitly claim that most AL research so far had been flawed.
 However, we don't want to invalidate AL literature as a whole, but rather push for a stronger separation of AL research (with validation set) and AL applications (without validation set).
-With our idea, we don't try to mask the high variance, but rather claim that **due** to the high variance we should consider fully optimizing our hyperparameters on validation to not make the effects of it worse.
+With our approach, we are not trying to mask the high variance; instead, we argue that **due** to the high variance, we should fully optimize our hyperparameters on validation to avoid exacerbating its effects.
 The core hypothesis is that a top-performing algorithm in an experiment with good hyperparameters also performs well in the application case with bad/worse hyperparameters.
@@
 We will include [r1] into Table 1 and our related work ([r2] and [r3] do not provide experiments on their own).
-AL benchmarks are still riddled with comparability issues, because their choice of datasets, domains and models ([r2] Sec. 4.1 and [r3] Tab. 1) often is completely disjoint.
+AL benchmarks are still riddled with comparability issues, because their choice of datasets, domains and models often is completely disjoint ([r2] Sec. 4.1 and [r3] Tab. 1).
 In our opinion, we need a benchmark that spans all domains and the most commonly used datasets and models to fix this problem.
-As described in our Section 5, we selected our datasets for maximum intersection with previous works.
-The same is true for our model selection, which either uses common models like ResNet18 or trivial to reproduce models like MLPs.
+As described in our Section 5, we selected our datasets for maximum overlap with previous works.
+The same is true for our model selection, which either uses common models like ResNet18 or trivial to reproduce models, like MLPs.
 [r2]: Zhang, Zhisong, Emma Strubell, and Eduard Hovy. "A survey of active learning for natural language processing." arXiv preprint arXiv:2210.10109 (2022).
@@
 ```
-Even though it is common knowledge that repeating an experiment enough times is crucial for ML research, in the field of Active Learning we still observe many authors not doing so. 
-To the best of our knowledge, we are the first to provide a first insight into exactly how often an AL experiment should be repeated in order to generate reliable results.
-The same is true for experimenting on different domains. Everyone knows that algorithms should be tested in as many domains as possible, but few actually apply that knowledge.
+Even though it is common knowledge that repeating an experiment enough times is crucial for ML research, in the field of Active Learning, we still observe many authors not doing so. 
+To the best of our knowledge, we are the first to provide insight into exactly how often an AL experiment should be repeated in order to generate reliable results.
+The same is true for experimenting across different domains. Everyone knows that algorithms should be tested in as many domains as possible, but few actually apply that knowledge.
 ```
@@
 ```
-We will revise the entire section about our oracle algorithm and work with more collegues from outside the paper to ensure that it can be understood without prior knowledge.
-Based on the suggestion of another reviewer, we will also generally extend our explanations about the oracle, placing a high level introduction of the algorithm early in the paper to introduce the reader to the concept.
+We will revise the entire section about our oracle algorithm and collaborate with more colleagues from outside the paper to ensure that it can be understood without prior knowledge.
+Based on the suggestion of another reviewer, we will also generally extend our explanations about the oracle, placing a high-level introduction of the algorithm early in the paper to introduce the reader to the concept.
 ```
@@
 ```
-As you have mentioned in your review, there exists a fair number of benchmarks for AL already.
-Each of those specialize on 1-2 domains and place high emphasis on their results.
-However, apart from [1], each benchmark has the experimental setup and technical details on low priority, leading to a situation, where everyone is doing their own framework in isolation and an informed discussion about best practices is very hard.
+As you have mentioned in your review, there are a fair number of benchmarks for AL already.
+Each of those focus on 1-2 domains and place high emphasis on their results.
+However, apart from [1], each benchmark has the experimental setup and technical details on low priority, leading to a situation where everyone is doing their own framework in isolation and an informed discussion about best practices is very difficult.
 We hope (in combination with [1]) to open the door for such discussion in the area of AL.
 Nonetheless, we will extend discussion of our results. 
-Based on the comment from another reviewer, we are planning to study our results from the dimension of class imbalance, which currently is missing in our work. 
+Based on the comment from another reviewer, we plan to study our results from the dimension of class imbalance, which currently is missing in our work. 
 [1]: Ji, Yilin, et al. "Randomness is the root of all evil: more reliable evaluation of deep active learning." Proceedings of the IEEE/CVF Winter Conference on Applications of Computer Vision. 2023.
@@
 ```
-As described in Section 5.2 we are not interested in beating the SOTA performance on any dataset, but to provide a fast and low-variance framework.
+As described in Section 5.2, we are not interested in beating the SOTA performance on any dataset, but rather in providing a fast and low-variance framework.
 The same idea applies to our choice of semi-supervised learning.
-For that reason we opted for pre-encoded datasets instead of more sophisticated approaches.
-Finally, the authors of [4] provide good evidence, that the ranking of algorithms does not change when the semi-supervised training changes (at least for the image domain).
+For that reason, we opted for pre-encoded datasets instead of more sophisticated approaches.
+Finally, the authors of [4] provide good evidence that the ranking of algorithms does not change when the semi-supervised training changes (at least for the image domain).
 [4]: Zhang, Jifan, et al. "LabelBench: A Comprehensive Framework for Benchmarking Adaptive Label-Efficient Learning." Journal of Data-centric Machine Learning Research (2024).
@@
 We have all accuracy scores in Appendix K, including standard deviations and the amount of wins per algorithm.
-We would like to point out that our benchmark aggregrates results from different datasets, which we cannot be done by simply averaging the accuracies to display them in a table, etc.
-For a fair comparison, we rely on the paired-t-test, which the Critical-Difference-Diagrams conveniently provide.
-Lastly, we have been unable to include a second results table besides Table 3 within the page limit and therefore located our accuracy values in the Appendix.
+We would like to point out that our benchmark aggregates results from different datasets, which we cannot be done by simply averaging the accuracies to display them in a table, etc.
+For a fair comparison, we rely on the paired-t test, which the Critical-Difference-Diagrams conveniently provide.
+Lastly, we were unable to include a second results table besides Table 3 within the page limit and therefore located our accuracy values in the Appendix.
 ```
@@
 ```
-We omitted information about class imbalance for our datasets, as we did not focus on that aspect.
-However, both our text, and 2 out of 3 tabular datasets are imbalanced by nature. Implementing GALAXY and discussing results along this dimension as well is a great suggestion. 
+We omitted information about class imbalance in our datasets, as we did not focus on that aspect.
+However, both our text and 2 out of 3 tabular datasets are imbalanced by nature. Implementing GALAXY and discussing results along this dimension as well is a great suggestion. 
 Since testing GALAXY on our entire benchmark is a significant computational effort, we can only provide experiments on the TopV2 dataset for this rebuttal (doc/img/micro_dna).
-However, GALAXY is indeed the top performing algorithm on the unbalanced TopV2 dataset.
-We thank the reviewer for this valuable addition to the benchmark and will fully incorporate GALAXY for the paper ready version.
+However, GALAXY is indeed the top-performing algorithm on the imbalanced TopV2 dataset.
+We thank the reviewer for this valuable addition to the benchmark and will fully incorporate GALAXY for the camera-ready version.
 ```

### 2024-10-10 doc/rebuttal.md
camera ready

@@
 ```
-Guarantees:
-Design Choices:
-The description of the Oracle was negatively pointed out by two other reviewers as well. 
-The entire section explaining the oracle will be revised for the camera ready version. 
-Apart from introducing the concept earlier in the paper, we aim to improve the understanding of readers without prior knowledge.
+We designed our oracle based in the following two ideas:
+- Have an algorithm that does not rely on search or excessive enumeration, so it can be computed efficiently 
+- Pose a performance that is "good enough", so that it beats every AL algorithm on average without wasting computational resources for diminishing lifts
+
+Since we are mainly interested in beating other AL algorithms and not a theoretically optimal upper bound we did not see the need derive a performance guarantee for our oracle.
+
+To improve the description of the oracle, we will add comments to the pseudocode in Appendix H and reference it more clearly in the main body:
+Remove Line 258: The pseudocode for our oracle can be found in App. H.
+Add: For a detailed description of the oracle algorithm and how it is used in the AL loop, please refer to App. H.
 ```
@@
 We agree that no comparable benchmark has shown a systematic conversion of small-model setup to large-model setups.
-However, we have evidence in domain-specific literature that supports our hypothesis:
+However, we have evidence in domain-specific literature that supports our hypothesis that our ranking transfers to larger models:
 [4] use larger image models like ViT-B32 and also find least confidence sampling and BADGE to be the best algorithms for images
@@
 ```
-The entire section explaining the oracle will be revised for the camera ready version. 
-Apart from introducing the concept earlier in the paper, we aim to improve the understanding for readers without prior knowledge (also mentioned by another reviewer).
+We will insert a sentence in line 55 to clarify the main idea of our oracle algorithm early on:
+"Our oracle relies on directly testing a small sample of points in every iteration if they induce an improvement in test accuracy and selects the optimal point from that small sample."
 ```
@@
 We have 9 datasets in our benchmark, 5 of which have a pre-encoded version (excluding text and synthetic) , which brings the total number of experiments to 14.
+We will include this description in the caption of Table 1.
 ```
@@
 ```
-Similar to the explanations about the oracle, we aim to revise this section to improve the understanding of readers without prior knowledge.
+We will add all visible curves to the legend in Figure 1.
+Additionally, we will revise the caption of Figure 1 to make it more accessible for the readers.
 ```
@@
 It is true that, by proposing to use a fully labeled validation set, we implicitly claim that most AL research so far had been flawed.
-However, we don't want to invalidate AL literature as a whole, but rather push for a stronger separation of AL research (with validation set) and AL applications (without validation set).
+However, we don't want to challenge AL literature as a whole, but rather push for a stronger separation of AL research (with validation set) and AL applications (without validation set).
 With our approach, we are not trying to mask the high variance; instead, we argue that **due** to the high variance, we should fully optimize our hyperparameters on validation to avoid exacerbating its effects.
@@
 ```
-We will revise the entire section about our oracle algorithm and collaborate with more colleagues from outside the paper to ensure that it can be understood without prior knowledge.
-Based on the suggestion of another reviewer, we will also generally extend our explanations about the oracle, placing a high-level introduction of the algorithm early in the paper to introduce the reader to the concept.
+See below
 ```
@@
 [1] and [2] will be integrated into Table 1 and our related work section.
-Based on the comment of another reviewer we aim to extend our discussion section. 
-This will also include a more in-depth comparison of our results with other benchmarks like [1].
+Based on the comment of another reviewer we aim to extend our discussion of our results in the appendix, which will also include a more in-depth comparison of our results with other benchmarks like [1].
 ```

### 2024-10-10 doc/al_benchmark_paper.tex
camera ready

@@
 
-%\title{Towards Comparable Active Learning}
 \title{A Cross-Domain Benchmark for Active Learning}
@@
 	Anonymous Authors
-%	Thorben Werner 
-%	\thanks{Institute of Computer Science - Information Systems and Machine Learning Lab (ISMLL)} 
-%	\\
-%	University of Hildesheim\\
-%	Universitätsplatz 1\, 31141 Hildesheim \\
-%	\texttt{werner@ismll.de} \\
-%	\And
-%	Johannes Burchert$^*$ \\
-%	University of Hildesheim\\
-%	Universitätsplatz 1, 31141 Hildesheim \\
-%	\texttt{burchert@ismll.de} \\
-%	\And
-%	Maximilian Stubbemann$^*$ \\
-%	University of Hildesheim\\
-%	Universitätsplatz 1, 31141 Hildesheim \\
-%	\texttt{stubbemann@ismll.de} \\
-%	\AND
-%	Lars Schmidt-Thieme$^*$ \\
-%	University of Hildesheim\\
-%	Universitätsplatz 1, 31141 Hildesheim \\
-%	\texttt{schmidt-thieme@ismll.uni-hildesheim.de}
+	Thorben Werner 
+	\thanks{Institute of Computer Science - Information Systems and Machine Learning Lab (ISMLL)} 
+	\\
+	University of Hildesheim\\
+	Universitätsplatz 1\, 31141 Hildesheim \\
+	\texttt{werner@ismll.de} \\
+	\And
+	Johannes Burchert$^*$ \\
+	University of Hildesheim\\
+	Universitätsplatz 1, 31141 Hildesheim \\
+	\texttt{burchert@ismll.de} \\
+	\And
+	Maximilian Stubbemann$^*$ \\
+	University of Hildesheim\\
+	Universitätsplatz 1, 31141 Hildesheim \\
+	\texttt{stubbemann@ismll.de} \\
+	\AND
+	Lars Schmidt-Thieme$^*$ \\
+	University of Hildesheim\\
+	Universitätsplatz 1, 31141 Hildesheim \\
+	\texttt{schmidt-thieme@ismll.uni-hildesheim.de}
 }
@@
 			Semi DNA    & Linear & 40 & o & o &&&&& \\
-			Semi Splice & Linear & 100 & o & o & o & o &&&\\
+			Semi Splice & Linear & 100 & o & o & o &&&&\\
 			TopV2       & BiLSTM & 200 & o & o & o & o &&& \\

### 2024-10-10 core/evaluation.py

@@ -2,7 +2,6 @@ import itertools, math
 import os
 from os.path import exists, join
 import matplotlib.pyplot as plt
-import seaborn as sns
 import pandas as pd
 import numpy as np
 from tqdm import tqdm
@@ -161,6 +160,7 @@ def _query_to_list(query, current_folder):
     else:
         raise ValueError(f"Query not recognized: {query}")
     return result_list
+
 
 def generate_rank_leaderboard(precision=1, add_std=True, subsample_runs=None):
     datasets_raw = ["Splice", "DNA", "USPS", "Cifar10", "FashionMnist", "TopV2", "News",]
@@ -354,6 +354,7 @@ def average_out_columns(df:pd.DataFrame, columns:list):
             result_list.append(sub_df)
         result_df = pd.concat(result_list)
     return result_df
+
 
 def std_for_column(df:pd.DataFrame, column:str):
     result_df = df.copy(deep=True)
